### 2017-06-11 paper.tex
061117

\documentclass[]{article}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{amsthm}

\newtheorem{thm}{Theorem}[subsection]

%opening
\title{Dynamics of Geodesic Flows on the Infinite Cube Step Surface}
\author{Pavel Javornik}

\begin{document}

\maketitle

\begin{abstract}

\end{abstract}

\section{Introduction}
\subsection{Background}
The purpose of the following paper is to study the non-compact, infinite surface generated entirely out of the edge-to-edge gluing of square faces, whose final product resembles cubes stacked in a cascading, pyramidal step pattern (or perhaps a rhombile tiling). (See Fig //) The surface we call the "Cube Step" is a member of the family of infinite Euclidean cone surfaces built out of rectangles with vertices of cone angle $3\pi$ and $\frac{3\pi}{2}$. We wish to study the surface in the case where we glue only squares with edges of length 1. Specifically, we want to study geodesic flows on this surface and to prove that properties of the initial trajectory angle determine whether or not the geodesic will close or drift periodically.

The determining factor is simply whether or not the trajectory angle vector direction $(a,b)$, relative to the bottom and left edges of any square face, has integer values $a$,$b$ (where $a$ and $b$ are relatively prime) are of the same parity (odd-odd slope) or different parities (even-odd slope). We choose to neglect cases where paths are incident with vertices, and proceed to treat them as singularities. With these conditions in place we found that any geodesic with an initial odd-odd trajectory closes on the surface, while the even-odd cases drift periodically, tending to infinity.

To better understand this surface, we will adapt methods used in the study of Veech surfaces to suit our needs. These methods have become incredibly valuable to anyone studying rational billiard flows on a variety of compact or (in the case of infinite surfaces) non-compact surfaces. Geodesic flows in these problems are characterized as point masses moving along frictionless surfaces that experience perfectly elastic collisions with the boundaries of the surface. The straight line path is then deformed by a reflection over the boundary in the opposite direction. Translation surfaces constructed in the following manner therefore take on symmetric properties in accordance to these transformations and the types of boundaries the point mass might encounter. That is to say that orienting these flows requires that the translation surface itself be reflected to realign the flow. These translation surfaces are often described as being "unfolded" by some billiard flow, or strung along a laundry line of polygons congruent to the base. In the later sections we will describe how our surface's canonical form deforms geodesic flows by rotation and how we can use these rotations and edge-crossings to determine whether or not a geodesic will close.

What inspired this paper was the work done by Hubert et al. [Hu] on the Ehrenfest-Wind Tree model. Any readers who are familiar with their paper or the Tree model will most likely notice the similarities between the quotient spaces of the two surfaces and the properties that determine whether or not geodesic paths close. While similar, these surfaces are not entirely identical. The effect that a boundary collision has on a geodesic alters the dynamical properties of the surface. Or perhaps they better demonstrate properties of the surface itself. Another fundamental difference between the two is that the Cube Step is a surface with cone angles surrounding an infinite number of singularities on the surface. The problem has to take into careful consideration what it means for a path to cross a saddle connection between these singularities of the translation surface.

\subsection{Procedure}
The Cube Step surface has an interesting construction process that lets us take each face of the cube and arrange them in any way we choose so long as we do not change how the edges are identified, or glued together. It is preferable to study the flows on this surface as though every cube has been flattened onto the plane. This means that we can consider the cube step to be tiled by infinitely many of L-shaped surfaces whose edges are identified by rotation.

Doing so greatly reduces the number of possible directions that a flow might travel. Since every edge of this L-shaped surface is identified to another either by translation or a 90 degree rotation, any straight line path on this flattened version of the Cube Step can experience only one of four possible directions at any given time, whereas the Cube Step would've had considerably more.

Moreover, every face of this L-shaped surface is given by rotating its associated face on the Cube Step about the x or y axis, affine translations, or a combination of the two. By cutting and gluing every half-cube of the Cube Step in a consistent fashion, we can preserve the areas of the faces and distances between any two points on the surface with the flattening procedure. We describe this procedure as piecewise maps that are comprised of compositions of Euclidean matrices and translations of points relative to the L-shape onto a subset of $\mathbb{R}^{2}$ with square shaped holes cut out. Since rotations and translations of points preserves distance between them, the path lengths and  associated properties of geodesic flows remain invariant.

The image of this map is a flattened version of the Cube Step. (See FIG [MATLAB MODEL]). From that surface we construct a new one made out of four copies of original, each rotated by some factor of 90 degrees so as to realign the geodesic flows and assure that they can only travel in one possible direction. This four-fold surface is then a cover of the original surface. We consider every possible direction to be a "phase" of the flow, and index the planes accordingly. Any edges in the holes of the surface are identified with an edge on an adjacent phase plane (By rotation of $\pm \frac{\pi}{4}$). Any point on this surface has a coordinate in $\mathbb{R}\times\mathbb{R}\times\mathbb{Z}_{4} $

Similarly, we take the quotient space of the flat Cube Step centered around each square shaped hole and realign its flows by rotation. The resulting surface is a compact branch cover of the quotient space. Singularities on the translation surface have a total angle that is some multiple of $2\pi$. We consider the translation surface to be a base surface of the four-fold cover, and study its properties to prove our main theorem. In particular, we wish to determine what effect a member of the translation surface's Veech group has on both the surface and geodesic flow in hopes that it would shine some light on odd-odd even-odd phenomenon.

 $SL_{2}(\mathbb{R})$

\section{Constructing the Surface}
This section will detail how the Cube Step is constructed out of an infinite number of squares, and explain the process by which we flatten the structure in such a way that preserves the metric, and makes it simpler for us to describe the behavior of geodesics at these edges.

\subsection{The Cube Step Surface and Isometric Variant }
To build the surface out of infinitely many squares, we take the union of the instances of three squares, each lying on planes orthogonal to the other two and sharing an edge with exactly one other square. The three squares have one vertex in common.

Every instance can be described as cube whose bottom, left, and rear faces have been removed. They will be arranged in such a way that every integer triple $m,n,p\in\mathbb{Z}$ will have this surface associated with it. To each face we ascribe a letter A, B, or C. This will be useful when defining the flattening map.
\begin{gather*}
\chi_{m,n,p} = ([m, m+1]\times[n,n+1]\times\{p\})
\\\cup  (\{m+1\}\times[n,n+1]\times[p,p-1]) 
\\\cup ([m,m+1]\times\{n+1\}\times[p,p-1])
\end{gather*}

is the family of these instances based at some integer triple. The surface will be comprised of the union of the members that satisfy the property, $m+n+p=0$. From here on we will refer to the Cube Step surface embedded form in $\mathbb{R}^{3}$ as $\mathbf{C}$. That is to say that

$$\mathbf{C} =\bigcup\big{\{}\chi_{m,n,p}  : m+n+p=0 \big{\}} .$$

$\mathbf{C}$ now resembles the structure in (FIG). Before describing the transformation that takes $\mathbf{C}$ to its flattened form, it would be best to describe the relations on a subset of the plane that preserves the identification of edges. We will call this surface $\mathbf{U}'$. However, the flattening of the surface takes $\mathbf{C}$ to a translation of $\mathbf{U}'$, we call $\mathbf{U}$ (See FIG). We will denote the metric preserving map from $\mathbf{C}$ to $\mathbf{U}$, by $\Psi$. Defining these edge identifying relations over $\mathbf{U}'$ is more convenient for us than defining them over $\mathbf{U}$, but defining $\Psi$ is more convenient when the map is taken to $\mathbf{U}$. Later on we will use $\mathbf{U}'$ when taking its quotient space.

Now we will describe exactly how to construct $\mathbf{U}'$. Take the subset of the real plane, $\mathbf{P} = \mathbb{R}^{2}-\{(u,v)|2m-\frac{1}{2}<u<2m+\frac{1}{2}, 2n-\frac{1}{2}<v<2n+\frac{1}{2} \}$, with 1$\times$1 open squares cut out at every even integer pair and let $\mathbf{U}'$ represent the flattened structure with the proper edge identifications with the first "hole" centered at the origin. We will describe this surface as a product of the auxiliary plane,$z=0$ and the image of a natural map from $\mathbf{P}$ to its quotient space with the following relation on $\mathbf{P}$:

\begin{gather*}
\text{Let } x_{0}=(u_{0},v_{0}),x_{1}=(u_{1},v_{1}) \in \mathbf{P}. \text{Let } \mathbf{R} \text { be a relation described as } x_{0}\mathbf{R}x_{1} \\ \text{ iff } x_{0}=x_{1}
 \text{ or if }x_{0},x_{1} \in {\partial} \left( \left[2m-\frac{1}{2},2m+\frac{1}{2}\right] \times \left[2n-\frac{1}{2},2n+\frac{1}{2}\right] \right)\\
  \left[\begin{array}{c}
u_{1} -2m
\\v_{1}-2n
\end{array}\right] = \left[\begin{matrix}
0 && 1\\
1 && 0
\end{matrix}\right]
\left[ \begin{array}{c}u_{0}-2m\\
v_{0}-2n
\end{array}\right]
\\\text{For some }m,n\in\mathbb{Z} \end{gather*}

The relation will either relate every point to itself, or (in the case where any two points lie on the boundary of the square shaped holes) take the center of each square shaped hole in the plane to the origin and reflect it over the line $y=x$ to match up these points accordingly.  Let $N: \mathbf{P} \rightarrow \mathbf{P}/\mathbf{R}$ denote the natural map from every point on $\mathbf{P}$ to its appropriate equivalence class. Now, let $\mathbf{U}' = Im N \text{ }\times\{0\}$ be the flat representation of the surface that has been translated so as to center the first square cut-out region over the origin. The map $\mathbf{U}\rightarrow\mathbf{U}'$ is given by an affine translation of the surface $(x,y,z) \mapsto (x-\frac{3}{2},y-\frac{3}{2},z)$. (The point at the origin of $\mathbf{C}$ is mapped to $(-\frac{3}{2},-\frac{3}{2},0) \text{ on } \mathbf{U}'$.)

\subsection{Flattening $\mathbf{C}$}
This section will describe the isomorphic map $\Psi:\mathbf{C}\rightarrow\mathbf{U}$, its inverse, $\Psi^{-1}:\mathbf{U}\rightarrow\mathbf{C}$, and prove that $\mathbf{C}$ and $\mathbf{U}$ are isometric.

$\Psi$ is defined piecewise, and denote each subset of $\mathbf{C}$ with the letters $A$, $B$, $C$. We do the same for $\mathbf{U}$ but use instead $A'$, $B'$, $C'$.

\begin{gather*}
	\text{Let }x,y,z\in\mathbb{R}. \text{ Then}
	\\A=\bigcup\big\{\mathbf{C}\cap\{(x,y,p)\in\mathbb{R}\}: p\in\mathbb{Z}\big\}
	\\B=\bigcup\big\{\mathbf{C}\cap\{(x,n,z)\in\mathbb{R}\}: n\in\mathbb{Z}\big\}
	\\C=\bigcup\big\{\mathbf{C}\cap\{(m,y,z)\in\mathbb{R}\}: m\in\mathbb{Z}\big\}
	\\\text{are the collection of the faces of cubes that are parallel to eachother.}
\end{gather*}

To describe the subsets $A'$, $B'$, and $C'$ properly we will introduce the function $\mu_{n}:\mathbb{R}\rightarrow \left[0,n\right) $ as shorthand for the family of functions that take a real number to its residue class of reals modulo n, for some integer n.

\begin{gather*}
	\text{Let }u,v\in\mathbb{R}. \text{Then }
	\\A'= \{(u,v,0): 0<\mu_{2}(u)\leq 1 ,\hspace{4mm} 0<\mu_{2}(v)\leq 1\}
	\\B'= \{(u,v,0): 1<\mu_{2}(u)\leq 2,\hspace{4mm} 0<\mu_{2}(v)\leq 1\}
	\\C'= \{(u,v,0): 0<\mu_{2}(u)\leq 1,\hspace{4mm} 1<\mu_{2}(v)\leq 2\}
	\\\text{are the collections of the associated faces of the flattened structure.}
\end{gather*}

Now we will define $\Psi$ piece-wise.

\begin{equation}
\Psi\left[\begin{array}{c}
	x\\y\\z
\end{array}\right] 
= 
\begin{cases}
	\left[ \hspace{2mm} \begin{matrix}
		1 & 0 & 0 \\
		0 & 1 & 0 \\
		0 & 0 & 1
	\end{matrix}\hspace{3mm}\right]

	\left[\begin{array}{c}
	x - \left\lfloor x \right\rfloor
	\\ y- \left\lfloor y \right\rfloor
	\\ -(z - \left\lfloor z \right\rfloor)
	\end{array} \right]
	+
	\left[\begin{array}{c}
		2 \left\lfloor x \right\rfloor
		\\ 2\left\lfloor y \right\rfloor
		\\ z - \left\lfloor z \right\rfloor
	\end{array} \right]
		& \text{if } (x,y,z)\in A	\vspace{2mm}
	\\

	\left[ \begin{matrix}
	0 & 0 & 1 \\
	0 & 1 & 0 \\
	-1 & 0 & 0
	\end{matrix}\hspace{2mm}\right]
	\left[\begin{array}{c}
		x - \left\lfloor x \right\rfloor
		\\ y- \left\lfloor y \right\rfloor
		\\ -(z - \left\lfloor z \right\rfloor)
		\end{array} \right]
	+
		\left[\begin{array}{c}
			2 \left\lfloor x \right\rfloor
			\\ 2\left\lfloor y \right\rfloor
			\\ x - \left\lfloor x \right\rfloor
		\end{array} \right]
			& \text{if } (x,y,z)\in B	\vspace{2mm}
	\\

		\left[ \begin{matrix}
		1 & 0 & 0 \\
		0 & 0 & 1 \\
		0 & -1 & 0
		\end{matrix}\hspace{2mm}\right]
		\left[\begin{array}{c}
			x - \left\lfloor x \right\rfloor
			\\ y- \left\lfloor y \right\rfloor
			\\ -(z - \left\lfloor z \right\rfloor)
			\end{array} \right]
		+
			\left[\begin{array}{c}
				2 \left\lfloor x \right\rfloor
				\\ 2\left\lfloor y \right\rfloor
				\\ y - \left\lfloor y \right\rfloor
			\end{array} \right]
				& \text{if } (x,y,z)\in C	\vspace{2mm}
\end{cases}
\end{equation}

We chose to preserve the matrix structure in order to emphasize the manner in which every square is taken to $\mathbf{U}$ from $\mathbf{C}$ and vice-versa. The negations of some of these variables are a result of this structure being built so as the integer triples lie on the plane $z=-(x+y)$. Defining the maps this way also makes it possible to generalize it for any surface constructed in a similar manner of taking unions of three faces of rectangular prisms in real space. (i.e. $\chi_{m,n,p}$ belongs to a larger family of rectangular arrangements parameterized by the dimensions of the rectangular prism. In this case, $\chi_{m,n,p}$ is a member of this family whose dimensions are $1\times1\times1$).

Now we will define the inverse map, $\Psi^{-1}$, in a similar manner.

\begin{equation}
	\Psi^{-1}\left[\begin{array}{c}
		x\\y\\z
	\end{array}\right] 
	= 
	\begin{cases}
		\left[ \hspace{2mm} \begin{matrix}
			1 & 0 & 0 \\
			0 & 1 & 0 \\
			0 & 0 & 1
		\end{matrix}\hspace{3mm}\right]

		\left[\begin{array}{c}
		x - 2\left\lfloor \frac{x}{2} \right\rfloor
		\\ y- 2\left\lfloor \frac{y}{2} \right\rfloor
		\\ z
		\end{array} \right]
		+
		\left[\begin{array}{c}
			\left\lfloor \frac{x}{2} \right\rfloor
			\\ \left\lfloor \frac{y}{2} \right\rfloor
			\\ -(\left\lfloor \frac{x}{2} \right\rfloor + \left\lfloor \frac{y}{2} \right\rfloor )
		\end{array} \right]
			& \text{if } (x,y,z)\in A'	\vspace{2mm}
		\\
		\left[ \hspace{2mm} \begin{matrix}
			0 & 0 & -1 \\
			0 & 1 & 0 \\
			1 & 0 & 0
		\end{matrix}\hspace{0mm}\right]

		\left[\begin{array}{c}
		-(x - 2\left\lfloor \frac{x}{2} \right\rfloor)
		\\ y- 2\left\lfloor \frac{y}{2} \right\rfloor
		\\ z
		\end{array} \right]
		+
		\left[\begin{array}{c}
			1+\left\lfloor \frac{x}{2} \right\rfloor
			\\ \left\lfloor \frac{y}{2} \right\rfloor
			\\ -(\left\lfloor \frac{x}{2} \right\rfloor + \left\lfloor \frac{y}{2} \right\rfloor -1)
		\end{array} \right]
			& \text{if } (x,y,z)\in B'	\vspace{2mm}

	\\

		\left[ \hspace{2mm} \begin{matrix}
			1 & 0 & 0 \\
			0 & 0 & -1 \\
			0 & 1 & 0
		\end{matrix}\hspace{0mm}\right]

		\left[\begin{array}{c}
		x - 2\left\lfloor \frac{x}{2} \right\rfloor
		\\ -(y- 2\left\lfloor \frac{y}{2} \right\rfloor)
		\\ z
		\end{array} \right]
		+
		\left[\begin{array}{c}
			\left\lfloor \frac{x}{2} \right\rfloor
			\\1+ \left\lfloor \frac{y}{2} \right\rfloor
			\\ -(\left\lfloor \frac{x}{2} \right\rfloor + \left\lfloor \frac{y}{2} \right\rfloor -1)
		\end{array} \right]
			& \text{if } (x,y,z)\in B'	\vspace{2mm}

	\end{cases}
\end{equation}

By using properties of the floor function, one can show that $\Psi$ is indeed a bijective, invertible map. All that is left to show is that $\Psi$ preserves distance.

\begin{thm}{$\mathbf{C}$ and $\mathbf{U}$ are isometric.}\\
Proof. Let $(\mathbf{C}, d)$ and $(\mathbf{U}, d')$ be topological spaces with their respective metrics. Let $a,b\in\mathbf{C}$, and let $\Psi(a)=a',\Psi(b)=b'\in\mathbf{U}$. We will denote the distance between $a$ and $b$ by $D=d(a,b)$, and the distance between $a'$ and $b'$ by $D'=d'(a',b')$. Since $\Psi$ and $\Psi^{-1}$ are linear transformations that act on subsets of auxiliary planes in $\mathbb{R}^{3}$, the images of any curves on $\mathbf{C}$ or $\mathbf{U}$ under these maps will not be deformed and retain their lengths. Thus, $D \leq d'(\Psi(a), \Psi(b)) = d'(a',b') = D'$. Likewise, $D' \leq d(\Psi^{-1}(a'),\Psi^{-1}(b')) = d(a,b) = D$. Therefore $D=D'$, and $\Psi$ and $\Psi^{-1}$ are metric preserving maps between $\mathbf{C}$ and $\mathbf{U}$. $\qed$
\end{thm}

As a consequence of this theorem, any dynamical properties of geodesic flow on the surface $\mathbf{U}'$ immediately applies to $\mathbf{C}$. ($\mathbf{U}'$ and $\mathbf{U}$ are trivially isometric).

\section{The Four-Fold Cover and Translation Surface}

\end{document}

### 2017-06-13 paper.tex
upload

@@
 \usepackage{amsthm}
-
+\usepackage{svg}
+\graphicspath{{images/}}
+\setsvg{svgpath={./images/}}
 \newtheorem{thm}{Theorem}[subsection]
+\newtheorem*{thm*}{Theorem}
 
 %opening
-\title{Dynamics of Geodesic Flows on the Infinite Cube Step Surface}
+\title{Geodesic Flow on the Necker Cube Surface}
+\date{}
 \author{Pavel Javornik}
@@
 
+\begin{center}
+\includesvg[width=2in]{cubesxyz}
+\end{center}
+
 \begin{abstract}
@@
 \section{Introduction}
+
 \subsection{Background}
-The purpose of the following paper is to study the non-compact, infinite surface generated entirely out of the edge-to-edge gluing of square faces, whose final product resembles cubes stacked in a cascading, pyramidal step pattern (or perhaps a rhombile tiling).
-(See Fig //) The surface we call the "Cube Step" is a member of the family of infinite Euclidean cone surfaces built out of rectangles with vertices of cone angle $3\pi$ and $\frac{3\pi}{2}$.
+The purpose of the following paper is to study the non-compact, infinite surface generated entirely out of the edge-to-edge gluing of square faces, whose final product resembles cubes stacked in a cascading, pyramidal pattern (or perhaps a rhombile tiling).
+(See Fig //) M.C.
+Escher famously adapted Louis Albert Necker's optical illusion known as the Necker Cube in 1937.
+The cubes appear in Escher's lithograph, "Metamorphosis I," in the form of tessellated rhombi.[CITE METAMORPHOSIS I] The print features a number of various tessellations transforming into simpler ones when traversed from left to right.
+The term Necker Cube is used interchangeably to refer to Escher's surface and Necker's optical illusion.
+For the purposes of the following paper, we will be using Necker Cube to refer to the infinite cube tiling that appears in Escher's work, as well as many other places.
+Those familiar with the arcade game known as Q*bert will most likely recognize the Necker Cube as the board whose colors you were meant to change in order to complete each level.
+
+\begin{center}
+\includegraphics[scale=0.5]{escher.jpg}
+\\ 
+\text{The Necker Cube as it appears briefly in Metamorphosis I}
+\end{center}
+
+The Necker Cube is a member of the family of infinite Euclidean cone surfaces built out of rectangles with vertices of cone angle $3\pi$ and $\frac{3\pi}{2}$.
 We wish to study the surface in the case where we glue only squares with edges of length 1.
@@
 With these conditions in place we found that any geodesic with an initial odd-odd trajectory closes on the surface, while the even-odd cases drift periodically, tending to infinity.
+We aim to prove the following:
+
+\begin{thm*}{Theorem}
+\\content...
+\end{thm*}
 
@@
 Or perhaps they better demonstrate properties of the surface itself.
-Another fundamental difference between the two is that the Cube Step is a surface with cone angles surrounding an infinite number of singularities on the surface.
+Another fundamental difference between the two is that the Necker Cube is a surface with cone angles surrounding an infinite number of singularities on the surface.
 The problem has to take into careful consideration what it means for a path to cross a saddle connection between these singularities of the translation surface.
@@
 \subsection{Procedure}
-The Cube Step surface has an interesting construction process that lets us take each face of the cube and arrange them in any way we choose so long as we do not change how the edges are identified, or glued together.
+The Necker Cube surface has an interesting construction process that lets us take each face of the cube and arrange them in any way we choose so long as we do not change how the edges are identified, or glued together.
 It is preferable to study the flows on this surface as though every cube has been flattened onto the plane.
@@
 Doing so greatly reduces the number of possible directions that a flow might travel.
-Since every edge of this L-shaped surface is identified to another either by translation or a 90 degree rotation, any straight line path on this flattened version of the Cube Step can experience only one of four possible directions at any given time, whereas the Cube Step would've had considerably more.
+Since every edge of this L-shaped surface is identified to another either by translation or a 90 degree rotation, any straight line path on this flattened version of the Necker Cube can experience only one of four possible directions at any given time, whereas the Necker Cube would've had considerably more.
 
-Moreover, every face of this L-shaped surface is given by rotating its associated face on the Cube Step about the x or y axis, affine translations, or a combination of the two.
-By cutting and gluing every half-cube of the Cube Step in a consistent fashion, we can preserve the areas of the faces and distances between any two points on the surface with the flattening procedure.
+Moreover, every face of this L-shaped surface is given by rotating its associated face on the Necker Cube about the x or y axis, affine translations, or a combination of the two.
+By cutting and gluing every half-cube of the Necker Cube in a consistent fashion, we can preserve the areas of the faces and distances between any two points on the surface with the flattening procedure.
 We describe this procedure as piecewise maps that are comprised of compositions of Euclidean matrices and translations of points relative to the L-shape onto a subset of $\mathbb{R}^{2}$ with square shaped holes cut out.
@@
 
-The image of this map is a flattened version of the Cube Step.
+The image of this map is a flattened version of the Necker Cube.
 (See FIG [MATLAB MODEL]).
@@
 
-Similarly, we take the quotient space of the flat Cube Step centered around each square shaped hole and realign its flows by rotation.
+Similarly, we take the quotient space of the flattened Necker Cube centered around each square shaped hole and realign its flows by rotation.
 The resulting surface is a compact branch cover of the quotient space.
@@
 \section{Constructing the Surface}
-This section will detail how the Cube Step is constructed out of an infinite number of squares, and explain the process by which we flatten the structure in such a way that preserves the metric, and makes it simpler for us to describe the behavior of geodesics at these edges.
+This section will detail how the Necker Cube is constructed out of an infinite number of squares, and explain the process by which we flatten the structure in such a way that preserves the metric, and makes it simpler for us to describe the behavior of geodesics at these edges.
 
-\subsection{The Cube Step Surface and Isometric Variant }
+\subsection{The Necker Cube Surface and Isometric Variant }
 To build the surface out of infinitely many squares, we take the union of the instances of three squares, each lying on planes orthogonal to the other two and sharing an edge with exactly one other square.
@@
 The surface will be comprised of the union of the members that satisfy the property, $m+n+p=0$.
-From here on we will refer to the Cube Step surface embedded form in $\mathbb{R}^{3}$ as $\mathbf{C}$.
+From here on we will refer to the Necker Cube surface embedded form in $\mathbb{R}^{3}$ lying on the plane $z=-(x+y) $ as $\mathbf{C}$.
 That is to say that
@@
 $\mathbf{C}$ now resembles the structure in (FIG).
-Before describing the transformation that takes $\mathbf{C}$ to its flattened form, it would be best to describe the relations on a subset of the plane that preserves the identification of edges.
+Before describing the transformation that takes $\mathbf{C}$ to its flattened form, it would be best to define the relations on a subset of the plane that preserves the identification of edges.
 We will call this surface $\mathbf{U}'$.
@@
 
+The flattening procedure can be represented as a series of cuts made along edges parallel to the z-axis. The gluing structure is preserved by the relations defined on $\textbf{U}$ and $\textbf{U}'$.
+
+\begin{center}
+\includesvg[width=2.3in]{cubesxyzcut}
+\includesvg[width=2.3in]{unfoldcut}
+\text{The images of } $\Psi^{-1}(\mathbf{U})$ \text{ and } $\Psi(\mathbf{C})$, \text{ respectively.}
+\end{center}
+
 $\Psi$ is defined piecewise, and denote each subset of $\mathbf{C}$ with the letters $A$, $B$, $C$. We do the same for $\mathbf{U}$ but use instead $A'$, $B'$, $C'$.
@@
 	\\C=\bigcup\big\{\mathbf{C}\cap\{(m,y,z)\in\mathbb{R}\}: m\in\mathbb{Z}\big\}
-	\\\text{are the collection of the faces of cubes that are parallel to eachother.}
+	\\\text{are the collection of the faces of cubes that lie on planes that are parallel to one another.}
+	\\\text{As shown here:}
 \end{gather*}
+
+\includegraphics[]{label.png}\centering
 
@@
 
+\subsection{Comparing Geodesic Behavior on Both Surfaces}
+Two immediate observations that can be made in regards to geodesic flows on the surface was that any trajectory with slope one closes, and any flows with horizontal or vertical trajectories drift.
+Observe the path behavior of three different flows on the surface:
+
+\begin{center}
+\includesvg[width=2.3in]{cubesxyzpaths} 
+\includesvg[width=2.3in]{unfoldcutpath}
+\\ \text{Here we can see paths a,b, and c on }$\mathbf{U}$\text{ and }$\mathbf{C}$,
+\\ \text{where their initial trajectories are (0,1), (1,0), and (1,1) respectively.}
+\end{center}
+
+The figure above shows us some very simple paths on the surfaces.
+We use the three vectors, (0,1), (1,0), and (1,1), as a sort of basis for proving our theorem.
+Included in the appendix are a few more simulated flows of various trajectories.
+We've color coded the trajectories in the image above for path c to represent one of four phases
+
 \section{The Four-Fold Cover and Translation Surface}

### 2017-06-13 images/svg/doc/examples/root.C

@@ -1,0 +1,27 @@
+void root() {
+  
+  // Set the style.
+  gStyle->SetTextFont(80);        gStyle->SetLabelFont(80,"XYZ");
+  gStyle->SetTitleFont(80,"");    gStyle->SetTitleFont(80,"XYZ"); 
+  gStyle->SetPalette(1);          gStyle->SetOptStat(0);
+
+  // Draw the plot.
+  TH2D *h = new TH2D("", "", 25, 0, 3.9, 25, 0, 3.9); TRandom r; 
+  for (int i = 0; i < 30000; i++) h->Fill(r.Gaus(2.,1), r.Gaus(2.,1));
+  h->GetXaxis()->CenterTitle(); h->GetXaxis()->SetTitleOffset(2.5);
+  h->GetYaxis()->CenterTitle(); h->GetYaxis()->SetTitleOffset(2.5);
+  h->GetXaxis()->SetTitle("\\larger[2]$x$"); 
+  h->GetYaxis()->SetTitle("\\larger[2]$y$");
+  h->Draw("LEGO2");
+
+  // Draw additional text.
+  TText *t = new TText(); t->SetTextAlign(31);
+  t->DrawText(0.7, 0.9, "\\larger[2]$z(x,y) =  \\frac{1}{\\sigma_x\\sigma_y"
+	      "\\sqrt{4\\pi^2}}\\exp\\left(- \\left(\\frac{(x-\\mu_x)^2}"
+	      "{2\\sigma_x^2} + \\frac{(y-\\mu_y)^2}{2\\sigma_y^2} \\right)"
+	      "\\right)$");
+
+  // Print the plot.
+  gPad->Print("root.svg");
+  
+}

### 2017-06-13 paper.tex
Added figure environment. Minor changes.

@@
 \maketitle
+\begin{abstract}
+This paper classifies the periodic and drift-periodic trajectories on the Necker cube surface, a periodic surface built out of squares popular in optical illusions.
+\end{abstract}
 
+\begin{figure}[b!]
 \begin{center}
@@
 \end{center}
-
-\begin{abstract}
-
-\end{abstract}
+\caption{The Necker cube surface.}
+\label{fig:Necker}
+\end{figure}
 
@@
 
-\subsection{Background}
 The purpose of the following paper is to study the non-compact, infinite surface generated entirely out of the edge-to-edge gluing of square faces, whose final product resembles cubes stacked in a cascading, pyramidal pattern (or perhaps a rhombile tiling).
-(See Fig //) M.C.
+(See Figure \ref{fig:Necker}) M.C.
 Escher famously adapted Louis Albert Necker's optical illusion known as the Necker Cube in 1937.
@@
 
+\begin{figure}
 \begin{center}
 \includegraphics[scale=0.5]{escher.jpg}
-\\ 
-\text{The Necker Cube as it appears briefly in Metamorphosis I}
+\caption{The Necker Cube as it appears briefly in Metamorphosis I}
+\label{fig:Escher}
 \end{center}
+\end{figure}
 
@@
 \\content...
+\begin{enumerate}
+\item Statement 1
+\item Statement 2
+\end{enumerate}
 \end{thm*}
+
+\subsection{Discussion of Proof}
+\subsection{Related work}
 

### 2017-06-20 paper.tex
62017

@@
 \documentclass[]{article}
+
 \usepackage{amsfonts}
@@
 \usepackage{svg}
+\usepackage{enumitem}
+\usepackage{color}
+\usepackage{float}
+
 \graphicspath{{images/}}
 \setsvg{svgpath={./images/}}
+
 \newtheorem{thm}{Theorem}[subsection]
+\newtheorem{Def}{Definition}[subsection]
 \newtheorem*{thm*}{Theorem}
+\newtheorem*{def*}{Definition}
+
+\newcommand{\compav}[1]{\textbf{\textcolor{blue}{#1}}}
+\newcommand{\compat}[1]{\textbf{\textcolor{red}{#1}}}
+\newcommand{\shiftleft}[2]{\makebox[0pt][r]{\makebox[#1][l]{#2}}}
+\newcommand{\tilda}{\big{\char"7E}}
 
@@
 \maketitle
+
+\begin{center}
+
+\includesvg[width=4.8in]{cubecoverphoto}\\
+\end{center}
+
 \begin{abstract}
@@
 
-\begin{figure}[b!]
-\begin{center}
-\includesvg[width=2in]{cubesxyz}
-\end{center}
-\caption{The Necker cube surface.}
-\label{fig:Necker}
-\end{figure}
-
+\newpage
 \section{Introduction}
 
-The purpose of the following paper is to study the non-compact, infinite surface generated entirely out of the edge-to-edge gluing of square faces, whose final product resembles cubes stacked in a cascading, pyramidal pattern (or perhaps a rhombile tiling).
-(See Figure \ref{fig:Necker}) M.C.
-Escher famously adapted Louis Albert Necker's optical illusion known as the Necker Cube in 1937.
-The cubes appear in Escher's lithograph, "Metamorphosis I," in the form of tessellated rhombi.[CITE METAMORPHOSIS I] The print features a number of various tessellations transforming into simpler ones when traversed from left to right.
-The term Necker Cube is used interchangeably to refer to Escher's surface and Necker's optical illusion.
-For the purposes of the following paper, we will be using Necker Cube to refer to the infinite cube tiling that appears in Escher's work, as well as many other places.
-Those familiar with the arcade game known as Q*bert will most likely recognize the Necker Cube as the board whose colors you were meant to change in order to complete each level.
+The purpose of the following paper is to study the geodesic flow on the periodic surface depicted above, which is formed from an infinite collection of unit squares.
+We call this surface the Necker cube surface after Louis Albert Necker's "Necker Cube$_{1}$," which became a popular optical illusion.
+M.C.
+Escher's adaptation of the cube in Metamorphosis I$_{2}$ briefly features the Necker cube surface (See Figure $\ref{fig:Escher}$).The term Necker Cube is used interchangeably to refer to Escher's surface adaptation and Necker's optical illusion.
+For the purposes of the following paper, we will be using either Necker Cube or $\mathbf{S}$ to refer to the periodic cube tiling that appears in Escher's work.
 
-\begin{figure}
+\begin{figure}[H]
 \begin{center}
 \includegraphics[scale=0.5]{escher.jpg}
-\caption{The Necker Cube as it appears briefly in Metamorphosis I}
+\caption{The Necker cube as it appears briefly in Metamorphosis I}
 \label{fig:Escher}
@@
 \end{figure}
+
+The initial (rational) trajectories of any given flow fall into only one of two categories.
+
+\begin{def*}
+A vector, $(x,y)\in\mathbb{Z}^{2}$, is considered \textbf{odd-odd} if x and y are relatively prime and both odd. The \textbf{set of all odd-odd vectors} is given as 
+$$ \mathfrak{O} = \{(x,y)\in\mathbb{Z}^{2} : \gcd(x,y)=1\text{ and x,y are odd} \}.$$
+\end{def*}
+
+\begin{def*}
+A vector, $(x,y)\in\mathbb{Z}^{2}$, is considered \textbf{even-odd} if x and y are relatively prime and, of x and y, exactly one is odd and the other is even.
+The \textbf{set of all even-odd vectors} is given as
+$$ \mathfrak{E} = \{(x,y)\in\mathbb{Z}^{2} : \gcd(x,y)=1\text{ and x is odd iff y is even} \}.$$
+\end{def*}
+
+We take every vector on the Necker cube to be relative to the standard basis of the plane of intersection in space. From there we prove the following:
+
+\begin{thm*}{Denote $\phi^{\theta}_{t}$ to be the flow in a vector direction $\theta$ to time t on the Necker cube. Then the following is true:}
+\begin{enumerate}[label=(\roman*)]
+\item If $\theta$ belongs to $\mathfrak{O}$, then $\phi^{\theta}_{t}$ is periodic on $\mathbf{S}$.
+
+\item If $\theta$ belongs to $\mathfrak{E}$, then $\phi^{\theta}_{t}$ is drift-periodic on $\mathbf{S}$.
+
+\item If $\theta$ is of the form (x,y) such that x or y are irrational, then $\phi^{\theta}_{t}$ is uniquely ergodic on $\mathbf{S}$.
+
+\end{enumerate}
+\end{thm*}
+
+\compav{Stopped here. Continued in Section 2.}
 
@@
 We aim to prove the following:
-
-\begin{thm*}{Theorem}
-\\content...
-\begin{enumerate}
-\item Statement 1
-\item Statement 2
-\end{enumerate}
-\end{thm*}
 
@@
 
-\section{Constructing the Surface}
-This section will detail how the Necker Cube is constructed out of an infinite number of squares, and explain the process by which we flatten the structure in such a way that preserves the metric, and makes it simpler for us to describe the behavior of geodesics at these edges.
+\section{Construction the Surface \& Isometric Variant}
+\compav{Started up here again.}
 
-\subsection{The Necker Cube Surface and Isometric Variant }
+This section will detail how the Necker cube (See Figure \ref{fig:cubexyz}) is constructed out of an infinite set of unit squares, and explain the process by which we flatten the structure in such a way that preserves distances, and makes it simpler for us to describe the behavior of the geodesic flows that come into contact with these edges.
+
+\begin{figure}[H]
+\begin{center}
+\includesvg[width=3in]{cubesxyz}
+\caption{The Necker cube embedded in space.}
+\label{fig:cubexyz}
+\end{center}
+\end{figure}
+
+\subsection{The Necker Cube Surface and its Flattened Form}
 To build the surface out of infinitely many squares, we take the union of the instances of three squares, each lying on planes orthogonal to the other two and sharing an edge with exactly one other square.
@@
 They will be arranged in such a way that every integer triple $m,n,p\in\mathbb{Z}$ will have this surface associated with it.
-To each face we ascribe a letter A, B, or C.
-This will be useful when defining the flattening map.
-\begin{gather*}
+To each face we ascribe a letter A, B, or C, and construct an instance as a union of these faces:
+
+\vspace{0.2in}
+\begin{tabular}{p{10cm}c}
+\begin{align*}
 \chi_{m,n,p} = ([m, m+1]\times[n,n+1]\times\{p\}) 
@@
 \\\cup ([m,m+1]\times\{n+1\}\times[p,p-1])
-\end{gather*}
+\end{align*}
+&
+\shiftleft{0.8in}{\raisebox{-1in}{\includegraphics[scale=1]{label.png}}}
+\end{tabular}
+\vspace{0.2in}
 
 is the family of these instances based at some integer triple.
-The surface will be comprised of the union of the members that satisfy the property, $m+n+p=0$.
-From here on we will refer to the Necker Cube surface embedded form in $\mathbb{R}^{3}$ lying on the plane $z=-(x+y) $ as $\mathbf{C}$.
+The surface will be comprised of the members that satisfy the property, $m+n+p=0$.
+From here on we will refer to the Necker cube surface's embedded form in $\mathbb{R}^{3}$ lying along the plane $z=-(x+y) $ as $\mathbf{S}$.
 That is to say that
 
-$$\mathbf{C} =\bigcup\big{\{}\chi_{m,n,p}  : m+n+p=0 \big{\}} .$$
+$$\mathbf{S} =\bigcup\big{\{}\chi_{m,n,p}  : m+n+p=0 \big{\}}.$$
 
-$\mathbf{C}$ now resembles the structure in (FIG).
-Before describing the transformation that takes $\mathbf{C}$ to its flattened form, it would be best to define the relations on a subset of the plane that preserves the identification of edges.
+$\mathbf{S}$ now resembles the structure in Figure \ref{fig:cubexyz}.
+Before describing the transformation that takes $\mathbf{S}$ to its flattened form, it would be best to define the relations on a subset of the plane that preserves the identification of edges.
 We will call this surface $\mathbf{U}'$.
-However, the flattening of the surface takes $\mathbf{C}$ to a translation of $\mathbf{U}'$, we call $\mathbf{U}$ (See FIG).
-We will denote the metric preserving map from $\mathbf{C}$ to $\mathbf{U}$, by $\Psi$.
-Defining these edge identifying relations over $\mathbf{U}'$ is more convenient for us than defining them over $\mathbf{U}$, but defining $\Psi$ is more convenient when the map is taken to $\mathbf{U}$.
-Later on we will use $\mathbf{U}'$ when taking its quotient space.
+However, the flattening of the surface takes $\mathbf{S}$ to a translation of $\mathbf{U}'$, we call $\mathbf{U}$ (See $\ref{fig:uprimeu}$).
+We will denote the metric preserving map from $\mathbf{S}$ to $\mathbf{U}$, by $\Psi$.
+Defining these edge identifying relations over $\mathbf{U}'$ is more convenient for us than defining them over $\mathbf{U}$, but defining $\Psi$ is more convenient when the map is taken to $\mathbf{U}$ due to the nature of the construction of $\mathbf{S}$.
 
-Now we will describe exactly how to construct $\mathbf{U}'$.
-Take the subset of the real plane, $\mathbf{P} = \mathbb{R}^{2}-\{(u,v)|2m-\frac{1}{2}<u<2m+\frac{1}{2}, 2n-\frac{1}{2}<v<2n+\frac{1}{2} \}$, with 1$\times$1 open squares cut out at every even integer pair and let $\mathbf{U}'$ represent the flattened structure with the proper edge identifications with the first "hole" centered at the origin.
-We will describe this surface as a product of the auxiliary plane,$z=0$ and the image of a natural map from $\mathbf{P}$ to its quotient space with the following relation on $\mathbf{P}$:
+\begin{figure}[H]
+\begin{center}
+\includesvg[width=1.5 in]{u}\hspace{0.6in}\includesvg[width=1.5in]{uprime}
+\end{center}
+\caption{$\mathbf{U}$ (left) and $\mathbf{U}'$ (right) differ only by translation.}
+\label{fig:uprimeu}
+\end{figure}
+
+Now we will describe exactly how to construct $\mathbf{U}'$. Take the subset of the real plane, 
+$$\mathbf{P} = \mathbb{R}^{2}-\{(u,v):2m-\frac{1}{2}<u<2m+\frac{1}{2}, 2n-\frac{1}{2}<v<2n+\frac{1}{2} \},$$ with open unit squares cut out at every even integer pair and let $\mathbf{U}'$ represent the flattened structure with the proper edge identifications with the first "hole" centered at the origin.
+We will describe this surface as a cross product of the auxiliary plane, $z=0$ and the image of the natural map from $\mathbf{P}$ to its quotient space with the following relation on $\mathbf{P}$:
 
 \begin{gather*}
-\text{Let } x_{0}=(u_{0},v_{0}),x_{1}=(u_{1},v_{1}) \in \mathbf{P}. \text{Let } \mathbf{R} \text { be a relation described as } x_{0}\mathbf{R}x_{1} \\ \text{ iff } x_{0}=x_{1}
+\text{Let } x_{0}=(u_{0},v_{0}),x_{1}=(u_{1},v_{1}) \in \mathbf{P}. \text{Let } \mathbf{R} \text { be a relation given, } x_{0}\mathbf{R}x_{1} \\ \text{ iff } x_{0}=x_{1}
  \text{ or if }x_{0},x_{1} \in {\partial} \left( \left[2m-\frac{1}{2},2m+\frac{1}{2}\right] \times \left[2n-\frac{1}{2},2n+\frac{1}{2}\right] \right)\\
@@
 \end{array}\right]
-\\\text{For some }m,n\in\mathbb{Z} \end{gather*}
+\\\text{For some }m,n\in\mathbb{Z}. \end{gather*}
 
-The relation will either relate every point to itself, or (in the case where any two points lie on the boundary of the square shaped holes) take the center of each square shaped hole in the plane to the origin and reflect it over the line $y=x$ to match up these points accordingly.
+The relation will either identify every point to itself or, in the case where any two points lie on the boundary of the square shaped holes, take the center of each square hole in the plane to the origin and reflect it over the line $y=x$ to match up these points accordingly.
 Let $N: \mathbf{P} \rightarrow \mathbf{P}/\mathbf{R}$ denote the natural map from every point on $\mathbf{P}$ to its appropriate equivalence class.
@@
 The map $\mathbf{U}\rightarrow\mathbf{U}'$ is given by an affine translation of the surface $(x,y,z) \mapsto (x-\frac{3}{2},y-\frac{3}{2},z)$.
-(The point at the origin of $\mathbf{C}$ is mapped to $(-\frac{3}{2},-\frac{3}{2},0) \text{ on } \mathbf{U}'$.)
+(The point at the origin of $\mathbf{S}$ is mapped to $(-\frac{3}{2},-\frac{3}{2},0) \text{ on } \mathbf{U}'$).
 
-\subsection{Flattening $\mathbf{C}$}
+\compav{Stopped here. Continued in Sec. 3.}
+
+\newpage
+\subsection{Flattening $\mathbf{S}$}
 This section will describe the isomorphic map $\Psi:\mathbf{C}\rightarrow\mathbf{U}$, its inverse, $\Psi^{-1}:\mathbf{U}\rightarrow\mathbf{C}$, and prove that $\mathbf{C}$ and $\mathbf{U}$ are isometric.
@@
 \includesvg[width=2.3in]{cubesxyzcut}
-\includesvg[width=2.3in]{unfoldcut}
-\text{The images of } $\Psi^{-1}(\mathbf{U})$ \text{ and } $\Psi(\mathbf{C})$, \text{ respectively.}
+\raisebox{0.4in}{\includesvg[width=2.3in]{u}}
+\text{The images of } $\Psi^{-1}(\mathbf{U})$ (left)\text{ and }$\Psi(\mathbf{S})$ (right)
 \end{center}
@@
 \end{gather*}
-
-\includegraphics[]{label.png}\centering
 
@@
 \begin{center}
-\includesvg[width=2.3in]{cubesxyzpaths} 
-\includesvg[width=2.3in]{unfoldcutpath}
-\\ \text{Here we can see paths a,b, and c on }$\mathbf{U}$\text{ and }$\mathbf{C}$,
+\includesvg[width=2.3in]{cubesxyzpaths}\hspace{0.1in}
+\raisebox{0.5in}{\includesvg[width=2.2in]{unfoldcutpath}}
+\\ \text{Here we can see paths a,b, and c on }$\mathbf{U}$\text{ and }$\mathbf{S}$,
 \\ \text{where their initial trajectories are (0,1), (1,0), and (1,1) respectively.}
@@
 
+\newpage
 \section{The Four-Fold Cover and Translation Surface}
+\compav{Started here again.}
+The following section is concerned with defining a quotient space on $\mathbf{U}'$ (See $\ref{fig:quotient}$), and using that space to construct our translation surface.
+In addition we will describe a cover made from four copies of $\mathbf{U}'$ that aligns geodesic flow and behaves as a cover of our translation surface.
+
+We will call this quotient space $\mathbf{M}$ and define it as such:
+\begin{Def}{$\mathbf{M}$.}\\
+$\mathbf{M}$ is a base surface of $\mathbf{U}'$ given by the quotient $\mathbf{U}'/2\mathbb{Z}\oplus2\mathbb{Z}$, where $2\mathbb{Z}\oplus2\mathbb{Z}\times\mathbf{U}'\rightarrow\mathbf{U}'$ is the group action on $\mathbf{U}'$ defined as $(m,n)\circ(u,v)=(u+m,v+n)$.
+\end{Def}
+
+\begin{figure}[H]
+\begin{center}
+\includesvg[width=2in.]{unfoldcutquotient}\hspace{0.2in}
+\includesvg[width=2in.]{quotient}
+\end{center}
+\caption{$\mathbf{U}'$ equivalence classes under group action (left), and $\mathbf{M}$ with identified edges and vertices (right).}
+\label{fig:quotient}
+\end{figure}
+
+The quotient space is a genus 1 surface that resembles a $2\times2$ torus with a unit square removed from it.
+Near the outer vertex it behaves as a torus might.
+The inner vertices $c_{1},c_{3}$ have cone angle $\frac{3\pi}{2}$, and $c_{2}$ has a cone angle of $3\pi$.
+None of these vertices are integral multiples of $2\pi$, so in order to create a translation surface we have to take a minimal number of copies of the quotient space so every vertex has a cone angle that is a multiple of $2\pi$.
+
+\newpage
+\subsection{Translation Surface}
+With $\mathbf{M}$ defined, we can define a branch cover of $\mathbf{M}$ composed of four copies.
+To each copy we associate a direction of the flow under indices \{0,1,2,3\}.
+Let $\mathbf{M}_{0} $ denote the quotient space with its vertices removed:
+
+\begin{center}
+\includesvg[width=2in]{branchvert}
+\end{center}
+
+Now denote the set of singularities of $\mathbf{M}_{0}$ by $\Sigma$, where $\Sigma=\{s_{i}:i=1,...,8\}$.
+We let $ H_{1} (\mathbf{M}_{0}, \Sigma ; \mathbb{Z} / 4 \mathbb{Z}) =\{\gamma,e\}$ be the homology group
+
+\compav{Little stuck on how to actually write this out.}
+
+\begin{figure}[H]
+\begin{center}
+\includesvg[width=4in]{branchcover}
+\end{center}
+\end{figure}
+
+\begin{figure}[H]
+\begin{center}
+\includesvg[width=4in]{branchcoverchopped}
+\end{center}
+\end{figure}
 

### 2017-06-20 Fig.1.tex
62017a

\documentclass []{article}
\usepackage {amsfonts}
\usepackage {amsmath}
\usepackage {graphicx}
\usepackage {amsthm}
\usepackage {svg}
\graphicspath {{images/}}
\setsvg {svgpath={./images/}}
\newtheorem {thm}{Theorem}[subsection]
\newtheorem *{thm*}{Theorem}

\title {Geodesic Flow on the Necker Cube Surface}
\date {}
\author {Pavel Javornik}

\makeatletter \@ifpackageloaded {import}{}{\RequirePackage {import}}\@ifpackageloaded {graphicx}{}{\RequirePackage {graphicx}}\@ifpackageloaded {transparent}{}{\RequirePackage {transparent}}\@ifpackageloaded {xcolor}{}{\RequirePackage {xcolor}}\makeatother 
\setlength {\pdfpagewidth }{166.2212pt}
\setlength {\pdfpageheight }{166.2148pt}
\setlength {\paperheight }{\pdfpageheight }
\setlength {\paperwidth }{\pdfpagewidth }
\setlength {\textheight }{\paperheight }
\setlength {\textwidth }{\paperwidth }
\setlength {\textheight }{\paperheight }
\setlength {\oddsidemargin }{-1in}
\setlength {\evensidemargin }{-1in}
\setlength {\topmargin }{-1in}
\setlength {\headheight }{0in}
\setlength {\headsep }{0in}
\setlength {\topskip }{0in}
\setlength {\footskip }{0in}
\setlength {\parindent }{0in}
\setlength {\parsep }{0in}
\setlength {\parskip }{0in}
\begin {document}
\pagestyle {empty}\begin {center}\def \svgwidth {0.99\textwidth }\import {./images/}{unfoldcutpath.pdf_tex}\end {center}
\end {document}

### 2017-06-20 paper.tex
Comments.

@@
 The purpose of the following paper is to study the geodesic flow on the periodic surface depicted above, which is formed from an infinite collection of unit squares.
-We call this surface the Necker cube surface after Louis Albert Necker's "Necker Cube$_{1}$," which became a popular optical illusion.
+We call this surface the Necker cube surface after Louis Albert Necker's ``Necker Cube$_{1}$,'' which became a popular optical illusion.
 M.C.
@@
 The initial (rational) trajectories of any given flow fall into only one of two categories.
+\compat{The definitions below seem redundant. Perhaps you want ${\mathcal O}$ and ${\mathcal E}$}
 
@@
 
-\begin{thm*}{Denote $\phi^{\theta}_{t}$ to be the flow in a vector direction $\theta$ to time t on the Necker cube. Then the following is true:}
+\begin{thm*}{Denote $\phi^{\theta}_{t}$ to be the flow in a vector direction $\theta$ to time t on the Necker cube.
+Then the following is true:} \compat{There is no flow in a fixed direction because when trajectories move to new squares they may be traveling different directions.}
 \begin{enumerate}[label=(\roman*)]
@@
 \item If $\theta$ is of the form (x,y) such that x or y are irrational, then $\phi^{\theta}_{t}$ is uniquely ergodic on $\mathbf{S}$.
-
+\compat{We should move the discussion of ergodicity outside the theorem.}
 \end{enumerate}
@@
 
-\section{Construction the Surface \& Isometric Variant}
+\section{Constructing the Surface \& Isometric Variant}
 \compav{Started up here again.}
@@
 
-\subsection{The Necker Cube Surface and its Flattened Form}
+\subsection{The Necker Cube Surface}
+
 To build the surface out of infinitely many squares, we take the union of the instances of three squares, each lying on planes orthogonal to the other two and sharing an edge with exactly one other square.
@@
 \chi_{m,n,p} = ([m, m+1]\times[n,n+1]\times\{p\}) 
-\\\cup  (\{m+1\}\times[n,n+1]\times[p,p-1]) 
-\\\cup ([m,m+1]\times\{n+1\}\times[p,p-1])
+\\\cup  (\{m+1\}\times[n,n+1]\times[p-1,p]) 
+\\\cup ([m,m+1]\times\{n+1\}\times[p-1,p])
 \end{align*}
@@
 \vspace{0.2in}
+\compat{I would define $A_{m,n,p}$, etc.}
 
+\noindent \compat{Inserted noindent}
 is the family of these instances based at some integer triple.
 The surface will be comprised of the members that satisfy the property, $m+n+p=0$.
-From here on we will refer to the Necker cube surface's embedded form in $\mathbb{R}^{3}$ lying along the plane $z=-(x+y) $ as $\mathbf{S}$.
+From here on we will refer to the Necker cube surface's embedded form in $\mathbb{R}^{3}$ lying on the plane $z=-(x+y) $ as $\mathbf{S}$.
 That is to say that
@@
 $$\mathbf{S} =\bigcup\big{\{}\chi_{m,n,p}  : m+n+p=0 \big{\}}.$$
+$\mathbf{S}$ now resembles the structure in Figure \ref{fig:cubexyz}. \compat{I would be more definitive. Say $S$ {\em is} the structure...}
 
-$\mathbf{S}$ now resembles the structure in Figure \ref{fig:cubexyz}.
+\subsection{The Flattened Form}
+\compat{I would use as an introductory paragraph: ``We will show that $\mathbf S$ is isometric to a topological quotient of a subset of the plane.''}
 Before describing the transformation that takes $\mathbf{S}$ to its flattened form, it would be best to define the relations on a subset of the plane that preserves the identification of edges.
@@
 Defining these edge identifying relations over $\mathbf{U}'$ is more convenient for us than defining them over $\mathbf{U}$, but defining $\Psi$ is more convenient when the map is taken to $\mathbf{U}$ due to the nature of the construction of $\mathbf{S}$.
+\compat{I found this paragraph awkward.
+I would just go ahead and define the region in the plane.}
 
@@
 \end{center}
-\caption{$\mathbf{U}$ (left) and $\mathbf{U}'$ (right) differ only by translation.}
+\caption{$\mathbf{U}$ (left) and $\mathbf{U}'$ (right) differ only by translation. \compat{Perhaps just use one.}}
 \label{fig:uprimeu}
@@
 Now we will describe exactly how to construct $\mathbf{U}'$. Take the subset of the real plane, 
-$$\mathbf{P} = \mathbb{R}^{2}-\{(u,v):2m-\frac{1}{2}<u<2m+\frac{1}{2}, 2n-\frac{1}{2}<v<2n+\frac{1}{2} \},$$ with open unit squares cut out at every even integer pair and let $\mathbf{U}'$ represent the flattened structure with the proper edge identifications with the first "hole" centered at the origin.
+$$\mathbf{P} = \mathbb{R}^{2}-\bigcup_{m,n \in {\mathbb Z}} \{(u,v):2m-\frac{1}{2}<u<2m+\frac{1}{2}, 2n-\frac{1}{2}<v<2n+\frac{1}{2} \},$$ 
+\compat{I would write $(m-\frac{1}{2},m+\frac{1}{2}) \times (n-\frac{1}{2},n+\frac{1}{2})$.}
+with open unit squares cut out at every even integer pair and let $\mathbf{U}'$ represent the flattened structure with the proper edge identifications with the first "hole" centered at the origin.
 We will describe this surface as a cross product of the auxiliary plane, $z=0$ and the image of the natural map from $\mathbf{P}$ to its quotient space with the following relation on $\mathbf{P}$:
@@
 
-Now we will define $\Psi$ piece-wise.
+Now we will define $\Psi$ piecewise.
 
@@
 \compav{Started here again.}
-The following section is concerned with defining a quotient space on $\mathbf{U}'$ (See $\ref{fig:quotient}$), and using that space to construct our translation surface.
+The following section is concerned with defining a quotient of $\mathbf{U}'$ (see Figure $\ref{fig:quotient}$), and using that space to construct a translation surface.
 In addition we will describe a cover made from four copies of $\mathbf{U}'$ that aligns geodesic flow and behaves as a cover of our translation surface.
+\compat{Revisit to make sure it summarizes the section}
+
+\subsection{The Euclidean cone surface $\mathbf{M}$}
+\compat{I would say: Observe that translations of ${\mathbb R}^2$ by pairs even integers preserves the subset ${\mathbf P}$ and respects the equivalence relation defining ${\mathbf U'}$.
+Thus this gives an action of $(2{\mathbb Z})^2$ on ${\mathbf U}'$.
+We define ${\mathbf M}$ to be the quotient.
+It is depicted in Figure ...}
 
@@
 None of these vertices are integral multiples of $2\pi$, so in order to create a translation surface we have to take a minimal number of copies of the quotient space so every vertex has a cone angle that is a multiple of $2\pi$.
+\compat{Move the discussion of building the translation surface to the next section.
+Otherwise this paragraph is good.}
 
@@
 \subsection{Translation Surface}
-With $\mathbf{M}$ defined, we can define a branch cover of $\mathbf{M}$ composed of four copies.
+\compat{You need to build the surface out of pieces rather than out of ``copies of $\mathbf{M}$.''} 
+We will now define a translation which is a four-fold branch cover of $\mathbf{M}$.
 To each copy we associate a direction of the flow under indices \{0,1,2,3\}.

### 2017-06-22 paper.tex
62117

@@
 \begin{def*}
-A vector, $(x,y)\in\mathbb{Z}^{2}$, is considered \textbf{odd-odd} if x and y are relatively prime and both odd. The \textbf{set of all odd-odd vectors} is given as 
-$$ \mathfrak{O} = \{(x,y)\in\mathbb{Z}^{2} : \gcd(x,y)=1\text{ and x,y are odd} \}.$$
+Let $v$ be a vector of the form $(x,y)\in\mathbb{Z}^{2}$.
+We call $v$ an \textbf{odd-odd} vector if x and y are relatively prime and both odd.
+We denote the \textbf{set of all odd-odd vectors} by $\mathcal{O}$.
+We say that $v$ is an  \textbf{even-odd} vector if x and y are relatively prime, and x is even if and only if y is odd.
+We denote the \textbf{set of all even-odd vectors} by $\mathcal{E}$.
 \end{def*}
 
-\begin{def*}
-A vector, $(x,y)\in\mathbb{Z}^{2}$, is considered \textbf{even-odd} if x and y are relatively prime and, of x and y, exactly one is odd and the other is even.
-The \textbf{set of all even-odd vectors} is given as
-$$ \mathfrak{E} = \{(x,y)\in\mathbb{Z}^{2} : \gcd(x,y)=1\text{ and x is odd iff y is even} \}.$$
-\end{def*}
+We consider every (rational) initial trajectory on the surface to fall into one of the two categories, otherwise it is a positive scalar multiple of some direction.
+Given some initial point $x_{0}$ on the surface embedded in real space and some vector direction in real space, $\bar{v}$, our initial trajectory,$v$, is taken as an isomorphism from the planar subspace of $\mathbb{R}^{3}$ that contains $\bar{v}$.
+The three possible cases are the auxiliary planes, $x=0, y=0, z=0$, whose projections onto $\mathbb{R}^{2}$ are given by $(0,y,z)\mapsto(-z,y)$, $(x,0,z)\mapsto(x,-z)$, and $(x,y,0)\mapsto(x,y)$ respectively and relative to the standard bases of $\mathbb{R}^{3}$.
+We take $-z$ to be positive, and consider vectors that lie outside of these subspaces as degenerate cases.
+From there we prove the following:
 
-We take every vector on the Necker cube to be relative to the standard basis of the plane of intersection in space. From there we prove the following:
+\begin{thm*}{Given an initial trajectory, $\theta\in\mathbb{R}^{2}$, and geodesic path, $\gamma$, on $\mathbf{S}$ with initial trajectory $\theta$, the following is true:} \compat{There is no flow in a fixed direction because when trajectories move to new squares they may be traveling different directions.}
+\compav{Can I use gamma as a path on S?}
+\begin{enumerate}[label=(\roman*)]
+\item If $\theta$ belongs to $\mathcal{O}$, then $\gamma$ is closed on $\mathbf{S}$ and homeomorphic to $\mathbb{R}/\mathbb{Z}$.
+\item If $\theta$ belongs to $\mathcal{E}$, then $\gamma$ is drift-periodic on $\mathbf{S}$.
 
-\begin{thm*}{Denote $\phi^{\theta}_{t}$ to be the flow in a vector direction $\theta$ to time t on the Necker cube.
-Then the following is true:} \compat{There is no flow in a fixed direction because when trajectories move to new squares they may be traveling different directions.}
-\begin{enumerate}[label=(\roman*)]
-\item If $\theta$ belongs to $\mathfrak{O}$, then $\phi^{\theta}_{t}$ is periodic on $\mathbf{S}$.
-
-\item If $\theta$ belongs to $\mathfrak{E}$, then $\phi^{\theta}_{t}$ is drift-periodic on $\mathbf{S}$.
-
-\item If $\theta$ is of the form (x,y) such that x or y are irrational, then $\phi^{\theta}_{t}$ is uniquely ergodic on $\mathbf{S}$.
-\compat{We should move the discussion of ergodicity outside the theorem.}
 \end{enumerate}
 \end{thm*}
+
+\compav{Archiving for later:\\If $\theta$ is of the form (x,y) such that x or y are irrational, then $\phi^{\theta}_{t}$ is uniquely ergodic on $\mathbf{S}$.
+\compat{We should move the discussion of ergodicity outside the theorem.}	}
 
@@
 \section{Constructing the Surface \& Isometric Variant}
-\compav{Started up here again.}
+\compav{Started up here again 6/21/17}
 
@@
 
-To build the surface out of infinitely many squares, we take the union of the instances of three squares, each lying on planes orthogonal to the other two and sharing an edge with exactly one other square.
-The three squares have one vertex in common.
+To build the surface out of infinitely many unit squares, we take the union of the instances of three squares, each lying on planes orthogonal to the other two and sharing an edge with exactly one other square.
+The three squares have only one vertex in common.
 
-Every instance can be described as cube whose bottom, left, and rear faces have been removed.
-They will be arranged in such a way that every integer triple $m,n,p\in\mathbb{Z}$ will have this surface associated with it.
-To each face we ascribe a letter A, B, or C, and construct an instance as a union of these faces:
+First we divide the planes $x,y,z=k$ for some integer $k$ in such a way that they are composed entirely out of unit squares.
+We want to index each of these squares by integer triples and some label A,B, or C.
+i.e.
+squares on planes of the form $z=k$ will have an A label, $x=k$ a B label, and $y=k$ a C label.
+Let $m,n,p\in\mathbb{Z}$.
+Then each square based at $(m,n,p)$ is defined as follows:
 
@@
 \begin{align*}
-\chi_{m,n,p} = ([m, m+1]\times[n,n+1]\times\{p\}) 
-\\\cup  (\{m+1\}\times[n,n+1]\times[p-1,p]) 
-\\\cup ([m,m+1]\times\{n+1\}\times[p-1,p])
+\mathbf{A}_{m,n,p} = [m, m+1]\times[n,n+1]\times\{p\} 
+\\\mathbf{B}_{m,n,p} = \{m+1\}\times[n,n+1]\times[p-1,p]
+\\\mathbf{C}_{m,n,p} [m,m+1]\times\{n+1\}\times[p-1,p]
 \end{align*}
@@
 \vspace{0.2in}
-\compat{I would define $A_{m,n,p}$, etc.}
+\compat{I would define $A_{m,n,p}$, etc.}\\
+\noindent For the sake of convenience, B and C are shifted in order to associate each integer triple that satisfies the linear condition $m+n+p=0$ with an instance of the cube that $\mathbf{S}$ is composed of.
+Furthermore, we can separate $\mathbf{S}$ into the sets:
 
-\noindent \compat{Inserted noindent}
-is the family of these instances based at some integer triple.
-The surface will be comprised of the members that satisfy the property, $m+n+p=0$.
-From here on we will refer to the Necker cube surface's embedded form in $\mathbb{R}^{3}$ lying on the plane $z=-(x+y) $ as $\mathbf{S}$.
-That is to say that
+\begin{align*}
+\mathbf{A}=\bigcup\big{\{}\mathbf{A}_{m,n,p}: m+n+p=0\big{\}}
+\\\mathbf{B}=\bigcup\big{\{}\mathbf{B}_{m,n,p}: m+n+p=0\big{\}}
+\\\mathbf{C}=\bigcup\big{\{}\mathbf{C}_{m,n,p}: m+n+p=0\big{\}}
+\end{align*}
 
-$$\mathbf{S} =\bigcup\big{\{}\chi_{m,n,p}  : m+n+p=0 \big{\}}.$$
-$\mathbf{S}$ now resembles the structure in Figure \ref{fig:cubexyz}. \compat{I would be more definitive. Say $S$ {\em is} the structure...}
+\noindent We can now give a formal definition of the Necker Cube.
+
+\begin{Def}{The Necker Cube.} 
+\\ The \textbf{Necker Cube}, denoted by $\mathbf{S}$, is the surface made up of unions of unit squares of the form $\mathbf{S} = \mathbf{A}\cup\mathbf{B}\cup\mathbf{C}$.\end{Def}
+
+\noindent $\mathbf{S}$ is the surface portrayed in Figure \ref{fig:cubexyz}.
 
 \subsection{The Flattened Form}
+We will show that $\mathbf S$ is isometric to a a topological quotient of a particular subset of the plane.
+We call this surface $\mathbf{U}$, and the subset of the plane $\mathbf P$ (See Figure $\ref{fig:uprimeu}$).
+
+\begin{Def}{$\mathbf{P}$}\\
+$\mathbf P$ is the region of the plane with unit squares removed:
+$$\mathbf{P} = \mathbb{R}^{2}\text{ }\backslash\bigcup_{m,n \in {\mathbb Z}} \big{\{}(u,v):u\in(2m-\frac{1}{2},2m+\frac{1}{2}),\text{ } v\in(2n-\frac{1}{2},2n+\frac{1}{2})\big{\}}$$
+\end{Def}
+
+\begin{figure}[H]
+\begin{center}
+\includesvg[width=1.5in]{uprime}
+\end{center}
+\caption{$\mathbf P$, the subset of the plane with unit squares removed at every even integer pair.} \compat{Perhaps just use one.}
+\label{fig:uprimeu}
+\end{figure}
+
+\noindent On $\mathbf P$ we define a relation that identifies the sides of the boundaries of each hole in the plane with another by rotating an edge around the bottom left or top right corners by $\pm\frac{\pi}{2}$.
+The result is an edge identification as seen here (centered at each even coordinate pair):
+
+\begin{figure}[H]\centering
+\includesvg[width=2in.]{quotient}
+\end{figure}
+\noindent We call this relation $\mathbf{R}$.
+What $\mathbf{R}$ essentially does is translates every hole to the origin and identifies the edges by reflecting over the line $y=x$.
+If these points coincide, then they are indeed the same point on the boundary.
+Otherwise, every point that is not on the boundary belongs to its own class.
+We can safely assume this wont identify edges improperly, since we translate every boundary so as to situate the unit holes over the origin.
+For example, if we took two edges that do not belong to the same hole then by that same translation and reflection the two edges couldn't possible coincide.
+
+\begin{Def}
+content...
+\end{Def}
+
+\begin{Def}{$\mathbf U$}\\
+$\mathbf U$ is equal to the quotient space $\mathbf{P}/\mathbf{R}\times\{0\}$
+\end{Def}
+
+\compav{Stopped here. 6/21/17}
+
 \compat{I would use as an introductory paragraph: ``We will show that $\mathbf S$ is isometric to a topological quotient of a subset of the plane.''}
+
 Before describing the transformation that takes $\mathbf{S}$ to its flattened form, it would be best to define the relations on a subset of the plane that preserves the identification of edges.
@@
 
-\begin{figure}[H]
-\begin{center}
-\includesvg[width=1.5 in]{u}\hspace{0.6in}\includesvg[width=1.5in]{uprime}
-\end{center}
-\caption{$\mathbf{U}$ (left) and $\mathbf{U}'$ (right) differ only by translation. \compat{Perhaps just use one.}}
-\label{fig:uprimeu}
-\end{figure}
-
 Now we will describe exactly how to construct $\mathbf{U}'$. Take the subset of the real plane, 
-$$\mathbf{P} = \mathbb{R}^{2}-\bigcup_{m,n \in {\mathbb Z}} \{(u,v):2m-\frac{1}{2}<u<2m+\frac{1}{2}, 2n-\frac{1}{2}<v<2n+\frac{1}{2} \},$$ 
+$$,$$ 
 \compat{I would write $(m-\frac{1}{2},m+\frac{1}{2}) \times (n-\frac{1}{2},n+\frac{1}{2})$.}

### 2017-06-22 paper.tex
Comments/minor changes.

@@
 \begin{align*}
-\mathbf{A}_{m,n,p} = [m, m+1]\times[n,n+1]\times\{p\} 
-\\\mathbf{B}_{m,n,p} = \{m+1\}\times[n,n+1]\times[p-1,p]
-\\\mathbf{C}_{m,n,p} [m,m+1]\times\{n+1\}\times[p-1,p]
+\mathbf{A}_{m,n,p} = [m, m+1]\times[n,n+1]\times\{p\}, 
+\\\mathbf{B}_{m,n,p} = \{m+1\}\times[n,n+1]\times[p-1,p],
+\\\mathbf{C}_{m,n,p}= [m,m+1]\times\{n+1\}\times[p-1,p].
 \end{align*}
@@
 \end{tabular}
-\vspace{0.2in}
-\compat{I would define $A_{m,n,p}$, etc.}\\
+\vspace{0.2in}\\
 \noindent For the sake of convenience, B and C are shifted in order to associate each integer triple that satisfies the linear condition $m+n+p=0$ with an instance of the cube that $\mathbf{S}$ is composed of.
-Furthermore, we can separate $\mathbf{S}$ into the sets:
-
+Furthermore, we  separate $\mathbf{S}$ into the sets:
 \begin{align*}
-\mathbf{A}=\bigcup\big{\{}\mathbf{A}_{m,n,p}: m+n+p=0\big{\}}
-\\\mathbf{B}=\bigcup\big{\{}\mathbf{B}_{m,n,p}: m+n+p=0\big{\}}
-\\\mathbf{C}=\bigcup\big{\{}\mathbf{C}_{m,n,p}: m+n+p=0\big{\}}
+\mathbf{A}=\bigcup\big{\{}\mathbf{A}_{m,n,p}: m+n+p=0\big{\}},
+\\\mathbf{B}=\bigcup\big{\{}\mathbf{B}_{m,n,p}: m+n+p=0\big{\}},
+\\\mathbf{C}=\bigcup\big{\{}\mathbf{C}_{m,n,p}: m+n+p=0\big{\}}.
 \end{align*}
@@
 
-\begin{Def}{The Necker Cube.} 
-\\ The \textbf{Necker Cube}, denoted by $\mathbf{S}$, is the surface made up of unions of unit squares of the form $\mathbf{S} = \mathbf{A}\cup\mathbf{B}\cup\mathbf{C}$.\end{Def}
+\begin{Def} The \textbf{Necker Cube}, denoted by $\mathbf{S}$, is the surface made up of unions of unit squares of the form $\mathbf{S} = \mathbf{A}\cup\mathbf{B}\cup\mathbf{C}$.\end{Def}
 
@@
 We will show that $\mathbf S$ is isometric to a a topological quotient of a particular subset of the plane.
-We call this surface $\mathbf{U}$, and the subset of the plane $\mathbf P$ (See Figure $\ref{fig:uprimeu}$).
+We call this quotient $\mathbf{U}$, and the subset of the plane $\mathbf P$ (See Figure $\ref{fig:uprimeu}$).
+The subset $\mathbf P$ is the region of the plane with unit squares removed:
+\begin{equation}
+\label{eq:P}
+\mathbf{P} = \mathbb{R}^{2}\text{ }\backslash\bigcup_{m,n \in {\mathbb Z}} \big{\{}(u,v):u\in(2m-\frac{1}{2},2m+\frac{1}{2}),\text{ } v\in(2n-\frac{1}{2},2n+\frac{1}{2})\big{\}}.
+\end{equation}
 
-\begin{Def}{$\mathbf{P}$}\\
-$\mathbf P$ is the region of the plane with unit squares removed:
-$$\mathbf{P} = \mathbb{R}^{2}\text{ }\backslash\bigcup_{m,n \in {\mathbb Z}} \big{\{}(u,v):u\in(2m-\frac{1}{2},2m+\frac{1}{2}),\text{ } v\in(2n-\frac{1}{2},2n+\frac{1}{2})\big{\}}$$
-\end{Def}
+\compat{Illustrating refering to equations: Recall $\mathbf{P}$ was defined in \eqref{eq:P}.}
 
@@
 \end{center}
-\caption{$\mathbf P$, the subset of the plane with unit squares removed at every even integer pair.} \compat{Perhaps just use one.}
+\caption{The subset $\mathbf P$ of the plane with unit squares removed at every even integer pair.} \compat{Perhaps just use one.}
 \label{fig:uprimeu}
@@
 
-\noindent On $\mathbf P$ we define a relation that identifies the sides of the boundaries of each hole in the plane with another by rotating an edge around the bottom left or top right corners by $\pm\frac{\pi}{2}$.
-The result is an edge identification as seen here (centered at each even coordinate pair):
+\noindent On $\mathbf P$ we define a relation $\sim$ that identifies the sides of the boundaries of each hole in pairs.
+Each side is identified with another side of the same hole by rotation about the bottom left or top right corners of the hole by $\pm\frac{\pi}{2}$.
+The result is an edge identification as seen here (centered at each even coordinate pair): \compat{I think overall this is very good and completely defines the relation.
+You may want to tweak it to make it even more clear.
+You expressed some concerns about the vertices.}
 
@@
 \noindent We call this relation $\mathbf{R}$.
-What $\mathbf{R}$ essentially does is translates every hole to the origin and identifies the edges by reflecting over the line $y=x$.
+\compat{How about $\sim$?} What $\mathbf{R}$ essentially does is translates every hole to the origin and identifies the edges by reflecting over the line $y=x$.
 If these points coincide, then they are indeed the same point on the boundary.
@@
 For example, if we took two edges that do not belong to the same hole then by that same translation and reflection the two edges couldn't possible coincide.
+\compat{This paragraph doesn't seem to add much.}
 

### 2017-07-01 paper.tex
63017

@@
 \subsection{The Euclidean cone surface $\mathbf{M}$}
-\compat{I would say: Observe that translations of ${\mathbb R}^2$ by pairs even integers preserves the subset ${\mathbf P}$ and respects the equivalence relation defining ${\mathbf U'}$.
+\compat{I would say: Observe that translations of ${\mathbb R}^2$ by pairs even integers preserves the subset ${\mathbf P}$ and respects the equivalence relation defining ${\mathbf U}$.
 Thus this gives an action of $(2{\mathbb Z})^2$ on ${\mathbf U}'$.
@@
 It is depicted in Figure ...}
+
+Observe that the translations by pairs of even integers preserves the subset $\mathbf P$ and respects the equivalence relation on $\mathbf U$
 
@@
 The inner vertices $c_{1},c_{3}$ have cone angle $\frac{3\pi}{2}$, and $c_{2}$ has a cone angle of $3\pi$.
-None of these vertices are integral multiples of $2\pi$, so in order to create a translation surface we have to take a minimal number of copies of the quotient space so every vertex has a cone angle that is a multiple of $2\pi$.
+None of these inner vertices have cone angles that are integral multiples of $2\pi$, so in order to create the necessary translation surfaces we require the minimal number of copies of both $\mathbf{U}$ and $\mathbf{M}$ to construct their associated phase spaces.
 \compat{Move the discussion of building the translation surface to the next section.
 Otherwise this paragraph is good.}
+
+\subsection{Four-Fold Cover of $\mathbf{U}$}
+A geodesic on our surface can travel in one of four possible directions at any given time.
+To each phase we associate a numeric label I,II,II, or IV, where I contains the initial and terminal points of a closed geodesic.
+(Similar to [author]'s construction of phases of the Wind-Tree Model).
+For every even integer pair that is centered around every hole of $\mathbf{P}$, assign an integer pair (m.n).
+Identify the edges in the following manner:
+
+\begin{center}
+\includesvg[width=4in]{utilda}
+\end{center}
+This is a branched cover of $\mathbf{U}$ of degree four that can be thought of as four copies of $\mathbf{P}$, all stacked on one another with the following identifications:
+\begin{center}
+$([2m-\frac{1}{4}, 2m+\frac{1}{4}], 2n-\frac{1}{4},j)\sim(2m-\frac{1}{4},[2n-\frac{1}{4}, 2n+\frac{1}{4}],j+\bar{1})$\\
+\vspace{0.2in}$([2m-\frac{1}{4}, 2m+\frac{1}{4}], 2n+\frac{1}{4},j)\sim(2m+\frac{1}{4},[2n-\frac{1}{4}, 2n+\frac{1}{4}],j+\bar{1})$\\
+\vspace{0.2in}For (m,n,j) $\in \mathbb{Z}^{2}\times\mathbb{Z}/4\mathbb{Z}$
+\end{center}
+There are four distinct vertices for each hole centered at (m,n) on  $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}/\sim$ of cone angle $6\pi$.
+To create  $\mathbf{U}$'s translation surface that still behaves as a proper cover of $\mathbf{U}$, we define a map  $p:(\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}/\sim) \rightarrow \tilde{\mathbf{U}}$ as
+
+\begin{equation}
+	p\left[\begin{array}{c}
+		u\\v\\z
+	\end{array}\right] =
+		\left[ \hspace{1mm} \begin{matrix}
+				1 & 0 \\
+				0 & -1 \\
+			\end{matrix}\hspace{1mm}\right]^{z}\left[\begin{array}{c}
+					u\\v
+				\end{array}\right]\times\{z\}.
+\end{equation}
+
+\compav{So I'm not sure if I can actually define p on}
+ $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}/\sim$
+
+p preserves the edge identification by affine transformations (rotating every phase plane about its origin according to its phase value).  $\tilde{\mathbf{U}}$ as a translation cover:
+
+\begin{figure}[H]
+\centering
+\includesvg[width=4in]{utildaprime}
+\end{figure}
+As one might expect, traveling in the same direction in every phase has a different effect on the integer pair (m,n) relative to first phase.
+A path in that sense is considered closed if it eventually returns to the exact same point, and periodic should it return to the same phase as well.
 
@@
 \compat{You need to build the surface out of pieces rather than out of ``copies of $\mathbf{M}$.''} 
-We will now define a translation which is a four-fold branch cover of $\mathbf{M}$.
-To each copy we associate a direction of the flow under indices \{0,1,2,3\}.
-Let $\mathbf{M}_{0} $ denote the quotient space with its vertices removed:
-
-\begin{center}
-\includesvg[width=2in]{branchvert}
-\end{center}
-
-Now denote the set of singularities of $\mathbf{M}_{0}$ by $\Sigma$, where $\Sigma=\{s_{i}:i=1,...,8\}$.
-We let $ H_{1} (\mathbf{M}_{0}, \Sigma ; \mathbb{Z} / 4 \mathbb{Z}) =\{\gamma,e\}$ be the homology group
-
-\compav{Little stuck on how to actually write this out.}
+With $\tilde{\mathbf{U}}$ defined, its topological quotient is given by $\tilde{\mathbf{M}}$ = $\tilde{\mathbf{U}}/(2\mathbb{Z})^{2}$.
 
 \begin{figure}[H]
-\begin{center}
-\includesvg[width=4in]{branchcover}
-\end{center}
+\centering
+\includesvg[width=4in]{mtilda}
+\end{figure}
+$\tilde{\mathbf{M}}$ is also a degree four branched cover of $\mathbf{M}$.
+A covering map $g:\tilde{\mathbf{M}} \rightarrow \mathbf{M}$ also individually rotates these squares about the centers of their holes according to their labels and maps their points to $\mathbf{M}$.
+
+$\tilde{\mathbf{M}}$ is a genus 5 translation surface with four singularities.
+
+\begin{figure}[H]
+\centering
+\includesvg[width=4in]{mtildastaircase}
 \end{figure}
@@
 \begin{figure}[H]
-\begin{center}
-\includesvg[width=4in]{branchcoverchopped}
-\end{center}
+\centering
+\includegraphics[width=4in]{cylinderdecomp.pdf}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includesvg[width=4in]{cylinderskew}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includesvg[pretex=\tiny, width=4in]{slopeonecylinder}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=2in]{commute.pdf}
 \end{figure}

### 2017-07-03 paper.tex
070217

@@
 A geodesic on our surface can travel in one of four possible directions at any given time.
-To each phase we associate a numeric label I,II,II, or IV, where I contains the initial and terminal points of a closed geodesic.
+To each trajectory phase of we associate a numeric label I,II,II, or IV, where I contains the initial and terminal points of a closed geodesic.
 (Similar to [author]'s construction of phases of the Wind-Tree Model).
-For every even integer pair that is centered around every hole of $\mathbf{P}$, assign an integer pair (m.n).
+For every even integer pair centered around every hole of $\mathbf{P}$, assign a pair of integers (m,n).
 Identify the edges in the following manner:
@@
 \end{center}
-This is a branched cover of $\mathbf{U}$ of degree four that can be thought of as four copies of $\mathbf{P}$, all stacked on one another with the following identifications:
-\begin{center}
-$([2m-\frac{1}{4}, 2m+\frac{1}{4}], 2n-\frac{1}{4},j)\sim(2m-\frac{1}{4},[2n-\frac{1}{4}, 2n+\frac{1}{4}],j+\bar{1})$\\
-\vspace{0.2in}$([2m-\frac{1}{4}, 2m+\frac{1}{4}], 2n+\frac{1}{4},j)\sim(2m+\frac{1}{4},[2n-\frac{1}{4}, 2n+\frac{1}{4}],j+\bar{1})$\\
-\vspace{0.2in}For (m,n,j) $\in \mathbb{Z}^{2}\times\mathbb{Z}/4\mathbb{Z}$
-\end{center}
-There are four distinct vertices for each hole centered at (m,n) on  $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}/\sim$ of cone angle $6\pi$.
-To create  $\mathbf{U}$'s translation surface that still behaves as a proper cover of $\mathbf{U}$, we define a map  $p:(\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}/\sim) \rightarrow \tilde{\mathbf{U}}$ as
+This is a branched cover of $\mathbf{U}$ of degree four that can be thought of as four copies of $\mathbf{P}$, all stacked on top of one another with the following identifications made:
+
+\begin{multline}
+([2m-\frac{1}{4}, 2m+\frac{1}{4}], 2n-\frac{1}{4},j)\sim(2m-\frac{1}{4},[2n-\frac{1}{4}, 2n+\frac{1}{4}],j+\bar{1})\\
+([2m-\frac{1}{4}, 2m+\frac{1}{4}], 2n+\frac{1}{4},j)\sim(2m+\frac{1}{4},[2n-\frac{1}{4}, 2n+\frac{1}{4}],j+\bar{1})\\
+\vspace{0.2in}For (m,n,j) \in \mathbb{Z}^{2}\times\mathbb{Z}/4\mathbb{Z}
+\label{eq:rel2}
+\end{multline}
+
+There are four distinct vertices for each hole centered at $(m,n,\mathbb{Z}/4\mathbb{Z})$ on  $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$ of cone angle $6\pi$ once the identifications have been made.
+Aligning the flow on  $\mathbf{U}$ requires a rotation on phase planes II,III, and IV by $-\frac{\pi}{2}, -\pi,$ and $ -\frac{3\pi}{2}$ in that order, about the origin of each respective plane.
+The choice in rotations are arbitrary as long as the edges are glued by translation.
+
+We express  $p:(\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}) \rightarrow (\mathbf{P}\times\mathbb{Z}/4\mathbb{Z})$ as
 
@@
 	p\left[\begin{array}{c}
-		u\\v\\z
+		u\\v
 	\end{array}\right] =
 		\left[ \hspace{1mm} \begin{matrix}
-				1 & 0 \\
-				0 & -1 \\
+				0 & 1\\
+				-1 & 0
 			\end{matrix}\hspace{1mm}\right]^{z}\left[\begin{array}{c}
 					u\\v
-				\end{array}\right]\times\{z\}.
+				\end{array}\right]\times\{\bar z\}
+\label{eq:p}
 \end{equation}
 
-\compav{So I'm not sure if I can actually define p on}
- $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}/\sim$
+\begin{Def} Let $\sim$ be the equivalence relation given at $\eqref{eq:rel2}$, and p be the automorphism from $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$ to itself at $\eqref{eq:p}$.
+Define the relation obtained from $\sim$ on the image of p by $\sim_{p} = \{(x_1,p(x_2)) : x_1 \sim x_2\}$.
+Then the $\mathbf{translation\text{ }surface}$  $\tilde{\mathbf{U}}$ = Im p$\slash\sim_{p}$.
+\end{Def}
 
-p preserves the edge identification by affine transformations (rotating every phase plane about its origin according to its phase value).  $\tilde{\mathbf{U}}$ as a translation cover:
+\begin{thm}{$\tilde{\mathbf{U}}$ is a branched cover of $\mathbf{U}$. }\newline
+
+\end{thm}
+
+p preserves the edge identification by affine transformations.  $\tilde{\mathbf{U}}$ as a translation cover:
 
@@
 \compat{You need to build the surface out of pieces rather than out of ``copies of $\mathbf{M}$.''} 
-With $\tilde{\mathbf{U}}$ defined, its topological quotient is given by $\tilde{\mathbf{M}}$ = $\tilde{\mathbf{U}}/(2\mathbb{Z})^{2}$.
+With $\tilde{\mathbf{U}}$ defined, its topological quotient is given as a union of phase planes modded out by translations.
+
+\begin{Def}
+$\tilde{\mathbf{M}}{}={}\left[\bigcup_{i=0}^{3} p(\mathbf{P},i)\slash(2\mathbb{Z})^{2} \right]/\sim_{p\vert_{X}}{}X{}={}(\left[-1,1\right]\times\left[-1,1\right])\times\mathbb{Z}\slash4\mathbb{Z}$
+\end{Def}
+
+$\tilde{\mathbf{M}}$ is also a degree four branched cover of $\mathbf{M}$.
+A covering map $\tilde{\mathbf{M}} \rightarrow \mathbf{M}$ individually rotates these squares about the centers of their holes according to their labels and maps their points to $\mathbf{M}$.
+This is a genus five translation surface with four singularities.
+$\tilde{\mathbf{M}}$ as a quotient of $\tilde{\mathbf{U}}$ can be represented as quotients of each plane:
 
@@
 \end{figure}
-$\tilde{\mathbf{M}}$ is also a degree four branched cover of $\mathbf{M}$.
-A covering map $g:\tilde{\mathbf{M}} \rightarrow \mathbf{M}$ also individually rotates these squares about the centers of their holes according to their labels and maps their points to $\mathbf{M}$.
 
-$\tilde{\mathbf{M}}$ is a genus 5 translation surface with four singularities.
-
-\begin{figure}[H]
-\centering
-\includesvg[width=4in]{mtildastaircase}
-\end{figure}
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=4in]{cylinderdecomp.pdf}
-\end{figure}
-
-\begin{figure}[H]
-\centering
-\includesvg[width=4in]{cylinderskew}
-\end{figure}
-
-\begin{figure}[H]
-\centering
-\includesvg[pretex=\tiny, width=4in]{slopeonecylinder}
-\end{figure}
+All in all, we have the following diagram:
 
@@
 
+\newpage
+
+\section{$\tilde{\mathbf{M}}$ as a Veech Surface}
+The following section studies the translation surface's various symmetries and associated Veech group.
+We use the Veech dichotomy to prove our main theorems that determine whether the geodesic is ultimately periodic or drift periodic.
+
+\subsection{The Staircase Model}
+At first glance, $\tilde{\mathbf{M}}$ does not seem to be a square-tiled surface.
+However, observe that the surface can be reconstructed by a sequence of cutting and gluing operations and realized in the following way:
+
+\begin{figure}[H]
+\centering
+\includesvg[width=4in]{mtildastaircase}
+\caption{$\tilde{\mathbf{M}}$ as a staircase with vertices and phases identified.}
+\end{figure}
+
+Every edge on the staircase model is glued by translation to another by at most a distance of 2. (Each square is unit). More importantly, the staircase has a very natural cylindrical decomposition.
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=4in]{cylinderdecomp.pdf}
+\caption{The staircase with edges identified.}
+\end{figure}
+
+The staircase is then decomposed into 6 uniform cylinders of moduli $\frac{1}{2}$ in both the vertical and horizontal directions.
+The modulus of the cylinder, $\mu$, is given by the ratio of the cylinder's circumference and width, $\mu{}={}w/c$.
+A surface that is composed entirely out of similar horizontal and vertical cylinders is stabilized by Dehn-twists of magnitude equal to that of the lowest common multiple of the inverse moduli, $\mu^{-1}$, of the horizontal and vertical decompositions.
+
+\begin{figure}[H]
+\centering
+\includesvg[width=2.6in]{cylinderskew}
+\caption{Stabilizing Dehn-twists of the horizontal/vertical cylinders}
+\end{figure}
+
+In this case we are lucky and the moduli are easily obtained and uniform throughout $\tilde{\mathbf{M}}$.
+We can see that any parabolic skews of magnitude 2 preserve the boundaries of the cylinders.
+Consequently, our surface is stabilized by the following matrices:
+
+\begin{equation}
+\left[ \hspace{1mm} \begin{matrix}
+				1 & \pm 2\\
+				0 & 1
+			\end{matrix}\hspace{1mm}\right],
+			\left[ \hspace{1mm} \begin{matrix}
+							1 & 0\\
+							\pm 2 & 1
+						\end{matrix}\hspace{1mm}\right]
+\end{equation}
+
+\newpage
+\subsection{Other symmetries of $\tilde{\mathbf{M}}$}
+
+\subsection{Horizontal, Vertical, and Slope One Cylinders of $\tilde{\mathbf{M}}$}
+\begin{figure}[H]
+\centering
+\includesvg[pretex=\tiny, width=4in]{slopeonecylinder}
+\end{figure}
+
+\subsection{Computing the Veech Group of $\tilde{\mathbf{M}}$}
+\subsection{Slope Normalization Algorithm}
+\subsection{Proof of Theorem}
+
 \end{document}

### 2017-07-05 paper.tex
070517

@@
 \subsection{The Staircase Model}
-At first glance, $\tilde{\mathbf{M}}$ does not seem to be a square-tiled surface.
-However, observe that the surface can be reconstructed by a sequence of cutting and gluing operations and realized in the following way:
+Observe that $\tilde{\mathbf{M}}$ can be reconstructed by a series of cutting and gluing operations and realized as a square tiled staircase:
 
@@
 \includesvg[width=4in]{mtildastaircase}
-\caption{$\tilde{\mathbf{M}}$ as a staircase with vertices and phases identified.}
+\caption{$\tilde{\mathbf{M}}$ as a staircase with vertices $x_i$ and phases identified.}
 \end{figure}
 
-Every edge on the staircase model is glued by translation to another by at most a distance of 2. (Each square is unit). More importantly, the staircase has a very natural cylindrical decomposition.
+Every edge pair on the staircase model has an exact distance of two. (Each square is unit). This gives way to a very natural cylindrical decomposition.
 
@@
 The staircase is then decomposed into 6 uniform cylinders of moduli $\frac{1}{2}$ in both the vertical and horizontal directions.
-The modulus of the cylinder, $\mu$, is given by the ratio of the cylinder's circumference and width, $\mu{}={}w/c$.
-A surface that is composed entirely out of similar horizontal and vertical cylinders is stabilized by Dehn-twists of magnitude equal to that of the lowest common multiple of the inverse moduli, $\mu^{-1}$, of the horizontal and vertical decompositions.
+All families of saddle connections on the surface intersect exactly two others.
+
+\begin{Def}
+The \textbf{family of horizontal cylinders} of $\tilde{\mathbf{M}}$, denoted $\Delta_{i}$ where $1\leq i \leq 6$, are cylinders formed by saddle connections made between singularities $((x_3,x_2)\cup(x_2,x_3)$, $(x_3,x_4))$ when i is odd, and $((x_1,x_4)\cup(x_2,x_4)$, $(x_1,x_2))$ when i is even.
+\newline
+Likewise, the \textbf{family of vertical cylinders}, denoted $\varepsilon_j$ $(1\leq j 
+\leq 6)$, are formed by connections $((x_2,x_3)$, $(x_2,x_1)\cup(x_1,x_2))$ when j is odd, and  $((x_4,x_1)$, $(x_4,x_3)\cup(x_3,x_4))$ when j is even.
+\end{Def}
+
+Horizontal and vertical cylinders of the surface are disjoint within their respective sets, and the union of both sets is the surface $\tilde{\mathbf{M}}$.
+
+  Recall that the modulus of a cylinder, $\mu$, is given by the ratio of the cylinder's width to its circumference, $\mu{}={}w/c$.
+A surface that is composed entirely out of similar horizontal and vertical cylinders is stabilized by parabolic matrices with magnitude $\lambda$ equal to that of lowest common multiple of the inverse moduli, $\mu^{-1}$, of both horizontal and vertical decompositions.
+[cite Veech 89?]
 
@@
 \caption{Stabilizing Dehn-twists of the horizontal/vertical cylinders}
+\label{fig:skew}
 \end{figure}
 
-In this case we are lucky and the moduli are easily obtained and uniform throughout $\tilde{\mathbf{M}}$.
-We can see that any parabolic skews of magnitude 2 preserve the boundaries of the cylinders.
-Consequently, our surface is stabilized by the following matrices:
+In this case, the moduli are easily obtained and uniform throughout $\tilde{\mathbf{M}}$.
+We can see that any parabolic skews of magnitude $\lambda=\pm2$, since $\mu^{-1}{}=2$, preserve the boundaries of the cylinders.
+Figure $\ref{fig:skew}$ showcases how Dehn-twists of this magnitude followed by an Affine transformation matches these boundaries again.
+Consequently, our surface is stabilized by the matrices
 
@@
 				0 & 1
-			\end{matrix}\hspace{1mm}\right],
+			\end{matrix}\hspace{1mm}\right],\text{ and }
 			\left[ \hspace{1mm} \begin{matrix}
@@
 							\pm 2 & 1
-						\end{matrix}\hspace{1mm}\right]
+						\end{matrix}\hspace{1mm}\right].
 \end{equation}
@@
 \subsection{Other symmetries of $\tilde{\mathbf{M}}$}
+A major benefit of $\tilde{\mathbf{M}}$ being tiled by unit squares is that the surface has all of the same symmetries of the square.
+The action of $\mathbb{D}_4$ on $\tilde{\mathbf{M}}$ preserves its boundaries up to translation.
+
+\subsection{Computing the Veech Group of $\tilde{\mathbf{M}}$}
 
@@
 
-\subsection{Computing the Veech Group of $\tilde{\mathbf{M}}$}
 \subsection{Slope Normalization Algorithm}

### 2017-07-07 paper.tex
070717

@@
 \subsection{Four-Fold Cover of $\mathbf{U}$}
-A geodesic on our surface can travel in one of four possible directions at any given time.
-To each trajectory phase of we associate a numeric label I,II,II, or IV, where I contains the initial and terminal points of a closed geodesic.
-(Similar to [author]'s construction of phases of the Wind-Tree Model).
-For every even integer pair centered around every hole of $\mathbf{P}$, assign a pair of integers (m,n).
-Identify the edges in the following manner:
+A geodesic as viewed as line segments on $\mathbf{P}$ can travel in one of four possible directions at any given time.
+On $\mathbf{U}$ a geodesic flow's direction is characterized as an element of the set of all possible directions as equivalence classes of $\mathbb{R}/2\pi\mathbb{Z}$ of the form $\mathbf{V}_{\theta}=\{\theta,\theta+\frac{\pi}{2},\theta+\pi, \theta+\frac{3\pi}{2} : \theta\in[0,\frac{\pi}{2})\}$.
+To each member of the class we associate an element of $\mathbb{Z}/4\mathbb{Z}$ to signify the current phase of the geodesic flow.
+We use this construction to create our four-fold cover of $\mathbf{U}$ as a topological quotient of $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$.
+For every even integer pair centered around every hole of $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$, assign a pair of integers (m,n).
+Edge identifications are made on the boundaries of every hole taking one phase plane to the next in the following manner:
 
@@
 \end{center}
-This is a branched cover of $\mathbf{U}$ of degree four that can be thought of as four copies of $\mathbf{P}$, all stacked on top of one another with the following identifications made:
+This is a branched cover of $\mathbf{U}$ of degree four,$\tilde{\mathbf{U}}_0$, with the obvious covering map taking every non-singular point and projecting it onto $\mathbf{U}$.
+We define the edge identifications as a relation on $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$.
+For $i\in[-\frac{1}{4},\frac{1}{4}]$:
 
 \begin{multline}
-([2m-\frac{1}{4}, 2m+\frac{1}{4}], 2n-\frac{1}{4},j)\sim(2m-\frac{1}{4},[2n-\frac{1}{4}, 2n+\frac{1}{4}],j+\bar{1})\\
-([2m-\frac{1}{4}, 2m+\frac{1}{4}], 2n+\frac{1}{4},j)\sim(2m+\frac{1}{4},[2n-\frac{1}{4}, 2n+\frac{1}{4}],j+\bar{1})\\
-\vspace{0.2in}For (m,n,j) \in \mathbb{Z}^{2}\times\mathbb{Z}/4\mathbb{Z}
+([2m+t], 2n-\frac{1}{4},j)\sim(2m-\frac{1}{4},[2n+t],j+1)\\
+([2m+t], 2n+\frac{1}{4},j)\sim(2m+\frac{1}{4},[2n+t],j+1)\\
+\vspace{0.2in}For (m,n;j) \in \mathbb{Z}^{2}\times\mathbb{Z}/4\mathbb{Z}
 \label{eq:rel2}
@@
 
-There are four distinct vertices for each hole centered at $(m,n,\mathbb{Z}/4\mathbb{Z})$ on  $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$ of cone angle $6\pi$ once the identifications have been made.
-Aligning the flow on  $\mathbf{U}$ requires a rotation on phase planes II,III, and IV by $-\frac{\pi}{2}, -\pi,$ and $ -\frac{3\pi}{2}$ in that order, about the origin of each respective plane.
-The choice in rotations are arbitrary as long as the edges are glued by translation.
+We define $\tilde{\mathbf{U}}_0$ to be the topological quotient, $(\mathbf{P}\times\mathbb{Z}/4\mathbb{Z})/\sim$.
+$\tilde{\mathbf{U}}_0$ is a Euclidean cone surface with four distinct vertices for each hole centered at $(m,n;\mathbb{Z}/4\mathbb{Z})$ on  $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$ of cone angle $6\pi$ with proper identification.
+Now we will show that $\tilde{\mathbf{U}}_0$ is isometric to a translation surface via an induced isometry on $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$.
 
-We express  $p:(\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}) \rightarrow (\mathbf{P}\times\mathbb{Z}/4\mathbb{Z})$ as
+We express $\phi:\mathbf{P}\times\mathbb{Z}/4\mathbb{Z} \rightarrow \mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$ as $\phi(u,v;j) = (u',v';j)$, where
 
 \begin{equation}
-	p\left[\begin{array}{c}
-		u\\v
+	\left[\begin{array}{c}
+		u'\\v'
 	\end{array}\right] =
@@
 				-1 & 0
-			\end{matrix}\hspace{1mm}\right]^{z}\left[\begin{array}{c}
+			\end{matrix}\hspace{1mm}\right]^{j}\left[\begin{array}{c}
 					u\\v
-				\end{array}\right]\times\{\bar z\}
-\label{eq:p}
+				\end{array}\right]
+\label{eq:phi}
 \end{equation}
 
-\begin{Def} Let $\sim$ be the equivalence relation given at $\eqref{eq:rel2}$, and p be the automorphism from $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$ to itself at $\eqref{eq:p}$.
-Define the relation obtained from $\sim$ on the image of p by $\sim_{p} = \{(x_1,p(x_2)) : x_1 \sim x_2\}$.
-Then the $\mathbf{translation\text{ }surface}$  $\tilde{\mathbf{U}}$ = Im p$\slash\sim_{p}$.
+\begin{Def} Let $\sim$ be the equivalence relation $\eqref{eq:rel2}$, and $\phi$ be the isomorphism on $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$ $\eqref{eq:phi}$.
+Define a new relation obtained from $\sim$ as:  $x_1\sim_{\phi}x_2$ if $\phi^{-1}(x_1) = \phi^{-1}(x_2)$ for $x_1,x_2\in\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$.
+We define $\tilde{\mathbf{U}}$ to be  $(\mathbf{P}\times\mathbb{Z}/4\mathbb{Z})/\sim_{\phi}$.
 \end{Def}
 
-\begin{thm}{$\tilde{\mathbf{U}}$ is a branched cover of $\mathbf{U}$. }\newline
-
+\begin{thm}{The following statements are true:}
+\begin{enumerate}[label=(\roman*)]
+\item $\tilde{\mathbf{U}}$ is a translation surface.
+\item The map $\phi$ induces a well-defined isometry $\bar{\phi}:\tilde{\mathbf{U}}_0 \rightarrow \tilde{\mathbf{U}}$
+\end{enumerate}
+\begin{proof}
+the proof is left as an exercise to the reader.
+\end{proof}
 \end{thm}
 
-p preserves the edge identification by affine transformations.  $\tilde{\mathbf{U}}$ as a translation cover:
+$\phi$ preserves the edge identifications on $\tilde{\mathbf{U}}_0$ by affine transformations.
+And since $\tilde{\mathbf{U}}_0$ is a cover of $\mathbf{U}$, $\tilde{\mathbf{U}}$ is a translation cover of $\mathbf{U}$.
+Observe that by rotating each phase plane, the manner in which the holes are indexed undergoes the same transformations from our perspective.
 
@@
 \end{figure}
-As one might expect, traveling in the same direction in every phase has a different effect on the integer pair (m,n) relative to first phase.
-A path in that sense is considered closed if it eventually returns to the exact same point, and periodic should it return to the same phase as well.
+
+Alternatively, the edge identifications made on $\mathbf{U}$ that gave us the set $\mathbf{V}_\theta$ let us define a flow $F^{t}:T_1(\mathbf{U})\rightarrow T_1(\mathbf{U})$ as $F^{t}(p,v)=(p+tv,v)$, where $(p,v)$ is a member of the unit tangent bundle on point $p$ of $\mathbf{U}$ in direction $v$ of $\mathbf{V}_\theta$.
+Defining a map from $d:T_1(\mathbf{U})\rightarrow \mathbf{V}_\theta$ that returns the direction $v$.
+$\tilde{\mathbf{U}}$ can be abstractly realized as the preimage of $d$, $d^{-1}(\mathbf{V}_\theta)$.
+
+A geodesic on this translation surface can be realized as a sequence of edge crossings on the phase planes.
+By taking a quotient of this surface, a closed straight-line path can be lifted from the quotient onto $\tilde{\mathbf{U}}$, and then projected onto $\mathbf{U}$ to study geodesic behavior of any given direction.
 
@@
 \compat{You need to build the surface out of pieces rather than out of ``copies of $\mathbf{M}$.''} 
-With $\tilde{\mathbf{U}}$ defined, its topological quotient is given as a union of phase planes modded out by translations.
+With $\tilde{\mathbf{U}}$ defined, we can describe its quotient as $\tilde{\mathbf{U}}$ modded out by translations.
+We use the induced isometry, $\bar\phi$, to respect our original notion of direction on $\mathbf{U}$ as a quotient of $\mathbf{P}$.
 
 \begin{Def}
-$\tilde{\mathbf{M}}{}={}\left[\bigcup_{i=0}^{3} p(\mathbf{P},i)\slash(2\mathbb{Z})^{2} \right]/\sim_{p\vert_{X}}{}X{}={}(\left[-1,1\right]\times\left[-1,1\right])\times\mathbb{Z}\slash4\mathbb{Z}$
+The translation surface $\tilde{\mathbf{M}}$ is given as the quotient $\tilde{\mathbf{U}}/T$, where $T:(2\mathbb{Z})^2\times\tilde{\mathbf{U}}\rightarrow\tilde{\mathbf{U}}$ is the group action defined as $T_{m,n}(x)=\bar{\phi}\left((m,n)+\bar{\phi}^{-1}(x)\right)$ for $x\in\tilde{\mathbf{U}}$.
 \end{Def}

### 2017-07-11 paper.tex
71117

@@
-\documentclass[]{article}
+j\documentclass[]{article}
 
@@
 
-\subsection{The Euclidean cone surface $\mathbf{M}$}
-\compat{I would say: Observe that translations of ${\mathbb R}^2$ by pairs even integers preserves the subset ${\mathbf P}$ and respects the equivalence relation defining ${\mathbf U}$.
-Thus this gives an action of $(2{\mathbb Z})^2$ on ${\mathbf U}'$.
-We define ${\mathbf M}$ to be the quotient.
-It is depicted in Figure ...}
-
-Observe that the translations by pairs of even integers preserves the subset $\mathbf P$ and respects the equivalence relation on $\mathbf U$
-
-We will call this quotient space $\mathbf{M}$ and define it as such:
-\begin{Def}{$\mathbf{M}$.}\\
-$\mathbf{M}$ is a base surface of $\mathbf{U}'$ given by the quotient $\mathbf{U}'/2\mathbb{Z}\oplus2\mathbb{Z}$, where $2\mathbb{Z}\oplus2\mathbb{Z}\times\mathbf{U}'\rightarrow\mathbf{U}'$ is the group action on $\mathbf{U}'$ defined as $(m,n)\circ(u,v)=(u+m,v+n)$.
-\end{Def}
+\subsection{Four-Fold Cover of $\mathbf{U}$}
+A geodesic on $\mathbf{U}$ viewed as a sequence of line segments on $\mathbf{P}$ travels in at most four directions which differ by rotation by a multiple of $\frac{\pi}{2}$.
+This is a consequence of the edge identifications taking the boundary of a hole on $\mathbf{P}$ to its counterpart by rotation of that factor.
+To each of these directions we associate an element of $\mathbb{Z}/4\mathbb{Z}$ to signify the current phase of the geodesic flow.
+We use this construction to create our four-fold cover of $\mathbf{U}$ as a topological quotient of $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$.
+For every integer pair centered around every hole of $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$, create the edge identifications taking one phase plane to the next in the following manner:
 
 \begin{figure}[H]
-\begin{center}
-\includesvg[width=2in.]{unfoldcutquotient}\hspace{0.2in}
-\includesvg[width=2in.]{quotient}
-\end{center}
-\caption{$\mathbf{U}'$ equivalence classes under group action (left), and $\mathbf{M}$ with identified edges and vertices (right).}
-\label{fig:quotient}
+\centering
+\includesvg[width=4in]{utilda}
+\label{fig:utilda0}
+\caption{$\tilde{\mathbf{U}}_0$ is a branched cover of $\mathbf{U}$ of degree four.}
 \end{figure}
 
-The quotient space is a genus 1 surface that resembles a $2\times2$ torus with a unit square removed from it.
-Near the outer vertex it behaves as a torus might.
-The inner vertices $c_{1},c_{3}$ have cone angle $\frac{3\pi}{2}$, and $c_{2}$ has a cone angle of $3\pi$.
-None of these inner vertices have cone angles that are integral multiples of $2\pi$, so in order to create the necessary translation surfaces we require the minimal number of copies of both $\mathbf{U}$ and $\mathbf{M}$ to construct their associated phase spaces.
-\compat{Move the discussion of building the translation surface to the next section.
-Otherwise this paragraph is good.}
-
-\subsection{Four-Fold Cover of $\mathbf{U}$}
-A geodesic as viewed as line segments on $\mathbf{P}$ can travel in one of four possible directions at any given time.
-On $\mathbf{U}$ a geodesic flow's direction is characterized as an element of the set of all possible directions as equivalence classes of $\mathbb{R}/2\pi\mathbb{Z}$ of the form $\mathbf{V}_{\theta}=\{\theta,\theta+\frac{\pi}{2},\theta+\pi, \theta+\frac{3\pi}{2} : \theta\in[0,\frac{\pi}{2})\}$.
-To each member of the class we associate an element of $\mathbb{Z}/4\mathbb{Z}$ to signify the current phase of the geodesic flow.
-We use this construction to create our four-fold cover of $\mathbf{U}$ as a topological quotient of $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$.
-For every even integer pair centered around every hole of $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$, assign a pair of integers (m,n).
-Edge identifications are made on the boundaries of every hole taking one phase plane to the next in the following manner:
-
-\begin{center}
-\includesvg[width=4in]{utilda}
-\end{center}
-This is a branched cover of $\mathbf{U}$ of degree four,$\tilde{\mathbf{U}}_0$, with the obvious covering map taking every non-singular point and projecting it onto $\mathbf{U}$.
+$\tilde{\mathbf{U}}_0$ has the obvious covering map that takes every point on $\tilde{\mathbf{U}}_0$ and projects it onto $\mathbf{U}$.
 We define the edge identifications as a relation on $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$.
@@
 \begin{multline}
-([2m+t], 2n-\frac{1}{4},j)\sim(2m-\frac{1}{4},[2n+t],j+1)\\
-([2m+t], 2n+\frac{1}{4},j)\sim(2m+\frac{1}{4},[2n+t],j+1)\\
-\vspace{0.2in}For (m,n;j) \in \mathbb{Z}^{2}\times\mathbb{Z}/4\mathbb{Z}
+([2m+i], 2n-\frac{1}{4},j)\sim(2m-\frac{1}{4},[2n+i],j+1)\\
+([2m+i], 2n+\frac{1}{4},j)\sim(2m+\frac{1}{4},[2n+i],j+1)\\
+\vspace{0.2in}\text{for } (m,n;j) \in \mathbb{Z}^{2}\times\mathbb{Z}/4\mathbb{Z}.
 \label{eq:rel2}
@@
 $\tilde{\mathbf{U}}_0$ is a Euclidean cone surface with four distinct vertices for each hole centered at $(m,n;\mathbb{Z}/4\mathbb{Z})$ on $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$ of cone angle $6\pi$ with proper identification.
-Now we will show that $\tilde{\mathbf{U}}_0$ is isometric to a translation surface via an induced isometry on $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$.
+
+\subsection{Symmetries of $\tilde{\mathbf{U}}_0$}
+There are certain translational and rotational symmetries of $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$ that we use to define isometric maps.
+These isometries then induce well-defined isometries on $\tilde{\mathbf{U}}_0$ that respect the edge identifications we made earlier.
+
+\begin{Def}Let $m,n\in\mathbb{Z}$.
+The \textbf{family of translations}, $\mathbf{T}^{m,n}_{0}$ on $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$, is given as $\mathbf{T}^{m,n}_{0}(u,v;j)=(u+2m,v+2n;j)$.\newline
+The \textbf{family of "rotational" isometries} on $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$ is given as powers of the automorphism $\mathbf{R}(u,v;j)=(u,v;j+1)$, where $\mathbf{R}$ as an action on $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$ forms a group of order 4.
+\end{Def}
+
+\begin{thm}$\mathbf{T}^{m,n}_{0}$ induces an isometry on $\tilde{\mathbf{U}}_0$.
+\begin{proof}
+This theorem follows from the fact that $\mathbf{T}^{m,n}_{0}$ preserves the edge identifications defined at $\eqref{eq:rel2}$.
+Take $\mathbf{T}^{m,n}_{0}(u,v;j)=(u+2m,v+2n;j)$ for arbitrary integers $m,n$, and $(u,v;j)\in\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$.
+Then, $\mathbf{T}^{m,n}_{0}(u,v;j)$ is trivially isometric for elements not on the boundary of $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$, since these points lie on a subset of a plane and are identified with themselves.
+Suppose then that $(u,v;j)\in\partial\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$.
+Then $\mathbf{T}^{m,n}_{0}(u,v;j)$ acting on the boundary gives a new relation:
+\begin{multline*}
+([2(m+m')+i], 2(n+n')-\frac{1}{4},j')\sim(2(m+m')-\frac{1}{4},[2(n+n')+i],j'+1)\\
+([2(m+m')+i], 2(n+n')+\frac{1}{4},j')\sim(2(m+m')+\frac{1}{4},[2(n+n')+i],j'+1)\\
+\vspace{0.2in}\text{for } (m',n';j') \in \mathbb{Z}^{2}\times\mathbb{Z}/4\mathbb{Z}\text{ and }i\in[-\frac{1}{4},\frac{1}{4}],
+\end{multline*}
+which preserves the lengths of the boundary edges and respects the edge identifications as every hole is centered around a unique pair of even integers, $(2(m+m'),2(n+n')$.
+We denote the induced isometry on $\tilde{\mathbf{U}}_0$ as $\bar{\mathbf{T}}^{m,n}_{0}$.
+\end{proof}
+\end{thm}
+
+\begin{thm}$\mathbf{R}_{0}$ induces an isometry on $\tilde{\mathbf{U}}_0$.
+\begin{proof}
+Similar to the previous theorem, consider a point $(u,v;j)\in\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$.
+Then $\mathbf{R}_{0}(u,v;j)=(u,v;j+1)$ respects edge identifications at $\eqref{eq:rel2}$.
+since $\mathbf{R}_{0}$ is itself a bijective permutation on $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$ that takes every pair $(u,v)$ to itself and permutes the $j$ coordinate by a factor of one.
+Therefore, $\mathbf{R}_{0}$ induces an isometry on $\tilde{\mathbf{U}}_0$.
+We denote this induced isometry by $\bar{\mathbf{R}}_{0}$.
+\end{proof}
+\end{thm}
+
+\subsection{Four-fold Translation Cover}
+
+ Now we will show that $\tilde{\mathbf{U}}_0$ is isometric to a translation surface via another induced isometry on $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$.
+We would like this isometry to rotate the individual phase planes of $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$ about the origin in such a way that edges one the boundaries are identified strictly by translation.
+Every phase characterized as an element of $\mathbb{Z}/4\mathbb{Z}$ becomes a power of a rotation matrix on the individual phase planes.
 
@@
 
-\begin{Def} Let $\sim$ be the equivalence relation $\eqref{eq:rel2}$, and $\phi$ be the isomorphism on $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$ $\eqref{eq:phi}$.
+\begin{Def} Let $\sim$ be the equivalence relation $\eqref{eq:rel2}$, and $\phi$ be the isometry on $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$ $\eqref{eq:phi}$.
 Define a new relation obtained from $\sim$ as:  $x_1\sim_{\phi}x_2$ if $\phi^{-1}(x_1) = \phi^{-1}(x_2)$ for $x_1,x_2\in\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$.
-We define $\tilde{\mathbf{U}}$ to be  $(\mathbf{P}\times\mathbb{Z}/4\mathbb{Z})/\sim_{\phi}$.
+We define $\tilde{\mathbf{U}}$ to be  $(\mathbf{P}\times\mathbb{Z}/4\mathbb{Z})/\sim_{\phi}$.\label{eq:simphi}
 \end{Def}
+
+The map $\phi$ preserves the edge identifications on $\tilde{\mathbf{U}}_0$ by these affine transformations.
+And since $\tilde{\mathbf{U}}_0$ is a cover of $\mathbf{U}$, $\tilde{\mathbf{U}}$ is a translation cover of $\mathbf{U}$.
+Observe that by rotating each phase plane about the origin, the manner in which the holes are indexed undergoes a rotation according to the j-coordinate.
+
+\begin{figure}[H]
+\centering
+\includesvg[width=3.3in]{utildaprime}
+\end{figure}
 
@@
 \begin{enumerate}[label=(\roman*)]
+\item The map $\phi$ induces a well-defined isometry $\bar{\phi}:\tilde{\mathbf{U}}_0 \rightarrow \tilde{\mathbf{U}}$
 \item $\tilde{\mathbf{U}}$ is a translation surface.
-\item The map $\phi$ induces a well-defined isometry $\bar{\phi}:\tilde{\mathbf{U}}_0 \rightarrow \tilde{\mathbf{U}}$
 \end{enumerate}
 \begin{proof}
-the proof is left as an exercise to the reader.
+The proof of this theorem also follows from the fact that $\phi \eqref{eq:phi}$ is an isometric map that respects edge identifications.
+Let $(u,v;j)\in\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$.
+Then $\phi(u,v;j)=(u',v';j)$.
+Let $M=\left[\begin{matrix}
+0 & 1\\
+-1 & 0
+\end{matrix}\right]$, $U_1=([2m+i],2n-\frac{1}{4})$, $U_2=([2m+i],2n+\frac{1}{4})$, $V_1=(2m-\frac{1}{4},[2n+i])$, and $V_2=(2m+\frac{1}{4},[2n+i])$ for $i\in[-\frac{1}{4},\frac{1}{4}]$ and $m,n\in\mathbb{Z}$.
+Observe that making the identification $M^{j}(X)\sim' M^{j+1}(X')$ for some set of vectors $X,X'$ is equivalent to $X\sim' M(X')$ since $M$ is an invertible, isometric linear map.
+Similarly, we can express the relations $\phi(U_1;j)\sim\phi(V_1;j+1)$ and $\phi(U_2;j)\sim\phi(V_2;j+1)$ as $(M^{j}(U_1);j)\sim (M^{j+1}(V_1);j+1)$ and $(M^{j}(U_2);j)\sim (M^{j+1}(V_2);j+1)$, or, more concisely, as
+
+\begin{multline*}
+([2m+i], 2n-\frac{1}{4},0)\sim([2n+i],-2m+\frac{1}{4},1)\\
+([2m+i], 2n+\frac{1}{4},0)\sim([2n+i],-2m-\frac{1}{4},1)\\
+\text{for } (m,n) \in \mathbb{Z}^{2}.
+\end{multline*}
+This is precisely how the relation $\sim_{\phi}$ in Definition $\eqref{eq:simphi}$ is described (taking $\phi^{-1}$ of two points and checking if $\sim$ relates them).
+Since $\phi$ respects edge identifications and preserves boundary lengths, there exists an induced map, $\bar{\phi}:\tilde{\mathbf{U}}_0 \rightarrow \tilde{\mathbf{U}}$ and \emph{(i)} is true.
+Finally, the relation $\sim_{\phi}$ as expressed this way obviously identifies edges from one phase plane to the next by translation, so \emph{(ii)} is true.
 \end{proof}
@@
 
-$\phi$ preserves the edge identifications on $\tilde{\mathbf{U}}_0$ by affine transformations.
-And since $\tilde{\mathbf{U}}_0$ is a cover of $\mathbf{U}$, $\tilde{\mathbf{U}}$ is a translation cover of $\mathbf{U}$.
-Observe that by rotating each phase plane, the manner in which the holes are indexed undergoes the same transformations from our perspective.
-
-\begin{figure}[H]
-\centering
-\includesvg[width=4in]{utildaprime}
-\end{figure}
-
-Alternatively, the edge identifications made on $\mathbf{U}$ that gave us the set $\mathbf{V}_\theta$ let us define a flow $F^{t}:T_1(\mathbf{U})\rightarrow T_1(\mathbf{U})$ as $F^{t}(p,v)=(p+tv,v)$, where $(p,v)$ is a member of the unit tangent bundle on point $p$ of $\mathbf{U}$ in direction $v$ of $\mathbf{V}_\theta$.
-Defining a map from $d:T_1(\mathbf{U})\rightarrow \mathbf{V}_\theta$ that returns the direction $v$.
-$\tilde{\mathbf{U}}$ can be abstractly realized as the preimage of $d$, $d^{-1}(\mathbf{V}_\theta)$.
-
-A geodesic on this translation surface can be realized as a sequence of edge crossings on the phase planes.
+%A geodesic on this translation surface can be realized as a sequence of edge crossings on the phase planes.
 By taking a quotient of this surface, a closed straight-line path can be lifted from the quotient onto $\tilde{\mathbf{U}}$, and then projected onto $\mathbf{U}$ to study geodesic behavior of any given direction.
 
-\newpage
-\subsection{Translation Surface}
-\compat{You need to build the surface out of pieces rather than out of ``copies of $\mathbf{M}$.''} 
-With $\tilde{\mathbf{U}}$ defined, we can describe its quotient as $\tilde{\mathbf{U}}$ modded out by translations.
-We use the induced isometry, $\bar\phi$, to respect our original notion of direction on $\mathbf{U}$ as a quotient of $\mathbf{P}$.
+\subsection{Symmetries of $\tilde{\mathbf{U}}$}
+In this section we use the induced isometries on $\tilde{\mathbf{U}}_0$, and the map $\bar{\phi}:\tilde{\mathbf{U}}_0\rightarrow\tilde{\mathbf{U}}$ to define a set of rotational and translational symmetries of $\tilde{\mathbf{U}}$ such that its quotient over the action of these maps are faithful representations of $\tilde{\mathbf{U}}_0$ in terms of how the holes in $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$ are indexed.
+It is briefly mention after the definition of $\tilde{\mathbf{U}}$ that $\bar{\phi}$ alters this scheme relative to $\tilde{\mathbf{U}}_0$.
+
+\begin{Def}We call the \textbf{family of faithful Translational symmetries} $\mathbf{T}^{m,n}$ on $\tilde{\mathbf{U}}$ and define it to be the composition of induced isometries, $\mathbf{T}^{m,n}$ = $\bar{\phi}\circ\bar{\mathbf{T}}^{m,n}\circ\bar{\phi^{-1}}$.
+\end{Def}
+
+We can see that $\mathbf{T}^{m,n}$ is indeed an isometry on $\tilde{\mathbf{U}}$ as it is composed of isometries.
+We take the quotient over the group generated from these translations, and show that this new surface is Veech.
 
 \begin{Def}
-The translation surface $\tilde{\mathbf{M}}$ is given as the quotient $\tilde{\mathbf{U}}/T$, where $T:(2\mathbb{Z})^2\times\tilde{\mathbf{U}}\rightarrow\tilde{\mathbf{U}}$ is the group action defined as $T_{m,n}(x)=\bar{\phi}\left((m,n)+\bar{\phi}^{-1}(x)\right)$ for $x\in\tilde{\mathbf{U}}$.
+The translation surface $\tilde{\mathbf{M}}$ is given as the quotient  $\tilde{\mathbf{U}}/\mathbf{T}^{m,n}$.
 \end{Def}
-
-$\tilde{\mathbf{M}}$ is also a degree four branched cover of $\mathbf{M}$.
-A covering map $\tilde{\mathbf{M}} \rightarrow \mathbf{M}$ individually rotates these squares about the centers of their holes according to their labels and maps their points to $\mathbf{M}$.
-This is a genus five translation surface with four singularities.
-$\tilde{\mathbf{M}}$ as a quotient of $\tilde{\mathbf{U}}$ can be represented as quotients of each plane:
 
@@
 \includesvg[width=4in]{mtilda}
+\caption{The surface $\tilde{\mathbf{M}}$ with labeled vertices and identified edges}
+\label{fig:mtilda}
 \end{figure}
 
-All in all, we have the following diagram:
+$\tilde{\mathbf{M}}$ is a genus 5 translation surface that is covered by $\tilde{\mathbf{U}}$.
+Thus we can take a path on $\tilde{\mathbf{M}}$, lift it onto $\tilde{\mathbf{U}}$, take it to $\tilde{\mathbf{U}}_0$, and project it onto $\mathbf U$.
+And since $\mathbf{U}\simeq\mathbf{S}$, the core of our argument is that geodesics on the translation surface $\tilde{\mathbf{M}}$ equipped with homology classes are representations of geodesics on $\mathbf{S}$ away from the singularities.
+The labeled vertices in Figure $\ref{fig:mtilda}$ are all cone singularities with total angles of $6\pi$.
 
-\begin{figure}[H]
-\centering
-\includegraphics[width=2in]{commute.pdf}
-\end{figure}
-
-\newpage
-
-\section{$\tilde{\mathbf{M}}$ as a Veech Surface}
-The following section studies the translation surface's various symmetries and associated Veech group.
-We use the Veech dichotomy to prove our main theorems that determine whether the geodesic is ultimately periodic or drift periodic.
-
-\subsection{The Staircase Model}
-Observe that $\tilde{\mathbf{M}}$ can be reconstructed by a series of cutting and gluing operations and realized as a square tiled staircase:
+Although it is not immediately obvious, $\tilde{\mathbf{M}}$ can be realized as a square-tiled surface in the form of a staircase as in $\ref{fig:staircase}$
 
@@
 \includesvg[width=4in]{mtildastaircase}
-\caption{$\tilde{\mathbf{M}}$ as a staircase with vertices $x_i$ and phases identified.}
+\caption{$\tilde{\mathbf{M}}$ as a square-tiled staircase with vertices and phases identified.}
+\label{fig:staircase}
 \end{figure}
 
-Every edge pair on the staircase model has an exact distance of two. (Each square is unit). This gives way to a very natural cylindrical decomposition.
+\subsection{$\tilde{\mathbf{M}}$ as a Veech Surface}
+We use basic properties of $\tilde{\mathbf{M}}$ to show that it is Veech, and its Veech group is a lattice in $SL_2(\mathbb{Z})$.
+Recall that translation surfaces are equipped with charts, $\omega$, of maps from the surface to $\mathbb{C}$.
+The Veech group of the surface $\tilde{\mathbf{M}}$ is given as the image of its stabilizer in the projective group $PSL_2(\mathbb{R})$.
+
+\begin{Def}
+Since $\tilde{\mathbf{M}}$ is a translation surface with four singularities, it is a Veech surface that locally behaves like a torus away from these cone singularities.
+$\tilde{\mathbf{M}}$ has a \textbf{Veech group}, denoted $\Gamma$.
+\end{Def}
+
+We use the following theorem of Gutkin and Judge to show that $\Gamma$ is lattice (and arithemtic) in $SL_2(\mathbb{Z})$:
+
+\begin{thm}(Gutkin-Judge) A translation surface $(S,\omega)$ is squaretiled
+if and only if its Veech group $V(S,\omega)$ shares a finite-index subgroup
+with $SL_2(\mathbb{Z}).$
+\end{thm}
+
+\begin{thm}The following are equivalent:
+\begin{enumerate}[label=(\roman*)]
+\item $\tilde{\mathbf{M}}$ is tiled by squares.
+\item $\Gamma$ is lattice in $SL_2(\mathbb{Z})$.
+\end{enumerate}
+\begin{proof}
+Figure $\ref{fig:staircase}$ explicitly demonstrates how $\tilde{\mathbf{M}}$ is tiled by squares. The proof then follows as an obvious result of the previous theorem.
+\end{proof}
+\end{thm}
+
+\newpage
+%\subsection{Translation Surface}
+%\compat{You need to build the surface out of pieces rather than out of ``copies of $\mathbf{M}$.''} 
+%With $\tilde{\mathbf{U}}$ defined, we can describe its quotient as $\tilde{\mathbf{U}}$ modded out by translations.
+We use the induced isometry, $\bar\phi$, to respect our original notion of direction on $\mathbf{U}$ as a quotient of $\mathbf{P}$.
+%
+%
+%
+%\begin{Def}
+%The translation surface $\tilde{\mathbf{M}}$ is given as the quotient $\tilde{\mathbf{U}}/T$, where $T:(2\mathbb{Z})^2\times\tilde{\mathbf{U}}\rightarrow\tilde{\mathbf{U}}$ is the group action defined as $T_{m,n}(x)=\bar{\phi}\left((m,n)+\bar{\phi}^{-1}(x)\right)$ for $x\in\tilde{\mathbf{U}}$.
+%\end{Def}
+%
+%$\tilde{\mathbf{M}}$ is also a degree four branched cover of $\mathbf{M}$.
+A covering map $\tilde{\mathbf{M}} \rightarrow \mathbf{M}$ individually rotates these squares about the centers of their holes according to their labels and maps their points to $\mathbf{M}$.
+This is a genus five translation surface with four singularities.
+$\tilde{\mathbf{M}}$ as a quotient of $\tilde{\mathbf{U}}$ can be represented as quotients of each plane:
+%
+%\begin{figure}[H]
+%\centering
+%\includesvg[width=4in]{mtilda}
+%\end{figure}
+%
+%All in all, we have the following diagram:
+
+%\begin{figure}[H]
+%\centering
+%\includegraphics[width=2in]{commute.pdf}
+%\end{figure}
+%
+%\newpage
+%\subsection{The Euclidean cone surface $\mathbf{M}$}
+%\compat{I would say: Observe that translations of ${\mathbb R}^2$ by pairs even integers preserves the subset ${\mathbf P}$ and respects the equivalence relation defining ${\mathbf U}$.
+Thus this gives an action of $(2{\mathbb Z})^2$ on ${\mathbf U}'$.
+We define ${\mathbf M}$ to be the quotient.
+It is depicted in Figure ...}
+%
+%Observe that the translations by pairs of even integers preserves the subset $\mathbf P$ and respects the equivalence relation on $\mathbf U$
+%
+%We will call this quotient space $\mathbf{M}$ and define it as such:
+%\begin{Def}{$\mathbf{M}$.}\\
+%$\mathbf{M}$ is a base surface of $\mathbf{U}'$ given by the quotient $\mathbf{U}'/2\mathbb{Z}\oplus2\mathbb{Z}$, where $2\mathbb{Z}\oplus2\mathbb{Z}\times\mathbf{U}'\rightarrow\mathbf{U}'$ is the group action on $\mathbf{U}'$ defined as $(m,n)\circ(u,v)=(u+m,v+n)$.
+%\end{Def}
+%
+%\begin{figure}[H]
+%\begin{center}
+%\includesvg[width=2in.]{unfoldcutquotient}\hspace{0.2in}
+%\includesvg[width=2in.]{quotient}
+%\end{center}
+%\caption{$\mathbf{U}'$ equivalence classes under group action (left), and $\mathbf{M}$ with identified edges and vertices (right).}
+%\label{fig:quotient}
+%\end{figure}
+%
+%The quotient space is a genus 1 surface that resembles a $2\times2$ torus with a unit square removed from it.
+Near the outer vertex it behaves as a torus might.
+The inner vertices $c_{1},c_{3}$ have cone angle $\frac{3\pi}{2}$, and $c_{2}$ has a cone angle of $3\pi$.
+None of these inner vertices have cone angles that are integral multiples of $2\pi$, so in order to create the necessary translation surfaces we require the minimal number of copies of both $\mathbf{U}$ and $\mathbf{M}$ to construct their associated phase spaces.
+\compat{Move the discussion of building the translation surface to the next section.
+Otherwise this paragraph is good.}
+
+\section{Proof of Main Theorem}
+The following section studies the translation surface's various symmetries and associated Veech group.
+We use the Veech dichotomy to prove our main theorems that determine whether the geodesic is ultimately periodic or drift periodic.
+
+\subsection{The Staircase Model's Symmetries}
+$\tilde{\mathbf{M}}$ has all the same symmetries of the square, as represented by the dihedral group $\mathbb{D}_4$.
+This is made obvious by the fact that $\tilde{\mathbf{M}}$ is tiled by squares.
+We want to only consider elements of $\mathbb{D}_4$ that preserve its orientation.
+So, let $\mathbb{D}^{+}_4$ be the subgroup of the dihedral group whose matrix representations are determinant one, the rotational matrices.
+It then isn't surprising that the order four group of rotations of the square is a subgroup of $\Gamma$, given the manner in which $\tilde{\mathbf{U}}$ is constructed.
+
+Next we show that the group of edge and orientation preserving  affine diffeomorphisms, $\text{Aff}^{+}(\tilde{\mathbf{M}})$, of $\tilde{\mathbf{M}}$ is contained in $\Gamma$
+
+Every pair of edge on the staircase model has an exact distance of two. (Each square is unit). This gives way to a very natural cylindrical decomposition.
 
@@
 \caption{The staircase with edges identified.}
+\label{fig:decomp}
 \end{figure}
@@
 The staircase is then decomposed into 6 uniform cylinders of moduli $\frac{1}{2}$ in both the vertical and horizontal directions.
-All families of saddle connections on the surface intersect exactly two others.
+All families of geodesics on the horizontal cylinders intersect exactly two other families of geodesics on the vertical cylinders $\tilde{\mathbf{M}}$.
 
 \begin{Def}
-The \textbf{family of horizontal cylinders} of $\tilde{\mathbf{M}}$, denoted $\Delta_{i}$ where $1\leq i \leq 6$, are cylinders formed by saddle connections made between singularities $((x_3,x_2)\cup(x_2,x_3)$, $(x_3,x_4))$ when i is odd, and $((x_1,x_4)\cup(x_2,x_4)$, $(x_1,x_2))$ when i is even.
+The \textbf{set of horizontal cylinders} of $\tilde{\mathbf{M}}$, denoted $\Delta_{i}$ $(1\leq i \leq 6)$, are the cylinders with boundaries formed by saddle connections made between unions of singularities in the horizontal direction.
 \newline
-Likewise, the \textbf{family of vertical cylinders}, denoted $\varepsilon_j$ $(1\leq j 
-\leq 6)$, are formed by connections $((x_2,x_3)$, $(x_2,x_1)\cup(x_1,x_2))$ when j is odd, and  $((x_4,x_1)$, $(x_4,x_3)\cup(x_3,x_4))$ when j is even.
+Likewise, the \textbf{set of vertical cylinders}, denoted $\varepsilon_j$ $(1\leq j 
+\leq 6)$, are the cylinders with boundaries formed by saddle connections made between unions of singularities in the vertical direction. \newline
+Figure $\ref{fig:decomp}$ shows how $\tilde{\mathbf{M}}$ is decomposed into these cylinders.
 \end{Def}
@@
 We can see that any parabolic skews of magnitude $\lambda=\pm2$, since $\mu^{-1}{}=2$, preserve the boundaries of the cylinders.
-Figure $\ref{fig:skew}$ showcases how Dehn-twists of this magnitude followed by an Affine transformation matches these boundaries again.
-Consequently, our surface is stabilized by the matrices
+Figure $\ref{fig:skew}$ showcases how Dehn-twists of this magnitude followed by an Affine transformation respects the boundaries of the cylinders that will overlap.
+Hence, our surface is stabilized by the matrices
 
@@
 
+Therefore, $\left<\left[ \hspace{1mm} \begin{matrix}
+				1 & \pm 2\\
+				0 & 1
+			\end{matrix}\hspace{1mm}\right], 
+			\left[ \hspace{1mm} \begin{matrix}
+							1 & 0\\
+							\pm 2 & 1
+						\end{matrix}\hspace{1mm}\right]\right>=\text{Aff}^{+}(\tilde{\mathbf{M}})\subset\Gamma$.
+
+\begin{Def}
+We call the \textbf{subgroup of $\Gamma$ containing the groups generated by rotational and skew isometries of $\tilde{\mathbf{M}}$} $(\text{Aff}^{+}(\tilde{\mathbf{M}}),$ and $\mathbb{D}^{+}_4)$ $\Gamma'$.
+\end{Def}
+
 \newpage
-\subsection{Other symmetries of $\tilde{\mathbf{M}}$}
-A major benefit of $\tilde{\mathbf{M}}$ being tiled by unit squares is that the surface has all of the same symmetries of the square.
-The action of $\mathbb{D}_4$ on $\tilde{\mathbf{M}}$ preserves its boundaries up to translation.
+\subsection{Homology classes of $\tilde{\mathbf{M}}$}
 
-\subsection{Computing the Veech Group of $\tilde{\mathbf{M}}$}
-
+\newpage
 \subsection{Horizontal, Vertical, and Slope One Cylinders of $\tilde{\mathbf{M}}$}
@@
 \end{figure}
+\newpage
+\subsection{Slope Normalization Algorithm}
 
-\subsection{Slope Normalization Algorithm}
+\newpage
 \subsection{Proof of Theorem}

### 2017-07-27 paper.tex
072717

@@
-j\documentclass[]{article}
+\documentclass[]{article}
 
@@
 
-\subsection{$\tilde{\mathbf{M}}$ as a Veech Surface}
-We use basic properties of $\tilde{\mathbf{M}}$ to show that it is Veech, and its Veech group is a lattice in $SL_2(\mathbb{Z})$.
+%\subsection{$\tilde{\mathbf{M}}$ as a Veech Surface}
+%We use basic properties of $\tilde{\mathbf{M}}$ to show that it is Veech, and its Veech group is a lattice in $SL_2(\mathbb{Z})$.
 Recall that translation surfaces are equipped with charts, $\omega$, of maps from the surface to $\mathbb{C}$.
 The Veech group of the surface $\tilde{\mathbf{M}}$ is given as the image of its stabilizer in the projective group $PSL_2(\mathbb{R})$.
-
-\begin{Def}
-Since $\tilde{\mathbf{M}}$ is a translation surface with four singularities, it is a Veech surface that locally behaves like a torus away from these cone singularities.
+%
+%\begin{Def}
+%Since $\tilde{\mathbf{M}}$ is a translation surface with four singularities, it is a Veech surface that locally behaves like a torus away from these cone singularities.
 $\tilde{\mathbf{M}}$ has a \textbf{Veech group}, denoted $\Gamma$.
-\end{Def}
-
-We use the following theorem of Gutkin and Judge to show that $\Gamma$ is lattice (and arithemtic) in $SL_2(\mathbb{Z})$:
-
-\begin{thm}(Gutkin-Judge) A translation surface $(S,\omega)$ is squaretiled
-if and only if its Veech group $V(S,\omega)$ shares a finite-index subgroup
-with $SL_2(\mathbb{Z}).$
-\end{thm}
-
-\begin{thm}The following are equivalent:
-\begin{enumerate}[label=(\roman*)]
-\item $\tilde{\mathbf{M}}$ is tiled by squares.
-\item $\Gamma$ is lattice in $SL_2(\mathbb{Z})$.
-\end{enumerate}
-\begin{proof}
-Figure $\ref{fig:staircase}$ explicitly demonstrates how $\tilde{\mathbf{M}}$ is tiled by squares. The proof then follows as an obvious result of the previous theorem.
-\end{proof}
-\end{thm}
+%\end{Def}
+%
+%We use the following theorem of Gutkin and Judge to show that $\Gamma$ is lattice (and arithemtic) in $SL_2(\mathbb{Z})$:
+%
+%\begin{thm}(Gutkin-Judge) A translation surface $(S,\omega)$ is squaretiled
+%if and only if its Veech group $V(S,\omega)$ shares a finite-index subgroup
+%with $SL_2(\mathbb{Z}).$
+%\end{thm}
+%
+%\begin{thm}The following are equivalent:
+%\begin{enumerate}[label=(\roman*)]
+%\item $\tilde{\mathbf{M}}$ is tiled by squares.
+%\item $\Gamma$ is lattice in $SL_2(\mathbb{Z})$.
+%\end{enumerate}
+%\begin{proof}
+%Figure $\ref{fig:staircase}$ explicitly demonstrates how $\tilde{\mathbf{M}}$ is tiled by squares. The proof then follows as an obvious result of the previous theorem.
+%\end{proof}
+%\end{thm}
 
@@
 The following section studies the translation surface's various symmetries and associated Veech group.
-We use the Veech dichotomy to prove our main theorems that determine whether the geodesic is ultimately periodic or drift periodic.
+In doing so, we establish what sorts of properties of $\tilde{\mathbf{M}}$ remain invariant when undergoing affine transformations.
+Since $\tilde{\mathbf{U}}$ is a cover of $\tilde{\mathbf{M}}$, we will study the first homology classes of $\tilde{\mathbf{M}}$ that realize $\tilde{\mathbf{U}}$ as a $\mathbb{Z}^2$ cover.
+The goal in the end is to construct a basis for the homotopy classes of paths on $\tilde{\mathbf{M}}$ that can be lifted onto $\tilde{\mathbf{U}}$.
+We generalize how the Veech group of $\tilde{\mathbf{M}}$ acts on homology, and show that closed geodesics on the base surface whose lifts onto the $\mathbb{Z}^2$-cover extend to a homomorphism of $\pi_1(\tilde{\mathbf{M}})$ to $\mathbb{Z}^2$.
+The kernel of this homomorphism determines the cover, and any element of the Veech group acting on $\tilde{\mathbf{M}}$ will stabilize the surface and extend to an isomorphism of its homology classes.
 
-\subsection{The Staircase Model's Symmetries}
-$\tilde{\mathbf{M}}$ has all the same symmetries of the square, as represented by the dihedral group $\mathbb{D}_4$.
+\subsection{Veech Group and Symmetries}
+Observe that $\tilde{\mathbf{M}}$ shares symmetries with the square (up to translation).
 This is made obvious by the fact that $\tilde{\mathbf{M}}$ is tiled by squares.
-We want to only consider elements of $\mathbb{D}_4$ that preserve its orientation.
-So, let $\mathbb{D}^{+}_4$ be the subgroup of the dihedral group whose matrix representations are determinant one, the rotational matrices.
-It then isn't surprising that the order four group of rotations of the square is a subgroup of $\Gamma$, given the manner in which $\tilde{\mathbf{U}}$ is constructed.
+Any rotations and reflections acting on $\tilde{\mathbf{M}}$ respects its boundaries.
+We use their matrix representations, and consider only the orientation preserving transformations of determinant one.
 
-Next we show that the group of edge and orientation preserving  affine diffeomorphisms, $\text{Aff}^{+}(\tilde{\mathbf{M}})$, of $\tilde{\mathbf{M}}$ is contained in $\Gamma$
+Theorems of Gutkin and Judge tell us that any Veech surface that is tiled by squares has a Veech group that is a finite-indexed subgroup of $SL_2(\mathbb Z)$.
+Therefore the group is a lattice in $SL_2(\mathbb R)$.
+We will call the Veech group of $\tilde{\mathbf{M}}$, $\Gamma$.
+$\Gamma$ will without a doubt be generated by parabolic elements, such as skew matrices, and $\tilde{\mathbf{M}}$'s various isometries.
 
-Every pair of edge on the staircase model has an exact distance of two. (Each square is unit). This gives way to a very natural cylindrical decomposition.
+It is easily shown that $SO_2(\mathbb Z)\subset\Gamma$.
+Consider a rotation about the center of any square.
+By a series of cutting and gluing operations, $\tilde{\mathbf{M}}$ can be reconstruced as a staircase again, sending one edge to exactly one other.
+
+Each square is unit, and any pair of edges have two squares between them. This gives rise to a very natural cylinder decomposition of $\tilde{\mathbf{M}}$ in both horizontal and vertical directions.
 
@@
 The staircase is then decomposed into 6 uniform cylinders of moduli $\frac{1}{2}$ in both the vertical and horizontal directions.
-All families of geodesics on the horizontal cylinders intersect exactly two other families of geodesics on the vertical cylinders $\tilde{\mathbf{M}}$.
+Any geodesic on the horizontal cylinders are incident with exactly two other geodesics on the vertical cylinders.
 
@@
 \leq 6)$, are the cylinders with boundaries formed by saddle connections made between unions of singularities in the vertical direction. \newline
-Figure $\ref{fig:decomp}$ shows how $\tilde{\mathbf{M}}$ is decomposed into these cylinders.
+Refer to Figure $\ref{fig:decomp}$ for the decomposition
 \end{Def}
 
-Horizontal and vertical cylinders of the surface are disjoint within their respective sets, and the union of both sets is the surface $\tilde{\mathbf{M}}$.
+Horizontal and vertical cylinders of the surface are disjoint within their respective sets, and their union is the surface $\tilde{\mathbf{M}}$.
 
   Recall that the modulus of a cylinder, $\mu$, is given by the ratio of the cylinder's width to its circumference, $\mu{}={}w/c$.
-A surface that is composed entirely out of similar horizontal and vertical cylinders is stabilized by parabolic matrices with magnitude $\lambda$ equal to that of lowest common multiple of the inverse moduli, $\mu^{-1}$, of both horizontal and vertical decompositions.
+A surface that is composed entirely out of similar horizontal and vertical cylinders is stabilized by Dehn-twists whose derivatives are matrices of magnitude $\lambda$ equal to that of lowest common multiple of the inverse moduli, $\mu^{-1}$.
 [cite Veech 89?]
@@
 
-In this case, the moduli are easily obtained and uniform throughout $\tilde{\mathbf{M}}$.
-We can see that any parabolic skews of magnitude $\lambda=\pm2$, since $\mu^{-1}{}=2$, preserve the boundaries of the cylinders.
-Figure $\ref{fig:skew}$ showcases how Dehn-twists of this magnitude followed by an Affine transformation respects the boundaries of the cylinders that will overlap.
-Hence, our surface is stabilized by the matrices
-
+In this case, the moduli are easily obtained and uniform throughout $\tilde{\mathbf{M}}$'s cylinders.
+We can see that any skews of magnitude $\lambda=\pm2$, since $\mu^{-1}{}=2$, are derivatives of Dehn-twists that preserve the boundaries of the cylinders.
+Figure $\ref{fig:skew}$ demonstrates a Dehn-twist of magnitude 2 acting on both horizontal and vertical cylinders.
+Hence the derivatives,
 \begin{equation}
@@
 							\pm 2 & 1
-						\end{matrix}\hspace{1mm}\right].
+						\end{matrix}\hspace{1mm}\right]
 \end{equation}
-
-Therefore, $\left<\left[ \hspace{1mm} \begin{matrix}
-				1 & \pm 2\\
-				0 & 1
-			\end{matrix}\hspace{1mm}\right], 
-			\left[ \hspace{1mm} \begin{matrix}
-							1 & 0\\
-							\pm 2 & 1
-						\end{matrix}\hspace{1mm}\right]\right>=\text{Aff}^{+}(\tilde{\mathbf{M}})\subset\Gamma$.
-
-\begin{Def}
-We call the \textbf{subgroup of $\Gamma$ containing the groups generated by rotational and skew isometries of $\tilde{\mathbf{M}}$} $(\text{Aff}^{+}(\tilde{\mathbf{M}}),$ and $\mathbb{D}^{+}_4)$ $\Gamma'$.
-\end{Def}
+are the parabolic generators of $\Gamma$. The fact that $\lambda=1$ does \emph{not} stabilize $\tilde{\mathbf{M}}$ is an absolutely integral part of the major proof.
 
 \newpage
-\subsection{Homology classes of $\tilde{\mathbf{M}}$}
+\subsection{Defining Homology Classes of $\tilde{\mathbf{M}}$}
+\compav{TODO:\\
+Construct homotopy class 12-gon\\
+Construct staircase with class intersection\\}
+The cylinder decomposition of $\tilde{\mathbf{M}}$ into 12 vertical and horizontal cylinders naturally produce a set of closed homotopic paths that travel along the cylinder's circumference.
+Because of $\tilde{\mathbf{M}}$'s translational symmetries, we consider every cylinder to contain a set of homologous paths, characterized by the geodesic running along the cylinder's meridian.
 
+Since none of these paths can't be contracted into a single point or deformed into another path without crossing over a singularity, they are non-trivial and span the set of homology classes of $\tilde{\mathbf{M}}$.
+We denote the set of closed paths on the cylinders by $\gamma_{0,..,11}$, where the even-indexed paths are horizontal and the odd-indexed paths are vertical.
+(see figure)
+
+From these paths we create the following homology classes that determine our $\mathbb{Z}^2$-cover:
+
+\begin{align*}
+u&= \{-\gamma_2, +\gamma_5,\gamma_8,-\gamma_{11}\},\\
+v&= \{+\gamma_0, -\gamma_3,-\gamma_6,+\gamma_{9}\},\\
+z&= \{+\gamma_1, +\gamma_4,-\gamma_7,-\gamma_{10}\}.
+\end{align*}
+
+\begin{figure}[H]
+\includegraphics[width=3in]{homologyclass.png}
+\centering
+\end{figure}
+
+Here $u,v,$ and $z$ belong to $H_1(\tilde{\mathbf{M}},\mathbb Z)$.
+These classes are generalizations of differential 1-forms, say $dx,dy,dz$, on the four-fold cover whose path integrals on $\tilde{\mathbf{M}}$ map directly to the cover.
+In our case, there is translational symmetry and it is understood that any geodesic with rational slope is periodic, so we need only consider $\tilde{\mathbf{M}}$'s homology classes.
+Let $\alpha$ be a closed path on $\tilde{\mathbf{M}}$ as an element of $\pi_1(\tilde{\mathbf{M}})$.
+The intersection number of any two homology classes is given as the bilinear form:
+
+\begin{align}
+i:H_1(\tilde{\mathbf{M}},\mathbb Z)\times H_1(\tilde{\mathbf{M}},\mathbb Z)\rightarrow \mathbb Z.
+\end{align}
+
+The $\mathbb{Z}^2$ cover, $\tilde{\mathbf{U}}$, is associated to the kernel of the homomorphism
+\begin{align}
+\Omega_{u,v}:\pi_1(\tilde{\mathbf{M}})\rightarrow \mathbb{Z}^2, \alpha\mapsto(i(u,[\alpha]),i(v,[\alpha])).
+\end{align}
+The classes u and v determine where the lifted geodesic terminates on $\tilde{\mathbf{U}}$.
+Careful experimentation and observation shows us that these classes correspond to moving along the horizontal and vertical directions relative to $\tilde{\mathbf{U}}_0$'s hole indexing convention.
+We suspect that the remaining z class is some remnant of how the quotient topology of $\tilde{\mathbf{U}}_0$ identifies a point on the boundary of a hole with another as an element of $\mathbb{Z}/4\mathbb Z$, as these are the cylinders of $\tilde{\mathbf{U}}_0$ that cross from one phase to the next (See Figure $\ref{fig:mtilda}$).
+Regardless, the classes $u,v$ are enough to determine $\tilde{\mathbf{U}}$'s fundamental group as they already take into consideration the effect that the rotations of $\tilde{\mathbf{U}}_0$ has on $\tilde{\mathbf{U}}$'s translational symmetries.
 \newpage
@@
 \subsection{Slope Normalization Algorithm}
-
 \newpage

### 2017-07-27 sanov/sanov.cpp

@@ -1,0 +1,64 @@
+#include <cstdio>
+#include <iostream>
+#include <vector>
+using std::vector;
+#include <string>
+using std::string;
+using std::endl;
+using std::cout;
+
+string seq;
+vector<int> c1;
+vector<int> c2;
+vector<int> c3;
+string A = "A";
+string Ainv = "Ainv";
+string B = "B";
+string Binv = "Binv";
+string R = "R";
+
+void a(vector<int>& vec){
+	vector<int> tmp;
+	while(vec != c1 && vec != c2 && vec != c3) {
+		tmp = vec;
+		if(vec[0] < 0 && vec[1] > 0) {
+			vec[0]=tmp[1];vec[1]=-tmp[0];
+			seq.insert(0,R);
+		}
+		else if(vec[0] < 0 && vec[1] < 0) {
+			vec[0]=-tmp[0];vec[1]=-tmp[1];
+			seq.insert(0,R);seq.insert(0,R);
+		}
+		else if(vec[0] > 0 && vec[1] < 0) {
+			vec[0]=-tmp[1];vec[1]=tmp[0];
+			seq.insert(0,R);seq.insert(0,R);seq.insert(0,R);
+		}
+		else if(vec[1] > vec[0]) {
+			vec[1] = (-2*tmp[0]) + tmp[1];
+			seq.insert(0,Binv);
+		}
+		else {
+			vec[0] = (-2*tmp[1]) + tmp[0];
+			seq.insert(0,Ainv);
+		}
+	}
+}
+
+int main(int argc, char* argv[]) {
+/*	if(argc < 3) {
+		std::cerr<<"Expected two integers"<<endl;
+		return 1;
+	}*/
+	c1.push_back(1);c1.push_back(1);
+	c2.push_back(1);c2.push_back(0);
+	c3.push_back(0);c3.push_back(1);
+	vector<int> v;
+	int x = atoi(argv[0]);
+	int y = atoi(argv[1]);
+	v.push_back(x);
+	v.push_back(y);
+	a(v);
+	cout<<"\n"<<seq<<endl;
+	cout<<"Final vector is ("<<v[0]<<", "<<v[1]<<")\n";
+	return 0;
+}

### 2017-08-01 paper.tex
080117

@@
 \newtheorem*{def*}{Definition}
+\newtheorem{lem}{Lemma}[subsection]
 
@@
 
-\begin{center}
-\includesvg[width=2.3in]{cubesxyzpaths}\hspace{0.1in}
-\raisebox{0.5in}{\includesvg[width=2.2in]{unfoldcutpath}}
-\\ \text{Here we can see paths a,b, and c on }$\mathbf{U}$\text{ and }$\mathbf{S}$,
-\\ \text{where their initial trajectories are (0,1), (1,0), and (1,1) respectively.}
-\end{center}
-
 The figure above shows us some very simple paths on the surfaces.
@@
 \subsection{Four-Fold Cover of $\mathbf{U}$}
-A geodesic on $\mathbf{U}$ viewed as a sequence of line segments on $\mathbf{P}$ travels in at most four directions which differ by rotation by a multiple of $\frac{\pi}{2}$.
-This is a consequence of the edge identifications taking the boundary of a hole on $\mathbf{P}$ to its counterpart by rotation of that factor.
-To each of these directions we associate an element of $\mathbb{Z}/4\mathbb{Z}$ to signify the current phase of the geodesic flow.
-We use this construction to create our four-fold cover of $\mathbf{U}$ as a topological quotient of $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$.
-For every integer pair centered around every hole of $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$, create the edge identifications taking one phase plane to the next in the following manner:
+A geodesic on $\mathbf{U}$ viewed as a sequence of line segments on $\mathbf{P}$ travels in at most four directions which differ by a rotation of a multiple of $\frac{\pi}{2}$.
+This is a consequence of the edge identifications taking the boundary of a hole on $\mathbf{P}$ to its counterpart by a rotation of ninety degrees.
+To each of these directions we associate an element of $\mathbb{Z}/4\mathbb{Z}$ to signify one of the four possible directions of the geodesic flow at any given time.
+Given that, we can create a four-fold cover of $\mathbf{U}$ as a topological quotient of $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$.
+For every integer pair centered around every hole of $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$, create the edge identifications that takes one plane to the next in the following manner:
 
@@
 We define the edge identifications as a relation on $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$.
-For $i\in[-\frac{1}{4},\frac{1}{4}]$:
+For $i\in[-\frac{1}{2},\frac{1}{2}]$:
 
-\begin{multline}
-([2m+i], 2n-\frac{1}{4},j)\sim(2m-\frac{1}{4},[2n+i],j+1)\\
-([2m+i], 2n+\frac{1}{4},j)\sim(2m+\frac{1}{4},[2n+i],j+1)\\
+\begin{equation}
+\begin{split}
+(2m+i, 2n-\frac{1}{2};j)\sim(2m-\frac{1}{2},2n+i;j+1)\\
+(2m+i, 2n+\frac{1}{2};j)\sim(2m+\frac{1}{2},2n+i;j+1)\\
 \vspace{0.2in}\text{for } (m,n;j) \in \mathbb{Z}^{2}\times\mathbb{Z}/4\mathbb{Z}.
 \label{eq:rel2}
-\end{multline}
+\end{split}
+\end{equation}
 
 We define $\tilde{\mathbf{U}}_0$ to be the topological quotient, $(\mathbf{P}\times\mathbb{Z}/4\mathbb{Z})/\sim$.
-$\tilde{\mathbf{U}}_0$ is a Euclidean cone surface with four distinct vertices for each hole centered at $(m,n;\mathbb{Z}/4\mathbb{Z})$ on $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$ of cone angle $6\pi$ with proper identification.
+$\tilde{\mathbf{U}}_0$ is a Euclidean cone surface with four vertices at each hole centered at $(2m,2n;\mathbb{Z}/4\mathbb{Z})$ of cone angle $6\pi$.
 
 \subsection{Symmetries of $\tilde{\mathbf{U}}_0$}
-There are certain translational and rotational symmetries of $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$ that we use to define isometric maps.
-These isometries then induce well-defined isometries on $\tilde{\mathbf{U}}_0$ that respect the edge identifications we made earlier.
+There are certain translational and rotational symmetries of $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$ that we use to define isometries.
+These isometries are well-defined and induce isometries on $\tilde{\mathbf{U}}_0$ that respect the edge identifications we made earlier.
 
 \begin{Def}Let $m,n\in\mathbb{Z}$.
-The \textbf{family of translations}, $\mathbf{T}^{m,n}_{0}$ on $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$, is given as $\mathbf{T}^{m,n}_{0}(u,v;j)=(u+2m,v+2n;j)$.\newline
-The \textbf{family of "rotational" isometries} on $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$ is given as powers of the automorphism $\mathbf{R}(u,v;j)=(u,v;j+1)$, where $\mathbf{R}$ as an action on $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$ forms a group of order 4.
+The {family of translation maps}, $\mathbf{T}^{m,n}_{0}$ on $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$, is given as $\mathbf{T}^{m,n}_{0}(u,v;j)=(u+2m,v+2n;j)$.\newline
+The {family of ``rotational" isometries} on $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$ is given as the automorphism $\mathbf{R}_0(u,v;j)=(u,v;j+1)$, where $\mathbf{R}_0$ is an action on $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$ of order 4.
 \end{Def}
 
-\begin{thm}$\mathbf{T}^{m,n}_{0}$ induces an isometry on $\tilde{\mathbf{U}}_0$.
+\begin{lem}$\mathbf{T}^{m,n}_{0}$ induces a well-defined isometry on $\tilde{\mathbf{U}}_0$.
 \begin{proof}
-This theorem follows from the fact that $\mathbf{T}^{m,n}_{0}$ preserves the edge identifications defined at $\eqref{eq:rel2}$.
-Take $\mathbf{T}^{m,n}_{0}(u,v;j)=(u+2m,v+2n;j)$ for arbitrary integers $m,n$, and $(u,v;j)\in\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$.
-Then, $\mathbf{T}^{m,n}_{0}(u,v;j)$ is trivially isometric for elements not on the boundary of $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$, since these points lie on a subset of a plane and are identified with themselves.
-Suppose then that $(u,v;j)\in\partial\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$.
-Then $\mathbf{T}^{m,n}_{0}(u,v;j)$ acting on the boundary gives a new relation:
-\begin{multline*}
-([2(m+m')+i], 2(n+n')-\frac{1}{4},j')\sim(2(m+m')-\frac{1}{4},[2(n+n')+i],j'+1)\\
-([2(m+m')+i], 2(n+n')+\frac{1}{4},j')\sim(2(m+m')+\frac{1}{4},[2(n+n')+i],j'+1)\\
-\vspace{0.2in}\text{for } (m',n';j') \in \mathbb{Z}^{2}\times\mathbb{Z}/4\mathbb{Z}\text{ and }i\in[-\frac{1}{4},\frac{1}{4}],
-\end{multline*}
-which preserves the lengths of the boundary edges and respects the edge identifications as every hole is centered around a unique pair of even integers, $(2(m+m'),2(n+n')$.
-We denote the induced isometry on $\tilde{\mathbf{U}}_0$ as $\bar{\mathbf{T}}^{m,n}_{0}$.
+It suffices to show that $\mathbf{T}^{m,n}_{0}$ preserves the edge identifications $\eqref{eq:rel2}$.
+Let $x,y\in\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$, where $x=(u_x,v_x;j_x)$ and $y=(u_y,v_y,j_y)$.
+Clearly, if $x=y$, then $\mathbf{T}^{m,n}_{0}(x)=\mathbf{T}^{m,n}_{0}(y)$.
+Hence, $T_0$ is bijective.
+Consider the case where $x\neq y$.
+Let $x\sim y$ and observe that if $x,y$ lie on any of the edges, then $\mathbf{T}^{m,n}_{0}(x)=(u_x+2m,v_x+2n;j_x)$ and $\mathbf{T}^{m,n}_{0}(y)=(u_y+2m,v_y+2n;j_y)$ for $m,n\in\mathbb{Z}$.
+Since $x\sim y$, either $j_y=j_x+1$, or $j_x=j_y+1$.
+Suppose the former is true.
+Then for $i$ in the interval $[-\frac{1}{2},\frac{1}{2}]$:
+
+\begin{equation*}
+\begin{split}
+(2m_x+2m+i, 2n_x+2n-\frac{1}{2};j_x)\sim(2m_x+2m-\frac{1}{2},2n_x+2n+i;j_y)\\\text{ and }\hspace{2in}\\
+(2m_x+2m+i, 2n_x+2n+\frac{1}{2};j_x)\sim(2m_x+2m+\frac{1}{2},2n_x+2n+i;j_y),\\
+\end{split}
+\end{equation*}
+which is the result of $T_0$ acting on the edges, and $m_x,n_x$ are integers that originally ensured that $x\sim y$.
+Since $m,n$ are arbitrarily chosen integers, $T_0$ maps one edge to exactly one other as the edges surrounding each hole are indexed by a unique pair of even integers and $T_0$  is a bijective map that respects that construction.
+A similar argument holds for the case where $j_x=j_y+1$.
+Therefore, $\mathbf{T}^{m,n}_{0}(x)\sim\mathbf{T}^{m,n}_{0}(y)$.
 \end{proof}
-\end{thm}
+\end{lem}
 
-\begin{thm}$\mathbf{R}_{0}$ induces an isometry on $\tilde{\mathbf{U}}_0$.
+We will call the induced isometry on $\tilde{\mathbf{U}}_0$, $\bar{\mathbf{T}}^{m,n}$.
+
+\begin{thm}$\mathbf{R}_{0}$ induces an isometry on $\tilde{\mathbf{U}}_0$
 \begin{proof}
-Similar to the previous theorem, consider a point $(u,v;j)\in\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$.
-Then $\mathbf{R}_{0}(u,v;j)=(u,v;j+1)$ respects edge identifications at $\eqref{eq:rel2}$.
-since $\mathbf{R}_{0}$ is itself a bijective permutation on $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$ that takes every pair $(u,v)$ to itself and permutes the $j$ coordinate by a factor of one.
-Therefore, $\mathbf{R}_{0}$ induces an isometry on $\tilde{\mathbf{U}}_0$.
-We denote this induced isometry by $\bar{\mathbf{R}}_{0}$.
+\compav{might rethink including $R_0$}.
 \end{proof}
@@
  Now we will show that $\tilde{\mathbf{U}}_0$ is isometric to a translation surface via another induced isometry on $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$.
-We would like this isometry to rotate the individual phase planes of $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$ about the origin in such a way that edges one the boundaries are identified strictly by translation.
-Every phase characterized as an element of $\mathbb{Z}/4\mathbb{Z}$ becomes a power of a rotation matrix on the individual phase planes.
+We would like this map to rotate each individual plane of $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$ about the origin in such a way that edges one the boundaries are identified strictly by translation.
+Every plane that is represented as an element of $\mathbb{Z}/4\mathbb{Z}$ is conveniently acted on by a ninety degree rotation matrix, whose power we take to be that very same element.
 
@@
 					u\\v
-				\end{array}\right]
+				\end{array}\right].
 \label{eq:phi}
 \end{equation}
+This is clearly an isometry, as rotations preserve distance and $j\mapsto j$.
 
-\begin{Def} Let $\sim$ be the equivalence relation $\eqref{eq:rel2}$, and $\phi$ be the isometry on $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$ $\eqref{eq:phi}$.
-Define a new relation obtained from $\sim$ as:  $x_1\sim_{\phi}x_2$ if $\phi^{-1}(x_1) = \phi^{-1}(x_2)$ for $x_1,x_2\in\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$.
-We define $\tilde{\mathbf{U}}$ to be  $(\mathbf{P}\times\mathbb{Z}/4\mathbb{Z})/\sim_{\phi}$.\label{eq:simphi}
+\begin{Def} Let $\sim$ be the equivalence relation $\eqref{eq:rel2}$, and $\phi$ be the isometric map on $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$ $\eqref{eq:phi}$.
+Define a new relation obtained from $\sim$ as $x_1\sim_{\phi}x_2$ iff $\phi^{-1}(x_1) \sim \phi^{-1}(x_2)$ for $x_1,x_2\in\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$.
+A new surface is obtained as the topological quotient $\tilde{\mathbf{U}}=$ $(\mathbf{P}\times\mathbb{Z}/4\mathbb{Z})/\sim_{\phi}$.\label{eq:simphi}
 \end{Def}
-
-The map $\phi$ preserves the edge identifications on $\tilde{\mathbf{U}}_0$ by these affine transformations.
-And since $\tilde{\mathbf{U}}_0$ is a cover of $\mathbf{U}$, $\tilde{\mathbf{U}}$ is a translation cover of $\mathbf{U}$.
-Observe that by rotating each phase plane about the origin, the manner in which the holes are indexed undergoes a rotation according to the j-coordinate.
 
@@
 \centering
-\includesvg[width=3.3in]{utildaprime}
+\includesvg[width=3.65in]{utildaprime}
+\caption{$\tilde{\mathbf{U}}$ with edges identified, and holes indexed relative to $\mathbf{P}\times\{0\}$.}
+\label{fig:utilda}
 \end{figure}
@@
 \begin{enumerate}[label=(\roman*)]
-\item The map $\phi$ induces a well-defined isometry $\bar{\phi}:\tilde{\mathbf{U}}_0 \rightarrow \tilde{\mathbf{U}}$
+\item The map $\phi$ induces a well-defined isometry from $\tilde{\mathbf{U}}_0$ to $\tilde{\mathbf{U}}$.
 \item $\tilde{\mathbf{U}}$ is a translation surface.
@@
 \begin{proof}
-The proof of this theorem also follows from the fact that $\phi \eqref{eq:phi}$ is an isometric map that respects edge identifications.
-Let $(u,v;j)\in\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$.
-Then $\phi(u,v;j)=(u',v';j)$.
-Let $M=\left[\begin{matrix}
-0 & 1\\
--1 & 0
-\end{matrix}\right]$, $U_1=([2m+i],2n-\frac{1}{4})$, $U_2=([2m+i],2n+\frac{1}{4})$, $V_1=(2m-\frac{1}{4},[2n+i])$, and $V_2=(2m+\frac{1}{4},[2n+i])$ for $i\in[-\frac{1}{4},\frac{1}{4}]$ and $m,n\in\mathbb{Z}$.
-Observe that making the identification $M^{j}(X)\sim' M^{j+1}(X')$ for some set of vectors $X,X'$ is equivalent to $X\sim' M(X')$ since $M$ is an invertible, isometric linear map.
-Similarly, we can express the relations $\phi(U_1;j)\sim\phi(V_1;j+1)$ and $\phi(U_2;j)\sim\phi(V_2;j+1)$ as $(M^{j}(U_1);j)\sim (M^{j+1}(V_1);j+1)$ and $(M^{j}(U_2);j)\sim (M^{j+1}(V_2);j+1)$, or, more concisely, as
-
-\begin{multline*}
-([2m+i], 2n-\frac{1}{4},0)\sim([2n+i],-2m+\frac{1}{4},1)\\
-([2m+i], 2n+\frac{1}{4},0)\sim([2n+i],-2m-\frac{1}{4},1)\\
-\text{for } (m,n) \in \mathbb{Z}^{2}.
-\end{multline*}
-This is precisely how the relation $\sim_{\phi}$ in Definition $\eqref{eq:simphi}$ is described (taking $\phi^{-1}$ of two points and checking if $\sim$ relates them).
-Since $\phi$ respects edge identifications and preserves boundary lengths, there exists an induced map, $\bar{\phi}:\tilde{\mathbf{U}}_0 \rightarrow \tilde{\mathbf{U}}$ and \emph{(i)} is true.
-Finally, the relation $\sim_{\phi}$ as expressed this way obviously identifies edges from one phase plane to the next by translation, so \emph{(ii)} is true.
+It will again suffice to show that $\phi$ preserves $\tilde{\mathbf{U}}_0$'s edge identifications to prove $(i)$.
+Let $x_1,x_2\in\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$.
+From $\tilde{\mathbf{U}}$'s definition $\eqref{eq:simphi}$, we see that $x_1\sim_{\phi} x_2$ if and only if $\phi^{-1}(x_1) \sim \phi^{-1}(x_2)$.
+By taking $\phi^{-1}(x_1)$ and $ \phi^{-1}(x_2)$ to be the pre-images of the quotient map onto $\tilde{\mathbf{U}}_0$, it is clear that $\phi$ preserves identifications.
+Otherwise, it would contradict the definition.
+Furthermore, since $\phi$ is a well-defined isometry, there is an induced isometry on the topological quotients, $\tilde{\mathbf{U}}_0$ and $\tilde{\mathbf{U}}$.\newline
+Let $x_1\sim_{\phi} x_2$.
+Then $\phi^{-1}(x_1) \sim \phi^{-1}(x_2)$.
+Let $\phi^{-1}(x_1)=(u_1,v_1;j_1)$ and $\phi^{-1}(x_2)=(u_2,v_2;j_2)$.
+Without loss of generality, suppose that $j_2=j_1+1$.
+Let $i$ vary over the interval $[-\frac{1}{2},\frac{1}{2}]$, and $m,n$ be integers such that they satisfy the relation $\phi^{-1}(x_1) \sim \phi^{-1}(x_2)$.
+If $\phi$ acts on these these two points, then $x_1 = (u_1',v_1';j_1)$ and $x_2 = (u_2',v_2';j_1+1)$.
+But we know that $(u_1',v_1')=M^{j_1}(u_1,v_1)$ and $M^{j_1+1}(u_2',v_2')=(u_2,v_2)$, where $M$ is the rotation matrix as in $\phi$.
+Thus, $M^{j_1}(u_1,v_1)\times\{j_1\}\sim_\phi M^{j_1+1}(u_2,v_2)\times\{j_1+1\}$.
+By multiplying through by $M^{-j_1}$ prior to taking the cross-product, we obtain $(u_1,v_1)\times\{j_1\}\sim_\phi M(u_2,v_2)\times\{j_1+1\}$.
+Since $M(u_2,v_2) = (v_2,-u_2)$, we can express the relation $\sim_\phi$ as
+\begin{equation*}
+\begin{split}
+(2m+i, 2n-\frac{1}{2};j_1)\sim_\phi(2n+i, -(2m-\frac{1}{2});j_1+1),\\
+\text{ and }\hspace{1.2in}\\
+(2m+i, 2n+\frac{1}{2};j_1)\sim_\phi(2n+i, -(2m+\frac{1}{2});j_1+1).\\
+\end{split}
+\end{equation*}
+It is clear from this relation and figure $\ref{fig:utilda}$ that $\tilde{\mathbf{U}}$ is a translation surface.
 \end{proof}
 \end{thm}
+We will denote this induced isometry as $\bar{\phi}$.
 
@@
 
-\subsection{Symmetries of $\tilde{\mathbf{U}}$}
+\subsection{Constructing $\tilde{\mathbf{U}}$'s base surface}
 In this section we use the induced isometries on $\tilde{\mathbf{U}}_0$, and the map $\bar{\phi}:\tilde{\mathbf{U}}_0\rightarrow\tilde{\mathbf{U}}$ to define a set of rotational and translational symmetries of $\tilde{\mathbf{U}}$ such that its quotient over the action of these maps are faithful representations of $\tilde{\mathbf{U}}_0$ in terms of how the holes in $\mathbf{P}\times\mathbb{Z}/4\mathbb{Z}$ are indexed.
@@
 
-\begin{Def}We call the \textbf{family of faithful Translational symmetries} $\mathbf{T}^{m,n}$ on $\tilde{\mathbf{U}}$ and define it to be the composition of induced isometries, $\mathbf{T}^{m,n}$ = $\bar{\phi}\circ\bar{\mathbf{T}}^{m,n}\circ\bar{\phi^{-1}}$.
+\begin{Def}We call the {family of $\tilde{\mathbf{U}}$'s translational symmetries} $\mathbf{T}^{m,n}$ and define it to be the composition of the induced isometric maps: $\mathbf{T}^{m,n}$ = $\bar{\phi}\circ\bar{\mathbf{T}}^{m,n}\circ\bar{\phi}^{-1}$.
 \end{Def}
 
-We can see that $\mathbf{T}^{m,n}$ is indeed an isometry on $\tilde{\mathbf{U}}$ as it is composed of isometries.
-We take the quotient over the group generated from these translations, and show that this new surface is Veech.
+We can see that $\mathbf{T}^{m,n}$ is indeed an isometry on $\tilde{\mathbf{U}}$ as it is composed of the induced isometries, $\bar{\phi}$ and $\bar{\mathbf{T}}^{m,n}$ .
+We can take the quotient of the translation surface $\tilde{\mathbf{U}}$ over the action of $\mathbf{T}^{m,n}$ to obtain a base surface that realizes $\tilde{\mathbf{U}}$ as an infinite $\mathbb{Z}^2$ cover.
 
 \begin{Def}
-The translation surface $\tilde{\mathbf{M}}$ is given as the quotient  $\tilde{\mathbf{U}}/\mathbf{T}^{m,n}$.
+The translation surface $\mathbf{M}$ is given as the quotient  $\tilde{\mathbf{U}}/\mathbf{T}^{m,n}$.
 \end{Def}
@@
 \includesvg[width=4in]{mtilda}
-\caption{The surface $\tilde{\mathbf{M}}$ with labeled vertices and identified edges}
+\caption{The surface $\mathbf{M}$ with labeled vertices and identified edges. Labels I, II, III, and IV identify the individual planes indexed by \{0,1,2,3\}, respectively.}
 \label{fig:mtilda}
@@
 
-$\tilde{\mathbf{M}}$ is a genus 5 translation surface that is covered by $\tilde{\mathbf{U}}$.
-Thus we can take a path on $\tilde{\mathbf{M}}$, lift it onto $\tilde{\mathbf{U}}$, take it to $\tilde{\mathbf{U}}_0$, and project it onto $\mathbf U$.
-And since $\mathbf{U}\simeq\mathbf{S}$, the core of our argument is that geodesics on the translation surface $\tilde{\mathbf{M}}$ equipped with homology classes are representations of geodesics on $\mathbf{S}$ away from the singularities.
-The labeled vertices in Figure $\ref{fig:mtilda}$ are all cone singularities with total angles of $6\pi$.
+The new surface, $\mathbf M$, is a genus 5 translation surface with the surjective quotient map $f:\tilde{\mathbf{U}}\rightarrow\mathbf M$ that sends every point of $\tilde{\mathbf{U}}$ to its modular equivalent.
+$\mathbf M$ is a Euclidean cone surface with four singularites of cone angle $6\pi$ on the corners of the square cut out from each of the squares labeled I,II,III, and IV.
+Those squares have their opposite edges identified by translation.
+$\mathbf M$ might be thought of as a collection of four $2\times2$ square tori with a unit square cut out from their centers, and identifications made on these edges so as to associate one torus with exactly two others.
+Although a single edge only relates two of the tori, the vertices in the centers are the same throughout $\mathbf M$.
+\newpage
 
-Although it is not immediately obvious, $\tilde{\mathbf{M}}$ can be realized as a square-tiled surface in the form of a staircase as in $\ref{fig:staircase}$
+\subsection{$\mathbf M$ as a Veech Surface}
+By a series of cutting and gluing operations, $\mathbf M$ has this symmetric representation as a square-tiled staircase (Figure $\ref{fig:staircase}$).
+The vertices of cone angle $2\pi$ can be found on the centers of the squares in the figure below where the black lines intersect.
+\begin{figure}[H]
+\centering
+\includesvg[width=4in]{mtildastaircase}
+\caption{$\mathbf{M}$ as a square-tiled staircase with cone singularities identified.}
+\label{fig:staircase}
+\end{figure}
+As a result of Gutkin-Judge [citation], $\mathbf M$ is a Veech surface whose Veech group is commensurable to $SL(2,\mathbb Z)$.
+In order to determine which elements of orientation preserving automorphisms, denoted Aut$^{+}(\mathbf M)$, have derivatives that belong to the Veech group of $\mathbf M$, SL($\mathbf M$), it is necessary to study $\mathbf M$'s various symmetries and cylinder decompositions.
+
+The derivatives of these automorphisms form the affine group, Aff$^+(\mathbf M)$.
+One such group contained in Aff$^+(\mathbf M)$ is SO$(2,\mathbb{Z})$, the group generated by rotations of order 4.
+Even though the other elements in $\mathbb{D}_4$ are derivatives of automorphisms of $\mathbf M$, we will only consider those of determinant +1.
+
+Observe from the following image that $\mathbf M$ has a uniform cylinder decomposition in both the horizontal and vertical directions:
 
@@
 \centering
-\includesvg[width=4in]{mtildastaircase}
-\caption{$\tilde{\mathbf{M}}$ as a square-tiled staircase with vertices and phases identified.}
-\label{fig:staircase}
+\includegraphics[width=4in]{cylinderdecomp.pdf}
+\caption{Horizontal and Verical Cylinders overlayed.}
+\label{fig:decomp}
 \end{figure}
 
+Both sets of cylinders are indexed $0\leq i, j \leq 5$.
+Every cylinder is composed of two unit squares with edges identified by translation.
+We start counting the cylinders from the bottom-most cylinder and increment the index as the cylinders fill the surface going "up" the staircase.
+$\small{\mathcal{E}}_0$ is the first cylinder at the bottom whose identifications are \em{not} \normalfont the solid triangles.
+$\small{\mathcal{E}}_5$ is the cylinder whose edges are paired by the solid triangles.
+$\Delta_{i}$ has the more standard representation starting from the bottom-most horizontal cylinder.
+
+We define the \emph{inverse modulus} of a cylinder to be the ratio of a cylinders circumference to its width, $\mu^{-1}=\frac{c}{h}$.
+A \emph{Dehn-twist} of a cylinder is an automorphism of the cylinder of the form  $(x,y)\mapsto(x,y+\mu^{-1} x \mod{h})$, where $x$ and $y$ are points on the cylinder relative to its circumference and height respectively.
+According to [need to cite something], a translation surface whose finite cylinder decompositions in orthogonal directions form a set of inverse moduli.
+The lowest (rational) common multiple of the set of all these values match up to give global affine diffeomorphisms of the surface, whose matrix representations are
+\begin{equation*}
+\left[ \hspace{1mm} \begin{matrix}
+				1 & \pm \lambda\\
+				0 & 1
+			\end{matrix}\hspace{1mm}\right] \text{ and }
+			\left[ \hspace{1mm} \begin{matrix}
+							1 & 0\\
+							\pm \lambda & 1
+						\end{matrix}\hspace{1mm}\right],
+\end{equation*}
+where $\lambda$ is the lowest common multiple.
+
+Fortunately, the cylinders $\Delta_{i}$ and $\small{\mathcal{E}}_j$, as i and j vary, all have the same inverse moduli.
+Therefore, $\lambda=$ l.c.m.($\mu^{-1}_{\Delta_i}\cup\mu^{-1}_{\mathcal{E}_j}$) $=2$.
+
+\begin{figure}[H]
+\centering
+\includesvg[width=2.6in]{cylinderskew}
+\caption{Stabilizing Dehn-twists of the horizontal/vertical cylinders}
+\label{fig:skew}
+\end{figure}
+
+Consequently, 
+\begin{equation}
+\left< \left[ \hspace{1mm} \begin{matrix}
+				1 &  2\\
+				0 & 1
+			\end{matrix}\hspace{1mm}\right] \text{ , }
+			\left[ \hspace{1mm} \begin{matrix}
+							1 & 0\\
+							 2 & 1
+						\end{matrix}\hspace{1mm}\right] \right>
+						\subset \text{SL}(\mathbf M).
+\end{equation}
+This group is also known as the \emph{Sanov Subgroup}, a free group of rank 2.
+We take the group generated by ninety degree rotations and these two elements to be the group $\mathbb{X}\subset \text{SL}(\mathbf M)$.
+\newpage
+\subsection{$\mathbb X$'s Action on $\mathbf M$'s Homology}
+The group $\mathbb X$ consists of the set of all derivatives of Aut$^+(\mathbf M)$, whose action on the absolute homology of $\mathbf{M}$ extends to an action on the homology of $\mathbf{M}$'s $\mathbb{Z}^2$-cover, $\tilde{\mathbf{U}}$.
+$\mathbf{M}$'s cylinder decomposition gives way to a set of homology classes that serve as a basis for $H_1(\mathbf{M},\mathbb Z)$.
+For a total of 12 cylinders, there are 12 curves:
+
+\begin{figure}[H]
+\includegraphics[width=3in]{homologyclass.png}
+\centering
+\end{figure}
+Every curve intersects two others.
+By constructing a $12\times12$ matrix where the rows represent each curve and the columns represent their intersections with other curves, we can see that the rank of said matrix is 10.
+This is to be expected as the genus of $\mathbf M$ is 5.
+From these curves we construct homology classes $u,v,$ and $z$.
+$u$ and $v$ are the curves that were the outermost edges of the tori in $\mathbf{M}$'s original representation (Figure $\ref{fig:mtilda}$).
+We can see from Figure $\ref{fig:utilda}$ that crossing over one of these edges corresponds to a shift of the even integer index pair that determines exactly which of the square shaped holes a point on the surface is near to.
+The $z$ class of cylinder curves lie on the cylinders that take a path from one plane to the next in our cover.
+Since one plane is related to exactly two others, it is impossible to go from the first to the third, or the second to the fourth, and vice versa.
+
+Regardless, $u$ and $v$ are the most important homology classes that we use to determine whether or not a path properly closes on the four-fold cover.
+Let $\alpha:[0,1]\rightarrow\mathbf M$ be a curve on the surface such that $\alpha(0)=\alpha(1)=x_0\in\mathbf{M}$.
+Take $f:\tilde{\mathbf U}\rightarrow\mathbf M$ to be the covering map that takes every point on $\tilde{\mathbf U}$ to its quotient in $\mathbf{M}$.
+We will fix a point $\tilde{x}_0\in\tilde{\mathbf U}$ such that the lifted path $\tilde\alpha(0)=\tilde{x}_0$ under the function $f$.
+$\alpha$ is a closed path in $\pi_1(\mathbf M,x_0)$.
+We denoted the class of paths $\alpha$ as [$\alpha$]$\in H_1(\tilde{\mathbf{M}},\mathbb Z)$, and define the function,
+
+\begin{align}
+i:H_1(\mathbf{M},\mathbb Z)\times H_1(\mathbf{M},\mathbb Z)\rightarrow \mathbb Z,
+\end{align}
+to return the intersection number of two homology classes.
+Out of convention, we consider a positive intersection to adhere to the right-hand rule and take the crossings to be relative to $u$ and $v$.
+With that being said, the signed homology classes of gamma curves are
+
+\begin{align*}
+u &= -\gamma_2 +\gamma_5 + \gamma_8 - \gamma_{11},\\
+v &= +\gamma_0 -\gamma_3 -\gamma_6 +\gamma_9,\\
+z &= +\gamma_1 +\gamma_4-\gamma_7-\gamma_{10}.
+\end{align*}
+
+The $\mathbb{Z}^2$ cover, $\tilde{\mathbf{U}}$, is associated to the kernel of the homomorphism
+\begin{align}
+\Omega_{u,v}:\pi_1(\mathbf{M},x_0)\rightarrow \mathbb{Z}^2, \text{ such that } \alpha\mapsto(i(u,[\alpha]),i(v,[\alpha])).
+\end{align}
+Suppose that $h:\mathbf{M}\rightarrow\mathbf{M}$ is an automorphism of the surface such that dir $h\in SL(\mathbf M)$.
+$h$ induces the isomorphism, $h_*:\pi_1(\mathbf{M},x_0)\rightarrow \pi_1(\mathbf{M},x_0)$.
+Likewise, $f$ induces the inclusion map $f_*:\pi_1(\tilde{\mathbf{U}},\tilde x_0)\hookrightarrow\pi_1(\mathbf{M},x_0)$.
+If $\pi_1(\tilde{\mathbf{U}},\tilde x_0) = $ Ker $\Omega_{u,v}$, then it must be true that the kernel of such a homomorphism is preserved under $h_*$.
+
+The actions of $h_*$ are dual to the actions of dir $h$, denoted $h'$, on the sets of odd-odd and even-odd $\mathbb{Z}^2$ vectors, $\mathcal{O}$ and $\mathcal{E}$.
+In particular, the action of $\mathbb X$ on $\mathcal{O}$ and $\mathcal{E}$ is self-contained and well-defined.
+This is vital as what we want to show is that $h'(\alpha')$, such that $\alpha'$ is a constant, is dual to $h_*([\alpha])$.
+In which case, it should also be true that $h_*(\alpha)\mapsto(i(u,h_*([\alpha])),i(v,h_*([\alpha])))$, under $i$.
+And since intersection number is a non-degenerate bilinear form,
+
+\begin{align*}
+\alpha\mapsto(i(h_*^{-1}(u),[\alpha]),i(h_*^{-1}(v),[\alpha])).
+\end{align*}
+
+This amounts to finding some such path whose lift belongs to the fundamental group of $\tilde{\mathbf U}$, noting the effect $h_*$ has on $\mathbf M$'s homology, and finding the new intersection number.
+Let $\beta$ be such a path.
+Suppose there is some sequence of affine maps $h_1,...,h_n\in$Aff$^+(\mathbf{M})$ such that $h_1' \circ ...
+\circ h_n'(\alpha')=\beta'$.
+We will show the following is true:
+
+\begin{align*}
+h_1 \circ ... \circ h_n(\mathbf{M})=\mathbf{M}'&\iff\\
+h_{*_1} \circ ... \circ h_{*_n}([\beta])=[\alpha] &\iff
+\end{align*}
 %\subsection{$\tilde{\mathbf{M}}$ as a Veech Surface}
@@
 
-\begin{figure}[H]
-\centering
-\includegraphics[width=4in]{cylinderdecomp.pdf}
-\caption{The staircase with edges identified.}
-\label{fig:decomp}
-\end{figure}
-
 The staircase is then decomposed into 6 uniform cylinders of moduli $\frac{1}{2}$ in both the vertical and horizontal directions.
@@
 [cite Veech 89?]
-
-\begin{figure}[H]
-\centering
-\includesvg[width=2.6in]{cylinderskew}
-\caption{Stabilizing Dehn-twists of the horizontal/vertical cylinders}
-\label{fig:skew}
-\end{figure}
 
@@
 
-\begin{align*}
-u&= \{-\gamma_2, +\gamma_5,\gamma_8,-\gamma_{11}\},\\
-v&= \{+\gamma_0, -\gamma_3,-\gamma_6,+\gamma_{9}\},\\
-z&= \{+\gamma_1, +\gamma_4,-\gamma_7,-\gamma_{10}\}.
-\end{align*}
-
-\begin{figure}[H]
-\includegraphics[width=3in]{homologyclass.png}
-\centering
-\end{figure}
-
 Here $u,v,$ and $z$ belong to $H_1(\tilde{\mathbf{M}},\mathbb Z)$.
@@
 
+\begin{figure}[H]
+\includesvg[width=2.3in]{cubesxyzpaths}\hspace{0.1in}
+\raisebox{0.5in}{\includesvg[width=2.2in]{unfoldcutpath}}
+\\ \text{Here we can see paths a,b, and c on }$\mathbf{U}$\text{ and }$\mathbf{S}$,
+\\ \text{where their initial trajectories are (0,1), (1,0), and (1,1) respectively.}
+\end{figure}
+
 \end{document}

### 2017-08-01 paper.tex
lastminute

@@
 $u$ and $v$ are the curves that were the outermost edges of the tori in $\mathbf{M}$'s original representation (Figure $\ref{fig:mtilda}$).
-We can see from Figure $\ref{fig:utilda}$ that crossing over one of these edges corresponds to a shift of the even integer index pair that determines exactly which of the square shaped holes a point on the surface is near to.
+We can see from Figure $\ref{fig:utilda}$ that crossing over one of these edges corresponds to a shift of the even integer index pair that determines exactly which of the quotients of $\tilde{\mathbf{U}}$ a terminal point of a lifted path on the surface belongs to.
+That point is identified with the triple $(m,n,\mathbb Z/4\mathbb{Z})$, with integers $m,n$.
 The $z$ class of cylinder curves lie on the cylinders that take a path from one plane to the next in our cover.
-Since one plane is related to exactly two others, it is impossible to go from the first to the third, or the second to the fourth, and vice versa.
-
-Regardless, $u$ and $v$ are the most important homology classes that we use to determine whether or not a path properly closes on the four-fold cover.
+Since one plane is related to exactly two others, it is impossible to go directly from the first to the third or the second to the fourth, and vice versa.
+
+Regardless, $u$ and $v$ are the most important homology classes that we use to determine whether or not a lifted path properly closes on the four-fold cover.
 Let $\alpha:[0,1]\rightarrow\mathbf M$ be a curve on the surface such that $\alpha(0)=\alpha(1)=x_0\in\mathbf{M}$.
@@
 $\alpha$ is a closed path in $\pi_1(\mathbf M,x_0)$.
-We denoted the class of paths $\alpha$ as [$\alpha$]$\in H_1(\tilde{\mathbf{M}},\mathbb Z)$, and define the function,
+We denoted the class of abelianized paths $\alpha$ as [$\alpha$]$\in H_1(\tilde{\mathbf{M}},\mathbb Z)$, and define the function,
 
@@
 \end{align}
+
 to return the intersection number of two homology classes.
@@
 Likewise, $f$ induces the inclusion map $f_*:\pi_1(\tilde{\mathbf{U}},\tilde x_0)\hookrightarrow\pi_1(\mathbf{M},x_0)$.
-If $\pi_1(\tilde{\mathbf{U}},\tilde x_0) = $ Ker $\Omega_{u,v}$, then it must be true that the kernel of such a homomorphism is preserved under $h_*$.
+If $\pi_1(\tilde{\mathbf{U}},\tilde x_0)$ is associated with the Ker $\Omega_{u,v}$, then it must be true that the kernel of such a homomorphism is preserved under $h_*$ and there exists a map from  $\pi_1(\mathbf M, x_0)$ to $\pi_1(\tilde{\mathbf{U}},\tilde x_0)$.
 
@@
 In particular, the action of $\mathbb X$ on $\mathcal{O}$ and $\mathcal{E}$ is self-contained and well-defined.
-This is vital as what we want to show is that $h'(\alpha')$, such that $\alpha'$ is a constant, is dual to $h_*([\alpha])$.
+This is vital as what we want to show is that $h'(\alpha')$, where $\alpha'$ is a constant and $\alpha$ is a closed geodesic, is dual to $h_*([\alpha])$.
 In which case, it should also be true that $h_*(\alpha)\mapsto(i(u,h_*([\alpha])),i(v,h_*([\alpha])))$, under $i$.

### 2017-08-04 paper.tex
080317

@@
 \newtheorem{lem}{Lemma}[subsection]
+\newtheorem*{rem}{Remark}
 
@@
 %opening
-\title{Geodesic Flow on the Necker Cube Surface}
+\title{Recurrence of Geodesic Flow on the Necker Cube Surface}
 \date{}
@@
 \begin{abstract}
-This paper classifies the periodic and drift-periodic trajectories on the Necker cube surface, a periodic surface built out of squares popular in optical illusions.
+\compav{Still working on the abstract}
+This paper classifies the periodic and drift-periodic rational trajectories on the Necker cube surface, a periodic surface built out of squares popular in optical illusions.
+By studying the branched cover of the Necker cube with translational and rotational symmetries, we can deduce from the structure of its quotient exactly which conditions of the initial trajectory angles are necessary to close on the Necker surface.
+The $\mathbb Z^2$-cover of the quotient is a branched cover of the Necker cube.
+As such, any closed paths lifted onto the cover will close on the surface as well.
+An affine automorphism of the quotient extends to an isomorphism of its absolute Homology, which is dual to its derivative, an element of the surface's Veech group, acting on a paths rational direction identified with $\mathbb{Z}^2$.
 \end{abstract}
@@
 
-The purpose of the following paper is to study the geodesic flow on the periodic surface depicted above, which is formed from an infinite collection of unit squares.
-We call this surface the Necker cube surface after Louis Albert Necker's ``Necker Cube$_{1}$,'' which became a popular optical illusion.
-M.C.
-Escher's adaptation of the cube in Metamorphosis I$_{2}$ briefly features the Necker cube surface (See Figure $\ref{fig:Escher}$).The term Necker Cube is used interchangeably to refer to Escher's surface adaptation and Necker's optical illusion.
-For the purposes of the following paper, we will be using either Necker Cube or $\mathbf{S}$ to refer to the periodic cube tiling that appears in Escher's work.
+\compav{Started here on Aug. 3 (I've made changes all the way until my next blue comment. )}
+
+The Necker cube[citation] has made numerous appearances throughout history.
+The artist famous for his use of optical illusions in his work, M.C.
+Escher, has occasionally used tilings of the Necker cube such as in the lithographs ``Metamorphosis I," and ``Convex and Concave." (Pictured below.) The crystallographer Louis Albert Necker was credited for having extensively studied the Necker cube's geometry[citation] and remarked on this simple, yet fascinating, optical illusion that appears to ``simultaneously protrude from and intrude into the page."[citation:wolfram] Those who feel nostalgic might have even realized it is the same board in which ``Q*bert" hops around on in the 1982 arcade game.
+[cite?]
 
@@
 \begin{center}
-\includegraphics[scale=0.5]{escher.jpg}
-\caption{The Necker cube as it appears briefly in Metamorphosis I}
+\includegraphics[scale=0.4]{escher.jpg}
+\includegraphics[scale=0.55]{escher2.jpg}
+\caption{The Necker cube tiling as it appears briefly in ``Metamorphosis I," and hung on a banner in ``Convex and Concave."[citation]}
 \label{fig:Escher}
@@
 
-The initial (rational) trajectories of any given flow fall into only one of two categories.
-\compat{The definitions below seem redundant. Perhaps you want ${\mathcal O}$ and ${\mathcal E}$}
+The tiling of the Necker cube is what we mean when we refer to the Necker cube surface, denoted $\mathbf S$.
+There are $\mathbb R^3$ translational symmetries of its embedded form in space that allows it to be tiled periodically.
+Studying such a surface naturally begs the question of whether or not straight-line paths traveling on the surface are ever periodic.
+And if so, is there a way to classify all such paths?
+Could there even be some intrinsic property
+\begin{figure}[H]
+\centering
+\includesvg[width=1.5in]{cubesxyzpaths2}
+\caption{Some periodic and drift-periodic trajectories show us that the inital direction of a geodesic most likely determines its periodicity.}
+\end{figure}
+
+In fact there is a remarkably simple solution to this problem. Let's begin by defining the following subsets of $\mathbb{Z}^2$:
 
@@
 
-We consider every (rational) initial trajectory on the surface to fall into one of the two categories, otherwise it is a positive scalar multiple of some direction.
-Given some initial point $x_{0}$ on the surface embedded in real space and some vector direction in real space, $\bar{v}$, our initial trajectory,$v$, is taken as an isomorphism from the planar subspace of $\mathbb{R}^{3}$ that contains $\bar{v}$.
-The three possible cases are the auxiliary planes, $x=0, y=0, z=0$, whose projections onto $\mathbb{R}^{2}$ are given by $(0,y,z)\mapsto(-z,y)$, $(x,0,z)\mapsto(x,-z)$, and $(x,y,0)\mapsto(x,y)$ respectively and relative to the standard bases of $\mathbb{R}^{3}$.
-We take $-z$ to be positive, and consider vectors that lie outside of these subspaces as degenerate cases.
-From there we prove the following:
+By a fixing a point on the surface $\mathbf S$ and taking a sufficiently small segment of a geodesic (such that it and the point lie on exactly one face of a cube), we construct classes of directions according to the vector direction that the segment travels.
+The direction is obtained by projecting the segment onto a plane parallel to the face it lies on.
 
-\begin{thm*}{Given an initial trajectory, $\theta\in\mathbb{R}^{2}$, and geodesic path, $\gamma$, on $\mathbf{S}$ with initial trajectory $\theta$, the following is true:} \compat{There is no flow in a fixed direction because when trajectories move to new squares they may be traveling different directions.}
-\compav{Can I use gamma as a path on S?}
-\begin{enumerate}[label=(\roman*)]
-\item If $\theta$ belongs to $\mathcal{O}$, then $\gamma$ is closed on $\mathbf{S}$ and homeomorphic to $\mathbb{R}/\mathbb{Z}$.
-\item If $\theta$ belongs to $\mathcal{E}$, then $\gamma$ is drift-periodic on $\mathbf{S}$.
+\begin{figure}[H]
+\centering
+\includesvg{vectorplane}
+\end{figure}
 
-\end{enumerate}
+The set of equivalence classes of directions on $\mathbf S$ is $\mathbb R/\frac{\pi}{2}\mathbb Z$ (a consequence of $\mathbf S$'s rotational symmetries), whose rational angles are then identified with the sets $\mathcal{O}$ and $\mathcal{E}$.
+Now if the geodesic has an initial direction identified with an element of $\mathcal{O}$, it is periodic on $\mathbf S$.
+But if that direction is identified with an element of $\mathcal{E}$, then the geodesic is drift-periodic.
+Otherwise, the geodesic has an irrational initial trajectory and is [yet to be decided] on $\mathbf S$.
+More formally stated, we have the following theorem:
+
+\begin{thm*}
+Yet to be written
 \end{thm*}
 
-\compav{Archiving for later:\\If $\theta$ is of the form (x,y) such that x or y are irrational, then $\phi^{\theta}_{t}$ is uniquely ergodic on $\mathbf{S}$.
-\compat{We should move the discussion of ergodicity outside the theorem.}	}
+\subsection{Some Background}
+This paper assumes that the reader is familiar with fundamental concepts in algebraic topology, such as covering space theory, and the study of Veech surfaces and their associated Veech groups.
+In order to prove this theorem, we take a number of covers of a surface isometric to $\mathbf S$.
+Our goal then is to take a branched cover of the surface as four "copies" of that surface, and rotate these copies so as to align the geodesic flow and identify edges solely by translation.
+The quotient space obtained by modding out this surface by its translational symmetries is translation and Veech.
 
-\compav{Stopped here. Continued in Section 2.}
+The Veech surface has an $SL(2,\mathbb{R})$-commensurable Veech group characterized as derivatives of affine maps of the translation surface.
+Such affine diffeomorphisms of the surface induce isomorphisms of its absolute homology (with coefficients in $\mathbb Q$).
+We obtain the direction of a closed geodesic flow from its holonomy vector and show that the odd-odd directions closing on the translation cover, and consequently $\mathbf S$, remains invariant given that the flow of a geodesic with slope one closes on the cover.
+This translation cover is considered a $\mathbb{Z}^2$-cover of the veech surface.
+We associate its fundamental group with the kernel of a homomorphism from the closed paths of its Veech surface quotient to $\mathbb{Z}^2$.
+There are numerous dualizations to be made between homologies, homology classes, and $SL(2,\mathbb{R})$ action on $\mathbb{Z}^2$ in order to solve this problem.
 
-The Necker Cube is a member of the family of infinite Euclidean cone surfaces built out of rectangles with vertices of cone angle $3\pi$ and $\frac{3\pi}{2}$.
-We wish to study the surface in the case where we glue only squares with edges of length 1.
-Specifically, we want to study geodesic flows on this surface and to prove that properties of the initial trajectory angle determine whether or not the geodesic will close or drift periodically.
+\begin{rem}
+The initial approach of taking four ``copies" of the surface as a branched cover of $\mathbf S$ was in many ways inspired by the method in which [such and such]'s work on the dynamics of geodesic flow on the Ehrenfest-Wind Tree Model.[citation].
+In fact the Wind Tree and the Necker cube surface have much in common in terms of which initial trajectory angles determine periodicity.
+The two differ, however, in regards to whether or not the initial point of the flow has an effect on periodicity.
+Although the translation surfaces obtained are near identical, the manners in which they are obtain differ greatly.
+A billiard flow is realigned by a series of reflections of the original surface by stringing copies of polygons along a ``laundry line" until a translation surface is obtained.
+On the other hand, $\mathbf S$ exhibits rotational symmetries that we exploit to obtain our translation surface.
+\end{rem}
+\subsection{Acknowledgements}
+to do:
 
-The determining factor is simply whether or not the trajectory angle vector direction $(a,b)$, relative to the bottom and left edges of any square face, has integer values $a$,$b$ (where $a$ and $b$ are relatively prime) are of the same parity (odd-odd slope) or different parities (even-odd slope).
-We choose to neglect cases where paths are incident with vertices, and proceed to treat them as singularities.
-With these conditions in place we found that any geodesic with an initial odd-odd trajectory closes on the surface, while the even-odd cases drift periodically, tending to infinity.
-We aim to prove the following:
+- Hooper
 
-\subsection{Discussion of Proof}
-\subsection{Related work}
+- Artigiani "Exceptional Ergodic Directions" for Notation
 
-To better understand this surface, we will adapt methods used in the study of Veech surfaces to suit our needs.
-These methods have become incredibly valuable to anyone studying rational billiard flows on a variety of compact or (in the case of infinite surfaces) non-compact surfaces.
-Geodesic flows in these problems are characterized as point masses moving along frictionless surfaces that experience perfectly elastic collisions with the boundaries of the surface.
-The straight line path is then deformed by a reflection over the boundary in the opposite direction.
-Translation surfaces constructed in the following manner therefore take on symmetric properties in accordance to these transformations and the types of boundaries the point mass might encounter.
-That is to say that orienting these flows requires that the translation surface itself be reflected to realign the flow.
-These translation surfaces are often described as being "unfolded" by some billiard flow, or strung along a laundry line of polygons congruent to the base.
-In the later sections we will describe how our surface's canonical form deforms geodesic flows by rotation and how we can use these rotations and edge-crossings to determine whether or not a geodesic will close.
+- Vincent, Ferran and Pascal if I can get a rough draft of their book
 
-What inspired this paper was the work done by Hubert et al.
-[Hu] on the Ehrenfest-Wind Tree model.
-Any readers who are familiar with their paper or the Tree model will most likely notice the similarities between the quotient spaces of the two surfaces and the properties that determine whether or not geodesic paths close.
-While similar, these surfaces are not entirely identical.
-The effect that a boundary collision has on a geodesic alters the dynamical properties of the surface.
-Or perhaps they better demonstrate properties of the surface itself.
-Another fundamental difference between the two is that the Necker Cube is a surface with cone angles surrounding an infinite number of singularities on the surface.
-The problem has to take into careful consideration what it means for a path to cross a saddle connection between these singularities of the translation surface.
+- Hubert, et. al. "Recurrence flow " paper on wind tree
 
-\subsection{Procedure}
-The Necker Cube surface has an interesting construction process that lets us take each face of the cube and arrange them in any way we choose so long as we do not change how the edges are identified, or glued together.
-It is preferable to study the flows on this surface as though every cube has been flattened onto the plane.
-This means that we can consider the cube step to be tiled by infinitely many of L-shaped surfaces whose edges are identified by rotation.
+- Ariel Mazur for assistance earlier on
 
-Doing so greatly reduces the number of possible directions that a flow might travel.
-Since every edge of this L-shaped surface is identified to another either by translation or a 90 degree rotation, any straight line path on this flattened version of the Necker Cube can experience only one of four possible directions at any given time, whereas the Necker Cube would've had considerably more.
+- Math dept at CCNY? or research internship? Not sure if apropos.
 
-Moreover, every face of this L-shaped surface is given by rotating its associated face on the Necker Cube about the x or y axis, affine translations, or a combination of the two.
-By cutting and gluing every half-cube of the Necker Cube in a consistent fashion, we can preserve the areas of the faces and distances between any two points on the surface with the flattening procedure.
-We describe this procedure as piecewise maps that are comprised of compositions of Euclidean matrices and translations of points relative to the L-shape onto a subset of $\mathbb{R}^{2}$ with square shaped holes cut out.
-Since rotations and translations of points preserves distance between them, the path lengths and  associated properties of geodesic flows remain invariant.
+\section{Constructing and Flattening $\mathbf{S}$}
 
-The image of this map is a flattened version of the Necker Cube.
-(See FIG [MATLAB MODEL]).
-From that surface we construct a new one made out of four copies of original, each rotated by some factor of 90 degrees so as to realign the geodesic flows and assure that they can only travel in one possible direction.
-This four-fold surface is then a cover of the original surface.
-We consider every possible direction to be a "phase" of the flow, and index the planes accordingly.
-Any edges in the holes of the surface are identified with an edge on an adjacent phase plane (By rotation of $\pm \frac{\pi}{4}$).
-Any point on this surface has a coordinate in $\mathbb{R}\times\mathbb{R}\times\mathbb{Z}_{4} $
-
-Similarly, we take the quotient space of the flattened Necker Cube centered around each square shaped hole and realign its flows by rotation.
-The resulting surface is a compact branch cover of the quotient space.
-Singularities on the translation surface have a total angle that is some multiple of $2\pi$.
-We consider the translation surface to be a base surface of the four-fold cover, and study its properties to prove our main theorem.
-In particular, we wish to determine what effect a member of the translation surface's Veech group has on both the surface and geodesic flow in hopes that it would shine some light on odd-odd even-odd phenomenon.
-
- $SL_{2}(\mathbb{R})$
-
-\section{Constructing the Surface \& Isometric Variant}
-\compav{Started up here again 6/21/17}
-
-This section will detail how the Necker cube (See Figure \ref{fig:cubexyz}) is constructed out of an infinite set of unit squares, and explain the process by which we flatten the structure in such a way that preserves distances, and makes it simpler for us to describe the behavior of the geodesic flows that come into contact with these edges.
+This section will detail how the Necker cube surfaces is constructed as the union of infinitely many Necker cubes.
+Each Necker cube is taken to be the union of unit squares, each belonging to one of three orthogonal auxiliary planes.
+An isometric variant of $\mathbf{S}$ can be described as a \emph{flattening} of $\mathbf{S}$ such that these unit squares all lie on the $x$-$y$ plane.
 
@@
 \begin{center}
-\includesvg[width=3in]{cubesxyz}
-\caption{The Necker cube embedded in space.}
+\includesvg[width=2in]{cubesxyz}
+\caption{The Necker cube surface embedded in space.}
 \label{fig:cubexyz}
@@
 
-To build the surface out of infinitely many unit squares, we take the union of the instances of three squares, each lying on planes orthogonal to the other two and sharing an edge with exactly one other square.
-The three squares have only one vertex in common.
-
-First we divide the planes $x,y,z=k$ for some integer $k$ in such a way that they are composed entirely out of unit squares.
-We want to index each of these squares by integer triples and some label A,B, or C.
-i.e.
-squares on planes of the form $z=k$ will have an A label, $x=k$ a B label, and $y=k$ a C label.
-Let $m,n,p\in\mathbb{Z}$.
-Then each square based at $(m,n,p)$ is defined as follows:
+The set $\mathbf{S}$ is built periodically out of infinitely many Necker cubes (a union of three unit squares).
+To each square fix a label, A,B, or C.
+For every integer triple, $(m,n,p)$, in $\mathbb R^3$, we have the following:
 
@@
 \end{tabular}
-\vspace{0.2in}\\
-\noindent For the sake of convenience, B and C are shifted in order to associate each integer triple that satisfies the linear condition $m+n+p=0$ with an instance of the cube that $\mathbf{S}$ is composed of.
-Furthermore, we  separate $\mathbf{S}$ into the sets:
+
+The Necker cube surface is obtained as the union of infinitely many Necker cubes that satisfy the condition that $m+n+p=0$. $\mathbf{S}$ can be separated into the following sets:
+
 \begin{align*}
@@
 
-\noindent We can now give a formal definition of the Necker Cube.
+\noindent We can now give a set definition of the Necker cube surface.
 
-\begin{Def} The \textbf{Necker Cube}, denoted by $\mathbf{S}$, is the surface made up of unions of unit squares of the form $\mathbf{S} = \mathbf{A}\cup\mathbf{B}\cup\mathbf{C}$.\end{Def}
-
-\noindent $\mathbf{S}$ is the surface portrayed in Figure \ref{fig:cubexyz}.
+\begin{Def} The \textbf{Necker cube surface}, denoted $\mathbf{S}$, is the surface made up of the union of unit squares of the form $\mathbf{S} = \mathbf{A}\cup\mathbf{B}\cup\mathbf{C}$.\end{Def}
 
 \subsection{The Flattened Form}
-We will show that $\mathbf S$ is isometric to a a topological quotient of a particular subset of the plane.
-We call this quotient $\mathbf{U}$, and the subset of the plane $\mathbf P$ (See Figure $\ref{fig:uprimeu}$).
+The Necker cube surface is path-isometric to a topological quotient of a particular subset of the plane.
+We call the flattened form $\mathbf{U}$, and the subset of the plane $\mathbf P$ (See figure $\ref{fig:uprimeu}$).
 The subset $\mathbf P$ is the region of the plane with unit squares removed:
@@
 
-\compat{Illustrating refering to equations: Recall $\mathbf{P}$ was defined in \eqref{eq:P}.}
-
 \begin{figure}[H]
@@
 \end{center}
-\caption{The subset $\mathbf P$ of the plane with unit squares removed at every even integer pair.} \compat{Perhaps just use one.}
+\caption{$\mathbf P$ is real plane with unit squares removed at every even integer pair.}
 \label{fig:uprimeu}
@@
 
-\noindent On $\mathbf P$ we define a relation $\sim$ that identifies the sides of the boundaries of each hole in pairs.
+\noindent We define a relation on $\mathbf P$ that identifies the sides of the boundaries of each hole in pairs.
 Each side is identified with another side of the same hole by rotation about the bottom left or top right corners of the hole by $\pm\frac{\pi}{2}$.
-The result is an edge identification as seen here (centered at each even coordinate pair): \compat{I think overall this is very good and completely defines the relation.
-You may want to tweak it to make it even more clear.
-You expressed some concerns about the vertices.}
+The result is the following edge identifications made on cut out square holes:
 
@@
 \end{figure}
-\noindent We call this relation $\mathbf{R}$.
-\compat{How about $\sim$?} What $\mathbf{R}$ essentially does is translates every hole to the origin and identifies the edges by reflecting over the line $y=x$.
-If these points coincide, then they are indeed the same point on the boundary.
-Otherwise, every point that is not on the boundary belongs to its own class.
-We can safely assume this wont identify edges improperly, since we translate every boundary so as to situate the unit holes over the origin.
-For example, if we took two edges that do not belong to the same hole then by that same translation and reflection the two edges couldn't possible coincide.
-\compat{This paragraph doesn't seem to add much.}
 
-\begin{Def}
-content...
-\end{Def}
-
-\begin{Def}{$\mathbf U$}\\
-$\mathbf U$ is equal to the quotient space $\mathbf{P}/\mathbf{R}\times\{0\}$
-\end{Def}
-
-\compav{Stopped here. 6/21/17}
-
-\compat{I would use as an introductory paragraph: ``We will show that $\mathbf S$ is isometric to a topological quotient of a subset of the plane.''}
-
-Before describing the transformation that takes $\mathbf{S}$ to its flattened form, it would be best to define the relations on a subset of the plane that preserves the identification of edges.
-We will call this surface $\mathbf{U}'$.
-However, the flattening of the surface takes $\mathbf{S}$ to a translation of $\mathbf{U}'$, we call $\mathbf{U}$ (See $\ref{fig:uprimeu}$).
-We will denote the metric preserving map from $\mathbf{S}$ to $\mathbf{U}$, by $\Psi$.
-Defining these edge identifying relations over $\mathbf{U}'$ is more convenient for us than defining them over $\mathbf{U}$, but defining $\Psi$ is more convenient when the map is taken to $\mathbf{U}$ due to the nature of the construction of $\mathbf{S}$.
-\compat{I found this paragraph awkward.
-I would just go ahead and define the region in the plane.}
-
-Now we will describe exactly how to construct $\mathbf{U}'$. Take the subset of the real plane, 
-$$,$$ 
-\compat{I would write $(m-\frac{1}{2},m+\frac{1}{2}) \times (n-\frac{1}{2},n+\frac{1}{2})$.}
-with open unit squares cut out at every even integer pair and let $\mathbf{U}'$ represent the flattened structure with the proper edge identifications with the first "hole" centered at the origin.
-We will describe this surface as a cross product of the auxiliary plane, $z=0$ and the image of the natural map from $\mathbf{P}$ to its quotient space with the following relation on $\mathbf{P}$:
-
+\begin{Def}{} $\mathbf U$ is the surface obtained as the topological quotient $\mathbf P/\mathbf R$, where $\mathbf R$ is defined as follows:
 \begin{gather*}
-\text{Let } x_{0}=(u_{0},v_{0}),x_{1}=(u_{1},v_{1}) \in \mathbf{P}. \text{Let } \mathbf{R} \text { be a relation given, } x_{0}\mathbf{R}x_{1} \\ \text{ iff } x_{0}=x_{1}
- \text{ or if }x_{0},x_{1} \in {\partial} \left( \left[2m-\frac{1}{2},2m+\frac{1}{2}\right] \times \left[2n-\frac{1}{2},2n+\frac{1}{2}\right] \right)\\
+\text{Let } x_{0}=(u_{0},v_{0}),x_{1}=(u_{1},v_{1}) \in \mathbf{P}.  \mathbf{ R} \text { is given as the relation } x_{0}\mathbf{R}x_{1} \\ \text{ iff } x_{0}=x_{1}
+ \text{ or, for }x_{0},x_{1} \in {\partial} \left( \left[2m-\frac{1}{2},2m+\frac{1}{2}\right] \times \left[2n-\frac{1}{2},2n+\frac{1}{2}\right] \right)\\
+ \text{ and }m,n\in\mathbb{Z},
   \left[\begin{array}{c}
@@
 \end{array}\right]
-\\\text{For some }m,n\in\mathbb{Z}. \end{gather*}
+. \end{gather*}
+\end{Def}
 
-The relation will either identify every point to itself or, in the case where any two points lie on the boundary of the square shaped holes, take the center of each square hole in the plane to the origin and reflect it over the line $y=x$ to match up these points accordingly.
-Let $N: \mathbf{P} \rightarrow \mathbf{P}/\mathbf{R}$ denote the natural map from every point on $\mathbf{P}$ to its appropriate equivalence class.
-Now, let $\mathbf{U}' = Im N \text{ }\times\{0\}$ be the flat representation of the surface that has been translated so as to center the first square cut-out region over the origin.
-The map $\mathbf{U}\rightarrow\mathbf{U}'$ is given by an affine translation of the surface $(x,y,z) \mapsto (x-\frac{3}{2},y-\frac{3}{2},z)$.
-(The point at the origin of $\mathbf{S}$ is mapped to $(-\frac{3}{2},-\frac{3}{2},0) \text{ on } \mathbf{U}'$).
+We can see that $\mathbf{U}$ is a Euclidean cone surface with infinitely many singularities of cone angles $\frac{3\pi}{2}$ and $6\pi$.
+Observe that $\mathbf{S}$ is also a cone surface with as many singularities of the same cone angles.
 
-\compav{Stopped here. Continued in Sec. 3.}
+\subsection{A Path-Isometric Map From $\mathbf S$ to $\mathbf U$}
+A path-isometric map from $\mathbf{S}$ to $\mathbf{P}$ induces an isometry from the Necker cube surface to $\mathbf{U}$.
+The path isometry $\Psi:\mathbf{S}\rightarrow\mathbf{P}$ is a piece-wise, well-defined map that is given as a linear transformation in which a square labeled $\mathbf{A}$, $\mathbf{B}$, or $\mathbf{C}$ is mapped to its corresponding square on $\mathbf{P}$.
+$\Psi$ is  essentially a transformation that cuts along the edges parallel with the z-axis.
+(See figure $\ref{fig:cut}$)
 
-\newpage
-\subsection{Flattening $\mathbf{S}$}
-This section will describe the isomorphic map $\Psi:\mathbf{C}\rightarrow\mathbf{U}$, its inverse, $\Psi^{-1}:\mathbf{U}\rightarrow\mathbf{C}$, and prove that $\mathbf{C}$ and $\mathbf{U}$ are isometric.
+\begin{figure}[H]
+\includesvg[width=2.1in]{cubesxyzcut}
+\raisebox{0.4in}{\includesvg[width=2.3in]{unfoldcut}}
+\caption{The images of $\mathbf S$ and $\Psi(\mathbf{S})$ with edges identified.}
+\label{fig:cut}
+\end{figure}
 
-The flattening procedure can be represented as a series of cuts made along edges parallel to the z-axis. The gluing structure is preserved by the relations defined on $\textbf{U}$ and $\textbf{U}'$.
-
-\begin{center}
-\includesvg[width=2.3in]{cubesxyzcut}
-\raisebox{0.4in}{\includesvg[width=2.3in]{u}}
-\text{The images of } $\Psi^{-1}(\mathbf{U})$ (left)\text{ and }$\Psi(\mathbf{S})$ (right)
-\end{center}
-
-$\Psi$ is defined piecewise, and denote each subset of $\mathbf{C}$ with the letters $A$, $B$, $C$. We do the same for $\mathbf{U}$ but use instead $A'$, $B'$, $C'$.
-
-\begin{gather*}
-	\text{Let }x,y,z\in\mathbb{R}. \text{ Then}
-	\\A=\bigcup\big\{\mathbf{C}\cap\{(x,y,p)\in\mathbb{R}\}: p\in\mathbb{Z}\big\}
-	\\B=\bigcup\big\{\mathbf{C}\cap\{(x,n,z)\in\mathbb{R}\}: n\in\mathbb{Z}\big\}
-	\\C=\bigcup\big\{\mathbf{C}\cap\{(m,y,z)\in\mathbb{R}\}: m\in\mathbb{Z}\big\}
-	\\\text{are the collection of the faces of cubes that lie on planes that are parallel to one another.}
-	\\\text{As shown here:}
-\end{gather*}
-
-To describe the subsets $A'$, $B'$, and $C'$ properly we will introduce the function $\mu_{n}:\mathbb{R}\rightarrow \left[0,n\right) $ as shorthand for the family of functions that take a real number to its residue class of reals modulo n, for some integer n.
-
-\begin{gather*}
-	\text{Let }u,v\in\mathbb{R}. \text{Then }
-	\\A'= \{(u,v,0): 0<\mu_{2}(u)\leq 1 ,\hspace{4mm} 0<\mu_{2}(v)\leq 1\}
-	\\B'= \{(u,v,0): 1<\mu_{2}(u)\leq 2,\hspace{4mm} 0<\mu_{2}(v)\leq 1\}
-	\\C'= \{(u,v,0): 0<\mu_{2}(u)\leq 1,\hspace{4mm} 1<\mu_{2}(v)\leq 2\}
-	\\\text{are the collections of the associated faces of the flattened structure.}
-\end{gather*}
-
-Now we will define $\Psi$ piecewise.
+\begin{Def}
+The bijective function $\Psi$ maps all of $\mathbf{S}$ to $\mathbf P$, and is given as:
+\end{Def}
 
@@
 	\left[\begin{array}{c}
-		2 \left\lfloor x \right\rfloor
-		\\ 2\left\lfloor y \right\rfloor
+		2 \left\lfloor x \right\rfloor - 1
+		\\ 2\left\lfloor y \right\rfloor - 1
 		\\ z - \left\lfloor z \right\rfloor
@@
 		\left[\begin{array}{c}
-			2 \left\lfloor x \right\rfloor
-			\\ 2\left\lfloor y \right\rfloor
+			2 \left\lfloor x \right\rfloor - 1
+			\\ 2\left\lfloor y \right\rfloor - 1
 			\\ x - \left\lfloor x \right\rfloor
@@
 			\left[\begin{array}{c}
-				2 \left\lfloor x \right\rfloor
-				\\ 2\left\lfloor y \right\rfloor
+				2 \left\lfloor x \right\rfloor - 1
+				\\ 2\left\lfloor y \right\rfloor - 1
 				\\ y - \left\lfloor y \right\rfloor
@@
 
-We chose to preserve the matrix structure in order to emphasize the manner in which every square is taken to $\mathbf{U}$ from $\mathbf{C}$ and vice-versa.
+Clearly, $\Psi$ preserves distances on the faces of each as it consists entirely of rotations and translations.
+As such, it is path-isometric and induces an isometry between the Necker Cube surface and $\mathbf U$.
+
+\begin{thm}{$\mathbf{S}$ is isometric to  $\mathbf{U}$.}
+\begin{proof}
+
+Let $(\mathbf{S}, d)$ and $(\mathbf{U}, d')$ be metric spaces with their respective metrics.
+Let $a,b\in\mathbf{S}$, and suppose that $\Psi(a)=a',\Psi(b)=b'\in\mathbf{U}$.
+We will denote the distance between $a$ and $b$ by $D=d(a,b)$, and the distance between $a'$ and $b'$ by $D'=d'(a',b')$.
+Since $\Psi$ and $\Psi^{-1}$ are linear transformations that act on subsets of auxiliary planes in $\mathbb{R}^{3}$, the images of any curves on $\mathbf{S}$ or $\mathbf{U}$ under these maps will not be deformed and retain their lengths.
+Thus $D \leq d'(\Psi(a), \Psi(b)) = d'(a',b') = D'$.
+Likewise $D' \leq d(\Psi^{-1}(a'),\Psi^{-1}(b')) = d(a,b) = D$.
+Therefore $D=D'$, and $\Psi$ and $\Psi^{-1}$ preserves distances between points.
+And so, $\mathbf{S}\simeq\mathbf{U}$.
+\end{proof} 
+\end{thm}
+
+\newpage
+%\subsection{Flattening $\mathbf{S}$}
+%This section will describe the isomorphic map $\Psi:\mathbf{C}\rightarrow\mathbf{U}$, its inverse, $\Psi^{-1}:\mathbf{U}\rightarrow\mathbf{C}$, and prove that $\mathbf{C}$ and $\mathbf{U}$ are isometric.
+%
+%The flattening procedure can be represented as a series of cuts made along edges parallel to the z-axis. The gluing structure is preserved by the relations defined on $\textbf{U}$ and $\textbf{U}'$. 
+%
+%
+%$\Psi$ is defined piecewise, and denote each subset of $\mathbf{C}$ with the letters $A$, $B$, $C$. We do the same for $\mathbf{U}$ but use instead $A'$, $B'$, $C'$. 
+%
+%\begin{gather*}
+%	\text{Let }x,y,z\in\mathbb{R}. \text{ Then}
+%	\\A=\bigcup\big\{\mathbf{C}\cap\{(x,y,p)\in\mathbb{R}\}: p\in\mathbb{Z}\big\}
+%	\\B=\bigcup\big\{\mathbf{C}\cap\{(x,n,z)\in\mathbb{R}\}: n\in\mathbb{Z}\big\}
+%	\\C=\bigcup\big\{\mathbf{C}\cap\{(m,y,z)\in\mathbb{R}\}: m\in\mathbb{Z}\big\}
+%	\\\text{are the collection of the faces of cubes that lie on planes that are parallel to one another.}
+%	\\\text{As shown here:}
+%\end{gather*}
+%
+%
+%
+%To describe the subsets $A'$, $B'$, and $C'$ properly we will introduce the function $\mu_{n}:\mathbb{R}\rightarrow \left[0,n\right) $ as shorthand for the family of functions that take a real number to its residue class of reals modulo n, for some integer n.
+%
+%\begin{gather*}
+%	\text{Let }u,v\in\mathbb{R}. \text{Then }
+%	\\A'= \{(u,v,0): 0<\mu_{2}(u)\leq 1 ,\hspace{4mm} 0<\mu_{2}(v)\leq 1\}
+%	\\B'= \{(u,v,0): 1<\mu_{2}(u)\leq 2,\hspace{4mm} 0<\mu_{2}(v)\leq 1\}
+%	\\C'= \{(u,v,0): 0<\mu_{2}(u)\leq 1,\hspace{4mm} 1<\mu_{2}(v)\leq 2\}
+%	\\\text{are the collections of the associated faces of the flattened structure.}
+%\end{gather*}
+%
+%
+%
+%Now we will define $\Psi$ piecewise.
+%
+%
+%
+%We chose to preserve the matrix structure in order to emphasize the manner in which every square is taken to $\mathbf{U}$ from $\mathbf{C}$ and vice-versa.
 The negations of some of these variables are a result of this structure being built so as the integer triples lie on the plane $z=-(x+y)$.
@@
 In this case, $\chi_{m,n,p}$ is a member of this family whose dimensions are $1\times1\times1$).
-
-Now we will define the inverse map, $\Psi^{-1}$, in a similar manner.
-
-\begin{equation}
-	\Psi^{-1}\left[\begin{array}{c}
-		x\\y\\z
-	\end{array}\right] 
-	= 
-	\begin{cases}
-		\left[ \hspace{2mm} \begin{matrix}
-			1 & 0 & 0 \\
-			0 & 1 & 0 \\
-			0 & 0 & 1
-		\end{matrix}\hspace{3mm}\right]
-
-		\left[\begin{array}{c}
-		x - 2\left\lfloor \frac{x}{2} \right\rfloor
-		\\ y- 2\left\lfloor \frac{y}{2} \right\rfloor
-		\\ z
-		\end{array} \right]
-		+
-		\left[\begin{array}{c}
-			\left\lfloor \frac{x}{2} \right\rfloor
-			\\ \left\lfloor \frac{y}{2} \right\rfloor
-			\\ -(\left\lfloor \frac{x}{2} \right\rfloor + \left\lfloor \frac{y}{2} \right\rfloor )
-		\end{array} \right]
-			& \text{if } (x,y,z)\in A'	\vspace{2mm}
-		\\
-		\left[ \hspace{2mm} \begin{matrix}
-			0 & 0 & -1 \\
-			0 & 1 & 0 \\
-			1 & 0 & 0
-		\end{matrix}\hspace{0mm}\right]
-
-		\left[\begin{array}{c}
-		-(x - 2\left\lfloor \frac{x}{2} \right\rfloor)
-		\\ y- 2\left\lfloor \frac{y}{2} \right\rfloor
-		\\ z
-		\end{array} \right]
-		+
-		\left[\begin{array}{c}
-			1+\left\lfloor \frac{x}{2} \right\rfloor
-			\\ \left\lfloor \frac{y}{2} \right\rfloor
-			\\ -(\left\lfloor \frac{x}{2} \right\rfloor + \left\lfloor \frac{y}{2} \right\rfloor -1)
-		\end{array} \right]
-			& \text{if } (x,y,z)\in B'	\vspace{2mm}
-
-	\\
-
-		\left[ \hspace{2mm} \begin{matrix}
-			1 & 0 & 0 \\
-			0 & 0 & -1 \\
-			0 & 1 & 0
-		\end{matrix}\hspace{0mm}\right]
-
-		\left[\begin{array}{c}
-		x - 2\left\lfloor \frac{x}{2} \right\rfloor
-		\\ -(y- 2\left\lfloor \frac{y}{2} \right\rfloor)
-		\\ z
-		\end{array} \right]
-		+
-		\left[\begin{array}{c}
-			\left\lfloor \frac{x}{2} \right\rfloor
-			\\1+ \left\lfloor \frac{y}{2} \right\rfloor
-			\\ -(\left\lfloor \frac{x}{2} \right\rfloor + \left\lfloor \frac{y}{2} \right\rfloor -1)
-		\end{array} \right]
-			& \text{if } (x,y,z)\in B'	\vspace{2mm}
-
-	\end{cases}
-\end{equation}
-
-By using properties of the floor function, one can show that $\Psi$ is indeed a bijective, invertible map. All that is left to show is that $\Psi$ preserves distance.
-
-\begin{thm}{$\mathbf{C}$ and $\mathbf{U}$ are isometric.}\\
-Proof.
-Let $(\mathbf{C}, d)$ and $(\mathbf{U}, d')$ be topological spaces with their respective metrics.
-Let $a,b\in\mathbf{C}$, and let $\Psi(a)=a',\Psi(b)=b'\in\mathbf{U}$.
-We will denote the distance between $a$ and $b$ by $D=d(a,b)$, and the distance between $a'$ and $b'$ by $D'=d'(a',b')$.
-Since $\Psi$ and $\Psi^{-1}$ are linear transformations that act on subsets of auxiliary planes in $\mathbb{R}^{3}$, the images of any curves on $\mathbf{C}$ or $\mathbf{U}$ under these maps will not be deformed and retain their lengths.
-Thus, $D \leq d'(\Psi(a), \Psi(b)) = d'(a',b') = D'$.
-Likewise, $D' \leq d(\Psi^{-1}(a'),\Psi^{-1}(b')) = d(a,b) = D$.
-Therefore $D=D'$, and $\Psi$ and $\Psi^{-1}$ are metric preserving maps between $\mathbf{C}$ and $\mathbf{U}$.
-$\qed$
-\end{thm}
-
-As a consequence of this theorem, any dynamical properties of geodesic flow on the surface $\mathbf{U}'$ immediately applies to $\mathbf{C}$. ($\mathbf{U}'$ and $\mathbf{U}$ are trivially isometric).
-
-\subsection{Comparing Geodesic Behavior on Both Surfaces}
-Two immediate observations that can be made in regards to geodesic flows on the surface was that any trajectory with slope one closes, and any flows with horizontal or vertical trajectories drift.
-Observe the path behavior of three different flows on the surface:
-
-The figure above shows us some very simple paths on the surfaces.
-We use the three vectors, (0,1), (1,0), and (1,1), as a sort of basis for proving our theorem.
-Included in the appendix are a few more simulated flows of various trajectories.
-We've color coded the trajectories in the image above for path c to represent one of four phases
-
-\newpage
+%
+%Now we will define the inverse map, $\Psi^{-1}$, in a similar manner. 
+%
+%\begin{equation}
+%	\Psi^{-1}\left[\begin{array}{c}
+%		x\\y\\z
+%	\end{array}\right] 
+%	= 
+%	\begin{cases}
+%		\left[ \hspace{2mm} \begin{matrix}
+%			1 & 0 & 0 \\
+%			0 & 1 & 0 \\
+%			0 & 0 & 1
+%		\end{matrix}\hspace{3mm}\right]
+%	
+%		\left[\begin{array}{c}
+%		x - 2\left\lfloor \frac{x}{2} \right\rfloor
+%		\\ y- 2\left\lfloor \frac{y}{2} \right\rfloor
+%		\\ z
+%		\end{array} \right]
+%		+
+%		\left[\begin{array}{c}
+%			\left\lfloor \frac{x}{2} \right\rfloor
+%			\\ \left\lfloor \frac{y}{2} \right\rfloor
+%			\\ -(\left\lfloor \frac{x}{2} \right\rfloor + \left\lfloor \frac{y}{2} \right\rfloor )
+%		\end{array} \right]
+%			& \text{if } (x,y,z)\in A'	\vspace{2mm}
+%		\\
+%		\left[ \hspace{2mm} \begin{matrix}
+%			0 & 0 & -1 \\
+%			0 & 1 & 0 \\
+%			1 & 0 & 0
+%		\end{matrix}\hspace{0mm}\right]
+%	
+%		\left[\begin{array}{c}
+%		-(x - 2\left\lfloor \frac{x}{2} \right\rfloor)
+%		\\ y- 2\left\lfloor \frac{y}{2} \right\rfloor
+%		\\ z
+%		\end{array} \right]
+%		+
+%		\left[\begin{array}{c}
+%			1+\left\lfloor \frac{x}{2} \right\rfloor
+%			\\ \left\lfloor \frac{y}{2} \right\rfloor
+%			\\ -(\left\lfloor \frac{x}{2} \right\rfloor + \left\lfloor \frac{y}{2} \right\rfloor -1)
+%		\end{array} \right]
+%			& \text{if } (x,y,z)\in B'	\vspace{2mm}
+%
+%	\\
+%	
+%		\left[ \hspace{2mm} \begin{matrix}
+%			1 & 0 & 0 \\
+%			0 & 0 & -1 \\
+%			0 & 1 & 0
+%		\end{matrix}\hspace{0mm}\right]
+%	
+%		\left[\begin{array}{c}
+%		x - 2\left\lfloor \frac{x}{2} \right\rfloor
+%		\\ -(y- 2\left\lfloor \frac{y}{2} \right\rfloor)
+%		\\ z
+%		\end{array} \right]
+%		+
+%		\left[\begin{array}{c}
+%			\left\lfloor \frac{x}{2} \right\rfloor
+%			\\1+ \left\lfloor \frac{y}{2} \right\rfloor
+%			\\ -(\left\lfloor \frac{x}{2} \right\rfloor + \left\lfloor \frac{y}{2} \right\rfloor -1)
+%		\end{array} \right]
+%			& \text{if } (x,y,z)\in B'	\vspace{2mm}
+%			
+%	\end{cases}
+%\end{equation}
 \section{The Four-Fold Cover and Translation Surface}
-\compav{Started here again.}
-The following section is concerned with defining a quotient of $\mathbf{U}'$ (see Figure $\ref{fig:quotient}$), and using that space to construct a translation surface.
-In addition we will describe a cover made from four copies of $\mathbf{U}'$ that aligns geodesic flow and behaves as a cover of our translation surface.
 \compat{Revisit to make sure it summarizes the section}
@@
 
-\begin{thm}$\mathbf{R}_{0}$ induces an isometry on $\tilde{\mathbf{U}}_0$
-\begin{proof}
-\compav{might rethink including $R_0$}.
-\end{proof}
-\end{thm}
-
 \subsection{Four-fold Translation Cover}
@@
 \includesvg[width=3.65in]{utildaprime}
-\caption{$\tilde{\mathbf{U}}$ with edges identified, and holes indexed relative to $\mathbf{P}\times\{0\}$.}
+\caption{$\tilde{\mathbf{U}}$ with edges identified, and holes indexed relative to $\mathbf{P}\times\{0\mod{4}\}$.}
 \label{fig:utilda}
@@
 We can see that $\mathbf{T}^{m,n}$ is indeed an isometry on $\tilde{\mathbf{U}}$ as it is composed of the induced isometries, $\bar{\phi}$ and $\bar{\mathbf{T}}^{m,n}$ .
-We can take the quotient of the translation surface $\tilde{\mathbf{U}}$ over the action of $\mathbf{T}^{m,n}$ to obtain a base surface that realizes $\tilde{\mathbf{U}}$ as an infinite $\mathbb{Z}^2$ cover.
+We can take the quotient of the translation surface $\tilde{\mathbf{U}}$ over the action of $\mathbf{T}^{m,n}$ to obtain a base surface that realizes $\tilde{\mathbf{U}}$ as an infinite $\mathbb{Z}^2$ cover, where $m,n$ is identified with an integer pair.
 
@@
 \subsection{$\mathbf M$ as a Veech Surface}
-By a series of cutting and gluing operations, $\mathbf M$ has this symmetric representation as a square-tiled staircase (Figure $\ref{fig:staircase}$).
+By a series of cutting and gluing operations, $\mathbf M$ has this symmetric representation as a square-tiled staircase (figure $\ref{fig:staircase}$).
 The vertices of cone angle $2\pi$ can be found on the centers of the squares in the figure below where the black lines intersect.
@@
 As a result of Gutkin-Judge [citation], $\mathbf M$ is a Veech surface whose Veech group is commensurable to $SL(2,\mathbb Z)$.
-In order to determine which elements of orientation preserving automorphisms, denoted Aut$^{+}(\mathbf M)$, have derivatives that belong to the Veech group of $\mathbf M$, SL($\mathbf M$), it is necessary to study $\mathbf M$'s various symmetries and cylinder decompositions.
+In order to determine which elements of orientation preserving affine diffeomorphisms, denoted Aff$^{+}(\mathbf M)$, have derivatives that belong to the Veech group, V($\mathbf M$), it is necessary to study $\mathbf M$'s various symmetries and cylinder decompositions.
+\vspace{0.1in}
 
-The derivatives of these automorphisms form the affine group, Aff$^+(\mathbf M)$.
-One such group contained in Aff$^+(\mathbf M)$ is SO$(2,\mathbb{Z})$, the group generated by rotations of order 4.
+\noindent\textbf{Rotational Symmetries}
+
+One group contained in $V(\mathbf M)$ is SO$(2,\mathbb{Z})$, the group generated by rotations of order 4.
 Even though the other elements in $\mathbb{D}_4$ are derivatives of automorphisms of $\mathbf M$, we will only consider those of determinant +1.
+We want the corresponding affine map of these rotation matrices to obey the convention that we \emph{fix} the second square on the bottom with label I.
+These maps are unique only up to translation and as such assigning this convention makes it easier to keep track of where the edges of $\mathbf M$ end up afterwards.
+The diffeomorphism with 90 degree counter-clockwise rotation fixed about that point at the center of that square is denoted $\mathbf{R}$.
+Its derivative, $\mathbf{R}'$, is the matrix $\left[ \hspace{1mm} \begin{matrix}
+							0 & -1\\
+							1 & 0
+						\end{matrix}\hspace{1mm}\right]$.
+
+\vspace{0.1in}
+
+\noindent\textbf{Translation Symmetries}
+
+The translational symmetries have no unique representation as an element of V($\mathbf M$), as their derivatives are represented as the identity.
+But they do have an important effect on the surface.
+Shifting the bottom left-most square to the right and reconstructing the rest by pairing edges is an order 2 homeomorphism denoted $\mathbf{H}$, for horizontal.
+Likewise, $\mathbf{V}$ is an order 2 homeomorphism that shifts that square upwards.
+Observe that together they satisfy the property $(\mathbf{H}\mathbf{V})^{6}=(\mathbf{V}\mathbf{H})^{6}=id$.
+This essentially moving the squares up the staircase until they come back to where they started.
+The group of translation deck transformations of the surface is isomorphic to a permutation group of 12 elements (each square) generated by these two translations that satisfy the relation.
+It is the group $\mathbb{D}_{12}$ generated by reflections about two adjacent vertices of a 12-gon.
+Thus it is possible to translate a square to any other on the surface as a composition of these two translations.
+
+\vspace{0.1in}
+
+\noindent\textbf{Skew Symmetries}
 
@@
 \includegraphics[width=4in]{cylinderdecomp.pdf}
-\caption{Horizontal and Verical Cylinders overlayed.}
+\caption{Horizontal and vertical cylinder decomposition as they overlay and cover the surface.}
 \label{fig:decomp}
@@
 Every cylinder is composed of two unit squares with edges identified by translation.
-We start counting the cylinders from the bottom-most cylinder and increment the index as the cylinders fill the surface going "up" the staircase.
+We start counting the cylinders from the bottom-most cylinder and increment the index as the cylinders fill the surface going ``up" the staircase.
 $\small{\mathcal{E}}_0$ is the first cylinder at the bottom whose identifications are \em{not} \normalfont the solid triangles.
@@
 We define the \emph{inverse modulus} of a cylinder to be the ratio of a cylinders circumference to its width, $\mu^{-1}=\frac{c}{h}$.
-A \emph{Dehn-twist} of a cylinder is an automorphism of the cylinder of the form  $(x,y)\mapsto(x,y+\mu^{-1} x \mod{h})$, where $x$ and $y$ are points on the cylinder relative to its circumference and height respectively.
+A \emph{Dehn-twist} of a cylinder is an automorphism of the cylinder of the form  $(x,y)\mapsto(x,y+\mu^{-1} x \mod{h})$, where $x$ and $y$ are points on the cylinder such that x varies along the its circumference.
 According to [need to cite something], a translation surface whose finite cylinder decompositions in orthogonal directions form a set of inverse moduli.
-The lowest (rational) common multiple of the set of all these values match up to give global affine diffeomorphisms of the surface, whose matrix representations are
+The smallest (rational) common multiple of the set of all these values match up to give a global affine diffeomorphisms of the surface, whose derivatives have the matrix representations
 \begin{equation*}
@@
 \end{equation*}
-where $\lambda$ is the lowest common multiple.
+such that $\lambda$ is that lowest common multiple.
 
-Fortunately, the cylinders $\Delta_{i}$ and $\small{\mathcal{E}}_j$, as i and j vary, all have the same inverse moduli.
+Let $\mu^{-1}_{\Delta_i},\mu^{-1}_{\mathcal{E}_j}$ denote the inverse moduli of the cylinders.
+The cylinders $\Delta_{i}$ and $\small{\mathcal{E}}_j$, as i and j vary, all have the same inverse moduli since they are all defined as two unit squares with identifications made by translation along the cylinders circumference.
 Therefore, $\lambda=$ l.c.m.($\mu^{-1}_{\Delta_i}\cup\mu^{-1}_{\mathcal{E}_j}$) $=2$.
@@
 						\end{matrix}\hspace{1mm}\right] \right>
-						\subset \text{SL}(\mathbf M).
+						\subset \text{V}(\mathbf M).
 \end{equation}
-This group is also known as the \emph{Sanov Subgroup}, a free group of rank 2.
-We take the group generated by ninety degree rotations and these two elements to be the group $\mathbb{X}\subset \text{SL}(\mathbf M)$.
-\newpage
-\subsection{$\mathbb X$'s Action on $\mathbf M$'s Homology}
-The group $\mathbb X$ consists of the set of all derivatives of Aut$^+(\mathbf M)$, whose action on the absolute homology of $\mathbf{M}$ extends to an action on the homology of $\mathbf{M}$'s $\mathbb{Z}^2$-cover, $\tilde{\mathbf{U}}$.
-$\mathbf{M}$'s cylinder decomposition gives way to a set of homology classes that serve as a basis for $H_1(\mathbf{M},\mathbb Z)$.
-For a total of 12 cylinders, there are 12 curves:
+This group is also known as the \emph{Sanov Subgroup} [cite], a free group of rank 2 that is commensurable to the Veech group of $\mathbf M$.
+We denote the \emph{horizontal} Dehn-twist of the surface with derivative $\left[ \hspace{1mm} \begin{matrix}
+				1 &  2\\
+				0 & 1
+			\end{matrix}\hspace{1mm}\right]$ by $\textbf{A}$. The \emph{vertical} Dehn-twist is denoted $\mathbf{B}$.
+
+\vspace{0.1in}
+
+\noindent\textbf{The Orientation-Preserving Symmetry Group of $\textbf{M}$}
+
+It is now possible to generate the \emph{group of symmetries} of $\textbf{M}$ as Deck transformations of the surface. We call this group $\mathbb X$ and say that
+
+\begin{align}
+\mathbb X= \left< \mathbf{A},\mathbf{B},\mathbf{R},\mathbf{H},\mathbf{V}\vert\mathbf{R}^4 = (\mathbf{H}\mathbf{V})^{6}=(\mathbf{V}\mathbf{H})^{6}=id  \right>.
+\end{align}
+
+The group of derivatives of $\mathbb X$ is $\mathbb X'$, and it is generated by elements of V$(\textbf{M})$.
+It is not apparent whether or not $\mathbb{X}=$Aff$^+(\mathbf{M})$ or if $\mathbb X'=$ V$(\mathbf{M})$.
+
+\subsection{Action of $\mathbb X$ on the Homology of $\mathbf M$}
+The group $\mathbb X$ consists of the set of Deck transformations of $\mathbf{M}$, whose induced isomorphisms of $\pi_1(\mathbf{M})$ acts on the absolute homology of $\mathbf{M}$ and extends to an action on the homology of $\mathbf{M}$'s $\mathbb{Z}^2$-cover, $\tilde{\mathbf{U}}$.
+$\mathbf{M}$'s cylinder decomposition gives us a natural set of homology classes that serve as a basis for $H_1(\mathbf{M},\mathbb Q)$.
+For a total of 12 cylinders, there are 12 core curves:
 
@@
 \centering
+\caption{Cylinder core curves with u,v, and z homology classes that determines the $\mathbb{Z}^2$-cover.}
 \end{figure}
-Every curve intersects two others.
-By constructing a $12\times12$ matrix where the rows represent each curve and the columns represent their intersections with other curves, we can see that the rank of said matrix is 10.
-This is to be expected as the genus of $\mathbf M$ is 5.
-From these curves we construct homology classes $u,v,$ and $z$.
-$u$ and $v$ are the curves that were the outermost edges of the tori in $\mathbf{M}$'s original representation (Figure $\ref{fig:mtilda}$).
-We can see from Figure $\ref{fig:utilda}$ that crossing over one of these edges corresponds to a shift of the even integer index pair that determines exactly which of the quotients of $\tilde{\mathbf{U}}$ a terminal point of a lifted path on the surface belongs to.
-That point is identified with the triple $(m,n,\mathbb Z/4\mathbb{Z})$, with integers $m,n$.
-The $z$ class of cylinder curves lie on the cylinders that take a path from one plane to the next in our cover.
-Since one plane is related to exactly two others, it is impossible to go directly from the first to the third or the second to the fourth, and vice versa.
+
+\begin{rem}
+Although the genus of $\mathbf{M}$ is 5, we use 12 curves as an overdetermined basis for $H_1(\mathbf{M},\mathbb Q)$.
+To show that they indeed span all of $H_1(\mathbf{M},\mathbb Q)$, it is possible to construct a $12\times12$ matrix of curve intersections and calculate its rank.
+The rank of the matrix, as to be expected, is 10.
+But to use only the linearly independent curves means that many of the same symmetries described in the previous section would not induce isomorphisms of $\pi_1(\mathbf{M})$ that dualize $\mathbf{M}$'s symmetries as the union of 12 identical squares.
+\end{rem}
+
+We will denote the group of isomorphisms of $\mathbf{M}$'s fundamental group we obtain from $\mathbb{X}$ as $\mathbb X_*$.
+The same notation will be used for the isomorphisms themselves.
+We denote the \emph{set of cylinder core curves} as $\Gamma$.
+
+\vspace{0.1in}
+
+\noindent\textbf{A 12-gon dualization of the $H_1(\mathbf{M},\mathbb Q)$}
+
+The 12-gon was mentioned in the previous section under the discussion of $\mathbf{M}$'s translational symmetries. $\mathbb X_*$ acting on absolute homology is better represented as a symmetric object.
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=1.7in]{12gon.pdf}
+\end{figure}
+
+This graph's representation also reflects how two core curves are incident with only two others.
+An element of $\mathbb X_*$ either permutes these curves or, in the case where $\mathbf A_*$ or $\mathbf B_*$ generates the element, adds/subtracts black vertices to/from the red ones.
+Recall that in the previous section we had certain conventions when it came to the elements $\mathbf{R}$, $\mathbf{H}$, and $\mathbf{V}$.
+This was so that their induced isomorphisms would \emph{fix} one or more vertices as they act on homology.
+But because of the translational symmetries, it is possible to rotate about the center of any square and essentially rotating the entire 12-gon with $\mathbf{H}_*$ and $\mathbf{V}_*$.
+
+\begin{minipage}{0.5\textwidth}
+\vspace{0.3in}
+\begin{itemize}
+\item[\textbf{\emph{$\mathbf{H}_*$\& $\mathbf{V}_*$}}] The effect that these two have on the 12-gon is a reflection about these lines.
+Observed by keeping track of the squares and core curves after $\mathbf{H}$ and $\mathbf{V}$ have acted on $\mathbf{M}$.
+\end{itemize}
+\end{minipage}
+\begin{minipage}{0.7\textwidth}
+\begin{figure}[H]
+\hspace{0.2in}\includegraphics[width=1.7in]{12gonHV.pdf}
+\end{figure}
+\end{minipage}
+
+The two preserve the signs of the core curves. It is to be expected as they are induced isomorphisms of translation maps. They are given as the following permutations of $\Gamma$:
+\space
+\begin{align*}
+\mathbf{H}_* &= (\gamma_0)(\gamma_1\hspace{1mm}\gamma_{11})(\gamma_2\hspace{1mm}\gamma_{10})(\gamma_3\hspace{1mm}\gamma_9)(\gamma_4\hspace{1mm}\gamma_8)(\gamma_5\hspace{1mm}\gamma_7)(\gamma_6)\\
+\mathbf{V}_* &=(\gamma_{11})(\gamma_0\hspace{1mm}\gamma_{10})(\gamma_1\hspace{1mm}\gamma_{9})(\gamma_2\hspace{1mm}\gamma_8)(\gamma_3\hspace{1mm}\gamma_7)(\gamma_4\hspace{1mm}\gamma_6)(\gamma_5)
+\end{align*}
+
+Rotation about the center of a square will not only change signs of the core curves, but permute elements in such a way that 4 consecutive rotations of $\mathbf{M}$ is an identity on $\Gamma$.
+
+\begin{minipage}{0.4\textwidth}
+\begin{figure}[H]
+\hspace{0}\includegraphics[width=1.7in]{12gonR.pdf}
+\end{figure}
+\end{minipage}
+\begin{minipage}{0.55\textwidth}
+\vspace{0in}
+\begin{itemize}
+\item[\textbf{\emph{$\mathbf{R}_*$}}] This action reflects about the line and maps one vertex to another.
+No vertex is fixed.
+However it has the additional effect of changing the sign of a curve only if it was originally black (or odd index).
+The pattern is peculiar, but $\mathbf{R}_*$ is still order four.
+\end{itemize}
+\end{minipage}
+
+\vspace{2mm}
+We can see the effect that it has on homology clearer as a permutation of $\Gamma$:
+\begin{align*}
+\mathbf{R}_*= \hspace{1mm} &(\gamma_{0}\hspace{1mm}\gamma_{1}\hspace{1mm}\text{-}\gamma_{0}\hspace{1mm}\text{-}\gamma_{1})
+(\gamma_{2}\hspace{1mm}\gamma_{11}\hspace{1mm}\text{-}\gamma_{2}\hspace{1mm}\text{-}\gamma_{11})
+(\gamma_{3}\hspace{1mm}\text{-}\gamma_{10}\hspace{1mm}\text{-}\gamma_{3}\hspace{1mm}\gamma_{10})\\
+&(\gamma_{4}\hspace{1mm}\gamma_{9}\hspace{1mm}\text{-}\gamma_{4}\hspace{1mm}\text{-}\gamma_{9})
+(\gamma_{5}\hspace{1mm}\text{-}\gamma_{8}\hspace{1mm}\text{-}\gamma_{5}\hspace{1mm}\gamma_{8})
+(\gamma_{6}\hspace{1mm}\gamma_{7}\hspace{1mm}\text{-}\gamma_{6}\hspace{1mm}\text{-}\gamma_{7}).
+\end{align*}
+
+\compav{Ended here August 3.}
 
@@
 %\begin{proof}
-%Figure $\ref{fig:staircase}$ explicitly demonstrates how $\tilde{\mathbf{M}}$ is tiled by squares. The proof then follows as an obvious result of the previous theorem.
+%figure $\ref{fig:staircase}$ explicitly demonstrates how $\tilde{\mathbf{M}}$ is tiled by squares. The proof then follows as an obvious result of the previous theorem.
 %\end{proof}
@@
 We define ${\mathbf M}$ to be the quotient.
-It is depicted in Figure ...}
+It is depicted in figure ...}
 %
@@
 
-\section{Proof of Main Theorem}
-The following section studies the translation surface's various symmetries and associated Veech group.
-In doing so, we establish what sorts of properties of $\tilde{\mathbf{M}}$ remain invariant when undergoing affine transformations.
-Since $\tilde{\mathbf{U}}$ is a cover of $\tilde{\mathbf{M}}$, we will study the first homology classes of $\tilde{\mathbf{M}}$ that realize $\tilde{\mathbf{U}}$ as a $\mathbb{Z}^2$ cover.
-The goal in the end is to construct a basis for the homotopy classes of paths on $\tilde{\mathbf{M}}$ that can be lifted onto $\tilde{\mathbf{U}}$.
-We generalize how the Veech group of $\tilde{\mathbf{M}}$ acts on homology, and show that closed geodesics on the base surface whose lifts onto the $\mathbb{Z}^2$-cover extend to a homomorphism of $\pi_1(\tilde{\mathbf{M}})$ to $\mathbb{Z}^2$.
-The kernel of this homomorphism determines the cover, and any element of the Veech group acting on $\tilde{\mathbf{M}}$ will stabilize the surface and extend to an isomorphism of its homology classes.
-
-\subsection{Veech Group and Symmetries}
-Observe that $\tilde{\mathbf{M}}$ shares symmetries with the square (up to translation).
-This is made obvious by the fact that $\tilde{\mathbf{M}}$ is tiled by squares.
-Any rotations and reflections acting on $\tilde{\mathbf{M}}$ respects its boundaries.
-We use their matrix representations, and consider only the orientation preserving transformations of determinant one.
-
-Theorems of Gutkin and Judge tell us that any Veech surface that is tiled by squares has a Veech group that is a finite-indexed subgroup of $SL_2(\mathbb Z)$.
-Therefore the group is a lattice in $SL_2(\mathbb R)$.
-We will call the Veech group of $\tilde{\mathbf{M}}$, $\Gamma$.
-$\Gamma$ will without a doubt be generated by parabolic elements, such as skew matrices, and $\tilde{\mathbf{M}}$'s various isometries.
-
-It is easily shown that $SO_2(\mathbb Z)\subset\Gamma$.
-Consider a rotation about the center of any square.
-By a series of cutting and gluing operations, $\tilde{\mathbf{M}}$ can be reconstruced as a staircase again, sending one edge to exactly one other.
-
-Each square is unit, and any pair of edges have two squares between them. This gives rise to a very natural cylinder decomposition of $\tilde{\mathbf{M}}$ in both horizontal and vertical directions.
-
-The staircase is then decomposed into 6 uniform cylinders of moduli $\frac{1}{2}$ in both the vertical and horizontal directions.
-Any geodesic on the horizontal cylinders are incident with exactly two other geodesics on the vertical cylinders.
-
-\begin{Def}
-The \textbf{set of horizontal cylinders} of $\tilde{\mathbf{M}}$, denoted $\Delta_{i}$ $(1\leq i \leq 6)$, are the cylinders with boundaries formed by saddle connections made between unions of singularities in the horizontal direction.
-\newline
-Likewise, the \textbf{set of vertical cylinders}, denoted $\varepsilon_j$ $(1\leq j 
-\leq 6)$, are the cylinders with boundaries formed by saddle connections made between unions of singularities in the vertical direction. \newline
-Refer to Figure $\ref{fig:decomp}$ for the decomposition
-\end{Def}
-
-Horizontal and vertical cylinders of the surface are disjoint within their respective sets, and their union is the surface $\tilde{\mathbf{M}}$.
-
-  Recall that the modulus of a cylinder, $\mu$, is given by the ratio of the cylinder's width to its circumference, $\mu{}={}w/c$.
-A surface that is composed entirely out of similar horizontal and vertical cylinders is stabilized by Dehn-twists whose derivatives are matrices of magnitude $\lambda$ equal to that of lowest common multiple of the inverse moduli, $\mu^{-1}$.
-[cite Veech 89?]
-
-In this case, the moduli are easily obtained and uniform throughout $\tilde{\mathbf{M}}$'s cylinders.
-We can see that any skews of magnitude $\lambda=\pm2$, since $\mu^{-1}{}=2$, are derivatives of Dehn-twists that preserve the boundaries of the cylinders.
-Figure $\ref{fig:skew}$ demonstrates a Dehn-twist of magnitude 2 acting on both horizontal and vertical cylinders.
-Hence the derivatives,
-\begin{equation}
-\left[ \hspace{1mm} \begin{matrix}
-				1 & \pm 2\\
-				0 & 1
-			\end{matrix}\hspace{1mm}\right],\text{ and }
-			\left[ \hspace{1mm} \begin{matrix}
-							1 & 0\\
-							\pm 2 & 1
-						\end{matrix}\hspace{1mm}\right]
-\end{equation}
-are the parabolic generators of $\Gamma$. The fact that $\lambda=1$ does \emph{not} stabilize $\tilde{\mathbf{M}}$ is an absolutely integral part of the major proof.
-
-\newpage
-\subsection{Defining Homology Classes of $\tilde{\mathbf{M}}$}
-\compav{TODO:\\
-Construct homotopy class 12-gon\\
-Construct staircase with class intersection\\}
-The cylinder decomposition of $\tilde{\mathbf{M}}$ into 12 vertical and horizontal cylinders naturally produce a set of closed homotopic paths that travel along the cylinder's circumference.
-Because of $\tilde{\mathbf{M}}$'s translational symmetries, we consider every cylinder to contain a set of homologous paths, characterized by the geodesic running along the cylinder's meridian.
-
-Since none of these paths can't be contracted into a single point or deformed into another path without crossing over a singularity, they are non-trivial and span the set of homology classes of $\tilde{\mathbf{M}}$.
-We denote the set of closed paths on the cylinders by $\gamma_{0,..,11}$, where the even-indexed paths are horizontal and the odd-indexed paths are vertical.
-(see figure)
-
-From these paths we create the following homology classes that determine our $\mathbb{Z}^2$-cover:
-
-Here $u,v,$ and $z$ belong to $H_1(\tilde{\mathbf{M}},\mathbb Z)$.
-These classes are generalizations of differential 1-forms, say $dx,dy,dz$, on the four-fold cover whose path integrals on $\tilde{\mathbf{M}}$ map directly to the cover.
-In our case, there is translational symmetry and it is understood that any geodesic with rational slope is periodic, so we need only consider $\tilde{\mathbf{M}}$'s homology classes.
-Let $\alpha$ be a closed path on $\tilde{\mathbf{M}}$ as an element of $\pi_1(\tilde{\mathbf{M}})$.
-The intersection number of any two homology classes is given as the bilinear form:
-
-\begin{align}
-i:H_1(\tilde{\mathbf{M}},\mathbb Z)\times H_1(\tilde{\mathbf{M}},\mathbb Z)\rightarrow \mathbb Z.
-\end{align}
-
-The $\mathbb{Z}^2$ cover, $\tilde{\mathbf{U}}$, is associated to the kernel of the homomorphism
-\begin{align}
-\Omega_{u,v}:\pi_1(\tilde{\mathbf{M}})\rightarrow \mathbb{Z}^2, \alpha\mapsto(i(u,[\alpha]),i(v,[\alpha])).
-\end{align}
-The classes u and v determine where the lifted geodesic terminates on $\tilde{\mathbf{U}}$.
-Careful experimentation and observation shows us that these classes correspond to moving along the horizontal and vertical directions relative to $\tilde{\mathbf{U}}_0$'s hole indexing convention.
-We suspect that the remaining z class is some remnant of how the quotient topology of $\tilde{\mathbf{U}}_0$ identifies a point on the boundary of a hole with another as an element of $\mathbb{Z}/4\mathbb Z$, as these are the cylinders of $\tilde{\mathbf{U}}_0$ that cross from one phase to the next (See Figure $\ref{fig:mtilda}$).
-Regardless, the classes $u,v$ are enough to determine $\tilde{\mathbf{U}}$'s fundamental group as they already take into consideration the effect that the rotations of $\tilde{\mathbf{U}}_0$ has on $\tilde{\mathbf{U}}$'s translational symmetries.
 \newpage
@@
 \raisebox{0.5in}{\includesvg[width=2.2in]{unfoldcutpath}}
-\\ \text{Here we can see paths a,b, and c on }$\mathbf{U}$\text{ and }$\mathbf{S}$,
-\\ \text{where their initial trajectories are (0,1), (1,0), and (1,1) respectively.}
 \end{figure}

### 2017-08-04 untitled.m

@@ -1,0 +1,2 @@
+mat=[sym(0), sym(-1), sym(0), sym(0), sym(0), sym(0), sym(0), sym(0), sym(0), sym(0), sym(0), sym(-1); sym(1), sym(0), sym(1), sym(0), sym(0), sym(0), sym(0), sym(0), sym(0), sym(0), sym(0), sym(0); sym(0), sym(-1), sym(0), sym(-1), sym(0), sym(0), sym(0), sym(0), sym(0), sym(0), sym(0), sym(0); sym(0), sym(0), sym(1), sym(0), sym(1), sym(0), sym(0), sym(0), sym(0), sym(0), sym(0), sym(0); sym(0), sym(0), sym(0), sym(-1), sym(0), sym(-1), sym(0), sym(0), sym(0), sym(0), sym(0), sym(0); sym(0), sym(0), sym(0), sym(0), sym(1), sym(0), sym(1), sym(0), sym(0), sym(0), sym(0), sym(0); sym(0), sym(0), sym(0), sym(0), sym(0), sym(-1), sym(0), sym(-1), sym(0), sym(0), sym(0), sym(0); sym(0), sym(0), sym(0), sym(0), sym(0), sym(0), sym(1), sym(0), sym(1), sym(0), sym(0), sym(0); sym(0), sym(0), sym(0), sym(0), sym(0), sym(0), sym(0), sym(-1), sym(0), sym(-1), sym(0), sym(0); sym(0), sym(0), sym(0), sym(0), sym(0), sym(0), sym(0), sym(0), sym(1), sym(0), sym(1), sym(0); sym(0), sym(0), sym(0), sym(0), sym(0), sym(0), sym(0), sym(0), sym(0), sym(-1), sym(0), sym(-1); sym(1), sym(0), sym(0), sym(0), sym(0), sym(0), sym(0), sym(0), sym(0), sym(0), sym(1), sym(0)];
+basis(mat)

### 2017-08-09 paper.tex
070917

@@
 
-\newtheorem{thm}{Theorem}[subsection]
-\newtheorem{Def}{Definition}[subsection]
+\newtheorem{thm}{Theorem}[section]
+\newtheorem{Def}{Definition}[section]
 \newtheorem*{thm*}{Theorem}
 \newtheorem*{def*}{Definition}
-\newtheorem{lem}{Lemma}[subsection]
+\newtheorem{lem}{Lemma}[section]
 \newtheorem*{rem}{Remark}
@@
 \subsection{Some Background}
-This paper assumes that the reader is familiar with fundamental concepts in algebraic topology, such as covering space theory, and the study of Veech surfaces and their associated Veech groups.
+This paper assumes that the reader is familiar with fundamental concepts in algebraic topology, such as covering space theory and the study of Veech surfaces and associated Veech groups.
 In order to prove this theorem, we take a number of covers of a surface isometric to $\mathbf S$.
@@
 
-\subsection{A Path-Isometric Map From $\mathbf S$ to $\mathbf U$}
+\subsection{A Path Isometry From $\mathbf S$ to $\mathbf U$}
 A path-isometric map from $\mathbf{S}$ to $\mathbf{P}$ induces an isometry from the Necker cube surface to $\mathbf{U}$.
@@
 \begin{Def}
-The translation surface $\mathbf{M}$ is given as the quotient  $\tilde{\mathbf{U}}/\mathbf{T}^{m,n}$.
+The translation surface $\mathbf{M}$ is given as a quotient of $\tilde{\mathbf{U}}$ equipped with a  holomorphic 1-form, $(\tilde{\mathbf{U}}/\mathbf{T}^{m,n},\omega)$.
 \end{Def}
@@
 By a series of cutting and gluing operations, $\mathbf M$ has this symmetric representation as a square-tiled staircase (figure $\ref{fig:staircase}$).
+This is a compact surface with an infinite $\mathbb Z$-cover as the Infinite Staircase.
+[cite] It shares many of the same symmetries, as well.
 The vertices of cone angle $2\pi$ can be found on the centers of the squares in the figure below where the black lines intersect.
@@
 As a result of Gutkin-Judge [citation], $\mathbf M$ is a Veech surface whose Veech group is commensurable to $SL(2,\mathbb Z)$.
-In order to determine which elements of orientation preserving affine diffeomorphisms, denoted Aff$^{+}(\mathbf M)$, have derivatives that belong to the Veech group, V($\mathbf M$), it is necessary to study $\mathbf M$'s various symmetries and cylinder decompositions.
+In order to determine which elements of orientation preserving affine diffeomorphisms, denoted Aff$^{+}(\mathbf M)$, have derivatives that belong to the Veech group, V($\mathbf M$), it is necessary to study $\mathbf M$'s various symmetries, as well as its cylinder decomposition.
 \vspace{0.1in}
@@
 
-One group contained in $V(\mathbf M)$ is SO$(2,\mathbb{Z})$, the group generated by rotations of order 4.
-Even though the other elements in $\mathbb{D}_4$ are derivatives of automorphisms of $\mathbf M$, we will only consider those of determinant +1.
+One group contained in $V(\mathbf M)$ is SO$(2,\mathbb{Z})$, the group generated by rotations of the square.
 We want the corresponding affine map of these rotation matrices to obey the convention that we \emph{fix} the second square on the bottom with label I.
-These maps are unique only up to translation and as such assigning this convention makes it easier to keep track of where the edges of $\mathbf M$ end up afterwards.
-The diffeomorphism with 90 degree counter-clockwise rotation fixed about that point at the center of that square is denoted $\mathbf{R}$.
+These maps are unique only up to a translation of the surface and assigning this convention makes it easier to keep track of where the edges of $\mathbf M$ end up.
+The 90 degree counter-clockwise rotation fixed about the point in the center of that square is denoted $\mathbf{R}$.
 Its derivative, $\mathbf{R}'$, is the matrix $\left[ \hspace{1mm} \begin{matrix}
@@
 Observe that together they satisfy the property $(\mathbf{H}\mathbf{V})^{6}=(\mathbf{V}\mathbf{H})^{6}=id$.
-This essentially moving the squares up the staircase until they come back to where they started.
-The group of translation deck transformations of the surface is isomorphic to a permutation group of 12 elements (each square) generated by these two translations that satisfy the relation.
-It is the group $\mathbb{D}_{12}$ generated by reflections about two adjacent vertices of a 12-gon.
-Thus it is possible to translate a square to any other on the surface as a composition of these two translations.
+This is essentially moving the squares up the staircase until they come back to where they started, unlike the the group of diagonal translations of the Infinite Staircase that is isomorphic to $\mathbb Z$.
+The group Trans($\mathbf M$) is isomorphic to the group generated by reflections of a 12-gon, $\mathbb{D}_{12}$.
 
@@
 
-Both sets of cylinders are indexed $0\leq i, j \leq 5$.
+Both sets of maximal cylinders are indexed $0\leq i, j \leq 5$.
 Every cylinder is composed of two unit squares with edges identified by translation.
+Each cylinder is obtained as a ramified cover of the square torus.
 We start counting the cylinders from the bottom-most cylinder and increment the index as the cylinders fill the surface going ``up" the staircase.
 $\small{\mathcal{E}}_0$ is the first cylinder at the bottom whose identifications are \em{not} \normalfont the solid triangles.
-$\small{\mathcal{E}}_5$ is the cylinder whose edges are paired by the solid triangles.
-$\Delta_{i}$ has the more standard representation starting from the bottom-most horizontal cylinder.
+Travel up the staircase until the laste one, $\small{\mathcal{E}}_5$, the cylinder whose edges are paired by the solid triangles and glue the 12th square with the 1st.
+$\Delta_{i}$ has the more standard representation starting from the bottom-most horizontal cylinder and going all the way up to the top.
 
@@
 \end{equation*}
-such that $\lambda$ is that lowest common multiple.
+such that $\lambda$ is that lowest common multiple. [cite]
 
 Let $\mu^{-1}_{\Delta_i},\mu^{-1}_{\mathcal{E}_j}$ denote the inverse moduli of the cylinders.
-The cylinders $\Delta_{i}$ and $\small{\mathcal{E}}_j$, as i and j vary, all have the same inverse moduli since they are all defined as two unit squares with identifications made by translation along the cylinders circumference.
+The cylinders $\Delta_{i}$ and $\small{\mathcal{E}}_j$, as i and j vary, all have the same inverse moduli since they are all defined as two unit squares with identifications made by translation.
 Therefore, $\lambda=$ l.c.m.($\mu^{-1}_{\Delta_i}\cup\mu^{-1}_{\mathcal{E}_j}$) $=2$.
@@
 
-It is now possible to generate the \emph{group of symmetries} of $\textbf{M}$ as Deck transformations of the surface. We call this group $\mathbb X$ and say that
+It is now possible to generate the \emph{group of symmetries} of $\textbf{M}$ with the elements of these affine transformations. We call this group $\mathbb X$ and say that
 
@@
 
-The group of derivatives of $\mathbb X$ is $\mathbb X'$, and it is generated by elements of V$(\textbf{M})$.
-It is not apparent whether or not $\mathbb{X}=$Aff$^+(\mathbf{M})$ or if $\mathbb X'=$ V$(\mathbf{M})$.
+The group of derivatives of $\mathbb X$ is the image of the map D: $\mathbb X\rightarrow SL(2,\mathbb{R})$, denoted $\mathbb X'$.
+We know that $\mathbb{X}\subset$Aff$^+(\mathbf{M})$ and $\mathbb X'\subset$ V$(\mathbf{M})$.
+The corresponding elements of $\mathbb X'$ will use the same prime notation.
 
 \subsection{Action of $\mathbb X$ on the Homology of $\mathbf M$}
-The group $\mathbb X$ consists of the set of Deck transformations of $\mathbf{M}$, whose induced isomorphisms of $\pi_1(\mathbf{M})$ acts on the absolute homology of $\mathbf{M}$ and extends to an action on the homology of $\mathbf{M}$'s $\mathbb{Z}^2$-cover, $\tilde{\mathbf{U}}$.
-$\mathbf{M}$'s cylinder decomposition gives us a natural set of homology classes that serve as a basis for $H_1(\mathbf{M},\mathbb Q)$.
-For a total of 12 cylinders, there are 12 core curves:
+The elements of $\mathbb X$ induces isomorphisms of $\pi_1(\mathbf{M})$ and extends to an action on the homology of $\mathbf{M}$'s $\mathbf{T}^{m,n}$-cover, $\tilde{\mathbf{U}}$.
+
+$\mathbf{M}$'s cylinder decomposition gives us a natural set of homology classes that serve as a basis for $H_1(\mathbf{M},\mathbb Q)$. For a total of 12 cylinders, there are 12 core curves:
 
@@
 \centering
-\caption{Cylinder core curves with u,v, and z homology classes that determines the $\mathbb{Z}^2$-cover.}
+\caption{Cylinder core curves with u,v, and z homology classes that determines the $\mathbf{T}^{m,n}$-cover.}
+\label{fig:homology}
 \end{figure}
@@
 The same notation will be used for the isomorphisms themselves.
-We denote the \emph{set of cylinder core curves} as $\Gamma$.
+We denote the \emph{set of cylinder core curves} as $\Gamma\subset H_1(\mathbf{M},\mathbb Q)$.
 
@@
 
-The 12-gon was mentioned in the previous section under the discussion of $\mathbf{M}$'s translational symmetries. $\mathbb X_*$ acting on absolute homology is better represented as a symmetric object.
+The 12-gon was mentioned in the previous section under the discussion of $\mathbf{M}$'s translational symmetries.
+$\mathbb X_*$ acting on absolute homology is better represented as this symmetric object.
 
@@
 This was so that their induced isomorphisms would \emph{fix} one or more vertices as they act on homology.
-But because of the translational symmetries, it is possible to rotate about the center of any square and essentially rotating the entire 12-gon with $\mathbf{H}_*$ and $\mathbf{V}_*$.
+Because of the translational symmetries, it is possible to rotate or reflect entire 12-gon with $\mathbf{H}_*$ and $\mathbf{V}_*$.
+These permutations of $\Gamma$ generate a group isometric to $\mathbb{D}_{12}$.
 
@@
 
-The two preserve the signs of the core curves. It is to be expected as they are induced isomorphisms of translation maps. They are given as the following permutations of $\Gamma$:
+The two preserve the signs of the core curves. It is to be expected as they are induced isomorphisms of translations. They are expressed as:
 \space
@@
 
-Rotation about the center of a square will not only change signs of the core curves, but permute elements in such a way that 4 consecutive rotations of $\mathbf{M}$ is an identity on $\Gamma$.
+Rotation about the center of a square will not only change signs of the core curves, but permute elements in such a way that the rotation element is order four.
 
 \begin{minipage}{0.4\textwidth}
-\begin{figure}[H]
-\hspace{0}\includegraphics[width=1.7in]{12gonR.pdf}
+\begin{figure}[H]\includegraphics[width=1.7in]{12gonR.pdf}
 \end{figure}
@@
 \begin{itemize}
-\item[\textbf{\emph{$\mathbf{R}_*$}}] This action reflects about the line and maps one vertex to another.
+\item[\textbf{\emph{$\mathbf{R}_*$}}] This action reflects the 12-gon about that line and maps one vertex to another.
 No vertex is fixed.
 However it has the additional effect of changing the sign of a curve only if it was originally black (or odd index).
-The pattern is peculiar, but $\mathbf{R}_*$ is still order four.
 \end{itemize}
@@
 \vspace{2mm}
-We can see the effect that it has on homology clearer as a permutation of $\Gamma$:
+We can see the effect that it has on homology clearer as a permutation group on $\Gamma$:
 \begin{align*}
@@
 
+\compav{TODO: Make the parabolic 12-gon image.}
+
+\begin{figure}[H]
+\centering
+\includesvg{12gonA}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includesvg{12gonABexample}
+\end{figure}
 \compav{Ended here August 3.}
 
-Regardless, $u$ and $v$ are the most important homology classes that we use to determine whether or not a lifted path properly closes on the four-fold cover.
-Let $\alpha:[0,1]\rightarrow\mathbf M$ be a curve on the surface such that $\alpha(0)=\alpha(1)=x_0\in\mathbf{M}$.
-Take $f:\tilde{\mathbf U}\rightarrow\mathbf M$ to be the covering map that takes every point on $\tilde{\mathbf U}$ to its quotient in $\mathbf{M}$.
-We will fix a point $\tilde{x}_0\in\tilde{\mathbf U}$ such that the lifted path $\tilde\alpha(0)=\tilde{x}_0$ under the function $f$.
-$\alpha$ is a closed path in $\pi_1(\mathbf M,x_0)$.
-We denoted the class of abelianized paths $\alpha$ as [$\alpha$]$\in H_1(\tilde{\mathbf{M}},\mathbb Z)$, and define the function,
+\subsection{$\mathbf{T}^{m,n}$-Cover of $\mathbf M$}
+When defining $\mathbf{M}$ as a quotient over the action of $\mathbf{T}^{m,n}$, the result was a finite-type translation surface whose cover indexes these four "cut-open tori" by the integer pair $(m,n)$.
+For this reason it is easy to consider $\tilde{\mathbf{U}}$ to be a kind of $\mathbb{Z}^2$-esque covering space of $\mathbf{M}$.
 
-\begin{align}
-i:H_1(\mathbf{M},\mathbb Z)\times H_1(\mathbf{M},\mathbb Z)\rightarrow \mathbb Z,
-\end{align}
+\begin{lem}
+The family of translational symmetries $\mathbf{T}^{m,n}$ is isomorphic to $\mathbb Z^2$.
+\begin{proof}
+$\mathbf{T}^{m,n}$ is induced by rotational and translational symmetries of the fundamental domain of $\tilde{\mathbf{U}}$, $\mathbf P \times \mathbb{Z}/4\mathbb{Z}$.
+It's clear that the rotations themselves are bijective maps.
+By definition, $\mathbf{T}_0^{m,n}=(x+2m,y+2n;j)$.
+As a group action, observe that the points $(x,y;j)$ are stabilized.
+The orbit of this triple is the set of infinitely many points in its domain parametrized by integers $m,n$.
+The obvious isomorphism identifies an element of $\mathbb{Z}^2$ with $(m,n)$.
+The orbit of any point is itself given as $\mathbf{T}_0^{\mathbb Z^2}$.
+Therefore, the induced map on $\tilde{\mathbf{U}}$, $\mathbf{T}^{m,n}$, is isomorphic to $\mathbb Z^2$ as well.
+\end{proof}
+\end{lem}
 
-to return the intersection number of two homology classes.
-Out of convention, we consider a positive intersection to adhere to the right-hand rule and take the crossings to be relative to $u$ and $v$.
-With that being said, the signed homology classes of gamma curves are
+From figure $\ref{fig:mtilda}$ we construct homology classes $u,v$ according to the outer edges of the tori.
+In addition, there is a class $z$ that doesn't play a role in determining the cover $\tilde{\mathbf U}$.
+This is because $z$ is the set of core curves that lie on cylinders that go from one plane to the next.
+Figure $\ref{fig:homology}$ labels them on the staircase model.
 
@@
 
-The $\mathbb{Z}^2$ cover, $\tilde{\mathbf{U}}$, is associated to the kernel of the homomorphism
+Let i be a bilinear, non-degenerate form given as
+
 \begin{align}
-\Omega_{u,v}:\pi_1(\mathbf{M},x_0)\rightarrow \mathbb{Z}^2, \text{ such that } \alpha\mapsto(i(u,[\alpha]),i(v,[\alpha])).
+i:H_1(\mathbf{M},\mathbb Q)\times H_1(\mathbf{M},\mathbb Q)\rightarrow \mathbb Z.
 \end{align}
-Suppose that $h:\mathbf{M}\rightarrow\mathbf{M}$ is an automorphism of the surface such that dir $h\in SL(\mathbf M)$.
-$h$ induces the isomorphism, $h_*:\pi_1(\mathbf{M},x_0)\rightarrow \pi_1(\mathbf{M},x_0)$.
-Likewise, $f$ induces the inclusion map $f_*:\pi_1(\tilde{\mathbf{U}},\tilde x_0)\hookrightarrow\pi_1(\mathbf{M},x_0)$.
-If $\pi_1(\tilde{\mathbf{U}},\tilde x_0)$ is associated with the Ker $\Omega_{u,v}$, then it must be true that the kernel of such a homomorphism is preserved under $h_*$ and there exists a map from  $\pi_1(\mathbf M, x_0)$ to $\pi_1(\tilde{\mathbf{U}},\tilde x_0)$.
 
-The actions of $h_*$ are dual to the actions of dir $h$, denoted $h'$, on the sets of odd-odd and even-odd $\mathbb{Z}^2$ vectors, $\mathcal{O}$ and $\mathcal{E}$.
-In particular, the action of $\mathbb X$ on $\mathcal{O}$ and $\mathcal{E}$ is self-contained and well-defined.
-This is vital as what we want to show is that $h'(\alpha')$, where $\alpha'$ is a constant and $\alpha$ is a closed geodesic, is dual to $h_*([\alpha])$.
-In which case, it should also be true that $h_*(\alpha)\mapsto(i(u,h_*([\alpha])),i(v,h_*([\alpha])))$, under $i$.
-And since intersection number is a non-degenerate bilinear form,
+Let $\beta\in \pi_1(\mathbf{M}, x_0)$, and $[\beta]$ be its abelianization in $H_1(\mathbf{M},\mathbb{Q})$.
+The function returns an \emph{intersection number} between two elements in $H_1(\mathbf{M},\mathbb Q)$.
+Assuming that $[\beta]$ is the second argument of i, a positive crossing occurs when $[\beta]$'s angle of intersection with the first homology class is positive.
+Fix a point $x_0\in\mathbf{M}$.
+$x_0$ lifts to $\tilde{x}_0\in\tilde{\mathbf{U}}$ by the covering map $f$, such that $\tilde{x}_0$ belongs to the quotient of $\tilde{\mathbf{U}}$ indexed $\mathbf{T}^{0,0}$.
 
-\begin{align*}
-\alpha\mapsto(i(h_*^{-1}(u),[\alpha]),i(h_*^{-1}(v),[\alpha])).
-\end{align*}
+$\tilde{\mathbf{U}}$ is associated to the kernel of the group homomorphism,
+\begin{align}
+\Omega_{u,v}:\pi_1(\mathbf{M},x_0)\rightarrow \mathbb{Z}^2 \text{; } \beta\mapsto(i(u,[\beta]),i(v,[\beta])).
+\end{align}
+Since $\mathbf{T}^{m,n}$ is isomorphic to $\mathbb{Z}^2$, the lift of $\beta\in \text{Ker } \Omega_{u,v}$ is the closed path, $\tilde{\beta}$.
+Let $F^t_\theta$ be a \emph{non-singular} geodesic flow on $\mathbf{M}\backslash$Sing$(\mathbf M)$ with initial point $x_0$.
+By [some theorem], $F^t_\theta$ is closed for for every rational direction $\theta$.
+The image of $F^t_\theta$ is an element of $\pi_1(\mathbf M, x_0)$, and homotopic to closed path $\beta$.
 
-This amounts to finding some such path whose lift belongs to the fundamental group of $\tilde{\mathbf U}$, noting the effect $h_*$ has on $\mathbf M$'s homology, and finding the new intersection number.
-Let $\beta$ be such a path.
-Suppose there is some sequence of affine maps $h_1,...,h_n\in$Aff$^+(\mathbf{M})$ such that $h_1' \circ ...
-\circ h_n'(\alpha')=\beta'$.
-We will show the following is true:
+\begin{Def} Identify $\mathbb C$ with $\mathbb{R}^2$ in the usual way.
+The closed path $\alpha\in\pi_1(\mathbf M, x_0)$ is the path contained in the maximal cylinder decomposition in direction $(1,1)$.
+\end{Def}
 
-\begin{align*}
-h_1 \circ ... \circ h_n(\mathbf{M})=\mathbf{M}'&\iff\\
-h_{*_1} \circ ... \circ h_{*_n}([\beta])=[\alpha] &\iff
-\end{align*}
-%\subsection{$\tilde{\mathbf{M}}$ as a Veech Surface}
-%We use basic properties of $\tilde{\mathbf{M}}$ to show that it is Veech, and its Veech group is a lattice in $SL_2(\mathbb{Z})$.
-Recall that translation surfaces are equipped with charts, $\omega$, of maps from the surface to $\mathbb{C}$.
-The Veech group of the surface $\tilde{\mathbf{M}}$ is given as the image of its stabilizer in the projective group $PSL_2(\mathbb{R})$.
-%
-%\begin{Def}
-%Since $\tilde{\mathbf{M}}$ is a translation surface with four singularities, it is a Veech surface that locally behaves like a torus away from these cone singularities.
-$\tilde{\mathbf{M}}$ has a \textbf{Veech group}, denoted $\Gamma$.
-%\end{Def}
-%
-%We use the following theorem of Gutkin and Judge to show that $\Gamma$ is lattice (and arithemtic) in $SL_2(\mathbb{Z})$:
-%
-%\begin{thm}(Gutkin-Judge) A translation surface $(S,\omega)$ is squaretiled
-%if and only if its Veech group $V(S,\omega)$ shares a finite-index subgroup
-%with $SL_2(\mathbb{Z}).$
-%\end{thm}
-%
-%\begin{thm}The following are equivalent:
-%\begin{enumerate}[label=(\roman*)]
-%\item $\tilde{\mathbf{M}}$ is tiled by squares.
-%\item $\Gamma$ is lattice in $SL_2(\mathbb{Z})$.
-%\end{enumerate}
-%\begin{proof}
-%figure $\ref{fig:staircase}$ explicitly demonstrates how $\tilde{\mathbf{M}}$ is tiled by squares. The proof then follows as an obvious result of the previous theorem.
-%\end{proof}
-%\end{thm}
-
-\newpage
-%\subsection{Translation Surface}
-%\compat{You need to build the surface out of pieces rather than out of ``copies of $\mathbf{M}$.''} 
-%With $\tilde{\mathbf{U}}$ defined, we can describe its quotient as $\tilde{\mathbf{U}}$ modded out by translations.
-We use the induced isometry, $\bar\phi$, to respect our original notion of direction on $\mathbf{U}$ as a quotient of $\mathbf{P}$.
-%
-%
-%
-%\begin{Def}
-%The translation surface $\tilde{\mathbf{M}}$ is given as the quotient $\tilde{\mathbf{U}}/T$, where $T:(2\mathbb{Z})^2\times\tilde{\mathbf{U}}\rightarrow\tilde{\mathbf{U}}$ is the group action defined as $T_{m,n}(x)=\bar{\phi}\left((m,n)+\bar{\phi}^{-1}(x)\right)$ for $x\in\tilde{\mathbf{U}}$.
-%\end{Def}
-%
-%$\tilde{\mathbf{M}}$ is also a degree four branched cover of $\mathbf{M}$.
-A covering map $\tilde{\mathbf{M}} \rightarrow \mathbf{M}$ individually rotates these squares about the centers of their holes according to their labels and maps their points to $\mathbf{M}$.
-This is a genus five translation surface with four singularities.
-$\tilde{\mathbf{M}}$ as a quotient of $\tilde{\mathbf{U}}$ can be represented as quotients of each plane:
-%
-%\begin{figure}[H]
-%\centering
-%\includesvg[width=4in]{mtilda}
-%\end{figure}
-%
-%All in all, we have the following diagram:
-
-%\begin{figure}[H]
-%\centering
-%\includegraphics[width=2in]{commute.pdf}
-%\end{figure}
-%
-%\newpage
-%\subsection{The Euclidean cone surface $\mathbf{M}$}
-%\compat{I would say: Observe that translations of ${\mathbb R}^2$ by pairs even integers preserves the subset ${\mathbf P}$ and respects the equivalence relation defining ${\mathbf U}$.
-Thus this gives an action of $(2{\mathbb Z})^2$ on ${\mathbf U}'$.
-We define ${\mathbf M}$ to be the quotient.
-It is depicted in figure ...}
-%
-%Observe that the translations by pairs of even integers preserves the subset $\mathbf P$ and respects the equivalence relation on $\mathbf U$
-%
-%We will call this quotient space $\mathbf{M}$ and define it as such:
-%\begin{Def}{$\mathbf{M}$.}\\
-%$\mathbf{M}$ is a base surface of $\mathbf{U}'$ given by the quotient $\mathbf{U}'/2\mathbb{Z}\oplus2\mathbb{Z}$, where $2\mathbb{Z}\oplus2\mathbb{Z}\times\mathbf{U}'\rightarrow\mathbf{U}'$ is the group action on $\mathbf{U}'$ defined as $(m,n)\circ(u,v)=(u+m,v+n)$.
-%\end{Def}
-%
-%\begin{figure}[H]
-%\begin{center}
-%\includesvg[width=2in.]{unfoldcutquotient}\hspace{0.2in}
-%\includesvg[width=2in.]{quotient}
-%\end{center}
-%\caption{$\mathbf{U}'$ equivalence classes under group action (left), and $\mathbf{M}$ with identified edges and vertices (right).}
-%\label{fig:quotient}
-%\end{figure}
-%
-%The quotient space is a genus 1 surface that resembles a $2\times2$ torus with a unit square removed from it.
-Near the outer vertex it behaves as a torus might.
-The inner vertices $c_{1},c_{3}$ have cone angle $\frac{3\pi}{2}$, and $c_{2}$ has a cone angle of $3\pi$.
-None of these inner vertices have cone angles that are integral multiples of $2\pi$, so in order to create the necessary translation surfaces we require the minimal number of copies of both $\mathbf{U}$ and $\mathbf{M}$ to construct their associated phase spaces.
-\compat{Move the discussion of building the translation surface to the next section.
-Otherwise this paragraph is good.}
-
-\newpage
-\subsection{Horizontal, Vertical, and Slope One Cylinders of $\tilde{\mathbf{M}}$}
 \begin{figure}[H]
 \centering
-\includesvg[pretex=\tiny, width=4in]{slopeonecylinder}
+\includesvg[pretex=\tiny, width=4.5in]{slopeonecylinder}
+\caption{In the slope one direction, $\mathbf M$ decomposes into two cylinders.}
 \end{figure}
-\newpage
-\subsection{Slope Normalization Algorithm}
-\newpage
-\subsection{Proof of Theorem}
+
+We can see that $hol(\alpha)=6+6i$ is the holonomy vector of the closed loop $\alpha$ contained in one of these cylinders. Call the unlabeled cylinder $C_1$(right), and the labeled one $C_2$(left).
+
+\begin{lem}
+$\alpha$ is homotopic to the image of the closed geodesic in direction $(1,1)$ on $\mathbf{M}\backslash\text{Sing}(\mathbf M)$.
+\begin{proof}
+Let $\chi$ be a separatrix contained in either $C_1$ or $C_2$.
+Since a separatrix does not admit singularities, it is the image of a closed geodesic on $\mathbf{M}\backslash\text{Sing}(\mathbf M)$ with initial point $x_0$ on the strips of $C_1$ and $C_2$ with boundaries removed, denoted $C_1',$ $C_2'$.
+Express $[\alpha]$ as $\Sigma^{11}_{j=0}$ $\frac{1}{2}\gamma_j$ (a closed path climbing up the staircase).
+We show that the intersection numbers of $[\alpha]$ and $[\chi]$ are the same for every core cylinder curve $\gamma$, i.e.
+$i([\gamma_k], \Sigma^{11}_{j=0} \text{ } \frac{1}{2}[\gamma_j])=i([\gamma_k],[\chi])\text{ } \forall k=0,\dots,11$.
+And since $x_0$ is in either $C_1'$ or $C_2'$, $\alpha$ cannot climb up the staircase without intersecting the core curves.
+Thus, the sum of $\frac{1}{2}$ core curves of $[\alpha]$ have to have to intersect at least one other core curve.
+\\
+\textbf{Case one:} k is even.
+If k is even, then every curve $\gamma_k$ is oriented to the right.
+Since $\chi$ intersects every curve once, $i([\gamma_k],[\chi])=1$.
+No even indexed curves intersect eachother, so we need only consider when $j$ is odd.
+Now if $j$ is odd, it is incident (positively crossing) with only two horizontal curves, namely $\gamma_{j+1},\gamma_{j-1}$.
+Therefore $i([\gamma_k],[\alpha])=i([\gamma_{j-1}],\frac{1}{2}[\gamma_{j}])+i([\gamma_{j+1}],\frac{1}{2}[\gamma_{j}])=\frac{1}{2}(i([\gamma_{j-1}],[\gamma_{j}])+i([\gamma_{j+1}],[\gamma_{j}]))=\frac{1}{2}(1+1)=1.$\\
+\textbf{Case two:} k is odd.
+If k is odd, then $[\chi]$ will have an intersection number of $-1$ with $[\gamma_k]$ since odd-indexed core curves are oriented upwards.
+Now since k is odd, we only consider when $j$ is even.
+Similarly, this means that $\gamma_j$ negatively intersects the two vertical core curves with adjacent indices.
+Hence, $i([\gamma_k],[\alpha])=i([\gamma_{j-1}],\frac{1}{2}[\gamma_{j}])+i([\gamma_{j+1}],\frac{1}{2}[\gamma_{j}])=\frac{1}{2}(i([\gamma_{j-1}],[\gamma_{j}])+i([\gamma_{j+1}],[\gamma_{j}]))=\frac{1}{2}(-1-1)=-1.$\\
+We know intersection number to be a non-degenerate bilinear form, and if $\alpha$ and $\chi$'s abelianizations admit the same intersection numbers for every curve in the basis set of $H_1(\textbf M, \mathbb{Q})$, then $\alpha\sim\chi$.
+\end{proof}
+\end{lem}
+
+We will abuse notation occasionally and refer to $\alpha$ as the geodesic of slope one.
+In the horizontal direction, (1,0), $\mathbf{M}$ has the 6 cylinder decomposition shown in figure $\ref{fig:decomp}$.
+A similar argument shows that any horizontal cylinder curve is homotopic to a curve with holonomy vector 2 on the cylinder.
+The core curves themselves are geodesics on the surface that intersect only two other curves, which is a property used in the previous theorem.
+Their intersection numbers are therefore zero everywhere except for the curves with adjacent indices.
+This is unlike $\alpha$, which intersects every curve exactly once.
+Oftentimes we consider $\alpha$ to travel around the 12-gon counterclockwise, intersecting negatively with odd-indexed $\gamma$ (black nodes), and positively with even-indexed $\gamma$ (red nodes).
+
+The translational and rotational symmetries allows one $\gamma$-curve to end up in the place of any other curve, with positive or negative orientation. For that reason we default to using $\gamma_0$
+
+\begin{rem}
+The Infinite Staircase that covers $\mathbf{M}$ has the same shear automorphisms as $\mathbf{M}$, and similar cylinder decompositions as well.
+In these odd-odd directions, it decomposes into two cylinders of infinite area.
+In even-odd directions, it will decompose into countably many cylinders of finite area.
+[cite]
+\end{rem}
+
+\begin{thm}
+$\alpha\in\pi_1(\mathbf{M}, x_0)$ lifts to a closed path $\tilde{\alpha}\in\pi_1(\tilde{\mathbf{U}},\tilde{x}_0)$.
+\end{thm}
+\begin{proof}
+As a result of the previous theorem, $i([\gamma_i],[\alpha])=1$ when i is even and $i([\gamma_i],[\alpha])=-1$ when i is odd.
+Therefore $i([u],[\alpha])=-i([\gamma_2],[\alpha] +i([\gamma_5],[\alpha]) + i([\gamma_8],[\alpha]) - i([\gamma_{11}],[\alpha])=-1+(-1)+1-(-1)=0$, and $i([v],[\alpha])=1-(-1)-1+(-1)=0$.
+So $\Omega_{u,v}(\alpha)=(0,0)\simeq \mathbf{T}^{0,0}$, by Lemma 3.2.
+As such $\alpha\in$ Ker $\Omega_{u,v}$, and lifts to a closed path on $\tilde{\mathbf{U}}$.
+\end{proof}
+
+\begin{thm}
+The lift of core curves $\Gamma$ on $\tilde{\mathbf{U}}$ are drift-periodic.
+\end{thm}
+\begin{proof}
+\compav{show that intersection with u,v will only get one of 1, 0, or -1. therefore i(x,y) is never (0,0)}
+\end{proof}
 
 \begin{figure}[H]
-\includesvg[width=2.3in]{cubesxyzpaths}\hspace{0.1in}
-\raisebox{0.5in}{\includesvg[width=2.2in]{unfoldcutpath}}
+\centering
+\includesvg[width=1.8in]{cubesxyzpaths}
+\includesvg[width=2.8in]{unfoldcutpath}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics{z2cover.pdf}
+\caption{\compav{Fix this!}}
 \end{figure}

### 2017-08-28 newpaper.tex
082717

\documentclass[]{article}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{amsthm}
\usepackage{svg}
\usepackage{enumitem}
\usepackage{color}
\usepackage{float}
\usepackage[utf8]{inputenc}

\begin{document}
\newtheorem{thm}{Theorem}[]
\newtheorem{Def}{Definition}[]
\newtheorem*{thm*}{Theorem}
\newtheorem*{def*}{Definition}
\newtheorem{lem}{Lemma}
\newtheorem*{rem}{Remark}
\newcommand{\shiftleft}[2]{\makebox[0pt][r]{\makebox[#1][l]{#2}}}
\newtheorem*{conj}{Conjecture}
\newtheorem{cor}{Corollary}[]

\newcommand{\compav}[1]{\textbf{\textcolor{blue}{#1}}}
\newcommand{\compat}[1]{\textbf{\textcolor{red}{#1}}}
\graphicspath{{images/}}
\setsvg{svgpath={./images/}}

% Title Page
\title{Summer Report 2017}
\author{Pavel Javornik}

\title{Dynamical Properties of Geodesic Flows on the Necker Cube Surface}
\author{Pavel Javornik}

\maketitle

\begin{figure}[H]
\centering
\includesvg[width=2.2in]{cubesxyzpaths2}
\end{figure}

%\begin{center}
%
%\includesvg[width=4.8in]{cubecoverphoto}\\
%\end{center}

\begin{abstract}
The Necker cube surface, $\mathbf{S}$, is an infinite topological surface obtained by the periodic tiling of unit cubes by $\mathbb{R}^3$ translational symmetries along a 2-dimensional subspace realized as a Euclidean cone surface with a discrete set of singularities of cone angle $3\pi$ and $\frac{3\pi}{2}$. An isometric ``flattening" of $\mathbf{S}$ is obtained by injective piecewise-linear maps onto the planar subspace. The fundamental domain of the surface admits infinitely unit square \emph{holes} centered on the  $(2\mathbb{Z})^2$ lattice in $\mathbb{C}$. The flat surface, $\mathbf{U}$, has rotational and translation symmetries and a unit tangent bundle, $UT(\mathbf{U})$, that is isomorphic to $\mathbb{Z}/\frac{\pi}{2}\mathbb{Z}$ as a consequence of the edge identifications. A \emph{flat} branched cover, $\tilde{\mathbf{U}}$, of the surface is obtained by a quotient topology defined on four copies of $\mathbf{U}$'s fundamental domain and indexed by elements in $\mathbb{Z}/4\mathbb{Z}$ for the purpose of aligning geodesic flows. It is a \emph{formally étale} cover of $\mathbf{U}$ whose unit tangent bundle is $\mathbb{Z}/2\pi\mathbb{Z}$ and obtained by the monodromy group of the covering map. $\tilde{\mathbf{U}}$ is a translation surface whose quotient by its translational symmetries is a \emph{Veech} surface whose \emph{Veech group} is $SL(2,\mathbb{Z})$-commensurable. This paper's aim is to study the relations between geodesic flows on base surfaces, Deck transformations of their covers, the orbit of rational directions in $\mathbb{Z}^2$ under Veech group action, and use group theoretic results to prove theorems about rational trajectories on the Necker cube surface.
\end{abstract}

\newpage
\section{Introduction}

Put your Problem statement here! Example of a Citation\cite[p.219]{Robotics}. Here's Another Citation\cite{Flueck}

The Necker cube[citation] has made numerous appearances throughout history. The artist famous for his use of optical illusions in his work, M.C. Escher, has occasionally used tilings of the Necker cube such as in the lithographs ``Metamorphosis I," and ``Convex and Concave." (Pictured below.) The crystallographer Louis Albert Necker was credited for having extensively studied the Necker cube's geometry[citation] and remarked on this simple, yet fascinating, optical illusion that appears to ``simultaneously protrude from and intrude into the page."[citation:wolfram] Some might even recognize it as the same surface in which ``Q*bert" hops around on in the 1982 arcade game. [cite?]

\begin{figure}[H]
\begin{center}
\includegraphics[scale=0.4]{escher.jpg}
\includegraphics[scale=0.55]{escher2.jpg}
\caption{The Necker cube tiling as it appears briefly in ``Metamorphosis I," and hung on a banner in ``Convex and Concave."[citation]}
\label{fig:Escher}
\end{center}
\end{figure}

This paper assumes the reader is familiar with fundamental concepts in algebraic topology such as covering space theory, induced homomorphisms on surface homologies of their symmetries, as well as some background in Veech structures of compact translation surfaces. 
\newpage

\subsection{Background}
We will introduce some objects, definitions, and theorems to be used later in the paper. First and foremost, we need to describe the construction of the Necker cube surface and give it a concrete structure as a subset of $\mathbb{R}^3$.

\noindent Take the following compact subsets of auxiliary planes:

\vspace{0.2in}
\begin{tabular}{p{10cm}c}
\begin{align*}
\mathbf{A}_{m,n,p} = [m, m+1]\times[n,n+1]\times\{p\}, 
\\\mathbf{B}_{m,n,p} = \{m+1\}\times[n,n+1]\times[p-1,p],
\\\mathbf{C}_{m,n,p}= [m,m+1]\times\{n+1\}\times[p-1,p].
\end{align*}
&
\shiftleft{0.8in}{\raisebox{-1in}{\includegraphics[scale=1]{label.png}}}
\end{tabular}

\noindent If we take the integer triple $(m,n,p)$ and require that the individual faces are based around them, we obtain the sets

\begin{align*}
\mathbf{A}=\bigcup\big{\{}\mathbf{A}_{m,n,p}: m+n+p=0\big{\}},
\\\mathbf{B}=\bigcup\big{\{}\mathbf{B}_{m,n,p}: m+n+p=0\big{\}},
\\\mathbf{C}=\bigcup\big{\{}\mathbf{C}_{m,n,p}: m+n+p=0\big{\}},
\end{align*}

\noindent and consider them to be piecewise linear maps of infinitely many compact topological spaces, we acquire subsets of the infinite Necker cube surface whose unions cover the surface.

\begin{Def} We denote the \textbf{Necker cube surface} by $\mathbf{S}$, and define it to be the surface of the union of unit squares of the form $\mathbf{S} = \mathbf{A}\cup\mathbf{B}\cup\mathbf{C}$.\end{Def}

The structure is then \emph{flattened} onto x-y plane by piece-wise isometric maps onto a subset of $\mathbb C$ with unit squares removed at every even integer pair in the complex plane:

\begin{figure}[H]
\centering
\includesvg[width=1.1in]{cubesxyzcut} \hspace{0.1in}\raisebox{1.0in}{\text{$\rightarrow$}}\hspace{0.1in}
\raisebox{0.4in}{\includesvg[width=1.5in]{unfoldcut}}
\caption{A piecewise-linear map from $\mathbf{S}$ to $\mathbb{C}$. }
\label{fig:Psi}
\end{figure}
\noindent The red/blue dotted lines are meant to show how the edge identifications are preserved after the transformation. We call the flattened version of the Necker cube surface $\mathbf U$. The map from one fundamental domain to the other is composed of piecewise-linear maps, $\Psi:\mathbf{S}\rightarrow\mathbb{C}$. We obtain an induced isometry between $\mathbf{S}$ and $\mathbf{U}$ from $\Psi$. Identify $\mathbb{C}$ with $\mathbb{R}^2$, and $\Psi$ is given as:

\begin{equation}
\Psi\left[\begin{array}{c}
	x\\y\\z
\end{array}\right] 
= 
\begin{cases}
	\left[ \hspace{2mm} \begin{matrix}
		1 & 0 & 0 \\
		0 & 1 & 0 \\
		0 & 0 & 1
	\end{matrix}\hspace{3mm}\right]

	\left[\begin{array}{c}
	x - \left\lfloor x \right\rfloor
	\\ y- \left\lfloor y \right\rfloor
	\\ -(z - \left\lfloor z \right\rfloor)
	\end{array} \right]
	+
	\left[\begin{array}{c}
		2 \left\lfloor x \right\rfloor - \frac{3}{2}
		\\ 2\left\lfloor y \right\rfloor - \frac{3}{2}
		\\ z - \left\lfloor z \right\rfloor
	\end{array} \right]
		& \text{if } (x,y,z)\in \mathbf{A}	\vspace{2mm}
	\\

	\left[ \begin{matrix}
	0 & 0 & 1 \\
	0 & 1 & 0 \\
	-1 & 0 & 0
	\end{matrix}\hspace{2mm}\right]
	\left[\begin{array}{c}
		x - \left\lfloor x \right\rfloor
		\\ y- \left\lfloor y \right\rfloor
		\\ -(z - \left\lfloor z \right\rfloor)
		\end{array} \right]
	+
		\left[\begin{array}{c}
			2 \left\lfloor x \right\rfloor - \frac{3}{2}
			\\ 2\left\lfloor y \right\rfloor - \frac{3}{2}
			\\ x - \left\lfloor x \right\rfloor
		\end{array} \right]
			& \text{if } (x,y,z)\in \mathbf{B}	\vspace{2mm}
	\\

		\left[ \begin{matrix}
		1 & 0 & 0 \\
		0 & 0 & 1 \\
		0 & -1 & 0
		\end{matrix}\hspace{2mm}\right]
		\left[\begin{array}{c}
			x - \left\lfloor x \right\rfloor
			\\ y- \left\lfloor y \right\rfloor
			\\ -(z - \left\lfloor z \right\rfloor)
			\end{array} \right]
		+
			\left[\begin{array}{c}
				2 \left\lfloor x \right\rfloor - \frac{3}{2}
				\\ 2\left\lfloor y \right\rfloor - \frac{3}{2}
				\\ y - \left\lfloor y \right\rfloor
			\end{array} \right]
				& \text{if } (x,y,z)\in \mathbf{C}	\vspace{2mm}
\end{cases}
\end{equation}

\noindent The fundamental domain of $\mathbf{U}$ is

\begin{equation}
\label{eq:P}
\mathbf{P} = \mathbb{C}\text{ }\backslash\bigcup_{m,n \in {\mathbb Z}} \big{\{} u+vi:u\in(2m-\frac{1}{2},2m+\frac{1}{2}),\text{ } v\in(2n-\frac{1}{2},2n+\frac{1}{2})\big{\}},
\end{equation}

\noindent and it is defined as follows (identify $\mathbf{P}$ with $\mathbb{R}^2$):

\begin{Def} $\mathbf U$ is the surface obtained as the topological quotient $\mathbf P/\mathbf{E}$, where $\mathbf E$ is a minimal relation on $\mathbf{P}$ defined as follows:
\begin{gather*}
\text{Let } x_{0}=(u_{0},v_{0}),x_{1}=(u_{1},v_{1}) \in \mathbf{P}.  \text{ }\mathbf{E} \text { is given as the relation } x_{0}\mathbf{E}x_{1} \\ \text{ iff } x_{0}=x_{1}
 \text{ or, for }x_{0},x_{1} \in {\partial} \left( \left[2m-\frac{1}{2},2m+\frac{1}{2}\right] \times \left[2n-\frac{1}{2},2n+\frac{1}{2}\right] \right)\\
 \text{ and }m,n\in\mathbb{Z},
  \left[\begin{array}{c}
u_{1} -2m
\\v_{1}-2n
\end{array}\right] = \left[\begin{matrix}
0 && 1\\
1 && 0
\end{matrix}\right]
\left[ \begin{array}{c}u_{0}-2m\\
v_{0}-2n
\end{array}\right]
. \end{gather*}
\end{Def}

\begin{thm}
The Necker cube surface $\mathbf{S}$ is isometric to $\mathbf{U}$.
\begin{proof}
To show that $\Psi$ is an isometry of the two surfaces, consider $(x,y,z)\in\mathbf{S}$. $\Psi$ is a bijection between sets $\mathbf{P}$ and $\mathbf{S}$ outside of the edges of each cube parallel to the z-axis (Figure $\ref{fig:Psi}$). These edges split into two edges in $\mathbf{P}$, but are identified to be one and the same in $\mathbf{U}$.\\
For $0\leq i\leq 1$, Let $S_1=\{(m+1,n,p-i)\}, S_2=\{(m+1,n+1,p-i)\}, S_3=\{(m,n+1,p-i)\}$ be these edges.
\compav{come back to this.}
\end{proof}
\end{thm}

\noindent We use $\mathbf{U}$ from here on, since it is a far simpler surface to work with. Below are representations of simple geodesics on $\mathbf{S}$ and $\mathbf{U}$.

\begin{figure}[H]
\centering
\includesvg[width=1.8in]{cubesxyzpaths}
\includesvg[width=2.8in]{unfoldcutpath}
\caption{Periodic(c) and drift-periodic (a,b) trajectories.}
\end{figure}

We see from the figure above that a geodesic flow in what would appear to be the $\frac{\pi}{4}$ direction closes up on the surface. The edges of $\mathbf{U}$ are identified by a $\pm\frac{\pi}{2}$ rotation about a corner of one of these squares. It is a Euclidean cone surface with countably many conic singularities of angle $\frac{3\pi}{2}$ and $3\pi$ (3 singularities per hole). The set of singularities is denoted $Sing(\mathbf{U})$.

\begin{figure}[H]
\centering
\includesvg[width=1.5in.]{quotient}
\caption{A $2\times2$ cut out section centered at each missing square. Edges and verticies identified.}
\label{fig:quotient}
\end{figure}

The surface $\mathbf{U}^0=\mathbf{U}\backslash Sing(\mathbf{U})$ is the surface with these cone singularities \textbf{removed}.

\begin{lem}
The unit tangent bundle of $\mathbf{U}^0$, $UT(\mathbf{U}^0)$, is isomorphic to $\mathbf{U} \times\mathbb{R}/\frac{\pi}{2}\mathbb{Z}$.
\begin{proof}
Let $T_x(\mathbf{U}^0)$ be the unit tangent fiber at a point $x\in\mathbf{U}^0$. If $x$ is not on an identified edge around a hole of $\mathbf{P}$, then $T_x\simeq\mathbb{R}/2\pi\mathbb{Z}$. Suppose that it lies on an edge of this unit square cut-out. Any vector $v\in\mathbb{C}$ is simultaneously in the direction $Arg(z)$ and $Arg(z)+\frac{\pi}{2}$. It follows then that for unit vectors of $\mathbb{C}$ $\frac{z}{|z|}\sim(-i)^n\frac{z}{|z|}$ for $n\in\mathbb{Z}$, and $UT(\mathbf{U}^0)\simeq\mathbb{R}/\frac{\pi}{2}\mathbb{Z}$.
\end{proof}
\end{lem}

\subsection{Discussion of Results}
This section will detail how an initial trajectory angle is obtained from $\mathbf{S}$, and how it relates to the periodicity of geodesic flow as a statement of the main theorem.

\begin{Def}
Take a sufficiently small straight-line segment in $\mathbb{R}^3$, $\ell:[0,1]\times\mathbf{S}\rightarrow\mathbb{R}^3$ such that $\ell(t,s)\subset \{X_{m,n,p}: X=\mathbf{A}, \mathbf{B}, \text{or }\mathbf{C}, \text{ and } m,n,p\in\mathbb{Z}\}$ for all $t\in[0,1]$ ($\ell$ is contained in one, and only one, face of $\mathbf{S}$). The \textbf{initial trajectory angle} of a geodesic on $\mathbf{S}$ is given as $\theta=Arg(\Psi(\ell(1,s))-\Psi(\ell(0,s)))$, where Arg is the principal argument of a vector in $\mathbb{C}$. The \textbf{initial trajectory} is the vector $\dot{v}=(\cos(\theta),\sin(\theta))\in\mathbb{R}^2$.
\begin{figure}[H]
\centering
\includesvg{vectorplane}
\end{figure}
\end{Def}

\noindent We have two subsets of $\mathbb Z^2$ directions that partition the set of possible rational directions $\theta$.
\begin{Def}
Let $v$ be a unit vector of the form $\frac{1}{k}(x,y)\simeq\mathbb{R}/2\pi\mathbb{R}$ for $x,y\in\mathbb{Z}$ and $k=\sqrt{x^2+y^2}\in\mathbb{R}$.  We call $v$ an \textbf{odd-odd} vector if x and y are relatively prime integers and both odd. We denote the \textbf{set of all odd-odd slopes} $(x,y)\in\mathcal{O}$. We say that $v$ is an \textbf{even-odd} vector if x and y are relatively prime integers, and x is even if and only if y is odd. We denote the \textbf{set of all even-odd slopes} $(x,y)\in\mathcal{E}$.
\end{Def}

If $v$ is rational, that is $v=k(x,y)$ for integers x,y, then we commonly refer to the pair $(x,y)$ as the \emph{slope} of a given initial trajectory. For example, the following trajectory modeled on the surface is a closed geodesic on $\mathbf{U}$ with initial trajectory (51,31):

\begin{figure}[H]
\centering
\includegraphics[width=4in.]{closed2.png}
\caption{Geodesic flow on $\mathbf{U}$ modeled using sage-flatsurf.}
\end{figure}

\begin{Def}
A \textbf{geodesic flow} on a topological surface, $X$, is a function $\Phi^{v}_t:UT(X)\times\mathbb{R}\rightarrow UT(X)$ such that $\frac{d\Phi}{dt}=v\in UT(X)$. 
\end{Def}

\noindent Experimental evidence indicates that geodesics in the odd-odd directions are periodic, while geodesics in the even-odd directions are drift-periodic. We will prove the theorem below.

\begin{thm}{\textbf{Dynamics of Geodesic Flow on the Necker cube surface.}}\\ Obtain $\theta$ and $v$ as described in Definition 3. Denote the non-singular unit-speed geodesic flow with initial point $s\in\mathbf{U}^0$ in direction $(\theta\mod{\frac{\pi}{2}})\sim\phi\in UT(\mathbf{U}^0)$ by $\Phi^{\phi}_t:UT(\mathbf{U}^0)\times\mathbb{R}\rightarrow UT(\mathbf{U}^0)$ such that $\frac{d\Phi}{dt}=\phi\in UT(X)$  Then the following is true:
\begin{enumerate}[label=(\roman*)]
\item (Periodic) There exists a $t_0 > 0$ such that $\Phi^{\phi}_{t+t_{0}}(s)=\Phi^{\phi}_{t}(s)$ if and only if $v\in\mathcal{O}$.
\item (Drift-Periodic) There exists a $t_0 > 0$ such that $\Phi^{\phi}_{t+t_{0}}(s)= \Phi^{\phi}_{t}\circ\Phi^{\phi}_{t_0}(s)\neq\Phi^{\phi}_t(s)$ if and only if $v\in\mathcal{E}$.
\end{enumerate}
\end{thm}

\subsection{Translation Surfaces}
We recall some common definitions, theorems, and properties of translation surfaces and their $\mathbb{Z}^d$-covers.

\begin{Def}
A translation surface is a pair $M=(S,\omega)$ formed of a connected
Riemann surface X and a holomorphic 1-form $\omega$ on $S$ which is not identically zero and satisfies:
\begin{enumerate}
\item $\Sigma\subset S$ is a discrete subset of cone singularities of $S$.
\item $\omega=z^kdz$ at points in $\Sigma$ of cone angle $2(k+1)\pi$.
\item $S\backslash\Sigma$ admits a translation surface $M^0=(S\backslash\Sigma,dz)$, where dz is a flat metric on $S\backslash\Sigma$.
\end{enumerate}
\end{Def}

The set of all orientation-preserving affine automorphisms of $M$ forms the group Aff$^+(M)$. The corresponding \emph{Veech group} of M is the image of the group morphism $D:\text{Aff}^+(M)\rightarrow SL(2,\mathbb{R})$, where $D$ is the derivative map of an element of Aff$^+(M)$, and denote it by $V(M)$. A surface is said to by \emph{Veech} if its Veech group is commensurable to $SL(2,\mathbb{R})$. The unit tangent bundle on $M\backslash\Sigma$, denoted $UT(M)$, is isomorphic to $M\times\mathbb{R}/2\pi\mathbb{Z}$. Let $\phi,\psi$ be two compatible charts between $M$ and $\mathbb{C}$ such that $\phi(x_0)\in\mathbb{C}$ for $x_0\in M\backslash\Sigma$, and let $F^t_\theta:z\mapsto z+te^{i\theta}$ be a translation flow on $\mathbb{C}$. The non-singular geodesic flow on the unit tangent bundle of $M$, $G^t:UT(M)\times\mathbb{R}\rightarrow UT(M)$, is well defined on and given by $G^t(x_0,\theta)=(\psi^{-1}\circ F^t_\theta\circ\phi(x_0),\theta)$. \\
A \emph{maximal geodesic} on $M$ is a re-parameterized closed geodesic $\gamma:[0,1]\rightarrow M$ obtained from $G_t$, and we consider $\gamma\in\pi_1(M,x_0)$. A normal subgroup $N\triangleleft\pi_1(M,x_0)$ is associated with the quotient group $\Delta=\pi_1(M,x_0)/N$. $\tilde{M}$ is called the $\Delta$-cover of $M$ if $p:\tilde{M}\rightarrow M$ is a cover of $M$ and for some fixed $\tilde{x}_0\in p^{-1}(x_0)$, $\pi_1(M,x_0)/\Delta=\pi_1(\tilde{M},\tilde{x}_0)$. $\Delta$ has a unique lift to the group of Deck transformations acting on $\tilde{x}_0\in\tilde{M}$, or its orbit under Aut$(p)$. Compact translation surfaces are oriented, hence why their fundamental groups are isomorphic to $\mathbb{Z}^{2g}$ (g is the genus of $M$). Therefore $\Delta\simeq \mathbb{Z}^d$, and $\tilde{M}$ is called a $\mathbb{Z}^d$ \emph{translation cover} of $M$. Translation covers have Deck transformation groups isomorphic to $\mathbb{Z}^d$. \\
When $\gamma\in\pi_1(M,x_0)$ is a closed curve, we denote its abelianization as $[\gamma]\in H_1(M,R)$, with coefficients in unit ring $R$, and its unique lift to $\tilde{M}$ under the covering map $p$ as $\tilde{\gamma}$. Suppose you have some set of homology classes, $\{a_1,\dots,a_n\}=\Gamma\subset H_1(M,R)$, such that span$(\Gamma)=H_1(M,R).$ We call $\Gamma$ the \emph{spanning set}, and express an element $[\beta]\in H_1(M,R)$ as $[\beta]=\Sigma_{j=1}^{n}x_j a_j$, where $x_1,\dots x_n\in R$.

\begin{Def}
Denote the non-degenerate, bi-linear \textbf{intersection number} between two homology classes $[\alpha],[\beta]$ as $i([\alpha],[\beta])$, where 
\begin{align*}
i:H_1(M,R)\times H_1(M, R)\rightarrow \mathbb{Z}.
\end{align*}
returns the total number of intersections and its sign corresponds to the sign of the angle that $[\beta]$ makes relative to $[\alpha]$.
\end{Def}

We mainly concern ourselves with $\mathbb{Z}^2$ covers for the purposes of this paper. Consider the set $Q\subset N\subset \pi_1(M,x_0)$ to be the spanning set of N.

\begin{Def}
Consider two sets, $q,r\subset\Gamma$, such that span$(q,r)$ is the abelianization of $N$. There exists a group morphism $s:N\rightarrow Aut(p)$ that takes a generator of $N$ to a generator of $Aut(p)$, a translational symmetry of $\tilde{M}$ isometric to $(m,n)\in\mathbb{Z}^2$, call it $T^{m,n}$. Note that dim$(N)=2g-2$, and $s$ is an epimorphism when $g>2$. Obtain $q$ as the abelianization of elements in $Q$ that are mapped to $T^{0,\pm 1}$, the vertical translations. Likewise, take $r$ to be the set of abelianized elements of $Q$ that are mapped to  $T^{\pm 1, 0}$, the horizontal translations. A path non-trivially intersecting with a curve in q corresponds to a horizontal translation, while an intersection with r corresponds to a translation in the vertical direction. This is because an intersection of a closed path $\beta$ with an element $a \in s^{-1}(T^{0,\pm 1})$ implies that $\beta$ at the time of intersection would have had to have been homotopic to $b\in s^{-1}(T^{\pm 1,0})$ instead (assuming that $a,b$ intersect).\\
For example, let $t$ be the time of that intersection, and $t_0$ be the time of the previous one. Then $\tilde{\beta}(t)=T^{\pm 1,0}(\tilde{\beta}(t_0))$ would have moved horizontally, and not vertically.\\
Since intersection number is bi-linear and non-degenerate, we   obtain the \textbf{group homomorphism as a representation of a lifted path} on $\tilde{M}$:

\begin{align*}
\Omega_{q,r}:\pi_1(M,x_0)\rightarrow \mathbb{Z}^2 \text{; } \beta\mapsto(i(q,[\beta]),i(r,[\beta])).
\end{align*}
\end{Def}

\begin{lem}
Let $\beta\in\pi_1(M,x_0)$. Then $\beta$ lifts to $\tilde{\beta}\in\pi_1(\tilde{M},\tilde{x}_0)$ if and only if $\beta\in\text{Ker }\Omega_{q,r}$.
\begin{proof}
do this later
\end{proof}
\end{lem}

This lemma becomes an integral part of the proof of the main theorem, and allows us to show how induced homomorphisms of a surface's homology by its affine automorphisms allows for us to generalize whenever a geodesic on the base surface lifts to a closed geodesic on the cover. 
\subsection{Acknowledgements}

\section{Four-fold Cover of $\mathbf{U}$}

\begin{figure}[H]
\centering
\includesvg[width=2.05in.]{fourfold}\hspace{0.5in}\includesvg[width=2.05in.]{fourfoldrotate}
\caption{Four-fold cover isometry and the preimage of a point in $\mathbf U \backslash Sing(\mathbf U)$.}
\end{figure}

\begin{figure}[H]
\centering
\includesvg[width=3.15in]{utilda}
\label{fig:utilda0}
\caption{Branched cover associating every direction with one plane.}
\end{figure}

\begin{figure}[H]
\centering
\includesvg[width=3.15in]{utildaprime}
\caption{Infinite-type translation surface obtained by rotating each copy of the fundamental domain accordingly.}
\label{fig:utilda}
\end{figure}

The quotient under the group action of translational symmetries is isomorphic to $\mathbb{Z}^2$ since the orbit of any point in the fundamental domain is a lattice in the space.

\begin{thm}
The translational symmetries of $\tilde{\mathbf{U}}$'s fundamental domain induce symmetries on the surface isomorphic to $\mathbb{Z}^2$.
\begin{proof}
Let $(z;j)\in\mathbb{C}\times\mathbb{Z}/4\mathbb{Z}$ and define the group action $T_0^{m,n}:\mathbb{C}\times\mathbb{Z}/4\mathbb{Z}\rightarrow\mathbb{C}\times\mathbb{Z}/4\mathbb{Z}$ as $T^{m,n}(z;j)=(z+2e^{-j\frac{i\pi}{2}}(m+in);j)$. This translation acts faithfully on the preimages of $\mathbf{U}\backslash Sing(\mathbf{U})$, and respects edge identifications of $\mathbf{\tilde{\mathbf{U}}}$, thereby making it an isometry of the surface. Consider a group homomorphism, $T_0^{m,n}\mapsto m+in$ onto the plane of Gaussian integers, $\mathbb{Z}[i]$. The exponential function is never zero, so the identity of the translation group is $T_0^{0,0}$. This is an isomorphism since it is clearly surjective and any non-trivial element of $T^{m,n}$ could not possibly map to the identity element of $\mathbb{Z}[i]$, regardless of the value of j. Since $\mathbb{Z}^2$ is isomorphic to $\mathbb{Z}[i]$, it is isomorphic to  $T_0^{m,n}$ as well.
\end{proof}
\label{thm:z2}
\end{thm}

\begin{Def}
The automorphism $T^{m,n}:\tilde{\mathbf{U}}\rightarrow\tilde{\mathbf{U}}$ is an  \textbf{induced translation} of $\tilde{\mathbf{U}}$ as a result of the previous theorem.
\end{Def}

We take the quotient of $\mathbf{\tilde{\mathbf{U}}}$ by the action of the induced isometry of $T^{m,n}$, to get a translation base surface, $\mathbf{M}$.

\begin{figure}[H]
\centering
\includesvg[width=2.6in]{mtilda}
\caption{Compact translation surface, $\mathbf{M}$, covered by the infinite surface with edges and cone singularities (1,2,3,4) identified. The Roman numerals are meant to identify each quotient with a plane in the cover, $\mathbf{\tilde{\mathbf{U}}}$.}
\label{fig:mtilda}
\end{figure}

It is clear that this is a square-tiled Veech surface when realized as a finite staircase:

\begin{figure}[H]
\centering
\includesvg[width=3.4in]{mtildastaircase}
\caption{The staircase Veech surface with directional planes and vertices identified. All edges are paired by translations. Two adjacent squares have opposite edges identified. The top edge of the bottom-left square is glued to the bottom edge of the top-right square (both labeled I). Likewise, the bottom edge of the bottom-left square is identified with the top edge of the top-right square.}
\label{fig:staircase}
\end{figure}

This surface is obtained as a ramified cover of the unit square torus. It is a translation surface and is therefore equipped with a \textbf{holomorphic one-form}, a collection of charts from neighborhoods of $\mathbf{M}$ to $\mathbb{C}$ such that any neighborhood away from Sing$(\mathbf{M})$ has a \emph{flat} induced Euclidean metric. A theorem of Gutkin and Judge tells us that its Veech group is commensurable to SL$(2,\mathbb{Z})$ and is therefore a Veech surface. We look at some of its affine maps, and generate a subgroup $\mathbb{X}\subset$ Aff$^+(\mathbf{M})$  by the following transformations:

\begin{enumerate}[label=(\roman*)]
\item Multi-twists of the surface as global diffeomorphisms given by Dehn-twists of its cylinder decomposition in horizontal and vertical directions with derivatives:
\begin{equation*}
\left\{ \left[ \hspace{1mm} \begin{matrix}
				1 &  \pm 2\\
				0 & 1
			\end{matrix}\hspace{1mm}\right] \text{ , }
			\left[ \hspace{1mm} \begin{matrix}
							1 & 0\\
							 \pm 2 & 1
						\end{matrix}\hspace{1mm}\right] \right\}.
\end{equation*}
we call $\mathbf{A}^{\pm 1}$, $\mathbf{B}^{\pm 1}$, respectively.\\
A Dehn-twist on each cylinder in the cylinder decomposition of $\mathbf{M}$ in horizontal and vertical directions gives way to these global affine diffeomorphisms:
\begin{figure}[H]
\centering
\includesvg[width=1.5in]{cylinderskew}
\label{fig:skew}
\end{figure}

\item Rotation group generated by a $+\frac{\pi}{2}$ rotation of the surface fixed about the center of the second square on the bottom of the staircase, an order four isometry on $\mathbf{M}$ denoted $\mathbf{R}$.

\item  Order 2 translation of the surface that moves the bottom left-most square to the square right next to it, denoted $\mathbf{H}$.
\item Order 2 translation of the surface that takes the bottom right-most square to the one right above it, denoted $\mathbf{V}$.
\end{enumerate}

\begin{Def}
The group $\mathbb{X}$ is the isometry group generated by affine maps $\mathbf{A}$,$\mathbf{B}$,$\mathbf{R}$, $\mathbf{H}$, and $\mathbf{V}$. The image of the derivative map on elements in $\mathbb{X}$ is denoted $\mathbb{X}'$ and generated by matrices
\begin{align*}
\left[\begin{matrix}
1 & 2 \\ 0 & 1
\end{matrix}\right],
\left[\begin{matrix}
1 & 0 \\ 2 & 1
\end{matrix}\right],
\left[\begin{matrix}
0 & -1 \\ 1 & 0
\end{matrix}\right],
\end{align*}
denoted $\mathbf{A}'$, $\mathbf{B}'$, and $\mathbf{R}'$ in that order.
\end{Def}

It is not immediately apparent if these affine maps generate Aff$^{+}(\mathbf{M})$, or if their derivatives generate $V(\mathbf{M})$, its Veech group. We use these to induce homomorphisms on $H_1(X, \mathbb Q)$. A spanning set of $H_1(\mathbf{M}, \mathbb Q)$ is obtained as the set of homology classes of the core curves of X's cylinder decompositions in both vertical and horizontal directions:

\begin{figure}[H]
\includegraphics[width=3in]{homologyclass.png}
\centering
\caption{Cylinder core curves with u,v, and z homology classes that determines the $\mathbf{Z}^{2}$-cover.}
\label{fig:homology}
\end{figure}

\begin{Def}
The set of abelianized cylinder core curves is denoted as $\Gamma=\{\gamma_i: i = 0,\dots,11\}\subset H_1(\mathbf{M},\mathbb{Q})$.
\end{Def}

\begin{rem}
We use 12 elements to span homology, although a basis requires only 10. It's not impossible to determine the relations between these core curve classes, but it is not necessary. A $12\times12$ matrix of these core curve cylinder decompositions to their intersection numbers with adjacent curves is rank 10, as to be expected.
\end{rem}

The induced homomorphisms of $H_1(\mathbf{M}, \mathbb Q)$ have come from affine maps that have various effects on these core-curves. We use a 12-gon to represent the set of curves, and show how these elements act on them. The multi-twists add curves to adjacent curves, and the translation maps permute them. The reader is encouraged to check these for themselves.

e.g. for $\mathbf{H,V}\in$Aff$^+(\mathbf{M})$,

\begin{minipage}{0.5\textwidth}
\vspace{0.3in}
\begin{itemize}
\item[\textbf{\emph{$\mathbf{H}$ \& $\mathbf{V}$}}] The effect that these two translations have on the 12-gon is a reflection about these lines. Observed by keeping track of the squares and core curves after $\mathbf{H}$ and $\mathbf{V}$ have acted on $\mathbf{X}$.
\end{itemize}
\end{minipage}
\begin{minipage}{0.7\textwidth}
\begin{figure}[H]
\hspace{0.2in}\includegraphics[width=1.7in]{12gonHV.pdf}
\end{figure}
\end{minipage}

\begin{Def}
The induced homomorphisms of $H_1(\mathbf{M},\mathbb{Q})$ are obtained from the affine subgroup $\mathbb{X}$ and denoted $\mathbb{X}_*$. The associated homomorphisms on the spanning set $\Gamma$ are given as:
\begin{align*}
\mathbf{A}^k_*\circ[\gamma_i]=&[\gamma_i] + \frac{k}{2}(1-(-1)^i)([\gamma_{i-1}]+[\gamma_{i+1}])\\
\mathbf{B}^k_*\circ[\gamma_i]=&[\gamma_i] + \frac{k}{2}(1+(-1)^i)([\gamma_{i-1}]+[\gamma_{i+1}])\\
\mathbf{R}_*\circ[\gamma_i]=&(-1)^i[\gamma_{1-i\text{ mod }12}]\\
\mathbf{H}_*\circ[\gamma_i]=&[\gamma_{12-i\text{ mod }12}]\\
\mathbf{V}_*\circ[\gamma_i]=&[\gamma_{10-i\text{ mod }12}]
\end{align*}
\end{Def}

\begin{Def}
The homology classes $u,v,z$ are given as the following sums of core curves:
\begin{align*}
[u] &= -[\gamma_2] +[\gamma_5] + [\gamma_8] - [\gamma_{11}],\\
[v] &= +[\gamma_0] -[\gamma_3] -[\gamma_6] +[\gamma_9],\\
[z] &= +[\gamma_1] +[\gamma_4]-[\gamma_7]-[\gamma_{10}].
\end{align*}
\end{Def}

\begin{thm}
The fundamental group of the $\mathbb{Z}^2$-cover is obtained by lifting the kernel of the closed paths of $\mathbf{M}$ of the homomorphism:

\begin{align*}
\Omega_{u,v}:\pi_1(\mathbf{M},x_0)\rightarrow \mathbb{Z}^2 \text{; } \beta\mapsto(i(u,[\beta]),i(v,[\beta])),\text{ where}
\end{align*}

\begin{align*}
i:H_1(\mathbf{M},\mathbb Q)\times H_1(\mathbf{M},\mathbb Q)\rightarrow \mathbb Z.
\end{align*}is the intersection number of two homology classes.
\begin{proof}
We know from Theorem $\ref{thm:z2}$ that the translational symmetries of $\tilde{\mathbf{U}}$ induced by $T^{m,n}$ is isometric to $\mathbb{Z}^2$. Since $\mathbf{M}$ is a genus 5 base surface, we know that $\pi_1(\mathbf{M},x_0)\simeq \mathbb{Z}^{10}$, and the associated cover satisfies $\mathbf{M}=\tilde{\mathbf{U}}/(\pi_1(\mathbf{M},x_0)/N)$, such that $N$ is a normal subgroup of $\pi_1(\mathbf{M},x_0)$. This means that $N\simeq \mathbb{Z}^8$. The eight core curve classes are the abelianized forms of  $\gamma_0,\gamma_2,\gamma_3,\gamma_5,\gamma_6,\gamma_8,\gamma_9,$ and $\gamma_{11}$ that span N. The classes and their signs are obtained from Figure $\ref{fig:mtilda}$ as the outer regions identified by the translations of $T^{m,n}$. Thus any closed path on $\mathbf{M}$ is lifted to a closed path on the cover under the quotient map only when a path has a trivial intersection number with the classes.
\end{proof}

\end{thm}
Two paths are homologous if they return the same intersection number with the classes of closed core cylinder curves of $\mathbf{U}$ that span its homology. The classes u and v are obtained from the group group action of $T^{m,n}$ on the cover.

\begin{Def}
$\mathbf{hol}:\mathbf{M}\backslash\text{Sing}(\mathbf{M})\rightarrow\mathbb{C}$ is the holonomy vector pulled back from a non-singular path $\gamma$ in $\mathbf{M}$ onto the complex plane given by $\mathbf{hol}(\gamma)=\int_{\gamma}dz$.
\end{Def}

We denote the \textbf{closed path} $\alpha$, such that $\mathbf{hol}(\alpha)=6+6i$, and show it is homologous to the closed geodesic with the same holonomy vector. The slope one direction also decomposes $\mathbf{M}$ into two cylinders by a series of saddle connections of length $\sqrt{2}$ between singularities:

\begin{figure}[H]
\centering
\includesvg[width=4.4in.]{slopeonecylinder}
\caption{The two right-most cylinders $C_1$ (labeled) and $C_2$(unlabeled).}
\end{figure}

The circumferences of these two cylinders are $6\sqrt{2}$. Geodesic flows on this surface are well defined, and rational directions

\begin{Def}
Let $\omega^{\theta}_t:[0,1]\times\mathbb{R}/2\pi\mathbb{Z}\rightarrow\mathbf{M}$ be the \textbf{maximal geodesic flow} on the surface in direction $\theta$ such that $\omega^{\theta}_0=\omega^{\theta}_1=x_0\in\mathbf{M}\backslash\text{Sing}(\mathbf{M})$.\\
$\omega^{\frac{\pi}{4}}_t$ is the geodesic flow in the \textbf{slope one direction}, and $\chi$ is its image in $\mathbf{M}$ and element of $\pi_1(\mathbf{M},x_0)$.
\end{Def}

\begin{lem}
$\alpha$ is homologous to $\chi$ in  $\mathbf{M}\backslash\text{Sing}(\mathbf{M})$.
\begin{proof}
Let $\chi$ be a geodesic contained in either $C_1$ or $C_2$. Since a geodesic does not admit singularities, it is the image of a closed path on $X\backslash\text{Sing}(X)$ with initial point $x_0$ on the strips of $C_1$ and $C_2$ with boundaries removed, denoted $C_1',$ $C_2'$. Express $[\alpha]$ as $\Sigma^{11}_{j=0}$ $\frac{1}{2}\gamma_j$ (a closed path climbing up the staircase). We show that the intersection numbers of $[\alpha]$ and $[\chi]$ are the same for every core cylinder curve $\gamma$, i.e. $i([\gamma_k], \Sigma^{11}_{j=0} \text{ } \frac{1}{2}[\gamma_j])=i([\gamma_k],[\chi])\text{ } \forall k=0,\dots,11$.\\
\textbf{Case one:} k is even. If k is even, then every curve $\gamma_k$ is oriented to the right. Since $\chi$ intersects every curve once, $i([\gamma_k],[\chi])=1$. No even indexed curves intersect eachother, so we need only consider when $j$ is odd. Now if $j$ is odd, it is incident (positively crossing) with only two horizontal curves, namely $\gamma_{j+1},\gamma_{j-1}$. Therefore $i([\gamma_k],[\alpha])=i([\gamma_{j-1}],\frac{1}{2}[\gamma_{j}])+i([\gamma_{j+1}],\frac{1}{2}[\gamma_{j}])=\frac{1}{2}(i([\gamma_{j-1}],[\gamma_{j}])+i([\gamma_{j+1}],[\gamma_{j}]))=\frac{1}{2}(1+1)=1.$\\
\textbf{Case two:} k is odd. If k is odd, then $[\chi]$ will have an intersection number of $-1$ with $[\gamma_k]$ since odd-indexed core curves are oriented upwards. Now since k is odd, we only consider when $j$ is even. Similarly, this means that $\gamma_j$ negatively intersects the two vertical core curves with adjacent indices. Hence, $i([\gamma_k],[\alpha])=i([\gamma_{j-1}],\frac{1}{2}[\gamma_{j}])+i([\gamma_{j+1}],\frac{1}{2}[\gamma_{j}])=\frac{1}{2}(i([\gamma_{j-1}],[\gamma_{j}])+i([\gamma_{j+1}],[\gamma_{j}]))=\frac{1}{2}(-1-1)=-1.$\\
We know intersection number to be bilinear and non-degenerate on homology. So if $\alpha$ and $\chi$'s abelianizations admit the same intersection numbers for every curve in the spanning set of $H_1(\mathbf{M}, \mathbb{Q})$, then $[\alpha]=[\chi]$.
\end{proof}
\end{lem}

\begin{thm}
$\chi\in\pi_1(\mathbf{M},x_0)$ lifts to $\tilde{\chi}\in\pi_1(\tilde{\mathbf{U}},\tilde{x}_0)$
\begin{proof}
From Lemma 2, $[\chi]=[\alpha]$, so $\Omega_{u,v}(\chi)=\Omega_{u,v}(\alpha)$. Since
$i([u],[\alpha])=-i([\gamma_2],[\alpha] +i([\gamma_5],[\alpha]) + i([\gamma_8],[\alpha]) - i([\gamma_{11}],[\alpha])=-1+(-1)+1-(-1)=0$ and $i([v],[\alpha])=1-(-1)-1+(-1)=0$, it follows that $\alpha,\chi\in\text{Ker }\Omega_{u,v}$, and $\chi$ lifts to a closed geodesic on $\tilde{\mathbf{U}}$.
\end{proof}
\end{thm}

\begin{cor}
content...
\end{cor}

From here, we use $\alpha$ to show that the \emph{only} trajectories that close on the Necker cube surface are those that are in vector direction $(a,b)$ such that $\gcd(a,b)=1$ and $a,b$ are both odd. We call these \textbf{odd-odd} directions. We can make this claim because the group generated by the matrices

\begin{equation*}
\left[ \hspace{1mm} \begin{matrix}
				1 &  \pm 2\\
				0 & 1
			\end{matrix}\hspace{1mm}\right] \text{ and }
			\left[ \hspace{1mm} \begin{matrix}
							1 & 0\\
							 \pm 2 & 1
						\end{matrix}\hspace{1mm}\right]
\end{equation*}
is the \emph{Sanov subgroup} of $SL(2,\mathbb{Z})$ and only sends elements in the odd-odd set to itself. There are dualizations made between how these matrices skew a geodesic direction,  and how their original affine transformations induce an effect homology. In a sense the kernel is obtained by the orbit of $\chi$ under $\mathbb{X}$ and its holonomy vector under $\mathbb{X}'$.

\begin{lem}
The actions of $\mathbb{X}'$ on $\mathcal{O}$ and $\mathcal{E}$ are closed in their respective sets.
\begin{proof}
Since $\mathbb{X}'$ is generated by the elements $\mathbf{A}'$, $\mathbf{B}'$, and $\mathbf{R}'$, any matrix $G'\in\mathbb{X}'$ is of the form $G' = (\mathbf{A}')^{1_1}\circ(\mathbf{B}')^{1_2}\circ(\mathbf{R}')^{1_3}\circ(\mathbf{A}')^{2_1}\circ\dots(\mathbf{A}')^{n_1}\circ(\mathbf{B}')^{n_2}\circ(\mathbf{R}')^{n_3}$, where $i_k\in\mathbb{Z}$ for $i=1,\dots,n$ and $k=1,2,3.$ Let $x=\left(\begin{matrix}p \\ q  \end{matrix}\right),y\in\mathcal{O}$, and consider the equation $G'x=y$. Observe that $\left(\begin{matrix}1 && 2 \\ 0 && 1\end{matrix}\right)^l x=\left(\begin{matrix}p+2jq \\ q  \end{matrix}\right),$ and $ \left(\begin{matrix}1 && 0 \\ 2 && 1\end{matrix}\right)^m x=\left(\begin{matrix}p \\ q+2mp  \end{matrix}\right)$ for any $l,m\in\mathbb{Z}$. Also note that for any $j\in\mathbb{Z}$, $(\mathbf{R}')^{m}x=\left(\begin{matrix}p \\ q  \end{matrix}\right),\left(\begin{matrix}-q\\ p  \end{matrix}\right),\left(\begin{matrix}-p \\ -q  \end{matrix}\right),\left(\begin{matrix}q \\ -p  \end{matrix}\right)$ when $j \mod{4}\equiv0,1,2,3$, respectively. In any case, the product of any power of a generator of $\mathbb{X}'$ and any $x\in\mathcal{O}$ is an element of $\mathcal{O}$ . By letting $l=i_1,m=i_2,$ and $j=i_3$, we first consider the base case when $i=n$. Let $G'=G'_1\circ\dots\circ G'_n$, such that $G'_i=(\mathbf{A}')^{i_1}\circ(\mathbf{B}')^{i_2}\circ(\mathbf{R}')^{i_3}$. Since $n_1,n_2,n_3$ are arbitrary integers, $G'_n x \in\mathcal{O}$. Suppose for some $b< n-1$, $G'_{n-b}\circ\dots\circ G'_n x =y'\in\mathcal{O}$. Therefore $y'=(G'_1\circ\dots\circ G'_{b})^{-1}y$, which implies that $(G'_1\circ\dots\circ G'_{b})^{-1}$ preserves the set $\mathcal{O}$. Otherwise, if $y\in\mathcal{E}$, there exists at least one $G'_i$ for $1< i< b$ and $\tau\in\mathcal{E}$ such that $G'^{-1}_i \tau = (\mathbf{R}')^{-i_3}\circ(\mathbf{B}')^{-i_2}\circ(\mathbf{A}')^{-i_1} \tau \in \mathcal{O}$, a contradiction. Since elements in $\mathbb{X}'$ are invertible, $G'_1\circ\dots\circ G'_{b}$ must also map $\mathcal{O}$ to itself. Left multiply both sides of the equation to show that $G'_1\circ\dots\circ G'_n x = G' x = y$. By the principle of strong induction, this holds for all $0<b\leq n$. Since $G'$ is invertible and an arbitrarily chosen element of $\mathbb{X}'$, it follows that $x\in\mathcal{O}$ if and only if $y\in\mathcal{O}$ and $\mathcal{O}$ is closed under $\mathbb{X}'$.\\
The proof for when $x\in\mathcal{E}$ is made in the same way.
\end{proof}
\end{lem}

Now a trajectory in the horizontal direction has a directional vector of (1,0). The orbit of this vector by the Veech group is the set of all \textbf{even-odd} vectors. We also know that in this direction a geodesic is drift-periodic (See figure 1). The Veech group of $\mathbf{M}$ preserves these properties. Suppose you had some closed geodesic on $\mathbf{M}\backslash Sing(\mathbf{M})$ called $\beta$ such that $\beta=h(\alpha)$, where $h\in$  Aff$^+(\mathbf{M})$, and $h_*$ is its induced homomorphism. Then we want to show that

\begin{align*}
(i([\beta],[u]), i([\beta],[v]))=(i([\alpha],h^{-1}_*[u]), i([\alpha],h^{-1}_*[v])) = (0,0).
\end{align*}

But first, we look at some of the properties of the group $\mathbb{X}_*$.

\begin{thm}
Let $\mathbb{X}_*$ be the group generated by $\mathbf{A}_*, \mathbf{B}_*, \mathbf{R}_*, \mathbf{H}_*,$ and $\mathbf{V}_*$. Let $G=\left< \mathbf{A}_*, \mathbf{B}_* \right>$, $T=\left< \mathbf{H}_*, \mathbf{V}_* \right>$, and $R=\left< \mathbf{R}_*\right>$. Then the following is true:
\begin{enumerate}[label=(\roman*)]
\item $G$ is a free subgroup of $\mathbb{X}_*$ of rank two.
\item $T$ is a finite cyclic subgroup of $\mathbb{X}_*$ and a centralizer of G.
\item $R$ is a finite cyclic subgroup of $\mathbb{X}_*$, and a normalizer of $G$.
\end{enumerate}
\begin{proof} Let $h_*^{j}=\mathbf{A}_*^{k_j}\circ \mathbf{B}_*^{g_j}\in G$ for $k_j, g_j \in\mathbb{Z}$, $j=1,\dots,n$.\\ \emph{(i)}. When $\mathbf{A}_*$ and $\mathbf{B}_*$ act on $\gamma_i$, it is only ever trivial if i is even for $\mathbf{A}_*$ or i is odd on $\mathbf{B}_*$. Since i cannot be both odd and even at the same time, there is no relation between the two generators and therefore $G$ is free.\\
\emph{(ii)} It is up to the reader to show that $T$ has the relations $\mathbf{H}_*^2=\mathbf{V}_*^2=(\mathbf{H}_*\mathbf{V}_*)^3=id_*$, and is isomorphic to the rotational group of the hexagon generated by reflections about adjacent vertices of a 12-gon. Observe that $\mathbf{H}_*\circ \mathbf{A}_*^{k_j}\circ[\gamma_i]=\mathbf{H}_*\circ[\gamma_i] + \frac{k_j}{2}(1-(-1)^i)(\mathbf{H}_*\circ[\gamma_{i-1}]+\mathbf{H}_*\circ[\gamma_{i+1}])=[\gamma_{-i}] + \frac{k_j}{2}(1-(-1)^i)([\gamma_{1-i}]+[\gamma_{-i-1}])=\mathbf{A}^{k_j}\circ[\gamma_-i]=\mathbf{A}^{k_j}\circ\mathbf{H}_*\circ[\gamma_i],$ and $\mathbf{V}_*\circ \mathbf{A}_*^{k_j}\circ[\gamma_i]=\mathbf{V}_*\circ[\gamma_i] + \frac{k_j}{2}(1-(-1)^i)(\mathbf{V}_*\circ[\gamma_{i-1}]+\mathbf{V}_*\circ[\gamma_{i+1}])=[\gamma_{10-i}] + \frac{k_j}{2}(1-(-1)^i)([\gamma_{11-i}]+[\gamma_{9-i}])=\mathbf{A}^{k_j}\circ[\gamma_10-i]=\mathbf{A}^{k_j}\circ\mathbf{V}_*\circ[\gamma_i].$ In the same way one can show this to be true for $\mathbf{B}_*^{g_j}$, and we can see that $T$ is a centralizer of $G$. \\
\emph{(iii)} $R$ is obviously cyclic and finite since an isomorphism is obtained as $\mathbf{R}_*\mapsto\mathbf{R}'\in SO(2,\mathbb{Z})$. \\Note that $\mathbf{R}_*\circ\mathbf{A}_*^{k_j}\circ[\gamma_i]=\mathbf{R}_*\circ[\gamma_i] + \frac{k_j}{2}(1-(-1)^i)(\mathbf{R}_*\circ[\gamma_{i-1}]+\mathbf{R}_*\circ[\gamma_{i+1}])\\
=(-1)^i[\gamma_{1-i}] + \frac{k_j}{2}(1-(-1)^i)((-1)^{i-1}[\gamma_{2-i}]+(-1)^{i+1}[\gamma_{-i}])\\
=(-1)^{1-i}([\gamma_{1-i}] - \frac{k_j}{2}(1+(-1)^{1-i})([\gamma_{2-i}]+[\gamma_{-i}]))\\
=(-1)^{1-i}\mathbf{B}_*^{-k_j}\circ[\gamma_{1-i}]=\mathbf{B}_*^{-k_j}\circ(-1)^{1-i}[\gamma_{1-i}]=\mathbf{B}_*^{-k_j}\circ\mathbf{R}_*\circ[\gamma_{i}].$\\
Likewise, $\mathbf{R}_*\circ\mathbf{B}_*^{g_j}\circ[\gamma_{i}]=\mathbf{A}_*^{-g_j}\circ\mathbf{R}_*\circ[\gamma_{i}]$.
\end{proof}
\end{thm}

\begin{rem}
It can be easily shown that $\mathbb{X}'$ has similar properties.
\end{rem}

\begin{lem}
$Let$ $h_*\in \left<\mathbf{A}_*,\mathbf{B}_*\right>. $ Then $h_*\circ[\alpha]\in H_1(\mathbf{M},\mathbb{Q})$ can be expressed as $h_*\circ[\alpha]=\frac{1}{2}(c_1\Sigma^{5}_{j=0}[\gamma_{2j}]+c_2\Sigma^{5}_{j=0}[\gamma_{2j+1}])$ for $c_1,c_2\in\mathbb{Z}$.
\begin{proof}
Let $\Sigma^{5}_{j=0}[\gamma_{2j}]=\Sigma\Gamma_{even}$, $\Sigma^{5}_{j=0}[\gamma_{2j+1}]= \Sigma\Gamma_{odd}$, and  $\Sigma^{11}_{j=0}[\gamma_{j}]= \Sigma\Gamma$. Let $h_*=h_*^n\circ\dots\circ h_ *^1$, and $h_*^{i}=\mathbf{A}_*^{k_i}\circ \mathbf{B}_*^{g_i}$ for $k_i, g_i \in\mathbb{Z}$, $i=1,\dots,n$. Compose these two homomorphisms and obtain $\mathbf{A}_*^{k_i}\circ \mathbf{B}_*^{g_i}(\Sigma\Gamma)=(4g_ik_i+2k_i)\Sigma\Gamma_{even}+2g_i\Sigma\Gamma_{odd}+\Sigma\Gamma$. Let $c_i^1=(4g_ik_i+2k_i), c_i^2=2g_i$, and solve for $h_*^{i+1}\circ h_*^i\circ\Sigma\Gamma$:
\begin{align*}
h_*^{i+1}\circ h_*^{i}\circ(\Sigma\Gamma)=h_*^{i+1}\circ(c_i^1\Sigma\Gamma_{even}+c_i^2\Sigma\Gamma_{odd}+\Sigma\Gamma)\\ =c_i^{1}h_*^{i+1}\circ(\Sigma\Gamma_{even})+c_i^2h_*^{i+1}\circ(\Sigma\Gamma_{odd})+h_*^{i+1}\circ(\Sigma\Gamma)\\ =2g_{i+1}\Sigma\Gamma_{odd}+(4g_{i+1}k_{i+1}+2k_{i+1})\Sigma\Gamma_{even}+\Sigma\Gamma\\+c_i^{1}(4g_{i+1}k_{i+1}\Sigma\Gamma_{even}+2g_{i+1}\Sigma\Gamma_{odd}+\Sigma\Gamma_{even})\\+c_{i}^2(2k_{i+1}\Sigma\Gamma_{even}+\Sigma\Gamma_{odd})\\
=\Sigma\Gamma+(c_i^1+(c_{i}^1+1)(4g_{i+1}k_{i+1})+(c_i^2+1)2k_{i+1})\Sigma\Gamma_{even}\\+(c_i^2+(c_i^1+1)2g_{i+1})\Sigma\Gamma_{odd}\\
\text{Let }c_{i+1}^1:=(c_i^1+(c_{i}^1+1)(4g_{i+1}k_{i+1})+(c_i^2+1)2k_{i+1}),\\
c_{i+1}^2:=(c_i^2+(c_i^1+1)2g_{i+1}).
\end{align*}
From these recursive definitions and a finite sequence of integers, $\{k\}_i,\{g\}_i$, observe then that\\ $h_*\circ[\alpha]=h_*\circ[\frac{1}{2}\Sigma\Gamma]=\frac{1}{2}h_*\circ[\Sigma\Gamma]=\frac{1}{2}[c_n^1\Sigma\Gamma_{even}+c_n^2\Sigma\Gamma_{odd}+\Sigma\Gamma]\\
=\frac{1}{2}[(c_n^1+1)\Sigma\Gamma_{even}+(c_n^2+1)\Sigma\Gamma_{odd}]$. Further simplify by letting $c_1=c_n^1+1,c_2=c_n^2+1$.
\end{proof}
\end{lem}

\begin{lem}
Let $h_*\circ[\alpha]\in H_1(\mathbf{M},\mathbb{Q})$. Then for $a\in\left<\mathbf{H}_*, \mathbf{V}_*\right>$ and $b\in\left<\mathbf{R}_*\right>$, the following is true:
\begin{align*}
& a\circ h_*\circ[\alpha]=h_*\circ[\alpha]\\
& b\circ h_*\circ[\alpha]=\frac{1}{2}[c'_1\Sigma\Gamma_{even}+c'_2\Sigma\Gamma_{odd}]\\
& h_*\circ b\circ[\alpha]=\frac{1}{2}[c''_1\Sigma\Gamma_{even}+c''_2\Sigma\Gamma_{odd}]
\end{align*}
\begin{proof}
By Theorem 4, a is a centralizer of the group so $a\circ h_*\circ[\alpha]=h_*\circ a \circ[\alpha]=h_*\circ \frac{1}{2}a \circ[\Sigma\Gamma].$ Since a is a cyclic permutation of the set $\Gamma$, it acts trivially on $\Sigma\Gamma$. Therefore, $a\circ h_*\circ[\alpha]=h_*\circ \frac{1}{2}[\Sigma\Gamma]=h_* \circ[\alpha]$.\\
By theorem 4, $a\circ\mathbf{A}_*^{k_i}\circ\mathbf{B}_*^{g_i}=\mathbf{B}_*^{-k_i}\circ\mathbf{A}_*^{-g_i}\circ a$. Extend this property to $h_*$, and denote the normalized element as $h_{**}$, such that $b\circ h_{*}=h_{**}\circ b$.  Note that $b(\Sigma\Gamma)=b(\Sigma\Gamma_{even}+\Sigma\Gamma_{odd})=\Sigma\Gamma_{odd}-\Sigma\Gamma_{even}$. $b\circ h_*\circ[\Sigma\Gamma]=c_1b\circ\Sigma\Gamma_{even}+c_2b\circ\Sigma\Gamma_{odd}=c_1\Sigma\Gamma_{odd}-c_2\Sigma\Gamma_{even}.$ So, $c_1'=-c_2$ and $c_2'=c_1$.\\
Since $h_*$ is arbitrary, let $h_{**}=g_{*}$ be generated by an integer sequence that defines the word and consider $h_*\circ b\circ[\Sigma\Gamma]=b\circ g_{*}\circ[\Sigma\Gamma]=c^*_1b\circ\Sigma\Gamma_{even}+c^*_2b\circ\Sigma\Gamma_{odd}=c^*_1\Sigma\Gamma_{odd}-c_2^*\Sigma\Gamma_{even}.$ So, $c_1''=-c_2^*$ and $c_2''=c_1^*$.
\end{proof}
\end{lem}

\noindent Now that every element in the orbit of $[\alpha]$ can be expressed as a linear combination of integers, it is simple to show they lift to a closed trajectory in the cover.

\begin{Def}
Let $\mathbf{dir}:UT(\mathbf{M}\backslash\text{Sing}(\mathbf{M}))\rightarrow\mathcal{O}\cup\mathcal{E}$ be the injective map from $\mathbb{R}/2\pi\mathbb{Z}$ to $\mathbb{Z}^2$ given as $\mathbf{dir}(\theta)=(k_1\cos(\theta),k_2\sin(\theta)), k_1,k_2\in\mathbb{R}$ such that $\gcd(k_1\cos(\theta),k_2\sin(\theta))=1$.
\end{Def}

\begin{thm}(Sketch)\\
Any geodesic, $\beta$, in $\mathbf{M}$ lifts to a closed geodesic $\tilde{\beta}$ on $\tilde{\mathbf{U}}$ if and only if $\mathbf{dir}(Arg(\mathbf{hol}(\beta)))\in\mathcal{O}$.
\begin{proof}
Call the quotient cover $p:\tilde{\mathbf{U}}\rightarrow\mathbf{M}$, and fix a point $\tilde{x_0}\in p^{-1}(x_0)$. Let $\beta=h(\chi)$, where $h\in\mathbb{X}$. We also obtain $[\beta]=h_*\circ[\alpha]$ from Lemma 2. Since $h$ sends geodesics to geodesics, h induces the following: $\mathbf{hol}(h(\chi))=h'(\mathbf{hol}(\chi))=h'(6+6i)$ for $h'=\left[\begin{matrix} a & b \\ c & d\end{matrix}\right]\in\mathbb{X}'$. So, Arg$(h'(\mathbf{hol}(\chi)))=$Arg$(6[(a+b)+i(c+d)])=$Arg$(6h'(1+i))$. Lemma 3 states that for any $h'\in\mathbb{X}'$, $h'(\mathcal{O})=\mathcal{O}$. Therefore there is no such geodesic of \textbf{even-odd} slope in the orbit of $\chi$. Otherwise $h',h\notin\mathbb{X}',\mathbb{X}$. Consequently, $\mathbf{dir}(Arg(\mathbf{hol}(\beta)))\in\mathcal{O}$.\\
From Lemma 5 we see that $[h(\chi)]=h_*\circ[\chi]=h_*\circ[\alpha]=\frac{1}{2}(c_1\Sigma^{5}_{j=0}[\gamma_{2j}]+c_2\Sigma^{5}_{j=0}[\gamma_{2j+1}])$ for $c_1,c_2\in\mathbb{Z}$. Denote the sums as $\Sigma\Gamma_{even}$ and $\Sigma\Gamma_{odd}.$\\
Therefore, $2i([u],h_*\circ[\alpha])=c_1i([u],\Sigma\Gamma_{even})+c_2i([u],\Sigma\Gamma_{odd})\\
=c_1(-i([\gamma_2],0) +i([\gamma_5],[\gamma_6]+[\gamma_4]) + i([\gamma_8],0) - i([\gamma_{11}],[\gamma_{10}]+[\gamma_{0}]))\\
+c_2(-i([\gamma_2],[\gamma_1]+[\gamma_3]) +i([\gamma_5],0) + i([\gamma_8],[\gamma_7]+[\gamma_9]) - i([\gamma_{11}],0))\\
=c_1(-(0)+(-1-1)+(0)-(-1-1))+c_2(-(1+1)+(0)+(1+1)-(0))=0.$\\\vspace{0.1in}\\
Similarly, $2i([v],h_*\circ[\alpha])=c_1i([v],\Sigma\Gamma_{even})+c_2i([v],\Sigma\Gamma_{odd})\\
=c_1(i([\gamma_0],0)-i([\gamma_3],[\gamma_2]+[\gamma_4])-i([\gamma_6],0)+i([\gamma_9],[\gamma_8]+[\gamma_{10}])\\
+c_2(i([\gamma_0],[\gamma_{11}]+[\gamma_1])-i([\gamma_3],0)-i([\gamma_6],[\gamma_5]+[\gamma_7])+i([\gamma_9],0)\\
=c_1((0)-(-2)-(0)+(-2))+c_2((2)-(0)-(2)+(0))=0.$\\
Therefore, $\Omega_{u,v}(h(\chi))=(0,0)$, and $h(\chi)=\beta\in\text{Ker }\Omega_{u,v}$ for all $h\in\mathbb{X}$. By Theorem 2, $\beta$ lifts to $\tilde{\beta}\in\pi_1(\tilde{\mathbf{U}},\tilde{x}_0)$. Let $\theta=$Arg$(\mathbf{hol}(\beta))$. Then $\omega_t^{\theta}$ at $x_0$ lifts to $\tilde{\omega}_t^{p^{-1}(\theta)}\in\tilde{\mathbf{U}}\backslash\text{Sing}(\tilde{\mathbf{U}})$.\\
Now suppose instead that $\beta=h(\gamma_{i})$. Then $\mathbf{dir}(\beta)=\frac{1}{2}(1+(-1)^i,1-(-1)^i)$. According to Lemma 3, $h'(\mathcal{E})=\mathcal{E}$. Thus we have no geodesic in the \textbf{odd-odd} directions obtained from the orbits of $(1,0)$ and $(0,1)$. For contradiction, suppose that $h(\gamma_{i})\in\textbf{Ker }\Omega{u,v}$. Then $(i(h_*\circ[\gamma_i], [u]),i(h_*\circ[\gamma_i], [v]))=(i([\gamma_i], h^{-1}_*\circ[u]),i([\gamma_i], h^{-1}_*\circ[v]))=(0,0).$ Let $h^{-1}_*\circ[u]=\Sigma^{11}_{j=0}x_j[\gamma_j]$, and $h^{-1}_*\circ[v]=\Sigma^{11}_{j=0}y_j[\gamma_j]$. Note that since $\gamma_i$ intersects $\gamma_{i\pm1}$, $i([\gamma_i], h^{-1}_*\circ[u])=(-1)^{i+1}(x_{i-1}+x_{i+1})$ and $i([\gamma_i], h^{-1}_*\circ[v])=(-1)^{i+1}(y_{i-1}+y_{i+1}).$
\\
$\mathbf{Unfinished.}.$
\end{proof}
\end{thm}

\begin{conj}{\textbf{Dynamics of Geodesic Flow on the Necker cube surface.}}\\ Obtain $\theta$ and $\vec{\theta}$ as described in Definition 3. Denote the non-singular unit-speed geodesic flow with initial point $s\in(\mathbf{U}\backslash Sing(U))$ in direction $[\theta]\sim\phi\in UT(\mathbf{U}\backslash Sing(\mathbf U))$ by $F_t^{\phi}:\mathbf{U}\times\mathbb{R}_0^+\rightarrow\mathbf{U}$ on $(\mathbf{U},\mu)$, where $\mu$ is a flow-invariant measure.  Then the following is true:
\begin{enumerate}[label=(\roman*)]
\item (Periodic) There exists a $t_0 > 0$ such that $F^{\phi}_{t+t_{0}}(s)=F^{\phi}_{t}(s)$ if and only if $\vec{\theta}\in\mathcal{O}$.
\item (Drift-Periodic) There exists a $t_0 > 0$ such that $F^{\phi}_{t+t_{0}}(s)= F^{\phi}_{t}(s)+c$, where $c\in \mathbf{U}$ is a non-trivial translation of a point in $\mathbf{U}$, if and only if $\vec{\theta}\in\mathcal{E}$.
\end{enumerate}
\begin{proof}(Sketch)\\
Denote the covering maps $f:\tilde{\mathbf{U}}\rightarrow\mathbf{U}$, $p:\tilde{\mathbf{U}}\rightarrow\mathbf{M}$, and fix a point $\tilde{x}_0\in f^{-1}(s),p^{-1}(x_0),$ for $x_0\in\mathbf{M}$. $f^{-1}([\theta])=\{x:x=\theta+n\frac{\pi}{2}, n\in\mathbb{Z}\}=[\theta]\subset UT(\tilde{\mathbf{U}}\backslash\text{Sing}(\tilde{\mathbf{U}}))$ given by the four-fold cover and rotations of each individual plane. This gives us a relation between the two tangent bundles, where the translation four-fold cover has the standard $\mathbb{R}/2\pi\mathbb{Z}$ unit tangent fiber. $\theta'$ is the direction associated to the flow $\omega_t^{\theta'}:[0,1]\rightarrow\mathbf{M}$ on the translation surface. Since the cover is translation, $p^{-1}(\theta')=\theta'=\theta+n\frac{\pi}{2}$.  First suppose that $\vec{\theta}\in\mathcal{O}$. Then $\theta$ is identified with the set of directions that close on $\tilde{\mathbf{U}}$. From Theorem 5, $\omega_t^{\theta'}$ lifts to a closed geodesic $\tilde{\omega}_t^{\theta'}$. Given $\mathbf{hol}(\omega)=\int_{\omega}dz$, we obtain a period for the unit-speed flow, $t_0=|\mathbf{hol}(\omega)|$. That is, $\tilde{F}_t^{\theta'}:\mathbb{R}^+_0\rightarrow\tilde{\mathbf{U}}$ such that  $\frac{d}{dt}\tilde{F}_t=\frac{1}{|\mathbf{hol}(\omega)|}$. Then $F_t^{\phi}=F_t^{[\theta']}=f\circ \tilde{F}^{\theta'}_t$. The period carries over since there is no concern over a trajectory returning to $\tilde{x_0}$ in a different direction. Otherwise, the geodesic $\omega_t^{\theta'}$ on $\mathbf{M}$ would have closed in $0<t<1$. Now suppose that $\vec{\theta}\in\mathcal{E}$. Identifying it with $\theta'$, we see that $\omega$ in direction $\theta'$ is not an element of $\textbf{Ker }\Omega_{u,v}$ from Theorem 5. Therefore, $\Omega_{u,v}(\omega)=(m,n)\simeq T^{m,n}$ and lifting the terminal point $\omega(1)$, $\tilde{\omega}(1)=T^{m,n}(\tilde{\omega}(0))=T^{m,n}(\tilde{x}_0).$ The period remains unchanged, in that $\tilde{F}^{\theta'}_{t+\mathbf{hol}(\omega)}=\tilde{F}^{\theta'}_{t}+T^{m,n}(\tilde{x}_0)$. Therefore, $F^{\phi}_{t+t_{0}}(s)=f\circ\tilde{F}^{\theta'}_{t}(\tilde{x}_0)+f\circ T^{m,n}(\tilde{x}_0).$\\
Conversely, suppose $F_t$ is periodic. Then $[\theta]=\phi=[\theta']$, which defines directional flows $\tilde{F}^{\phi}_t$. According to Theorem 5, $\tilde{F}^{\phi}_t$ will close if and only if $\phi\subset\mathcal{O}$. $\phi$ is the orbit of $\vec{\theta}'$ under the 90 degree rotational matrix. This matrix does not alter the length or period of a geodesic. Thus, $F^{\phi}_t$ is exactly one of the flows $\tilde{F}^{\phi}_t$. Likewise, if $F^\phi_t$ is drift-periodic then $F^\phi_{t+t_0}=f\circ \tilde{F}^\phi_t+f\circ T^{m,n}$. $T^{m,n}$ is trivial if and only if $\theta'\in\mathcal{O}$. Therefore, $\theta'\in\mathcal{E}$, and $[\theta']=\phi$.
\end{proof}
\end{conj}

There is still much work to do in terms of cleaning up the proofs and organizing the final paper.

\newpage
\section*{Conclusion}
What I ultimately aim to do is port these results on X's homology back to the Necker cube surface. I want do it in such a way that the final theorem is bi-conditional. To do so, I imagine I can take a vector image of a small segment of a geodesic in $\mathbb{R}^3$ and project it onto the isometric flattening of the Necker Cube surface to obtain a direction (or classes of equivalent directions), and relate it to the unit tangent bundle of $\mathbf{M}$.

In addition, I would also like to find a formula for the arc-length of a geodesic based on direction alone. Knowing that hol$(\alpha)=6+6i$ means that the induced Euclidean metric on $\mathbf{M}\backslash$Sing($\mathbf{M}$) gives the geodesic an arc-length of $6\sqrt{2}$. I would like to show that:

\begin{align*}
\int_{\beta}|dz|=|hol(h(\alpha))|=|h'(hol(\alpha))|,
\end{align*}
\noindent where $h'\in V(\mathbf{M})$ is the derivative of h, and $\beta=h(\alpha)$. We know that is true on the translation surface, but it's a matter of then showing the translation quotient, branch-cover, and the Necker cube surface have the same induced Euclidean metric of these non-singular geodesics. (It would not be surprising considering that the surface is built out of subsets of planes.) Even more of a problem is finding a way to solve for a matrix in the Sanov subgroup that brings (1,1) to the desired odd-odd slope.

Even though the main theorem remains unproven for now, it is an enjoyable property of the surface and venturing in its understanding involves exploring some fascinating niches in algebraic topology and dynamical systems.

Needless to say, it has been a very busy summer for me. I am grateful for the opportunity to really take some time to delve deep into numerous topics that I have been very interested in for a while. I sincerely doubt it would have been possible without the summer internship, and I have the Rich scholarship committee to thank for that. I would also like to extend my gratitude to Dr. Hooper. He has been incredibly patient and kind throughout these last few months, and I would not have come this far in so short a time without his help.

\newpage
\begin{thebibliography}{9}
\bibitem{Robotics} Fred G. Martin \emph{Robotics Explorations: A Hands-On Introduction to Engineering}. New Jersey: Prentice Hall.
\bibitem{Flueck}  Flueck, Alexander J. 2005. \emph{ECE 100}[online]. Chicago: Illinois Institute of Technology, Electrical and Computer Engineering Department, 2005 [cited 30
August 2005]. Available from World Wide Web: (http://www.ece.iit.edu/~flueck/ece100).
\end{thebibliography}
\end{document}

### 2017-08-28 paper.tex
082717

@@
 \newtheorem*{rem}{Remark}
+\newtheorem*{conj}{Conjecture}
 
@@
 \begin{Def} Identify $\mathbb C$ with $\mathbb{R}^2$ in the usual way.
-The closed path $\alpha\in\pi_1(\mathbf M, x_0)$ is the path contained in the maximal cylinder decomposition in direction $(1,1)$.
+The closed path $\alpha\in\pi_1(\mathbf M, x_0)$ is the path contained in the maximal cylinder decomposition of $\mathbf M$ in direction $(1,1)$ with hol$(\alpha)=6+6i$.
 \end{Def}
@@
 \begin{lem}
-$\alpha$ is homotopic to the image of the closed geodesic in direction $(1,1)$ on $\mathbf{M}\backslash\text{Sing}(\mathbf M)$.
+$\alpha$ is homologous to the image of the closed geodesic in direction $(1,1)$ on $\mathbf{M}\backslash\text{Sing}(\mathbf M)$.
 \begin{proof}
@@
 $i([\gamma_k], \Sigma^{11}_{j=0} \text{ } \frac{1}{2}[\gamma_j])=i([\gamma_k],[\chi])\text{ } \forall k=0,\dots,11$.
-And since $x_0$ is in either $C_1'$ or $C_2'$, $\alpha$ cannot climb up the staircase without intersecting the core curves.
-Thus, the sum of $\frac{1}{2}$ core curves of $[\alpha]$ have to have to intersect at least one other core curve.
 \\
@@
 Hence, $i([\gamma_k],[\alpha])=i([\gamma_{j-1}],\frac{1}{2}[\gamma_{j}])+i([\gamma_{j+1}],\frac{1}{2}[\gamma_{j}])=\frac{1}{2}(i([\gamma_{j-1}],[\gamma_{j}])+i([\gamma_{j+1}],[\gamma_{j}]))=\frac{1}{2}(-1-1)=-1.$\\
-We know intersection number to be a non-degenerate bilinear form, and if $\alpha$ and $\chi$'s abelianizations admit the same intersection numbers for every curve in the basis set of $H_1(\textbf M, \mathbb{Q})$, then $\alpha\sim\chi$.
+We know intersection number to be non-degenerate, and since $\alpha$ and $\chi$'s abelianizations admit the same intersection numbers for every curve in the spanning set of $H_1(\textbf M, \mathbb{Q})$,  $[\alpha]=[\chi]$.
 \end{proof}
@@
 \centering
-\includesvg[width=1.8in]{cubesxyzpaths}
-\includesvg[width=2.8in]{unfoldcutpath}
-\end{figure}
-
-\begin{figure}[H]
-\centering
 \includegraphics{z2cover.pdf}
@@
 
+\begin{conj}{\textbf{Periodicity of Geodesic Flow on the Necker cube surface.}}\\ Let $\theta$ be a direction obtained described as in Definition 3.
+Denote the geodesic flow on $\mathbf{S}$ by $F^t_{[\theta]}:T^1_{\bar{\Psi}(s)}(\mathbf{U}\backslash Sing(\mathbf U))\times\mathbb{R}_0^+\rightarrow\mathbf{S}$, where $F^t_{[\theta]}=\bar{\Psi}^{-1}(\bar{\Psi}(s)+t[\theta])$ for initial point $s\in\mathbf{S}$.
+Then the following is true:
+\begin{enumerate}[label=(\roman*)]
+\item If $\theta$ is rational and identified with an element in $\mathcal{O}$, then $F^t_{[\theta]}$ is periodic.
+\item If $\theta$ is rational and identified with an element in $\mathcal{E}$, then $F^t_{[\theta]}$ is drift-periodic.
+\end{enumerate}
+\end{conj}
+
+\newpage
+\begin{lem}
+The group $\mathbb{X}'$ acting on $\mathbb{Z}^2$ sends $\mathcal{O}$ to $\mathcal{O}$, and $\mathcal{E}$ to $\mathcal{E}$.
+\begin{proof}
+Since $\mathbb{X}'$ is generated by the elements $\mathbf{A}'$, $\mathbf{B}'$, and $\mathbf{R}'$, any matrix $G'\in\mathbb{X}'$ is of the form $G' = (\mathbf{A}')^{1_1}\circ(\mathbf{B}')^{1_2}\circ(\mathbf{R}')^{1_3}\circ(\mathbf{A}')^{2_1}\circ\dots(\mathbf{A}')^{n_1}\circ(\mathbf{B}')^{n_2}\circ(\mathbf{R}')^{n_3}$, where $i_k\in\mathbb{Z}$ for $i=1,\dots,n$ and $k=1,2,3.$ Let $x=\left(\begin{matrix}p \\ q  \end{matrix}\right),y\in\mathcal{O}$, and consider the equation $G'x=y$.
+Observe that $\left(\begin{matrix}1 && 2 \\ 0 && 1\end{matrix}\right)^l x=\left(\begin{matrix}p+2jq \\ q  \end{matrix}\right),$ and $ \left(\begin{matrix}1 && 0 \\ 2 && 1\end{matrix}\right)^m x=\left(\begin{matrix}p \\ q+2mp  \end{matrix}\right)$ for any $l,m\in\mathbb{Z}$.
+Also note that for any $j\in\mathbb{Z}$, $(\mathbf{R}')^{m}x=\left(\begin{matrix}p \\ q  \end{matrix}\right),\left(\begin{matrix}-q\\ p  \end{matrix}\right),\left(\begin{matrix}-p \\ -q  \end{matrix}\right),\left(\begin{matrix}q \\ -p  \end{matrix}\right)$ when $j \mod{4}\equiv0,1,2,3$, respectively.
+In any case, the product of any power of a generator of $\mathbb{X}'$ and any $x\in\mathcal{O}$ is an element of $\mathcal{O}$ .
+By letting $l=i_1,m=i_2,$ and $j=i_3$, we first consider the base case when $i=n$.
+Let $G'=G'_1\circ\dots\circ G'_n$, such that $G'_i=(\mathbf{A}')^{i_1}\circ(\mathbf{B}')^{i_2}\circ(\mathbf{R}')^{i_3}$.
+Since $n_1,n_2,n_3$ are arbitrary integers, $G'_n x \in\mathcal{O}$.
+Suppose for some $b< n-1$, $G'_{n-b}\circ\dots\circ G'_n x =y'\in\mathcal{O}$.
+Therefore $y'=(G'_1\circ\dots\circ G'_{b})^{-1}y$, which implies that $(G'_1\circ\dots\circ G'_{b})^{-1}$ preserves the set $\mathcal{O}$.
+Otherwise, if $y\in\mathcal{E}$, there exists at least one $G'_i$ for $1< i< b$ and $\tau\in\mathcal{E}$ such that $G'^{-1}_i \tau = (\mathbf{R}')^{-i_3}\circ(\mathbf{B}')^{-i_2}\circ(\mathbf{A}')^{-i_1} \tau \in \mathcal{O}$, a contradiction.
+Since elements in $\mathbb{X}'$ are invertible, $G'_1\circ\dots\circ G'_{b}$ must also preserve $\mathcal{O}$.
+By inductive reasoning, we conclude that  $G'_1\circ\dots\circ G'_n x = G' x = y$.
+$x\in\mathcal{O}$ if and only if $y\in\mathcal{O}$ since $G'$ is invertible.\\
+Proving that $G'(\mathcal{E})=\mathcal{E}$ is done in the same way.
+\end{proof}
+\end{lem}
+
 \end{document}

### 2017-09-13 newpaper.tex
091217

@@
 \usepackage[utf8]{inputenc}
+
+%%%%% Alphabets %%%%% 
+\def\cA{\mathcal{A}}\def\cB{\mathcal{B}}\def\cC{\mathcal{C}}\def\cD{\mathcal{D}}\def\cE{\mathcal{E}}\def\cF{\mathcal{F}}\def\cG{\mathcal{G}}\def\cH{\mathcal{H}}\def\cI{\mathcal{I}}\def\cJ{\mathcal{J}}\def\cK{\mathcal{K}}\def\cL{\mathcal{L}}\def\cM{\mathcal{M}}\def\cN{\mathcal{N}}\def\cO{\mathcal{O}}\def\cP{\mathcal{P}}\def\cQ{\mathcal{Q}}\def\cR{\mathcal{R}}\def\cS{\mathcal{S}}\def\cT{\mathcal{T}}\def\cU{\mathcal{U}}\def\cV{\mathcal{V}}\def\cW{\mathcal{W}}\def\cX{\mathcal{X}}\def\cY{\mathcal{Y}}\def\cZ{\mathcal{Z}}
+
+\def\AA{\mathbb{A}} \def\BB{\mathbb{B}} \def\CC{\mathbb{C}} \def\DD{\mathbb{D}} \def\EE{\mathbb{E}} \def\FF{\mathbb{F}} \def\GG{\mathbb{G}} \def\HH{\mathbb{H}} \def\II{\mathbb{I}} \def\JJ{\mathbb{J}} \def\KK{\mathbb{K}} \def\LL{\mathbb{L}} \def\MM{\mathbb{M}} \def\NN{\mathbb{N}} \def\OO{\mathbb{O}} \def\PP{\mathbb{P}} \def\QQ{\mathbb{Q}} \def\RR{\mathbb{R}} \def\SS{\mathbb{S}} \def\TT{\mathbb{T}} \def\UU{\mathbb{U}} \def\VV{\mathbb{V}} \def\WW{\mathbb{W}} \def\XX{\mathbb{X}} \def\YY{\mathbb{Y}} \def\ZZ{\mathbb{Z}}
+
+\def\fa{\mathfrak{a}} \def\fb{\mathfrak{b}} \def\fc{\mathfrak{c}} \def\fd{\mathfrak{d}} \def\fe{\mathfrak{e}} \def\ff{\mathfrak{f}} \def\fg{\mathfrak{g}} \def\fh{\mathfrak{h}} \def\fj{\mathfrak{j}} \def\fk{\mathfrak{k}} \def\fl{\mathfrak{l}} \def\fm{\mathfrak{m}} \def\fn{\mathfrak{n}} \def\fo{\mathfrak{o}} \def\fp{\mathfrak{p}} \def\fq{\mathfrak{q}} \def\fr{\mathfrak{r}} \def\fs{\mathfrak{s}} \def\ft{\mathfrak{t}} \def\fu{\mathfrak{u}} \def\fv{\mathfrak{v}} \def\fw{\mathfrak{w}} \def\fx{\mathfrak{x}} \def\fy{\mathfrak{y}} \def\fz{\mathfrak{z}}
+\def\fgl{\mathfrak{gl}}  \def\fsl{\mathfrak{sl}}  \def\fso{\mathfrak{so}}  \def\fsp{\mathfrak{sp}}  
+\def\GL{\mathrm{GL}} \def\SL{\mathrm{SL}}  \def\SP{\mathrm{SL}}
+
+\def\<{\langle} \def\>{\rangle}
+\def\ad{\mathrm{ad}} 
+\def\Aut{\mathrm{Aut}}
+\def\dim{\mathrm{dim}} 
+\def\End{\mathrm{End}} 
+\def\ev{\mathrm{ev}} 
+\def\half{\hbox{$\frac12$}}
+\def\Hom{\mathrm{Hom}} 
+\def\qtr{\mathrm{qtr}} 
+\def\tr{\mathrm{tr}} 
+\def\Tr{\mathrm{Tr}} 
+\def\vep{\varepsilon}
+
+\def\ZZn{\ZZ/n\ZZ}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 % Title Page
-\title{Summer Report 2017}
-\author{Pavel Javornik}
 
@@
 \begin{abstract}
+The Necker cube surface is an infinite topological surface obtained by the periodic edge-to-edge tiling of unit cubes in $\RR^3$ along a 2-dimensional subspace realized as a Euclidean cone surface with a discrete, countable set of singularities of cone angles $3\pi$ and $\frac{3\pi}{2}$.
+Its various symmetries allow for the categorization of periodic and drift-periodic straight line trajectories in terms of their initial trajectory angles.
+Results are obtained from studying the Teichmuller geodesic flow on a branched four-fold cover of the surface, where the group of deck transformations of the cover preserves the trajectory angle properties, which we use to determine periodicity on what is an infinite translation surface.
+The initial rational angle is obtained by the ratio of relatively prime integer vectors of the form $(a,b)$.
+When $a,b$ are both odd the flow is periodic, and whenever $a$ or $b$ is even, the flow is drift-periodic.
+\\
+\end{abstract}
+
+\newpage
+\section{Introduction}
+
+\compav{ignore:}
 The Necker cube surface, $\mathbf{S}$, is an infinite topological surface obtained by the periodic tiling of unit cubes by $\mathbb{R}^3$ translational symmetries along a 2-dimensional subspace realized as a Euclidean cone surface with a discrete set of singularities of cone angle $3\pi$ and $\frac{3\pi}{2}$.
@@
 This paper's aim is to study the relations between geodesic flows on base surfaces, Deck transformations of their covers, the orbit of rational directions in $\mathbb{Z}^2$ under Veech group action, and use group theoretic results to prove theorems about rational trajectories on the Necker cube surface.
-\end{abstract}
-
-\newpage
-\section{Introduction}
-
-Put your Problem statement here! Example of a Citation\cite[p.219]{Robotics}. Here's Another Citation\cite{Flueck}
+\compav{ignore$\uparrow$}\\
+Put your Problem statement here! Example of a Citation\cite[p.219]{Robotics}. Here's Another Citation \cite{Flueck}
 
@@
 The artist famous for his use of optical illusions in his work, M.C.
-Escher, has occasionally used tilings of the Necker cube such as in the lithographs ``Metamorphosis I," and ``Convex and Concave." (Pictured below.) The crystallographer Louis Albert Necker was credited for having extensively studied the Necker cube's geometry[citation] and remarked on this simple, yet fascinating, optical illusion that appears to ``simultaneously protrude from and intrude into the page."[citation:wolfram] Some might even recognize it as the same surface in which ``Q*bert" hops around on in the 1982 arcade game.
-[cite?]
+Escher, has occasionally used tilings of the Necker cube such as in the lithographs ``Metamorphosis I," and ``Convex and Concave." (Pictured below.) The crystallographer Louis Albert Necker was credited for having extensively studied the Necker cube's geometry[citation] and remarked on this simple, yet fascinating, optical illusion that appears to ``simultaneously protrude from and intrude into the page."[citation:wolfram]\compav{Find new citation} Some might even recognize it as the same surface in which ``Q*bert" hops around on in the 1982 arcade game published by Gottlieb.
 
@@
 This paper assumes the reader is familiar with fundamental concepts in algebraic topology such as covering space theory, induced homomorphisms on surface homologies of their symmetries, as well as some background in Veech structures of compact translation surfaces.
+\compav{cite something}
 \newpage
@@
 
-\noindent Take the following compact subsets of auxiliary planes:
+\noindent Take the following unit squares in $\RR^3$:
 
@@
 
-The structure is then \emph{flattened} onto x-y plane by piece-wise isometric maps onto a subset of $\mathbb C$ with unit squares removed at every even integer pair in the complex plane:
-
+The structure can be \emph{flattened} onto plane by piecewise isometric maps onto a subset of $\mathbb C$ with unit squares removed at every even integer pair in the complex plane: 
 \begin{figure}[H]
@@
 \raisebox{0.4in}{\includesvg[width=1.5in]{unfoldcut}}
-\caption{A piecewise-linear map from $\mathbf{S}$ to $\mathbb{C}$. }
+\caption{A piecewise isometry from $\mathbf{S}$ to $\mathbb{C}$. }
 \label{fig:Psi}
 \end{figure}
-\noindent The red/blue dotted lines are meant to show how the edge identifications are preserved after the transformation.
-We call the flattened version of the Necker cube surface $\mathbf U$.
-The map from one fundamental domain to the other is composed of piecewise-linear maps, $\Psi:\mathbf{S}\rightarrow\mathbb{C}$.
-We obtain an induced isometry between $\mathbf{S}$ and $\mathbf{U}$ from $\Psi$.
-Identify $\mathbb{C}$ with $\mathbb{R}^2$, and $\Psi$ is given as:
+\noindent The red/blue dotted lines represent the edges that are split on the plane. The map from one surface to the other is composed of piecewise isometries, $\Psi:\mathbf{S}\rightarrow\RR^3$.
 
@@
 
-\noindent The fundamental domain of $\mathbf{U}$ is
+The flattened surface is contained entirely in $\{(x,y,z)\in\RR~|~z=0\}$, which is isometric to $\CC$. $\mathbf{S}$ is recovered as a topological quotient on the domain
 
@@
 \label{eq:P}
-\mathbf{P} = \mathbb{C}\text{ }\backslash\bigcup_{m,n \in {\mathbb Z}} \big{\{} u+vi:u\in(2m-\frac{1}{2},2m+\frac{1}{2}),\text{ } v\in(2n-\frac{1}{2},2n+\frac{1}{2})\big{\}},
+\mathbf{P} = \mathbb{C}\text{ }\backslash\bigcup_{m,n \in {\mathbb Z}} \big{\{} u+vi:u\in(2m-\frac{1}{2},2m+\frac{1}{2}),\text{ } v\in(2n-\frac{1}{2},2n+\frac{1}{2})\big{\}}.
 \end{equation}
 
-\noindent and it is defined as follows (identify $\mathbf{P}$ with $\mathbb{R}^2$):
+\noindent We denote the relation by $\sim_{\mathbf{P}}$, and the isometric
 
-\begin{Def} $\mathbf U$ is the surface obtained as the topological quotient $\mathbf P/\mathbf{E}$, where $\mathbf E$ is a minimal relation on $\mathbf{P}$ defined as follows:
+\begin{Def} $\mathbf U$ is the surface obtained as the topological quotient $\mathbf{P}/\sim_{\mathbf{P}}$, where $\sim_{\mathbf{P}}$ is a minimal relation on $\mathbf{P}$ defined as follows:
 \begin{gather*}
-\text{Let } x_{0}=(u_{0},v_{0}),x_{1}=(u_{1},v_{1}) \in \mathbf{P}.  \text{ }\mathbf{E} \text { is given as the relation } x_{0}\mathbf{E}x_{1} \\ \text{ iff } x_{0}=x_{1}
- \text{ or, for }x_{0},x_{1} \in {\partial} \left( \left[2m-\frac{1}{2},2m+\frac{1}{2}\right] \times \left[2n-\frac{1}{2},2n+\frac{1}{2}\right] \right)\\
- \text{ and }m,n\in\mathbb{Z},
+\text{Let } x_{0}=(u_{0},v_{0}),x_{1}=(u_{1},v_{1}) \in \mathbf{P}.  \text{ }\sim_{\mathbf{P}} \text { is given as the relation }\\ x_{0}\sim_{\mathbf{P}}x_{1}  \text{ iff } x_{0}=x_{1}
+ \\\text{ or, for some }m,n\in\ZZ~x_{0},x_{1} \in {\partial} \left( \left[2m-\frac{1}{2},2m+\frac{1}{2}\right] \times \left[2n-\frac{1}{2},2n+\frac{1}{2}\right] \right)\\
   \left[\begin{array}{c}
@@
 
-\noindent We use $\mathbf{U}$ from here on, since it is a far simpler surface to work with. Below are representations of simple geodesics on $\mathbf{S}$ and $\mathbf{U}$.
-
 \begin{figure}[H]
@@
 \includesvg[width=2.8in]{unfoldcutpath}
-\caption{Periodic(c) and drift-periodic (a,b) trajectories.}
+\caption{Simple periodic (c) and drift-periodic (a,b) trajectories represented as line segments on $\mathbf{S}$ and $\mathbf{U}$.}
 \end{figure}
@@
 
-\begin{lem}
-The unit tangent bundle of $\mathbf{U}^0$, $UT(\mathbf{U}^0)$, is isomorphic to $\mathbf{U} \times\mathbb{R}/\frac{\pi}{2}\mathbb{Z}$.
-\begin{proof}
-Let $T_x(\mathbf{U}^0)$ be the unit tangent fiber at a point $x\in\mathbf{U}^0$.
-If $x$ is not on an identified edge around a hole of $\mathbf{P}$, then $T_x\simeq\mathbb{R}/2\pi\mathbb{Z}$.
-Suppose that it lies on an edge of this unit square cut-out.
-Any vector $v\in\mathbb{C}$ is simultaneously in the direction $Arg(z)$ and $Arg(z)+\frac{\pi}{2}$.
-It follows then that for unit vectors of $\mathbb{C}$ $\frac{z}{|z|}\sim(-i)^n\frac{z}{|z|}$ for $n\in\mathbb{Z}$, and $UT(\mathbf{U}^0)\simeq\mathbb{R}/\frac{\pi}{2}\mathbb{Z}$.
-\end{proof}
-\end{lem}
-
 \subsection{Discussion of Results}
@@
 \begin{Def}
-Take a sufficiently small straight-line segment in $\mathbb{R}^3$, $\ell:[0,1]\times\mathbf{S}\rightarrow\mathbb{R}^3$ such that $\ell(t,s)\subset \{X_{m,n,p}: X=\mathbf{A}, \mathbf{B}, \text{or }\mathbf{C}, \text{ and } m,n,p\in\mathbb{Z}\}$ for all $t\in[0,1]$ ($\ell$ is contained in one, and only one, face of $\mathbf{S}$).
+Take a trivial straight-line path in $S\subset\mathbb{R}^3$, $\ell:[0,1]\rightarrow\mathbf{S}$ such that for some $m,n,p\in\mathbb{Z},~ \ell(t)$ is contained in exactly one of $\mathbf{A}_{m,n,p},\mathbf{B}_{m,n,p},~or~\mathbf{C}_{m,n,p}$ for all $t\in[0,1]$.
 The \textbf{initial trajectory angle} of a geodesic on $\mathbf{S}$ is given as $\theta=Arg(\Psi(\ell(1,s))-\Psi(\ell(0,s)))$, where Arg is the principal argument of a vector in $\mathbb{C}$.
-The \textbf{initial trajectory} is the vector $\dot{v}=(\cos(\theta),\sin(\theta))\in\mathbb{R}^2$.
+The \textbf{initial trajectory} is the vector $\dot{v}=(\cos(\theta),\sin(\theta))\in\mathbb{R}^2$.\\
+\compav{Not sure if I should proceed with derivative or map}
 \begin{figure}[H]
@@
 \begin{Def}
-Let $v$ be a unit vector of the form $\frac{1}{k}(x,y)\simeq\mathbb{R}/2\pi\mathbb{R}$ for $x,y\in\mathbb{Z}$ and $k=\sqrt{x^2+y^2}\in\mathbb{R}$.
+Let $v$ be a unit vector of the form $\frac{1}{k}(x,y)\simeq\mathbb{R}/2\pi\mathbb{Z}$ for $x,y\in\mathbb{Z}$ and $k=\sqrt{x^2+y^2}\in\mathbb{R}$.
 We call $v$ an \textbf{odd-odd} vector if x and y are relatively prime integers and both odd.
-We denote the \textbf{set of all odd-odd slopes} $(x,y)\in\mathcal{O}$.
+We denote the \textbf{set of all odd-odd directions} $(x,y)\in\mathcal{O}$.
 We say that $v$ is an \textbf{even-odd} vector if x and y are relatively prime integers, and x is even if and only if y is odd.
-We denote the \textbf{set of all even-odd slopes} $(x,y)\in\mathcal{E}$.
+We denote the \textbf{set of all even-odd directions} $(x,y)\in\mathcal{E}$.
 \end{Def}

### 2017-10-02 newpaper.tex
100117

@@
 \usepackage[utf8]{inputenc}
+\usepackage{wrapfig}
 
@@
 \def\AA{\mathbb{A}} \def\BB{\mathbb{B}} \def\CC{\mathbb{C}} \def\DD{\mathbb{D}} \def\EE{\mathbb{E}} \def\FF{\mathbb{F}} \def\GG{\mathbb{G}} \def\HH{\mathbb{H}} \def\II{\mathbb{I}} \def\JJ{\mathbb{J}} \def\KK{\mathbb{K}} \def\LL{\mathbb{L}} \def\MM{\mathbb{M}} \def\NN{\mathbb{N}} \def\OO{\mathbb{O}} \def\PP{\mathbb{P}} \def\QQ{\mathbb{Q}} \def\RR{\mathbb{R}} \def\SS{\mathbb{S}} \def\TT{\mathbb{T}} \def\UU{\mathbb{U}} \def\VV{\mathbb{V}} \def\WW{\mathbb{W}} \def\XX{\mathbb{X}} \def\YY{\mathbb{Y}} \def\ZZ{\mathbb{Z}}
+
+\def\bU{\mathbf{U}} \def\btU{\tilde{\bU}} \def\bUs{\bU^\circ}
+\def\btUos{\btU_0^\circ}
+\def\bG{\mathbf{G}} \def\bGs{\mathbf{G}^\circ}
 
@@
 
-\title{Dynamical Properties of Geodesic Flows on the Necker Cube Surface}
+\title{Periodicity of Geodesic Flows on the Necker Cube Surface}
 \author{Pavel Javornik}
@@
 \maketitle
-
-\begin{figure}[H]
-\centering
-\includesvg[width=2.2in]{cubesxyzpaths2}
-\end{figure}
 
@@
 \begin{abstract}
-The Necker cube surface is an infinite topological surface obtained by the periodic edge-to-edge tiling of unit cubes in $\RR^3$ along a 2-dimensional subspace realized as a Euclidean cone surface with a discrete, countable set of singularities of cone angles $3\pi$ and $\frac{3\pi}{2}$.
-Its various symmetries allow for the categorization of periodic and drift-periodic straight line trajectories in terms of their initial trajectory angles.
-Results are obtained from studying the Teichmuller geodesic flow on a branched four-fold cover of the surface, where the group of deck transformations of the cover preserves the trajectory angle properties, which we use to determine periodicity on what is an infinite translation surface.
-The initial rational angle is obtained by the ratio of relatively prime integer vectors of the form $(a,b)$.
-When $a,b$ are both odd the flow is periodic, and whenever $a$ or $b$ is even, the flow is drift-periodic.
+The Necker cube surface is an infinite topological surface obtained by the periodic tiling of unit cubes in $\RR^3$ along a 2-dimensional subspace realized as a Euclidean cone surface with a discrete, countable set of singularities of cone angles $3\pi$ and $\frac{3\pi}{2}$.
+When punctured at these points it is an infinite genus flat surface equipped with a flat metric with nontrivial holonomy.
+We employ techniques most commonly used in studying geodesic flow on finite flat surfaces and adapt them to this one.
 \\
 \end{abstract}
+\begin{figure}[H]
+\centering
+\includesvg[width=2.2in]{cubesxyzpaths2}
+\label{fig:front}
+\caption{Periodic and drift-periodic flows on the Necker cube surface}
+\end{figure}
 
@@
 
-\compav{ignore:}
-The Necker cube surface, $\mathbf{S}$, is an infinite topological surface obtained by the periodic tiling of unit cubes by $\mathbb{R}^3$ translational symmetries along a 2-dimensional subspace realized as a Euclidean cone surface with a discrete set of singularities of cone angle $3\pi$ and $\frac{3\pi}{2}$.
-An isometric ``flattening" of $\mathbf{S}$ is obtained by injective piecewise-linear maps onto the planar subspace.
-The fundamental domain of the surface admits infinitely unit square \emph{holes} centered on the  $(2\mathbb{Z})^2$ lattice in $\mathbb{C}$.
-The flat surface, $\mathbf{U}$, has rotational and translation symmetries and a unit tangent bundle, $UT(\mathbf{U})$, that is isomorphic to $\mathbb{Z}/\frac{\pi}{2}\mathbb{Z}$ as a consequence of the edge identifications.
-A \emph{flat} branched cover, $\tilde{\mathbf{U}}$, of the surface is obtained by a quotient topology defined on four copies of $\mathbf{U}$'s fundamental domain and indexed by elements in $\mathbb{Z}/4\mathbb{Z}$ for the purpose of aligning geodesic flows.
-It is a \emph{formally étale} cover of $\mathbf{U}$ whose unit tangent bundle is $\mathbb{Z}/2\pi\mathbb{Z}$ and obtained by the monodromy group of the covering map.
-$\tilde{\mathbf{U}}$ is a translation surface whose quotient by its translational symmetries is a \emph{Veech} surface whose \emph{Veech group} is $SL(2,\mathbb{Z})$-commensurable.
-This paper's aim is to study the relations between geodesic flows on base surfaces, Deck transformations of their covers, the orbit of rational directions in $\mathbb{Z}^2$ under Veech group action, and use group theoretic results to prove theorems about rational trajectories on the Necker cube surface.
-\compav{ignore$\uparrow$}\\
 Put your Problem statement here! Example of a Citation\cite[p.219]{Robotics}. Here's Another Citation \cite{Flueck}
@@
 The Necker cube[citation] has made numerous appearances throughout history.
-The artist famous for his use of optical illusions in his work, M.C.
-Escher, has occasionally used tilings of the Necker cube such as in the lithographs ``Metamorphosis I," and ``Convex and Concave." (Pictured below.) The crystallographer Louis Albert Necker was credited for having extensively studied the Necker cube's geometry[citation] and remarked on this simple, yet fascinating, optical illusion that appears to ``simultaneously protrude from and intrude into the page."[citation:wolfram]\compav{Find new citation} Some might even recognize it as the same surface in which ``Q*bert" hops around on in the 1982 arcade game published by Gottlieb.
+For example, M.C.
+Escher, the artist most famously known for his use of optical illusions in his artwork, has occasionally used tilings of the Necker cube such as in the lithographs ``Metamorphosis I," and ``Convex and Concave." (Pictured below.) The crystallographer Louis Albert Necker was credited for having extensively studied the lone Necker cube's geometry[citation] and remarked on this simple, optical illusion.
+Some might even recognize it as the same surface in which ``Q*bert" hops around on in the 1982 arcade game published by Gottlieb.
 
@@
 
-This paper assumes the reader is familiar with fundamental concepts in algebraic topology such as covering space theory, induced homomorphisms on surface homologies of their symmetries, as well as some background in Veech structures of compact translation surfaces.
-\compav{cite something}
+\begin{rem}
+Familiarity is assumed on the part of the reader with covering space theory, translation (Veech) surfaces, and their associated Veech groups.
+For a general take on these topics check [cite] and [cite].
+\end{rem}
+
+\subsection{Discussion of Results}
+Initial experiments suggested there was a correlation between choice of trajectory angle and periodicity.
+On a surface composed of infinitely many cubes, there is certainly a flat metric endowed on smaller neighborhoods of the surface that are locally isometric to the plane.
+Via parallel transports of tangent vectors over the edges of the cubes connecting one face to the next, it is possible to describe geodesic behavior on the Necker cube surface as a series of straight line segments that traverse faces and continuously flow in a fixed direction (see figure $\ref{fig:front}$).
+\\\\
+
+\begin{wrapfigure}{l}{2.2in.}
+\includesvg{vectorplane}
+\end{wrapfigure}
+We consider classes of directions obtained from unit vectors in $\RR^3$ by projecting one of these line segments onto a plane parallel to some face.
+These vectors retain all necessary information about their directions and let us consider rational and irrational angles relative to a choice of basis in $\RR^2$.
+This is what we refer to as an \emph{initial trajectory angle}.
+\\\\
+Rational angle trajectories are categorized according to the parities of their components.
+\begin{Def}
+Let $v$ be a unit vector of the form $\frac{1}{k}(x,y)\simeq\mathbb{R}/2\pi\mathbb{Z}$ for $x,y\in\mathbb{Z}$ and $k=\sqrt{x^2+y^2}\in\mathbb{R}$.
+We call $v$ an \textbf{odd-odd} vector if x and y are relatively prime integers and both odd.
+We denote the \textbf{set of all odd-odd directions} $(x,y)\in\mathcal{O}$.
+We say that $v$ is an \textbf{even-odd} vector if x and y are relatively prime integers, and x is even if and only if y is odd.
+We denote the \textbf{set of all even-odd directions} $(x,y)\in\mathcal{E}$.
+\end{Def}
+If $v$ does not fall into one of these two categories, it is an \emph{irrational} direction. We prove the following is true of all rational trajectory angles:
+
+\begin{thm}
+Let $\Phi_t$ be the unit tangent flow on the Necker cube surface, and $\Phi_0=(x_0,v')$ be its initial point in direction $v'\in\RR^3$.
+Let $v$ be a unit tangent vector in $\RR^2$ identified with $v'$.
+Then the following is true:
+\begin{enumerate}[label=(\roman*)]
+\item $\Phi$ is periodic with period $T\in\RR$ if and only if $v\in\cO$.
+\item $\Phi$ is drift-periodic with period $T\in\RR$ if and only if $v\in\cE$.
+\end{enumerate}
+\end{thm}
+
+The proof of this theorem is in section X.
+\\
+\compav{theorem about most directions being recurrent?}
+
+\subsection{Acknowledgements}
+-Pat Hooper\\
+- Vincent Delecroix, Ferrán Valdez, pascal hubert\\
+- \\
+
 \newpage
 
-\subsection{Background}
-We will introduce some objects, definitions, and theorems to be used later in the paper.
-First and foremost, we need to describe the construction of the Necker cube surface and give it a concrete structure as a subset of $\mathbb{R}^3$.
+\section{Construction}
+This section details the construction of the Necker cube surface and how two different interpretations of the surface are one and the same.
+\subsection{Geometric Interpretation}
+Consider the following surface embedded in $\RR^3$:
+\begin{figure}[H]
+\centering
+\includesvg[width=2in.]{neckercube}
+\end{figure}
 
-\noindent Take the following unit squares in $\RR^3$:
+We call this surface $\bG$, the \emph{Necker cube}.
+$\bG$ is obtained by taking three copies of the unit square and identifying the edges so as to resemble a cube when viewed from the front.
+$\bG$ has a total of three vertices.
+Every vertex has a cone angle of either $3\pi$ or $\frac{3\pi}{2}$ and is therefore a conical point/singularity.
+It is apparent that $\bG$ has an induced flat metric and non-trivial linear holonomy, so geodesic flows on its unit tangent bundle are well-defined via parallel transports over the edges.
+We denote the surface without singularities by $\bGs$.
+The periodic tiling of $\bG$ will take infinitely many copies of $\bG$ and glue them translationally along the solid triangles and create our Necker cube surface, denoted $\mathbf{S}$.\\
 
+\noindent Take the following unit squares in $\RR^3$:\\
 \vspace{0.2in}
@@
 
-\noindent If we take the integer triple $(m,n,p)$ and require that the individual faces are based around them, we obtain the sets
+\noindent If we take the integer triple $(m,n,p)$ and require that the individual faces based around them  has the triple lie on the plane $x+y+z=0$, $\mathbf{S}$ is broken down into the following sets:
 
@@
 \\\mathbf{B}=\bigcup\big{\{}\mathbf{B}_{m,n,p}: m+n+p=0\big{\}},
-\\\mathbf{C}=\bigcup\big{\{}\mathbf{C}_{m,n,p}: m+n+p=0\big{\}},
-\end{align*}
+\\\mathbf{C}=\bigcup\big{\{}\mathbf{C}_{m,n,p}: m+n+p=0\big{\}}.
+\end{align*}\\
+The sets $\mathbf{A},\mathbf{B},\mathbf{C}$ are contained in $\mathbf{S}$ their union is the Necker cube surface.
 
-\noindent and consider them to be piecewise linear maps of infinitely many compact topological spaces, we acquire subsets of the infinite Necker cube surface whose unions cover the surface.
+\begin{figure}[H]
+\centering
+\includesvg[width=2in]{cubesxyzcut}
+\caption{The tiling of $\bG$ where the $\mathbf{A}$ faces are copies of copies of the front-facing square of $\bG$.(Edges that were not used in the tiling are labeled).}
+\end{figure}
 
-\begin{Def} We denote the \textbf{Necker cube surface} by $\mathbf{S}$, and define it to be the surface of the union of unit squares of the form $\mathbf{S} = \mathbf{A}\cup\mathbf{B}\cup\mathbf{C}$.\end{Def}
+\subsection{Algebraic Interpretation}
+Rigid motions of the faces of $\bGs$ allow us to use a series of cutting and gluing operations to obtain an L-shaped surface homeomorphic to the two-torus with three points removed,\\
+\begin{figure}[H]
+\centering
+\includesvg[width=3in]{LTorus}
+\end{figure}
+which is then cut along the dotted lines to obtain a surface that resembles a torus with a unit square shaped hole cut out:\\
+\begin{figure}[H]
+\centering
+\includesvg[width=3in]{BranchTorus}
+\caption{$\bGs$ with 4 closed paths labeled identified}
+\end{figure}
+We assert that $\pi_1(\bGs)\cong \<a,b,c,d\>$ since it is homeomorphic to the thrice-branched torus.
+We tile this surface so that the associated deck group, $\Delta_{\bGs}$, of the cover is isomorphic to $\ZZ^2$.
+\begin{Def}
+Let $\varphi:\pi_1(\bGs)\rightarrow \ZZ^2$ be the map between groups such that \\$\varphi(1)=\varphi(c)=\varphi(d)=(0,0)$\\ $\varphi(a)=(1,0)$\\$\varphi(b)=(0,1)$\\ $\varphi(ab)=(1,1)=\varphi(ba)$.
+\end{Def}
 
-The structure can be \emph{flattened} onto plane by piecewise isometric maps onto a subset of $\mathbb C$ with unit squares removed at every even integer pair in the complex plane: 
+\begin{lem}
+$\varphi$ is a group homomorphism.
+\begin{proof}
+$\varphi$ preserves identities and is well defined on the generators of $\pi_1(\bGs)$, and is therefore a homomorphism.
+\end{proof}
+\end{lem}
+
+\begin{Def}
+$\bUs$ is the cover of $\bGs$ such that $\pi_1(\bUs)=\ker\varphi$.
+\end{Def}
+
+The kernel of the homomorphism is the free product of the commutator subgroup $[a,b]$ and free group $\<c,d\>$, so $\Delta_{\bGs}=\pi_1(\bGs)/\pi_1(\bUs)=\<a,b,c,d\>/\<c,d\>\otimes [a,b]\cong\ZZ^2$.
+The covering map that induces the inclusion $k_{\varphi*}:\pi_1(\bUs)\hookrightarrow\pi_1(\bGs)$ is the function $k_{\varphi}:\bUs\hookrightarrow\bGs$ that takes every point on $\bUs$ to its modular equivalent under these translational symmetries.
+$\bUs$ is realized as an infinite $\ZZ^2$-tiling of $\bGs$ in $\CC$:
+\begin{figure}[H]
+\centering
+\includesvg[width=2in]{uprime}
+\caption{The infinite surface $\bUs$ embedded in $\CC$ with edges identified.}
+\end{figure}
+
+\subsection{Isometry From $\mathbf{S}$ to $\bU$}
+The Necker cube surface can be \emph{flattened} onto the plane by piecewise isometric maps onto a subspace of $\RR^3$ with (open) unit squares removed at every even integer pair in the plane, what we claim to be $\bU$.
 \begin{figure}[H]
@@
 \raisebox{0.4in}{\includesvg[width=1.5in]{unfoldcut}}
-\caption{A piecewise isometry from $\mathbf{S}$ to $\mathbb{C}$. }
+\caption{A piecewise isometry from $\mathbf{S}$ to $\bU$. }
 \label{fig:Psi}
@@
 
-The flattened surface is contained entirely in $\{(x,y,z)\in\RR~|~z=0\}$, which is isometric to $\CC$. $\mathbf{S}$ is recovered as a topological quotient on the domain
+The flattened surface is contained entirely in $(x,y,0)\in\RR^3$, which is isometric to $\CC$. $\bU$ is recovered as a topological quotient on the domain
 
@@
 
-\noindent We denote the relation by $\sim_{\mathbf{P}}$, and the isometric
-
-\begin{Def} $\mathbf U$ is the surface obtained as the topological quotient $\mathbf{P}/\sim_{\mathbf{P}}$, where $\sim_{\mathbf{P}}$ is a minimal relation on $\mathbf{P}$ defined as follows:
+\begin{Def} $\mathbf U$ is the surface obtained as the topological quotient $\mathbf{P}/\sim_{\mathbf{P}}$, where $\CC$ is identified with $\RR^2$ in the usual way and $\sim_{\mathbf{P}}$ is a minimal relation on $\mathbf{P}$ defined as follows:
 \begin{gather*}
@@
 
+\begin{rem}
+While $\Psi$ does send edges and vertices to different line segments and points $(x,y,0)\sim\mathbf{P}$ depending on the choice of subset of $\mathbf{S}$, the surface $\bU$ is recovered as a topological quotient.
+This being an isometry follows from $\Psi$ being composed of Euclidean matrices
+\end{rem}
+
+\subsection{Four-Fold Cover}
+We denote the \emph{unit tangent bundle} on the surface $\bUs$ by $UT(\bUs)=\bUs\times\RR^2$, and describe a lift by $SO(2,\ZZ)$ action on the fibers of $\bUs$.
+Initial experiments have shown us that geodesics viewed as discontinuous line segments on $\mathbf{P}$ flow in any one of four directions at a time.
+\begin{figure}[H]
+\centering
+\includesvg[width=1.8in]{cubesxyzpaths}
+\includesvg[width=2.8in]{unfoldcutpath}
+\caption{Simple periodic (c) and drift-periodic (a,b) trajectories represented as line segments on $\mathbf{S}$ and $\mathbf{U}$.}
+\label{fig:holonomypath}
+\end{figure}
+Observe that a closed loop on $\bUs$ will act on the vectors by a rotation of $\pm\frac{\pi}{2}$ radians.
+Let $\eta:UT(\bUs)\rightarrow\bUs$ be the projection map onto $\bUs$, and $\gamma:[0,1]\rightarrow\bUs$ be a closed path based at $x_0\in\bUs$.
+We lift $\gamma$ to $\tilde{\gamma}$ under $\eta$ such that $\tilde{\gamma}(0)=p=(x_0,v_0)$, where $v_0$ is an arbitrary vector in $\RR^2$.
+A holonomy map $\tau:\pi_1(\bUs)\rightarrow$ Hol$_{p}$ takes $c\mapsto \begin{pmatrix}0 & -1\\ 1 & 0\end{pmatrix}=R$, $d\mapsto \begin{pmatrix}0 & 1\\ -1 & 0\end{pmatrix}=R^3$, and maps $a,b$ trivially.
+Then $\tau(\pi_1(\bUs))=\tau([a,b]\otimes\<c,d\>)=[I_2,I_2]\otimes\<R,R^3\>\cong SO(2,\ZZ)$.
+$A\in SO(2,\ZZ)$ acts on $(x,v)\in UT(\bUs)$ by the left action: $A\cdot(x,v)=(x,Av)$ that corresponds to the effect that looping around these cone singularities has on a vector.
+The corresponding lifted path thus returns as $\tilde{\gamma}(1)=\tau([\gamma])\cdot(x_0,v_0)=(x_0,\tau([\gamma])v_0)$.
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=2.5in]{monodromy.png}
+\caption{Effect a nontrivial loop has on arbitrary basis (i,j).}
+\label{fig:loop}
+\end{figure}
+Thus, $\tilde{\gamma}$ is closed if and only if $\gamma$ has trivial linear holonomy. 
+\begin{Def}
+The four-fold cover $\btUos$ is the surface with fundamental group $\pi_1(\btUos)=\ker\tau$.
+\end{Def}
+
+We consider this to be a four-fold cover, as this is a cyclic cover whose sole purpose is to have a trivial holonomy on all \emph{closed} paths in order to lift closed geodesics to closed geodesics.
+Arbitrary paths do \emph{not} have trivial holonomy by parallel transport over the edges:
+\begin{figure}[H]
+\centering
+\includegraphics[width=3in]{monogroup.png}
+\label{fig:arbitrarylift}
+\caption{Arbitrary lifts to fibers }
+\end{figure}
+The inclusion map $k_{\tau*}:\pi_1(\btUos)\hookrightarrow\pi(\bUs)$ is induced by the projection $k_{\tau}:\btUos\hookrightarrow\bUs$.
+A simple calculation also shows us that $\ker\tau=[a,b]\otimes\<c^4,d^4\>$.
+Naturally, this also extends to a monodromy map from the fundamental group to the cyclic group of permutations of elements in the fiber $k^{-1}_{\tau}(x_0)$.
 \begin{thm}
-The Necker cube surface $\mathbf{S}$ is isometric to $\mathbf{U}$.
+A periodic geodesic on $\bUs$ lifts to a periodic geodesic on $\btUos$.
 \begin{proof}
-To show that $\Psi$ is an isometry of the two surfaces, consider $(x,y,z)\in\mathbf{S}$.
-$\Psi$ is a bijection between sets $\mathbf{P}$ and $\mathbf{S}$ outside of the edges of each cube parallel to the z-axis (Figure $\ref{fig:Psi}$).
-These edges split into two edges in $\mathbf{P}$, but are identified to be one and the same in $\mathbf{U}$.\\
-For $0\leq i\leq 1$, Let $S_1=\{(m+1,n,p-i)\}, S_2=\{(m+1,n+1,p-i)\}, S_3=\{(m,n+1,p-i)\}$ be these edges.
-\compav{come back to this.}
+Let $\gamma:[0,1]\rightarrow \bUs$ be a closed, periodic geodeisc in unit direction $\gamma'$.
+Suppose that $\gamma$ does not lift to a closed path on $\btUos$.
+Then $[\gamma]\notin\ker\tau$, which implies that $\tau([\gamma])\neq I_2$.
+Now lift $\gamma$ to $UT(\bUs)$ under $\eta$ as $\tilde{\gamma}:[0,1]\rightarrow UT(\bUs)$ with base point $(\gamma(0),\gamma')$.
+Since $\gamma$ is periodic it must be true that the parallel transport of a vector on the unit tangent bundle returns unaltered.
+But because $\bUs$ has nontrivial linear holonomy, $\tilde{\gamma}(1)=\tau([\gamma])\cdot(\gamma(1),\gamma')=(\gamma(1),\tau([\gamma])\gamma')\neq (\gamma(1),\gamma')$ since $\tau([\gamma])\neq I_2$.
+Hence $[\gamma]\in\ker\tau=\pi_1(\btUos)$ and $\gamma$ lifts to a closed geodesic on $\btUos$.
 \end{proof}
@@
 \centering
-\includesvg[width=1.8in]{cubesxyzpaths}
-\includesvg[width=2.8in]{unfoldcutpath}
-\caption{Simple periodic (c) and drift-periodic (a,b) trajectories represented as line segments on $\mathbf{S}$ and $\mathbf{U}$.}
+\includesvg[width=1.5in.]{quotient}
+\caption{A $2\times2$ cut out section centered at each missing square. Edges and vertices identified.}
+\label{fig:quotient}
 \end{figure}
-
-We see from the figure above that a geodesic flow in what would appear to be the $\frac{\pi}{4}$ direction closes up on the surface.
-The edges of $\mathbf{U}$ are identified by a $\pm\frac{\pi}{2}$ rotation about a corner of one of these squares.
-It is a Euclidean cone surface with countably many conic singularities of angle $\frac{3\pi}{2}$ and $3\pi$ (3 singularities per hole).
-The set of singularities is denoted $Sing(\mathbf{U})$.
 
@@
 \centering
-\includesvg[width=1.5in.]{quotient}
-\caption{A $2\times2$ cut out section centered at each missing square. Edges and verticies identified.}
-\label{fig:quotient}
+\includegraphics[width=4in.]{closed2.png}
+\caption{Geodesic flow on $\mathbf{U}$ modeled using sage-flatsurf.}
+\label{fig:complicated}
 \end{figure}
 
-The surface $\mathbf{U}^0=\mathbf{U}\backslash Sing(\mathbf{U})$ is the surface with these cone singularities \textbf{removed}.
+\subsection{Translation Surface}
 
-\subsection{Discussion of Results}
-This section will detail how an initial trajectory angle is obtained from $\mathbf{S}$, and how it relates to the periodicity of geodesic flow as a statement of the main theorem.
+A better picture of $\btUos$ is obtained by making the following edge identifications on $\mathbf{P'}=\mathbf{P}\times\ZZ/4\ZZ$:
+
+\begin{figure}[H]
+\centering
+\includesvg[width=4in]{utilda}
+\label{fig:utilda0}
+\caption{Branched cover of $\mathbf{U}$ of degree four.}
+\end{figure}
+
+\begin{Def}
+$\btUos$ is the surface obtained as the quotient $\mathbf{P}'/\sim_{\mathbf{P}'}$.
+Denote paths $\zeta$ and $\eta$  on $\mathbf{P}$, parameterized by integers $m,n\in\ZZ$ and $t\in[-\frac{1}{4},\frac{1}{4}]$, and defined:\\ $\zeta(t)=(2m+t)+i(2n-\frac{1}{2})$,\\
+$\eta(t)=(2m+t)+i(2n+\frac{1}{2})$.\\ Let $\aleph=2m+i2n\in\CC$. The minimal relation $\sim_{\mathbf{P}'}$ is given as:
+\begin{equation}
+\begin{split}
+(\zeta(t);j)\sim_{\mathbf{P'}}(e^{i\frac{\pi}{2}}~\overline{\zeta(t)-\aleph};j+1)\\
+(\eta(t);j)\sim_{\mathbf{P'}}(e^{i\frac{\pi}{2}}~\overline{\eta(t)-\aleph};j+1)
+\label{eq:rel2}
+\end{split}
+\end{equation}
+\end{Def}
+This relation is similar to $\sim_\mathbf{P}$ in that it translates line segments of squares surrounding even integer pairs to the origin and relates points on one edge of a square to points on an adjacent edge.
+Where it differs is that these adjacent edges now belong to a different copy of $\mathbf{P}$.
+It is in this cyclic manner that edges are glued that allows for trivial linear holonomy on arbitrary paths by rotating each copy of $\mathbf{P}$ accordingly.
+For example, here is a path on a section of the surface in the neighborhood of $P=2m+i2n\in\CC$ $(m,n\in\ZZ)$ after rotation:
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=4.2in]{coverdirection.png}
+\end{figure}
+
+\noindent and its subsequent projection onto $\bUs:$
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=2in]{overlay.png}
+\end{figure}
+
+\section{Construction}
+
+The surface itself has a variety of constructions. One interpretation is the periodic tiling of ``\emph{half-cubes}," or sets of three squares sharing a single vertex.
+
+We will introduce some objects and definitions to be used later in the paper.
+We give the Necker cube surface a concrete structure as a subset of $\mathbb{R}^3$.
+Take the following unit squares in $\RR^3$:
+
+\noindent If we take the integer triple $(m,n,p)$ and require that the individual faces based around them  has the triple lie on the plane $x+y+z=0$, we obtain the sets
+
+\begin{align*}
+\mathbf{A}=\bigcup\big{\{}\mathbf{A}_{m,n,p}: m+n+p=0\big{\}},
+\\\mathbf{B}=\bigcup\big{\{}\mathbf{B}_{m,n,p}: m+n+p=0\big{\}},
+\\\mathbf{C}=\bigcup\big{\{}\mathbf{C}_{m,n,p}: m+n+p=0\big{\}}.
+\end{align*}
+
+\begin{Def} We denote the \textbf{Necker cube surface} by $\mathbf{S}$, and define it to be the surface of the union of unit squares of the form $\mathbf{S} = \mathbf{A}\cup\mathbf{B}\cup\mathbf{C}$.\end{Def}
+
+%\begin{thm}
+%The Necker cube surface $\mathbf{S}$ is isometric to $\mathbf{U}$.
+%\begin{proof}
+%To show that $\Psi$ is an isometry of the two surfaces, consider $(x,y,z)\in\mathbf{S}$.
+$\Psi$ is a bijection between sets $\mathbf{P}$ and $\mathbf{S}$ outside of the edges of each cube parallel to the z-axis (Figure $\ref{fig:Psi}$).
+These edges split into two edges in $\mathbf{P}$, but are identified to be one and the same in $\mathbf{U}$.\\
+%For $0\leq i\leq 1$, Let $S_1=\{(m+1,n,p-i)\}, S_2=\{(m+1,n+1,p-i)\}, S_3=\{(m,n+1,p-i)\}$ be these edges.
+%\compav{come back to this.}
+%\end{proof}
+%\end{thm}
+
+\subsection{Four-Fold Cover}
+This section describes how a branched translation cover of $\bU$ is obtained by winding around cone singularities of the surface.
+
+We denote the surface $\bU$ with its countably many cone singularities removed by $\bUs$.
+$\bUs$ is an infinite-genus surface obtainable as a translation cover of a thrice-punctured torus (figure $\ref{fig:quotient}$).
+
+There is a nontrivial linear holonomy as can be seen in figure $\ref{fig:holonomypath}$.
+The reader is encouraged to check that the total angle around each singularity is either $\frac{3\pi}{2}$ or $3\pi$ radians.
+
+As to be expected, a parallel transport of a geodesic flow over an edge will act on the unit tangent bundle by a rotation of $\pm\frac{\pi}{2}$ radians.
+This is best exemplified by the periodic flow labeled $c$ in figure $\ref{fig:holonomypath}$, where the geodesic flowing in the $\frac{1}{\sqrt{2}}(1,1)$ can be seen as line segments on $\mathbf{P}$.
+When a unit direction is fixed, it is mapped to its equivalence class of directions in the orbit under $SO(2,\ZZ)$ action.
+
+If we begin flowing a unit vector in some direction there are only one of four possible directions in the ambient space of the surface (See figure $\ref{fig:complicated}$ for further convincing).
+This is primarily what motivated taking four copies of the surface so as to lift a flow to a surface whose rotational symmetries could align the flow by associating every one of the four possible directions with its own ``copy" of $\mathbf{P}$.
+$\btU$ will become what is known as an infinite translation surface, equipped with a flat metric and trivial linear holonomy.
+
+Surrounding each pair of even integers in $(2\ZZ+2\ZZ i)^2\subset \CC$, there is an open unit square removed.
+Four copies of the domain $\mathbf{P}$ are indexed by elements in $\ZZ/4\ZZ$, and we consider the set of points $\mathbf{P}'=\mathbf{P}\times\ZZ/4\ZZ$.
+It is on $\mathbf{P}'$ that we define a minimal relation $\sim_{\mathbf{P}'}$ that is used to identify edges in the following manner:
+
+We call this surface $\btU_0$.
+It is another infinite surface constructed in a cyclic manner that complements $\bUs$'s holonomy.
+Winding around each cone singularity shows us that there are 4 distinct vertices for each removed square centered at $(2\ZZ+2\ZZ i)^2\times \ZZ/4\ZZ$, each of cone angle $6\pi$.
+The surface with these singularities removed is denoted $\btU_0^\circ$.
+The obvious covering map $p:\btU_0^\circ\rightarrow\bUs$ is taken to be a projection map onto $\bUs$ induced by the map $\mathbf{P}'\hookrightarrow \mathbf{P}$, where $(z;j)\mapsto(z)$.
+
+\subsection{Monodromy Action on Fibers of $\bUs$}
+Let $x_0$ be a point on $\bU^\circ$, and fix a point $\tilde{x}_0\in\btU_0^\circ$ in the fiber $p^{-1}(x_0)=\{\tilde{x}_0,\tilde{x}_1,\tilde{x}_2,\tilde{x}_3\}$.
+Consider the closed paths around a hole in $\bUs$ in figure $\ref{fig:loop}$, and their subsequent lifts onto $\btU_0^\circ$:
+
+The kernel of the associated monodromy map $m:\pi_1(\bUs,x_0)\times p^{-1}(x_0)\rightarrow p^{-1}(x_0)$ of the cover $p$ is a cyclic permutation of 4 elements isomorphic to $SO(2,\ZZ)$.
+The lifted kernel of this map is equal to $\pi_1(\btU_0^\circ,\tilde{x}_0)$.
+
+\subsection{Translation Surface}
+Now we consider rotating each plane about their origins so that the parallel transport of a unit vector over an edge acts trivially on that vector.
+We consider \emph{North} on this new surface, call it $\btU$, obtained in this way to respect $\bU$'s notion of vertical direction (relative to the first plane).
+Let $P=2m+i2n\in \CC$ for some $m,n\in\ZZ$.
+Then the following diagrams illustrate the rotational relationships between directions in $\btU$ and $\bU$:
+
+\begin{lem}
+A geodesic on $\bU$ is periodic if and only if it's lifted to a closed geodesic on $\btU_0$.
+\begin{proof}
+
+\compav{Redo this maybe define the cover by monodromy map}
+Since a geodesic is undefined at conical singularities, it suffices to work with $\bUs$ and $\btU_0^\circ$.
+Let $\gamma:[0,1]\rightarrow UT(\bUs)$ be a closed geodesic defined on the unit tangent bundle of $\bUs$ such that $\gamma(0)=\gamma(1)=(x_0,v_0)$ for constant unit vector $v_0\in\RR^3$.
+The covering map $p$ induces a map $p_*:UT(\btUos)\hookrightarrow UT(\bUs)$.
+The fiber $p_*^{-1}(x_0,v_0)=\{(\tilde{x}_0,V),(\tilde{x}_1,V),(\tilde{x}_2,V),(\tilde{x}_3,V)\}$ with $v_0\in V$.
+Suppose that the lift $\tilde{\gamma}$ with $\tilde{\gamma}(0)=\tilde{x_0}$ does not close, then $\tilde{\gamma}(1)\in\{\tilde{x}_1,\tilde{x}_2,\tilde{x}_3\}$ Then $[\gamma]\notin\ker m$ and $m[\gamma]\neq \mathbf{I}_2$.
+Since $\btUos$ is simply connected, there is some $\tilde{\alpha}:[0,1]\rightarrow \btUos$ lifted from $\alpha\in\bUs$ such that $\tilde{x}_i\mapsto\tilde{x_0}$ for $i=1,2,3$.
+Then $m[\alpha\circ\gamma]=m[\alpha]\cdot m[\gamma]\neq m[\alpha]\neq \mathbf{I}_2$ implies that $[\alpha]\in \ker m$, which is a contradiction since then $\tilde{\alpha}$ does not lift to a closed path.\\
+Conversely let $\tilde{\gamma}$ be a closed geodesic on 
+\end{proof}
+\end{lem}
+
+\newpage
 
@@
 \noindent We have two subsets of $\mathbb Z^2$ directions that partition the set of possible rational directions $\theta$.
-\begin{Def}
-Let $v$ be a unit vector of the form $\frac{1}{k}(x,y)\simeq\mathbb{R}/2\pi\mathbb{Z}$ for $x,y\in\mathbb{Z}$ and $k=\sqrt{x^2+y^2}\in\mathbb{R}$.
-We call $v$ an \textbf{odd-odd} vector if x and y are relatively prime integers and both odd.
-We denote the \textbf{set of all odd-odd directions} $(x,y)\in\mathcal{O}$.
-We say that $v$ is an \textbf{even-odd} vector if x and y are relatively prime integers, and x is even if and only if y is odd.
-We denote the \textbf{set of all even-odd directions} $(x,y)\in\mathcal{E}$.
-\end{Def}
 
@@
 For example, the following trajectory modeled on the surface is a closed geodesic on $\mathbf{U}$ with initial trajectory (51,31):
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=4in.]{closed2.png}
-\caption{Geodesic flow on $\mathbf{U}$ modeled using sage-flatsurf.}
-\end{figure}
 

### 2017-10-16 newpaper.tex
101517

--- 2017-09-13 newpaper.tex
@@
 \def\AA{\mathbb{A}} \def\BB{\mathbb{B}} \def\CC{\mathbb{C}} \def\DD{\mathbb{D}} \def\EE{\mathbb{E}} \def\FF{\mathbb{F}} \def\GG{\mathbb{G}} \def\HH{\mathbb{H}} \def\II{\mathbb{I}} \def\JJ{\mathbb{J}} \def\KK{\mathbb{K}} \def\LL{\mathbb{L}} \def\MM{\mathbb{M}} \def\NN{\mathbb{N}} \def\OO{\mathbb{O}} \def\PP{\mathbb{P}} \def\QQ{\mathbb{Q}} \def\RR{\mathbb{R}} \def\SS{\mathbb{S}} \def\TT{\mathbb{T}} \def\UU{\mathbb{U}} \def\VV{\mathbb{V}} \def\WW{\mathbb{W}} \def\XX{\mathbb{X}} \def\YY{\mathbb{Y}} \def\ZZ{\mathbb{Z}}
+
+\def\bU{\mathbf{U}} \def\btU{\tilde{\bU}} \def\bUs{\bU^\circ}
+\def\btUos{\btU_0^\circ}
+\def\bG{\mathbf{G}} \def\bGs{\mathbf{G}^\circ}
+\def\hol{\mathbf{hol}}
+\def\bM{\mathbf{M}}
 
@@
 \def\ZZn{\ZZ/n\ZZ}
+\def\acts{\rotatebox[origin=c]{-90}{$\circlearrowright$}}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
@@
 \begin{abstract}
-The Necker cube surface is an infinite topological surface obtained by the periodic edge-to-edge tiling of unit cubes in $\RR^3$ along a 2-dimensional subspace realized as a Euclidean cone surface with a discrete, countable set of singularities of cone angles $3\pi$ and $\frac{3\pi}{2}$.
-Its various symmetries allow for the categorization of periodic and drift-periodic straight line trajectories in terms of their initial trajectory angles.
-Results are obtained from studying the Teichmuller geodesic flow on a branched four-fold cover of the surface, where the group of deck transformations of the cover preserves the trajectory angle properties, which we use to determine periodicity on what is an infinite translation surface.
-The initial rational angle is obtained by the ratio of relatively prime integer vectors of the form $(a,b)$.
-When $a,b$ are both odd the flow is periodic, and whenever $a$ or $b$ is even, the flow is drift-periodic.
-\\
+We study dynamical properties of geodesic flows on a flat, periodically constructed Euclidean cone surface by obtaining an infinite-type translation cover branched over conical singularities, of which admits a flat induced metric and a connection between any two points in the tangent vector bundle of the surface has trivial holonomy.
+A unit vector flow on the unit tangent bundle of such a surface is well-defined with a canonical vector representation in $\RR^2$.
+We study this infinite-type surface as a $\ZZ^2$ cover of a compact Veech surface belonging to the stratum of $\cH(2,2,2,2)$ translation surfaces, and use its $SL(2,\RR)-$commensurable Veech group to prove results that relate directions of flows to the periodicity and ergodicity of their lifts in the cover.
 \end{abstract}
+\begin{figure}[H]
+\centering
+\includesvg[width=2.2in]{cubesxyzpaths2}
+\label{fig:front}
+\caption{Periodic and drift-periodic flows on the Necker cube surface}
+\end{figure}
 
@@
 
-\compav{ignore:}
-The Necker cube surface, $\mathbf{S}$, is an infinite topological surface obtained by the periodic tiling of unit cubes by $\mathbb{R}^3$ translational symmetries along a 2-dimensional subspace realized as a Euclidean cone surface with a discrete set of singularities of cone angle $3\pi$ and $\frac{3\pi}{2}$.
-An isometric ``flattening" of $\mathbf{S}$ is obtained by injective piecewise-linear maps onto the planar subspace.
-The fundamental domain of the surface admits infinitely unit square \emph{holes} centered on the  $(2\mathbb{Z})^2$ lattice in $\mathbb{C}$.
-The flat surface, $\mathbf{U}$, has rotational and translation symmetries and a unit tangent bundle, $UT(\mathbf{U})$, that is isomorphic to $\mathbb{Z}/\frac{\pi}{2}\mathbb{Z}$ as a consequence of the edge identifications.
-A \emph{flat} branched cover, $\tilde{\mathbf{U}}$, of the surface is obtained by a quotient topology defined on four copies of $\mathbf{U}$'s fundamental domain and indexed by elements in $\mathbb{Z}/4\mathbb{Z}$ for the purpose of aligning geodesic flows.
-It is a \emph{formally étale} cover of $\mathbf{U}$ whose unit tangent bundle is $\mathbb{Z}/2\pi\mathbb{Z}$ and obtained by the monodromy group of the covering map.
-$\tilde{\mathbf{U}}$ is a translation surface whose quotient by its translational symmetries is a \emph{Veech} surface whose \emph{Veech group} is $SL(2,\mathbb{Z})$-commensurable.
-This paper's aim is to study the relations between geodesic flows on base surfaces, Deck transformations of their covers, the orbit of rational directions in $\mathbb{Z}^2$ under Veech group action, and use group theoretic results to prove theorems about rational trajectories on the Necker cube surface.
-\compav{ignore$\uparrow$}\\
 Put your Problem statement here! Example of a Citation\cite[p.219]{Robotics}. Here's Another Citation \cite{Flueck}
 
-The Necker cube[citation] has made numerous appearances throughout history.
-The artist famous for his use of optical illusions in his work, M.C.
-Escher, has occasionally used tilings of the Necker cube such as in the lithographs ``Metamorphosis I," and ``Convex and Concave." (Pictured below.) The crystallographer Louis Albert Necker was credited for having extensively studied the Necker cube's geometry[citation] and remarked on this simple, yet fascinating, optical illusion that appears to ``simultaneously protrude from and intrude into the page."[citation:wolfram]\compav{Find new citation} Some might even recognize it as the same surface in which ``Q*bert" hops around on in the 1982 arcade game published by Gottlieb.
+Historically, the Necker cube[citation] has made numerous appearances in the work of mathematicians, crystallographers, and scientists interested in human visual systems prior to it being popularized in the works of illusionary artist M.C.
+Escher (pictured below).
+The crystallographer Louis Albert Necker was credited for having discovered the optical illusion, and studying its geometry[cite].
+The solid presentation of the cube, when rendered as a flat surface with its back faces removed, achieves a similar effect when its three visible faces are shaded in a particular way.
+We are interested in the infinite tiling of this structure (figure $\ref{fig:front}$), which has the appearance of a rhombile tiling of the Euclidean plane, and refer to it as the Necker cube surface.
+Video game fans might even recognize it as the same surface that appears in the 1982 arcade game by Gottlieb, ``Q*bert."
 
@@
 \includegraphics[scale=0.55]{escher2.jpg}
-\caption{The Necker cube tiling as it appears briefly in ``Metamorphosis I," and hung on a banner in ``Convex and Concave."[citation]}
+\caption{The Necker cube tiling as it appears briefly in ``Metamorphosis I," and impressed into a banner in ``Convex and Concave."[citation]}
 \label{fig:Escher}
@@
 
-This paper assumes the reader is familiar with fundamental concepts in algebraic topology such as covering space theory, induced homomorphisms on surface homologies of their symmetries, as well as some background in Veech structures of compact translation surfaces.
-\compav{cite something}
+Flat, periodic surfaces such as the Necker cube surface are of much interest to differential geometers for two reasons in particular.
+The first is that, locally, neighborhoods are isometric to the plane, making it remarkably easy to describe parallel transports between two points in tangent vector bundles over the surface by linear group actions.
+The second is that Riemann surfaces constructed out of infinitely many polyhedra can be \emph{flattened} and contained in the plane in such a way that the metric is preserved.
+The advantage in this case is that one obtains something akin to infinite billiard tables where a geodesic flows are represented by series of straight line segments in the plane.
+This paper will use such methods to prove dynamical results about Geodesic flows on the Necker cube surface.
+
+\begin{rem}
+Familiarity is assumed on the part of the reader with covering space theory, translation (Veech) surfaces, and their associated Veech groups. For general surveys on these topics: [cite], [cite].
+\end{rem}
+
+\subsection{Discussion of Results}
+Initial experiments suggested there was a correlation between choice of trajectory angle and periodicity.
+On a surface composed of infinitely many cubes, there is certainly a flat metric endowed on smaller neighborhoods of the surface that are locally isometric to the plane.
+Via parallel transports of tangent vectors over the edges of the cubes connecting one face to the next, it is possible to describe geodesic behavior on the Necker cube surface as a series of straight line segments that traverse faces and continuously flow in a fixed direction (see figure $\ref{fig:front}$).
+\\\\
+
+\begin{wrapfigure}{l}{2.2in.}
+\includesvg{vectorplane}
+\end{wrapfigure}
+We consider classes of directions obtained from unit vectors in $\RR^3$ by projecting a sufficiently small segment of a flow contained onto a parallel plane.
+These vectors retain all necessary information about their directions and let us consider rational and irrational angles relative to a choice of basis in $\RR^2$.
+This is what we refer to as the \emph{initial trajectory angle}.
+\\\\
+Rational component trajectory angles are then further categorized according to the parities of their components.
+\begin{Def}
+Let $v$ be a unit vector of the form $\frac{1}{k}(x,y)\in\RR^2$ for $x,y\in\mathbb{Z}$ and $k=\sqrt{x^2+y^2}\in\mathbb{R}$.
+We call $v$ an \textbf{odd-odd} vector if x,y are relatively prime and both odd.
+We denote the \textbf{set of all odd-odd directions} by $\mathcal{O}$.
+We say that $v$ is an \textbf{even-odd} vector if x and y are relatively prime, and x is even if and only if y is odd.
+We denote the \textbf{set of all even-odd directions} by $\mathcal{E}$.
+\end{Def}
+We prove the following is true of all rational directions:
+
+\begin{thm*}
+Let $\Phi_t$ be the geodesic flow defined on the unit tangent bundle of the Necker cube surface for $t\in\RR$, and initial point $p=(x_0,v_0)$ with $v_0\in\RR$. Then the following is true:
+\begin{enumerate}[label=(\roman*)]
+\item $\Phi$ is periodic with period $T\in\RR$ if and only if $v\in\cO$.
+\item $\Phi$ is drift-periodic with period $T\in\RR$ if and only if $v\in\cE$.
+\end{enumerate}
+\end{thm*}
+
+The proof for this theorem can be found in XXX and follows from Theorems YYYY and ZZZZZ.
+\\
+\compav{theorem about most directions being recurrent?}
+
+\subsection{Acknowledgements}
+-Pat Hooper\\
+- Vincent Delecroix, Ferrán Valdez, pascal hubert\\
+- \\
+
 \newpage
 
-\subsection{Background}
-We will introduce some objects, definitions, and theorems to be used later in the paper.
-First and foremost, we need to describe the construction of the Necker cube surface and give it a concrete structure as a subset of $\mathbb{R}^3$.
+\section{Construction}
+This section details how we construct the Necker cube surface as an infinite cover of a compact Riemann surface.
+\subsection{Geometric Interpretation}
+Consider the following surface:
+\begin{figure}[H]
+\centering
+\includesvg[width=2in.]{neckercube}
+\end{figure}
 
-\noindent Take the following unit squares in $\RR^3$:
+We call this surface $\bG$, the \emph{Necker cube}.
+$\bG$ is obtained by taking three copies of the unit square and identifying the edges in order to resemble the visible portion of a cube.
+$\bG$ has a total of three vertices.
+Every vertex has a cone angle of either $3\pi$ or $\frac{3\pi}{2}$ and are \emph{conical points} (or \emph{singularities}) of the surface.
+We denote the surface without singularities by $\bGs$.
+The periodic tiling of $\bG$ will take infinitely many copies of $\bG$ and tile them along the edges labeled with solid triangles.
+This cover of $\bG$ is the Necker cube surface we denote by $\mathbf{S}$.
+Likewise, the surface tiled by $\bGs$ we denote by $\mathbf{S}^\circ$\\
 
+\noindent Take the following unit squares in $\RR^3$:\\
 \vspace{0.2in}
@@
 \\\mathbf{B}=\bigcup\big{\{}\mathbf{B}_{m,n,p}: m+n+p=0\big{\}},
-\\\mathbf{C}=\bigcup\big{\{}\mathbf{C}_{m,n,p}: m+n+p=0\big{\}},
-\end{align*}
+\\\mathbf{C}=\bigcup\big{\{}\mathbf{C}_{m,n,p}: m+n+p=0\big{\}}.
+\end{align*}\\
+The sets $\mathbf{A},\mathbf{B},\mathbf{C}$ are contained in $\mathbf{S}$. Their union is also the Necker cube surface.
 
-\noindent and consider them to be piecewise linear maps of infinitely many compact topological spaces, we acquire subsets of the infinite Necker cube surface whose unions cover the surface.
+\begin{figure}[H]
+\centering
+\includesvg[width=2in]{cubesxyzcut}
+\caption{The tiling of $\bG$ where the $\mathbf{A}$ faces are copies of copies of the front-facing square of $\bG$.(Edges that were not used in the tiling are labeled).}
+\end{figure}
 
-\begin{Def} We denote the \textbf{Necker cube surface} by $\mathbf{S}$, and define it to be the surface of the union of unit squares of the form $\mathbf{S} = \mathbf{A}\cup\mathbf{B}\cup\mathbf{C}$.\end{Def}
+\subsection{Algebraic Interpretation}
+We use a series of cutting and gluing operations to obtain an L-shaped surface homeomorphic to the torus branched at three points:\\
+\begin{figure}[H]
+\centering
+\includesvg[width=3in]{LTorus}
+\end{figure}
+\noindent Which is then cut along these dotted lines and pieced together to resemble the torus with a unit-square removed from the center:\\
+\begin{figure}[H]
+\centering
+\includesvg[width=3in]{BranchTorus}
+\caption{$\bGs$ with 4 independent, closed paths labeled identified}
+\end{figure}
+The labeled paths are independent from eachother, and a quick calculation shows us that its fundamental group $\pi_1(\bGs)\cong \<a,b,c,d\>$, the free group of four generators (denote identity by $\mathbf{1}$).
+We describe an infinite cover of $\bGs$ such that the associated deck group is isomorphic to $\ZZ^2$
+\begin{Def}
+We define $\varphi_1:\pi_1(\bGs)\rightarrow \ZZ^2$ as the group homomorphism on generators $a,b,c,d$: \\$c,d\mapsto(0,0)$\\ $a\mapsto(1,0)$\\$b\mapsto(0,1)$\\ $ab,ba\mapsto(1,1)$.
+\end{Def}
 
-The structure can be \emph{flattened} onto plane by piecewise isometric maps onto a subset of $\mathbb C$ with unit squares removed at every even integer pair in the complex plane: 
+\begin{Def}
+$\bUs$ is the cover of $\bGs$ such that $\pi_1(\bUs)=\ker\varphi$.
+\end{Def}
+
+Our cover of $\bGs$ has the fundamental group, $\pi_1(\bUs)=\<\<a^{-1}b^{-1}ab, c, d\>\>$, the conjugate subgroup of elements that map trivially under $\varphi_1$.
+The covering map $k_{\varphi_1}:\bUs\hookrightarrow\bGs$ takes every point on $\bUs$ to its modular equivalent under these translational symmetries.
+$\bUs$ is realized as an infinite $\ZZ^2$-tiling of $\bGs$ in $\CC$:
+\begin{figure}[H]
+\centering
+\includesvg[width=2in]{uprime}
+\caption{The infinite surface $\bUs$ embedded in $\CC$ with edges identified.}
+\end{figure}
+
+\begin{Def}
+$\bU$ is the completion of $\bUs$ that includes the branched vertices. 
+\end{Def}
+
+\begin{rem}
+The surface $\bG$ is homeomorphic to the torus, and its cover $\bU$ is its universal cover as the paths labeled $c,d$ are only non-trivial when $\bG$ is branched at its three conical points.
+\end{rem}
+
+\subsection{Isometry From $\mathbf{S}$ to $\bU$}
+The Necker cube surface can be \emph{flattened} onto the plane by piecewise isometric maps onto a subspace of $\RR^3$ with (open) unit squares removed at every even integer pair in the plane, what we claim to be $\bU$.
 \begin{figure}[H]
@@
 		\end{array} \right]
-			& \text{if } (x,y,z)\in \mathbf{B}	\vspace{2mm}
+			& \text{if } (x,y,z)\in \mathbf{B}\backslash(\mathbf{A}\cup\mathbf{C})	\vspace{2mm}
 	\\
@@
 			\end{array} \right]
-				& \text{if } (x,y,z)\in \mathbf{C}	\vspace{2mm}
+				& \text{if } (x,y,z)\in \mathbf{C}\backslash(\mathbf{A}\cup\mathbf{B})	\vspace{2mm}
 \end{cases}
@@
 \end{array}\right]
-. \end{gather*}
-\end{Def}
+.\end{gather*}\end{Def}
 
-\begin{thm}
-The Necker cube surface $\mathbf{S}$ is isometric to $\mathbf{U}$.
-\begin{proof}
-To show that $\Psi$ is an isometry of the two surfaces, consider $(x,y,z)\in\mathbf{S}$.
-$\Psi$ is a bijection between sets $\mathbf{P}$ and $\mathbf{S}$ outside of the edges of each cube parallel to the z-axis (Figure $\ref{fig:Psi}$).
-These edges split into two edges in $\mathbf{P}$, but are identified to be one and the same in $\mathbf{U}$.\\
-For $0\leq i\leq 1$, Let $S_1=\{(m+1,n,p-i)\}, S_2=\{(m+1,n+1,p-i)\}, S_3=\{(m,n+1,p-i)\}$ be these edges.
-\compav{come back to this.}
-\end{proof}
-\end{thm}
+\begin{rem}
+Our map $\Psi$ induces an isometry between $\mathbf{S}$ and $\bU$, and its restriction to the branched surfaces induces an isometry between $\mathbf{S}^\circ$ and $\bUs$.
+This follows from these piecewise Euclidean transformations that preserves our induced flat metric.
+\end{rem}
+From here on we use $\bUs$ and $\bU$ instead of $\mathbf{S}$ and $\mathbf{S}^\circ$.
 
+\subsection{Four-Fold Cover}
+We denote the \emph{unit tangent bundle} on the surface $\bUs$ by $T^1\bUs$, and construct a cover with trivial holonomy.
+Initial experiments have shown us that geodesics viewed as discontinuous line segments on $\mathbf{P}$ flow in any one of four directions at a time.
 \begin{figure}[H]
@@
 \includesvg[width=2.8in]{unfoldcutpath}
-\caption{Simple periodic (c) and drift-periodic (a,b) trajectories represented as line segments on $\mathbf{S}$ and $\mathbf{U}$.}
+\caption{Simple periodic (c) and drift-periodic (a,b) trajectories represented as line segments on $\mathbf{S}$ and $\mathbf{U}$.\compav{RELABEL THESE}}
+\label{fig:holonomypath}
 \end{figure}
-
-We see from the figure above that a geodesic flow in what would appear to be the $\frac{\pi}{4}$ direction closes up on the surface.
-The edges of $\mathbf{U}$ are identified by a $\pm\frac{\pi}{2}$ rotation about a corner of one of these squares.
-It is a Euclidean cone surface with countably many conic singularities of angle $\frac{3\pi}{2}$ and $3\pi$ (3 singularities per hole).
-The set of singularities is denoted $Sing(\mathbf{U})$.
+Observe that a closed loop on $\bUs$ will act on vectors tangent to the surface by a rotation of $\pm\frac{\pi}{2}$ radians (since the surface is flat and embedded in $\CC$, we work with vectors in $\RR^2$).
 
@@
 \centering
-\includesvg[width=1.5in.]{quotient}
-\caption{A $2\times2$ cut out section centered at each missing square. Edges and verticies identified.}
-\label{fig:quotient}
+\includegraphics[width=2.5in]{monodromy.png}
+\caption{Effect a nontrivial loop has on arbitrary basis (i,j).}
+\label{fig:loop}
 \end{figure}
 
-The surface $\mathbf{U}^0=\mathbf{U}\backslash Sing(\mathbf{U})$ is the surface with these cone singularities \textbf{removed}.
-
-\subsection{Discussion of Results}
-This section will detail how an initial trajectory angle is obtained from $\mathbf{S}$, and how it relates to the periodicity of geodesic flow as a statement of the main theorem.
+Let $\omega$ be a flat, Riemannian metric connection on $T^1\bUs$.
+This is allowable since $\bUs$ is obtained as the cover of a flat Riemann surface.
+We say that $p=(x_0,v_0)\in T^1\bUs$ is a fixed point in the fiber over $x_0$ in the unit tangent bundle of the surface, with $x_0\in\bUs$ and $v_0\in\RR^2$ being unit.
+Denote the \emph{holonomy group} of the connection with base point p by $Hol_p(\omega)=\{A\in SO(2,\RR)~:~ p \equiv A\cdot p\}$ where $p$ differs from $A\cdot p$ by some rotation of $v_0$, and the \emph{holonomy bundle}  by $H(p)$, the set of points $p'\in T^1\bUs$ where a closed loop on $\gamma:[0,1]\rightarrow\bUs$ lifts to $\gamma^*:[0,1]\rightarrow T^1\bUs$ such that $p=\gamma^*(0)$ and $p'=\gamma^*(1)$.
 
 \begin{Def}
-Take a trivial straight-line path in $S\subset\mathbb{R}^3$, $\ell:[0,1]\rightarrow\mathbf{S}$ such that for some $m,n,p\in\mathbb{Z},~ \ell(t)$ is contained in exactly one of $\mathbf{A}_{m,n,p},\mathbf{B}_{m,n,p},~or~\mathbf{C}_{m,n,p}$ for all $t\in[0,1]$.
-The \textbf{initial trajectory angle} of a geodesic on $\mathbf{S}$ is given as $\theta=Arg(\Psi(\ell(1,s))-\Psi(\ell(0,s)))$, where Arg is the principal argument of a vector in $\mathbb{C}$.
-The \textbf{initial trajectory} is the vector $\dot{v}=(\cos(\theta),\sin(\theta))\in\mathbb{R}^2$.\\
-\compav{Not sure if I should proceed with derivative or map}
-\begin{figure}[H]
-\centering
-\includesvg{vectorplane}
-\end{figure}
+Define $\varphi_2:\pi_1(\bGs)\rightarrow SO(2,\ZZ)$ as the group homomorphism on generators $a,b,c,d$ such that:
+\begin{align*}
+a,b\mapsto \begin{pmatrix}1 && 0 \\ 0 && 1\end{pmatrix}=I_2,~~~
+c\mapsto \begin{pmatrix}0 && -1 \\ 1 && 0\end{pmatrix}=R,~~~
+d\mapsto \begin{pmatrix}0 && 1 \\ -1 && 0\end{pmatrix}=R^3
+\end{align*}
 \end{Def}
+The homomorphism restricted to $\pi_1(\bUs)$ factors through $\pi_1(\bGs)$ and we show that:
+\begin{lem}
+The group $\varphi_2(\pi_1(\bUs,x_0))=Hol_p(\omega)$.
+\begin{proof}
+Since $c\mapsto R$ and $d\mapsto R^3$, the image of $\varphi_2$ is generated by $R,R^3$ and isomorphic to $SO(2,\ZZ)$.
+We see from figure $\ref{fig:holonomypath}$ that parallel transports of vectors by non-trivial paths produce clockwise/counter-clockwise rotations equal to that of of the cone angles of the singularities they loop around, all integer multiples of $\frac{\pi}{2}$ that coincide with paths generated by elements $c,d$.
+Further, if there is an element in $\pi_1(\bUs,x_0)$ conjugated by $[a,b]$, the effect this would have on holonomy would be trivial as the total cone angle sum around each cut out square is an integer multiple of $2\pi$, which agrees with $a$ and $b$'s trivial images under $\phi_2$.
+\end{proof}
+\end{lem}
 
-\noindent We have two subsets of $\mathbb Z^2$ directions that partition the set of possible rational directions $\theta$.
-\begin{Def}
-Let $v$ be a unit vector of the form $\frac{1}{k}(x,y)\simeq\mathbb{R}/2\pi\mathbb{Z}$ for $x,y\in\mathbb{Z}$ and $k=\sqrt{x^2+y^2}\in\mathbb{R}$.
-We call $v$ an \textbf{odd-odd} vector if x and y are relatively prime integers and both odd.
-We denote the \textbf{set of all odd-odd directions} $(x,y)\in\mathcal{O}$.
-We say that $v$ is an \textbf{even-odd} vector if x and y are relatively prime integers, and x is even if and only if y is odd.
-We denote the \textbf{set of all even-odd directions} $(x,y)\in\mathcal{E}$.
-\end{Def}
-
-If $v$ is rational, that is $v=k(x,y)$ for integers x,y, then we commonly refer to the pair $(x,y)$ as the \emph{slope} of a given initial trajectory.
-For example, the following trajectory modeled on the surface is a closed geodesic on $\mathbf{U}$ with initial trajectory (51,31):
+Hence, we obtain a natural monodromy representation with the map $m:\pi_1(\bUs,x_0)\rightarrow SO(2,\ZZ)\cong Hol_p(\omega)$, where for $[\gamma]\in\pi_1(\bUs,x_0)$ we have that $\gamma^*(1)=(x_0,m([\gamma])v_0)$.
+It follows that since $\omega$ is a flat connection, trivial loops have trivial holonomy and $Hol_p(\omega)$ acts on $H(p)$.
 
@@
 \caption{Geodesic flow on $\mathbf{U}$ modeled using sage-flatsurf.}
+\label{fig:complicated}
 \end{figure}
@@
 \begin{Def}
-A \textbf{geodesic flow} on a topological surface, $X$, is a function $\Phi^{v}_t:UT(X)\times\mathbb{R}\rightarrow UT(X)$ such that $\frac{d\Phi}{dt}=v\in UT(X)$. 
+Let $\btUos$ be the cover of $\bUs$ with fundamental group $\pi_1(\btUos)=\ker\varphi_2|_{\pi_1(\bUs)}\trianglelefteq \pi_1(\bUs)$.
 \end{Def}
 
-\noindent Experimental evidence indicates that geodesics in the odd-odd directions are periodic, while geodesics in the even-odd directions are drift-periodic. We will prove the theorem below.
+\begin{lem}
+Let $\Phi_t:\bUs\times\RR\rightarrow \bUs$ be a unit-speed geodesic flow on $\bUs$, with a parallel transport map induced by $\omega$. Then the following is true:
+\begin{enumerate}[label=(\roman*)]
+\item $\Phi$ is periodic if and only if its lift to $\btUos$ is.
+\item $\Phi$ is drift-periodic if and only if its lift to $\btUos$ is. 
+\end{enumerate}
+\begin{proof}
+Let $x_0=\Phi(0)$ be an initial point in $\bUs$, and let $v_0=\frac{d}{dt}\Phi|_{t=0}\in\RR^2$ be its initial direction.
+Then $\Phi^*:T^1\bUs\times\RR\rightarrow T^1\bUs$ is a well-defined lift to the unit tangent bundle with initial point $p=(x_0,v_0)$.
+\\\emph{(i)}.
+Suppose $\Phi$ is periodic with period $T\in\RR$.
+Then $\Phi$ is reparameterized as $\alpha:[0,1]\rightarrow \bUs$, where $[\alpha]\in\pi_1(\bUs,x_0)$.
+It follows then that if $\alpha$ does not lift to a closed path, then $\alpha$ must have non-trivial holonomy since $[\alpha]\notin\ker\varphi_2$.
+That is when lifted to the unit tangent bundle with base point $p$, $\alpha^*(1)\neq p$ since $m([\alpha])\neq I_2$.
+But this implies that $\Phi^*(T)\neq p$, which is impossible.
+Hence $[\alpha]\in\ker\varphi_2=\pi_1(\btUos,\tilde{x}_0)$, where $\tilde{x}_0$ belongs to the fiber over $x_0$ under the covering map $\btUos\hookrightarrow\bUs$.
+The converse holds trivially.
+\\\emph{(ii)}.
+Suppose $\Phi$ is drift-periodic with period $T\in\RR$ and non-trivial $f:\bUs\rightarrow\bUs \in Trans(\bUs)\cong\pi_1(\bGs)/\pi_1(\bUs)\cong \ZZ^2$ such that $\Phi(T)=f(x_0)$.
+Re-parameterize this as $\alpha:[0,1]\rightarrow\bUs$ with $[\alpha]\in\pi_1(\bGs)$.
+When lifted to $T^1\bUs$, $\Phi^*(0)=p$ and $\Phi^*(T)=(f(x_0),v_0)$.
+Let $\alpha:[0,1]\rightarrow \bUs$ be its reparameterization up to time $T$.
+Thus $[\alpha]\notin \pi_1(\bUs,x_0),\pi_1(\btUos,\tilde{x}_0)$.
+Further, $f$ has a unique, lift to $\tilde{f}\in Trans(\btUos)$ because the space is connected.
+(a bit stuck here).
+\\
+Conversely, ...
+\end{proof}
+\end{lem}
 
-\begin{thm}{\textbf{Dynamics of Geodesic Flow on the Necker cube surface.}}\\ Obtain $\theta$ and $v$ as described in Definition 3.
-Denote the non-singular unit-speed geodesic flow with initial point $s\in\mathbf{U}^0$ in direction $(\theta\mod{\frac{\pi}{2}})\sim\phi\in UT(\mathbf{U}^0)$ by $\Phi^{\phi}_t:UT(\mathbf{U}^0)\times\mathbb{R}\rightarrow UT(\mathbf{U}^0)$ such that $\frac{d\Phi}{dt}=\phi\in UT(X)$  Then the following is true:
-\begin{enumerate}[label=(\roman*)]
-\item (Periodic) There exists a $t_0 > 0$ such that $\Phi^{\phi}_{t+t_{0}}(s)=\Phi^{\phi}_{t}(s)$ if and only if $v\in\mathcal{O}$.
-\item (Drift-Periodic) There exists a $t_0 > 0$ such that $\Phi^{\phi}_{t+t_{0}}(s)= \Phi^{\phi}_{t}\circ\Phi^{\phi}_{t_0}(s)\neq\Phi^{\phi}_t(s)$ if and only if $v\in\mathcal{E}$.
-\end{enumerate}
-\end{thm}
+A visual representation of $\btUos$ is as a four-fold cover of $\bUs$ with trivial holonomy on all closed paths. Arbitrary paths do \emph{not} have trivial holonomy:
+\begin{figure}[H]
+\centering
+\includegraphics[width=3in]{monogroup.png}
+\label{fig:arbitrarylift}
+\caption{Arbitrary lifts to fibers }
+\end{figure}
 
-\subsection{Translation Surfaces}
-We recall some common definitions, theorems, and properties of translation surfaces and their $\mathbb{Z}^d$-covers.
+\begin{figure}[H]
+\centering
+\includesvg[width=1.5in.]{quotient}
+\caption{A $2\times2$ cut out section centered at each missing square. Edges and vertices identified.}
+\label{fig:quotient}
+\end{figure}
+
+\subsection{Translation Surface}
+
+A better picture of $\btUos$ is obtained by making cyclic edge identifications on $\mathbf{P'}=\mathbf{P}\times\ZZ/4\ZZ$.
+This comes from $\btUos$ inheriting the topological properties of $\bUs\times \pi_1(\bUs)/\pi_1(\btUos)\cong \bUs\times SO(2,\ZZ)$.
+
+\begin{figure}[H]
+\centering
+\includesvg[width=4in]{utilda}
+\label{fig:utilda0}
+\caption{Branched cover of $\mathbf{U}$ of degree four.}
+\end{figure}
 
 \begin{Def}
-A translation surface is a pair $M=(S,\omega)$ formed of a connected
-Riemann surface X and a holomorphic 1-form $\omega$ on $S$ which is not identically zero and satisfies:
-\begin{enumerate}
-\item $\Sigma\subset S$ is a discrete subset of cone singularities of $S$.
-\item $\omega=z^kdz$ at points in $\Sigma$ of cone angle $2(k+1)\pi$.
-\item $S\backslash\Sigma$ admits a translation surface $M^0=(S\backslash\Sigma,dz)$, where dz is a flat metric on $S\backslash\Sigma$.
-\end{enumerate}
+$\btU_0$ is the surface obtained as the quotient $\mathbf{P}'/\sim_{\mathbf{P}'}$.
+Denote paths $\zeta$ and $\eta$  on $\mathbf{P}$, parameterized by integers $m,n\in\ZZ$ and $t\in[-\frac{1}{4},\frac{1}{4}]$, and defined:\\ $\zeta(t)=(2m+t)+i(2n-\frac{1}{2})$,\\
+$\eta(t)=(2m+t)+i(2n+\frac{1}{2})$.\\ Let $\aleph=2m+i2n\in\CC$. The minimal relation $\sim_{\mathbf{P}'}$ is given as:
+\begin{equation}
+\begin{split}
+(\zeta(t);j)\sim_{\mathbf{P'}}(e^{i\frac{\pi}{2}}~\overline{\zeta(t)-\aleph};j+1)\\
+(\eta(t);j)\sim_{\mathbf{P'}}(e^{i\frac{\pi}{2}}~\overline{\eta(t)-\aleph};j+1)
+\label{eq:rel2}
+\end{split}
+\end{equation}
+\end{Def}
+This relation is similar to $\sim_\mathbf{P}$ in that it translates line segments of squares surrounding even integer pairs to the origin and relates points on one edge of a square to points on an adjacent edge.
+Where it differs is that these adjacent edges now belong to a different ``copy" of $\mathbf{P}$.
+It is in this cyclic manner that edges are glued that allows for trivial linear holonomy on arbitrary paths by rotating each copy of $\mathbf{P}$ accordingly.
+For example, here is a path on a section of the surface in the neighborhood of $P=2m+i2n\in\CC$ $(m,n\in\ZZ)$ after rotation:
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=4.2in]{coverdirection.png}
+\end{figure}
+
+\noindent and its subsequent projection onto $\bUs:$
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=2in]{overlay.png}
+\end{figure}
+
+\begin{Def}
+Let $r:\mathbf{P}'\rightarrow\mathbf{P}'$ be the isometric map of $\btU_0$'s domain given  $r(z;\bar{j})=(e^{(-j)i\frac{\pi}{2}}z;\bar{j})$.
 \end{Def}
 
-The set of all orientation-preserving affine automorphisms of $M$ forms the group Aff$^+(M)$.
-The corresponding \emph{Veech group} of M is the image of the group morphism $D:\text{Aff}^+(M)\rightarrow SL(2,\mathbb{R})$, where $D$ is the derivative map of an element of Aff$^+(M)$, and denote it by $V(M)$.
-A surface is said to by \emph{Veech} if its Veech group is commensurable to $SL(2,\mathbb{R})$.
-The unit tangent bundle on $M\backslash\Sigma$, denoted $UT(M)$, is isomorphic to $M\times\mathbb{R}/2\pi\mathbb{Z}$.
-Let $\phi,\psi$ be two compatible charts between $M$ and $\mathbb{C}$ such that $\phi(x_0)\in\mathbb{C}$ for $x_0\in M\backslash\Sigma$, and let $F^t_\theta:z\mapsto z+te^{i\theta}$ be a translation flow on $\mathbb{C}$.
-The non-singular geodesic flow on the unit tangent bundle of $M$, $G^t:UT(M)\times\mathbb{R}\rightarrow UT(M)$, is well defined on and given by $G^t(x_0,\theta)=(\psi^{-1}\circ F^t_\theta\circ\phi(x_0),\theta)$.
-\\
-A \emph{maximal geodesic} on $M$ is a re-parameterized closed geodesic $\gamma:[0,1]\rightarrow M$ obtained from $G_t$, and we consider $\gamma\in\pi_1(M,x_0)$.
-A normal subgroup $N\triangleleft\pi_1(M,x_0)$ is associated with the quotient group $\Delta=\pi_1(M,x_0)/N$.
-$\tilde{M}$ is called the $\Delta$-cover of $M$ if $p:\tilde{M}\rightarrow M$ is a cover of $M$ and for some fixed $\tilde{x}_0\in p^{-1}(x_0)$, $\pi_1(M,x_0)/\Delta=\pi_1(\tilde{M},\tilde{x}_0)$.
-$\Delta$ has a unique lift to the group of Deck transformations acting on $\tilde{x}_0\in\tilde{M}$, or its orbit under Aut$(p)$.
-Compact translation surfaces are oriented, hence why their fundamental groups are isomorphic to $\mathbb{Z}^{2g}$ (g is the genus of $M$).
-Therefore $\Delta\simeq \mathbb{Z}^d$, and $\tilde{M}$ is called a $\mathbb{Z}^d$ \emph{translation cover} of $M$.
-Translation covers have Deck transformation groups isomorphic to $\mathbb{Z}^d$.
-\\
-When $\gamma\in\pi_1(M,x_0)$ is a closed curve, we denote its abelianization as $[\gamma]\in H_1(M,R)$, with coefficients in unit ring $R$, and its unique lift to $\tilde{M}$ under the covering map $p$ as $\tilde{\gamma}$.
-Suppose you have some set of homology classes, $\{a_1,\dots,a_n\}=\Gamma\subset H_1(M,R)$, such that span$(\Gamma)=H_1(M,R).$ We call $\Gamma$ the \emph{spanning set}, and express an element $[\beta]\in H_1(M,R)$ as $[\beta]=\Sigma_{j=1}^{n}x_j a_j$, where $x_1,\dots x_n\in R$.
+Observe that when $r$ acts on the relations defined on $\ref{eq:rel2}$ we get a new relation:
+
+\begin{equation}
+\begin{split}
+(e^{(-j)i\frac{\pi}{2}}\zeta(t);j)\sim_{r(\mathbf{P'})}(e^{(-j)i\frac{\pi}{2}}~{\zeta(t)-\aleph};j+1)\\
+(e^{(-j)i\frac{\pi}{2}}\eta(t);j)\sim_{r(\mathbf{P'})}(e^{(-j)i\frac{\pi}{2}}~{\eta(t)-\aleph};j+1)
+\end{split}
+\end{equation}
+
+$\btU$ is then recovered as $r\cdot(\btU_0)=\mathbf{P'}/\sim_{r(\mathbf{P}')}$, where all identifications are translations made on copies of $\mathbf{P}$.
+
+\subsection{Translation Surfaces and $\ZZ^2$-Covers}
+We recall some definitions and properties of translation surfaces and their translation covers.
+A \emph{translation surface}, $S$, is a compact, flat surface that admits a discrete number of conical singularities and has trivial linear holonomy on arbitrary paths.
+We denote the surface without singularities by $S^\circ$.
+Translation surfaces belong to a stratum or family of surfaces characterized by the number of conical points and their degrees $(\text{cone angle}-2\pi)$.
+A $Z^2$-cover is a cover of $S$ given by a map $f:\tilde{S}\rightarrow S$, such that the associated deck group of the cover, $\Delta=\pi_1(S)/\pi_1(\tilde{S})=\pi_1(S^\circ)/\pi_1(\tilde{S}^\circ)$, is isomorphic to $\ZZ^2$.
 
 \begin{Def}
-Denote the non-degenerate, bi-linear \textbf{intersection number} between two homology classes $[\alpha],[\beta]$ as $i([\alpha],[\beta])$, where 
+Intersection number is a non-degenerate bilinear form:
 \begin{align*}
-i:H_1(M,R)\times H_1(M, R)\rightarrow \mathbb{Z}.
+i:H_1(S,\mathbf{R})\times H_1(S,\mathbf{R})\rightarrow\mathbf{R},
 \end{align*}
-returns the total number of intersections and its sign corresponds to the sign of the angle that $[\beta]$ makes relative to $[\alpha]$.
+where $\mathbf{R}$ is abelian and for $[\gamma],[\beta]\in H_1(S,\mathbf{R})$, $i([\beta],[\gamma])$ returns the intersection number as an element of $\mathbf{R}$, where a positive crossing has $[\gamma]$ make a positive angle relative to $[\beta]$.
 \end{Def}
 
-We mainly concern ourselves with $\mathbb{Z}^2$ covers for the purposes of this paper. Consider the set $Q\subset N\subset \pi_1(M,x_0)$ to be the spanning set of N.
-
 \begin{Def}
-Consider two sets, $q,r\subset\Gamma$, such that span$(q,r)$ is the abelianization of $N$.
-There exists a group morphism $s:N\rightarrow Aut(p)$ that takes a generator of $N$ to a generator of $Aut(p)$, a translational symmetry of $\tilde{M}$ isometric to $(m,n)\in\mathbb{Z}^2$, call it $T^{m,n}$.
-Note that dim$(N)=2g-2$, and $s$ is an epimorphism when $g>2$.
-Obtain $q$ as the abelianization of elements in $Q$ that are mapped to $T^{0,\pm 1}$, the vertical translations.
-Likewise, take $r$ to be the set of abelianized elements of $Q$ that are mapped to  $T^{\pm 1, 0}$, the horizontal translations.
-A path non-trivially intersecting with a curve in q corresponds to a horizontal translation, while an intersection with r corresponds to a translation in the vertical direction.
-This is because an intersection of a closed path $\beta$ with an element $a \in s^{-1}(T^{0,\pm 1})$ implies that $\beta$ at the time of intersection would have had to have been homotopic to $b\in s^{-1}(T^{\pm 1,0})$ instead (assuming that $a,b$ intersect).\\
-For example, let $t$ be the time of that intersection, and $t_0$ be the time of the previous one.
-Then $\tilde{\beta}(t)=T^{\pm 1,0}(\tilde{\beta}(t_0))$ would have moved horizontally, and not vertically.\\
-Since intersection number is bi-linear and non-degenerate, we   obtain the \textbf{group homomorphism as a representation of a lifted path} on $\tilde{M}$:
-
+Let $u,v\in H_1(S,\mathbf{R})$ be homology classes of curves on $S$ that span $H_1(\tilde{S},\mathbf{R})$. Then there is a group homomorphism from $\pi_1(S)$ to $\mathbf{R}^2$ defined as:
 \begin{align*}
-\Omega_{q,r}:\pi_1(M,x_0)\rightarrow \mathbb{Z}^2 \text{; } \beta\mapsto(i(q,[\beta]),i(r,[\beta])).
+\Omega_{u,v}:\pi_1(S)\rightarrow \mathbf{R}^2 \text{; } \beta\mapsto(i(u,[\beta]),i(v,[\beta])).
 \end{align*}
@@
 
+The set of all orientation-preserving affine diffeomorphisms of $S$ forms the group Aff$^+(S)$.
+The corresponding \emph{Veech group}, $V(S)$ of $S$ is the image of the group morphism $D:\text{Aff}^+(S)\rightarrow SL(2,\mathbb{R})$ that takes an affine map to its derivative.
+A surface is said to be \emph{Veech} if its Veech group is commensurable to $SL(2,\mathbb{R})$.
+It is well known that origami, or square-tiled, surfaces have Veech groups commensurable to $SL(2,\ZZ)$.
+[cite] When $\mathbf{R}=\ZZ$, $\Omega_{u,v}$ takes an element of $\pi_1(S)$ to $\Delta$.
+Thus, $\gamma\in\pi_1(S)$ lifts to $\tilde{\gamma}\in\pi_1(\tilde{S})$ if and only if $\gamma\in\ker\Omega_{u,v}$.
+
+\begin{Def}
+Let $\alpha:[0,1]\rightarrow S$ be a closed, non-singular geodesic path on $S$.
+The holonomy map $\mathbf{hol}:H_1(S, \mathbf{R})\rightarrow\CC$ returns the holonomy vector of a closed path as a difference of the starting and endpoints of a flow by\\
+\begin{align*}
+\mathbf{hol}([\alpha])=\int_{\alpha}dz.
+\end{align*}
+Since $\alpha$ is non-singular, it can be mapped to $S^\circ$ which admits a flat holomorphic one-form $dz$.
+Let $\theta=Arg(\mathbf{hol}([\alpha]))$.
+We say that $\phi_t^\theta:[0,\infty)\rightarrow S^\circ$ is the unit-speed geodesic flow on $S^\circ$ given by $[\alpha]$ such that $\phi^\theta_0=\alpha(0)$.
+\end{Def}
+
 \begin{lem}
-Let $\beta\in\pi_1(M,x_0)$. Then $\beta$ lifts to $\tilde{\beta}\in\pi_1(\tilde{M},\tilde{x}_0)$ if and only if $\beta\in\text{Ker }\Omega_{q,r}$.
+$\phi_t^\theta$ has a period of $T=|\mathbf{hol}([\alpha])|$
 \begin{proof}
-do this later
+This just follows from the fact that $\phi_t^\theta$ flows at unit-speed in the direction of $\frac{\mathbf{hol}([\alpha])}{|\mathbf{hol}([\alpha])|}$. 
+\end{proof}
+\end{lem}
+And so the length of a vector determines the period of a flow on $S$. More importantly, we have the following:
+
+\begin{lem}
+Denote the lifted flow of $\phi_t^\theta$ on $\tilde{S}$ by $\tilde{\phi}_t^\theta$.
+Then $\tilde{\phi}_t^\theta$ is periodic on $\tilde{S}$ and $\tilde{S}^\circ$ if and only if $[\alpha]\in\ker\Omega_{u,v}$.
+Further, $\tilde{\phi}_t^\theta$ has period $T=|\mathbf{hol}([\alpha])|$.
+\begin{proof}
+Suppose that $[\alpha]\notin\ker\Omega_{u,v}$.
+Since $\alpha([0,1])=\phi^\theta_{[0,\hol([\alpha])]}$, their homology classes are equivalent.
+Hence $\tilde{\phi}_t^\theta$ could not close on $\tilde{S}$ or $\tilde{S}^\circ$, or else $\Omega_{u,v}(k[\alpha])=k(i(u,[\alpha]),i(v,[\alpha]))=(0,0)$ for some non-zero $k\in\ZZ$.
+Conversely, suppose $[\alpha]\in\ker\Omega_{u,v}$.
+Then
 \end{proof}
@@
 
-This lemma becomes an integral part of the proof of the main theorem, and allows us to show how induced homomorphisms of a surface's homology by its affine automorphisms allows for us to generalize whenever a geodesic on the base surface lifts to a closed geodesic on the cover.
-\subsection{Acknowledgements}
+\begin{cor}
+If $[\alpha]\notin\ker\Omega_{u,v}$, then $\tilde{\phi}_t^\theta$ is drift-periodic with period $T=\hol([\alpha])$.
+\begin{proof}
+This follows immediately from the previous lemma since $\tilde{S},\tilde{S}^\circ$ have translational $\ZZ^2$ symmetries and a well-defined $\ZZ^2$ action on an element in the fiber of a basepoint in $S$ under $f$.
+The period is $T$ since a geodesic closes on $S$ with period $T$.
+\end{proof}
+\end{cor}
+
+\begin{lem}
+Let $h\in\text{Aff}^+(S)$. If $h(\beta)=\alpha$ for closed geodesics $\alpha,\beta$ on $S$ and $\beta$ lifts to a closed path on $\tilde{S}$, then so does $\alpha$. (not sure about this)
+\begin{proof}
+If the premise is true, then $[\beta]\in\ker\Omega_{u,v}$.
+Denote the group automorphism induced by $h$ as $h_*$.
+Then $\Omega_{u,v}([\alpha])=\Omega_{u,v}([h(\beta)])=\Omega_{u,v}(h_*\cdot[\beta])=(i([\beta],h_*^{-1}\cdot u),i([\beta],h_*^{-1}\cdot v))=(0,0)$ as automorphisms .
+\end{proof}
+\end{lem}
+
+\begin{lem}
+Let $\beta=h(\alpha)$ as before. If $D(h)=h'\in V(S)$ and $S$ is Veech, then $\tilde{\phi}^\theta_t$ has period $T=h'\cdot\hol([\beta])$. (also not sure)
+\begin{proof}
+
+\end{proof}
+\end{lem}
+
+\subsection{Veech Staircase Model}
+Let $f:\btU\rightarrow\mathbf{M}$ be a projection onto the quotient surface that takes every point to its modular equivalent by its $\ZZ^2$ symmetries, and what we obtain is the following compact translation surface belonging to the stratum $\cH(2,2,2,2):$
+
+\begin{figure}[H]
+\centering
+\includesvg[width=2.6in]{mtilda}
+\caption{Compact translation surface, $\mathbf{M}$, covered by the infinite surface with edges and cone singularities (1,2,3,4) identified.
+The Roman numerals are meant to identify each quotient with a plane in the cover, $\mathbf{\tilde{\mathbf{U}}}$.}
+\label{fig:mtilda}
+\end{figure}
+
+It is clear that this is a square-tiled Veech surface when realized as a finite staircase:
+
+\begin{figure}[H]
+\centering
+\includesvg[width=3.4in]{mtildastaircase}
+\caption{The staircase Veech surface with directional planes and vertices identified.
+All edges are paired by translation.
+Two adjacent squares have opposite edges identified.
+The top edge of the bottom-left square is glued to the bottom edge of the top-right square (both labeled I).
+Likewise, the bottom edge of the bottom-left square is identified with the top edge of the top-right square.}
+\label{fig:staircase}
+\end{figure}
+
+Since $\pi_1(\bM)\leq\pi_1(\bM^\circ)$, we may work with the absolute homology of $\bM$ to prove results on $M^\circ$. Consider the following cylinder curves on $\bM$:
+
+\begin{figure}[H]
+\includegraphics[width=3in]{homologyclass.png}
+\centering
+\caption{$\bM$'s cylinder core curves with u,v, and z homology class labels.}
+\label{fig:homology}
+\end{figure}
+
+\begin{Def}
+The set of abelianized cylinder core curves is $\Gamma=\{\gamma_i~: ~0\leq i\leq11\}\subset H_1(\bM,\ZZ)\subset H_1(\mathbf{M},\mathbb{Q})$.
+\end{Def}
+
+It can be shown that $\Gamma$ spans $H_1(\bM,\QQ)$ by constructing a $12\times12$ matrix $M$ such that $M_{j,k}=i(\gamma_k,\gamma_j)$.
+Every row has two entries that alternate between $1$ when $j$ is odd and $-1$ when $j$ is even.
+Such a matrix is rank 10, precisely the dimension needed for a genus 5 surface.
+
+\begin{Def}
+The homology classes $u,v,z$ are given as the following sums of core curves:
+\begin{align*}
+u &= -[\gamma_2] +[\gamma_5] + [\gamma_8] - [\gamma_{11}],\\
+v &= +[\gamma_0] -[\gamma_3] -[\gamma_6] +[\gamma_9],\\
+z &= +[\gamma_1] +[\gamma_4]-[\gamma_7]-[\gamma_{10}].
+\end{align*}
+\end{Def}
+
+\begin{Def}
+Define the group homomorphism $\Omega_{u,v}:\pi_1(\bM)\rightarrow \QQ^2$, where $\beta\mapsto(i(u,[\beta]),i(v,[\beta]))$.
+\end{Def}
+
+\begin{thm}
+
+\end{thm}
+
+\subsection{Symmetries of $\bM$}
+\begin{figure}[H]
+\centering
+\includesvg[width=2in]{skew}
+\end{figure}
 
@@
 \end{Def}
-
-We take the quotient of $\mathbf{\tilde{\mathbf{U}}}$ by the action of the induced isometry of $T^{m,n}$, to get a translation base surface, $\mathbf{M}$.
-
-\begin{figure}[H]
-\centering
-\includesvg[width=2.6in]{mtilda}
-\caption{Compact translation surface, $\mathbf{M}$, covered by the infinite surface with edges and cone singularities (1,2,3,4) identified.
-The Roman numerals are meant to identify each quotient with a plane in the cover, $\mathbf{\tilde{\mathbf{U}}}$.}
-\label{fig:mtilda}
-\end{figure}
-
-It is clear that this is a square-tiled Veech surface when realized as a finite staircase:
-
-\begin{figure}[H]
-\centering
-\includesvg[width=3.4in]{mtildastaircase}
-\caption{The staircase Veech surface with directional planes and vertices identified.
-All edges are paired by translations.
-Two adjacent squares have opposite edges identified.
-The top edge of the bottom-left square is glued to the bottom edge of the top-right square (both labeled I).
-Likewise, the bottom edge of the bottom-left square is identified with the top edge of the top-right square.}
-\label{fig:staircase}
-\end{figure}
 
@@
 
-Even though the main theorem remains unproven for now, it is an enjoyable property of the surface and venturing in its understanding involves exploring some fascinating niches in algebraic topology and dynamical systems.
-
-Needless to say, it has been a very busy summer for me.
-I am grateful for the opportunity to really take some time to delve deep into numerous topics that I have been very interested in for a while.
-I sincerely doubt it would have been possible without the summer internship, and I have the Rich scholarship committee to thank for that.
-I would also like to extend my gratitude to Dr.
-Hooper.
-He has been incredibly patient and kind throughout these last few months, and I would not have come this far in so short a time without his help.
-
 \newpage
 \begin{thebibliography}{9}
-\bibitem{Robotics} Fred G. Martin \emph{Robotics Explorations: A Hands-On Introduction to Engineering}. New Jersey: Prentice Hall.
-\bibitem{Flueck}  Flueck, Alexander J. 2005. \emph{ECE 100}[online]. Chicago: Illinois Institute of Technology, Electrical and Computer Engineering Department, 2005 [cited 30
-August 2005]. Available from World Wide Web: (http://www.ece.iit.edu/~flueck/ece100).
+\bibitem{Necker} 
+\bibitem{Albert} Necker, L.A., \emph{Observations on some remarkable optical phaenomena seen in Switzerland; and on an optical phaenomenon which occurs on viewing a figure of a crystal or geometrical solid.} London and Edinburgh Philosophical Magazine and Journal of Science.
+1832.
+doi:10.1080/14786443208647909
+\bibitem{Qbert} Gottlieb., \emph{Q*bert} [Arcade], Gottlieb, 1982.
+\bibitem{Escher1} Escher, M.C., \emph{Metamorphosis I}, 1937.
+\bibitem{Escher2} Escher, M.C., \emph{Convex and Concave}, 1955.
 \end{thebibliography}

### 2017-10-25 newpaper.tex
102417

@@
 \usepackage{amsmath}
+\usepackage{amssymb}
 \usepackage{graphicx}
@@
 We denote the \emph{unit tangent bundle} on the surface $\bUs$ by $T^1\bUs$, and construct a cover with trivial holonomy.
-Initial experiments have shown us that geodesics viewed as discontinuous line segments on $\mathbf{P}$ flow in any one of four directions at a time.
+Initial experiments have shown us that any geodesic viewed as discontinuous line segments on $\mathbf{P}$ moves in at most four directions.
 \begin{figure}[H]
@@
 \end{figure}
-Observe that a closed loop on $\bUs$ will act on vectors tangent to the surface by a rotation of $\pm\frac{\pi}{2}$ radians (since the surface is flat and embedded in $\CC$, we work with vectors in $\RR^2$).
+Observe that the parallel transport of a vector around a closed loop on $\bUs$ will act on vectors tangent to the surface by a rotation of an integer multiple of $\frac{\pi}{2}$ radians (since the surface is flat and embedded in $\CC$, we work with vectors in $\RR^2$).
 
@@
 
-Let $\omega$ be a flat, Riemannian metric connection on $T^1\bUs$.
-This is allowable since $\bUs$ is obtained as the cover of a flat Riemann surface.
-We say that $p=(x_0,v_0)\in T^1\bUs$ is a fixed point in the fiber over $x_0$ in the unit tangent bundle of the surface, with $x_0\in\bUs$ and $v_0\in\RR^2$ being unit.
-Denote the \emph{holonomy group} of the connection with base point p by $Hol_p(\omega)=\{A\in SO(2,\RR)~:~ p \equiv A\cdot p\}$ where $p$ differs from $A\cdot p$ by some rotation of $v_0$, and the \emph{holonomy bundle}  by $H(p)$, the set of points $p'\in T^1\bUs$ where a closed loop on $\gamma:[0,1]\rightarrow\bUs$ lifts to $\gamma^*:[0,1]\rightarrow T^1\bUs$ such that $p=\gamma^*(0)$ and $p'=\gamma^*(1)$.
+Let $\triangledown$ be a flat, Riemannian metric connection on $T^1\bUs$.
+This is allowable since $\bUs$ is obtained as the cover of a flat Riemann surface (\compav{you said of a translation surface?}).
+Let $p=(x_0,v_0)\in E_{x_0}\subset T^1\bUs$ be an element in the fiber over $x_0$ the usual projection from the tangent bundle to the surface, with $x_0\in\bUs$ and $v_0\in\RR^2$.
+Denote the \emph{holonomy group} of the connection with base point $p$ by $Hol_p(\triangledown)=\{A\in SO(E_{x_0})~:~ p \equiv A\cdot p\}$ where $p$ differs from $A\cdot p$ by some rotation of $v_0$.
+That is, $p\equiv p'$ if $p'=(x_0,A v_0)$ for $A\in Hol_p(\triangledown)$.
+Denote the \emph{holonomy bundle} based at $p$ by $H(p)$, the set of points $p'\in E_{x_0}$ where a closed loop $\gamma:[0,1]\rightarrow\bUs$ describes a horizontal lift to $\gamma^*:[0,1]\rightarrow T^1\bUs$ such that $p=\gamma^*(0)$ and $p'=\gamma^*(1)$.
 
@@
 
-Observe that when $r$ acts on the relations defined on $\ref{eq:rel2}$ we get a new relation:
+Observe that when $r$ acts on the relations $\eqref{eq:rel2}$ we get the following:
 
@@
 $\btU$ is then recovered as $r\cdot(\btU_0)=\mathbf{P'}/\sim_{r(\mathbf{P}')}$, where all identifications are translations made on copies of $\mathbf{P}$.
-
-\subsection{Translation Surfaces and $\ZZ^2$-Covers}
-We recall some definitions and properties of translation surfaces and their translation covers.
-A \emph{translation surface}, $S$, is a compact, flat surface that admits a discrete number of conical singularities and has trivial linear holonomy on arbitrary paths.
-We denote the surface without singularities by $S^\circ$.
-Translation surfaces belong to a stratum or family of surfaces characterized by the number of conical points and their degrees $(\text{cone angle}-2\pi)$.
-A $Z^2$-cover is a cover of $S$ given by a map $f:\tilde{S}\rightarrow S$, such that the associated deck group of the cover, $\Delta=\pi_1(S)/\pi_1(\tilde{S})=\pi_1(S^\circ)/\pi_1(\tilde{S}^\circ)$, is isomorphic to $\ZZ^2$.
-
-\begin{Def}
-Intersection number is a non-degenerate bilinear form:
-\begin{align*}
-i:H_1(S,\mathbf{R})\times H_1(S,\mathbf{R})\rightarrow\mathbf{R},
-\end{align*}
-where $\mathbf{R}$ is abelian and for $[\gamma],[\beta]\in H_1(S,\mathbf{R})$, $i([\beta],[\gamma])$ returns the intersection number as an element of $\mathbf{R}$, where a positive crossing has $[\gamma]$ make a positive angle relative to $[\beta]$.
-\end{Def}
-
-\begin{Def}
-Let $u,v\in H_1(S,\mathbf{R})$ be homology classes of curves on $S$ that span $H_1(\tilde{S},\mathbf{R})$. Then there is a group homomorphism from $\pi_1(S)$ to $\mathbf{R}^2$ defined as:
-\begin{align*}
-\Omega_{u,v}:\pi_1(S)\rightarrow \mathbf{R}^2 \text{; } \beta\mapsto(i(u,[\beta]),i(v,[\beta])).
-\end{align*}
-\end{Def}
-
-The set of all orientation-preserving affine diffeomorphisms of $S$ forms the group Aff$^+(S)$.
-The corresponding \emph{Veech group}, $V(S)$ of $S$ is the image of the group morphism $D:\text{Aff}^+(S)\rightarrow SL(2,\mathbb{R})$ that takes an affine map to its derivative.
-A surface is said to be \emph{Veech} if its Veech group is commensurable to $SL(2,\mathbb{R})$.
-It is well known that origami, or square-tiled, surfaces have Veech groups commensurable to $SL(2,\ZZ)$.
-[cite] When $\mathbf{R}=\ZZ$, $\Omega_{u,v}$ takes an element of $\pi_1(S)$ to $\Delta$.
-Thus, $\gamma\in\pi_1(S)$ lifts to $\tilde{\gamma}\in\pi_1(\tilde{S})$ if and only if $\gamma\in\ker\Omega_{u,v}$.
-
-\begin{Def}
-Let $\alpha:[0,1]\rightarrow S$ be a closed, non-singular geodesic path on $S$.
-The holonomy map $\mathbf{hol}:H_1(S, \mathbf{R})\rightarrow\CC$ returns the holonomy vector of a closed path as a difference of the starting and endpoints of a flow by\\
-\begin{align*}
-\mathbf{hol}([\alpha])=\int_{\alpha}dz.
-\end{align*}
-Since $\alpha$ is non-singular, it can be mapped to $S^\circ$ which admits a flat holomorphic one-form $dz$.
-Let $\theta=Arg(\mathbf{hol}([\alpha]))$.
-We say that $\phi_t^\theta:[0,\infty)\rightarrow S^\circ$ is the unit-speed geodesic flow on $S^\circ$ given by $[\alpha]$ such that $\phi^\theta_0=\alpha(0)$.
-\end{Def}
-
-\begin{lem}
-$\phi_t^\theta$ has a period of $T=|\mathbf{hol}([\alpha])|$
-\begin{proof}
-This just follows from the fact that $\phi_t^\theta$ flows at unit-speed in the direction of $\frac{\mathbf{hol}([\alpha])}{|\mathbf{hol}([\alpha])|}$. 
-\end{proof}
-\end{lem}
-And so the length of a vector determines the period of a flow on $S$. More importantly, we have the following:
-
-\begin{lem}
-Denote the lifted flow of $\phi_t^\theta$ on $\tilde{S}$ by $\tilde{\phi}_t^\theta$.
-Then $\tilde{\phi}_t^\theta$ is periodic on $\tilde{S}$ and $\tilde{S}^\circ$ if and only if $[\alpha]\in\ker\Omega_{u,v}$.
-Further, $\tilde{\phi}_t^\theta$ has period $T=|\mathbf{hol}([\alpha])|$.
-\begin{proof}
-Suppose that $[\alpha]\notin\ker\Omega_{u,v}$.
-Since $\alpha([0,1])=\phi^\theta_{[0,\hol([\alpha])]}$, their homology classes are equivalent.
-Hence $\tilde{\phi}_t^\theta$ could not close on $\tilde{S}$ or $\tilde{S}^\circ$, or else $\Omega_{u,v}(k[\alpha])=k(i(u,[\alpha]),i(v,[\alpha]))=(0,0)$ for some non-zero $k\in\ZZ$.
-Conversely, suppose $[\alpha]\in\ker\Omega_{u,v}$.
-Then
-\end{proof}
-\end{lem}
-
-\begin{cor}
-If $[\alpha]\notin\ker\Omega_{u,v}$, then $\tilde{\phi}_t^\theta$ is drift-periodic with period $T=\hol([\alpha])$.
-\begin{proof}
-This follows immediately from the previous lemma since $\tilde{S},\tilde{S}^\circ$ have translational $\ZZ^2$ symmetries and a well-defined $\ZZ^2$ action on an element in the fiber of a basepoint in $S$ under $f$.
-The period is $T$ since a geodesic closes on $S$ with period $T$.
-\end{proof}
-\end{cor}
-
-\begin{lem}
-Let $h\in\text{Aff}^+(S)$. If $h(\beta)=\alpha$ for closed geodesics $\alpha,\beta$ on $S$ and $\beta$ lifts to a closed path on $\tilde{S}$, then so does $\alpha$. (not sure about this)
-\begin{proof}
-If the premise is true, then $[\beta]\in\ker\Omega_{u,v}$.
-Denote the group automorphism induced by $h$ as $h_*$.
-Then $\Omega_{u,v}([\alpha])=\Omega_{u,v}([h(\beta)])=\Omega_{u,v}(h_*\cdot[\beta])=(i([\beta],h_*^{-1}\cdot u),i([\beta],h_*^{-1}\cdot v))=(0,0)$ as automorphisms .
-\end{proof}
-\end{lem}
-
-\begin{lem}
-Let $\beta=h(\alpha)$ as before. If $D(h)=h'\in V(S)$ and $S$ is Veech, then $\tilde{\phi}^\theta_t$ has period $T=h'\cdot\hol([\beta])$. (also not sure)
-\begin{proof}
-
-\end{proof}
-\end{lem}
-
 \subsection{Veech Staircase Model}
@@
 
-It is clear that this is a square-tiled Veech surface when realized as a finite staircase:
+We rearrange $\bM$ as such:
 
@@
 
-Since $\pi_1(\bM)\leq\pi_1(\bM^\circ)$, we may work with the absolute homology of $\bM$ to prove results on $M^\circ$. Consider the following cylinder curves on $\bM$:
-
+\noindent We recall some basic definitions and theorems about $\ZZ^2-$covers of translation surfaces as they apply to $\bM$.
+
+\begin{Def}
+Algebraic intersection number is a non-degenerate bilinear form:
+\begin{align*}
+i:H_1(\bM,\QQ)\times H_1(\bM,\QQ)\rightarrow\QQ,
+\end{align*}
+for $[\gamma],[\beta]\in H_1(\mathbf{M},\QQ)$, $i([\beta],[\gamma])$ returns the signed intersection number of two homology classes.
+We say a crossing at the instance of an intersection is positive if $\gamma$ makes a positive angle relative to $\beta$.
+\end{Def}
+The set of all affine diffeomorphisms of $\bM$ form the group $\text{Aff}^+(\bM)$.
+The $\emph{Veech group}$ of $\bM$ is the image in the co-domain of the homomorphism $D:\text{Aff}^+(\bM)\rightarrow SL(2,\RR)$ that takes every affine map to its derivative, which we denote $V(\bM)$.
+$\bM$ is a square-tiled translation surface whose Veech group is a finite index subgroup of $SL(2,\ZZ)$.
+We use automorphisms of the fundamental group in $\text{Aut}(\pi_1(\bM))$ induced by affine maps to prove our main results.
+But first, we define a group homomorphism from $\pi_1(\bM)$ to $\pi_1(\bM)/\pi_1(\btU)$ by algebraic intersection numbers pver a sum of linearly independent homology classes.
+Consider the following cylinder core curves on $\bM$ labeled $\gamma_i$ for $0\leq i\leq 11$:
 \begin{figure}[H]
@@
 
-\begin{Def}
-The set of abelianized cylinder core curves is $\Gamma=\{\gamma_i~: ~0\leq i\leq11\}\subset H_1(\bM,\ZZ)\subset H_1(\mathbf{M},\mathbb{Q})$.
-\end{Def}
-
-It can be shown that $\Gamma$ spans $H_1(\bM,\QQ)$ by constructing a $12\times12$ matrix $M$ such that $M_{j,k}=i(\gamma_k,\gamma_j)$.
-Every row has two entries that alternate between $1$ when $j$ is odd and $-1$ when $j$ is even.
-Such a matrix is rank 10, precisely the dimension needed for a genus 5 surface.
+Every curve intersects two others and if we take $i(\Sigma_{i=0}^{11}[\gamma_i])$
+\compav{show that $\gamma_0,\dots,\gamma_{11}$ form a basis for $H(\bM,\QQ)$}\\
+\compav{show that u,v linearly independent}
 
@@
 
-\begin{thm}
-
-\end{thm}
+\begin{lem} $\Omega_{u,v}(\pi_1(\bM))=\ZZ^2$.
+\end{lem}
+\begin{lem}
+\compav{show somehow that u,v determine the cover}
+\end{lem}
+
+\subsection{Induced Automorphisms of $H_1(\bM,\QQ)$}
+We look at some important automorphisms induced by Affine maps on $\bM$.
+Observe that $\bM$ has a uniform cylinder decomposition in both horizontal and vertical directions as in figure $15$.
+We define the \emph{modulus}, $\mu$, of a cylinder to be the ratio of the cylinder's width to its circumference, $\frac{w}{c}$.
+The $\emph{Dehn-twist}$ of a cylinder is an affine diffeomorphism that skews the cylinder and sends every vertex to itself.
+
+\begin{figure}[H]
+\centering
+\includesvg[width=3in]{cylinderskew}
+\label{fig:skew}
+\caption{Dehn-twist of a cylinder in $\bM$'s cylinder decomposition.}
+\end{figure}
+
+Such a map would have derivative $\left[~\begin{matrix}1 && \pm\mu^{-1}\\0 && 1\end{matrix}~\right]$.
+On $\bM$ every cylinder in both vertical and horizontal decompositions has a modulus of $\frac{1}{2}$.
+These give way to global diffeomorphisms as $\emph{multi-twists}$ of $\bM$.
+
+\begin{Def}
+We call the global affine diffeomorphisms in $\text{Aff}^+(\bM)$ obtained as \textbf{multi-twists of the surface in horizontal and vertical directions} $\mathbf{A}$ and $\mathbf{B}$, respectively.
+\\We define the \textbf{derivatives of} $\mathbf{A},\mathbf{B}$ as the matrices 
+$D(\mathbf{A})=\mathbf{A}'=\left[ \hspace{1mm} \begin{matrix}
+				1 &   2\\
+				0 & 1
+			\end{matrix}\hspace{1mm}\right] \text{ , and }
+			D(\mathbf{B})=\mathbf{B}'=\left[ \hspace{1mm} \begin{matrix}
+							1 & 0\\
+							 2 & 1
+						\end{matrix}\hspace{1mm}\right]$, where $D:\text{Aff}^+(\bM)\rightarrow V(\bM)$.
+\end{Def}
+These are parabolic elements of $V(\bM)$ that are directly related to the periodic/drift-periodic dichotomy of rational trajectories on the Necker cube surface.
+To show this directly, we look at how skewing the surface affects our core cylinder curves:
+
+\begin{figure}[H]
+\centering
+\includesvg[width=2in]{skew}
+\end{figure}
+
+When skewing the surface in the horizontal direction, the horizontal curves are preserved, but the vertical curves obtain two additional positive intersections with adjacent horizontal curves.
+We work out the $\mathbf{A}, \mathbf{B}$ induced automorphisms of $H_1(\bM,\QQ)$ to be:
+
+\begin{align*}
+\mathbf{A}^k_*([\gamma_i])=&[\gamma_i] + \frac{k}{2}(1-(-1)^i)([\gamma_{i-1}]+[\gamma_{i+1}])\\
+\mathbf{B}^k_*([\gamma_i])=&[\gamma_i] + \frac{k}{2}(1+(-1)^i)([\gamma_{i-1}]+[\gamma_{i+1}])\\
+\end{align*}
+
+\section{Proof of Main Theorem}
+
+\subsection{Translation Surfaces and $\ZZ^2$-Covers}
+We recall some definitions and properties of translation surfaces and their translation covers.
+A \emph{translation surface}, $S$, is a compact, flat surface that admits a discrete number of conical singularities and has trivial linear holonomy on arbitrary paths.
+We denote the surface without singularities by $S^\circ$.
+Translation surfaces belong to a stratum or family of surfaces characterized according to the number of conical points and their degrees.
+If the total angle of a closed loop around the singularity is given as the quantity $2\pi(d+1)$, we say the \emph{degree} of the singularity is $d$.
+A $\ZZ^2$-cover is a normal cover of $S$ given by the map $f:\tilde{S}\rightarrow S$, such that the associated deck group of the cover, $\Delta=\pi_1(S)/\pi_1(\tilde{S})$, is isomorphic to $\ZZ^2$.
+
+\begin{Def}
+Algebraic intersection number is a non-degenerate bilinear form:
+\begin{align*}
+i:H_1(S,\mathbf{R})\times H_1(S,\mathbf{R})\rightarrow\mathbf{R},
+\end{align*}
+where $\mathbf{R}$ is a ring and for $[\gamma],[\beta]\in H_1(S,\mathbf{R})$, $i([\beta],[\gamma])$ returns the intersection number of two homology classes.
+\end{Def}
+Algebraic intersections are signed and follow some convention such as the right-hand rule.
+
+\begin{Def}
+Let $u,v\in H_1(S,\QQ)$ be linearly independent homology classes of curves on $S$. Then the group homomorphism from $\pi_1(S)$ to $\QQ^2$ is given as:
+\begin{align*}
+\Omega_{u,v}:\pi_1(S)\rightarrow \mathbf{R}^2 \text{; } \beta\mapsto(i(u,[\beta]),i(v,[\beta])).
+\end{align*}
+\end{Def}
+
+The set of all orientation-preserving affine diffeomorphisms of $S$ forms the group Aff$^+(S)$.
+The corresponding \emph{Veech group}, $V(S)$ of $S$ is the image of the group morphism $D:\text{Aff}^+(S)\rightarrow SL(2,\mathbb{R})$ that takes an affine map to its derivative.
+A surface is said to be \emph{Veech} if its Veech group is commensurable to $SL(2,\mathbb{R})$.
+It is well known that origami, or square-tiled, surfaces have Veech groups commensurable to $SL(2,\ZZ)$.
+[cite] When $\mathbf{R}=\ZZ$, $\Omega_{u,v}$ takes an element of $\pi_1(S)$ to $\Delta$.
+Thus, $\gamma\in\pi_1(S)$ lifts to $\tilde{\gamma}\in\pi_1(\tilde{S})$ if and only if $\gamma\in\ker\Omega_{u,v}$.
+
+\begin{Def}
+Let $\alpha:[0,1]\rightarrow S$ be a closed, non-singular geodesic path on $S$.
+The holonomy map $\mathbf{hol}:H_1(S, \mathbf{R})\rightarrow\CC$ returns the holonomy vector of a closed path as a difference of the starting and endpoints of a flow by\\
+\begin{align*}
+\mathbf{hol}([\alpha])=\int_{\alpha}dz.
+\end{align*}
+\end{Def}
+Since $\alpha$ is non-singular, it can be mapped to $S^\circ$ which admits a flat holomorphic one-form $dz$.
+Let $\theta=Arg(\mathbf{hol}([\alpha]))$.
+We say that $\phi_t^\theta:\RR\times S^\circ \rightarrow S^\circ$ is the unit-speed geodesic flow on $S^\circ$ in direction $\theta$ given by the  $[\alpha]$ such that $\phi^\theta_0=\alpha(0)$.
+In local coordinates this corresponds to $z+te^{i\theta}\in \CC$.
+
+\begin{lem}
+$\phi_t^\theta$ has a period of $T=|\mathbf{hol}([\alpha])|$
+\begin{proof}
+This just follows from the fact that $\phi_t^\theta$ flows at unit-speed in the direction of $\frac{\mathbf{hol}([\alpha])}{|\mathbf{hol}([\alpha])|}$. 
+\end{proof}
+\end{lem}
+And so the length of a vector determines the period of a flow on $S$. More importantly, we have the following:
+
+\begin{lem}
+Denote the lifted flow of $\phi_t^\theta$ on $\tilde{S}$ by $\tilde{\phi}_t^\theta$.
+Then $\tilde{\phi}_t^\theta$ is periodic on $\tilde{S}$ and $\tilde{S}^\circ$ if and only if $[\alpha]\in\ker\Omega_{u,v}$.
+Further, $\tilde{\phi}_t^\theta$ has period $T=|\mathbf{hol}([\alpha])|$.
+\begin{proof}
+Suppose that $[\alpha]\notin\ker\Omega_{u,v}$.
+Since $\alpha([0,1])=\phi^\theta_{[0,\hol([\alpha])]}$, their homology classes are equivalent.
+Hence $\tilde{\phi}_t^\theta$ could not close on $\tilde{S}$ or $\tilde{S}^\circ$, or else $\Omega_{u,v}(k[\alpha])=k(i(u,[\alpha]),i(v,[\alpha]))=(0,0)$ for some non-zero $k\in\ZZ$.
+Conversely, suppose $[\alpha]\in\ker\Omega_{u,v}$.
+Then
+\end{proof}
+\end{lem}
+
+\begin{cor}
+If $[\alpha]\notin\ker\Omega_{u,v}$, then $\tilde{\phi}_t^\theta$ is drift-periodic with period $T=\hol([\alpha])$.
+\begin{proof}
+This follows immediately from the previous lemma since $\tilde{S},\tilde{S}^\circ$ have translational $\ZZ^2$ symmetries and a well-defined $\ZZ^2$ action on an element in the fiber of a basepoint in $S$ under $f$.
+The period is $T$ since a geodesic closes on $S$ with period $T$.
+\end{proof}
+\end{cor}
+
+\begin{lem}
+Let $h\in\text{Aff}^+(S)$. If $h(\beta)=\alpha$ for closed geodesics $\alpha,\beta$ on $S$ and $\beta$ lifts to a closed path on $\tilde{S}$, then so does $\alpha$. (not sure about this)
+\begin{proof}
+If the premise is true, then $[\beta]\in\ker\Omega_{u,v}$.
+Denote the group automorphism induced by $h$ as $h_*$.
+Then $\Omega_{u,v}([\alpha])=\Omega_{u,v}([h(\beta)])=\Omega_{u,v}(h_*\cdot[\beta])=(i([\beta],h_*^{-1}\cdot u),i([\beta],h_*^{-1}\cdot v))=(0,0)$ as automorphisms .
+\end{proof}
+\end{lem}
+
+\begin{lem}
+Let $\beta=h(\alpha)$ as before. If $D(h)=h'\in V(S)$ and $S$ is Veech, then $\tilde{\phi}^\theta_t$ has period $T=h'\cdot\hol([\beta])$. (also not sure)
+\begin{proof}
+
+\end{proof}
+\end{lem}
 
 \subsection{Symmetries of $\bM$}
-\begin{figure}[H]
-\centering
-\includesvg[width=2in]{skew}
-\end{figure}
 
@@
 A Dehn-twist on each cylinder in the cylinder decomposition of $\mathbf{M}$ in horizontal and vertical directions gives way to these global affine diffeomorphisms:
-\begin{figure}[H]
-\centering
-\includesvg[width=1.5in]{cylinderskew}
-\label{fig:skew}
-\end{figure}
 

### 2017-10-25 newpaper.tex
102517

@@
 \end{Def}
-These are parabolic elements of $V(\bM)$ that are directly related to the periodic/drift-periodic dichotomy of rational trajectories on the Necker cube surface.
-To show this directly, we look at how skewing the surface affects our core cylinder curves:
+These are parabolic elements of $V(\bM)$ that are directly related to the dynamical dichotomy of rational trajectories on the Necker cube surface.
+To show this algebraically, we look at how skewing the surface affects our homology spanning core curves:
 
@@
 
-When skewing the surface in the horizontal direction, the horizontal curves are preserved, but the vertical curves obtain two additional positive intersections with adjacent horizontal curves.
-We work out the $\mathbf{A}, \mathbf{B}$ induced automorphisms of $H_1(\bM,\QQ)$ to be:
+When skewing the surface under $\mathbf{A}$ in the horizontal direction, the horizontal curves are preserved, but the vertical curves (odd $\gamma_i$ index) obtain two additional positive intersections with adjacent index (mod 12) horizontal curves.
+Similarly, the vertical $\mathbf{B}$ skews preserve the vertical curves, but the horizontal curves (even $\gamma_i$ index) obtain two additional \emph{negative} intersections with adjacent index (mod 12) vertical curves.
+The formulaic expressions of the $\mathbf{A}, \mathbf{B}$ induced automorphisms of $H_1(\bM,\QQ)$ as powers of $k\in\ZZ$ are then:
 
@@
 
+With all of these in mind we generate the following subgroups:
+
+\begin{Def}
+We say $\XX$ is the subgroup of $\text{Aff}^+(\bM)$ generated by $\mathbf{A}$, $\mathbf{B}$.
+\end{Def}
+
+\begin{Def}
+The images of $\XX$ under their derivative maps to $V(\bM)$ is the subgroup of $V(\bM)$ generated by matrices $\mathbf{A}',$ $\mathbf{B}'$. We denote this group by $\XX'$.
+\end{Def}
+
+\begin{Def}
+The induced automorphisms of $\XX$ contained in $Aut(H_1(\bM,\QQ))$ is the subgroup generated by $\mathbf{A}_*$, $\mathbf{B}_*$.
+\end{Def}
 \section{Proof of Main Theorem}

### 2017-11-01 newpaper.tex
110117

@@
 \def\bM{\mathbf{M}}
+\def\bMs{\mathbf{M}^\circ}
 
@@
 \def\GL{\mathrm{GL}} \def\SL{\mathrm{SL}}  \def\SP{\mathrm{SL}}
+\def\SO{\mathrm{SO}}
 
@@
 \subsection{Discussion of Results}
-Initial experiments suggested there was a correlation between choice of trajectory angle and periodicity.
-On a surface composed of infinitely many cubes, there is certainly a flat metric endowed on smaller neighborhoods of the surface that are locally isometric to the plane.
-Via parallel transports of tangent vectors over the edges of the cubes connecting one face to the next, it is possible to describe geodesic behavior on the Necker cube surface as a series of straight line segments that traverse faces and continuously flow in a fixed direction (see figure $\ref{fig:front}$).
-\\\\
+Our initial experiments strongly supported the theory that there would be a correlation between a choice of trajectory angle and dynamical properties of a geodesic on such a symmetric object.
+Rightfully so, a surface composed of infinitely many cubes an induced flat metric where every neighborhood is locally isometric to the Euclidean plane.
+Via the parallel transport of a unit tangent vector over the sharp edges of the surface, it becomes obvious to think of the geodesic as a sequence of line segments contained in the faces of each cube on the surface's embedded form in $\RR^3$.
+$\ref{fig:front}$.
 
@@
 \end{wrapfigure}
-We consider classes of directions obtained from unit vectors in $\RR^3$ by projecting a sufficiently small segment of a flow contained onto a parallel plane.
-These vectors retain all necessary information about their directions and let us consider rational and irrational angles relative to a choice of basis in $\RR^2$.
-This is what we refer to as the \emph{initial trajectory angle}.
+\noindent Let $x_0$ be a point on the surface that is contained in an open neighborhood of a smooth section of the surface and consider a tangent unit vector $v\in\RR^3$ protruding from $x_0$.
+Assuming the faces are parallel to every $2$-dimensional subspace of $\RR^3$ spanned by standard basis vectors, exactly one component of $v$ would be 0.
+Projecting this vector to a parallel plane retains all necessary information about direction.
+Call this vector $v_0\in\RR^2$.
 \\\\
-Rational component trajectory angles are then further categorized according to the parities of their components.
-\begin{Def}
-Let $v$ be a unit vector of the form $\frac{1}{k}(x,y)\in\RR^2$ for $x,y\in\mathbb{Z}$ and $k=\sqrt{x^2+y^2}\in\mathbb{R}$.
-We call $v$ an \textbf{odd-odd} vector if x,y are relatively prime and both odd.
+We say that $v_0$ is the \emph{initial trajectory} of the geodesic and the angle it makes relative to our choice of basis is its \emph{initial trajectory angle}.
+We consider a rational direction to fall into one of two categories.
+\begin{Def}
+Let $v_0$ be a unit vector of the form $\frac{1}{k}(x,y)\in\RR^2$ with $x,y\in\mathbb{Z}$ and $k=\sqrt{x^2+y^2}\in\mathbb{R}$.
+We say $v_0$ is an \textbf{odd-odd} vector if its components are relatively prime and odd.
 We denote the \textbf{set of all odd-odd directions} by $\mathcal{O}$.
-We say that $v$ is an \textbf{even-odd} vector if x and y are relatively prime, and x is even if and only if y is odd.
+We say that $v_0$ is an \textbf{even-odd} vector if its components are relatively prime and of opposite parity.
 We denote the \textbf{set of all even-odd directions} by $\mathcal{E}$.
 \end{Def}
-We prove the following is true of all rational directions:
+\noindent This paper will demonstrate how one reaches the following conclusion about the relationship between parity and periodicity of geodesic behavior on the Necker cube surface, denoted $\mathbf{S}$:
 
 \begin{thm*}
-Let $\Phi_t$ be the geodesic flow defined on the unit tangent bundle of the Necker cube surface for $t\in\RR$, and initial point $p=(x_0,v_0)$ with $v_0\in\RR$. Then the following is true:
+(Directional Dichotomy of Periodic Geodesics on $\mathbf{S}$) Let $\Phi_t$ be a unit-speed geodesic flowing on the unit tangent bundle of $\mathbf{S}$ for $t\in\RR$ with initial point and vector $\Phi_0=(x_0,v)$.
+Then the following is true:
 \begin{enumerate}[label=(\roman*)]
-\item $\Phi$ is periodic with period $T\in\RR$ if and only if $v\in\cO$.
-\item $\Phi$ is drift-periodic with period $T\in\RR$ if and only if $v\in\cE$.
+\item $\Phi$ is periodic with period $T\in\RR$ if and only if $v_0\in\cO$.
+\item $\Phi$ is drift-periodic with period $T\in\RR$ if and only if $v_0\in\cE$.
 \end{enumerate}
@@
 
-\section{Construction}
-This section details how we construct the Necker cube surface as an infinite cover of a compact Riemann surface.
-\subsection{Geometric Interpretation}
-Consider the following surface:
+\section{Periodic Tiling of Necker Cubes}
+This section will detail how the Necker Cube surface is obtained as an infinite regular cover of a branched torus. Consider the Necker cube with the following identifications:
 \begin{figure}[H]
@@
 
-We call this surface $\bG$, the \emph{Necker cube}.
-$\bG$ is obtained by taking three copies of the unit square and identifying the edges in order to resemble the visible portion of a cube.
-$\bG$ has a total of three vertices.
-Every vertex has a cone angle of either $3\pi$ or $\frac{3\pi}{2}$ and are \emph{conical points} (or \emph{singularities}) of the surface.
-We denote the surface without singularities by $\bGs$.
-The periodic tiling of $\bG$ will take infinitely many copies of $\bG$ and tile them along the edges labeled with solid triangles.
-This cover of $\bG$ is the Necker cube surface we denote by $\mathbf{S}$.
-Likewise, the surface tiled by $\bGs$ we denote by $\mathbf{S}^\circ$\\
-
-\noindent Take the following unit squares in $\RR^3$:\\
+We denote this surface $\bG$.
+$\bG$ is obtained by taking three copies of the unit square and identifying the edges in order to resemble the visible portion of a unit cube.
+$\bG$ is a genus one surface with three vertices homeomorphic to $\TT^2$.
+Check that every vertex has a cone angle of either $3\pi$ or $\frac{3\pi}{2}$.
+These are the conical singularities of the surface, $\Sigma$.
+We denote the surface without its singularities $\bG\backslash\Sigma = \bGs$.
+Thus, $\pi_1(\bGs)\cong\pi_1((thrice\text{-}branched~torus))\cong\pi_1(S^1\vee S^1\vee S^1\vee S^1)\cong\FF_4$, the free group of four generators.
+\begin{Def}
+The \textbf{Necker cube surface} $\mathbf{S}$ is obtained as the regular cover of $\bG$ by periodically tiling along the edges with solid triangles.\\
+The \textbf{branched Necker cube surface} $\mathbf{S}^\circ$ is obtained similarly by tiling $\bGs$.
+\end{Def}
+
+\noindent Consider the following families of unit squares in $\RR^3$:\\
 \vspace{0.2in}
@@
 
-\noindent If we take the integer triple $(m,n,p)$ and require that the individual faces based around them  has the triple lie on the plane $x+y+z=0$, $\mathbf{S}$ is broken down into the following sets:
-
+With the proper identifications on  $\mathbf{A}_{0,0,0}\cup\mathbf{B}_{0,0,0}\cup\mathbf{C}_{0,0,0}$, we have $\bG$'s embedded form in $\RR^3$ and a map $s:\bG\hookrightarrow\RR^3$.
+Similarly, removing the integer points from this set is an embedding of $\bGs$.
+Check that tiling along the edges as previously described determines a set of points in $\RR^3$ that coincides with points on $\mathbf{S}$ with $s^*$ as an extension of $s$ to the cover.
+That is, $s(\mathbf{S})=\bigcup\left\{\mathbf{A}_{m,n,p}\cup\mathbf{B}_{m,n,p}\cup\mathbf{C}_{m,n,p}~:~m+n+p=0\right\}$.
+Thus a point $x_0\in\mathbf{S}$ is in one-to-one correspondence with this set.
+To avoid ambiguity we say $U$ is a \emph{neighborhood} of $s(x_0)$ if it is contained in exactly one of $\mathbf{A}_{m,n,p}, \mathbf{B}_{m,n,p}, \mathbf{C}_{m,n,p}$.
+As $U$ must contain an open ball contained on a single square there cannot be a neighborhood where $s(x_0)\in\ZZ^3$ and $x_0$ is a conical point.
+It follows then that a neighborhood in $\mathbf{S}$ is a neighborhood in $\mathbf{S}^\circ$ as well.
+
+\begin{figure}[H]
+\centering
+\includesvg[width=1in]{cubesxyzcut}
+\caption{A section of $\mathbf{S}$}
+\end{figure}
+
+\subsection{Flattening $\mathbf{S}$}
+At the moment it is difficult to describe how the parallel transport of a vector tangent to the surface along an arbitrary path on the surface acts on the vector in $\RR^3$.
+To simplify this problem we take $\mathbf{S}$ to an isometric variant embedded in $\RR^2$ by piecewise linear transformations on the sets
 \begin{align*}
@@
 \\\mathbf{C}=\bigcup\big{\{}\mathbf{C}_{m,n,p}: m+n+p=0\big{\}}.
-\end{align*}\\
-The sets $\mathbf{A},\mathbf{B},\mathbf{C}$ are contained in $\mathbf{S}$. Their union is also the Necker cube surface.
-
-\begin{figure}[H]
-\centering
-\includesvg[width=2in]{cubesxyzcut}
-\caption{The tiling of $\bG$ where the $\mathbf{A}$ faces are copies of copies of the front-facing square of $\bG$.(Edges that were not used in the tiling are labeled).}
-\end{figure}
-
-\subsection{Algebraic Interpretation}
-We use a series of cutting and gluing operations to obtain an L-shaped surface homeomorphic to the torus branched at three points:\\
-\begin{figure}[H]
-\centering
-\includesvg[width=3in]{LTorus}
-\end{figure}
-\noindent Which is then cut along these dotted lines and pieced together to resemble the torus with a unit-square removed from the center:\\
-\begin{figure}[H]
-\centering
-\includesvg[width=3in]{BranchTorus}
-\caption{$\bGs$ with 4 independent, closed paths labeled identified}
-\end{figure}
-The labeled paths are independent from eachother, and a quick calculation shows us that its fundamental group $\pi_1(\bGs)\cong \<a,b,c,d\>$, the free group of four generators (denote identity by $\mathbf{1}$).
-We describe an infinite cover of $\bGs$ such that the associated deck group is isomorphic to $\ZZ^2$
-\begin{Def}
-We define $\varphi_1:\pi_1(\bGs)\rightarrow \ZZ^2$ as the group homomorphism on generators $a,b,c,d$: \\$c,d\mapsto(0,0)$\\ $a\mapsto(1,0)$\\$b\mapsto(0,1)$\\ $ab,ba\mapsto(1,1)$.
-\end{Def}
-
-\begin{Def}
-$\bUs$ is the cover of $\bGs$ such that $\pi_1(\bUs)=\ker\varphi$.
-\end{Def}
-
-Our cover of $\bGs$ has the fundamental group, $\pi_1(\bUs)=\<\<a^{-1}b^{-1}ab, c, d\>\>$, the conjugate subgroup of elements that map trivially under $\varphi_1$.
-The covering map $k_{\varphi_1}:\bUs\hookrightarrow\bGs$ takes every point on $\bUs$ to its modular equivalent under these translational symmetries.
-$\bUs$ is realized as an infinite $\ZZ^2$-tiling of $\bGs$ in $\CC$:
-\begin{figure}[H]
-\centering
-\includesvg[width=2in]{uprime}
-\caption{The infinite surface $\bUs$ embedded in $\CC$ with edges identified.}
-\end{figure}
-
-\begin{Def}
-$\bU$ is the completion of $\bUs$ that includes the branched vertices. 
-\end{Def}
-
-\begin{rem}
-The surface $\bG$ is homeomorphic to the torus, and its cover $\bU$ is its universal cover as the paths labeled $c,d$ are only non-trivial when $\bG$ is branched at its three conical points.
-\end{rem}
-
-\subsection{Isometry From $\mathbf{S}$ to $\bU$}
-The Necker cube surface can be \emph{flattened} onto the plane by piecewise isometric maps onto a subspace of $\RR^3$ with (open) unit squares removed at every even integer pair in the plane, what we claim to be $\bU$.
-\begin{figure}[H]
-\centering
-\includesvg[width=1.1in]{cubesxyzcut} \hspace{0.1in}\raisebox{1.0in}{\text{$\rightarrow$}}\hspace{0.1in}
-\raisebox{0.4in}{\includesvg[width=1.5in]{unfoldcut}}
-\caption{A piecewise isometry from $\mathbf{S}$ to $\bU$. }
-\label{fig:Psi}
-\end{figure}
-\noindent The red/blue dotted lines represent the edges that are split on the plane. The map from one surface to the other is composed of piecewise isometries, $\Psi:\mathbf{S}\rightarrow\RR^3$.
-
+\end{align*}
+
+\begin{Def}
+Let $\Psi:s^*(\mathbf{S})\rightarrow\RR^3$ be given as
 \begin{equation}
@@
 \end{equation}
+\end{Def}
+\noindent This map behaves as such:
+\begin{figure}[H]
+\centering
+\includesvg[width=1.1in]{cubesxyzcut} \hspace{0.1in}\raisebox{1.0in}{\text{$\rightarrow$}}\hspace{0.1in}
+\raisebox{0.4in}{\includesvg[width=1.5in]{unfoldcut}}
+\caption{An isometry of the surface $\mathbf{S}$.}
+\label{fig:Psi}
+\end{figure}
+
+\begin{lem}
+$\Psi\circ s^*$ is an isometry of the surface $\mathbf{S}$
+\begin{proof}
+$\Psi$ is well-defined on its domain $s^*(\mathbf{S})$, and $s^*$ is an embedding of the surface that is a bijection when restricting its co-domain to its image.
+Since $\Psi$ is a linear transformation composed of Euclidean matrices and translations it is also invertible and bijective.
+\end{proof}
+\end{lem}
+
+\subsection{Translation Surface Cover of $\bGs$}
+Reconstruct $\bG$ and $\bGs$ to resemble a quotient of the flattened Necker cube surface by a series of cutting and gluing operations: \\
+\begin{figure}[H]
+\centering
+\includesvg[width=3in]{LTorus}
+\end{figure}
+\noindent This L-shaped torus is then cut along the dotted lines and pieced together to look like a $2\times2$ torus with a unit square removed from the center:\\
+\begin{figure}[H]
+\centering
+\includesvg[width=2in]{BranchTorus}
+\caption{$\bGs$ with 4 independent, closed paths labeled identified}
+\end{figure}
+The paths labeled $a,b,c,d$ are independent.
+Let $\pi_1(\bGs)=\<a,b,c,d\>$.
+$\bGs$ has a $\ZZ^2$ tiling by traversing along paths $a,b$ and attaching infinitely many copies of the surface.
+It's apparent that this $\ZZ^2$ tiling is the flattened Necker cube surface.
+We use group homomorphisms on this free group to construct translation surfaces and $\ZZ^2$-covers to better understand geodesic behavior on the Necker cube surface.
+
+\begin{Def}
+Let $\varphi_1:\pi_1(\bGs)\rightarrow \ZZ^2$ be the group homomorphism defined on generators $a,b,c,d$: \\$c,d\mapsto(0,0)$\\ $a\mapsto(1,0)$\\$b\mapsto(0,1)$\\ $ab,ba\mapsto(1,1)$
+\end{Def}
+
+\begin{Def}
+Let $\varphi_2:\pi_1(\bGs)\rightarrow SO(2,\ZZ)$ be the group homomorphism defined on the generators $a,b,c,d$:
+\begin{align*}
+a,b & \mapsto \begin{pmatrix}
+1 && 0 \\ 0 && 1\end{pmatrix}=I_2\\
+c & \mapsto\begin{pmatrix}0 && -1\\1 && 0\end{pmatrix}=R\\ d & \mapsto\begin{pmatrix}0 && 1\\-1 && 0\end{pmatrix}=R^3
+\end{align*}
+\end{Def}
+
+Recall that a \emph{translation surface} is a compact, locally flat surface that admits a discrete number of conical singularities and has trivial linear holonomy on any arbitrary path.
+If the total cone angle of the singularity is some integer multiple of $2\pi$, $2\pi(d+1)$ for $d\geq 0$, we say that singularity has degree $d$.
+Two translation surfaces belong to the same stratum, or family, of surfaces characterized by the number and degree of conical singularities.
+Consider the effect that paths $c,d$ have on holonomy:
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=2.5in]{monodromy.png}
+\caption{Effect a nontrivial loop has on arbitrary basis vectors (i,j).}
+\label{fig:loop}
+\end{figure}
+
+We construct a cover of $\bGs$ by taking four copies of it and cyclically pairing edges, and show that this cover has trivial holonomy.
+Denote the unit tangent bundle of $\bGs$ by $T^1\bGs$.
+Because $\bGs$ can be contained in the plane, we consider vectors tangent to $\bGs$ to be in $\RR^2$.
+Let $x_0\in\bGs$.
+The fiber over $x_0$ under the projection $T^1\bGs\twoheadrightarrow\bGs$ is the set $E_{x_0}$.
+Let $\triangledown$ be a flat metric connection on $T^1\bGs$ and $p=(x_0,v_0)\in E_{x_0}$ be a single element in the fiber over $x_0$ with $v_0\in\RR^2$.
+We denote the \emph{holonomy group} of the connection with base point $p$ by $Hol_p(\triangledown)=\{A\in SO(E_{x_0})~:~ p = A\cdot p\}$ where $p$ differs from $A\cdot p$ by some rotation of $v_0$.
+An element in $SO(E_{x_0})$ acts on $p$ by rotation of the vector $v_0$.
+Denote the \emph{holonomy bundle} based at $p$ by $H(p)=\{p'\in E_{x_0}~:~ p'= A\cdot p \}$, the set of elements in the orbit of $p$ where a closed loop $\gamma:[0,1]\rightarrow\bGs$ defines a horizontal lift to $\gamma^*:[0,1]\rightarrow T^1\bGs$ by parallel transport of $v_0$ along the path such that $p=\gamma^*(0)$, and $p'=\gamma^*(1)$.
+Since $\bGs$ is flat and trivial paths have trivial holonomy, $Hol_p(\triangledown)$ acts on $H(p)$ the monodromy group is discrete and well-defined on $\bGs$.
+
+\begin{lem}
+$\varphi_2:\pi_1(\bGs)\rightarrow \mathrm{SO}(2,\ZZ)$ is a monodromy representation of the surface.
+\begin{proof}
+It is clear from Figure 6 that this is the case as concatenating a closed loop with paths $c$ or $d$ would rotate a vector exactly by the cone angles of the singularities that they loops around: $\pm \frac{3\pi}{2}$.
+\end{proof}
+\end{lem}
+
+\begin{Def}
+Let $\bMs$ be the regular cover of $\bGs$ with fundamental group $\pi_1(\bMs)=\ker\varphi_2$.
+\end{Def}
+Showing that $\bMs$ is a degree four cover of $\bGs$ with trivial holonomy and deck group $\SO(2,\ZZ)$ follows immediately from its construction.
+\begin{figure}[H]
+\centering
+\includegraphics[width=3in]{monogroup.png}
+\label{fig:arbitrarylift}
+\caption{Lift of paths $c,d$ to $\bMs$ (\compav{RELABEL EDGES})}
+\end{figure}
+
+\begin{Def}
+The surface $\bUs$ is the regular cover of $\bGs$ such that $\pi_1(\bUs)=\ker\varphi_1$. 
+\end{Def}
+
+The deck group of this cover is isomorphic to $\ZZ^2$.
+By reversing the cutting and gluing done previously to $\bGs$ and $\bG$, one can easily see that a cover defined as the kernel of this homomorphism is identical to the tiling of Necker cubes along two of its edges.
+For this reason, $\bU\cong\mathbf{S}$.
+
+\subsection{A Four-Fold Cover of $\bUs$}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=3in]{coverdirection.png}
+\end{figure}
+
+\newpage
+\begin{Def}
+$\bUs$ is the cover of $\bGs$ such that $\pi_1(\bUs)=\ker\varphi$.
+\end{Def}
+
+Our cover of $\bGs$ has the fundamental group, $\pi_1(\bUs)=\<\<a^{-1}b^{-1}ab, c, d\>\>$, the conjugate subgroup of elements that map trivially under $\varphi_1$.
+The covering map $k_{\varphi_1}:\bUs\hookrightarrow\bGs$ takes every point on $\bUs$ to its modular equivalent under these translational symmetries.
+$\bUs$ is realized as an infinite $\ZZ^2$-tiling of $\bGs$ in $\CC$:
+\begin{figure}[H]
+\centering
+\includesvg[width=2in]{uprime}
+\caption{The infinite surface $\bUs$ embedded in $\CC$ with edges identified.}
+\end{figure}
+
+\begin{Def}
+$\bU$ is the completion of $\bUs$ that includes the branched vertices. 
+\end{Def}
+
+\begin{rem}
+The surface $\bG$ is homeomorphic to the torus, and its cover $\bU$ is its universal cover as the paths labeled $c,d$ are only non-trivial when $\bG$ is branched at its three conical points.
+\end{rem}
+
+\subsection{}
+
+\subsection{Isometry From $\mathbf{S}$ to $\bU$}
+The Necker cube surface can be \emph{flattened} onto the plane by piecewise isometric maps onto a subspace of $\RR^3$ with (open) unit squares removed at every even integer pair in the plane, what we claim to be $\bU$.
+
+\noindent The red/blue dotted lines represent the edges that are split on the plane. The map from one surface to the other is composed of piecewise isometries, $\Psi:\mathbf{S}\rightarrow\RR^3$.
 
@@
 Observe that the parallel transport of a vector around a closed loop on $\bUs$ will act on vectors tangent to the surface by a rotation of an integer multiple of $\frac{\pi}{2}$ radians (since the surface is flat and embedded in $\CC$, we work with vectors in $\RR^2$).
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=2.5in]{monodromy.png}
-\caption{Effect a nontrivial loop has on arbitrary basis (i,j).}
-\label{fig:loop}
-\end{figure}
-
-Let $\triangledown$ be a flat, Riemannian metric connection on $T^1\bUs$.
-This is allowable since $\bUs$ is obtained as the cover of a flat Riemann surface (\compav{you said of a translation surface?}).
-Let $p=(x_0,v_0)\in E_{x_0}\subset T^1\bUs$ be an element in the fiber over $x_0$ the usual projection from the tangent bundle to the surface, with $x_0\in\bUs$ and $v_0\in\RR^2$.
-Denote the \emph{holonomy group} of the connection with base point $p$ by $Hol_p(\triangledown)=\{A\in SO(E_{x_0})~:~ p \equiv A\cdot p\}$ where $p$ differs from $A\cdot p$ by some rotation of $v_0$.
-That is, $p\equiv p'$ if $p'=(x_0,A v_0)$ for $A\in Hol_p(\triangledown)$.
-Denote the \emph{holonomy bundle} based at $p$ by $H(p)$, the set of points $p'\in E_{x_0}$ where a closed loop $\gamma:[0,1]\rightarrow\bUs$ describes a horizontal lift to $\gamma^*:[0,1]\rightarrow T^1\bUs$ such that $p=\gamma^*(0)$ and $p'=\gamma^*(1)$.
 
@@
 A visual representation of $\btUos$ is as a four-fold cover of $\bUs$ with trivial holonomy on all closed paths. Arbitrary paths do \emph{not} have trivial holonomy:
-\begin{figure}[H]
-\centering
-\includegraphics[width=3in]{monogroup.png}
-\label{fig:arbitrarylift}
-\caption{Arbitrary lifts to fibers }
-\end{figure}
 
@@
 
-\begin{figure}[H]
-\centering
-\includegraphics[width=4.2in]{coverdirection.png}
-\end{figure}
-
 \noindent and its subsequent projection onto $\bUs:$
@@
 \subsection{Translation Surfaces and $\ZZ^2$-Covers}
-We recall some definitions and properties of translation surfaces and their translation covers.
-A \emph{translation surface}, $S$, is a compact, flat surface that admits a discrete number of conical singularities and has trivial linear holonomy on arbitrary paths.
-We denote the surface without singularities by $S^\circ$.
-Translation surfaces belong to a stratum or family of surfaces characterized according to the number of conical points and their degrees.
-If the total angle of a closed loop around the singularity is given as the quantity $2\pi(d+1)$, we say the \emph{degree} of the singularity is $d$.
-A $\ZZ^2$-cover is a normal cover of $S$ given by the map $f:\tilde{S}\rightarrow S$, such that the associated deck group of the cover, $\Delta=\pi_1(S)/\pi_1(\tilde{S})$, is isomorphic to $\ZZ^2$.
 

### 2017-11-13 newpaper.tex
110617
111217

@@
 \usepackage{wrapfig}
+\usepackage[arrow]{xy}
+\usepackage{tikz-cd}
 
@@
 \def\bMs{\mathbf{M}^\circ}
+\def\btUs{\btU^\circ} 
+\def\xtild{\tilde{x}_0}
+\def\utild{\tilde{u}_0}
+\def\gamtild{\tilde{\gamma}}
 
@@
 \centering
-\includesvg[width=2in.]{neckercube}
+\includesvg[width=1.6in.]{neckercube}\includesvg[width=2in.]{neckercubepath}
 \end{figure}
@@
 We denote the surface without its singularities $\bG\backslash\Sigma = \bGs$.
-Thus, $\pi_1(\bGs)\cong\pi_1((thrice\text{-}branched~torus))\cong\pi_1(S^1\vee S^1\vee S^1\vee S^1)\cong\FF_4$, the free group of four generators.
-\begin{Def}
-The \textbf{Necker cube surface} $\mathbf{S}$ is obtained as the regular cover of $\bG$ by periodically tiling along the edges with solid triangles.\\
-The \textbf{branched Necker cube surface} $\mathbf{S}^\circ$ is obtained similarly by tiling $\bGs$.
-\end{Def}
-
+Thus, $\pi_1(\bGs)\cong\pi_1((\TT^2 \text{ with 3 punctures}))\cong\pi_1(S^1\vee S^1\vee S^1\vee S^1)\cong\FF_4$, the free group of four generators.
+We may flatten $\bG$ and $\bGs$ into an L-shaped torus: \\
+\begin{figure}[H]
+\centering
+\includesvg[width=3in]{LTorus}
+\end{figure}
+\noindent This L-shaped torus is then cut along the dotted lines and pieced together to resemble a $2\times2$ torus with a unit square removed from the center:\\
+\begin{figure}[H]
+\centering
+\includesvg[width=2in]{BranchTorus}
+\caption{$\bGs$ with 4 independent, closed paths labeled identified}
+\end{figure}
+Denote its fundamental group by $\pi_1(\bGs)=\<a,b,c,d\>$, where $a,b,c,d$ are the labeled independent paths on the surface. Consider the following map:
+
+\begin{Def}
+Let $\varphi_1:\pi_1(\bGs)\rightarrow \ZZ^2$ be the group homomorphism defined on generators $a,b,c,d$: \\$c,d\mapsto(0,0)$\\ $a\mapsto(1,0)$\\$b\mapsto(0,1)$
+\end{Def}
+
+\begin{Def}
+The punctured \textbf{Necker Cube Surface} is a regular cover of $\bGs$, and is denoted $\bUs$. It's fundamental group presentation is the subgroup of $\pi_1(\bGs)$, $\pi_1(\bUs)=\ker\varphi_1$.
+\end{Def}
+
+Thus $\Delta_{\varphi_1}=\pi_1(\bGs)/\pi_1(\bUs)\cong \varphi_1(\pi_1(\bGs))=\ZZ^2$ is the associated \emph{deck group} of the cover. $\bUs$ is the primary object of study.
+
+\subsection{Monodromy Group Representation}
+In this section we describe a translation cover of $\bGs$ which and $\bUs$ which is factored through as the kernel of a homomorphism from $\pi_1(\bGs)\rightarrow \SO(2,\ZZ)$.
+This homomorphism represents the monodromy group as the fundamental group acting on the unit tangent bundles $T^1\bGs$ and $T^1\bUs$.
+
+\begin{Def}
+Let $\varphi_2:\pi_1(\bGs)\rightarrow SO(2,\ZZ)$ be the group homomorphism defined on the generators $a,b,c,d$:
+\begin{align*}
+a,b & \mapsto \begin{pmatrix}
+1 && 0 \\ 0 && 1\end{pmatrix}=I_2\\
+c & \mapsto\begin{pmatrix}0 && -1\\1 && 0\end{pmatrix}=R\\ d & \mapsto\begin{pmatrix}0 && 1\\-1 && 0\end{pmatrix}=R^3
+\end{align*}
+\end{Def}
+
+Recall that a \emph{translation surface} is a compact, locally flat surface that admits a discrete number of conical singularities and has trivial linear holonomy on any arbitrary path.
+If the total cone angle of the singularity is some integer multiple of $2\pi$, $2\pi(d+1)$ for $d\geq 0$, we say that singularity has degree $d$.
+Two translation surfaces belong to the same stratum, or family, of surfaces characterized by the number and degree of conical singularities.
+Consider the effect that paths $c,d$ have on holonomy:
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=2.5in]{monodromy.png}
+\caption{Effect a nontrivial loop has on arbitrary basis vectors (i,j).}
+\label{fig:loop}
+\end{figure}
+
+We construct a cover of $\bGs$ by taking four copies of it and cyclically pairing edges, and show that this cover has trivial holonomy.
+Denote the unit tangent bundle of $\bGs$ by $T^1\bGs$.
+Because $\bGs$ can be contained in the plane, we consider vectors tangent to $\bGs$ as vectors in $\RR^2$.
+Let $x_0\in\bGs$.
+The fiber over $x_0$ under the projection $T^1\bGs\twoheadrightarrow\bGs$ is the set $E_{x_0}$.
+Let $\triangledown$ be a flat metric connection on $T^1\bGs$ and $p=(x_0,v_0)\in E_{x_0}$ be a single element in the fiber over $x_0$ with $v_0\in\RR^2$.
+We denote the \emph{holonomy group} of the connection with base point $p$ by $Hol_p(\triangledown)=\{A\in SO(E_{x_0})~:~ p = A\cdot p\}$ where $p$ differs from $A\cdot p$ by some rotation of $v_0$.
+An element in $SO(E_{x_0})$ acts on $p$ by rotation of the vector $v_0$.
+Denote the \emph{holonomy bundle} based at $p$ by $H(p)=\{p'\in E_{x_0}~:~ p'= A\cdot p \}$, the set of elements in the orbit of $p$ where a closed loop $\gamma:[0,1]\rightarrow\bGs$ defines a horizontal lift to $\gamma^*:[0,1]\rightarrow T^1\bGs$ by parallel transport of $v_0$ along the path such that $p=\gamma^*(0)$, and $p'=\gamma^*(1)$.
+Since $\bGs$ is flat and trivial paths have trivial holonomy, $Hol_p(\triangledown)$ acts on $H(p)$ monodromy group is discrete and well-defined on $\bGs$.
+
+\begin{lem}
+$\varphi_2:\pi_1(\bGs)\rightarrow \mathrm{SO}(2,\ZZ)$ is a monodromy representation of the surface.
+\begin{proof}
+It is clear from Figure 6 that this is the case as concatenating a closed loop with paths $c$ or $d$ would rotate a vector exactly by the cone angles of the singularities that they loops around: $\pm \frac{3\pi}{2}$.
+\end{proof}
+\end{lem}
+
+\begin{Def}
+Let $\bMs$ be the regular cover of $\bGs$ with fundamental group $\pi_1(\bMs)=\ker\varphi_2$.
+\end{Def}
+Showing that $\bMs$ is a degree four cover of $\bGs$ with trivial holonomy and deck group $\SO(2,\ZZ)$ follows immediately from its construction.
+\begin{figure}[H]
+\centering
+\includegraphics[width=3in]{monogroup.png}
+\label{fig:arbitrarylift}
+\caption{Lift of paths $c,d$ to $\bMs$ (\compav{RELABEL EDGES})}
+\end{figure}
+
+\begin{Def}
+Let $\btUs$ be the cover of $\bUs$ with fundamental group $\pi_1(\btUs)=\ker\varphi_2\big\vert_{\pi_1(\bUs)}$.
+\end{Def}
+
+\compav{Prove that $\ker\varphi_1\big\vert_{\pi_1(\bMs)}=\ker\varphi_2\big\vert_{\pi_1(\bUs)}$}
+
+$\btUs$ is also a cover of $\bMs$ with fundamental group $\ker\varphi_1\big\vert_{\pi_1(\bMs)}$.
+In the former case the deck group of the cover is $\SO(2,\ZZ)$, while in the latter case it is $\ZZ^2$.
+The deck group of the cover $\btUs\rightarrow\bGs$ is $\ZZ^2\rtimes\SO(2,\ZZ)$.
+\compav{worth talking about?}.
+$\btUs$ is a four-fold cover of $\bUs$ with trivial holonomy visualized as four copies of $\bUs$, which can be thought of as countably many copies of .
+We can identify a closed path on $\bGs$ as an element of the deck group $\ZZ^2\times \SO(2,\ZZ)$ or, equivalently, $ 2\ZZ+2\ZZ i\times \ZZ/4\ZZ$ if we want to consider a metric on the ambient space $\CC$ that is consistent with the tiling of $\bGs$.
+Further, we can rotate everything in $\CC\times \ZZ/4\ZZ$ along with $\btUs$ according to the element of $\SO(2,\ZZ)$ that acts non-trivially on the holonomy of $\bUs$.
+For example, let $P\in2\ZZ+2\ZZ i$ and consider the following curve on $\bUs$:
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=2in]{overlay.png}
+\end{figure}
+
+\noindent If we lift this curve to $\btUs$ we can identify it with an element of the deck group isomorphic to $2\ZZ+2\ZZ i \times \ZZ/4\ZZ$.
+Observe that it will only close if the curve acts trivially on the holonomy of $\bUs$:
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=3in]{coverdirection.png}
+\end{figure}
+
+\subsection{Geodesics on $\bUs$}
+The key connection made between surfaces $\bUs$ and $\btUs$ is that geodesics on $\bUs$ and $\btUs$ are identical in the sense that their images are in a one-to-one correspondence.
+Unless stated otherwise, a \emph{geodesic} is a \emph{unit-speed} locally distance minimizing curve on a smooth topological surface $\gamma:[0,\infty)\rightarrow S$ such that $\gamma'(0)=v_0\in\RR^2$ and $|| v_0||=1$.
+A geodesic is \emph{periodic} if for some $T\in\RR^+$ $\gamma(t)=\gamma(t+T)$ for all $t$.
+A geodesic is \emph{drift-periodic} if for some $T\in\RR^+$, there is some non-trivial translation of the surface that takes $\gamma(t)\mapsto \gamma(t+T)$ for all $t$.
+In either case we say that $T$ is the \emph{period} or \emph{length} of a segment; in the case where a geodesic does have a period, we restrict $\gamma$ to the closed interval $[0,T]$ and call this a \emph{geodesic segment}.
+In this paper we do not consider cases where a geodesic hits a cone singularity.
+For example, the following is a periodic geodesic on the surface with initial direction $kv_0\in\cO$:
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=4in.]{closed2.png}
+\caption{Periodic geodesic path on $\bUs$ modeled with sage-flatsurf.}
+\label{fig:complicated}
+\end{figure}
+
+We refer later to the following commuting diagram of branched covering maps and inclusion maps. It is meant to illustrate the relationship between all of these spaces.
+
+\begin{equation}
+\begin{tikzcd}
+\arrow[two heads, bend left]{rrr}{f^\circ}[swap]{} \arrow[hook]{d}{pr_b} \btUs \arrow[two heads]{r}{r^\circ} & \bUs \arrow[hook]{d}{pr_c} \arrow[two heads]{r}{q^\circ} & \bGs \arrow[hook]{d}{pr_d} &\arrow[two heads]{l}{w^\circ} \bMs \arrow[hook]{d}{pr_a}  \\
+\arrow[two heads, bend right]{rrr}{}[swap]{f} \btU \arrow[two heads]{r}{r} & \bU \arrow[two heads]{r}{q} & \bG & \arrow[two heads]{l}{w}\bM  \\
+\end{tikzcd}
+\label{eq:commute}
+\end{equation}
+
+Let $x_0\in G^\circ$ and fix $u_0\in q^{\circ-1}(x_0)$ Let $\alpha:[0,b]\rightarrow\bGs$ be a closed geodesic segment on $\bGs$ with initial and terminal point $x_0$, and $\alpha'(0)=v_0$.
+Denote $\alpha$'s unique lift to $\bUs$ under $q^\circ$ as $\gamma:[0,b]$, where $\gamma(0)=u_0$, $\gamma'(0)=\gamma'(b)=v_0$, and $\gamma(b)\in q^{\circ-1}(x_0)$.
+Fix an arbitrary point $\utild\in r^{\circ -1}(u_0)$.
+We define the unique lift of $\gamma$ under $r$ as the curve $\tilde{\gamma}:[0,b]\rightarrow\btUs$ with $\tilde{\gamma}(0)=\utild$ and $\tilde{\gamma}'(0)=\tilde{\gamma}'(b)=v_0$.
+Denote their homotopy class representatives as $[\alpha],[\gamma],$ and $[\tilde{\gamma}]$.
+
+A geodesic must have trivial holonomy upon returning to its initial point on $\bGs$.
+If a geodesic segment does not lift to a closed curve on $\bUs$, then $\gamma(b)$ would be a translation of $u_0$ by some element of the deck group of the cover, i.e.
+$\varphi_1([\alpha])\in\ZZ^2$ is non-trivial, but $\varphi_2( [ \alpha] ) =I_2$ as $\varphi_2$ is a monodromy map whose image acts on $Hol_p (\triangledown)$ of $\bUs$.
+Therefore $\gamma'(b)=v_0$ as well.
+We can naturally extend these segments to the domain $[0,\infty)$ by continuously concatenating the segments.
+
+\begin{Def}
+Let $m\in\NN$. The concatenation of a geodesic segment lifted from $\alpha$ onto $\bUs$ is done in the following manner:\\
+\begin{equation}
+\gamma^m=
+\begin{cases}
+\gamma(t)  & ~~~~~~~~~~0\leq  t\leq b\\
+\varphi_1([\alpha])\cdot \gamma(t)  & ~~~~~~~~~~b\leq  t\leq 2b\\
+\vdots\\
+\varphi_1([\alpha]^{m-1})\cdot \gamma(t)  & (m-1)b\leq t\leq mb\\
+\end{cases}
+\label{eq:concat}
+\end{equation}
+ \\
+The new segment is a map $\gamma^m:[0,mb]\rightarrow\bUs$ that is well-defined on the interval.
+Let $\Phi=\lim\limits_{m\to\infty} \gamma^m$.
+We call $\Phi_t:[0,\infty)\rightarrow\bUs$ a \textbf{geodesic} on the Necker cube surface.
+\end{Def}
+
+\begin{Def}
+The geodesic $\tilde{\Phi}_t:[0,\infty)\rightarrow\btUs$ is a geodesic obtained in a similar manner to $\eqref{eq:concat}$ by lifting $\gamma$ under $r^\circ$ and continuously concatenating $\tilde{\gamma}$ with the exception that the element of the deck group acting after every iteration is $(\varphi_2([\alpha]),\varphi_1([\alpha]))\in\SO(2,\ZZ)\times \ZZ^2$.
+\end{Def}
+
+The geodesic $\Phi$ obtained in this manner has a period that depends entirely on the length of the closed segment on $\bGs$.
+In addition, $\Phi(t+mb)=m\varphi_1([\alpha])\cdot\Phi(t)$ and $tilde{\Phi}(t+mb)=(m\varphi_1([\alpha]),I_2)\cdot\tilde{\Phi}(t)$  for all $t,m\geq 0$.
+The second statement holds since $\alpha$ and $\gamma$ must have trivial holonomy at their terminal points.
+We make the following observations about these lifts:
+
+\begin{thm}
+$\Phi$ is periodic if and only if $\tilde{\Phi}$ is periodic. Further, both geodesics have period $||\alpha||$.
+\begin{proof}
+Suppose $\Phi$ is periodic.
+Then $\gamma(0)=\gamma(b)=u_0$.
+Since $\gamma$ has trivial holonomy, $\varphi_2\big\vert_{\ker\varphi_1}([\gamma])=I_2$, and so $[\tilde{\gamma}]\in\pi_1(\btUs)$.
+Hence $\gamtild(0)=\gamtild(b)=\utild$.
+Since $\tilde{\Phi}$ is obtained by concatenating $\gamtild$ under the action of the trivial deck group element of $[\alpha]$, $\gamtild^m=\gamtild(t)$ for all $t\in[0,mb]$.
+It follows that $\tilde{\Phi}$ is periodic as well.
+Conversely, the proof is trivial since $\btUs$ is a finite degree cover of $\bUs.$ Since $||\alpha||=b$, there can be no smaller period for its lifted paths as that would imply either $\alpha$ has non-trivial holonomy or does not lift to a closed segment on $\bUs$.
+\end{proof}
+\end{thm}
+
+\begin{thm}
+$\Phi$ is drift-periodic if and only if $\tilde{\Phi}$ is drift-periodic. Both geodesics have period $||\alpha||$.
+\begin{proof}
+Suppose $\Phi$ is drift-periodic.
+Then $\varphi_1([\alpha])=(x,y)$ is non-trivial, but $\varphi_2([\alpha])=I_2$ is.
+The element of the deck group acting on $\gamtild^m(t)$ for $m>1$ is $(I_2,n(x,y))$ for $1<n<m$.
+As $t\to mb$ and $m\to\infty$, $d(\tilde\Phi(t),\utild)=d((I_2,m(x,y))\cdot \utild, \utild)\to\infty$, and $\tilde{\Phi}(t+b)=(I_2,(x,y))\cdot \tilde{\Phi}(t)$ is periodically translated by some element in $\ZZ^2$.
+Conversely, if $\tilde{\Phi}$ is drift-periodic there are three cases:\\
+Case one: $\varphi_2([\alpha])$ is trivial, but $\varphi_1([\alpha])$ is not. Then $\Phi$ must be drift-periodic as well since $\gamma$ would terminate on $\varphi_1([\alpha])\cdot u_0\neq u_0$.\\
+Case two: $\varphi_1([\alpha])$ is trivial, but $\varphi_2([\alpha])$ is not. This is a contradiction since that would imply $\alpha$ has non-trivial holonomy.
+Case three:$\varphi_1([\alpha]),\varphi([\alpha])$ are both non-trivial. Again this is not possible since $\alpha,\gamma$ must both have non-trivial holonomy.\\
+Now both geodesics must have the same period since they are obtained by concatenating unique lifts of $\alpha$.
+\end{proof}
+\end{thm}
+
+These theorems essentially tell us that these kinds of geodesics behave identically on the Necker cube surface and its cover.
+Now we show that this is true for $\btUs$ and $\btU$.
+For these proofs we will refer to the induced homomorphisms from ($\ref{eq:commute}$):
+
+\begin{equation}
+\begin{tikzcd}
+\arrow[hookrightarrow, bend left]{rrr}{f_*^\circ}[swap]{} \arrow[two heads]{d}{PR_b} \pi_1(\btUs) \arrow[hook]{r}{r_*^\circ} & \pi_1(\bUs) \arrow[two heads]{d}{PR_c} \arrow[hook]{r}{q_*^\circ} & \pi_1(\bGs) \arrow[two heads]{d}{PR_d} & \ar[hookrightarrow]{l}{}[swap]{w_*^\circ} \pi_1(\bMs) \arrow[two heads]{d}{PR_a} \\
+\arrow[hookrightarrow, bend right]{rrr}{}[swap]{f_*}  \pi_1(\btU) \arrow[hook]{r}{r_*} & \pi_1(\bU) \arrow[hook]{r}{q_*} & \pi_1(\bG) & \ar[hookrightarrow]{l}{}[swap]{w_*} \pi_1(\bM)\\
+\end{tikzcd}
+\label{eq:homomorphism}
+\end{equation}
+
+\begin{Def}
+Let $\tilde{\phi}:[0,\infty)\rightarrow\btUs$ be a geodesic with initial direction $v_\in$ 
+\end{Def}
+
+\begin{thm}
+content...
+\end{thm}
+
+\subsection{Translation Surface}
+Consider the projection $\btU\rightarrow\mathbf{M}$ onto a quotient of the surface under its translational symmetries.
+What we obtain is the following compact translation surface belonging to the stratum $\cH(2,2,2,2):$
+
+\begin{figure}[H]
+\centering
+\includesvg[width=2.6in]{mtilda}
+\caption{Compact translation surface, $\mathbf{M}$ with edges and cone singularities (1,2,3,4) identified.
+The Roman numerals are meant to identify every ``copy" of $\bG$ with a direction differing from $v_0$ by rotation of some element in $\SO(2,\ZZ)$.
+The individual copies are then rotated so we have a canonical notion of direction }
+\label{fig:mtilda}
+\end{figure}
+
+We rearrange $\bM$ as such:
+
+\begin{figure}[H]
+\centering
+\includesvg[width=3.4in]{mtildastaircase}
+\caption{The staircase Veech surface with directional planes and vertices identified.
+All edges are paired by translation.
+Two adjacent squares have opposite edges identified.
+The top edge of the bottom-left square is glued to the bottom edge of the top-right square (both labeled I).
+Likewise, the bottom edge of the bottom-left square is identified with the top edge of the top-right square.}
+\label{fig:staircase}
+\end{figure}
+
+\noindent We recall some basic definitions and theorems about $\ZZ^2-$covers of translation surfaces as they apply to $\bM$.
+
+\begin{Def}
+Algebraic intersection number is a non-degenerate bilinear form:
+\begin{align*}
+i:H_1(\bM,\QQ)\times H_1(\bM,\QQ)\rightarrow\QQ,
+\end{align*}
+for $[\gamma],[\beta]\in H_1(\mathbf{M},\QQ)$, $i([\beta],[\gamma])$ returns the signed intersection number of two homology classes.
+We say a crossing at the instance of an intersection is positive if $\gamma$ makes a positive angle relative to $\beta$.
+\end{Def}
+The set of all affine diffeomorphisms of $\bM$ form the group $\text{Aff}^+(\bM)$.
+The $\emph{Veech group}$ of $\bM$ is the image in the co-domain of the homomorphism $D:\text{Aff}^+(\bM)\rightarrow SL(2,\RR)$ that takes every affine map to its derivative, which we denote $V(\bM)$.
+$\bM$ is a square-tiled translation surface whose Veech group is a finite index subgroup of $SL(2,\ZZ)$.
+We use automorphisms of the fundamental group in $\text{Aut}(\pi_1(\bM))$ induced by affine maps to prove our main results.
+But first, we define a group homomorphism from $\pi_1(\bM)$ to $\pi_1(\bM)/\pi_1(\btU)$ by algebraic intersection numbers pver a sum of linearly independent homology classes.
+Consider the following cylinder core curves on $\bM$ labeled $\gamma_i$ for $0\leq i\leq 11$:
+\begin{figure}[H]
+\includegraphics[width=3in]{homologyclass.png}
+\centering
+\caption{$\bM$'s cylinder core curves with u,v, and z homology class labels.}
+\label{fig:homology}
+\end{figure}
+
+Every curve intersects two others and if we take $i(\Sigma_{i=0}^{11}[\gamma_i])$
+\compav{show that $\gamma_0,\dots,\gamma_{11}$ form a basis for $H(\bM,\QQ)$}\\
+\compav{show that u,v linearly independent}
+
+\begin{Def}
+The homology classes $u,v,z$ are given as the following sums of core curves:
+\begin{align*}
+u &= -[\gamma_2] +[\gamma_5] + [\gamma_8] - [\gamma_{11}],\\
+v &= +[\gamma_0] -[\gamma_3] -[\gamma_6] +[\gamma_9],\\
+z &= +[\gamma_1] +[\gamma_4]-[\gamma_7]-[\gamma_{10}].
+\end{align*}
+\end{Def}
+
+\begin{Def}
+Define the group homomorphism $\Omega_{u,v}:\pi_1(\bM)\rightarrow \QQ^2$, where $\beta\mapsto(i(u,[\beta]),i(v,[\beta]))$.
+\end{Def}
+
+\begin{lem} $\Omega_{u,v}(\pi_1(\bM))=\ZZ^2$.
+\end{lem}
+\begin{lem}
+\compav{show somehow that u,v determine the cover}
+\end{lem}
+
+\subsection{Induced Automorphisms of $H_1(\bM,\QQ)$}
+We look at some important automorphisms induced by Affine maps on $\bM$.
+Observe that $\bM$ has a uniform cylinder decomposition in both horizontal and vertical directions as in figure $15$.
+We define the \emph{modulus}, $\mu$, of a cylinder to be the ratio of the cylinder's width to its circumference, $\frac{w}{c}$.
+The $\emph{Dehn-twist}$ of a cylinder is an affine diffeomorphism that skews the cylinder and sends every vertex to itself.
+
+\begin{figure}[H]
+\centering
+\includesvg[width=3in]{cylinderskew}
+\label{fig:skew}
+\caption{Dehn-twist of a cylinder in $\bM$'s cylinder decomposition.}
+\end{figure}
+
+Such a map would have derivative $\left[~\begin{matrix}1 && \pm\mu^{-1}\\0 && 1\end{matrix}~\right]$.
+On $\bM$ every cylinder in both vertical and horizontal decompositions has a modulus of $\frac{1}{2}$.
+These give way to global diffeomorphisms as $\emph{multi-twists}$ of $\bM$.
+
+\begin{Def}
+We call the global affine diffeomorphisms in $\text{Aff}^+(\bM)$ obtained as \textbf{multi-twists of the surface in horizontal and vertical directions} $\mathbf{A}$ and $\mathbf{B}$, respectively.
+\\We define the \textbf{derivatives of} $\mathbf{A},\mathbf{B}$ as the matrices 
+$D(\mathbf{A})=\mathbf{A}'=\left[ \hspace{1mm} \begin{matrix}
+				1 &   2\\
+				0 & 1
+			\end{matrix}\hspace{1mm}\right] \text{ , and }
+			D(\mathbf{B})=\mathbf{B}'=\left[ \hspace{1mm} \begin{matrix}
+							1 & 0\\
+							 2 & 1
+						\end{matrix}\hspace{1mm}\right]$, where $D:\text{Aff}^+(\bM)\rightarrow V(\bM)$.
+\end{Def}
+These are parabolic elements of $V(\bM)$ that are directly related to the dynamical dichotomy of rational trajectories on the Necker cube surface.
+To show this algebraically, we look at how skewing the surface affects our homology spanning core curves:
+
+\begin{figure}[H]
+\centering
+\includesvg[width=2in]{skew}
+\end{figure}
+
+When skewing the surface under $\mathbf{A}$ in the horizontal direction, the horizontal curves are preserved, but the vertical curves (odd $\gamma_i$ index) obtain two additional positive intersections with adjacent index (mod 12) horizontal curves.
+Similarly, the vertical $\mathbf{B}$ skews preserve the vertical curves, but the horizontal curves (even $\gamma_i$ index) obtain two additional \emph{negative} intersections with adjacent index (mod 12) vertical curves.
+The formulaic expressions of the $\mathbf{A}, \mathbf{B}$ induced automorphisms of $H_1(\bM,\QQ)$ as powers of $k\in\ZZ$ are then:
+
+\begin{align*}
+\mathbf{A}^k_*([\gamma_i])=&[\gamma_i] + \frac{k}{2}(1-(-1)^i)([\gamma_{i-1}]+[\gamma_{i+1}])\\
+\mathbf{B}^k_*([\gamma_i])=&[\gamma_i] + \frac{k}{2}(1+(-1)^i)([\gamma_{i-1}]+[\gamma_{i+1}])\\
+\end{align*}
+
+With all of these in mind we generate the following subgroups:
+
+\begin{Def}
+We say $\XX$ is the subgroup of $\text{Aff}^+(\bM)$ generated by $\mathbf{A}$, $\mathbf{B}$.
+\end{Def}
+
+\begin{Def}
+The images of $\XX$ under their derivative maps to $V(\bM)$ is the subgroup of $V(\bM)$ generated by matrices $\mathbf{A}',$ $\mathbf{B}'$. We denote this group by $\XX'$.
+\end{Def}
+
+\begin{Def}
+The induced automorphisms of $\XX$ contained in $Aut(H_1(\bM,\QQ))$ is the subgroup generated by $\mathbf{A}_*$, $\mathbf{B}_*$.
+\end{Def}
+\section{Proof of Main Theorem}
+
+$S=\cO\cup \cE$.
+
+\begin{lem}
+Let $\XX'$ act on $\cS$ by left matrix multiplication.
+Then $\XX'$ acts faithfully and transitively on integer pairs in $\cO$, partitions $\cE$ into two sets and acts faithfully and transitively on them, and acts faithfully on $\cS$.
+\begin{proof}
+Let $v=kv_0=\begin{pmatrix}x \\ y\end{pmatrix}\in\cS$ with $||v_0||=1$ and $k\in\RR$, and let $A=\begin{pmatrix}a & b \\ c & d
+\end{pmatrix}\in\XX'$.
+The kernel of the homomorphism $SL(2, \ZZ)\rightarrow \SL(2,\ZZ/2\ZZ)$ that maps a matrix to its $(mod~2)$ equivalent is precisely $\XX'$ since it maps the generators trivially.
+[quote this paper for the proof of this maybe?
+http://www.fysik.su.se/~ingemar/SL2.pdf ]\\
+$x,y$ are relatively prime, so Bezout's identity states there exist $m,n\in\ZZ$ such that $mx+ny=1~(*)$.
+It's obvious that this action is faithful on the sets since the stabilizer of any vector in $\RR^2$ is trivial.
+Suppose $v\in\cO$.
+Consider the orbit of $\begin{pmatrix}1 \\1 \end{pmatrix}$.
+Then there is some matrix $A\in\XX'$ such that $A\begin{pmatrix}1\\1\end{pmatrix}=v$.
+Since $A\in\SL(2,\ZZ)$, $\det A = ad-bc=1$.
+Obtain from $1=(x-b)(y-c)-bc=xy-xc-by$ that $xy-1=xc+by$.
+From $(*)$ we see that $xy-1=x(u(xy-1))+y(v(xy-1))$.
+Let $c=u(xy-1)$ and $b=v(xy-1)$.
+Since $xy-1$ is even, $c,b$ are even.
+It follows then that $a=x-b$ and $d=y-c$ are odd.
+Therefore $A\in \XX'$ and the action is transitive on $\cO$.
+If $v\in\cE$, then either $x$ or $y$ is even.
+In either case this would mean that either $a$ or $d$ is even, which would mean $A\notin\XX'$, a contradiction.\\
+Now consider the orbit of $\begin{pmatrix}1\\0\end{pmatrix}$.
+Left multiplying by $A$ would mean $a=x$ and $c=y$.
+Split $\cE$ into sets $\cE_1=\{(x,y)\in\cE :x\in2\ZZ,~y\in(2\ZZ+1) \}$ and $\cE_2=\{(x,y)\in\cE:x\in(2\ZZ+1),~y\in2\ZZ \}$.
+Since $\det A=1=da-bc=dx-by=mx+ny$, we have that $d=m$ and $b=-n$.
+We rule out when $v\in\cO,\cE_1$ since $c\neq 0 (\mod 2)$ in those cases.
+Consider $v\in\cE_2$.
+Then $x$ is odd and $y$ is even.
+$d$ must be odd since $mx+ny=1$ and $y$ is even.
+And $b$ must be even since $ad-bc=1$\\
+\compav{finish this and say the proof for $\cE_1$ is analogous.}
+
+\end{proof}
+\end{lem}
+
+\begin{lem}
+Let $v_0\RR^2$. Then for any geodesic $\alpha$
+\end{lem}
+
+\subsection{empty}
 \noindent Consider the following families of unit squares in $\RR^3$:\\
@@
 \end{tabular}
-
-With the proper identifications on  $\mathbf{A}_{0,0,0}\cup\mathbf{B}_{0,0,0}\cup\mathbf{C}_{0,0,0}$, we have $\bG$'s embedded form in $\RR^3$ and a map $s:\bG\hookrightarrow\RR^3$.
-Similarly, removing the integer points from this set is an embedding of $\bGs$.
-Check that tiling along the edges as previously described determines a set of points in $\RR^3$ that coincides with points on $\mathbf{S}$ with $s^*$ as an extension of $s$ to the cover.
-That is, $s(\mathbf{S})=\bigcup\left\{\mathbf{A}_{m,n,p}\cup\mathbf{B}_{m,n,p}\cup\mathbf{C}_{m,n,p}~:~m+n+p=0\right\}$.
-Thus a point $x_0\in\mathbf{S}$ is in one-to-one correspondence with this set.
-To avoid ambiguity we say $U$ is a \emph{neighborhood} of $s(x_0)$ if it is contained in exactly one of $\mathbf{A}_{m,n,p}, \mathbf{B}_{m,n,p}, \mathbf{C}_{m,n,p}$.
-As $U$ must contain an open ball contained on a single square there cannot be a neighborhood where $s(x_0)\in\ZZ^3$ and $x_0$ is a conical point.
-It follows then that a neighborhood in $\mathbf{S}$ is a neighborhood in $\mathbf{S}^\circ$ as well.
 
@@
 \subsection{Translation Surface Cover of $\bGs$}
-Reconstruct $\bG$ and $\bGs$ to resemble a quotient of the flattened Necker cube surface by a series of cutting and gluing operations: \\
-\begin{figure}[H]
-\centering
-\includesvg[width=3in]{LTorus}
-\end{figure}
-\noindent This L-shaped torus is then cut along the dotted lines and pieced together to look like a $2\times2$ torus with a unit square removed from the center:\\
-\begin{figure}[H]
-\centering
-\includesvg[width=2in]{BranchTorus}
-\caption{$\bGs$ with 4 independent, closed paths labeled identified}
-\end{figure}
-The paths labeled $a,b,c,d$ are independent.
-Let $\pi_1(\bGs)=\<a,b,c,d\>$.
-$\bGs$ has a $\ZZ^2$ tiling by traversing along paths $a,b$ and attaching infinitely many copies of the surface.
-It's apparent that this $\ZZ^2$ tiling is the flattened Necker cube surface.
-We use group homomorphisms on this free group to construct translation surfaces and $\ZZ^2$-covers to better understand geodesic behavior on the Necker cube surface.
-
-\begin{Def}
-Let $\varphi_1:\pi_1(\bGs)\rightarrow \ZZ^2$ be the group homomorphism defined on generators $a,b,c,d$: \\$c,d\mapsto(0,0)$\\ $a\mapsto(1,0)$\\$b\mapsto(0,1)$\\ $ab,ba\mapsto(1,1)$
-\end{Def}
-
-\begin{Def}
-Let $\varphi_2:\pi_1(\bGs)\rightarrow SO(2,\ZZ)$ be the group homomorphism defined on the generators $a,b,c,d$:
-\begin{align*}
-a,b & \mapsto \begin{pmatrix}
-1 && 0 \\ 0 && 1\end{pmatrix}=I_2\\
-c & \mapsto\begin{pmatrix}0 && -1\\1 && 0\end{pmatrix}=R\\ d & \mapsto\begin{pmatrix}0 && 1\\-1 && 0\end{pmatrix}=R^3
-\end{align*}
-\end{Def}
-
-Recall that a \emph{translation surface} is a compact, locally flat surface that admits a discrete number of conical singularities and has trivial linear holonomy on any arbitrary path.
-If the total cone angle of the singularity is some integer multiple of $2\pi$, $2\pi(d+1)$ for $d\geq 0$, we say that singularity has degree $d$.
-Two translation surfaces belong to the same stratum, or family, of surfaces characterized by the number and degree of conical singularities.
-Consider the effect that paths $c,d$ have on holonomy:
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=2.5in]{monodromy.png}
-\caption{Effect a nontrivial loop has on arbitrary basis vectors (i,j).}
-\label{fig:loop}
-\end{figure}
-
-We construct a cover of $\bGs$ by taking four copies of it and cyclically pairing edges, and show that this cover has trivial holonomy.
-Denote the unit tangent bundle of $\bGs$ by $T^1\bGs$.
-Because $\bGs$ can be contained in the plane, we consider vectors tangent to $\bGs$ to be in $\RR^2$.
-Let $x_0\in\bGs$.
-The fiber over $x_0$ under the projection $T^1\bGs\twoheadrightarrow\bGs$ is the set $E_{x_0}$.
-Let $\triangledown$ be a flat metric connection on $T^1\bGs$ and $p=(x_0,v_0)\in E_{x_0}$ be a single element in the fiber over $x_0$ with $v_0\in\RR^2$.
-We denote the \emph{holonomy group} of the connection with base point $p$ by $Hol_p(\triangledown)=\{A\in SO(E_{x_0})~:~ p = A\cdot p\}$ where $p$ differs from $A\cdot p$ by some rotation of $v_0$.
-An element in $SO(E_{x_0})$ acts on $p$ by rotation of the vector $v_0$.
-Denote the \emph{holonomy bundle} based at $p$ by $H(p)=\{p'\in E_{x_0}~:~ p'= A\cdot p \}$, the set of elements in the orbit of $p$ where a closed loop $\gamma:[0,1]\rightarrow\bGs$ defines a horizontal lift to $\gamma^*:[0,1]\rightarrow T^1\bGs$ by parallel transport of $v_0$ along the path such that $p=\gamma^*(0)$, and $p'=\gamma^*(1)$.
-Since $\bGs$ is flat and trivial paths have trivial holonomy, $Hol_p(\triangledown)$ acts on $H(p)$ the monodromy group is discrete and well-defined on $\bGs$.
-
-\begin{lem}
-$\varphi_2:\pi_1(\bGs)\rightarrow \mathrm{SO}(2,\ZZ)$ is a monodromy representation of the surface.
-\begin{proof}
-It is clear from Figure 6 that this is the case as concatenating a closed loop with paths $c$ or $d$ would rotate a vector exactly by the cone angles of the singularities that they loops around: $\pm \frac{3\pi}{2}$.
-\end{proof}
-\end{lem}
-
-\begin{Def}
-Let $\bMs$ be the regular cover of $\bGs$ with fundamental group $\pi_1(\bMs)=\ker\varphi_2$.
-\end{Def}
-Showing that $\bMs$ is a degree four cover of $\bGs$ with trivial holonomy and deck group $\SO(2,\ZZ)$ follows immediately from its construction.
-\begin{figure}[H]
-\centering
-\includegraphics[width=3in]{monogroup.png}
-\label{fig:arbitrarylift}
-\caption{Lift of paths $c,d$ to $\bMs$ (\compav{RELABEL EDGES})}
-\end{figure}
-
-\begin{Def}
-The surface $\bUs$ is the regular cover of $\bGs$ such that $\pi_1(\bUs)=\ker\varphi_1$. 
-\end{Def}
-
-The deck group of this cover is isomorphic to $\ZZ^2$.
-By reversing the cutting and gluing done previously to $\bGs$ and $\bG$, one can easily see that a cover defined as the kernel of this homomorphism is identical to the tiling of Necker cubes along two of its edges.
-For this reason, $\bU\cong\mathbf{S}$.
 
 \subsection{A Four-Fold Cover of $\bUs$}
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=3in]{coverdirection.png}
-\end{figure}
 
@@
 It follows that since $\omega$ is a flat connection, trivial loops have trivial holonomy and $Hol_p(\omega)$ acts on $H(p)$.
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=4in.]{closed2.png}
-\caption{Geodesic flow on $\mathbf{U}$ modeled using sage-flatsurf.}
-\label{fig:complicated}
-\end{figure}
 
@@
 
-\begin{figure}[H]
-\centering
-\includegraphics[width=2in]{overlay.png}
-\end{figure}
-
 \begin{Def}
@@
 $\btU$ is then recovered as $r\cdot(\btU_0)=\mathbf{P'}/\sim_{r(\mathbf{P}')}$, where all identifications are translations made on copies of $\mathbf{P}$.
-\subsection{Veech Staircase Model}
-Let $f:\btU\rightarrow\mathbf{M}$ be a projection onto the quotient surface that takes every point to its modular equivalent by its $\ZZ^2$ symmetries, and what we obtain is the following compact translation surface belonging to the stratum $\cH(2,2,2,2):$
-
-\begin{figure}[H]
-\centering
-\includesvg[width=2.6in]{mtilda}
-\caption{Compact translation surface, $\mathbf{M}$, covered by the infinite surface with edges and cone singularities (1,2,3,4) identified.
-The Roman numerals are meant to identify each quotient with a plane in the cover, $\mathbf{\tilde{\mathbf{U}}}$.}
-\label{fig:mtilda}
-\end{figure}
-
-We rearrange $\bM$ as such:
-
-\begin{figure}[H]
-\centering
-\includesvg[width=3.4in]{mtildastaircase}
-\caption{The staircase Veech surface with directional planes and vertices identified.
-All edges are paired by translation.
-Two adjacent squares have opposite edges identified.
-The top edge of the bottom-left square is glued to the bottom edge of the top-right square (both labeled I).
-Likewise, the bottom edge of the bottom-left square is identified with the top edge of the top-right square.}
-\label{fig:staircase}
-\end{figure}
-
-\noindent We recall some basic definitions and theorems about $\ZZ^2-$covers of translation surfaces as they apply to $\bM$.
-
-\begin{Def}
-Algebraic intersection number is a non-degenerate bilinear form:
-\begin{align*}
-i:H_1(\bM,\QQ)\times H_1(\bM,\QQ)\rightarrow\QQ,
-\end{align*}
-for $[\gamma],[\beta]\in H_1(\mathbf{M},\QQ)$, $i([\beta],[\gamma])$ returns the signed intersection number of two homology classes.
-We say a crossing at the instance of an intersection is positive if $\gamma$ makes a positive angle relative to $\beta$.
-\end{Def}
-The set of all affine diffeomorphisms of $\bM$ form the group $\text{Aff}^+(\bM)$.
-The $\emph{Veech group}$ of $\bM$ is the image in the co-domain of the homomorphism $D:\text{Aff}^+(\bM)\rightarrow SL(2,\RR)$ that takes every affine map to its derivative, which we denote $V(\bM)$.
-$\bM$ is a square-tiled translation surface whose Veech group is a finite index subgroup of $SL(2,\ZZ)$.
-We use automorphisms of the fundamental group in $\text{Aut}(\pi_1(\bM))$ induced by affine maps to prove our main results.
-But first, we define a group homomorphism from $\pi_1(\bM)$ to $\pi_1(\bM)/\pi_1(\btU)$ by algebraic intersection numbers pver a sum of linearly independent homology classes.
-Consider the following cylinder core curves on $\bM$ labeled $\gamma_i$ for $0\leq i\leq 11$:
-\begin{figure}[H]
-\includegraphics[width=3in]{homologyclass.png}
-\centering
-\caption{$\bM$'s cylinder core curves with u,v, and z homology class labels.}
-\label{fig:homology}
-\end{figure}
-
-Every curve intersects two others and if we take $i(\Sigma_{i=0}^{11}[\gamma_i])$
-\compav{show that $\gamma_0,\dots,\gamma_{11}$ form a basis for $H(\bM,\QQ)$}\\
-\compav{show that u,v linearly independent}
-
-\begin{Def}
-The homology classes $u,v,z$ are given as the following sums of core curves:
-\begin{align*}
-u &= -[\gamma_2] +[\gamma_5] + [\gamma_8] - [\gamma_{11}],\\
-v &= +[\gamma_0] -[\gamma_3] -[\gamma_6] +[\gamma_9],\\
-z &= +[\gamma_1] +[\gamma_4]-[\gamma_7]-[\gamma_{10}].
-\end{align*}
-\end{Def}
-
-\begin{Def}
-Define the group homomorphism $\Omega_{u,v}:\pi_1(\bM)\rightarrow \QQ^2$, where $\beta\mapsto(i(u,[\beta]),i(v,[\beta]))$.
-\end{Def}
-
-\begin{lem} $\Omega_{u,v}(\pi_1(\bM))=\ZZ^2$.
-\end{lem}
-\begin{lem}
-\compav{show somehow that u,v determine the cover}
-\end{lem}
-
-\subsection{Induced Automorphisms of $H_1(\bM,\QQ)$}
-We look at some important automorphisms induced by Affine maps on $\bM$.
-Observe that $\bM$ has a uniform cylinder decomposition in both horizontal and vertical directions as in figure $15$.
-We define the \emph{modulus}, $\mu$, of a cylinder to be the ratio of the cylinder's width to its circumference, $\frac{w}{c}$.
-The $\emph{Dehn-twist}$ of a cylinder is an affine diffeomorphism that skews the cylinder and sends every vertex to itself.
-
-\begin{figure}[H]
-\centering
-\includesvg[width=3in]{cylinderskew}
-\label{fig:skew}
-\caption{Dehn-twist of a cylinder in $\bM$'s cylinder decomposition.}
-\end{figure}
-
-Such a map would have derivative $\left[~\begin{matrix}1 && \pm\mu^{-1}\\0 && 1\end{matrix}~\right]$.
-On $\bM$ every cylinder in both vertical and horizontal decompositions has a modulus of $\frac{1}{2}$.
-These give way to global diffeomorphisms as $\emph{multi-twists}$ of $\bM$.
-
-\begin{Def}
-We call the global affine diffeomorphisms in $\text{Aff}^+(\bM)$ obtained as \textbf{multi-twists of the surface in horizontal and vertical directions} $\mathbf{A}$ and $\mathbf{B}$, respectively.
-\\We define the \textbf{derivatives of} $\mathbf{A},\mathbf{B}$ as the matrices 
-$D(\mathbf{A})=\mathbf{A}'=\left[ \hspace{1mm} \begin{matrix}
-				1 &   2\\
-				0 & 1
-			\end{matrix}\hspace{1mm}\right] \text{ , and }
-			D(\mathbf{B})=\mathbf{B}'=\left[ \hspace{1mm} \begin{matrix}
-							1 & 0\\
-							 2 & 1
-						\end{matrix}\hspace{1mm}\right]$, where $D:\text{Aff}^+(\bM)\rightarrow V(\bM)$.
-\end{Def}
-These are parabolic elements of $V(\bM)$ that are directly related to the dynamical dichotomy of rational trajectories on the Necker cube surface.
-To show this algebraically, we look at how skewing the surface affects our homology spanning core curves:
-
-\begin{figure}[H]
-\centering
-\includesvg[width=2in]{skew}
-\end{figure}
-
-When skewing the surface under $\mathbf{A}$ in the horizontal direction, the horizontal curves are preserved, but the vertical curves (odd $\gamma_i$ index) obtain two additional positive intersections with adjacent index (mod 12) horizontal curves.
-Similarly, the vertical $\mathbf{B}$ skews preserve the vertical curves, but the horizontal curves (even $\gamma_i$ index) obtain two additional \emph{negative} intersections with adjacent index (mod 12) vertical curves.
-The formulaic expressions of the $\mathbf{A}, \mathbf{B}$ induced automorphisms of $H_1(\bM,\QQ)$ as powers of $k\in\ZZ$ are then:
-
-\begin{align*}
-\mathbf{A}^k_*([\gamma_i])=&[\gamma_i] + \frac{k}{2}(1-(-1)^i)([\gamma_{i-1}]+[\gamma_{i+1}])\\
-\mathbf{B}^k_*([\gamma_i])=&[\gamma_i] + \frac{k}{2}(1+(-1)^i)([\gamma_{i-1}]+[\gamma_{i+1}])\\
-\end{align*}
-
-With all of these in mind we generate the following subgroups:
-
-\begin{Def}
-We say $\XX$ is the subgroup of $\text{Aff}^+(\bM)$ generated by $\mathbf{A}$, $\mathbf{B}$.
-\end{Def}
-
-\begin{Def}
-The images of $\XX$ under their derivative maps to $V(\bM)$ is the subgroup of $V(\bM)$ generated by matrices $\mathbf{A}',$ $\mathbf{B}'$. We denote this group by $\XX'$.
-\end{Def}
-
-\begin{Def}
-The induced automorphisms of $\XX$ contained in $Aut(H_1(\bM,\QQ))$ is the subgroup generated by $\mathbf{A}_*$, $\mathbf{B}_*$.
-\end{Def}
-\section{Proof of Main Theorem}
 

### 2017-11-22 newpaper.tex
112217

@@
 \def\gamtild{\tilde{\gamma}}
+\def\phitild{\tilde{\Phi}}
+\def\tildphi{\tilde{\phi}}
 
@@
 Rightfully so, a surface composed of infinitely many cubes an induced flat metric where every neighborhood is locally isometric to the Euclidean plane.
-Via the parallel transport of a unit tangent vector over the sharp edges of the surface, it becomes obvious to think of the geodesic as a sequence of line segments contained in the faces of each cube on the surface's embedded form in $\RR^3$.
+Via the parallel transport of a unit tangent vector over the sharp edges of the surface denoted $\bUs$, it becomes necessary to think of the geodesic as a sequence of line segments contained in the faces of each cube on the surface's embedded form in $\RR^3$.
 $\ref{fig:front}$.
@@
 \end{wrapfigure}
-\noindent Let $x_0$ be a point on the surface that is contained in an open neighborhood of a smooth section of the surface and consider a tangent unit vector $v\in\RR^3$ protruding from $x_0$.
-Assuming the faces are parallel to every $2$-dimensional subspace of $\RR^3$ spanned by standard basis vectors, exactly one component of $v$ would be 0.
-Projecting this vector to a parallel plane retains all necessary information about direction.
-Call this vector $v_0\in\RR^2$.
+\noindent Let $x_0$ be a point on the surface that is contained in the interior of a face and consider a tangent unit vector $v\in\RR^3$ based at a point $u_0$.
+Each face is parallel to a $2$-dimensional subspace of $\RR^3$ spanned by exactly two basis vectors, so one component of $v$ is 0.
+Call this projected vector $v_0\in\RR^2$.
 \\\\
-We say that $v_0$ is the \emph{initial trajectory} of the geodesic and the angle it makes relative to our choice of basis is its \emph{initial trajectory angle}.
-We consider a rational direction to fall into one of two categories.
+The angle $v_0$ makes relative to a choice of basis is the \emph{initial trajectory angle} of the unit-speed geodesic on the surface which we denote by $\Phi_t:\RR\rightarrow\bUs$.
+A rational initial trajectory in $\RR^2$ falls into one of two categories:
 \begin{Def}
 Let $v_0$ be a unit vector of the form $\frac{1}{k}(x,y)\in\RR^2$ with $x,y\in\mathbb{Z}$ and $k=\sqrt{x^2+y^2}\in\mathbb{R}$.
-We say $v_0$ is an \textbf{odd-odd} vector if its components are relatively prime and odd.
+We say $v_0$ is an \textbf{odd-odd} vector if its components are relatively prime and both odd.
 We denote the \textbf{set of all odd-odd directions} by $\mathcal{O}$.
@@
 \end{Def}
-\noindent This paper will demonstrate how one reaches the following conclusion about the relationship between parity and periodicity of geodesic behavior on the Necker cube surface, denoted $\mathbf{S}$:
+\noindent This paper aims to depict the relationship between initial trajectory angles and dynamical properties of a geodesic on the surface.
+Denote $\bUs$'s translational symmetries by $\text{Trans}(\bUs)$.
+Recall that a geodesic on the surface is \emph{periodic} if there exists some $T>0$ such that $\Phi(t+T)=\Phi(t)$ for all $t\in\RR$ and a geodesic is \emph{drift-periodic} if there exists some $T>0$ and non-trivial $p\in \text{Trans}(\bUs)$ such that $\Phi(t+T)=p(\Phi(t))$ for all $t\in\RR$.
+In either case we call $T$ the \emph{period} of $\Phi$ if it is the \emph{smallest} possible value that satisfies these criteria.
+We additionally require that at no point does $\Phi$ encounter a cone singularity of the surface, for which geodesic behavior is non-deterministic.
 
 \begin{thm*}
-(Directional Dichotomy of Periodic Geodesics on $\mathbf{S}$) Let $\Phi_t$ be a unit-speed geodesic flowing on the unit tangent bundle of $\mathbf{S}$ for $t\in\RR$ with initial point and vector $\Phi_0=(x_0,v)$.
+(Dynamical Properties of a Geodesic on $\bUs$) Let $\Phi_t:\RR\rightarrow\bUs$ be a non-singular unit-speed geodesic on the Necker cube surface such that $\Phi(0)=u_0$ and $\Phi'(0)=v v_0\in\RR^2$.
 Then the following is true:
 \begin{enumerate}[label=(\roman*)]
-\item $\Phi$ is periodic with period $T\in\RR$ if and only if $v_0\in\cO$.
-\item $\Phi$ is drift-periodic with period $T\in\RR$ if and only if $v_0\in\cE$.
+\item $\Phi$ is periodic with period $T\in\RR$ if  $v_0\in\cO$.
+\item $\Phi$ is drift-periodic with period $T\in\RR$ if $v_0\in\cE$.
 \end{enumerate}
@@
 Thus $\Delta_{\varphi_1}=\pi_1(\bGs)/\pi_1(\bUs)\cong \varphi_1(\pi_1(\bGs))=\ZZ^2$ is the associated \emph{deck group} of the cover. $\bUs$ is the primary object of study.
+
+\subsection{Deck Transformations of the $\ZZ^2$ Cover}
+The set of Deck transformations of the covering map $\bUs\twoheadrightarrow\bGs$ is isomorphic to the deck group $\ZZ^2$ that we induce from $\text{Trans}(\CC)$.
+First we remove open unit squares on an even integer lattice in the plane:\\
+
+\noindent  Let $m,n\in\ZZ$.
+\begin{equation}
+\label{eq:P}
+\mathbf{P} = \mathbb{C}\text{ }\backslash\bigcup_{m,n \in {\mathbb Z}} \big{\{} u+vi:~|u-2m|<\frac{1}{2},~|v-2n|<\frac{1}{2},~u=2m\pm\frac{1}{2}~v=2n\pm\frac{1}{2} \big{\}}.
+\end{equation}
+
+\noindent Now make the following identifications on $\mathbf{P}:$
+\\ Let $\aleph_{m,n}=2m+i2n\in\CC$ be the center of a removed square. Now let $\zeta_{m,n}(t)=(2m+t)+i(2n-\frac{1}{2})$ and 
+$\xi_{m,n}(t)=(2m+t)+i(2n+\frac{1}{2})$ be families of paths on $\mathbf{P}$ that run along the edges of the squares, parameterized by $t\in (\frac{-1}{2},\frac{1}{2})$.
+$\bUs$ is recovered as a topological quotient by making the following edge identifications:
+
+\begin{equation}
+\begin{split}
+(\zeta(t))\sim(e^{i\frac{\pi}{2}}~\overline{\zeta(t)-\aleph})\\
+(\eta(t))\sim_(e^{i\frac{\pi}{2}}~\overline{\eta(t)-\aleph})
+\label{eq:rel2}
+\end{split}
+\end{equation}
+
+These spaces and identifications could be adapted to recover $\bGs$ by restricting $\mathbf{P}$ to a $2\times2$ square about the origin, only allowing $m=n=0$, and identifying the outer edges of the square much like a 2-torus.
+The translations $+2,+2i\in \text{Trans}(\CC)$ induce horizontal and vertical maps $f_h,f_v\in\text{Trans}(\bUs)$.
+The group $\<f_h,f_v\>$ is isomorphic to $\ZZ^2$.
+
+\begin{figure}[H]
+\centering
+\includesvg[width=2.3in]{tiling}
+\end{figure}
 
@@
 
+\subsection{A Four-Fold Cover}
+
 \begin{Def}
@@
 The key connection made between surfaces $\bUs$ and $\btUs$ is that geodesics on $\bUs$ and $\btUs$ are identical in the sense that their images are in a one-to-one correspondence.
-Unless stated otherwise, a \emph{geodesic} is a \emph{unit-speed} locally distance minimizing curve on a smooth topological surface $\gamma:[0,\infty)\rightarrow S$ such that $\gamma'(0)=v_0\in\RR^2$ and $|| v_0||=1$.
-A geodesic is \emph{periodic} if for some $T\in\RR^+$ $\gamma(t)=\gamma(t+T)$ for all $t$.
-A geodesic is \emph{drift-periodic} if for some $T\in\RR^+$, there is some non-trivial translation of the surface that takes $\gamma(t)\mapsto \gamma(t+T)$ for all $t$.
-In either case we say that $T$ is the \emph{period} or \emph{length} of a segment; in the case where a geodesic does have a period, we restrict $\gamma$ to the closed interval $[0,T]$ and call this a \emph{geodesic segment}.
-In this paper we do not consider cases where a geodesic hits a cone singularity.
+Unless stated otherwise, a \emph{geodesic} is a \emph{unit-speed} locally distance minimizing curve on a smooth topological surface as a function parameterized by an element in $\RR$.
+A \emph{geodesic segment} is a restriction of the geodesic to some closed interval in $\RR$.
+A geodesic has an initial trajectory defined at 0, represented by a unit vector $v_0\in\RR^2$.
+If the vector can be scaled by $k\in\RR$ such that $kv_0\in\cO\cup\cE$, the geodesic has a \emph{rational} initial trajectory.
 For example, the following is a periodic geodesic on the surface with initial direction $kv_0\in\cO$:
@@
 
-Let $x_0\in G^\circ$ and fix $u_0\in q^{\circ-1}(x_0)$ Let $\alpha:[0,b]\rightarrow\bGs$ be a closed geodesic segment on $\bGs$ with initial and terminal point $x_0$, and $\alpha'(0)=v_0$.
-Denote $\alpha$'s unique lift to $\bUs$ under $q^\circ$ as $\gamma:[0,b]$, where $\gamma(0)=u_0$, $\gamma'(0)=\gamma'(b)=v_0$, and $\gamma(b)\in q^{\circ-1}(x_0)$.
-Fix an arbitrary point $\utild\in r^{\circ -1}(u_0)$.
-We define the unique lift of $\gamma$ under $r$ as the curve $\tilde{\gamma}:[0,b]\rightarrow\btUs$ with $\tilde{\gamma}(0)=\utild$ and $\tilde{\gamma}'(0)=\tilde{\gamma}'(b)=v_0$.
-Denote their homotopy class representatives as $[\alpha],[\gamma],$ and $[\tilde{\gamma}]$.
-
-A geodesic must have trivial holonomy upon returning to its initial point on $\bGs$.
-If a geodesic segment does not lift to a closed curve on $\bUs$, then $\gamma(b)$ would be a translation of $u_0$ by some element of the deck group of the cover, i.e.
-$\varphi_1([\alpha])\in\ZZ^2$ is non-trivial, but $\varphi_2( [ \alpha] ) =I_2$ as $\varphi_2$ is a monodromy map whose image acts on $Hol_p (\triangledown)$ of $\bUs$.
-Therefore $\gamma'(b)=v_0$ as well.
-We can naturally extend these segments to the domain $[0,\infty)$ by continuously concatenating the segments.
-
-\begin{Def}
-Let $m\in\NN$. The concatenation of a geodesic segment lifted from $\alpha$ onto $\bUs$ is done in the following manner:\\
-\begin{equation}
-\gamma^m=
-\begin{cases}
-\gamma(t)  & ~~~~~~~~~~0\leq  t\leq b\\
-\varphi_1([\alpha])\cdot \gamma(t)  & ~~~~~~~~~~b\leq  t\leq 2b\\
-\vdots\\
-\varphi_1([\alpha]^{m-1})\cdot \gamma(t)  & (m-1)b\leq t\leq mb\\
-\end{cases}
-\label{eq:concat}
-\end{equation}
+Let $x_0\in G^\circ$ and fix $u_0\in\bUs$, an element in the fiber of the covering map.
+A \emph{geodesic} on $G^\circ$ is a closed unit-speed straight-line path given as $\eta:\RR\rightarrow\bGs$.
+The \emph{period}/\emph{length} of the geodesic is some strictly positive $b\in\RR$.
+Let $\eta|_{[0,b]}=\alpha:[0,b]\rightarrow\bGs$ be a closed geodesic segment on $\bGs$ with initial and terminal point $x_0$.
+Denote $\alpha$'s unique lift to $\bUs$ by $\gamma:[0,b]\rightarrow\bUs$, where $\gamma(0)=u_0$, and $\gamma(b)$ in the fiber over $x_0$.
+Fix a point $\utild\in\btUs$ in the fiber over $u_0$.
+We define the unique lift of $\gamma$ to be the curve $\tilde{\gamma}:[0,b]\rightarrow\btUs$ with $\tilde{\gamma}(0)=\utild$.
+Denote their homotopy class representatives by $[\alpha]$ and (if they exist) $[\gamma],~[\tilde{\gamma}]$.
+
+The geodesic segment on $\bGs$ has the property that it returns to $x_0$ with a trivial effect on the holonomy of the curve.
+On the other hand, the lifted paths will close only if their homotopy classes belong in their fundamental groups, i.e.
+the kernel of homomorphisms $\varphi_1$, $\varphi_2|_{\ker\varphi_1}$.
+Otherwise they are mapped non-trivially to $\ZZ^2$ or $\ZZ^2\times\SO(2,\ZZ)$.
+Let $v_0\in\RR^2$ be a unit vector such that $\alpha's$ image under the derivative map, $\alpha'$, is $v_0$ at times $0$ and $b$.
+The derivatives of the lifts $\gamma', \gamtild'$ are also $v_0$ at their endpoints.
+
+\begin{Def}
+Let $\Phi_t:\RR\rightarrow\bUs$ be a unit-speed geodesic on the \emph{Necker cube surface} with $\Phi(0)=u_0$ and $\Phi'(0)=v_0$.
+\\Similarly, let $\tilde{\Phi}_t:\RR\rightarrow\btUs$ be a unit-speed geodesic on $\bUs$'s four-fold cover with $\phitild(0)=\utild$ and $\phitild'(0)=v_0$.
+\end{Def}
+
+On the Necker cube surface, a horizontal translation, $f_h$, is induced by $2\in\text{Trans}(\CC)$, and a vertical translation, $f_v$, is induced by $2i\in\text{Trans}(\CC)$.
+The group of deck transformations of the surface given by the cover $\bUs\rightarrow\bGs$ is isomorphic to the group $\Delta_{\varphi_1}=\pi_1(\bGs)/\pi_(\bUs)\cong\ZZ^2$.
+Take $\ZZ^2$ to the deck group by sending $(1,0)$ to $f_h$ and $(0,1)$ to $f_v$.
+Similarly, $\btUs$ has deck transformations given by the cover $\btUs\rightarrow\bGs$ induced by elements in $\text{Trans}(\CC\times\SO(2,\ZZ))$.
+These are isomorphic to the deck group given by the cover $\btUs\rightarrow\bUs\rightarrow\bGs$ containing elements of the form $((m,n),M)\in\ZZ^2\rtimes_{\eta}\SO(2,\ZZ)$, where $\eta:\SO(2,\ZZ)\rightarrow \Aut(\ZZ^2)$.
+$\SO(2,\ZZ)$ already belongs in $\Aut(\ZZ^2)$ so $\eta$ is an inclusion endomorphism, and the binary operation of this group is defined as  $((m,n),M)\cdot((m',n'),M')=((m,n)+M(m',n'),MM')$.\\
 \\
-The new segment is a map $\gamma^m:[0,mb]\rightarrow\bUs$ that is well-defined on the interval.
-Let $\Phi=\lim\limits_{m\to\infty} \gamma^m$.
-We call $\Phi_t:[0,\infty)\rightarrow\bUs$ a \textbf{geodesic} on the Necker cube surface.
-\end{Def}
-
-\begin{Def}
-The geodesic $\tilde{\Phi}_t:[0,\infty)\rightarrow\btUs$ is a geodesic obtained in a similar manner to $\eqref{eq:concat}$ by lifting $\gamma$ under $r^\circ$ and continuously concatenating $\tilde{\gamma}$ with the exception that the element of the deck group acting after every iteration is $(\varphi_2([\alpha]),\varphi_1([\alpha]))\in\SO(2,\ZZ)\times \ZZ^2$.
-\end{Def}
-
-The geodesic $\Phi$ obtained in this manner has a period that depends entirely on the length of the closed segment on $\bGs$.
-In addition, $\Phi(t+mb)=m\varphi_1([\alpha])\cdot\Phi(t)$ and $tilde{\Phi}(t+mb)=(m\varphi_1([\alpha]),I_2)\cdot\tilde{\Phi}(t)$  for all $t,m\geq 0$.
-The second statement holds since $\alpha$ and $\gamma$ must have trivial holonomy at their terminal points.
-We make the following observations about these lifts:
+Let $\varphi:\pi_1(\bGs)\rightarrow \ZZ^2\rtimes_{\eta}\SO(2,\ZZ)$ be given by $[g]\mapsto(\varphi_1([g]),\varphi_2([g]))$. $\varphi$ is a
 
 \begin{thm}
-$\Phi$ is periodic if and only if $\tilde{\Phi}$ is periodic. Further, both geodesics have period $||\alpha||$.
-\begin{proof}
-Suppose $\Phi$ is periodic.
-Then $\gamma(0)=\gamma(b)=u_0$.
-Since $\gamma$ has trivial holonomy, $\varphi_2\big\vert_{\ker\varphi_1}([\gamma])=I_2$, and so $[\tilde{\gamma}]\in\pi_1(\btUs)$.
-Hence $\gamtild(0)=\gamtild(b)=\utild$.
-Since $\tilde{\Phi}$ is obtained by concatenating $\gamtild$ under the action of the trivial deck group element of $[\alpha]$, $\gamtild^m=\gamtild(t)$ for all $t\in[0,mb]$.
-It follows that $\tilde{\Phi}$ is periodic as well.
-Conversely, the proof is trivial since $\btUs$ is a finite degree cover of $\bUs.$ Since $||\alpha||=b$, there can be no smaller period for its lifted paths as that would imply either $\alpha$ has non-trivial holonomy or does not lift to a closed segment on $\bUs$.
-\end{proof}
+Let $\delta=\varphi_1([\alpha])\in\ZZ^2$ be an element of $\Delta_{\varphi_2}$ and $\tilde{\delta}=\varphi([\alpha])\in\ZZ^2\rtimes_{\eta}\SO(2,\ZZ)$ be an element of $\Delta_{\varphi_{2}\vert_{\pi_1(\bUs)}}$.
+Then:\\
+$\Phi(t+b)=\delta\cdot\Phi(t)$ for all $t\in\RR$, and \\
+$\phitild(t+b)=\tilde{\delta}\cdot\phitild(t)$ for all $t\in\RR$.
 \end{thm}
 
-\begin{thm}
-$\Phi$ is drift-periodic if and only if $\tilde{\Phi}$ is drift-periodic. Both geodesics have period $||\alpha||$.
-\begin{proof}
-Suppose $\Phi$ is drift-periodic.
-Then $\varphi_1([\alpha])=(x,y)$ is non-trivial, but $\varphi_2([\alpha])=I_2$ is.
-The element of the deck group acting on $\gamtild^m(t)$ for $m>1$ is $(I_2,n(x,y))$ for $1<n<m$.
-As $t\to mb$ and $m\to\infty$, $d(\tilde\Phi(t),\utild)=d((I_2,m(x,y))\cdot \utild, \utild)\to\infty$, and $\tilde{\Phi}(t+b)=(I_2,(x,y))\cdot \tilde{\Phi}(t)$ is periodically translated by some element in $\ZZ^2$.
-Conversely, if $\tilde{\Phi}$ is drift-periodic there are three cases:\\
-Case one: $\varphi_2([\alpha])$ is trivial, but $\varphi_1([\alpha])$ is not. Then $\Phi$ must be drift-periodic as well since $\gamma$ would terminate on $\varphi_1([\alpha])\cdot u_0\neq u_0$.\\
-Case two: $\varphi_1([\alpha])$ is trivial, but $\varphi_2([\alpha])$ is not. This is a contradiction since that would imply $\alpha$ has non-trivial holonomy.
-Case three:$\varphi_1([\alpha]),\varphi([\alpha])$ are both non-trivial. Again this is not possible since $\alpha,\gamma$ must both have non-trivial holonomy.\\
-Now both geodesics must have the same period since they are obtained by concatenating unique lifts of $\alpha$.
-\end{proof}
-\end{thm}
+\begin{cor}
+The following are equivalent:
+\begin{enumerate}
+\item $\Phi$ is periodic.
+\item $\phitild$ is periodic.
+\item $\delta=(0,0)$.
+\end{enumerate}
+\end{cor}
+
+\begin{cor}
+The following are equivalent:
+\begin{enumerate}
+\item $\Phi$ is drift-periodic.
+\item $\phitild$ is drift-periodic.
+\item $\delta\neq(0,0)$.
+\end{enumerate}
+\end{cor}
 
@@
 \begin{Def}
-Let $\tilde{\phi}:[0,\infty)\rightarrow\btUs$ be a geodesic with initial direction $v_\in$ 
+Let $\tildphi:\RR\rightarrow\btU$ be a geodesic on the unpunctured four-fold surface such that $\tildphi(0)=\utild$ and $\tildphi'(0)=v_0$.
 \end{Def}
@@
 \begin{thm}
-content...
+$\tildphi$ is periodic if and only if $\phitild$ is periodic.\\
+$\tildphi$ is drift-periodic if and only if $\phitild$ is drift-periodic.
 \end{thm}
@@
 
-\begin{equation}
-\label{eq:P}
-\mathbf{P} = \mathbb{C}\text{ }\backslash\bigcup_{m,n \in {\mathbb Z}} \big{\{} u+vi:u\in(2m-\frac{1}{2},2m+\frac{1}{2}),\text{ } v\in(2n-\frac{1}{2},2n+\frac{1}{2})\big{\}}.
-\end{equation}
-
 \begin{Def} $\mathbf U$ is the surface obtained as the topological quotient $\mathbf{P}/\sim_{\mathbf{P}}$, where $\CC$ is identified with $\RR^2$ in the usual way and $\sim_{\mathbf{P}}$ is a minimal relation on $\mathbf{P}$ defined as follows:
@@
 Denote paths $\zeta$ and $\eta$  on $\mathbf{P}$, parameterized by integers $m,n\in\ZZ$ and $t\in[-\frac{1}{4},\frac{1}{4}]$, and defined:\\ $\zeta(t)=(2m+t)+i(2n-\frac{1}{2})$,\\
-$\eta(t)=(2m+t)+i(2n+\frac{1}{2})$.\\ Let $\aleph=2m+i2n\in\CC$. The minimal relation $\sim_{\mathbf{P}'}$ is given as:
+$\eta(t)=(2m+t)+i(2n+\frac{1}{2})$.\\
+
+Let $\aleph=2m+i2n\in\CC$. The minimal relation $\sim_{\mathbf{P}'}$ is given as:
 \begin{equation}
@@
 (\eta(t);j)\sim_{\mathbf{P'}}(e^{i\frac{\pi}{2}}~\overline{\eta(t)-\aleph};j+1)
-\label{eq:rel2}
 \end{split}

### 2017-12-04 newpaper.tex
120417

@@
 
-\title{Periodicity of Geodesic Flows on the Necker Cube Surface}
+\title{Periodicity of Geodesics on the Necker Cube Surface}
 \author{Pavel Javornik}
@@
 \begin{abstract}
-We study dynamical properties of geodesic flows on a flat, periodically constructed Euclidean cone surface by obtaining an infinite-type translation cover branched over conical singularities, of which admits a flat induced metric and a connection between any two points in the tangent vector bundle of the surface has trivial holonomy.
-A unit vector flow on the unit tangent bundle of such a surface is well-defined with a canonical vector representation in $\RR^2$.
-We study this infinite-type surface as a $\ZZ^2$ cover of a compact Veech surface belonging to the stratum of $\cH(2,2,2,2)$ translation surfaces, and use its $SL(2,\RR)-$commensurable Veech group to prove results that relate directions of flows to the periodicity and ergodicity of their lifts in the cover.
+\noindent Although it is most recognizable as M.C.
+Escher's adaptation of an optical illusion popularized by Louis Albert Necker, the Necker cube surface has a complex structure and is, in analaytic language, described as an infinite, flat (locally isometric to $\RR^2$ on smooth sections), $\RR^3$-embeddable Euclidean cone surface with countably many singularities of cone angles $3\pi$ and $\frac{3\pi}{2}$.
+Its geometric construction is described as the gluing of infinitely many unit cubes along their edges so that every face shares an edge with exactly four others.
+A unit-speed geodesic on this surface has an initial trajectory angle $\theta\in\RR/2\pi\ZZ$ given up by the surface's local plane isometry.
+Given a rational vector direction in $\RR^2$ (components are relatively prime integers) obtained from $\theta$, one can classify every periodic/drift-periodic rational geodesic on the surface.
+In addition its total period/length and, in the case where it is drift-periodic, translational drift as Euclidean distance in its ambient space is given by solving a system of simple linear Diophantine equations.
 \end{abstract}
@@
 \centering
-\includesvg[width=2.2in]{cubesxyzpaths2}
+\includesvg[width=4in]{cubesxyzpaths2bwv2b}
 \label{fig:front}
@@
 
-Historically, the Necker cube[citation] has made numerous appearances in the work of mathematicians, crystallographers, and scientists interested in human visual systems prior to it being popularized in the works of illusionary artist M.C.
-Escher (pictured below).
-The crystallographer Louis Albert Necker was credited for having discovered the optical illusion, and studying its geometry[cite].
-The solid presentation of the cube, when rendered as a flat surface with its back faces removed, achieves a similar effect when its three visible faces are shaded in a particular way.
-We are interested in the infinite tiling of this structure (figure $\ref{fig:front}$), which has the appearance of a rhombile tiling of the Euclidean plane, and refer to it as the Necker cube surface.
-Video game fans might even recognize it as the same surface that appears in the 1982 arcade game by Gottlieb, ``Q*bert."
-
-\begin{figure}[H]
-\begin{center}
-\includegraphics[scale=0.4]{escher.jpg}
-\includegraphics[scale=0.55]{escher2.jpg}
-\caption{The Necker cube tiling as it appears briefly in ``Metamorphosis I," and impressed into a banner in ``Convex and Concave."[citation]}
-\label{fig:Escher}
-\end{center}
-\end{figure}
-
-Flat, periodic surfaces such as the Necker cube surface are of much interest to differential geometers for two reasons in particular.
-The first is that, locally, neighborhoods are isometric to the plane, making it remarkably easy to describe parallel transports between two points in tangent vector bundles over the surface by linear group actions.
-The second is that Riemann surfaces constructed out of infinitely many polyhedra can be \emph{flattened} and contained in the plane in such a way that the metric is preserved.
-The advantage in this case is that one obtains something akin to infinite billiard tables where a geodesic flows are represented by series of straight line segments in the plane.
-This paper will use such methods to prove dynamical results about Geodesic flows on the Necker cube surface.
+The Necker cube has made numerous appearances in the work of mathematicians, crystallographers, and scientists interested in human visual systems before and after appearing in the works of M.C.
+Escher [cite].
+Louis Albert Necker was credited for having discovered the optical illusion, and studying its geometric properties [cite].
+The solid presentation of the cube, when rendered as a flat surface with three rear faces removed, achieves a similar effect when its three visible faces are shaded in a particular way.
+The Necker cube surface is then obtained by gluing infinitely many copies along its outer edges as in Figure 1.
+
+Dynamical classification of geodesics on flat, periodic surfaces such as the Necker cube surface ultimately boils down to  studying the surface's various symmetries and how the rotational holonomy group acts on the unit tangent bundle of the surface as the unit vector (derivative of the geodesic) is parallel transported along its path.
+Loops on polyhedra punctured at its cone singularities tend to have discrete holonomy groups as its relative holonomy (holonomy of contractible loops) is trivial.
 
@@
 $\ref{fig:front}$.
-
-\begin{wrapfigure}{l}{2.2in.}
+\\
+\begin{wrapfigure}{l}{2in.}
 \includesvg{vectorplane}
 \end{wrapfigure}
-\noindent Let $x_0$ be a point on the surface that is contained in the interior of a face and consider a tangent unit vector $v\in\RR^3$ based at a point $u_0$.
+\noindent Let $u_0$ be a point on the surface that is contained in the interior of a face and consider a tangent unit vector $v\in\RR^3$ based at point $u_0$.
 Each face is parallel to a $2$-dimensional subspace of $\RR^3$ spanned by exactly two basis vectors, so one component of $v$ is 0.
 Call this projected vector $v_0\in\RR^2$.
-\\\\
-The angle $v_0$ makes relative to a choice of basis is the \emph{initial trajectory angle} of the unit-speed geodesic on the surface which we denote by $\Phi_t:\RR\rightarrow\bUs$.
-A rational initial trajectory in $\RR^2$ falls into one of two categories:
+\\\\\\
+The angle $v_0$ makes relative to a choice of basis is the \emph{initial trajectory angle} of the unit-speed geodesic on the surface, $\Phi_t:\RR\rightarrow\bUs$.
+A rational initial trajectory angle in $\RR^2$ falls into one of two categories:
 \begin{Def}
@@
 \end{Def}
-\noindent This paper aims to depict the relationship between initial trajectory angles and dynamical properties of a geodesic on the surface.
-Denote $\bUs$'s translational symmetries by $\text{Trans}(\bUs)$.
-Recall that a geodesic on the surface is \emph{periodic} if there exists some $T>0$ such that $\Phi(t+T)=\Phi(t)$ for all $t\in\RR$ and a geodesic is \emph{drift-periodic} if there exists some $T>0$ and non-trivial $p\in \text{Trans}(\bUs)$ such that $\Phi(t+T)=p(\Phi(t))$ for all $t\in\RR$.
-In either case we call $T$ the \emph{period} of $\Phi$ if it is the \emph{smallest} possible value that satisfies these criteria.
-We additionally require that at no point does $\Phi$ encounter a cone singularity of the surface, for which geodesic behavior is non-deterministic.
+\noindent For convenience we call the disjoint union of these sets $\cS$.
+This paper aims to depict the relationship between initial trajectory angles and dynamical properties of a geodesic on the surface.
+Denote $\bUs$'s translational isometries by $\text{Trans}(\bUs)$.
+Recall that a geodesic on the surface is \emph{periodic} if there exists some $T>0$ such that $\Phi(t+T)=\Phi(t)$ for all $t\in\RR$ and a geodesic is \emph{drift-periodic} if there exists some $T'>0$ and non-trivial $p\in \text{Trans}(\bUs)$ such that $\Phi(t+T')=p(\Phi(t))$ for all $t\in\RR$.
+In either case we call this value the \emph{period} of $\Phi$ if it is the \emph{smallest} possible value that satisfies our definition.
+We additionally require that at no point does $\Phi$ encounter a cone singularity of the surface, in which case the geodesic behavior is non-deterministic.
+
+\begin{align}
+\text{Let }\XX'=\left\< \left[ \hspace{1mm} \begin{matrix}
+							1 & 2\\
+							 0 & 1
+						\end{matrix}\hspace{1mm}\right],~
+\left[ \hspace{1mm} \begin{matrix}
+							1 & 0\\
+							 2 & 1
+						\end{matrix}\hspace{1mm}\right]\right\>
+\end{align}
+
+\noindent Section [blank] explains the significance of this group. Let $M=\left[ \hspace{1mm} \begin{matrix}
+							s & u\\
+							 w & z
+						\end{matrix}\hspace{1mm}\right]\in\XX'$.
 
 \begin{thm*}
-(Dynamical Properties of a Geodesic on $\bUs$) Let $\Phi_t:\RR\rightarrow\bUs$ be a non-singular unit-speed geodesic on the Necker cube surface such that $\Phi(0)=u_0$ and $\Phi'(0)=v v_0\in\RR^2$.
+(Dynamical Properties of a Geodesic on $\bUs$) \\Let $\Phi_t:\RR\rightarrow\bUs$ be a non-singular unit-speed geodesic on the Necker cube surface such that $\Phi(0)=u_0$ and $\Phi'(0)=v\in\RR^3$, isometric to $v_0\in\RR^2$.
+Let $\overline{u}_t\in\RR^3$ be $\Phi_t$'s coordinate in $\bUs$'s ambient space at time $t$.
 Then the following is true:
 \begin{enumerate}[label=(\roman*)]
-\item $\Phi$ is periodic with period $T\in\RR$ if  $v_0\in\cO$.
-\item $\Phi$ is drift-periodic with period $T\in\RR$ if $v_0\in\cE$.
+\item $\Phi$ is periodic if $v_0\in\cO$.
+\item Suppose $(i)$ is true. If $M (kv_0)=(1,1)$ then $\Phi$ has period $T=6\sqrt{2}||kv_0||$.
+\item $\Phi$ is drift-periodic if $v_0\in\cE$. 
+\item Suppose $(iii)$ is true. If $M (kv_0)=(1,0)$ or $M (kv_0)=(0,1)$ then $\Phi$ has period $T=2||kv_0||$, and $d_E(\overline{u}_t,\overline{u}_{t+T})=||kv_0||\sqrt{2}$.
 \end{enumerate}
 \end{thm*}
-
+\compav{Remember to prove that such an M exists.}
+
+Here $||~||$ is the standard Euclidean norm in $\RR^2$, and $d_E:\RR^3\times\RR^3\rightarrow\RR$ is the Euclidean distance.
 The proof for this theorem can be found in XXX and follows from Theorems YYYY and ZZZZZ.
-\\
-\compav{theorem about most directions being recurrent?}
 
@@
 \section{Periodic Tiling of Necker Cubes}
-This section will detail how the Necker Cube surface is obtained as an infinite regular cover of a branched torus. Consider the Necker cube with the following identifications:
-\begin{figure}[H]
-\centering
-\includesvg[width=1.6in.]{neckercube}\includesvg[width=2in.]{neckercubepath}
-\end{figure}
-
-We denote this surface $\bG$.
-$\bG$ is obtained by taking three copies of the unit square and identifying the edges in order to resemble the visible portion of a unit cube.
-$\bG$ is a genus one surface with three vertices homeomorphic to $\TT^2$.
-Check that every vertex has a cone angle of either $3\pi$ or $\frac{3\pi}{2}$.
-These are the conical singularities of the surface, $\Sigma$.
-We denote the surface without its singularities $\bG\backslash\Sigma = \bGs$.
-Thus, $\pi_1(\bGs)\cong\pi_1((\TT^2 \text{ with 3 punctures}))\cong\pi_1(S^1\vee S^1\vee S^1\vee S^1)\cong\FF_4$, the free group of four generators.
-We may flatten $\bG$ and $\bGs$ into an L-shaped torus: \\
+This section will detail how the Necker Cube surface is constructed as a covering space of a thrice-punctured torus. Make the following identifications on 3 unit squares:
+\begin{figure}[H]
+\centering
+\includesvg[width=1.83in]{neckercube}\includesvg[width=2in.]{neckercubepath}
+\caption{$\bGs$ with and without paths $a,b,c,d$.}
+\end{figure}
+
+We will call this ``half-cube" $\bG$.
+It is a genus one piecewise smooth surface homeomorphic to $\TT^2$.
+Every vertex has a cone angle of either $3\pi$ or $\frac{3\pi}{2}$.
+Let $\Sigma_{\bG}\subset\bG$ be the set of these three vertices.
+These are conical singularities of the surface.
+We denote the surface punctured at these points by $\bG\backslash\Sigma_{\bG} = \bGs$.
+The paths labeled $a,b,c,d$ are independent of one another and span $\bGs$'s first fundamental group, $\pi_1(\bGs)\cong \<a,b,c,d \>$, the free group of four generators.
+The unpunctured Necker cube surface $\bU$ is embedded in $\RR^3$ by tiling along $\bG$'s outer edges:
+
+\begin{figure}[H]
+\centering
+\includesvg[width=4in]{neckercubepathsurface}
+\end{figure}
+From this diagram we see that $\bU$ is a universal cover of $\TT^2$.
+In this paper we are studying geodesic behavior on $\bU$ and in order to do so, it is necessary to determine how the parallel transport of a unit vector around a cone singularity acts on the unit tangent bundle of the surface.
+But to study the path monodromy we require that these loops (see $c,d$ in Fig 3) aren't contractible on the cube's sharp corners.
+Denote the countably infinite set of singularities of $\bU$ by $\Sigma_{\bU}$.
+The \emph{Necker cube surface} is the punctured surface $\bUs=\bU\backslash\Sigma_{\bU}$.
+Observe from the previous diagram that $\pi_1(\bUs)$ is the kernel of the following group homomorphism:
+
+\begin{Def}
+Denote $\varphi_1:\pi_1(\bGs)\rightarrow \ZZ^2$ as the group homomorphism such that $c,d\mapsto(0,0)$, $a\mapsto(1,0)$, and $b\mapsto(0,1)$.
+\end{Def}
+
+Take $\pi_1(\bUs)$ to be $\ker\varphi_1$.
+Then $\Delta_{\varphi_1}=\pi_1(\bGs)/\pi_1(\bUs)\cong\ZZ^2$ is the \emph{deck group} of the covering map $\varphi_1^*:\bUs\rightarrow\bGs$.
+Now $\bGs$ is unfolded and flattened in order to study monodromy as it transports a vector in $\RR^2$ along a path.
+Apply the following cutting and gluing operations to the surface: \\
 \begin{figure}[H]
@@
 \end{figure}
-\noindent This L-shaped torus is then cut along the dotted lines and pieced together to resemble a $2\times2$ torus with a unit square removed from the center:\\
+\noindent This L-shaped torus is then taken apart and recovered as $2\times 2$ torus missing a unit square with the following identifications:\\
 \begin{figure}[H]
@@
 \includesvg[width=2in]{BranchTorus}
-\caption{$\bGs$ with 4 independent, closed paths labeled identified}
-\end{figure}
-Denote its fundamental group by $\pi_1(\bGs)=\<a,b,c,d\>$, where $a,b,c,d$ are the labeled independent paths on the surface. Consider the following map:
-
-\begin{Def}
-Let $\varphi_1:\pi_1(\bGs)\rightarrow \ZZ^2$ be the group homomorphism defined on generators $a,b,c,d$: \\$c,d\mapsto(0,0)$\\ $a\mapsto(1,0)$\\$b\mapsto(0,1)$
-\end{Def}
-
-\begin{Def}
-The punctured \textbf{Necker Cube Surface} is a regular cover of $\bGs$, and is denoted $\bUs$. It's fundamental group presentation is the subgroup of $\pi_1(\bGs)$, $\pi_1(\bUs)=\ker\varphi_1$.
-\end{Def}
-
-Thus $\Delta_{\varphi_1}=\pi_1(\bGs)/\pi_1(\bUs)\cong \varphi_1(\pi_1(\bGs))=\ZZ^2$ is the associated \emph{deck group} of the cover. $\bUs$ is the primary object of study.
-
-\subsection{Deck Transformations of the $\ZZ^2$ Cover}
-The set of Deck transformations of the covering map $\bUs\twoheadrightarrow\bGs$ is isomorphic to the deck group $\ZZ^2$ that we induce from $\text{Trans}(\CC)$.
-First we remove open unit squares on an even integer lattice in the plane:\\
-
-\noindent  Let $m,n\in\ZZ$.
+\caption{$\bGs$ and images of paths after reconstruction. The outermost vertex is \emph{not} a puncture.}
+\end{figure}
+
+Consider a $2\times2$ sheet missing an open unit square and that square's vertices (punctures).
+Center this structure around $0\in\CC$ and recover $\bGs$ with the identifications made in Fig 3.
+Let $m,n\in\ZZ$.
+Similar identifications can be made on the following subset of $\CC$ to recover $\bUs$:
+
 \begin{equation}
@@
 
-\noindent Now make the following identifications on $\mathbf{P}:$
-\\ Let $\aleph_{m,n}=2m+i2n\in\CC$ be the center of a removed square. Now let $\zeta_{m,n}(t)=(2m+t)+i(2n-\frac{1}{2})$ and 
-$\xi_{m,n}(t)=(2m+t)+i(2n+\frac{1}{2})$ be families of paths on $\mathbf{P}$ that run along the edges of the squares, parameterized by $t\in (\frac{-1}{2},\frac{1}{2})$.
-$\bUs$ is recovered as a topological quotient by making the following edge identifications:
+\noindent Let $\aleph_{m,n}=2m+i2n\in\CC$ be the center of a removed square. Now let $\zeta_{m,n}(t)=(2m+t)+i(2n-\frac{1}{2})$ and 
+$\xi_{m,n}(t)=(2m+t)+i(2n+\frac{1}{2})$ be paths on $\mathbf{P}$ that run along the edges of the squares, parameterized by $t\in (-\frac{1}{2},\frac{1}{2})$ and $(m,n)\in\ZZ^2$.
+$\bUs$ is recovered as a topological quotient $\mathbf{P}/\sim$, where $\sim$ is the minimal relation on $\mathbf{P}$:
 
@@
 \begin{split}
-(\zeta(t))\sim(e^{i\frac{\pi}{2}}~\overline{\zeta(t)-\aleph})\\
-(\eta(t))\sim_(e^{i\frac{\pi}{2}}~\overline{\eta(t)-\aleph})
+\zeta(t)\sim e^{i\frac{\pi}{2}}~\overline{\zeta(t)-\aleph}\\
+\eta(t)\sim e^{i\frac{\pi}{2}}~\overline{\eta(t)-\aleph}
 \label{eq:rel2}
@@
 \end{equation}
-
-These spaces and identifications could be adapted to recover $\bGs$ by restricting $\mathbf{P}$ to a $2\times2$ square about the origin, only allowing $m=n=0$, and identifying the outer edges of the square much like a 2-torus.
-The translations $+2,+2i\in \text{Trans}(\CC)$ induce horizontal and vertical maps $f_h,f_v\in\text{Trans}(\bUs)$.
-The group $\<f_h,f_v\>$ is isomorphic to $\ZZ^2$.
+The translations, $\pm 2,\pm 2i$, on $\mathbf{P}$ are isometries that preserve these identifications and induce isometries of $\bUs$.
+The group generated by these horizontal and vertical translations form the \emph{group of Deck transformations}, $\text{Deck}(\varphi_1^*)$, that  is isomorphic to $\ZZ^2$:
 
@@
 
+Identify 2 with $f_h$, and $2i$ with $f_v$.
+Let $x_0\in\bGs$ and fix a point in the fiber,  $u_0\in\varphi_1^*{-1}(x_0)\subset\bUs$.
+The action of $\ZZ^2$ on the fiber over $x_0$ is given by $(m,n)\cdot u_0=(f_h^m\circ f_v^n)(u_0)$.
+The distance between $u_0$ and $(f_h^m\circ f_v^n)(u_0)$ is measured in its ambient space and taken to be the magnitude of $2m+i2n\in\CC$.
+
+\begin{lem}
+Let $[\alpha]$ be homotopy class not in the kernel of $\varphi_1$. It will lift to an unclosed path terminating on $\varphi_1([\alpha])\cdot u_0$.
+\end{lem}
+
 \subsection{Monodromy Group Representation}
-In this section we describe a translation cover of $\bGs$ which and $\bUs$ which is factored through as the kernel of a homomorphism from $\pi_1(\bGs)\rightarrow \SO(2,\ZZ)$.
-This homomorphism represents the monodromy group as the fundamental group acting on the unit tangent bundles $T^1\bGs$ and $T^1\bUs$.
-
-\begin{Def}
-Let $\varphi_2:\pi_1(\bGs)\rightarrow SO(2,\ZZ)$ be the group homomorphism defined on the generators $a,b,c,d$:
-\begin{align*}
-a,b & \mapsto \begin{pmatrix}
-1 && 0 \\ 0 && 1\end{pmatrix}=I_2\\
-c & \mapsto\begin{pmatrix}0 && -1\\1 && 0\end{pmatrix}=R\\ d & \mapsto\begin{pmatrix}0 && 1\\-1 && 0\end{pmatrix}=R^3
-\end{align*}
-\end{Def}
-
-Recall that a \emph{translation surface} is a compact, locally flat surface that admits a discrete number of conical singularities and has trivial linear holonomy on any arbitrary path.
-If the total cone angle of the singularity is some integer multiple of $2\pi$, $2\pi(d+1)$ for $d\geq 0$, we say that singularity has degree $d$.
-Two translation surfaces belong to the same stratum, or family, of surfaces characterized by the number and degree of conical singularities.
-Consider the effect that paths $c,d$ have on holonomy:
+A \emph{rotational monodromy map} is a morphism from the first fundamental group of a surface to the \emph{monodromy group} acting on the unit tangent bundle via parallel transport of a vector along its path.
+Existence of a parallel transport map is a consequence of the metric connection on a surface with complex structure.
+The monodromy map is well defined when the \emph{holonomy group} is \emph{discrete}.
+When the \emph{relative holonomy} of contractible loops is trivial, the holonomy group is the set of attainable rotations of a unit vector upon returning to the base point after transporting it along a non-trivial loop.
+A \emph{flat surface} like $\bGs$ has trivial relative holonomy as its neighborhoods are isometric to the plane.
+Denote the \emph{unit tangent bundles} of $\bUs$ and $\bGs$ by $T^1\bUs$ and $T^1\bGs$.
+If $\alpha$ is a loop on $\bGs$, then $[\alpha]$ acts on a fixed vector $v_0\in\RR^2$ by rotation.
+In the case where $\alpha$ is \emph{not} a geodesic segment on the surface of length $L$, $v_0=\alpha'(0)\in\RR^2$ may differ from $\alpha'(L)$ by some rotation.
+
+\begin{Def}
+Define $\varphi_2:\pi_1(\bGs)\rightarrow SO(2,\ZZ)$ to be the group homomorphism where $a,b \mapsto I_2$, $c  \mapsto R$, and $d  \mapsto R^3$, where $I_2$ is the identity matrix and $R$ is a positive $90^\circ$ rotation matrix.
+\end{Def}
 
@@
 \includegraphics[width=2.5in]{monodromy.png}
-\caption{Effect a nontrivial loop has on arbitrary basis vectors (i,j).}
+\caption{Effect a nontrivial loop has on basis vectors $(i,j)\in\RR^2$.}
 \label{fig:loop}
@@
 
-We construct a cover of $\bGs$ by taking four copies of it and cyclically pairing edges, and show that this cover has trivial holonomy.
-Denote the unit tangent bundle of $\bGs$ by $T^1\bGs$.
-Because $\bGs$ can be contained in the plane, we consider vectors tangent to $\bGs$ as vectors in $\RR^2$.
-Let $x_0\in\bGs$.
-The fiber over $x_0$ under the projection $T^1\bGs\twoheadrightarrow\bGs$ is the set $E_{x_0}$.
-Let $\triangledown$ be a flat metric connection on $T^1\bGs$ and $p=(x_0,v_0)\in E_{x_0}$ be a single element in the fiber over $x_0$ with $v_0\in\RR^2$.
-We denote the \emph{holonomy group} of the connection with base point $p$ by $Hol_p(\triangledown)=\{A\in SO(E_{x_0})~:~ p = A\cdot p\}$ where $p$ differs from $A\cdot p$ by some rotation of $v_0$.
-An element in $SO(E_{x_0})$ acts on $p$ by rotation of the vector $v_0$.
-Denote the \emph{holonomy bundle} based at $p$ by $H(p)=\{p'\in E_{x_0}~:~ p'= A\cdot p \}$, the set of elements in the orbit of $p$ where a closed loop $\gamma:[0,1]\rightarrow\bGs$ defines a horizontal lift to $\gamma^*:[0,1]\rightarrow T^1\bGs$ by parallel transport of $v_0$ along the path such that $p=\gamma^*(0)$, and $p'=\gamma^*(1)$.
-Since $\bGs$ is flat and trivial paths have trivial holonomy, $Hol_p(\triangledown)$ acts on $H(p)$ monodromy group is discrete and well-defined on $\bGs$.
-
-\begin{lem}
-$\varphi_2:\pi_1(\bGs)\rightarrow \mathrm{SO}(2,\ZZ)$ is a monodromy representation of the surface.
-\begin{proof}
-It is clear from Figure 6 that this is the case as concatenating a closed loop with paths $c$ or $d$ would rotate a vector exactly by the cone angles of the singularities that they loops around: $\pm \frac{3\pi}{2}$.
-\end{proof}
-\end{lem}
+\begin{lem}
+$\varphi_2:\pi_1(\bGs)\rightarrow \mathrm{SO}(2,\ZZ)$ is the rotational monodromy representation of $\bGs$.
+\begin{proof}
+From Figure 4 it's clear that $c,d$ will rotate a unit vector by $\pm 90^\circ$ by parallel transport along their paths.
+These values are determined by the cone angles of the singularities they loop around.
+The parallel transport of a vector along $a$ or $b$ would be trivial since the outer edges they pass through are glued by translation.
+\end{proof}
+\end{lem}
+
+\begin{lem}
+Any element in $\text{Deck}(\varphi_1^*)$ acting on $T^1\bUs$ has a trivial effect on holonomy.
+\begin{proof}
+Since $\bGs=\bUs/\ZZ^2$, the deck group action lifts naturally to an action on their tangent vector spaces and unit tangent bundles via pushforward of deck transformations $f_h,f_v$ on the bundle maps.
+i.e.
+$T^1\bGs=T^1\bUs/\ZZ^2$.
+This effect must be trivial since $\varphi_1(a)\sim f_h$ and $\varphi_1(b)\sim f_v$, and both homotopy classes $a,b$ have trivial holonomy.
+\\
+How do $f_h,f_v$ parallel transport a vector?\\
+Well if $\varphi_1([a])\cdot u_0=f_h(u_0)$, then for 
+\end{proof}
+\end{lem}
+
+\begin{thm}
+Let $m=\varphi_2\vert_{\pi_1(\bUs)}$. Then $m$ is the rotational monodromy map of $\bUs$.
+\begin{proof}
+
+\end{proof}
+\end{thm}
+
+\begin{cor}
+Let $\alpha:[0,1]\rightarrow\bGs$ be a loop on $\bGs$ with trivial holonomy and $\tilde{\alpha}$ its lift to $\bUs$. Then $\tilde{\alpha}$ has trivial holonomy as well.
+\begin{proof}
+
+\end{proof}
+\end{cor}
 
@@
 \subsection{A Four-Fold Cover}
-
+We now take these covers we built to construct a third one that covers them all. This is a four-fold cover of $\bUs$ and a $\ZZ^2$ translation cover of $\bMs$. 
 \begin{Def}
@@
 
-\compav{Prove that $\ker\varphi_1\big\vert_{\pi_1(\bMs)}=\ker\varphi_2\big\vert_{\pi_1(\bUs)}$}
-
-$\btUs$ is also a cover of $\bMs$ with fundamental group $\ker\varphi_1\big\vert_{\pi_1(\bMs)}$.
-In the former case the deck group of the cover is $\SO(2,\ZZ)$, while in the latter case it is $\ZZ^2$.
-The deck group of the cover $\btUs\rightarrow\bGs$ is $\ZZ^2\rtimes\SO(2,\ZZ)$.
-\compav{worth talking about?}.
-$\btUs$ is a four-fold cover of $\bUs$ with trivial holonomy visualized as four copies of $\bUs$, which can be thought of as countably many copies of .
-We can identify a closed path on $\bGs$ as an element of the deck group $\ZZ^2\times \SO(2,\ZZ)$ or, equivalently, $ 2\ZZ+2\ZZ i\times \ZZ/4\ZZ$ if we want to consider a metric on the ambient space $\CC$ that is consistent with the tiling of $\bGs$.
-Further, we can rotate everything in $\CC\times \ZZ/4\ZZ$ along with $\btUs$ according to the element of $\SO(2,\ZZ)$ that acts non-trivially on the holonomy of $\bUs$.
-For example, let $P\in2\ZZ+2\ZZ i$ and consider the following curve on $\bUs$:
+It follows that $\ker\varphi_1\big\vert_{\pi_1(\bMs)}=\ker\varphi_2\big\vert_{\pi_1(\bUs)}$ and that $\btUs$ is a cover of these two surfaces.
+Edge identifications are made on four copies of $\mathbf{P}$, possibly indexed by the finite component of the direct product $\mathbf{P}\times\ZZ/4\ZZ$.
+The Deck group of the covering map from $\btUs$ to $\bGs$ is the set $\ZZ^2\times \SO(2,\ZZ)$.
+Consider a loop on $\bUs$:
 
@@
 
-\noindent If we lift this curve to $\btUs$ we can identify it with an element of the deck group isomorphic to $2\ZZ+2\ZZ i \times \ZZ/4\ZZ$.
-Observe that it will only close if the curve acts trivially on the holonomy of $\bUs$:
+\noindent If we lift this curve to $\btUs$ we can identify it with an element of the deck group, and check if its lift from $\bGs$ belongs to the kernel of both $\varphi_1$ and $\varphi_2$.
+Fix a direction on this surface pointing north that respects $\bUs$'s compass and rotate every plane by either $90^\circ,$ $180^\circ,$ or $270^\circ$ as the curve travels from one plane (copy of $\mathbf{P}$) to the next.
+Observe that it will not only close if the curve acts trivially on the holonomy of $\bUs$, but with this construction we can see that $\btUs$ is actually an infinite-type translation surface:
 
@@
 \includegraphics[width=3in]{coverdirection.png}
-\end{figure}
+\caption{With $P\in 2\ZZ+2\ZZ i$.}
+\end{figure}
+
+\noindent
 
 \subsection{Geodesics on $\bUs$}
-The key connection made between surfaces $\bUs$ and $\btUs$ is that geodesics on $\bUs$ and $\btUs$ are identical in the sense that their images are in a one-to-one correspondence.
+The key connection made between surfaces $\bUs$ and $\btUs$ is that geodesics on $\bUs$ and $\btUs$ behave identically.
 Unless stated otherwise, a \emph{geodesic} is a \emph{unit-speed} locally distance minimizing curve on a smooth topological surface as a function parameterized by an element in $\RR$.
 A \emph{geodesic segment} is a restriction of the geodesic to some closed interval in $\RR$.
-A geodesic has an initial trajectory defined at 0, represented by a unit vector $v_0\in\RR^2$.
-If the vector can be scaled by $k\in\RR$ such that $kv_0\in\cO\cup\cE$, the geodesic has a \emph{rational} initial trajectory.
+A geodesic has an initial trajectory defined at time 0, represented by a unit vector $v_0\in\RR^2$.
+If the vector can be scaled by $k\in\RR$ such that $kv_0\in\cS$, the geodesic has a \emph{rational} initial trajectory.
 For example, the following is a periodic geodesic on the surface with initial direction $kv_0\in\cO$:
@@
 Otherwise they are mapped non-trivially to $\ZZ^2$ or $\ZZ^2\times\SO(2,\ZZ)$.
-Let $v_0\in\RR^2$ be a unit vector such that $\alpha's$ image under the derivative map, $\alpha'$, is $v_0$ at times $0$ and $b$.
-The derivatives of the lifts $\gamma', \gamtild'$ are also $v_0$ at their endpoints.
+Let $v_0\in\RR^2$ be a unit vector such that $\alpha's$ derivative is $v_0$ at times $0$ and $b$.
+The derivative of $\gamma$ is also $v_0$ at these times as per Corollary 1, as $\alpha$ is a geodesic with trivial holonomy.
+$\gamtild'$ are also $v_0$ at their endpoints.
+
+The Deck group of the map $\btUs\rightarrow\bUs$ is $\SO(2,\ZZ)$.
+This deck group acts naturally on the tangent bundles via push-forward of rotational matrices.
+We obtain the identity $T^1\btUs/\SO(2,\ZZ)\cong T^1\bUs$.
+This action is \emph{non-trivial}.
+Essentially, this means that there are 4 choices of a vector on the four-fold cover given a vector $v_0$ on $\bUs$.
+The choice of vector does not matter in the case where we flow it along a geodesic path, as a geodesic must return to a point in the fiber of $u_0$ in the same direction.
+That is to say that no matter where we start by lifting this path, we will always end up on the same plane (copy of $\mathbf{P}$) that we started in.
+In the case where a geodesic is periodic on $\bUs$, it would return to the same point in the fiber of $u_0$ as well.
+For simplicity's sake we will just lift to the vector mod the identity matrix, and call it $v_0$ as well.
 
@@
 \end{Def}
-
-On the Necker cube surface, a horizontal translation, $f_h$, is induced by $2\in\text{Trans}(\CC)$, and a vertical translation, $f_v$, is induced by $2i\in\text{Trans}(\CC)$.
-The group of deck transformations of the surface given by the cover $\bUs\rightarrow\bGs$ is isomorphic to the group $\Delta_{\varphi_1}=\pi_1(\bGs)/\pi_(\bUs)\cong\ZZ^2$.
-Take $\ZZ^2$ to the deck group by sending $(1,0)$ to $f_h$ and $(0,1)$ to $f_v$.
-Similarly, $\btUs$ has deck transformations given by the cover $\btUs\rightarrow\bGs$ induced by elements in $\text{Trans}(\CC\times\SO(2,\ZZ))$.
-These are isomorphic to the deck group given by the cover $\btUs\rightarrow\bUs\rightarrow\bGs$ containing elements of the form $((m,n),M)\in\ZZ^2\rtimes_{\eta}\SO(2,\ZZ)$, where $\eta:\SO(2,\ZZ)\rightarrow \Aut(\ZZ^2)$.
-$\SO(2,\ZZ)$ already belongs in $\Aut(\ZZ^2)$ so $\eta$ is an inclusion endomorphism, and the binary operation of this group is defined as  $((m,n),M)\cdot((m',n'),M')=((m,n)+M(m',n'),MM')$.\\
-\\
-Let $\varphi:\pi_1(\bGs)\rightarrow \ZZ^2\rtimes_{\eta}\SO(2,\ZZ)$ be given by $[g]\mapsto(\varphi_1([g]),\varphi_2([g]))$. $\varphi$ is a
-
+Let $\varphi=\varphi_1\times\varphi_2:\pi_1(\bGs)\rightarrow \ZZ^2\times\SO(2,\ZZ)$ be given by $[g]\mapsto(\varphi_1([g]),\varphi_2([g]))$. $\varphi$ is a
+product of group homomorphisms that takes a loop to 
 \begin{thm}
-Let $\delta=\varphi_1([\alpha])\in\ZZ^2$ be an element of $\Delta_{\varphi_2}$ and $\tilde{\delta}=\varphi([\alpha])\in\ZZ^2\rtimes_{\eta}\SO(2,\ZZ)$ be an element of $\Delta_{\varphi_{2}\vert_{\pi_1(\bUs)}}$.
-Then:\\
-$\Phi(t+b)=\delta\cdot\Phi(t)$ for all $t\in\RR$, and \\
+Let $\delta=\varphi_1([\alpha])\in\ZZ^2$ be an element of $\Delta_{\varphi_1}$ and $\tilde{\delta}=\varphi([\alpha])\in\ZZ^2\times \SO(2,\ZZ)$ be an element of $\Delta_{\varphi}$. Then:\\
+$\Phi(t+b)=\delta\cdot\Phi(t)$ for all $t\in\RR$, and 
 $\phitild(t+b)=\tilde{\delta}\cdot\phitild(t)$ for all $t\in\RR$.
@@
 
-These theorems essentially tell us that these kinds of geodesics behave identically on the Necker cube surface and its cover.
-Now we show that this is true for $\btUs$ and $\btU$.
-For these proofs we will refer to the induced homomorphisms from ($\ref{eq:commute}$):
+These theorems essentially tell us that these types of geodesics behave identically on the Necker cube surface and its cover.
+Now we show that this is true for $\btUs$ and its metric completion, $\btU$.
 
@@
 
+This follows readily from the fact that geodesics are undefined on cone singularities. A geodesic on $\btUs$ can be mapped bijectively to a geodesic on  $\btU$ and vice-a-versa.
+
 \subsection{Translation Surface}
@@
 \centering
-\includesvg[width=2.6in]{mtilda}
+\includesvg[width=2.6in]{mtildabw}
 \caption{Compact translation surface, $\mathbf{M}$ with edges and cone singularities (1,2,3,4) identified.
@@
 \centering
-\includesvg[width=3.4in]{mtildastaircase}
-\caption{The staircase Veech surface with directional planes and vertices identified.
+\includesvg[width=3.4in]{mtildastaircasebw}
+\caption{The staircase surface with directional planes and vertices identified.
 All edges are paired by translation.
-Two adjacent squares have opposite edges identified.
+Two adjacent squares have opposite edges (A-M) identified.
 The top edge of the bottom-left square is glued to the bottom edge of the top-right square (both labeled I).
@@
 
-\noindent We recall some basic definitions and theorems about $\ZZ^2-$covers of translation surfaces as they apply to $\bM$.
+\noindent We recall some basic definitions and theorems concerning $\ZZ^2$covers of translation surfaces as they apply to $\bM$.
 
@@
 for $[\gamma],[\beta]\in H_1(\mathbf{M},\QQ)$, $i([\beta],[\gamma])$ returns the signed intersection number of two homology classes.
-We say a crossing at the instance of an intersection is positive if $\gamma$ makes a positive angle relative to $\beta$.
+We say a crossing at the moment of intersection is positive if $\gamma$ makes a positive angle relative to $\beta$.
 \end{Def}
@@
 The $\emph{Veech group}$ of $\bM$ is the image in the co-domain of the homomorphism $D:\text{Aff}^+(\bM)\rightarrow SL(2,\RR)$ that takes every affine map to its derivative, which we denote $V(\bM)$.
-$\bM$ is a square-tiled translation surface whose Veech group is a finite index subgroup of $SL(2,\ZZ)$.
+$\bM$ is a square-tiled translation surface whose Veech group is a finite index subgroup of $SL(2,\ZZ)$, thereby making it a \emph{Veech surface}.
 We use automorphisms of the fundamental group in $\text{Aut}(\pi_1(\bM))$ induced by affine maps to prove our main results.
-But first, we define a group homomorphism from $\pi_1(\bM)$ to $\pi_1(\bM)/\pi_1(\btU)$ by algebraic intersection numbers pver a sum of linearly independent homology classes.
+But first, we define a group homomorphism from $\pi_1(\bM)$ to $\pi_1(\bM)/\pi_1(\btU)$ by algebraic intersection numbers over a sum of linearly independent homology classes.
 Consider the following cylinder core curves on $\bM$ labeled $\gamma_i$ for $0\leq i\leq 11$:
@@
 
-Every curve intersects two others and if we take $i(\Sigma_{i=0}^{11}[\gamma_i])$
-\compav{show that $\gamma_0,\dots,\gamma_{11}$ form a basis for $H(\bM,\QQ)$}\\
-\compav{show that u,v linearly independent}
+The rank of a $12\times12$ intersection matrix of these curves is 10. Therefore these curves span $H_1(\bM,\QQ)$.
 
@@
 \begin{Def}
-Define the group homomorphism $\Omega_{u,v}:\pi_1(\bM)\rightarrow \QQ^2$, where $\beta\mapsto(i(u,[\beta]),i(v,[\beta]))$.
-\end{Def}
-
-\begin{lem} $\Omega_{u,v}(\pi_1(\bM))=\ZZ^2$.
-\end{lem}
+Define the group homomorphism $\Omega_{u,v}:\pi_1(\bM)\rightarrow \ZZ^2$, where $\beta\mapsto(i(u,[\beta]),i(v,[\beta]))$.
+\end{Def}
+
 \begin{lem}
@@
 
-Such a map would have derivative $\left[~\begin{matrix}1 && \pm\mu^{-1}\\0 && 1\end{matrix}~\right]$.
-On $\bM$ every cylinder in both vertical and horizontal decompositions has a modulus of $\frac{1}{2}$.
+Such a map has derivative $\left[~\begin{matrix}1 && \pm\mu^{-1}\\0 && 1\end{matrix}~\right]$.
+On $\bM$ all cylinders in both vertical and horizontal decompositions have moduli $\mu=\frac{1}{2}$.
 These give way to global diffeomorphisms as $\emph{multi-twists}$ of $\bM$.
@@
 \begin{Def}
-We call the global affine diffeomorphisms in $\text{Aff}^+(\bM)$ obtained as \textbf{multi-twists of the surface in horizontal and vertical directions} $\mathbf{A}$ and $\mathbf{B}$, respectively.
-\\We define the \textbf{derivatives of} $\mathbf{A},\mathbf{B}$ as the matrices 
-$D(\mathbf{A})=\mathbf{A}'=\left[ \hspace{1mm} \begin{matrix}
+We define the global affine diffeomorphisms of $\text{Aff}^+(\bM)$ obtained as \textbf{multi-twists of the surface in horizontal and vertical directions} $\psi_h$ and $\psi_v$, respectively.
+\end{Def}
+
+The \emph{derivative map} of the affine group is the homomorphism $D:\text{Aff}^+(\bM)\rightarrow \SL(2,\RR)$.
+The \emph{Veech group} of the surface is the image of the affine group under this map, $D(\text{Aff}^+(\bM))$, denoted $V(\bM)$.
+The twists $\psi_h$ and $\psi_v$ under $D$ are mapped to the same matrices as the Dehn-twists on its individual cylinders. i.e.
+\begin{align*}
+D(\psi_h)=\psi_h'=\left[ \hspace{1mm} \begin{matrix}
 				1 &   2\\
 				0 & 1
-			\end{matrix}\hspace{1mm}\right] \text{ , and }
-			D(\mathbf{B})=\mathbf{B}'=\left[ \hspace{1mm} \begin{matrix}
+			\end{matrix}\hspace{1mm}\right] \hspace{0.5in}
+			D(\psi_v)=\psi_v'=\left[ \hspace{1mm} \begin{matrix}
 							1 & 0\\
 							 2 & 1
-						\end{matrix}\hspace{1mm}\right]$, where $D:\text{Aff}^+(\bM)\rightarrow V(\bM)$.
-\end{Def}
-These are parabolic elements of $V(\bM)$ that are directly related to the dynamical dichotomy of rational trajectories on the Necker cube surface.
-To show this algebraically, we look at how skewing the surface affects our homology spanning core curves:
+						\end{matrix}\hspace{1mm}\right]
+\end{align*}
+
+These are parabolic elements of $V(\bM)$ that generate a free group of rank 2. [cite].
+\begin{Def}
+We call $\XX$ the generated parabolic subgroup of $\text{Aff}^+(\bM)$, where $\XX=\<\psi_h,\psi_v \>$.
+\end{Def}
+
+The image of this group under $D$ is itself a subgroup of $V(\bM)$ and denoted $D(\XX)=\XX'=\<\psi_h',\psi_v' \>$ as in the introduction.
+Note that when skewing the surface and piecing it back together, the cylinder curves intersect each other:
 
@@
 
-When skewing the surface under $\mathbf{A}$ in the horizontal direction, the horizontal curves are preserved, but the vertical curves (odd $\gamma_i$ index) obtain two additional positive intersections with adjacent index (mod 12) horizontal curves.
-Similarly, the vertical $\mathbf{B}$ skews preserve the vertical curves, but the horizontal curves (even $\gamma_i$ index) obtain two additional \emph{negative} intersections with adjacent index (mod 12) vertical curves.
-The formulaic expressions of the $\mathbf{A}, \mathbf{B}$ induced automorphisms of $H_1(\bM,\QQ)$ as powers of $k\in\ZZ$ are then:
-
-\begin{align*}
-\mathbf{A}^k_*([\gamma_i])=&[\gamma_i] + \frac{k}{2}(1-(-1)^i)([\gamma_{i-1}]+[\gamma_{i+1}])\\
-\mathbf{B}^k_*([\gamma_i])=&[\gamma_i] + \frac{k}{2}(1+(-1)^i)([\gamma_{i-1}]+[\gamma_{i+1}])\\
-\end{align*}
-
-With all of these in mind we generate the following subgroups:
-
-\begin{Def}
-We say $\XX$ is the subgroup of $\text{Aff}^+(\bM)$ generated by $\mathbf{A}$, $\mathbf{B}$.
-\end{Def}
-
-\begin{Def}
-The images of $\XX$ under their derivative maps to $V(\bM)$ is the subgroup of $V(\bM)$ generated by matrices $\mathbf{A}',$ $\mathbf{B}'$. We denote this group by $\XX'$.
-\end{Def}
-
-\begin{Def}
-The induced automorphisms of $\XX$ contained in $Aut(H_1(\bM,\QQ))$ is the subgroup generated by $\mathbf{A}_*$, $\mathbf{B}_*$.
-\end{Def}
+By skewing in the horizontal direction, $\psi_h$ preserves the even indexed cylinder curves, but the odd indexed vertical curves $\gamma_i$ intersect the two horizontal curves (positive intersection), $\gamma_{i-1}$ and $\gamma_{i+1}$.
+A similar observation can be made from applying $\psi_v$ where the vertical curves are preserved, but the horizontal curves intersect with adjacent vertical curves (negative intersection).
+Let $i\in\ZZ/12\ZZ$.
+Then this convenient indexing scheme gives us the following formulas for $\psi_h$ and $\psi_v$'s counterparts in $\Aut(H_1(\bM,\QQ))$ raised to $k\in\ZZ\backslash\{0\}$:
+
+\begin{align*}
+\psi^{*k}_h([\gamma_i])=&[\gamma_i] + \frac{k}{2}(1-(-1)^i)([\gamma_{i-1}]+[\gamma_{i+1}])\\
+\psi^{*k}_v([\gamma_i])=&[\gamma_i] - \frac{k}{2}(1+(-1)^i)([\gamma_{i-1}]+[\gamma_{i+1}])\\
+\end{align*}
+
+We call this generated group $\XX^*=\<\psi^{*}_h,\psi^{*}_v \>$.
+There are no relations between its generators since $i$ being even or odd has one act trivially and the other non-trivially.
+Therefore this group is also free and rank 2.
+Hence $\XX^*\cong\XX'\cong\XX$.
+Let $g\in\XX$, $g'\in\XX'$, $g^*\in\XX^*$ be isomorphic elements in their respective groups.
+Let $g=\prod_{j=0}^{n}\psi_h^{k_j}\circ\psi_v^{m_j}$ be a word in $\XX$ as a finite left product of multi-twist pairs with $k_j,m_j\in\ZZ$.
+Let $\psi_j=\psi_h^{k_j}\circ\psi_v^{m_j}$ be the $j^{th}$ word pair.
+Elements $g',g^*$ have identical representations.
+\\\\
+The $j^{th}$ element pair that is mapped into $\Aut(H_1(\bM,\QQ))$ can be expressed as
+
+\begin{align*}
+\psi_j^*([\gamma_i])=&[\gamma_i]+\frac{k_j(1-(-1)^i)}{2}([\gamma_{i-1}]+[\gamma_{i+1}])\\&-\frac{m_j(1+(-1)^i)}{2}(k_j(2[\gamma_i]+[\gamma_{i-2}]+[\gamma_{i+2}]) +[\gamma_{i-1}]+[\gamma_{i+1}]).
+\end{align*}
+
+\noindent Let $\sum\Gamma=\sum_{i=0}^{11}[\gamma_i]$, and denote the odd and even indexed sums as $\sum\Gamma_o$ and $\sum\Gamma_e$, respectively.
+We consider three kinds of closed geodesic segments on $\bM$ and don't consider initial/terminal points for the time being.
+These are the slope one, horizontal, and vertical loops.\\
+
+\noindent Call the geodesic in direction $\theta=\frac{\pi}{4}$ that runs up the staircase intersecting every $\gamma_i$ once $\chi$.
+Denote its homotopy class by $[\chi]$.
+Observe from Figure 10 that $\chi$ is homologous to the sum of cylinder curves divided by 2.
+In Figure 1, this is the closed geodesic that lifts to a closed loop that looks like a 12-gon on $\bUs$.
+
+\begin{lem}
+$\Omega_{u,v}([\chi])=(0,0).$
+\begin{proof}
+Observe that $\Omega_{u,v}([\chi])=\Omega_{u,v}(\frac{1}{2}\sum\Gamma)=\frac{1}{2}(i(u,\sum\Gamma),i(v,\sum\Gamma)).$ Since only adjacent curves intersect, $i(u,\sum\Gamma)=i(-[\gamma_2],[\gamma_{1}]+[\gamma_{3}]) +i([\gamma_5],[\gamma_{4}]+[\gamma_{6}]) + i([\gamma_8],[\gamma_{7}]+[\gamma_{9}]) - i([\gamma_{11}],[\gamma_{10}]+[\gamma_{0}])=-2+(-2)+2-(-2)=0$.\\
+Similarly, $i(u,\sum\Gamma)=i([\gamma_0],[\gamma_{11}]+[\gamma_{1}]) -i([\gamma_3],[\gamma_{2}]+[\gamma_{4}]) -i([\gamma_6],[\gamma_{5}]+[\gamma_{7}]) +i([\gamma_9],[\gamma_{8}]+[\gamma_{10}])=2-(-2)-2+(-2)=0$.
+Therefore, $\Omega_{u,v}([\chi])=(0,0)$.
+\end{proof}
+\end{lem}
+
+\begin{lem}
+Let $g$ be as above. If $\alpha=g(\chi)$, then $[\alpha]=\frac{1}{2}(c_1\sum_e+c_2\sum_o)$ for $c_1,c_2\in\ZZ$.
+\begin{proof}(By Induction on $\XX^*$).\\
+Consider $\psi_0(\chi)$.\\
+$[\psi_0(\chi)]=\psi_0^*[\chi]=\frac{1}{2}\psi_0^*(\sum\Gamma)=\frac{1}{2}\psi_0^*(\sum\Gamma_e+\sum\Gamma_o)=\frac{1}{2}(\psi_0^*(\sum\Gamma_e)+\psi_0^*(\sum\Gamma_o))$.
+Now\\ $\psi_0^*(\sum\Gamma_e)=\sum_{i=0}^{11}\psi_0^*([\gamma_{2i}])=\sum_{i=0}^{11}( [\gamma_{2i}]-2m_0k_0[\gamma_{2i}]-m_0k_0[\gamma_{2i-2}] $\\$ -m_0k_0[\gamma_{2i+2}]-m_0[\gamma_{2i+1}]-m_0[\gamma_{2i-1}] )=(1-4m_0k_0)\sum\Gamma_e+(-2m_0)\sum\Gamma_o$.\\
+Now\\ $\psi_0^*(\sum\Gamma_o)=\sum_{i=0}^{11}\psi_0([\gamma_{2i+1}])=\sum_{i=0}^{11}([\gamma_{2i+1}] +k_0([\gamma_{2i}]+[\gamma_{2i+2}]))=\sum\Gamma_o+2k_0\sum\Gamma_e$.\\
+And so $\frac{1}{2}\psi_0^*(\sum\Gamma)=\frac{1}{2}((1+2k_0-4k_0j_0)\sum\Gamma_e +(1-2m_0)\sum\Gamma_o )$.\\\\
+Induction: Suppose $\frac{1}{2}\psi^*_j(\sum\Gamma)=c_1\sum\Gamma_e+c_2\sum\Gamma_o$.\\
+Then $\psi^*_{j+1}(\psi^*_{j}(\sum\Gamma))=c_1\psi^*_{j+1}(\sum\Gamma_e)+c_2\psi^*_{j+1}(\sum\Gamma_o)=\\
+c_1((1-4m_{j+1}k_{j+1})\sum\Gamma_e+(-2m_{j+1})\sum\Gamma_o)+c_2(\sum\Gamma_o+2k_{j+1}\sum\Gamma_e)=\\
+(c_1(1-4m_{j+1}k_{j+1}) + 2k_{j+1}c_2)\sum\Gamma_e+(-2m_{j+1}c_1+c_2)\sum\Gamma_o$.
+
+\end{proof}
+\end{lem}
+
+\begin{thm}
+Let $g$ be as above. Suppose $\alpha=g(\chi)$. Then $\Omega_{u,v}([\alpha])=(0,0)$. 
+\begin{proof}
+$\Omega_{u,v}([\alpha])=\Omega_{u,v}([g(\chi)])=\Omega_{u,v}(g^{*}([\chi]))=\frac{1}{2}(i(u,g^*(\sum\Gamma)) i(v,g^*(\sum\Gamma))$.\\
+$i(u,g^*(\sum\Gamma)=c_1 i(u,\sum
+\Gamma_e)+c_2 i(u,\sum\Gamma_o)=\\ c_1(i([\gamma_5],[\gamma_{4}] - i([\gamma_{11}],[\gamma_{10}]+[\gamma_{0}]))\\+c_2(i(-[\gamma_2],[\gamma_{1}]+[\gamma_{3}]) + i([\gamma_8],[\gamma_{7}]+[\gamma_{9}]) )=\\c_1((-2)-(-2))+c_2(-2+ 2 )=0\\$
+Holds similarly for $i(v,g^*(\sum\Gamma)).$\\
+$\therefore \Omega_{u,v}([\alpha])=(0,0)$.
+\end{proof}
+\end{thm}
+
+\begin{lem}
+$\Omega_{u,v}([\gamma_i])\neq (0,0)$.
+\begin{proof}
+every curve intersects nontrivially with exactly one of either $u$ or $v$.
+\end{proof}
+\end{lem}
+
+\begin{thm}
+Suppose $\beta=g(\gamma_i)$. Then $\Omega_{u,v}([\beta])\neq (0,0)$.
+\begin{proof}
+$g^*([\gamma_i])=[\gamma_i]+C$.
+If $i(u,g^*([\gamma_i]))=i(u,[\gamma_i]+C)=i(u,[\gamma_i])+i(u,C)=0$, then 
+\end{proof}
+\end{thm}
+
 \section{Proof of Main Theorem}
@@
 $S=\cO\cup \cE$.
-
-\begin{lem}
-Let $\XX'$ act on $\cS$ by left matrix multiplication.
-Then $\XX'$ acts faithfully and transitively on integer pairs in $\cO$, partitions $\cE$ into two sets and acts faithfully and transitively on them, and acts faithfully on $\cS$.
+Let $\XX'$ act on $\cS$ by linear transformations.
+
+\begin{lem}
+
+\end{lem}
+
+\begin{lem}
+$\XX'$ acts faithfully and transitively on integer pairs in $\cO$, partitions $\cE$ into two sets and acts faithfully and transitively on them, and acts faithfully on $\cS$.
 \begin{proof}
@@
 
-\begin{lem}
-Let $v_0\RR^2$. Then for any geodesic $\alpha$
-\end{lem}
+\begin{Def}
+Let $\phi_t^\theta:\RR\rightarrow\bM$ be a unit-speed geodesic on $\bM$ in direction $\theta$.
+\end{Def}
 

### 2017-12-08 newpaper.tex
120717

@@
 \def\utild{\tilde{u}_0}
+\def\mtild{\tilde{m}_0}
 \def\gamtild{\tilde{\gamma}}
@@
 Given a rational vector direction in $\RR^2$ (components are relatively prime integers) obtained from $\theta$, one can classify every periodic/drift-periodic rational geodesic on the surface.
-In addition its total period/length and, in the case where it is drift-periodic, translational drift as Euclidean distance in its ambient space is given by solving a system of simple linear Diophantine equations.
+In addition its total period/length and, in the case where it is drift-periodic, translational drift as Euclidean distance in its ambient space is given by solving a system of Diophantine equations seemingly related to Farey sequences.
 \end{abstract}
@@
 \label{fig:front}
-\caption{Periodic and drift-periodic flows on the Necker cube surface}
-\end{figure}
-
-\newpage
+\caption{Periodic and drift-periodic geodesics on the Necker cube surface}
+\end{figure}
 \section{Introduction}
-
-Put your Problem statement here! Example of a Citation\cite[p.219]{Robotics}. Here's Another Citation \cite{Flueck}
-
-The Necker cube has made numerous appearances in the work of mathematicians, crystallographers, and scientists interested in human visual systems before and after appearing in the works of M.C.
+The Necker cube has made numerous appearances in the work of mathematicians, crystallographers, and scientists interested in human visual systems even before appearing in the works of M.C.
 Escher [cite].
-Louis Albert Necker was credited for having discovered the optical illusion, and studying its geometric properties [cite].
-The solid presentation of the cube, when rendered as a flat surface with three rear faces removed, achieves a similar effect when its three visible faces are shaded in a particular way.
-The Necker cube surface is then obtained by gluing infinitely many copies along its outer edges as in Figure 1.
-
-Dynamical classification of geodesics on flat, periodic surfaces such as the Necker cube surface ultimately boils down to  studying the surface's various symmetries and how the rotational holonomy group acts on the unit tangent bundle of the surface as the unit vector (derivative of the geodesic) is parallel transported along its path.
-Loops on polyhedra punctured at its cone singularities tend to have discrete holonomy groups as its relative holonomy (holonomy of contractible loops) is trivial.
+Louis Albert Necker was credited for having discovered this optical illusion and studying its various properties [cite].
+The solid presentation of the cube, when rendered as a flat surface with three rear faces removed, achieves a similar effect when its three visible faces are shaded in a particular way not unlike the illusion given off by the ``wire-frame" cube.[cite] The Necker cube surface is then obtained by gluing infinitely many copies of the semi-cube surface along its outer edges (Fig 1).
+
+Dynamical classification of geodesics on flat, infinite surfaces admitting cone singularities such as the Necker cube surface ultimately boils down to studying the surface's and, if it exists, base surface's various symmetries.
+The rotational holonomy group acts on the unit tangent bundle of the surface as a unit vector (derivative of the geodesic at any given time) undergoes parallel transport along the geodesic path.
+If a surface is periodically constructed and obtained as a normal cover of some elementary base surface (like a polygon with edge identifications), this holonomy group might be given up by observing how the push-forward of elements in the group of deck transformations acts on tangent spaces.
+In particular, loops on polyhedral surfaces punctured at its  cone singularities (cone angles that are rational multiples of $\pi$) tend to have discrete holonomy groups as its relative holonomy (holonomy of contractible loops) is trivial.
 
 \begin{rem}
-Familiarity is assumed on the part of the reader with covering space theory, translation (Veech) surfaces, and their associated Veech groups. For general surveys on these topics: [cite], [cite].
+Familiarity is assumed on the part of the reader with covering space theory, translation (particularly Veech) surfaces, and their associated Veech groups.
+For general surveys on these topics we refer readers to: [cite], [cite].
 \end{rem}
@@
 \subsection{Discussion of Results}
-Our initial experiments strongly supported the theory that there would be a correlation between a choice of trajectory angle and dynamical properties of a geodesic on such a symmetric object.
-Rightfully so, a surface composed of infinitely many cubes an induced flat metric where every neighborhood is locally isometric to the Euclidean plane.
-Via the parallel transport of a unit tangent vector over the sharp edges of the surface denoted $\bUs$, it becomes necessary to think of the geodesic as a sequence of line segments contained in the faces of each cube on the surface's embedded form in $\RR^3$.
-$\ref{fig:front}$.
+Our initial experiments supported the theory that there would be a strong correlation between a choice of initial trajectory angle and the dynamical properties of a geodesic traveling on the surface.
+[throw in some trajectories in the appendix] A surface composed of infinitely many cubes is endowed with a flat metric, i.e.
+every neighborhood is locally isometric to $\RR^2$.
+Via the parallel transport of a unit vector tangent to the surface in the direction of the geodesic, we view its image as a sequence of straight lines on the (smooth) faces of the surface that tightly wrap around its sharp edges (Fig 1).
 \\
@@
 \noindent Let $u_0$ be a point on the surface that is contained in the interior of a face and consider a tangent unit vector $v\in\RR^3$ based at point $u_0$.
-Each face is parallel to a $2$-dimensional subspace of $\RR^3$ spanned by exactly two basis vectors, so one component of $v$ is 0.
-Call this projected vector $v_0\in\RR^2$.
+Each face is parallel to a $2$-dimensional subspace of $\RR^3$ spanned by exactly two basis vectors, so one component of $v$ is always 0.
+Thus $v$ is is isometric to $v_0\in\RR^2$.
 \\\\\\
-The angle $v_0$ makes relative to a choice of basis is the \emph{initial trajectory angle} of the unit-speed geodesic on the surface, $\Phi_t:\RR\rightarrow\bUs$.
+Denote the Necker cube surface $\bUs$.
+The angle $v_0$ makes relative to a choice of basis is the \emph{initial trajectory angle} of a \emph{unit-speed geodesic} based at $u_0$ on the surface, $\Phi_t:\RR\rightarrow\bUs$.
+The initial trajectory angle is said to be $\emph{rational}$ if there is some $k\in\RR$ such that $kv_0$'s components are relatively prime integers.
 A rational initial trajectory angle in $\RR^2$ falls into one of two categories:
 \begin{Def}
-Let $v_0$ be a unit vector of the form $\frac{1}{k}(x,y)\in\RR^2$ with $x,y\in\mathbb{Z}$ and $k=\sqrt{x^2+y^2}\in\mathbb{R}$.
-We say $v_0$ is an \textbf{odd-odd} vector if its components are relatively prime and both odd.
+Let $v_0$ be a rational unit vector of the form $\frac{1}{k}(x,y)\in\RR^2$ with $x,y\in\mathbb{Z}$ and $k=\sqrt{x^2+y^2}\in\mathbb{R}$.
+We say $kv_0$ is an \textbf{odd-odd} vector if $x,y$ are relatively prime and both odd.
 We denote the \textbf{set of all odd-odd directions} by $\mathcal{O}$.
-We say that $v_0$ is an \textbf{even-odd} vector if its components are relatively prime and of opposite parity.
+We say that $kv_0$ is an \textbf{even-odd} vector if $x,y$ are relatively prime and of opposite parity.
 We denote the \textbf{set of all even-odd directions} by $\mathcal{E}$.
 \end{Def}
-\noindent For convenience we call the disjoint union of these sets $\cS$.
+\noindent Let $\cS=\cO\sqcup\cE$ be the disjoint union of the two.
 This paper aims to depict the relationship between initial trajectory angles and dynamical properties of a geodesic on the surface.
@@
 Recall that a geodesic on the surface is \emph{periodic} if there exists some $T>0$ such that $\Phi(t+T)=\Phi(t)$ for all $t\in\RR$ and a geodesic is \emph{drift-periodic} if there exists some $T'>0$ and non-trivial $p\in \text{Trans}(\bUs)$ such that $\Phi(t+T')=p(\Phi(t))$ for all $t\in\RR$.
-In either case we call this value the \emph{period} of $\Phi$ if it is the \emph{smallest} possible value that satisfies our definition.
-We additionally require that at no point does $\Phi$ encounter a cone singularity of the surface, in which case the geodesic behavior is non-deterministic.
+In either case we call this value the \emph{period} of $\Phi$ if it is the \emph{smallest} possible value that satisfies this definition.
+We additionally require that at no point does $\Phi$ encounter a cone singularity of the surface.
+Otherwise the geodesic behavior is indeterminable.
+We call $\Phi$ \emph{non-singular} if this holds.
 
@@
 
-\noindent Section [blank] explains the significance of this group. Let $M=\left[ \hspace{1mm} \begin{matrix}
-							s & u\\
-							 w & z
-						\end{matrix}\hspace{1mm}\right]\in\XX'$.
+\noindent Section [blank] explains the significance of this group. Let $M\in\XX'$.
 
 \begin{thm*}
-(Dynamical Properties of a Geodesic on $\bUs$) \\Let $\Phi_t:\RR\rightarrow\bUs$ be a non-singular unit-speed geodesic on the Necker cube surface such that $\Phi(0)=u_0$ and $\Phi'(0)=v\in\RR^3$, isometric to $v_0\in\RR^2$.
+(Dynamical Classification of Geodesics on the Necker Cube Surface) \\Let $\Phi_t:\RR\rightarrow\bUs$ be a non-singular unit-speed geodesic on the Necker cube surface such that $\Phi(0)=u_0$ and $\Phi'(0)=v\in\RR^3$, isometric to $v_0\in\RR^2$.
 Let $\overline{u}_t\in\RR^3$ be $\Phi_t$'s coordinate in $\bUs$'s ambient space at time $t$.
+Let $k\in\RR$ such that $kv_0\in\cS$.
 Then the following is true:
 \begin{enumerate}[label=(\roman*)]
-\item $\Phi$ is periodic if $v_0\in\cO$.
-\item Suppose $(i)$ is true. If $M (kv_0)=(1,1)$ then $\Phi$ has period $T=6\sqrt{2}||kv_0||$.
+\item $\Phi$ is periodic if $kv_0\in\cO$.
+\item Suppose $\Phi$ is periodic. If $M (kv_0)=(1,1)$ then $\Phi$ has period $T=6\sqrt{2}||kv_0||$.
 \item $\Phi$ is drift-periodic if $v_0\in\cE$. 
-\item Suppose $(iii)$ is true. If $M (kv_0)=(1,0)$ or $M (kv_0)=(0,1)$ then $\Phi$ has period $T=2||kv_0||$, and $d_E(\overline{u}_t,\overline{u}_{t+T})=||kv_0||\sqrt{2}$.
+\item Suppose $\Phi$ is drift-periodic. If $M (kv_0)=(1,0)$ or $M (kv_0)=(0,1)$ then $\Phi$ has period $T=2||kv_0||$, and $d_E(\overline{u}_t,\overline{u}_{t+T})=||kv_0||\sqrt{2}$.
 \end{enumerate}
@@
 \subsection{Translation Surface}
-Consider the projection $\btU\rightarrow\mathbf{M}$ onto a quotient of the surface under its translational symmetries.
-What we obtain is the following compact translation surface belonging to the stratum $\cH(2,2,2,2):$
+Consider the projection $\btU\rightarrow\mathbf{M}$ onto a quotient of the surface under its translational symmetries of each plane (see Fig 6).
+What we obtain is a genus 5 finite translation surface $\bM-$ also a branched cover of $\bG$.
+$\bMs$ was obtained as a cover of $\bGs$ from the rotational monodromy map on $\pi_1(\bGs)$, so $\bM$ is just its metric completion by reintroducing these vertices.
+There are four cone singularities on this surface, as opposed to the three on $\bG$, each of cone angle $6\pi$.
+Keeping track of $\btU$'s compass pointing \emph{true} north, the individual sections of $\bM$, labeled $I,II,III,$ and $IV$, are rotated in such a way to \emph{trivialize} \emph{holonomy} on arbitrary paths (Fig 8).
+Observe that $\bM$ is really a square-tiled/origami surface constructed out of unit squares (Fig 9).
 
@@
 The Roman numerals are meant to identify every ``copy" of $\bG$ with a direction differing from $v_0$ by rotation of some element in $\SO(2,\ZZ)$.
-The individual copies are then rotated so we have a canonical notion of direction }
+The individual copies are then rotated so every edge is identified by translation.}
 \label{fig:mtilda}
 \end{figure}
-
-We rearrange $\bM$ as such:
 
@@
 
-\noindent We recall some basic definitions and theorems concerning $\ZZ^2$covers of translation surfaces as they apply to $\bM$.
-
-\begin{Def}
-Algebraic intersection number is a non-degenerate bilinear form:
-\begin{align*}
-i:H_1(\bM,\QQ)\times H_1(\bM,\QQ)\rightarrow\QQ,
-\end{align*}
-for $[\gamma],[\beta]\in H_1(\mathbf{M},\QQ)$, $i([\beta],[\gamma])$ returns the signed intersection number of two homology classes.
-We say a crossing at the moment of intersection is positive if $\gamma$ makes a positive angle relative to $\beta$.
-\end{Def}
-The set of all affine diffeomorphisms of $\bM$ form the group $\text{Aff}^+(\bM)$.
-The $\emph{Veech group}$ of $\bM$ is the image in the co-domain of the homomorphism $D:\text{Aff}^+(\bM)\rightarrow SL(2,\RR)$ that takes every affine map to its derivative, which we denote $V(\bM)$.
-$\bM$ is a square-tiled translation surface whose Veech group is a finite index subgroup of $SL(2,\ZZ)$, thereby making it a \emph{Veech surface}.
-We use automorphisms of the fundamental group in $\text{Aut}(\pi_1(\bM))$ induced by affine maps to prove our main results.
-But first, we define a group homomorphism from $\pi_1(\bM)$ to $\pi_1(\bM)/\pi_1(\btU)$ by algebraic intersection numbers over a sum of linearly independent homology classes.
-Consider the following cylinder core curves on $\bM$ labeled $\gamma_i$ for $0\leq i\leq 11$:
+\noindent We recall some basic properties of $\ZZ^2$ covers of translation surfaces as they apply to $\bM$.
+
+The set of all orientation preserving affine diffeomorphisms of $\bM$ form the group $\text{Aff}^+(\bM)$.
+The \emph{Veech group} of $\bM$ is the image of $\text{Aff}^+(\bM)$ in $SL(2,\RR)$ under the derivative map, $D:\text{Aff}^+(\bM)\rightarrow SL(2,\RR)$.
+We denote this group $V(\bM)$.
+It is a well-known fact that square-tiled surfaces like $\bM$ are \emph{Veech} and have Veech groups of finite index in $\SL(2,\ZZ)$ [Gutkin-Judge].
+We use $V(M)$ and the $\text{Aff}^+(\bM)$-induced automorphism group, $\text{Aut}(\pi_1(\bM))$, as they act on geodesics and their homotopy classes.
+When we reconstructed $\bM$ as a staircase, we kept track of the outer edges of the surface that were identified with one another under the projection map defined on $\btU$.
+These now form cylinder core curves that are divided into homology classes $u,v,z$ (Fig 10).
+We denote the \emph{set of all core curve homology classes} by $\Gamma=\{[\gamma_i]~:~0\leq i \leq 11\}$.
+
 \begin{figure}[H]
@@
 
-The rank of a $12\times12$ intersection matrix of these curves is 10. Therefore these curves span $H_1(\bM,\QQ)$.
-
-\begin{Def}
-The homology classes $u,v,z$ are given as the following sums of core curves:
+We assign $u,v$ classes integer weights according to what translational effect \emph{crossing} the curve would have on a loop being lifted to the cover $\btU$.
+For example, a path crossing $[\gamma_0]$ and making a positive angle relative to the direction $\gamma_0$ is moving in (right) corresponds to moving ``up" once on $\btU$ from whatever section in the fiber of $\bM$ under the projection map on $\btU$ the lifted path might have belonged to prior to intersecting with $\gamma_0$.
+We would then assign to $[\gamma_0]$ the weight $(0,+1)\in\ZZ^2$.
+Crossing a curve in the $z$ class really does not take a path to another element in the fiber.
+
+\begin{Def}
+Algebraic intersection number is a non-degenerate bilinear form:
+\begin{align*}
+i:H_1(\bM,\QQ)\times H_1(\bM,\QQ)\rightarrow\ZZ.
+\end{align*}
+For $[\alpha],[\beta]\in H_1(\mathbf{M},\QQ)$, $i([\alpha],[\beta])$ returns the signed intersection number of two homology classes.
+We say a crossing at the moment of intersection is positive (+1) if $\beta$ makes a positive angle relative to the direction $\alpha$ is traveling in.
+A crossing is negative (-1) if $\beta$ makes a negative angle relative to $\alpha$.
+\end{Def}
+
+By observation, these signed weights are determined as follows:
+
+\begin{Def}
+The homology classes $u,v,z$ are given as the following weighted sums of elements of $\Gamma$:
 \begin{align*}
@@
 
-\begin{Def}
-Define the group homomorphism $\Omega_{u,v}:\pi_1(\bM)\rightarrow \ZZ^2$, where $\beta\mapsto(i(u,[\beta]),i(v,[\beta]))$.
-\end{Def}
-
-\begin{lem}
-\compav{show somehow that u,v determine the cover}
-\end{lem}
+\begin{rem}
+To be precise, we can think of $\btU$ as a $\ZZ^2$ cover only because we take 8 homology classes and apply the necessary transformations (change of sign) that agree with the deck group of the covering map from $\btUs$ to $\bGs$, $\ZZ^2\times\SO(2,\ZZ)$.
+Note that this group is isomorphic to $\ZZ^8$.
+\end{rem}
+
+By constructing a $12\times12$ intersection matrix and calculating its rank (rank=10), we can see that $\Gamma$ does span $H_1(\bM,\QQ)\cong\ZZ^{10}$ and does not need every cylinder curve.
+Certain relations can be discerned, such as $u+v=z$, to reduce the set to ten elements but the indexing scheme of $\Gamma$ here is far easier to work with, being in agreement with $\bM$'s cylinder decomposition in vertical and horizontal directions.
+This comes at the cost of the vector space being overdetermined and not having a coordinate basis, but it is not necessary to have one here.
+
+\begin{Def}
+Fix $m_0\in\bM$ and let $\mtild\in\btU$ be a point in the fiber of $m_0$.
+Let $\Omega_{u,v}:\pi_1(\bM)\rightarrow \ZZ^2$ be the homomorphism relating an $m_0$-based loop's intersection number with homology classes $u,v$ to the terminating point of its lifted path based at $\mtild$ on $\btU$.
+If $\beta$ is such a loop, then
+\begin{align*}
+\Omega_{u,v}([\beta])=(i(u,[\beta]),~i(v,[\beta]))
+\end{align*}
+\end{Def}
+
+\begin{thm}
+The cover $\btU$ has first fundamental group $\ker\Omega_{u,v}$.
+\begin{proof}
+\end{proof}
+\end{thm}
 
@@
 We look at some important automorphisms induced by Affine maps on $\bM$.
-Observe that $\bM$ has a uniform cylinder decomposition in both horizontal and vertical directions as in figure $15$.
+Observe that $\bM$ has a uniform cylinder decomposition in both horizontal and vertical directions (Fig 15).
 We define the \emph{modulus}, $\mu$, of a cylinder to be the ratio of the cylinder's width to its circumference, $\frac{w}{c}$.
-The $\emph{Dehn-twist}$ of a cylinder is an affine diffeomorphism that skews the cylinder and sends every vertex to itself.
+The $\emph{Dehn-twist}$ of a cylinder is an affine diffeomorphism that skews the cylinder and preserves edges and vertices.
 
@@
 On $\bM$ all cylinders in both vertical and horizontal decompositions have moduli $\mu=\frac{1}{2}$.
-These give way to global diffeomorphisms as $\emph{multi-twists}$ of $\bM$.
+These give up global diffeomorphisms as $\emph{multi-twists}$ of $\bM$ with the same derivative.
+[cite]
 
@@
 
-The \emph{derivative map} of the affine group is the homomorphism $D:\text{Aff}^+(\bM)\rightarrow \SL(2,\RR)$.
-The \emph{Veech group} of the surface is the image of the affine group under this map, $D(\text{Aff}^+(\bM))$, denoted $V(\bM)$.
-The twists $\psi_h$ and $\psi_v$ under $D$ are mapped to the same matrices as the Dehn-twists on its individual cylinders. i.e.
+The twists $\psi_h$ and $\psi_v$ under derivative map $D$ are mapped to the same matrices as the Dehn-twists. i.e.
 \begin{align*}
@@
 \begin{Def}
-We call $\XX$ the generated parabolic subgroup of $\text{Aff}^+(\bM)$, where $\XX=\<\psi_h,\psi_v \>$.
-\end{Def}
-
-The image of this group under $D$ is itself a subgroup of $V(\bM)$ and denoted $D(\XX)=\XX'=\<\psi_h',\psi_v' \>$ as in the introduction.
-Note that when skewing the surface and piecing it back together, the cylinder curves intersect each other:
+Let $\XX$ be the generated parabolic subgroup of $\text{Aff}^+(\bM)$, where $\XX=\<\psi_h,\psi_v \>$.
+\end{Def}
+
+The image of this group under $D$ is itself a subgroup of $V(\bM)$ and denoted $D(\XX)=\XX'=\<\psi_h',\psi_v' \>$ as in page 2.
+When skewing the surface and piecing it back together under $\psi_h$ or $\psi_v$, the cylinder curves intersect each other in pairs:
 
@@
 By skewing in the horizontal direction, $\psi_h$ preserves the even indexed cylinder curves, but the odd indexed vertical curves $\gamma_i$ intersect the two horizontal curves (positive intersection), $\gamma_{i-1}$ and $\gamma_{i+1}$.
-A similar observation can be made from applying $\psi_v$ where the vertical curves are preserved, but the horizontal curves intersect with adjacent vertical curves (negative intersection).
+A similar observation can be made from applying $\psi_v$, where the vertical curves are preserved but the horizontal curves intersect with adjacent vertical curves (negative intersection).
 Let $i\in\ZZ/12\ZZ$.
-Then this convenient indexing scheme gives us the following formulas for $\psi_h$ and $\psi_v$'s counterparts in $\Aut(H_1(\bM,\QQ))$ raised to $k\in\ZZ\backslash\{0\}$:
+Then the following formulas hold for $\psi_h$ and $\psi_v$'s counterparts in $\Aut(H_1(\bM,\QQ))$ raised to $k\in\ZZ\backslash\{0\}$:
 
@@
 
-We call this generated group $\XX^*=\<\psi^{*}_h,\psi^{*}_v \>$.
-There are no relations between its generators since $i$ being even or odd has one act trivially and the other non-trivially.
-Therefore this group is also free and rank 2.
-Hence $\XX^*\cong\XX'\cong\XX$.
-Let $g\in\XX$, $g'\in\XX'$, $g^*\in\XX^*$ be isomorphic elements in their respective groups.
-Let $g=\prod_{j=0}^{n}\psi_h^{k_j}\circ\psi_v^{m_j}$ be a word in $\XX$ as a finite left product of multi-twist pairs with $k_j,m_j\in\ZZ$.
-Let $\psi_j=\psi_h^{k_j}\circ\psi_v^{m_j}$ be the $j^{th}$ word pair.
-Elements $g',g^*$ have identical representations.
+If $k=0$, then it's just the identity.
+We generate the group $\XX^*=\<\psi^{*}_h,\psi^{*}_v \>$.
+There are no relations between its generators since $i$, being even or odd, has one automorphism act trivially while the other acts non-trivially.
+Therefore this group is also a free group of rank 2.
+All of our related groups are free groups, so we have that $\XX^*\cong\XX'\cong\XX$.
+
+Let $\tau\in\XX$.
+Let $D(\tau)=\tau'\in\XX'$ and $\tau^*\in\XX^*$ be isomorphic elements in their respective groups.
+We say $\tau=\prod_{j=0}^{n}\psi_h^{k_j}\circ\psi_v^{m_j}$ is a word in $\XX$ as a finite left concatenated product of multi-twist pairs with $k_j,m_j\in\ZZ$.
+Let $\psi_j=\psi_h^{k_j}\circ\psi_v^{m_j}$ be the $j^{th}$ multi-twist pair.
+Since $\tau',\tau^*$ are isomorphic, we can use their expressions interchangeably.
 \\\\
-The $j^{th}$ element pair that is mapped into $\Aut(H_1(\bM,\QQ))$ can be expressed as
+The $j^{th}$ automorphism pair is an element $\psi_j^*\in\Aut(H_1(\bM,\QQ))$ and can be expressed as
 
@@
 
-\noindent Let $\sum\Gamma=\sum_{i=0}^{11}[\gamma_i]$, and denote the odd and even indexed sums as $\sum\Gamma_o$ and $\sum\Gamma_e$, respectively.
-We consider three kinds of closed geodesic segments on $\bM$ and don't consider initial/terminal points for the time being.
-These are the slope one, horizontal, and vertical loops.\\
-
-\noindent Call the geodesic in direction $\theta=\frac{\pi}{4}$ that runs up the staircase intersecting every $\gamma_i$ once $\chi$.
+\noindent Similarly,
+
+\begin{align*}
+\psi_j'=\left[ \hspace{1mm} \begin{matrix}
+				1 &   2\\
+				0 & 1
+			\end{matrix}\hspace{1mm}\right]^{k_j}\left[ \hspace{1mm} \begin{matrix}
+							1 &   0\\
+							2 & 1
+						\end{matrix}\hspace{1mm}\right]^{m_j}=\left[ \hspace{1mm} \begin{matrix}
+										1+4m_jk_j &   2k_j\\
+										2m_j & 1
+									\end{matrix}\hspace{1mm}\right]
+\end{align*}
+
+\noindent Let $\sum\Gamma=\sum_{i=0}^{11}[\gamma_i]$, and denote the odd and even indexed sums of core curves $\sum\Gamma_o$ and $\sum\Gamma_e$, respectively.
+We look at three kinds of closed geodesic segments on $\bM$ and don't consider initial/terminal points for the time being.
+These are the geodesics in directions $\theta=0,\frac{\pi}{2},$ and $\frac{\pi}{4}$.
+The aim is to show that these linear automorphisms take $\sum\Gamma\mapsto c_1\sum\Gamma_e+c_2\sum\Gamma_o$ and an individual $[\gamma_i]$ to an element that maps non-trivially to $\ZZ^2$ under $\Omega_{u,v}$.
+\\
+
+\noindent Let $\chi$ be the geodesic segment in direction $\theta=\frac{\pi}{4}$ that runs up the staircase intersecting every $\gamma_i$ exactly once.
 Denote its homotopy class by $[\chi]$.
-Observe from Figure 10 that $\chi$ is homologous to the sum of cylinder curves divided by 2.
-In Figure 1, this is the closed geodesic that lifts to a closed loop that looks like a 12-gon on $\bUs$.
+Observe from Figure 10 that $\chi$ is homologous to $\frac{1}{2}\sum\Gamma$, a curve that climbs up the staircase.
+When lifted to $\btU$ and projected onto $\bUs$, this is a closed segment on the Necker cube surface (Fig 1).
 
@@
 
-\begin{lem}
-Let $g$ be as above. If $\alpha=g(\chi)$, then $[\alpha]=\frac{1}{2}(c_1\sum_e+c_2\sum_o)$ for $c_1,c_2\in\ZZ$.
-\begin{proof}(By Induction on $\XX^*$).\\
+Thus $\chi$ lifts to a closed geodesic segment on $\btU$.
+Now any geodesic segment in the orbit of $\chi$ under $\XX$ is taken into consideration, and we see why taking sums of gamma curves to sums of even and odd curves are so important.
+
+\begin{lem}
+Let $\tau$ be as above. If $\alpha=\tau(\chi)$ is a closed geodesic segment on $\bM$, then $[\alpha]=\frac{1}{2}(c_1\sum_e+c_2\sum_o)$ for $c_1,c_2\in\ZZ$.
+\begin{proof}(By Induction on $\psi^*_j\in\XX^*$).\\
 Consider $\psi_0(\chi)$.\\
@@
 c_1((1-4m_{j+1}k_{j+1})\sum\Gamma_e+(-2m_{j+1})\sum\Gamma_o)+c_2(\sum\Gamma_o+2k_{j+1}\sum\Gamma_e)=\\
-(c_1(1-4m_{j+1}k_{j+1}) + 2k_{j+1}c_2)\sum\Gamma_e+(-2m_{j+1}c_1+c_2)\sum\Gamma_o$.
+(c_1(1-4m_{j+1}k_{j+1}) + 2k_{j+1}c_2)\sum\Gamma_e+(-2m_{j+1}c_1+c_2)\sum\Gamma_o$.\\
+Therefore, $[\alpha]=\frac{1}{2}(c_1'\sum_e+c_2'\sum_o)$
 
@@
 \begin{thm}
-Let $g$ be as above. Suppose $\alpha=g(\chi)$. Then $\Omega_{u,v}([\alpha])=(0,0)$. 
-\begin{proof}
-$\Omega_{u,v}([\alpha])=\Omega_{u,v}([g(\chi)])=\Omega_{u,v}(g^{*}([\chi]))=\frac{1}{2}(i(u,g^*(\sum\Gamma)) i(v,g^*(\sum\Gamma))$.\\
-$i(u,g^*(\sum\Gamma)=c_1 i(u,\sum
-\Gamma_e)+c_2 i(u,\sum\Gamma_o)=\\ c_1(i([\gamma_5],[\gamma_{4}] - i([\gamma_{11}],[\gamma_{10}]+[\gamma_{0}]))\\+c_2(i(-[\gamma_2],[\gamma_{1}]+[\gamma_{3}]) + i([\gamma_8],[\gamma_{7}]+[\gamma_{9}]) )=\\c_1((-2)-(-2))+c_2(-2+ 2 )=0\\$
-Holds similarly for $i(v,g^*(\sum\Gamma)).$\\
+Let $\tau$ be as above. Suppose $\alpha=\tau(\chi)$. Then $\Omega_{u,v}([\alpha])=(0,0)$. 
+\begin{proof}
+$\Omega_{u,v}([\alpha])=\Omega_{u,v}([\tau(\chi)])=\Omega_{u,v}(\tau^{*}([\chi]))=\frac{1}{2}(i(u,\tau^*(\sum\Gamma)) i(v,\tau^*(\sum\Gamma))$.\\
+First consider $i(u,\tau^*(\sum\Gamma)=c_1 i(u,\sum
+\Gamma_e)+c_2 i(u,\sum\Gamma_o)$ by previous Lemma.
+$\\= c_1(i([\gamma_5],[\gamma_{4}]+[\gamma_6]) - i([\gamma_{11}],[\gamma_{10}]+[\gamma_{0}]))\\+c_2(-i([\gamma_2],[\gamma_{1}]+[\gamma_{3}]) + i([\gamma_8],[\gamma_{7}]+[\gamma_{9}]) )=\\c_1((-2)-(-2))+c_2(-2+ 2 )=0\\$
+The same holds for $i(v,\tau^*(\sum\Gamma)).$\\
 $\therefore \Omega_{u,v}([\alpha])=(0,0)$.
@@
 
+Now we know that $\chi$ serves as something of a basis for a class of lifted closed segments.
+To show that these are the only (rational direction) segments that lift to closed segments on $\btU$, we look at the orbit of segments in directions $\theta=0,$ and $\frac{\pi}{2}$ that are homologous to some $\gamma_i$.
+
 \begin{lem}
@@
 \begin{proof}
-every curve intersects nontrivially with exactly one of either $u$ or $v$.
+Every curve, $\gamma_i$, intersects exactly one adjacent index curve that is an element of exactly one of either $u$ or $v$, or with one of both $u$ \emph{and} $v$.
+Therefore, it is always taken to some non-trivial element of $\ZZ^2$ under $\Omega_{u,v}$.
 \end{proof}
@@
 \section{Proof of Main Theorem}
-
-$S=\cO\cup \cE$.
-Let $\XX'$ act on $\cS$ by linear transformations.
-
-\begin{lem}
-
-\end{lem}
+Knowing that geodesic segments in the orbit of $\chi$ and an element of $\Gamma$ are mapped either trivially or non-trivially, we use the fact that $\XX\cong\XX'$ to show that every rational direction is attainable from these elements.
+Recall that $\frac{1}{k}(x,y)\in S=\cO\cup \cE$ is a unit vector, and $(x,y)$ is an element of $\ZZ^2$ with $x,y$ relatively prime.
+Further partition $\cE$ into sets $\cE_1$ and $\cE_2$, where integer pairs $\frac{1}{k}(x,y)\in\cE$ are in $\cE_1$ if $x$ is odd or in $\cE_2$ if $x$ is even.
+These are disjoint since $x$ and $y$ cannot both be even or odd at the same time.
+Let $\XX'$ act on $\cS$ by linear transformations on $\ZZ^2$.
+We make use of Bezout's identity frequently.
+That is for any relatively prime $x,y$ there exist $m,n\in\ZZ$ such that $mx+ny=1~(*)$.
+
+\subsection{Orbit of Rational Directions}
+
+\begin{lem}
+$\XX'=\ker p$ where $p:M\mapsto M~(mod~2)$. [cite]
+\end{lem}
+
+\begin{lem}
+$\XX'$ acts faithfully and transitively on $\cO$.
+\begin{proof}
+Let $v_0=\begin{pmatrix}x \\ y\end{pmatrix}\in\cO$, where $x,y$ are both odd and relatively prime integers. Let $M=\begin{pmatrix}a & b \\ c & d
+\end{pmatrix}\in\XX'$.
+Let $u=\begin{pmatrix}1 \\ 1 \end{pmatrix}\in\cO$.
+It follows from the previous lemma that this action is faithful since $ax+by$ and $cx+dy$ cannot sum up to an element in $\cE$, as $a,d$ are odd and $b,c$ are even.
+We show that there is a matrix $M$ such that $Mu=v_0$, i.e.
+$a+b=x$ and $c+d=y$.
+We know $\det M=ad-bc=1$ and from $(*)$ that there exist $m,n\in\ZZ$ such that $1=mx+ny$.
+Let $a=(x-b)$, $b=n(xy-1)$, $c=m(xy-1)$, and $d=(y-c)$.
+Then
+\begin{align*}
+\det M&=(x-n(xy-1))(y-m(xy-1))-n(xy-1)m(xy-1)\\
+&=xy-mx(xy-1)-ny(xy-1)+mn(xy-1)^2-mn(xy-1)^2\\
+&=xy-x^2my-y^2nx+(mx+ny)=xy-xy(mx+ny)+1\\
+&=xy-xy+1=1.
+\end{align*}
+Therefore $M\in\SL(2,\ZZ)$.
+Now its simple to show that $a+b=(x-b)+b=x$ and $c+d=c+(y-c)=y$.
+Now $b=n(xy-1)$ and $c=m(xy-1)$ are even since $xy-1$ is even, and so $a=x-b$ and $d=y-c$ are odd.
+Hence, $M\in\ker p$ and $Mu=v_0$.
+$M$ is invertible, so $M^{-1}v_0=u.$ Apply the same argument for some other $v_0'\in\cO$, and obtain that $N^{-1}v_0'=u.$ We conclude then that $NM^{-1}v_0=v_0'$ for any $v_0,v_0'\in\cO$.
+\end{proof}
+\end{lem}
+
+\begin{lem}
+$\XX'$ acts faithfully and transitively on $\cE_1$ and $\cE_2$.
+\begin{proof}
+Likewise, faithfulness follows immediately from Lemma [something].
+Let $v_0=\begin{pmatrix}x \\ y\end{pmatrix}\in\cE$, where $x,y$ are both odd and relatively prime integers.
+Let $M=\begin{pmatrix}a & b \\ c & d
+\end{pmatrix}\in\XX'$. Let $u_1=\begin{pmatrix}1 \\ 0 \end{pmatrix}\in\cE_1$ and $u_1=\begin{pmatrix}0 \\ 1 \end{pmatrix}\in\cE_2$.\\\\
+Case I: $v_0\in cE_1$. Then let $a=x$, $b=-n$, $c=y$, and $d=m$, where $mx+ny=1$ by $(*)$. Then
+\begin{align*}
+\det M &= ad-bc=xm-(-ny)=mx+ny=1.
+\end{align*}
+Hence, $Mu_1=v_0$. Repeating this argument for $u_1'\in \cE_1$, we get that $Nu_1'=v_0$ and so $u_1'=N^{-1}Mu_1$.\\\\
+Case II: $v_0\in cE_1$. The proof for this is analogous to Case I.
+\end{proof}
+\end{lem}
+
+\subsection{Directions and Orbit of Geodesic}
+We can always pullback a geodesic segment on $\bM$ by associating with it a vector in $\RR^2$ because it is a translation surface.
+Let $\hol:\pi_1(\bM)\rightarrow\RR^2$ be the translational holonomy map that takes a closed path to its endpoint in local coordinates on the plane starting from the origin.
+For example, for the slope one geodesic segment $\chi$, $\hol([\chi])=(6,6)$.
+
+\begin{lem}
+Let $\alpha=\tau(\chi)$ be a geodesic segment in the orbit of $\chi$ under $\XX$. Then $\hol([\alpha])=\tau'\cdot\hol([\chi])$ with $\tau'=D(\tau)$. 
+\end{lem}
+
+\begin{lem}
+Let $\beta_h=\tau(\gamma_0)$ and $\beta_v=\tau(\gamma_{11})$ be geodesic segments in the orbit of $\gamma_0$, $\gamma_{11}$ under $\XX$.
+Then $\hol([\beta_h])=\tau'\cdot\hol([\gamma_0])$ and $ \hol([\beta_v])=\tau'\cdot\hol([\gamma_{11}])$ with $\tau'=D(\tau)$.
+\end{lem}
+
+\begin{Def}
+Let $\phi_t^\theta:\RR\rightarrow\bM$ be a unit-speed geodesic with base point $m_0$ in direction $\theta\in \RR/2\pi\ZZ$.
+Denote $\phi$'s geodesic lift onto $\btU$ $\tildphi_t^\theta:\RR\rightarrow\btU$, where $\tildphi^\theta(0)=\mtild$ is a fixed base point in the fiber over $m_0$ under the projection map.
+Let $v_\theta=(\cos\theta,\sin\theta)\in\RR^2$ be a unit-vector in direction $\theta$.
+\end{Def}
+
+\begin{thm}
+Let $\tildphi_t^\theta$ be a geodesic on $\btU$.
+Suppose there is some $k\in\RR$ such that $kv_\theta\in\cS$.
+Then $\tildphi_t^\theta$ is periodic with period $T=6\tau'\cdot (1,1)$ for some $\tau'\in\XX'$ if and only if $kv_\theta\in\cO$.
+\end{thm}
+
+\begin{thm}
+Let $\tildphi_t^\theta$ be a geodesic on $\btU$.
+Suppose there is some $k\in\RR$ such that $kv_\theta\in\cS$.
+Then $\tildphi_t^\theta$ is drift-periodic with period $T=2\tau'\cdot (1,0)$ or $2\tau'\cdot(0,1)$ for some $\tau'\in\XX'$ if and only if $kv_\theta\in\cE$.
+\end{thm}
+
+\subsection{Embedding the Necker Cube Surface in $\RR^3$}
+this section is about the map from $\RR^2$ to $\RR^3$ that embeds the cube surface. We finally show the relationship between euclidean distances in its embedding and $\Delta_{\varphi_1}$
+
+\subsection{Dynamical Classification Theorem}
+\begin{thm*}
+(Dynamical Classification of Geodesics on the Necker Cube Surface) \\Let $\Phi_t:\RR\rightarrow\bUs$ be a non-singular unit-speed geodesic on the Necker cube surface such that $\Phi(0)=u_0$ and $\Phi'(0)=v\in\RR^3$, isometric to $v_0\in\RR^2$.
+Let $\overline{u}_t\in\RR^3$ be $\Phi_t$'s coordinate in $\bUs$'s ambient space at time $t$.
+Let $k\in\RR$ such that $kv_0\in\cS$.
+Then the following is true:
+\begin{enumerate}[label=(\roman*)]
+\item $\Phi$ is periodic if $kv_0\in\cO$.
+\item Suppose $\Phi$ is periodic. If $M (kv_0)=(1,1)$ then $\Phi$ has period $T=6\sqrt{2}||kv_0||$.
+\item $\Phi$ is drift-periodic if $kv_0\in\cE$. 
+\item Suppose $\Phi$ is drift-periodic. If $M (kv_0)=(1,0)$ or $M (kv_0)=(0,1)$ then $\Phi$ has period $T=2||kv_0||$, and $d_E(\overline{u}_t,\overline{u}_{t+T})=||kv_0||\sqrt{2}$.
+\end{enumerate}
+\end{thm*}
+
+\newpage
+~
+\newpage
 
@@
 \begin{Def}
-Let $\phi_t^\theta:\RR\rightarrow\bM$ be a unit-speed geodesic on $\bM$ in direction $\theta$.
-\end{Def}
+Let $\phi_t^\theta:\RR\rightarrow\bM$ be a non-singular unit-speed geodesic on $\bM$ in direction $\theta$ based at $y_0$.
+\end{Def}
+
+Every rational direction geodesic closes so $\:\pi_1(\bM)\rightarrow\RR^2$ is well-defined for $\theta$ rational.
+Let $\alpha:[0,L]\rightarrow\bM$ be a closed geodesic segment of length L with image $\alpha([0,L])=\phi(\RR)$.
+Let $v_0=(\cos\theta,\sin\theta)$.
+
+\begin{thm}
+content...
+\end{thm}
 

### 2018-01-04 backup.tex
010318

--- 2017-12-08 newpaper.tex
@@
 
+\setlength{\topmargin}{1in} %top/bot margins
+\setlength{\oddsidemargin}{\topmargin} %sidemargins
+
+\setlength{\textheight}{11in} \setlength{\textwidth}{8.5in}
+\setlength{\hoffset}{-1in} \setlength{\voffset}{-1in} \setlength{\evensidemargin}{\oddsidemargin} \addtolength{\textheight}{-2 \topmargin}\addtolength{\textwidth}{-2\oddsidemargin}
+\setlength{\headheight}{0pt} \setlength{\headsep}{20pt} \setlength{\footskip}{20pt}
+\addtolength{\textheight}{-\footskip} \addtolength{\textheight}{-\headheight} \addtolength{\textheight}{-\headsep}
+
 \title{Periodicity of Geodesics on the Necker Cube Surface}
@@
 Each face is parallel to a $2$-dimensional subspace of $\RR^3$ spanned by exactly two basis vectors, so one component of $v$ is always 0.
-Thus $v$ is is isometric to $v_0\in\RR^2$.
+Thus $v$ is identified with $v_0\in\RR^2$.
 \\\\\\
@@
 The angle $v_0$ makes relative to a choice of basis is the \emph{initial trajectory angle} of a \emph{unit-speed geodesic} based at $u_0$ on the surface, $\Phi_t:\RR\rightarrow\bUs$.
-The initial trajectory angle is said to be $\emph{rational}$ if there is some $k\in\RR$ such that $kv_0$'s components are relatively prime integers.
+The initial trajectory angle is said to be $\emph{rational}$ if there is some $k\in\RR$ such that $kv_0$'s nonzero components are relatively prime integers.
 A rational initial trajectory angle in $\RR^2$ falls into one of two categories:
@@
 This paper aims to depict the relationship between initial trajectory angles and dynamical properties of a geodesic on the surface.
-Denote $\bUs$'s translational isometries by $\text{Trans}(\bUs)$.
-Recall that a geodesic on the surface is \emph{periodic} if there exists some $T>0$ such that $\Phi(t+T)=\Phi(t)$ for all $t\in\RR$ and a geodesic is \emph{drift-periodic} if there exists some $T'>0$ and non-trivial $p\in \text{Trans}(\bUs)$ such that $\Phi(t+T')=p(\Phi(t))$ for all $t\in\RR$.
+Recall that a geodesic on the surface is \emph{periodic} if there exists some $T>0$ such that $\Phi(t+T)=\Phi(t)$ for all $t\in\RR$ and a geodesic is \emph{drift-periodic} if there exists some $T'>0$ and non-trivial translation, $p$, such that $\Phi(t+T')=p(\Phi(t))$ for all $t\in\RR$.
 In either case we call this value the \emph{period} of $\Phi$ if it is the \emph{smallest} possible value that satisfies this definition.
 We additionally require that at no point does $\Phi$ encounter a cone singularity of the surface.
-Otherwise the geodesic behavior is indeterminable.
+Otherwise the geodesic behavior is undefined.
 We call $\Phi$ \emph{non-singular} if this holds.
@@
 \begin{thm*}
-(Dynamical Classification of Geodesics on the Necker Cube Surface) \\Let $\Phi_t:\RR\rightarrow\bUs$ be a non-singular unit-speed geodesic on the Necker cube surface such that $\Phi(0)=u_0$ and $\Phi'(0)=v\in\RR^3$, isometric to $v_0\in\RR^2$.
-Let $\overline{u}_t\in\RR^3$ be $\Phi_t$'s coordinate in $\bUs$'s ambient space at time $t$.
+(Dynamical Classification of Geodesics on the Necker Cube Surface) \\Let $\Phi_t:\RR\rightarrow\bUs$ be a non-singular unit-speed geodesic on the Necker cube surface such that $\Phi(0)=u_0$ and $\Phi'(0)=v\in\RR^3$.
+Let $v_0\RR^2$ be a distance-preserving projection of $v$ onto the plane tangent to the the surface at $u_0$.
+Let $\overline{u}_t\in\RR^3$ be $\Phi_t$'s position at time $t$.
 Let $k\in\RR$ such that $kv_0\in\cS$.
@@
 We will call this ``half-cube" $\bG$.
-It is a genus one piecewise smooth surface homeomorphic to $\TT^2$.
+It is a piecewise smooth surface homeomorphic to the two-torus, $\TT^2$.
 Every vertex has a cone angle of either $3\pi$ or $\frac{3\pi}{2}$.
@@
 We denote the surface punctured at these points by $\bG\backslash\Sigma_{\bG} = \bGs$.
-The paths labeled $a,b,c,d$ are independent of one another and span $\bGs$'s first fundamental group, $\pi_1(\bGs)\cong \<a,b,c,d \>$, the free group of four generators.
+The paths labeled $a,b,c,d$ are independent of one another and span $\bGs$'s fundamental group, $\pi_1(\bGs)\cong \<a,b,c,d \>$, the free group of four generators.
 The unpunctured Necker cube surface $\bU$ is embedded in $\RR^3$ by tiling along $\bG$'s outer edges:
@@
 \end{figure}
-From this diagram we see that $\bU$ is a universal cover of $\TT^2$.
+From this diagram we see that $\bU$ is a universal cover of $\bG$.
 In this paper we are studying geodesic behavior on $\bU$ and in order to do so, it is necessary to determine how the parallel transport of a unit vector around a cone singularity acts on the unit tangent bundle of the surface.
-But to study the path monodromy we require that these loops (see $c,d$ in Fig 3) aren't contractible on the cube's sharp corners.
+But to study a given path's rotational holonomy, we require that these loops (see $c,d$ in Fig 3) aren't contractible through the cube's sharp corners.
+To that effect we puncture the surface at these points.
 Denote the countably infinite set of singularities of $\bU$ by $\Sigma_{\bU}$.
@@
 \begin{Def}
-Denote $\varphi_1:\pi_1(\bGs)\rightarrow \ZZ^2$ as the group homomorphism such that $c,d\mapsto(0,0)$, $a\mapsto(1,0)$, and $b\mapsto(0,1)$.
-\end{Def}
-
-Take $\pi_1(\bUs)$ to be $\ker\varphi_1$.
-Then $\Delta_{\varphi_1}=\pi_1(\bGs)/\pi_1(\bUs)\cong\ZZ^2$ is the \emph{deck group} of the covering map $\varphi_1^*:\bUs\rightarrow\bGs$.
-Now $\bGs$ is unfolded and flattened in order to study monodromy as it transports a vector in $\RR^2$ along a path.
+Let $\varphi_1:\pi_1(\bGs)\rightarrow \ZZ^2$ be the group homomorphism given as $c,d\mapsto(0,0)$, $a\mapsto(1,0)$, and $b\mapsto(0,1)$.
+\end{Def}
+
+The cover of $\bGs$ with fundamental group $\ker\varphi_1$ is the space $\bUs=\tilde{\bGs}/\ker\varphi_1$ where $\ker\varphi_1$ acts on $\bGs$'s universal cover, $\tilde{\bGs}$, by deck transformations.
+We say $\Delta_{\varphi_1}=\pi_1(\bGs)/\pi_1(\bUs)\cong\ZZ^2$ is the \emph{deck group} of the covering map $\varphi_1^*:\bUs\rightarrow\bGs$.
+Now $\bGs$ is unfolded and flattened to better facilitate the study of its rotational holonomy.
 Apply the following cutting and gluing operations to the surface: \\
@@
 
-\noindent Let $\aleph_{m,n}=2m+i2n\in\CC$ be the center of a removed square. Now let $\zeta_{m,n}(t)=(2m+t)+i(2n-\frac{1}{2})$ and 
-$\xi_{m,n}(t)=(2m+t)+i(2n+\frac{1}{2})$ be paths on $\mathbf{P}$ that run along the edges of the squares, parameterized by $t\in (-\frac{1}{2},\frac{1}{2})$ and $(m,n)\in\ZZ^2$.
+\compav{Include diagram of P}
+
+\noindent Let $2m+i2n\in\CC$ be the center of a removed square with $m,n\in\ZZ$. Now let $\zeta_{m,n}(t)=(2m+t)+i(2n-\frac{1}{2})$ and 
+$\xi_{m,n}(t)=(2m+t)+i(2n+\frac{1}{2})$ be paths on $\mathbf{P}$ that run along the top and bottom edges of the squares, parameterized by $t\in (-\frac{1}{2},\frac{1}{2})$ and $(m,n)\in\ZZ^2$.
 $\bUs$ is recovered as a topological quotient $\mathbf{P}/\sim$, where $\sim$ is the minimal relation on $\mathbf{P}$:
@@
 \begin{split}
-\zeta(t)\sim e^{i\frac{\pi}{2}}~\overline{\zeta(t)-\aleph}\\
-\eta(t)\sim e^{i\frac{\pi}{2}}~\overline{\eta(t)-\aleph}
+\zeta(t)\sim e^{i\frac{\pi}{2}}~\overline{\zeta(t)-(2m+i2n)}\\
+\eta(t)\sim e^{i\frac{\pi}{2}}~\overline{\eta(t)-(2m+i2n)}
 \label{eq:rel2}
@@
 The translations, $\pm 2,\pm 2i$, on $\mathbf{P}$ are isometries that preserve these identifications and induce isometries of $\bUs$.
-The group generated by these horizontal and vertical translations form the \emph{group of Deck transformations}, $\text{Deck}(\varphi_1^*)$, that  is isomorphic to $\ZZ^2$:
+We denote the isometry induced by $+2$ as $f_h$, for horizontal, and the isometry induced by $+2i$ as $f_v$, for vertical.
+The group generated by these horizontal and vertical translations form the \emph{group of Deck transformations}, $\text{Deck}(\varphi_1^*)$, isomorphic to $\ZZ^2$:
 
@@
 
-Identify 2 with $f_h$, and $2i$ with $f_v$.
-Let $x_0\in\bGs$ and fix a point in the fiber,  $u_0\in\varphi_1^*{-1}(x_0)\subset\bUs$.
+One can see that the Let $x_0\in\bGs$ and fix a point in the fiber,  $u_0\in\varphi_1^*{-1}(x_0)\subset\bUs$.
 The action of $\ZZ^2$ on the fiber over $x_0$ is given by $(m,n)\cdot u_0=(f_h^m\circ f_v^n)(u_0)$.
+
+\compav{Discuss:}
 The distance between $u_0$ and $(f_h^m\circ f_v^n)(u_0)$ is measured in its ambient space and taken to be the magnitude of $2m+i2n\in\CC$.
@@
 \begin{lem}
-Let $[\alpha]$ be homotopy class not in the kernel of $\varphi_1$. It will lift to an unclosed path terminating on $\varphi_1([\alpha])\cdot u_0$.
+Let $\alpha:[0,1]\rightarrow \bGs$ be a loop on the surface.
+Let $[\alpha]\in\pi_1(\bGs)$ be its homotopy class representative.
+Denote its lift to $\bUs$ based at $u_0$ by $\tilde{\alpha}$.
+Then $\tilde{\alpha}(1)=\varphi([\alpha])\cdot\tilde{\alpha(0)}$.
 \end{lem}
@@
 \subsection{Monodromy Group Representation}
-A \emph{rotational monodromy map} is a morphism from the first fundamental group of a surface to the \emph{monodromy group} acting on the unit tangent bundle via parallel transport of a vector along its path.
+A \emph{rotational monodromy map} is a morphism from the first fundamental group of a surface to the \emph{monodromy group} acting on the unit tangent space via parallel transport of a vector along its path.
 Existence of a parallel transport map is a consequence of the metric connection on a surface with complex structure.
+\compav{Discuss:}\\
 The monodromy map is well defined when the \emph{holonomy group} is \emph{discrete}.
@@
 \begin{Def}
-Define $\varphi_2:\pi_1(\bGs)\rightarrow SO(2,\ZZ)$ to be the group homomorphism where $a,b \mapsto I_2$, $c  \mapsto R$, and $d  \mapsto R^3$, where $I_2$ is the identity matrix and $R$ is a positive $90^\circ$ rotation matrix.
+Define $\varphi_2:\pi_1(\bGs)\rightarrow SO(2,\ZZ)$ to be the group homomorphism where $a,b \mapsto I_2$, $c  \mapsto R$, and $d  \mapsto R^3$, where $I_2=\begin{bmatrix}1 & 0\\0 & 1\end{bmatrix}$ and $R=\begin{bmatrix}0 & -1 \\1 & 0\end{bmatrix}$.
 \end{Def}
@@
 
-\begin{lem}
+\begin{thm}
 $\varphi_2:\pi_1(\bGs)\rightarrow \mathrm{SO}(2,\ZZ)$ is the rotational monodromy representation of $\bGs$.
@@
 \end{proof}
-\end{lem}
-
+\end{thm}
+
+\compav{Discuss:}
 \begin{lem}
@@
 
-\begin{thm}
-Let $m=\varphi_2\vert_{\pi_1(\bUs)}$. Then $m$ is the rotational monodromy map of $\bUs$.
-\begin{proof}
-
-\end{proof}
-\end{thm}
-
-\begin{cor}
-Let $\alpha:[0,1]\rightarrow\bGs$ be a loop on $\bGs$ with trivial holonomy and $\tilde{\alpha}$ its lift to $\bUs$. Then $\tilde{\alpha}$ has trivial holonomy as well.
-\begin{proof}
-
-\end{proof}
-\end{cor}
-
-\begin{Def}
-Let $\bMs$ be the regular cover of $\bGs$ with fundamental group $\pi_1(\bMs)=\ker\varphi_2$.
-\end{Def}
-Showing that $\bMs$ is a degree four cover of $\bGs$ with trivial holonomy and deck group $\SO(2,\ZZ)$ follows immediately from its construction.
+A ramified cover of $\bGs$ is obtained by cyclically gluing 4 copies of neighborhoods surrounding its cone singularities.
+This cover is ramified over three singularities of cone angle $\frac{3\pi}{2}$ and $3\pi$, so 4 is the minimal number of copies needed to give every singularity in the cover a cone angle that is a multiple of $2\pi$.
+We call this cover $\bMs$.
+
 \begin{figure}[H]
@@
 \end{figure}
+
+\begin{Def}
+Let $\bMs$ be the regular cover of $\bGs$ with fundamental group $\pi_1(\bMs)=\ker\varphi_2$.
+\end{Def}
+\compav{Mark the puncutres}\\
+That $\bMs$ is a degree four cover of $\bGs$ with trivial rotational holonomy and deck group $\SO(2,\ZZ)$ follows immediately from its construction.
+The metric completion of $\bMs$ is the compact translation surface $\bM$ obtained by reinserting its cone singularities.
+A \emph{translation surface} is a topological surface with a discrete set of cone singularities and a maximal translation atlas on the surface with these singularities removed.
+Neighborhoods of the surface are naturally isometric to $\CC$, and a metric on the translation surface is obtained by pulling back the Euclidean metric in $\CC$.
 
@@
 The key connection made between surfaces $\bUs$ and $\btUs$ is that geodesics on $\bUs$ and $\btUs$ behave identically.
-Unless stated otherwise, a \emph{geodesic} is a \emph{unit-speed} locally distance minimizing curve on a smooth topological surface as a function parameterized by an element in $\RR$.
+A \emph{geodesic} is a \emph{unit-speed} locally distance minimizing curve on a smooth surface as a continuous function parameterized by an element in $\RR$.
 A \emph{geodesic segment} is a restriction of the geodesic to some closed interval in $\RR$.
-A geodesic has an initial trajectory defined at time 0, represented by a unit vector $v_0\in\RR^2$.
-If the vector can be scaled by $k\in\RR$ such that $kv_0\in\cS$, the geodesic has a \emph{rational} initial trajectory.
+A geodesic has derivative defined at time 0, represented by a unit vector $v_0\in\RR^2$.
+If the vector can be scaled by $k\in\RR$ such that $kv_0\in\cS$ (defined in the introduction), the geodesic has a \emph{rational} initial trajectory angle.
 For example, the following is a periodic geodesic on the surface with initial direction $kv_0\in\cO$:
-
+\\\compav{find this direction}
 \begin{figure}[H]
@@
 \subsection{Translation Surface}
-Consider the projection $\btU\rightarrow\mathbf{M}$ onto a quotient of the surface under its translational symmetries of each plane (see Fig 6).
-What we obtain is a genus 5 finite translation surface $\bM-$ also a branched cover of $\bG$.
-$\bMs$ was obtained as a cover of $\bGs$ from the rotational monodromy map on $\pi_1(\bGs)$, so $\bM$ is just its metric completion by reintroducing these vertices.
+The surface $\bMs$ was obtained as a cover of $\bGs$ from the rotational monodromy map on $\pi_1(\bGs)$, and $\bM$ is its metric completion as well as a branch cover of $\bGs$.
 There are four cone singularities on this surface, as opposed to the three on $\bG$, each of cone angle $6\pi$.
-Keeping track of $\btU$'s compass pointing \emph{true} north, the individual sections of $\bM$, labeled $I,II,III,$ and $IV$, are rotated in such a way to \emph{trivialize} \emph{holonomy} on arbitrary paths (Fig 8).
-Observe that $\bM$ is really a square-tiled/origami surface constructed out of unit squares (Fig 9).
+The individual sections of $\bM$, labeled $I,II,III,$ and $IV$, are rotated in such a way to trivialize holonomy on arbitrary paths (Fig 8).
+Though not immediately apparent, this surface can be realized as a finite ``staircase" surface.
+(Fig 9).
 
@@
 
-\noindent We recall some basic properties of $\ZZ^2$ covers of translation surfaces as they apply to $\bM$.
-
-The set of all orientation preserving affine diffeomorphisms of $\bM$ form the group $\text{Aff}^+(\bM)$.
+\noindent The set of all orientation preserving affine diffeomorphisms of $\bM$ form the group $\text{Aff}^+(\bM)$.
 The \emph{Veech group} of $\bM$ is the image of $\text{Aff}^+(\bM)$ in $SL(2,\RR)$ under the derivative map, $D:\text{Aff}^+(\bM)\rightarrow SL(2,\RR)$.
 We denote this group $V(\bM)$.
-It is a well-known fact that square-tiled surfaces like $\bM$ are \emph{Veech} and have Veech groups of finite index in $\SL(2,\ZZ)$ [Gutkin-Judge].
-We use $V(M)$ and the $\text{Aff}^+(\bM)$-induced automorphism group, $\text{Aut}(\pi_1(\bM))$, as they act on geodesics and their homotopy classes.
-When we reconstructed $\bM$ as a staircase, we kept track of the outer edges of the surface that were identified with one another under the projection map defined on $\btU$.
-These now form cylinder core curves that are divided into homology classes $u,v,z$ (Fig 10).
-We denote the \emph{set of all core curve homology classes} by $\Gamma=\{[\gamma_i]~:~0\leq i \leq 11\}$.
+We have two transverse cylinder decompositions of this surface in vertical and horizontal directions.
+Let $\mathfrak{H}=\{\delta_A,\delta_C,~\cdots,~\delta_K \}$ and $\mathfrak{V}=\{\epsilon_B,\epsilon_D,~\cdots,~\epsilon_{L/M} \}$ be the set of maximal cylinder decompositions of $\bM$ in the horizontal and vertical directions respectively.
+Their subscripts denote which edges the cylinder waist curves pass through.
+These cylinders are uniform in size, making $\bM$ a particularly symmetric surface.
+A convenient way to index the set of waist curves is in the following manner:
 
@@
 \centering
-\caption{$\bM$'s cylinder core curves with u,v, and z homology class labels.}
+\caption{$\bM$'s core curves with u,v, and z homology class labels.}
 \label{fig:homology}
@@
 
-We assign $u,v$ classes integer weights according to what translational effect \emph{crossing} the curve would have on a loop being lifted to the cover $\btU$.
-For example, a path crossing $[\gamma_0]$ and making a positive angle relative to the direction $\gamma_0$ is moving in (right) corresponds to moving ``up" once on $\btU$ from whatever section in the fiber of $\bM$ under the projection map on $\btU$ the lifted path might have belonged to prior to intersecting with $\gamma_0$.
-We would then assign to $[\gamma_0]$ the weight $(0,+1)\in\ZZ^2$.
-Crossing a curve in the $z$ class really does not take a path to another element in the fiber.
+These curves associate an even indexed $\gamma$ with a horizontal cylinder, and an odd indexed $\gamma$ with a vertical cylinder.
+Let $\Gamma=\{[\gamma_i]~:~0\leq i \leq 11\}$ be the set of homology classes of these curves.
+We will want to work with $\bM's$ homology group, $H_1(\bM,\QQ)$.
+
+\begin{lem}
+$\Gamma$ generates $H_1(\bM,\QQ)$.
+\begin{proof}
+One can show this is true by constructing a $12\times 12$ intersection matrix and calculating its rank.
+The genus of $\bM$ is 5, so $H_1(\bM,\QQ)\cong\QQ^{10}$.
+The rank of such a matrix is 10, which agrees with the dimension of its homology.
+\end{proof}
+\end{lem}
+
+\noindent $\Gamma$ is further divided into three classes:
+\begin{Def}
+The homology classes $u,v,z$ are given as the following weighted sums of elements of $\Gamma$:
+\begin{align*}
+u &= -[\gamma_2] +[\gamma_5] + [\gamma_8] - [\gamma_{11}],\\
+v &= +[\gamma_0] -[\gamma_3] -[\gamma_6] +[\gamma_9],\\
+z &= +[\gamma_1] +[\gamma_4]-[\gamma_7]-[\gamma_{10}].
+\end{align*}
+\end{Def}
+
+The curves composing $u$ and $v$ are linearly independent.
+By observation, these curves are weighted according to how a loop $\alpha:[0,1]\rightarrow\bM$ would lift to a path on $\btU$ by its intersections with these waist curves.
+For example, a path crossing $[\gamma_0]$ and making a positive angle relative to $\gamma_0$'s direction corresponds to moving ``up" a section on $\btU$.
+Because we consider up to be positive, $[\gamma_0]$ is signed positively, and we would give it the weight $(0,+1)\in\ZZ^2$.
 
@@
 
-By observation, these signed weights are determined as follows:
-
-\begin{Def}
-The homology classes $u,v,z$ are given as the following weighted sums of elements of $\Gamma$:
-\begin{align*}
-u &= -[\gamma_2] +[\gamma_5] + [\gamma_8] - [\gamma_{11}],\\
-v &= +[\gamma_0] -[\gamma_3] -[\gamma_6] +[\gamma_9],\\
-z &= +[\gamma_1] +[\gamma_4]-[\gamma_7]-[\gamma_{10}].
-\end{align*}
-\end{Def}
-
-\begin{rem}
-To be precise, we can think of $\btU$ as a $\ZZ^2$ cover only because we take 8 homology classes and apply the necessary transformations (change of sign) that agree with the deck group of the covering map from $\btUs$ to $\bGs$, $\ZZ^2\times\SO(2,\ZZ)$.
-Note that this group is isomorphic to $\ZZ^8$.
-\end{rem}
-
-By constructing a $12\times12$ intersection matrix and calculating its rank (rank=10), we can see that $\Gamma$ does span $H_1(\bM,\QQ)\cong\ZZ^{10}$ and does not need every cylinder curve.
-Certain relations can be discerned, such as $u+v=z$, to reduce the set to ten elements but the indexing scheme of $\Gamma$ here is far easier to work with, being in agreement with $\bM$'s cylinder decomposition in vertical and horizontal directions.
-This comes at the cost of the vector space being overdetermined and not having a coordinate basis, but it is not necessary to have one here.
-
 \begin{Def}
@@
 Let $\Omega_{u,v}:\pi_1(\bM)\rightarrow \ZZ^2$ be the homomorphism relating an $m_0$-based loop's intersection number with homology classes $u,v$ to the terminating point of its lifted path based at $\mtild$ on $\btU$.
-If $\beta$ is such a loop, then
+If $\beta$ is such a loop on $\bM$ with $[\beta]\in\pi_1(\bM)$, then
 \begin{align*}
@@
 \end{Def}
+
+This homomorphism takes $\pi_1(\bM)$ to an abelian group and as a result can be factored through $H_1(\bM,\QQ)$.
 
@@
 \subsection{Induced Automorphisms of $H_1(\bM,\QQ)$}
-We look at some important automorphisms induced by Affine maps on $\bM$.
-Observe that $\bM$ has a uniform cylinder decomposition in both horizontal and vertical directions (Fig 15).
-We define the \emph{modulus}, $\mu$, of a cylinder to be the ratio of the cylinder's width to its circumference, $\frac{w}{c}$.
-The $\emph{Dehn-twist}$ of a cylinder is an affine diffeomorphism that skews the cylinder and preserves edges and vertices.
+We have previously remarked on the uniformity of $\bM$'s cylinder decompositions, $\mathfrak{H},~\mathfrak{V}$, but have yet to show the correlation between these cylinders and $\text{Aff}^+(\bM)$.
+We say the \emph{modulus}, $\mu_i$, of the cylinder containing the $i^{th}$ waist curve to be the ratio of its width to its circumference, $\frac{w_i}{c_i}$.
+The $\emph{Dehn-twist}$ of a cylinder in either set is an affine diffeomorphism that skews the cylinder and preserves edges and vertices.
 
@@
 \label{fig:skew}
-\caption{Dehn-twist of a cylinder in $\bM$'s cylinder decomposition.}
-\end{figure}
-
-Such a map has derivative $\left[~\begin{matrix}1 && \pm\mu^{-1}\\0 && 1\end{matrix}~\right]$.
-On $\bM$ all cylinders in both vertical and horizontal decompositions have moduli $\mu=\frac{1}{2}$.
-These give up global diffeomorphisms as $\emph{multi-twists}$ of $\bM$ with the same derivative.
+\caption{Dehn-twist of a cylinder in $\bM$'s horizontal cylinder decomposition.}
+\end{figure}
+
+Such a map has derivative $\left[~\begin{matrix}1 && \pm\mu_{2i}^{-1}\\0 && 1\end{matrix}~\right]$ in local coordinates.
+On $\bM$ the cylinders in both vertical and horizontal decompositions have moduli $\mu_i=\frac{1}{2}$.
+Dehn-twists on $\mathfrak{H}$ and $\mathfrak{V}$ give up global diffeomorphisms as $\emph{multi-twists}$ of $\bM$ with the same derivatives.
 [cite]
@@
 \begin{Def}
-We define the global affine diffeomorphisms of $\text{Aff}^+(\bM)$ obtained as \textbf{multi-twists of the surface in horizontal and vertical directions} $\psi_h$ and $\psi_v$, respectively.
-\end{Def}
-
-The twists $\psi_h$ and $\psi_v$ under derivative map $D$ are mapped to the same matrices as the Dehn-twists. i.e.
+We denote the global affine diffeomorphisms of $\text{Aff}^+(\bM)$ obtained as \textbf{multi-twists} of the surface in \textbf{horizontal and vertical directions} $\psi_h$ and $\psi_v$, respectively.
+\end{Def}
+
+The images of the twists $\psi_h$ and $\psi_v$ under the derivative map $D$ are taken to the same matrices as the Dehn-twists in local coordinates. i.e.
 \begin{align*}
@@
 
-The image of this group under $D$ is itself a subgroup of $V(\bM)$ and denoted $D(\XX)=\XX'=\<\psi_h',\psi_v' \>$ as in page 2.
+The image of this group under $D$ is itself a subgroup of $V(\bM)$ and denoted $D(\XX)=\XX'=\<\psi_h',\psi_v' \>$..
 When skewing the surface and piecing it back together under $\psi_h$ or $\psi_v$, the cylinder curves intersect each other in pairs:
@@
 A similar observation can be made from applying $\psi_v$, where the vertical curves are preserved but the horizontal curves intersect with adjacent vertical curves (negative intersection).
+Let $\Aut(H_1(\bM,\QQ))$ be the group of automorphisms of $\bM$'s homology.
+By extension of these affine maps to the automorphism group, we may express the \emph{induced} automorphisms as they affect the homology generating set $\Gamma$.
+In turn, these automorphisms act linearly on a cycle $\mathfrak{c}=\sum_{i=0}^{11}k_i[\gamma_i]\in H_1(\bM,\QQ)$.
 Let $i\in\ZZ/12\ZZ$.
-Then the following formulas hold for $\psi_h$ and $\psi_v$'s counterparts in $\Aut(H_1(\bM,\QQ))$ raised to $k\in\ZZ\backslash\{0\}$:
+Then the following formulas hold for $\psi_h$ and $\psi_v$'s counterparts in $\Aut(H_1(\bM,\QQ))$ raised to $k\in\ZZ$:
 
@@
 
-If $k=0$, then it's just the identity.
 We generate the group $\XX^*=\<\psi^{*}_h,\psi^{*}_v \>$.
-There are no relations between its generators since $i$, being even or odd, has one automorphism act trivially while the other acts non-trivially.
+There are no relations between its generators since $i$, being either even or odd, has one automorphism act trivially while the other acts non-trivially.
 Therefore this group is also a free group of rank 2.
-All of our related groups are free groups, so we have that $\XX^*\cong\XX'\cong\XX$.
+All of our related groups are free groups, so we conclude $\XX^*\cong\XX'\cong\XX$.
 
 Let $\tau\in\XX$.
-Let $D(\tau)=\tau'\in\XX'$ and $\tau^*\in\XX^*$ be isomorphic elements in their respective groups.
+Let $D(\tau)=\tau'\in\XX'$ and $\tau^*\in\XX^*$ be its isomorphic counterparts belonging to their respective groups.
 We say $\tau=\prod_{j=0}^{n}\psi_h^{k_j}\circ\psi_v^{m_j}$ is a word in $\XX$ as a finite left concatenated product of multi-twist pairs with $k_j,m_j\in\ZZ$.
 Let $\psi_j=\psi_h^{k_j}\circ\psi_v^{m_j}$ be the $j^{th}$ multi-twist pair.
-Since $\tau',\tau^*$ are isomorphic, we can use their expressions interchangeably.
+Since $\tau',\tau^*$ are isomorphic, we can use these expressions interchangeably.
 \\\\
-The $j^{th}$ automorphism pair is an element $\psi_j^*\in\Aut(H_1(\bM,\QQ))$ and can be expressed as
+The corresponding $j^{th}$ automorphism pair is an element $\psi_j^*\in\Aut(H_1(\bM,\QQ))$ and is expressed as
 
@@
 These are the geodesics in directions $\theta=0,\frac{\pi}{2},$ and $\frac{\pi}{4}$.
-The aim is to show that these linear automorphisms take $\sum\Gamma\mapsto c_1\sum\Gamma_e+c_2\sum\Gamma_o$ and an individual $[\gamma_i]$ to an element that maps non-trivially to $\ZZ^2$ under $\Omega_{u,v}$.
+The aim is to show that these linear automorphisms take $\sum\Gamma\mapsto c_1\sum\Gamma_e+c_2\sum\Gamma_o$, and acts faithfully on the kernel of $\Omega_{u,v}$.
 \\
@@
 Denote its homotopy class by $[\chi]$.
-Observe from Figure 10 that $\chi$ is homologous to $\frac{1}{2}\sum\Gamma$, a curve that climbs up the staircase.
+Observe from Figure 10 that $\chi$ is homologous to $\frac{1}{2}\sum\Gamma$, a curve that climbs up the staircase by traversing each waist curve exactly halfway.
 When lifted to $\btU$ and projected onto $\bUs$, this is a closed segment on the Necker cube surface (Fig 1).
@@
 \begin{proof}
-Observe that $\Omega_{u,v}([\chi])=\Omega_{u,v}(\frac{1}{2}\sum\Gamma)=\frac{1}{2}(i(u,\sum\Gamma),i(v,\sum\Gamma)).$ Since only adjacent curves intersect, $i(u,\sum\Gamma)=i(-[\gamma_2],[\gamma_{1}]+[\gamma_{3}]) +i([\gamma_5],[\gamma_{4}]+[\gamma_{6}]) + i([\gamma_8],[\gamma_{7}]+[\gamma_{9}]) - i([\gamma_{11}],[\gamma_{10}]+[\gamma_{0}])=-2+(-2)+2-(-2)=0$.\\
+Consider $\Omega_{u,v}([\chi])=\Omega_{u,v}(\frac{1}{2}\sum\Gamma)=\frac{1}{2}(i(u,\sum\Gamma),i(v,\sum\Gamma)).$ Since only adjacent curves intersect, $i(u,\sum\Gamma)=i(-[\gamma_2],[\gamma_{1}]+[\gamma_{3}]) +i([\gamma_5],[\gamma_{4}]+[\gamma_{6}]) + i([\gamma_8],[\gamma_{7}]+[\gamma_{9}]) - i([\gamma_{11}],[\gamma_{10}]+[\gamma_{0}])=-2+(-2)+2-(-2)=0$.\\
 Similarly, $i(u,\sum\Gamma)=i([\gamma_0],[\gamma_{11}]+[\gamma_{1}]) -i([\gamma_3],[\gamma_{2}]+[\gamma_{4}]) -i([\gamma_6],[\gamma_{5}]+[\gamma_{7}]) +i([\gamma_9],[\gamma_{8}]+[\gamma_{10}])=2-(-2)-2+(-2)=0$.
@@
 
-Thus $\chi$ lifts to a closed geodesic segment on $\btU$.
-Now any geodesic segment in the orbit of $\chi$ under $\XX$ is taken into consideration, and we see why taking sums of gamma curves to sums of even and odd curves are so important.
-
-\begin{lem}
-Let $\tau$ be as above. If $\alpha=\tau(\chi)$ is a closed geodesic segment on $\bM$, then $[\alpha]=\frac{1}{2}(c_1\sum_e+c_2\sum_o)$ for $c_1,c_2\in\ZZ$.
+Thus, $\chi$ lifts to a closed geodesic segment on $\btU$.
+Now any geodesic segment in the orbit of $\chi$ under $\XX$ is taken into consideration, and we see why taking sums of gamma curves to sums of even/odd indexed curves are so important.
+
+\begin{lem}
+Let $\tau\in\XX$. If $\alpha=\tau(\chi)$ is a closed geodesic segment on $\bM$, then $[\alpha]=\frac{1}{2}(c_1\sum_e+c_2\sum_o)$ for $c_1,c_2\in\ZZ$.
 \begin{proof}(By Induction on $\psi^*_j\in\XX^*$).\\
@@
 And so $\frac{1}{2}\psi_0^*(\sum\Gamma)=\frac{1}{2}((1+2k_0-4k_0j_0)\sum\Gamma_e +(1-2m_0)\sum\Gamma_o )$.\\\\
-Induction: Suppose $\frac{1}{2}\psi^*_j(\sum\Gamma)=c_1\sum\Gamma_e+c_2\sum\Gamma_o$.\\
+Induction: Suppose $\psi^*_j(\sum\Gamma)=c_1\sum\Gamma_e+c_2\sum\Gamma_o$.\\
 Then $\psi^*_{j+1}(\psi^*_{j}(\sum\Gamma))=c_1\psi^*_{j+1}(\sum\Gamma_e)+c_2\psi^*_{j+1}(\sum\Gamma_o)=\\
@@
 $\Omega_{u,v}([\alpha])=\Omega_{u,v}([\tau(\chi)])=\Omega_{u,v}(\tau^{*}([\chi]))=\frac{1}{2}(i(u,\tau^*(\sum\Gamma)) i(v,\tau^*(\sum\Gamma))$.\\
-First consider $i(u,\tau^*(\sum\Gamma)=c_1 i(u,\sum
-\Gamma_e)+c_2 i(u,\sum\Gamma_o)$ by previous Lemma.
-$\\= c_1(i([\gamma_5],[\gamma_{4}]+[\gamma_6]) - i([\gamma_{11}],[\gamma_{10}]+[\gamma_{0}]))\\+c_2(-i([\gamma_2],[\gamma_{1}]+[\gamma_{3}]) + i([\gamma_8],[\gamma_{7}]+[\gamma_{9}]) )=\\c_1((-2)-(-2))+c_2(-2+ 2 )=0\\$
-The same holds for $i(v,\tau^*(\sum\Gamma)).$\\
+Now by the previous lemma, $i(u,\tau^*(\sum\Gamma)=c_1 i(u,\sum
+\Gamma_e)+c_2 i(u,\sum\Gamma_o)$ $\\= c_1(i([\gamma_5],[\gamma_{4}]+[\gamma_6]) - i([\gamma_{11}],[\gamma_{10}]+[\gamma_{0}]))\\+c_2(-i([\gamma_2],[\gamma_{1}]+[\gamma_{3}]) + i([\gamma_8],[\gamma_{7}]+[\gamma_{9}]) )=\\c_1((-2)-(-2))+c_2(-2+ 2 )=0\\$
+The proof that $i(v,\tau^*(\sum\Gamma))=0$ is analogous.\\
 $\therefore \Omega_{u,v}([\alpha])=(0,0)$.
@@
 Now we know that $\chi$ serves as something of a basis for a class of lifted closed segments.
-To show that these are the only (rational direction) segments that lift to closed segments on $\btU$, we look at the orbit of segments in directions $\theta=0,$ and $\frac{\pi}{2}$ that are homologous to some $\gamma_i$.
+To show that these are the only (rational direction) segments that lift to closed segments on $\btU$, we look at the orbit of segments in directions $\theta=0$ and $\frac{\pi}{2}$ that are homologous to some $\gamma_{2i}$ and $\gamma_{2i+1}$, respectively.
 
@@
 \begin{thm}
-Suppose $\beta=g(\gamma_i)$. Then $\Omega_{u,v}([\beta])\neq (0,0)$.
+Let $\sigma\in\XX$ Suppose $\beta=\sigma(\gamma_i)$. Then $\Omega_{u,v}([\beta])\neq (0,0)$.
 \begin{proof}
@@
 Knowing that geodesic segments in the orbit of $\chi$ and an element of $\Gamma$ are mapped either trivially or non-trivially, we use the fact that $\XX\cong\XX'$ to show that every rational direction is attainable from these elements.
-Recall that $\frac{1}{k}(x,y)\in S=\cO\cup \cE$ is a unit vector, and $(x,y)$ is an element of $\ZZ^2$ with $x,y$ relatively prime.
-Further partition $\cE$ into sets $\cE_1$ and $\cE_2$, where integer pairs $\frac{1}{k}(x,y)\in\cE$ are in $\cE_1$ if $x$ is odd or in $\cE_2$ if $x$ is even.
+Recall that $(x,y)\in S=\cO\cup \cE$ is an element of $\ZZ^2$ with $x,y$ relatively prime.
+Further partition $\cE$ into sets $\cE_1$ and $\cE_2$, where integer pairs $(x,y)\in\cE$ are in $\cE_1$ if $x$ is odd or in $\cE_2$ if $x$ is even.
 These are disjoint since $x$ and $y$ cannot both be even or odd at the same time.
 Let $\XX'$ act on $\cS$ by linear transformations on $\ZZ^2$.
-We make use of Bezout's identity frequently.
-That is for any relatively prime $x,y$ there exist $m,n\in\ZZ$ such that $mx+ny=1~(*)$.
+We make use of Bezout's identity$(*)$ often; for any relatively prime $x,y$ there exist $m,n\in\ZZ$ such that $mx+ny=1$.
 
@@
 Therefore $M\in\SL(2,\ZZ)$.
-Now its simple to show that $a+b=(x-b)+b=x$ and $c+d=c+(y-c)=y$.
-Now $b=n(xy-1)$ and $c=m(xy-1)$ are even since $xy-1$ is even, and so $a=x-b$ and $d=y-c$ are odd.
+Now its clear that $a+b=(x-b)+b=x$ and $c+d=c+(y-c)=y$.
+Further, $b=n(xy-1)$ and $c=m(xy-1)$ are even since $xy-1$ is even, and so $a=x-b$ and $d=y-c$ are odd.
 Hence, $M\in\ker p$ and $Mu=v_0$.
@@
 \begin{proof}
-Likewise, faithfulness follows immediately from Lemma [something].
-Let $v_0=\begin{pmatrix}x \\ y\end{pmatrix}\in\cE$, where $x,y$ are both odd and relatively prime integers.
-Let $M=\begin{pmatrix}a & b \\ c & d
-\end{pmatrix}\in\XX'$. Let $u_1=\begin{pmatrix}1 \\ 0 \end{pmatrix}\in\cE_1$ and $u_1=\begin{pmatrix}0 \\ 1 \end{pmatrix}\in\cE_2$.\\\\
-Case I: $v_0\in cE_1$. Then let $a=x$, $b=-n$, $c=y$, and $d=m$, where $mx+ny=1$ by $(*)$. Then
+Likewise, faithfulness follows immediately from [lem]. Let $v_0=\begin{pmatrix}x \\ y\end{pmatrix}\in\cE$, where $x,y$ are both odd and relatively prime integers. Let $M=\begin{pmatrix}a & b \\ c & d
+\end{pmatrix}\in\XX'$. Let $u_1=\begin{pmatrix}1 \\ 0 \end{pmatrix}\in\cE_1$ and $u_2=\begin{pmatrix}0 \\ 1 \end{pmatrix}\in\cE_2$.\\\\
+Case I: $v_0\in \cE_1$. Then let $a=x$, $b=-n$, $c=y$, and $d=m$, where $mx+ny=1$ by $(*)$. Then
 \begin{align*}
@@
 \end{align*}
-Hence, $Mu_1=v_0$. Repeating this argument for $u_1'\in \cE_1$, we get that $Nu_1'=v_0$ and so $u_1'=N^{-1}Mu_1$.\\\\
+Hence, $Mu_1=v_0$. Repeating this argument for $v_0'\in \cE_1$, we get that $Nu_1=v_0'$ and so $v_0=MN^{-1}v_0'$.\\\\
 Case II: $v_0\in cE_1$. The proof for this is analogous to Case I.
@@
 \subsection{Directions and Orbit of Geodesic}
-We can always pullback a geodesic segment on $\bM$ by associating with it a vector in $\RR^2$ because it is a translation surface.
-Let $\hol:\pi_1(\bM)\rightarrow\RR^2$ be the translational holonomy map that takes a closed path to its endpoint in local coordinates on the plane starting from the origin.
-For example, for the slope one geodesic segment $\chi$, $\hol([\chi])=(6,6)$.
-
-\begin{lem}
-Let $\alpha=\tau(\chi)$ be a geodesic segment in the orbit of $\chi$ under $\XX$. Then $\hol([\alpha])=\tau'\cdot\hol([\chi])$ with $\tau'=D(\tau)$. 
-\end{lem}
-
-\begin{lem}
-Let $\beta_h=\tau(\gamma_0)$ and $\beta_v=\tau(\gamma_{11})$ be geodesic segments in the orbit of $\gamma_0$, $\gamma_{11}$ under $\XX$.
-Then $\hol([\beta_h])=\tau'\cdot\hol([\gamma_0])$ and $ \hol([\beta_v])=\tau'\cdot\hol([\gamma_{11}])$ with $\tau'=D(\tau)$.
-\end{lem}
+Let $\hol:H_1(\bM,\QQ)\rightarrow\RR^2$ be the translational holonomy vector map that associates each cycle with a vector in $\RR^2$.
+The holonomy map preserves distances and is an accurate representation of a geodesic segment on $\bM$.
+In local coordinates an element of $V(\bM)$ factors through $\hol$.
+For example $\chi$ is a geodesic segment in direction $\theta=\frac{\pi}{4}$ with $\hol([\chi])=(6,6)$.
+After applying one multi-twist in the horizontal direction, we obtain $\hol([\psi_h(\chi)])=\hol(\psi_h^*([\chi])=\psi_h'\cdot\hol([\chi])=\psi_h'\cdot(6,6)=(18,6)=6(3,1)$.
+
+We denote the set of closed non-singular geodesics (separatrices) based at $x\in\bMs$ as $V_{x}$.
+The collection of these geodesics is the set $\cV=\bigcup_{x\in\bMs} V_{x}$.
+Let $\tau\in\XX\subseteq \text{Aff}^{+}(\bM)$.
+Then $\tau$ induces an automorphism of $\cV$ that maps $V_{x}$ to $V_{\tau(x)}$.
+In addition it is a parabolic isometry of the surface that takes a geodesic in $V_{x}$ in direction $\theta$ to a geodesic in $V_{\tau(x)}$, not necessarily in the same direction.
+$V_{x}$ forms a set of homology classes, $[V_{x}]$, such that $\hol([V_{x}])$ and $\hol([V_{\tau(x)}])$ are indistinguishable, discrete subsets of $\RR^2$.
+The faithful, transitive action of $D(\tau)=\tau'$ on $\cO$ and $\cE$ extends to $\hol([V_{x}])$.
+Because $\tau$ is isometric to $\tau^*\in\XX^*\subseteq \Aut(H_1(\bM,\QQ))$, the relationship between a choice of direction of a given geodesic, $\alpha$, based at $m_0$ and the orbit of its homology class becomes clear.
 
@@
 Let $\phi_t^\theta:\RR\rightarrow\bM$ be a unit-speed geodesic with base point $m_0$ in direction $\theta\in \RR/2\pi\ZZ$.
-Denote $\phi$'s geodesic lift onto $\btU$ $\tildphi_t^\theta:\RR\rightarrow\btU$, where $\tildphi^\theta(0)=\mtild$ is a fixed base point in the fiber over $m_0$ under the projection map.
+Denote the geodesic on $\btU$ in the same direction $\tildphi_t^\theta:\RR\rightarrow\btU$, where $\tildphi^\theta(0)=\mtild$ is a fixed base point in the fiber over $m_0$ under the projection map.
 Let $v_\theta=(\cos\theta,\sin\theta)\in\RR^2$ be a unit-vector in direction $\theta$.
-\end{Def}
+If this direction is rational, then $v_\theta$ is scaled by some $k\in\RR$ where $kv_\theta\in\cS$.
+\end{Def}
+
+We will use $\alpha,\beta$ to refer to two kinds of geodesic segments obtained from $\phi_t^\theta$.
+We will call their $\btUs$ lifts based at $\mtild$ $\tilde{\alpha}$ and $\tilde{\beta}$, and denote their homology class representatives as $[\alpha]$ and $[\beta]$.
+Similar notation will be used for the slope one, or $\frac{\pi}{4}$ direction, geodesic $\chi$ and horizontal/vertical direction geodesics $\gamma_h$, $\gamma_v$.
+
+\begin{thm}
+Let $kv_\theta\in\cO$.
+Let $\chi$ be the geodesic contained in the two cylinder decomposition obtained by flowing $\bM$ in the direction $\frac{\pi}{4}$.
+Every unit-speed, $m_0$ based, geodesic segment $\alpha:[0,T]\rightarrow\bM$ in direction $\theta$ is homologous to an element in the orbit of $\chi$ under $\XX$.
+Further, the length of $\alpha$ is given as $T=6||kv_\theta||$.
+\begin{figure}[H]
+\centering
+\includesvg[width=4.6in.]{slopeonecylinder}
+\end{figure}
+\begin{proof}
+Let $\alpha:[0,T]\rightarrow\bM$ be a closed unit-speed geodesic segment based at $m_0$ with $\alpha'=v_\theta$ and $T=||\hol([\alpha])||$.
+By transitivity of $\XX'$ acting on $\cO$, we can find some $\tau'$ such that $6kv_\theta=6\tau'\cdot(1,1)=\tau'\cdot(6,6)=\tau'\cdot\hol([\chi])$.
+$\tau'$ is isomorphic to $\tau\in\XX$, so $\tau'\cdot\hol([\chi])=\hol([\tau(\chi)])$.
+Let $\chi:[0,T']\rightarrow\bM$ be a geodesic in one of these cylinders based at $\tau^{-1}(m_0)$.
+Then $\tau(\chi)$ is a geodesic in the same direction as $\alpha$, based at $m_0$.
+$\tau$ is an isometry of $\bM$ that sends geodesics to geodesics, so $\tau(\chi)([0,T'])=\alpha([0,T])$.
+Therefore $[\tau(\chi)]=[\alpha]\in H_1(\bM,\QQ)$.\\
+Show $T=||6kv_\theta||$.
+\end{proof}
+\end{thm}
+
+\begin{cor}
+Let $\alpha$ be as before. Then $\alpha$ lifts to a closed geodesic $\tilde{\alpha}$ on $\btU$ of the same period.
+\begin{proof}
+Since $\alpha$ is homologous to $\tau(\chi)$, we have that $\Omega_{u,v}([\alpha])=\Omega_{u,v}([\tau(\chi)])=(0,0)$ by [lem].
+Therefore, $[\alpha]$ is in the kernel of $\Omega_{u,v}$ and its lift is an element of $\pi_1(\btU,\mtild)$.
+The equality of their periods follows from $\btU$ being a $\ZZ^2$ cover of a translation surface.
+\end{proof}
+\end{cor}
+
+\begin{thm}
+Let $kv_\theta\in\cE$.
+Every unit-speed, $m_0$ based geodesic segment $\beta:[0,T]\rightarrow\bM$ in direction $\theta$ is homologous to some cylinder waist curve $\gamma_i$.
+Furthermore, the length of $\beta$ is given as $T=2||kv_\theta||$.
+\begin{proof}
+Case I: $kv_\theta\in\cE_1$.
+By transitivity, there exists some $\sigma_h'\in\XX'$ such that $2kv_\theta=2\sigma_h'\cdot(1,0)=\sigma_h'\cdot(2,0)=\sigma_h'\cdot\hol([\gamma_{2i}])=\hol([\sigma_h(\gamma_{2i})])$.
+Let $\gamma_h:[0,T_h]\rightarrow\bM$ be a horizontal geodesic based at $\sigma_h^{-1}(m_0)$ contained in a horizontal cylinder in $\mathfrak{H}$ that is homologous to some $\gamma_{2i}$.
+Then $\sigma_h(\gamma_h)$ is a geodesic based at $m_0$ with $\hol([\sigma_h(\gamma_h)])=\sigma_h'\cdot\hol([\gamma_h])=\sigma'\cdot\hol([\gamma_{2i}])=2kv_\theta$.
+Therefore, $\sigma_h(\gamma_h)$ is a geodesic based at the same point and in the same direction as $\beta$, and the two are homologous.\\
+Case II: $kv_\theta\in\cE_1$.
+The proof for this is analogous with the exception that the vertical curve $\beta$ is mapped to is homologous to the waist curve of a vertical cylinder in $\mathfrak{V}$.\\
+Show $T=2||kv_0||$.
+\end{proof}
+\end{thm}
+
+\begin{cor}
+Let $\beta$ be as before. Then $\beta$ lifts to a  geodesic $\tilde{\beta}$ on $\btU$ of the same period that terminates at a point differing from $\mtild$ by some translation of $\btU$.
+\begin{proof}
+Since $\beta$ is homologous to $\sigma(\gamma_i)$ for some $\sigma\in\XX$, from [lem] we obtain that $\Omega_{u,v}([\beta])\neq(0,0)$.
+It follows that $\tilde{\beta}(T)=\Omega_{u,v}([\beta])\cdot(\tilde{\beta}(0))=\Omega_{u,v}([\beta])\cdot\mtild$ is some non-trivial $\ZZ^2$ translation of $\mtild$ that $\tilde{\beta}$ terminates at.
+The equality of the period follows from $\btU$ being a $\ZZ^2$ cover of $\bM$.
+\end{proof}
+\end{cor}
 
@@
 Suppose there is some $k\in\RR$ such that $kv_\theta\in\cS$.
-Then $\tildphi_t^\theta$ is periodic with period $T=6\tau'\cdot (1,1)$ for some $\tau'\in\XX'$ if and only if $kv_\theta\in\cO$.
+Then $\tildphi_t^\theta$ is periodic if and only if $kv_\theta\in\cO$.
+Further, $\tildphi$ has period $T=6||kv_\theta||$.
+\begin{proof}
+Suppose $\tildphi_t^\theta$ is periodic with period $T>0$.
+Let $\tilde{\alpha}=\tildphi_t^\theta\vert_{[0,T]}$ be a closed segment on $\btU$.
+Let $\alpha:[0,T]\rightarrow\bM$ be an $m_0$ based geodesic in the direction of $\theta$ with the same image as $\tilde{\alpha}$'s projection onto $\bM$.
+We've shown in [lem] that $T=6||kv_\theta||$ in both cases.
+Suppose for contradiction that $kv_\theta\notin\cO$.
+Then $kv_\theta\in\cE$.
+By [lem], this would imply $\alpha$ lifts to a geodesic that is not closed.
+If the lift is based at $\mtild=\tilde{\alpha}(0)$, then these are two identical geodesics on the surface of the same length.
+This would mean that $\tilde{\alpha}(T)\neq \mtild$, which is a contradiction.
+Therefore, $kv_\theta\in\cO$.\\
+Conversely suppose $kv_\theta\in\cO$.
+Then it follows from [lem] that a geodesic $\alpha$ on $\bM$ lifts to a closed geodesic on $\btU$ of the length $T=6||kv_\theta||$, a restriction of $\tildphi$ to $[0,T]$.
+It follows that $T$ must be the period of $\tildphi$ since it is the smallest possible value such that $\alpha$'s lift closes up.
+\end{proof}
 \end{thm}
-
+~\\\\\\\\\\\\\\\\
 \begin{thm}
@@
 Suppose there is some $k\in\RR$ such that $kv_\theta\in\cS$.
-Then $\tildphi_t^\theta$ is drift-periodic with period $T=2\tau'\cdot (1,0)$ or $2\tau'\cdot(0,1)$ for some $\tau'\in\XX'$ if and only if $kv_\theta\in\cE$.
+Then $\tildphi_t^\theta$ is drift-periodic if and only if $kv_\theta\in\cE$.
+Further, $\tildphi_t^\theta$ has period $T=2||kv_\theta||$.
+\begin{proof}
+
+\end{proof}
+\end{thm}
+
+\newpage
+
+\begin{thm}
+Let $\tildphi_t^\theta$ be a geodesic on $\btU$.
+Suppose there is some $k\in\RR$ such that $kv_\theta\in\cS$.
+Then $\tildphi_t^\theta$ is periodic if and only if $kv_\theta\in\cO$.
+Further, $\tildphi$ has period $T=||6kv_\theta||$.
+\begin{proof}
+If $\tildphi_t^\theta$ is periodic, there is some $T>0$ such that $\tilde{\alpha}_t=\tildphi_t^\theta\vert_{[0,T]}$ is a closed path on $\btU$.
+As a translation cover, $\tilde{\alpha}$ is projected onto $\alpha\subset\bM$ of the same period.
+Therefore, $\tilde{\alpha}$ is a lifted geodesic segment of $\alpha$.
+If $kv_\theta\notin\cO$, then $kv_\theta\in\cE$.
+Then by transitivity of $\XX'$ there exists some $\sigma'\in\XX'$ such that $\sigma'\cdot\hol([\alpha])=\hol(\sigma^*([\alpha]))=\hol([\gamma_i])=(0,2),~(2,0)$.
+This would imply $\sigma^{*-1}([\gamma_i])\in\ker\Omega_{u,v}$, which is a contradiction.
+Therefore $kv_\theta\in\cO$.
+\\ Conversely, if $kv_\theta\in\cO$ there is some $\tau'\in\XX'$ such that $6\tau'\cdot(1,1)=6kv_\theta$.
+Consider the geodesic $\xi_t^{\frac{\pi}{4}}:\RR\rightarrow\bM$ based at $\tau^{-1}(m_0)$.
+Let $\chi=\xi_t^{\frac{\pi}{4}}\vert_{[0,T']}$ be a closed geodesic segment on the surface where $T'$ is the period of $\xi_t^{\frac{\pi}{4}}$.
+But this segment is contained one of two cylinders obtained by flowing $\bM$ in the direction $\frac{\pi}{4}$:
+$\chi$ is homologous to $\frac{1}{2}\sum\Gamma$, so $\hol([\chi])=(6,6)$ and $T'=6\sqrt{2}$.
+Hence $T=6kv_\theta=\tau'\cdot\hol([\chi])=\hol(\tau^*([\chi]))=\hol([\tau(\chi)])$, and $\tau(\chi):[0,T']\rightarrow\bM$ is a geodesic segment in the direction of $\theta$.
+This segment is re-parametrized as a unit-speed segment $\alpha:[0,T]\rightarrow\bM$ where $T$ is its total length.
+$\alpha(0)=\tau(\chi(0))=\tau(\tau^{-1}(m_0))=m_0$, so $\alpha$ is based at $m_0$ and $\alpha'=v_\theta$.
+Since $[\alpha]=[\tau(\chi)]=\tau^*([\chi])\in \pi_1(\bM)$, $[\alpha]\in\ker\Omega_{u,v}$ and lifts to a closed geodesic segment of the same length,$\tilde{\alpha}$.
+$\tilde{\alpha}:[0,T]\rightarrow\btU$ based at $\tilde{m}_0$ and is just a restriction of $\tildphi_t^\theta$ to the interval $[0,T]$.
+Therefore $\tildphi$ is periodic with period $T=||6kv_\theta||$.
+\end{proof}
+
+\end{thm}
+
+\begin{thm}
+Let $\tildphi_t^\theta$ be a geodesic on $\btU$.
+Suppose there is some $k\in\RR$ such that $kv_\theta\in\cS$.
+Then $\tildphi_t^\theta$ is drift-periodic if and only if $kv_\theta\in\cE$.
+Further, $\tildphi_t^\theta$ has period $T=||2kv_\theta||$.
+\begin{proof}
+Suppose $kv_\theta\notin\cE$.
+Then $kv_\theta\in\cO$.
+Suppose $\tildphi_t^\theta$ is drift-periodic with period $T>0$.
+Let $\tilde{\beta}_t=\tildphi_t^\theta\vert_{[0,T]}$.
+Let $\beta:[0,T]\rightarrow\bM$ be the projected geodesic segment in the same direction based at $m_0$.
+Let $\chi:[0,T']\rightarrow\bM$ be a unit-speed geodesic based at $\tau^{-1}(m_0)$ in direction $\frac{\pi}{4}$.
+By [lem] there exists some $\tau'\in\XX'$ such that $\hol([\beta])=\hol([\tau(\chi)])=\tau'\cdot\hol([\chi])=6\tau'\cdot(1,1)=6kv_\theta$.
+By [lem] this implies $\Omega_{u,v}([\tau(\chi)])=(0,0)$, which is a contradiction.
+Therefore, $kv_\theta\in\cE$.\\
+Conversely, suppose $kv_\theta\in\cE$.\\
+Case I: $kv_\theta\in\cE_1$.
+Then there exists some $\sigma'$ such that $\hol([\beta])=\hol([\sigma(\gamma_{2i})])=\hol(\sigma^*([\gamma_{2i}]))=\sigma'\cdot\hol([\gamma_{2i}])=2\sigma'\cdot(1,0)=2kv_\theta$.
+\end{proof}
 \end{thm}
@@
 The slope one direction also decomposes $\mathbf{M}$ into two cylinders by a series of saddle connections of length $\sqrt{2}$ between singularities:
-
-\begin{figure}[H]
-\centering
-\includesvg[width=4.4in.]{slopeonecylinder}
-\caption{The two right-most cylinders $C_1$ (labeled) and $C_2$(unlabeled).}
-\end{figure}
 

### 2018-01-04 finalpaper.tex
010318

--- 2017-12-04 newpaper.tex
@@
-\documentclass[]{article}
+\documentclass[a4paper, 11pt]{article}
+\usepackage{comment} % enables the us`e of multi-line comments (\ifx \fi) 
+\usepackage{lipsum} %This package just generates Lorem Ipsum filler text. 
+\usepackage{fullpage} % changes the margin
 \usepackage{amsfonts}
@@
 
-\title{Periodicity of Geodesics on the Necker Cube Surface}
+\setlength{\topmargin}{1in} %top/bot margins
+\setlength{\oddsidemargin}{\topmargin} %sidemargins
+
+\setlength{\textheight}{11in} \setlength{\textwidth}{8.5in}
+\setlength{\hoffset}{-1in} \setlength{\voffset}{-1in} \setlength{\evensidemargin}{\oddsidemargin} \addtolength{\textheight}{-2 \topmargin}\addtolength{\textwidth}{-2\oddsidemargin}
+\setlength{\headheight}{0pt} \setlength{\headsep}{20pt} \setlength{\footskip}{20pt}
+\addtolength{\textheight}{-\footskip} \addtolength{\textheight}{-\headheight} \addtolength{\textheight}{-\headsep}
+
+\title{On the Characterizations of Rational Geodesics on the Infinite Necker Cube Surface}
 \author{Pavel Javornik}
-
+\date{}
 \maketitle
@@
 %\end{center}
-
 \begin{abstract}
-\noindent Although it is most recognizable as M.C.
-Escher's adaptation of an optical illusion popularized by Louis Albert Necker, the Necker cube surface has a complex structure and is, in analaytic language, described as an infinite, flat (locally isometric to $\RR^2$ on smooth sections), $\RR^3$-embeddable Euclidean cone surface with countably many singularities of cone angles $3\pi$ and $\frac{3\pi}{2}$.
-Its geometric construction is described as the gluing of infinitely many unit cubes along their edges so that every face shares an edge with exactly four others.
-A unit-speed geodesic on this surface has an initial trajectory angle $\theta\in\RR/2\pi\ZZ$ given up by the surface's local plane isometry.
-Given a rational vector direction in $\RR^2$ (components are relatively prime integers) obtained from $\theta$, one can classify every periodic/drift-periodic rational geodesic on the surface.
-In addition its total period/length and, in the case where it is drift-periodic, translational drift as Euclidean distance in its ambient space is given by solving a system of simple linear Diophantine equations.
+\noindent The deterministic quality of the dynamical characteristics, up to periodicity and drift periodicity, of a non-singular, rational geodesic on the infinite Euclidean cone surface obtained from the spatial edge-to-edge arrangement of infinitely many unit cubes comes down to the choice of an initial trajectory angle as a $\ZZ^2$ vector contained in the plane tangent to the surface at the geodesics initial point.
+Moreover these integer components give up the total period of a unit-speed geodesic, and their parities ultimately dictate whether or not it will be periodic or drift periodic.
+Using the surface and its translation cover's various symmetries, we prove these results and show that, when applicable, the Euclidean drift distance per period is given up by an element in the solution space of a system of $\ZZ$-linear equations.
 \end{abstract}
+\section{Introduction}
+The Necker cube has been studied extensively due to its interesting effects on human visual perception.
+A popular adaptation of the Necker cube is the edge-to-edge pairing of multiple solid cubes giving the illusion of a rhombile tiling of the plane when viewed from a fixed angle.
+(Fig \ref{fig:front}) It is most recognizable in the works of M.C.
+Escher \cite{Escher1}\cite{Escher2} and as the game board in the 80s cabinet arcade classic, $\text{Q}^*\text{bert}$.\cite{Qbert} The cube was named after the Swiss crystallographer Louis A.
+Necker for having been one of the first to study the wire-frame cube.\cite{Albert}
 \begin{figure}[H]
 \centering
-\includesvg[width=4in]{cubesxyzpaths2bwv2b}
+\includesvg[width=5.5in]{oddoddtrajectory}
+\caption{Periodic Geodesic on the Necker Cube Surface}
 \label{fig:front}
-\caption{Periodic and drift-periodic flows on the Necker cube surface}
 \end{figure}
-
-\newpage
-\section{Introduction}
-
-Put your Problem statement here! Example of a Citation\cite[p.219]{Robotics}. Here's Another Citation \cite{Flueck}
-
-The Necker cube has made numerous appearances in the work of mathematicians, crystallographers, and scientists interested in human visual systems before and after appearing in the works of M.C.
-Escher [cite].
-Louis Albert Necker was credited for having discovered the optical illusion, and studying its geometric properties [cite].
-The solid presentation of the cube, when rendered as a flat surface with three rear faces removed, achieves a similar effect when its three visible faces are shaded in a particular way.
-The Necker cube surface is then obtained by gluing infinitely many copies along its outer edges as in Figure 1.
-
-Dynamical classification of geodesics on flat, periodic surfaces such as the Necker cube surface ultimately boils down to  studying the surface's various symmetries and how the rotational holonomy group acts on the unit tangent bundle of the surface as the unit vector (derivative of the geodesic) is parallel transported along its path.
-Loops on polyhedra punctured at its cone singularities tend to have discrete holonomy groups as its relative holonomy (holonomy of contractible loops) is trivial.
-
-\begin{rem}
-Familiarity is assumed on the part of the reader with covering space theory, translation (Veech) surfaces, and their associated Veech groups. For general surveys on these topics: [cite], [cite].
-\end{rem}
-
-\subsection{Discussion of Results}
-Our initial experiments strongly supported the theory that there would be a correlation between a choice of trajectory angle and dynamical properties of a geodesic on such a symmetric object.
-Rightfully so, a surface composed of infinitely many cubes an induced flat metric where every neighborhood is locally isometric to the Euclidean plane.
-Via the parallel transport of a unit tangent vector over the sharp edges of the surface denoted $\bUs$, it becomes necessary to think of the geodesic as a sequence of line segments contained in the faces of each cube on the surface's embedded form in $\RR^3$.
-$\ref{fig:front}$.
-\\
-\begin{wrapfigure}{l}{2in.}
-\includesvg{vectorplane}
-\end{wrapfigure}
-\noindent Let $u_0$ be a point on the surface that is contained in the interior of a face and consider a tangent unit vector $v\in\RR^3$ based at point $u_0$.
-Each face is parallel to a $2$-dimensional subspace of $\RR^3$ spanned by exactly two basis vectors, so one component of $v$ is 0.
-Call this projected vector $v_0\in\RR^2$.
-\\\\\\
-The angle $v_0$ makes relative to a choice of basis is the \emph{initial trajectory angle} of the unit-speed geodesic on the surface, $\Phi_t:\RR\rightarrow\bUs$.
-A rational initial trajectory angle in $\RR^2$ falls into one of two categories:
-\begin{Def}
-Let $v_0$ be a unit vector of the form $\frac{1}{k}(x,y)\in\RR^2$ with $x,y\in\mathbb{Z}$ and $k=\sqrt{x^2+y^2}\in\mathbb{R}$.
-We say $v_0$ is an \textbf{odd-odd} vector if its components are relatively prime and both odd.
-We denote the \textbf{set of all odd-odd directions} by $\mathcal{O}$.
-We say that $v_0$ is an \textbf{even-odd} vector if its components are relatively prime and of opposite parity.
-We denote the \textbf{set of all even-odd directions} by $\mathcal{E}$.
-\end{Def}
-\noindent For convenience we call the disjoint union of these sets $\cS$.
-This paper aims to depict the relationship between initial trajectory angles and dynamical properties of a geodesic on the surface.
-Denote $\bUs$'s translational isometries by $\text{Trans}(\bUs)$.
-Recall that a geodesic on the surface is \emph{periodic} if there exists some $T>0$ such that $\Phi(t+T)=\Phi(t)$ for all $t\in\RR$ and a geodesic is \emph{drift-periodic} if there exists some $T'>0$ and non-trivial $p\in \text{Trans}(\bUs)$ such that $\Phi(t+T')=p(\Phi(t))$ for all $t\in\RR$.
-In either case we call this value the \emph{period} of $\Phi$ if it is the \emph{smallest} possible value that satisfies our definition.
-We additionally require that at no point does $\Phi$ encounter a cone singularity of the surface, in which case the geodesic behavior is non-deterministic.
-
-\begin{align}
-\text{Let }\XX'=\left\< \left[ \hspace{1mm} \begin{matrix}
-							1 & 2\\
-							 0 & 1
-						\end{matrix}\hspace{1mm}\right],~
-\left[ \hspace{1mm} \begin{matrix}
-							1 & 0\\
-							 2 & 1
-						\end{matrix}\hspace{1mm}\right]\right\>
-\end{align}
-
-\noindent Section [blank] explains the significance of this group. Let $M=\left[ \hspace{1mm} \begin{matrix}
-							s & u\\
-							 w & z
-						\end{matrix}\hspace{1mm}\right]\in\XX'$.
-
-\begin{thm*}
-(Dynamical Properties of a Geodesic on $\bUs$) \\Let $\Phi_t:\RR\rightarrow\bUs$ be a non-singular unit-speed geodesic on the Necker cube surface such that $\Phi(0)=u_0$ and $\Phi'(0)=v\in\RR^3$, isometric to $v_0\in\RR^2$.
-Let $\overline{u}_t\in\RR^3$ be $\Phi_t$'s coordinate in $\bUs$'s ambient space at time $t$.
-Then the following is true:
-\begin{enumerate}[label=(\roman*)]
-\item $\Phi$ is periodic if $v_0\in\cO$.
-\item Suppose $(i)$ is true. If $M (kv_0)=(1,1)$ then $\Phi$ has period $T=6\sqrt{2}||kv_0||$.
-\item $\Phi$ is drift-periodic if $v_0\in\cE$. 
-\item Suppose $(iii)$ is true. If $M (kv_0)=(1,0)$ or $M (kv_0)=(0,1)$ then $\Phi$ has period $T=2||kv_0||$, and $d_E(\overline{u}_t,\overline{u}_{t+T})=||kv_0||\sqrt{2}$.
-\end{enumerate}
-\end{thm*}
-\compav{Remember to prove that such an M exists.}
-
-Here $||~||$ is the standard Euclidean norm in $\RR^2$, and $d_E:\RR^3\times\RR^3\rightarrow\RR$ is the Euclidean distance.
-The proof for this theorem can be found in XXX and follows from Theorems YYYY and ZZZZZ.
-
-\subsection{Acknowledgements}
--Pat Hooper\\
-- Vincent Delecroix, Ferrán Valdez, pascal hubert\\
-- \\
-
-\newpage
-
-\section{Periodic Tiling of Necker Cubes}
-This section will detail how the Necker Cube surface is constructed as a covering space of a thrice-punctured torus. Make the following identifications on 3 unit squares:
+We take the three visible faces of a single cube viewed from this angle and make identifications on the parallel edges by translation.
+(Fig \ref{fig:halfcube}) These are the outer edges that are glued to the opposite edges of other cubes.
+We call this half-cube $\bG$.
+$\bG$ has a set of three cone singularities, $\Sigma_\bG$, two with cone angle $3\pi$ and one with cone angle $\frac{3\pi}{2}$.
+The half-cube punctured at these points is $\bGs=\bG\backslash\Sigma_\bG$, and we call this our \emph{Necker cube}.
 \begin{figure}[H]
@@
 \includesvg[width=1.83in]{neckercube}\includesvg[width=2in.]{neckercubepath}
-\caption{$\bGs$ with and without paths $a,b,c,d$.}
+\caption{The surface $\bGs$ with and without loops $a,b,c,d$ based at $x_0$.}
+\label{fig:halfcube}
 \end{figure}
-
-We will call this ``half-cube" $\bG$.
-It is a genus one piecewise smooth surface homeomorphic to $\TT^2$.
-Every vertex has a cone angle of either $3\pi$ or $\frac{3\pi}{2}$.
-Let $\Sigma_{\bG}\subset\bG$ be the set of these three vertices.
-These are conical singularities of the surface.
-We denote the surface punctured at these points by $\bG\backslash\Sigma_{\bG} = \bGs$.
-The paths labeled $a,b,c,d$ are independent of one another and span $\bGs$'s first fundamental group, $\pi_1(\bGs)\cong \<a,b,c,d \>$, the free group of four generators.
-The unpunctured Necker cube surface $\bU$ is embedded in $\RR^3$ by tiling along $\bG$'s outer edges:
+The set of paths labeled above generate the fundamental group of the surface, $\pi_1(\bGs,x_0)$, that is isomorphic to the free group of four generators since $\bG$ is homeomorphic to the genus 1 torus.
+The \emph{Necker cube surface} itself is a cover of $\bGs$ with a deck group isomorphic to $\ZZ^2$:
 
@@
 \centering
-\includesvg[width=4in]{neckercubepathsurface}
-\end{figure}
-From this diagram we see that $\bU$ is a universal cover of $\TT^2$.
-In this paper we are studying geodesic behavior on $\bU$ and in order to do so, it is necessary to determine how the parallel transport of a unit vector around a cone singularity acts on the unit tangent bundle of the surface.
-But to study the path monodromy we require that these loops (see $c,d$ in Fig 3) aren't contractible on the cube's sharp corners.
-Denote the countably infinite set of singularities of $\bU$ by $\Sigma_{\bU}$.
-The \emph{Necker cube surface} is the punctured surface $\bUs=\bU\backslash\Sigma_{\bU}$.
-Observe from the previous diagram that $\pi_1(\bUs)$ is the kernel of the following group homomorphism:
-
-\begin{Def}
-Denote $\varphi_1:\pi_1(\bGs)\rightarrow \ZZ^2$ as the group homomorphism such that $c,d\mapsto(0,0)$, $a\mapsto(1,0)$, and $b\mapsto(0,1)$.
-\end{Def}
-
-Take $\pi_1(\bUs)$ to be $\ker\varphi_1$.
-Then $\Delta_{\varphi_1}=\pi_1(\bGs)/\pi_1(\bUs)\cong\ZZ^2$ is the \emph{deck group} of the covering map $\varphi_1^*:\bUs\rightarrow\bGs$.
-Now $\bGs$ is unfolded and flattened in order to study monodromy as it transports a vector in $\RR^2$ along a path.
-Apply the following cutting and gluing operations to the surface: \\
-\begin{figure}[H]
-\centering
-\includesvg[width=3in]{LTorus}
-\end{figure}
-\noindent This L-shaped torus is then taken apart and recovered as $2\times 2$ torus missing a unit square with the following identifications:\\
-\begin{figure}[H]
-\centering
-\includesvg[width=2in]{BranchTorus}
-\caption{$\bGs$ and images of paths after reconstruction. The outermost vertex is \emph{not} a puncture.}
-\end{figure}
-
-Consider a $2\times2$ sheet missing an open unit square and that square's vertices (punctures).
-Center this structure around $0\in\CC$ and recover $\bGs$ with the identifications made in Fig 3.
-Let $m,n\in\ZZ$.
-Similar identifications can be made on the following subset of $\CC$ to recover $\bUs$:
-
-\begin{equation}
-\label{eq:P}
-\mathbf{P} = \mathbb{C}\text{ }\backslash\bigcup_{m,n \in {\mathbb Z}} \big{\{} u+vi:~|u-2m|<\frac{1}{2},~|v-2n|<\frac{1}{2},~u=2m\pm\frac{1}{2}~v=2n\pm\frac{1}{2} \big{\}}.
-\end{equation}
-
-\noindent Let $\aleph_{m,n}=2m+i2n\in\CC$ be the center of a removed square. Now let $\zeta_{m,n}(t)=(2m+t)+i(2n-\frac{1}{2})$ and 
-$\xi_{m,n}(t)=(2m+t)+i(2n+\frac{1}{2})$ be paths on $\mathbf{P}$ that run along the edges of the squares, parameterized by $t\in (-\frac{1}{2},\frac{1}{2})$ and $(m,n)\in\ZZ^2$.
-$\bUs$ is recovered as a topological quotient $\mathbf{P}/\sim$, where $\sim$ is the minimal relation on $\mathbf{P}$:
-
-\begin{equation}
-\begin{split}
-\zeta(t)\sim e^{i\frac{\pi}{2}}~\overline{\zeta(t)-\aleph}\\
-\eta(t)\sim e^{i\frac{\pi}{2}}~\overline{\eta(t)-\aleph}
-\label{eq:rel2}
-\end{split}
-\end{equation}
-The translations, $\pm 2,\pm 2i$, on $\mathbf{P}$ are isometries that preserve these identifications and induce isometries of $\bUs$.
-The group generated by these horizontal and vertical translations form the \emph{group of Deck transformations}, $\text{Deck}(\varphi_1^*)$, that  is isomorphic to $\ZZ^2$:
-
-\begin{figure}[H]
-\centering
-\includesvg[width=2.3in]{tiling}
-\end{figure}
-
-Identify 2 with $f_h$, and $2i$ with $f_v$.
-Let $x_0\in\bGs$ and fix a point in the fiber,  $u_0\in\varphi_1^*{-1}(x_0)\subset\bUs$.
-The action of $\ZZ^2$ on the fiber over $x_0$ is given by $(m,n)\cdot u_0=(f_h^m\circ f_v^n)(u_0)$.
-The distance between $u_0$ and $(f_h^m\circ f_v^n)(u_0)$ is measured in its ambient space and taken to be the magnitude of $2m+i2n\in\CC$.
-
-\begin{lem}
-Let $[\alpha]$ be homotopy class not in the kernel of $\varphi_1$. It will lift to an unclosed path terminating on $\varphi_1([\alpha])\cdot u_0$.
-\end{lem}
-
-\subsection{Monodromy Group Representation}
-A \emph{rotational monodromy map} is a morphism from the first fundamental group of a surface to the \emph{monodromy group} acting on the unit tangent bundle via parallel transport of a vector along its path.
-Existence of a parallel transport map is a consequence of the metric connection on a surface with complex structure.
-The monodromy map is well defined when the \emph{holonomy group} is \emph{discrete}.
-When the \emph{relative holonomy} of contractible loops is trivial, the holonomy group is the set of attainable rotations of a unit vector upon returning to the base point after transporting it along a non-trivial loop.
-A \emph{flat surface} like $\bGs$ has trivial relative holonomy as its neighborhoods are isometric to the plane.
-Denote the \emph{unit tangent bundles} of $\bUs$ and $\bGs$ by $T^1\bUs$ and $T^1\bGs$.
-If $\alpha$ is a loop on $\bGs$, then $[\alpha]$ acts on a fixed vector $v_0\in\RR^2$ by rotation.
-In the case where $\alpha$ is \emph{not} a geodesic segment on the surface of length $L$, $v_0=\alpha'(0)\in\RR^2$ may differ from $\alpha'(L)$ by some rotation.
-
-\begin{Def}
-Define $\varphi_2:\pi_1(\bGs)\rightarrow SO(2,\ZZ)$ to be the group homomorphism where $a,b \mapsto I_2$, $c  \mapsto R$, and $d  \mapsto R^3$, where $I_2$ is the identity matrix and $R$ is a positive $90^\circ$ rotation matrix.
-\end{Def}
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=2.5in]{monodromy.png}
-\caption{Effect a nontrivial loop has on basis vectors $(i,j)\in\RR^2$.}
-\label{fig:loop}
-\end{figure}
-
-\begin{lem}
-$\varphi_2:\pi_1(\bGs)\rightarrow \mathrm{SO}(2,\ZZ)$ is the rotational monodromy representation of $\bGs$.
-\begin{proof}
-From Figure 4 it's clear that $c,d$ will rotate a unit vector by $\pm 90^\circ$ by parallel transport along their paths.
-These values are determined by the cone angles of the singularities they loop around.
-The parallel transport of a vector along $a$ or $b$ would be trivial since the outer edges they pass through are glued by translation.
-\end{proof}
-\end{lem}
-
-\begin{lem}
-Any element in $\text{Deck}(\varphi_1^*)$ acting on $T^1\bUs$ has a trivial effect on holonomy.
-\begin{proof}
-Since $\bGs=\bUs/\ZZ^2$, the deck group action lifts naturally to an action on their tangent vector spaces and unit tangent bundles via pushforward of deck transformations $f_h,f_v$ on the bundle maps.
-i.e.
-$T^1\bGs=T^1\bUs/\ZZ^2$.
-This effect must be trivial since $\varphi_1(a)\sim f_h$ and $\varphi_1(b)\sim f_v$, and both homotopy classes $a,b$ have trivial holonomy.
-\\
-How do $f_h,f_v$ parallel transport a vector?\\
-Well if $\varphi_1([a])\cdot u_0=f_h(u_0)$, then for 
-\end{proof}
-\end{lem}
-
-\begin{thm}
-Let $m=\varphi_2\vert_{\pi_1(\bUs)}$. Then $m$ is the rotational monodromy map of $\bUs$.
-\begin{proof}
-
-\end{proof}
-\end{thm}
-
-\begin{cor}
-Let $\alpha:[0,1]\rightarrow\bGs$ be a loop on $\bGs$ with trivial holonomy and $\tilde{\alpha}$ its lift to $\bUs$. Then $\tilde{\alpha}$ has trivial holonomy as well.
-\begin{proof}
-
-\end{proof}
-\end{cor}
-
-\begin{Def}
-Let $\bMs$ be the regular cover of $\bGs$ with fundamental group $\pi_1(\bMs)=\ker\varphi_2$.
-\end{Def}
-Showing that $\bMs$ is a degree four cover of $\bGs$ with trivial holonomy and deck group $\SO(2,\ZZ)$ follows immediately from its construction.
-\begin{figure}[H]
-\centering
-\includegraphics[width=3in]{monogroup.png}
-\label{fig:arbitrarylift}
-\caption{Lift of paths $c,d$ to $\bMs$ (\compav{RELABEL EDGES})}
-\end{figure}
-
-\subsection{A Four-Fold Cover}
-We now take these covers we built to construct a third one that covers them all. This is a four-fold cover of $\bUs$ and a $\ZZ^2$ translation cover of $\bMs$. 
-\begin{Def}
-Let $\btUs$ be the cover of $\bUs$ with fundamental group $\pi_1(\btUs)=\ker\varphi_2\big\vert_{\pi_1(\bUs)}$.
-\end{Def}
-
-It follows that $\ker\varphi_1\big\vert_{\pi_1(\bMs)}=\ker\varphi_2\big\vert_{\pi_1(\bUs)}$ and that $\btUs$ is a cover of these two surfaces.
-Edge identifications are made on four copies of $\mathbf{P}$, possibly indexed by the finite component of the direct product $\mathbf{P}\times\ZZ/4\ZZ$.
-The Deck group of the covering map from $\btUs$ to $\bGs$ is the set $\ZZ^2\times \SO(2,\ZZ)$.
-Consider a loop on $\bUs$:
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=2in]{overlay.png}
-\end{figure}
-
-\noindent If we lift this curve to $\btUs$ we can identify it with an element of the deck group, and check if its lift from $\bGs$ belongs to the kernel of both $\varphi_1$ and $\varphi_2$.
-Fix a direction on this surface pointing north that respects $\bUs$'s compass and rotate every plane by either $90^\circ,$ $180^\circ,$ or $270^\circ$ as the curve travels from one plane (copy of $\mathbf{P}$) to the next.
-Observe that it will not only close if the curve acts trivially on the holonomy of $\bUs$, but with this construction we can see that $\btUs$ is actually an infinite-type translation surface:
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=3in]{coverdirection.png}
-\caption{With $P\in 2\ZZ+2\ZZ i$.}
-\end{figure}
-
-\noindent
-
-\subsection{Geodesics on $\bUs$}
-The key connection made between surfaces $\bUs$ and $\btUs$ is that geodesics on $\bUs$ and $\btUs$ behave identically.
-Unless stated otherwise, a \emph{geodesic} is a \emph{unit-speed} locally distance minimizing curve on a smooth topological surface as a function parameterized by an element in $\RR$.
-A \emph{geodesic segment} is a restriction of the geodesic to some closed interval in $\RR$.
-A geodesic has an initial trajectory defined at time 0, represented by a unit vector $v_0\in\RR^2$.
-If the vector can be scaled by $k\in\RR$ such that $kv_0\in\cS$, the geodesic has a \emph{rational} initial trajectory.
-For example, the following is a periodic geodesic on the surface with initial direction $kv_0\in\cO$:
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=4in.]{closed2.png}
-\caption{Periodic geodesic path on $\bUs$ modeled with sage-flatsurf.}
-\label{fig:complicated}
-\end{figure}
-
-We refer later to the following commuting diagram of branched covering maps and inclusion maps. It is meant to illustrate the relationship between all of these spaces.
-
-\begin{equation}
-\begin{tikzcd}
-\arrow[two heads, bend left]{rrr}{f^\circ}[swap]{} \arrow[hook]{d}{pr_b} \btUs \arrow[two heads]{r}{r^\circ} & \bUs \arrow[hook]{d}{pr_c} \arrow[two heads]{r}{q^\circ} & \bGs \arrow[hook]{d}{pr_d} &\arrow[two heads]{l}{w^\circ} \bMs \arrow[hook]{d}{pr_a}  \\
-\arrow[two heads, bend right]{rrr}{}[swap]{f} \btU \arrow[two heads]{r}{r} & \bU \arrow[two heads]{r}{q} & \bG & \arrow[two heads]{l}{w}\bM  \\
-\end{tikzcd}
-\label{eq:commute}
-\end{equation}
-
-Let $x_0\in G^\circ$ and fix $u_0\in\bUs$, an element in the fiber of the covering map.
-A \emph{geodesic} on $G^\circ$ is a closed unit-speed straight-line path given as $\eta:\RR\rightarrow\bGs$.
-The \emph{period}/\emph{length} of the geodesic is some strictly positive $b\in\RR$.
-Let $\eta|_{[0,b]}=\alpha:[0,b]\rightarrow\bGs$ be a closed geodesic segment on $\bGs$ with initial and terminal point $x_0$.
-Denote $\alpha$'s unique lift to $\bUs$ by $\gamma:[0,b]\rightarrow\bUs$, where $\gamma(0)=u_0$, and $\gamma(b)$ in the fiber over $x_0$.
-Fix a point $\utild\in\btUs$ in the fiber over $u_0$.
-We define the unique lift of $\gamma$ to be the curve $\tilde{\gamma}:[0,b]\rightarrow\btUs$ with $\tilde{\gamma}(0)=\utild$.
-Denote their homotopy class representatives by $[\alpha]$ and (if they exist) $[\gamma],~[\tilde{\gamma}]$.
-
-The geodesic segment on $\bGs$ has the property that it returns to $x_0$ with a trivial effect on the holonomy of the curve.
-On the other hand, the lifted paths will close only if their homotopy classes belong in their fundamental groups, i.e.
-the kernel of homomorphisms $\varphi_1$, $\varphi_2|_{\ker\varphi_1}$.
-Otherwise they are mapped non-trivially to $\ZZ^2$ or $\ZZ^2\times\SO(2,\ZZ)$.
-Let $v_0\in\RR^2$ be a unit vector such that $\alpha's$ derivative is $v_0$ at times $0$ and $b$.
-The derivative of $\gamma$ is also $v_0$ at these times as per Corollary 1, as $\alpha$ is a geodesic with trivial holonomy.
-$\gamtild'$ are also $v_0$ at their endpoints.
-
-The Deck group of the map $\btUs\rightarrow\bUs$ is $\SO(2,\ZZ)$.
-This deck group acts naturally on the tangent bundles via push-forward of rotational matrices.
-We obtain the identity $T^1\btUs/\SO(2,\ZZ)\cong T^1\bUs$.
-This action is \emph{non-trivial}.
-Essentially, this means that there are 4 choices of a vector on the four-fold cover given a vector $v_0$ on $\bUs$.
-The choice of vector does not matter in the case where we flow it along a geodesic path, as a geodesic must return to a point in the fiber of $u_0$ in the same direction.
-That is to say that no matter where we start by lifting this path, we will always end up on the same plane (copy of $\mathbf{P}$) that we started in.
-In the case where a geodesic is periodic on $\bUs$, it would return to the same point in the fiber of $u_0$ as well.
-For simplicity's sake we will just lift to the vector mod the identity matrix, and call it $v_0$ as well.
-
-\begin{Def}
-Let $\Phi_t:\RR\rightarrow\bUs$ be a unit-speed geodesic on the \emph{Necker cube surface} with $\Phi(0)=u_0$ and $\Phi'(0)=v_0$.
-\\Similarly, let $\tilde{\Phi}_t:\RR\rightarrow\btUs$ be a unit-speed geodesic on $\bUs$'s four-fold cover with $\phitild(0)=\utild$ and $\phitild'(0)=v_0$.
-\end{Def}
-Let $\varphi=\varphi_1\times\varphi_2:\pi_1(\bGs)\rightarrow \ZZ^2\times\SO(2,\ZZ)$ be given by $[g]\mapsto(\varphi_1([g]),\varphi_2([g]))$. $\varphi$ is a
-product of group homomorphisms that takes a loop to 
-\begin{thm}
-Let $\delta=\varphi_1([\alpha])\in\ZZ^2$ be an element of $\Delta_{\varphi_1}$ and $\tilde{\delta}=\varphi([\alpha])\in\ZZ^2\times \SO(2,\ZZ)$ be an element of $\Delta_{\varphi}$. Then:\\
-$\Phi(t+b)=\delta\cdot\Phi(t)$ for all $t\in\RR$, and 
-$\phitild(t+b)=\tilde{\delta}\cdot\phitild(t)$ for all $t\in\RR$.
-\end{thm}
-
-\begin{cor}
-The following are equivalent:
-\begin{enumerate}
-\item $\Phi$ is periodic.
-\item $\phitild$ is periodic.
-\item $\delta=(0,0)$.
-\end{enumerate}
-\end{cor}
-
-\begin{cor}
-The following are equivalent:
-\begin{enumerate}
-\item $\Phi$ is drift-periodic.
-\item $\phitild$ is drift-periodic.
-\item $\delta\neq(0,0)$.
-\end{enumerate}
-\end{cor}
-
-These theorems essentially tell us that these types of geodesics behave identically on the Necker cube surface and its cover.
-Now we show that this is true for $\btUs$ and its metric completion, $\btU$.
-
-\begin{equation}
-\begin{tikzcd}
-\arrow[hookrightarrow, bend left]{rrr}{f_*^\circ}[swap]{} \arrow[two heads]{d}{PR_b} \pi_1(\btUs) \arrow[hook]{r}{r_*^\circ} & \pi_1(\bUs) \arrow[two heads]{d}{PR_c} \arrow[hook]{r}{q_*^\circ} & \pi_1(\bGs) \arrow[two heads]{d}{PR_d} & \ar[hookrightarrow]{l}{}[swap]{w_*^\circ} \pi_1(\bMs) \arrow[two heads]{d}{PR_a} \\
-\arrow[hookrightarrow, bend right]{rrr}{}[swap]{f_*}  \pi_1(\btU) \arrow[hook]{r}{r_*} & \pi_1(\bU) \arrow[hook]{r}{q_*} & \pi_1(\bG) & \ar[hookrightarrow]{l}{}[swap]{w_*} \pi_1(\bM)\\
-\end{tikzcd}
-\label{eq:homomorphism}
-\end{equation}
-
-\begin{Def}
-Let $\tildphi:\RR\rightarrow\btU$ be a geodesic on the unpunctured four-fold surface such that $\tildphi(0)=\utild$ and $\tildphi'(0)=v_0$.
-\end{Def}
-
-\begin{thm}
-$\tildphi$ is periodic if and only if $\phitild$ is periodic.\\
-$\tildphi$ is drift-periodic if and only if $\phitild$ is drift-periodic.
-\end{thm}
-
-This follows readily from the fact that geodesics are undefined on cone singularities. A geodesic on $\btUs$ can be mapped bijectively to a geodesic on  $\btU$ and vice-a-versa.
-
-\subsection{Translation Surface}
-Consider the projection $\btU\rightarrow\mathbf{M}$ onto a quotient of the surface under its translational symmetries.
-What we obtain is the following compact translation surface belonging to the stratum $\cH(2,2,2,2):$
-
-\begin{figure}[H]
-\centering
-\includesvg[width=2.6in]{mtildabw}
-\caption{Compact translation surface, $\mathbf{M}$ with edges and cone singularities (1,2,3,4) identified.
-The Roman numerals are meant to identify every ``copy" of $\bG$ with a direction differing from $v_0$ by rotation of some element in $\SO(2,\ZZ)$.
-The individual copies are then rotated so we have a canonical notion of direction }
-\label{fig:mtilda}
-\end{figure}
-
-We rearrange $\bM$ as such:
-
-\begin{figure}[H]
-\centering
-\includesvg[width=3.4in]{mtildastaircasebw}
-\caption{The staircase surface with directional planes and vertices identified.
-All edges are paired by translation.
-Two adjacent squares have opposite edges (A-M) identified.
-The top edge of the bottom-left square is glued to the bottom edge of the top-right square (both labeled I).
-Likewise, the bottom edge of the bottom-left square is identified with the top edge of the top-right square.}
-\label{fig:staircase}
-\end{figure}
-
-\noindent We recall some basic definitions and theorems concerning $\ZZ^2$covers of translation surfaces as they apply to $\bM$.
-
-\begin{Def}
-Algebraic intersection number is a non-degenerate bilinear form:
-\begin{align*}
-i:H_1(\bM,\QQ)\times H_1(\bM,\QQ)\rightarrow\QQ,
-\end{align*}
-for $[\gamma],[\beta]\in H_1(\mathbf{M},\QQ)$, $i([\beta],[\gamma])$ returns the signed intersection number of two homology classes.
-We say a crossing at the moment of intersection is positive if $\gamma$ makes a positive angle relative to $\beta$.
-\end{Def}
-The set of all affine diffeomorphisms of $\bM$ form the group $\text{Aff}^+(\bM)$.
-The $\emph{Veech group}$ of $\bM$ is the image in the co-domain of the homomorphism $D:\text{Aff}^+(\bM)\rightarrow SL(2,\RR)$ that takes every affine map to its derivative, which we denote $V(\bM)$.
-$\bM$ is a square-tiled translation surface whose Veech group is a finite index subgroup of $SL(2,\ZZ)$, thereby making it a \emph{Veech surface}.
-We use automorphisms of the fundamental group in $\text{Aut}(\pi_1(\bM))$ induced by affine maps to prove our main results.
-But first, we define a group homomorphism from $\pi_1(\bM)$ to $\pi_1(\bM)/\pi_1(\btU)$ by algebraic intersection numbers over a sum of linearly independent homology classes.
-Consider the following cylinder core curves on $\bM$ labeled $\gamma_i$ for $0\leq i\leq 11$:
-\begin{figure}[H]
-\includegraphics[width=3in]{homologyclass.png}
-\centering
-\caption{$\bM$'s cylinder core curves with u,v, and z homology class labels.}
-\label{fig:homology}
-\end{figure}
-
-The rank of a $12\times12$ intersection matrix of these curves is 10. Therefore these curves span $H_1(\bM,\QQ)$.
-
-\begin{Def}
-The homology classes $u,v,z$ are given as the following sums of core curves:
-\begin{align*}
-u &= -[\gamma_2] +[\gamma_5] + [\gamma_8] - [\gamma_{11}],\\
-v &= +[\gamma_0] -[\gamma_3] -[\gamma_6] +[\gamma_9],\\
-z &= +[\gamma_1] +[\gamma_4]-[\gamma_7]-[\gamma_{10}].
-\end{align*}
-\end{Def}
-
-\begin{Def}
-Define the group homomorphism $\Omega_{u,v}:\pi_1(\bM)\rightarrow \ZZ^2$, where $\beta\mapsto(i(u,[\beta]),i(v,[\beta]))$.
-\end{Def}
-
-\begin{lem}
-\compav{show somehow that u,v determine the cover}
-\end{lem}
-
-\subsection{Induced Automorphisms of $H_1(\bM,\QQ)$}
-We look at some important automorphisms induced by Affine maps on $\bM$.
-Observe that $\bM$ has a uniform cylinder decomposition in both horizontal and vertical directions as in figure $15$.
-We define the \emph{modulus}, $\mu$, of a cylinder to be the ratio of the cylinder's width to its circumference, $\frac{w}{c}$.
-The $\emph{Dehn-twist}$ of a cylinder is an affine diffeomorphism that skews the cylinder and sends every vertex to itself.
-
-\begin{figure}[H]
-\centering
-\includesvg[width=3in]{cylinderskew}
-\label{fig:skew}
-\caption{Dehn-twist of a cylinder in $\bM$'s cylinder decomposition.}
-\end{figure}
-
-Such a map has derivative $\left[~\begin{matrix}1 && \pm\mu^{-1}\\0 && 1\end{matrix}~\right]$.
-On $\bM$ all cylinders in both vertical and horizontal decompositions have moduli $\mu=\frac{1}{2}$.
-These give way to global diffeomorphisms as $\emph{multi-twists}$ of $\bM$.
-
-\begin{Def}
-We define the global affine diffeomorphisms of $\text{Aff}^+(\bM)$ obtained as \textbf{multi-twists of the surface in horizontal and vertical directions} $\psi_h$ and $\psi_v$, respectively.
-\end{Def}
-
-The \emph{derivative map} of the affine group is the homomorphism $D:\text{Aff}^+(\bM)\rightarrow \SL(2,\RR)$.
-The \emph{Veech group} of the surface is the image of the affine group under this map, $D(\text{Aff}^+(\bM))$, denoted $V(\bM)$.
-The twists $\psi_h$ and $\psi_v$ under $D$ are mapped to the same matrices as the Dehn-twists on its individual cylinders. i.e.
-\begin{align*}
-D(\psi_h)=\psi_h'=\left[ \hspace{1mm} \begin{matrix}
-				1 &   2\\
-				0 & 1
-			\end{matrix}\hspace{1mm}\right] \hspace{0.5in}
-			D(\psi_v)=\psi_v'=\left[ \hspace{1mm} \begin{matrix}
-							1 & 0\\
-							 2 & 1
-						\end{matrix}\hspace{1mm}\right]
-\end{align*}
-
-These are parabolic elements of $V(\bM)$ that generate a free group of rank 2. [cite].
-\begin{Def}
-We call $\XX$ the generated parabolic subgroup of $\text{Aff}^+(\bM)$, where $\XX=\<\psi_h,\psi_v \>$.
-\end{Def}
-
-The image of this group under $D$ is itself a subgroup of $V(\bM)$ and denoted $D(\XX)=\XX'=\<\psi_h',\psi_v' \>$ as in the introduction.
-Note that when skewing the surface and piecing it back together, the cylinder curves intersect each other:
-
-\begin{figure}[H]
-\centering
-\includesvg[width=2in]{skew}
-\end{figure}
-
-By skewing in the horizontal direction, $\psi_h$ preserves the even indexed cylinder curves, but the odd indexed vertical curves $\gamma_i$ intersect the two horizontal curves (positive intersection), $\gamma_{i-1}$ and $\gamma_{i+1}$.
-A similar observation can be made from applying $\psi_v$ where the vertical curves are preserved, but the horizontal curves intersect with adjacent vertical curves (negative intersection).
-Let $i\in\ZZ/12\ZZ$.
-Then this convenient indexing scheme gives us the following formulas for $\psi_h$ and $\psi_v$'s counterparts in $\Aut(H_1(\bM,\QQ))$ raised to $k\in\ZZ\backslash\{0\}$:
-
-\begin{align*}
-\psi^{*k}_h([\gamma_i])=&[\gamma_i] + \frac{k}{2}(1-(-1)^i)([\gamma_{i-1}]+[\gamma_{i+1}])\\
-\psi^{*k}_v([\gamma_i])=&[\gamma_i] - \frac{k}{2}(1+(-1)^i)([\gamma_{i-1}]+[\gamma_{i+1}])\\
-\end{align*}
-
-We call this generated group $\XX^*=\<\psi^{*}_h,\psi^{*}_v \>$.
-There are no relations between its generators since $i$ being even or odd has one act trivially and the other non-trivially.
-Therefore this group is also free and rank 2.
-Hence $\XX^*\cong\XX'\cong\XX$.
-Let $g\in\XX$, $g'\in\XX'$, $g^*\in\XX^*$ be isomorphic elements in their respective groups.
-Let $g=\prod_{j=0}^{n}\psi_h^{k_j}\circ\psi_v^{m_j}$ be a word in $\XX$ as a finite left product of multi-twist pairs with $k_j,m_j\in\ZZ$.
-Let $\psi_j=\psi_h^{k_j}\circ\psi_v^{m_j}$ be the $j^{th}$ word pair.
-Elements $g',g^*$ have identical representations.
-\\\\
-The $j^{th}$ element pair that is mapped into $\Aut(H_1(\bM,\QQ))$ can be expressed as
-
-\begin{align*}
-\psi_j^*([\gamma_i])=&[\gamma_i]+\frac{k_j(1-(-1)^i)}{2}([\gamma_{i-1}]+[\gamma_{i+1}])\\&-\frac{m_j(1+(-1)^i)}{2}(k_j(2[\gamma_i]+[\gamma_{i-2}]+[\gamma_{i+2}]) +[\gamma_{i-1}]+[\gamma_{i+1}]).
-\end{align*}
-
-\noindent Let $\sum\Gamma=\sum_{i=0}^{11}[\gamma_i]$, and denote the odd and even indexed sums as $\sum\Gamma_o$ and $\sum\Gamma_e$, respectively.
-We consider three kinds of closed geodesic segments on $\bM$ and don't consider initial/terminal points for the time being.
-These are the slope one, horizontal, and vertical loops.\\
-
-\noindent Call the geodesic in direction $\theta=\frac{\pi}{4}$ that runs up the staircase intersecting every $\gamma_i$ once $\chi$.
-Denote its homotopy class by $[\chi]$.
-Observe from Figure 10 that $\chi$ is homologous to the sum of cylinder curves divided by 2.
-In Figure 1, this is the closed geodesic that lifts to a closed loop that looks like a 12-gon on $\bUs$.
-
-\begin{lem}
-$\Omega_{u,v}([\chi])=(0,0).$
-\begin{proof}
-Observe that $\Omega_{u,v}([\chi])=\Omega_{u,v}(\frac{1}{2}\sum\Gamma)=\frac{1}{2}(i(u,\sum\Gamma),i(v,\sum\Gamma)).$ Since only adjacent curves intersect, $i(u,\sum\Gamma)=i(-[\gamma_2],[\gamma_{1}]+[\gamma_{3}]) +i([\gamma_5],[\gamma_{4}]+[\gamma_{6}]) + i([\gamma_8],[\gamma_{7}]+[\gamma_{9}]) - i([\gamma_{11}],[\gamma_{10}]+[\gamma_{0}])=-2+(-2)+2-(-2)=0$.\\
-Similarly, $i(u,\sum\Gamma)=i([\gamma_0],[\gamma_{11}]+[\gamma_{1}]) -i([\gamma_3],[\gamma_{2}]+[\gamma_{4}]) -i([\gamma_6],[\gamma_{5}]+[\gamma_{7}]) +i([\gamma_9],[\gamma_{8}]+[\gamma_{10}])=2-(-2)-2+(-2)=0$.
-Therefore, $\Omega_{u,v}([\chi])=(0,0)$.
-\end{proof}
-\end{lem}
-
-\begin{lem}
-Let $g$ be as above. If $\alpha=g(\chi)$, then $[\alpha]=\frac{1}{2}(c_1\sum_e+c_2\sum_o)$ for $c_1,c_2\in\ZZ$.
-\begin{proof}(By Induction on $\XX^*$).\\
-Consider $\psi_0(\chi)$.\\
-$[\psi_0(\chi)]=\psi_0^*[\chi]=\frac{1}{2}\psi_0^*(\sum\Gamma)=\frac{1}{2}\psi_0^*(\sum\Gamma_e+\sum\Gamma_o)=\frac{1}{2}(\psi_0^*(\sum\Gamma_e)+\psi_0^*(\sum\Gamma_o))$.
-Now\\ $\psi_0^*(\sum\Gamma_e)=\sum_{i=0}^{11}\psi_0^*([\gamma_{2i}])=\sum_{i=0}^{11}( [\gamma_{2i}]-2m_0k_0[\gamma_{2i}]-m_0k_0[\gamma_{2i-2}] $\\$ -m_0k_0[\gamma_{2i+2}]-m_0[\gamma_{2i+1}]-m_0[\gamma_{2i-1}] )=(1-4m_0k_0)\sum\Gamma_e+(-2m_0)\sum\Gamma_o$.\\
-Now\\ $\psi_0^*(\sum\Gamma_o)=\sum_{i=0}^{11}\psi_0([\gamma_{2i+1}])=\sum_{i=0}^{11}([\gamma_{2i+1}] +k_0([\gamma_{2i}]+[\gamma_{2i+2}]))=\sum\Gamma_o+2k_0\sum\Gamma_e$.\\
-And so $\frac{1}{2}\psi_0^*(\sum\Gamma)=\frac{1}{2}((1+2k_0-4k_0j_0)\sum\Gamma_e +(1-2m_0)\sum\Gamma_o )$.\\\\
-Induction: Suppose $\frac{1}{2}\psi^*_j(\sum\Gamma)=c_1\sum\Gamma_e+c_2\sum\Gamma_o$.\\
-Then $\psi^*_{j+1}(\psi^*_{j}(\sum\Gamma))=c_1\psi^*_{j+1}(\sum\Gamma_e)+c_2\psi^*_{j+1}(\sum\Gamma_o)=\\
-c_1((1-4m_{j+1}k_{j+1})\sum\Gamma_e+(-2m_{j+1})\sum\Gamma_o)+c_2(\sum\Gamma_o+2k_{j+1}\sum\Gamma_e)=\\
-(c_1(1-4m_{j+1}k_{j+1}) + 2k_{j+1}c_2)\sum\Gamma_e+(-2m_{j+1}c_1+c_2)\sum\Gamma_o$.
-
-\end{proof}
-\end{lem}
-
-\begin{thm}
-Let $g$ be as above. Suppose $\alpha=g(\chi)$. Then $\Omega_{u,v}([\alpha])=(0,0)$. 
-\begin{proof}
-$\Omega_{u,v}([\alpha])=\Omega_{u,v}([g(\chi)])=\Omega_{u,v}(g^{*}([\chi]))=\frac{1}{2}(i(u,g^*(\sum\Gamma)) i(v,g^*(\sum\Gamma))$.\\
-$i(u,g^*(\sum\Gamma)=c_1 i(u,\sum
-\Gamma_e)+c_2 i(u,\sum\Gamma_o)=\\ c_1(i([\gamma_5],[\gamma_{4}] - i([\gamma_{11}],[\gamma_{10}]+[\gamma_{0}]))\\+c_2(i(-[\gamma_2],[\gamma_{1}]+[\gamma_{3}]) + i([\gamma_8],[\gamma_{7}]+[\gamma_{9}]) )=\\c_1((-2)-(-2))+c_2(-2+ 2 )=0\\$
-Holds similarly for $i(v,g^*(\sum\Gamma)).$\\
-$\therefore \Omega_{u,v}([\alpha])=(0,0)$.
-\end{proof}
-\end{thm}
-
-\begin{lem}
-$\Omega_{u,v}([\gamma_i])\neq (0,0)$.
-\begin{proof}
-every curve intersects nontrivially with exactly one of either $u$ or $v$.
-\end{proof}
-\end{lem}
-
-\begin{thm}
-Suppose $\beta=g(\gamma_i)$. Then $\Omega_{u,v}([\beta])\neq (0,0)$.
-\begin{proof}
-$g^*([\gamma_i])=[\gamma_i]+C$.
-If $i(u,g^*([\gamma_i]))=i(u,[\gamma_i]+C)=i(u,[\gamma_i])+i(u,C)=0$, then 
-\end{proof}
-\end{thm}
-
-\section{Proof of Main Theorem}
-
-$S=\cO\cup \cE$.
-Let $\XX'$ act on $\cS$ by linear transformations.
-
-\begin{lem}
-
-\end{lem}
-
-\begin{lem}
-$\XX'$ acts faithfully and transitively on integer pairs in $\cO$, partitions $\cE$ into two sets and acts faithfully and transitively on them, and acts faithfully on $\cS$.
-\begin{proof}
-Let $v=kv_0=\begin{pmatrix}x \\ y\end{pmatrix}\in\cS$ with $||v_0||=1$ and $k\in\RR$, and let $A=\begin{pmatrix}a & b \\ c & d
-\end{pmatrix}\in\XX'$.
-The kernel of the homomorphism $SL(2, \ZZ)\rightarrow \SL(2,\ZZ/2\ZZ)$ that maps a matrix to its $(mod~2)$ equivalent is precisely $\XX'$ since it maps the generators trivially.
-[quote this paper for the proof of this maybe?
-http://www.fysik.su.se/~ingemar/SL2.pdf ]\\
-$x,y$ are relatively prime, so Bezout's identity states there exist $m,n\in\ZZ$ such that $mx+ny=1~(*)$.
-It's obvious that this action is faithful on the sets since the stabilizer of any vector in $\RR^2$ is trivial.
-Suppose $v\in\cO$.
-Consider the orbit of $\begin{pmatrix}1 \\1 \end{pmatrix}$.
-Then there is some matrix $A\in\XX'$ such that $A\begin{pmatrix}1\\1\end{pmatrix}=v$.
-Since $A\in\SL(2,\ZZ)$, $\det A = ad-bc=1$.
-Obtain from $1=(x-b)(y-c)-bc=xy-xc-by$ that $xy-1=xc+by$.
-From $(*)$ we see that $xy-1=x(u(xy-1))+y(v(xy-1))$.
-Let $c=u(xy-1)$ and $b=v(xy-1)$.
-Since $xy-1$ is even, $c,b$ are even.
-It follows then that $a=x-b$ and $d=y-c$ are odd.
-Therefore $A\in \XX'$ and the action is transitive on $\cO$.
-If $v\in\cE$, then either $x$ or $y$ is even.
-In either case this would mean that either $a$ or $d$ is even, which would mean $A\notin\XX'$, a contradiction.\\
-Now consider the orbit of $\begin{pmatrix}1\\0\end{pmatrix}$.
-Left multiplying by $A$ would mean $a=x$ and $c=y$.
-Split $\cE$ into sets $\cE_1=\{(x,y)\in\cE :x\in2\ZZ,~y\in(2\ZZ+1) \}$ and $\cE_2=\{(x,y)\in\cE:x\in(2\ZZ+1),~y\in2\ZZ \}$.
-Since $\det A=1=da-bc=dx-by=mx+ny$, we have that $d=m$ and $b=-n$.
-We rule out when $v\in\cO,\cE_1$ since $c\neq 0 (\mod 2)$ in those cases.
-Consider $v\in\cE_2$.
-Then $x$ is odd and $y$ is even.
-$d$ must be odd since $mx+ny=1$ and $y$ is even.
-And $b$ must be even since $ad-bc=1$\\
-\compav{finish this and say the proof for $\cE_1$ is analogous.}
-
-\end{proof}
-\end{lem}
-
-\begin{Def}
-Let $\phi_t^\theta:\RR\rightarrow\bM$ be a unit-speed geodesic on $\bM$ in direction $\theta$.
-\end{Def}
-
-\subsection{empty}
-\noindent Consider the following families of unit squares in $\RR^3$:\\
-\vspace{0.2in}
-\begin{tabular}{p{10cm}c}
-\begin{align*}
-\mathbf{A}_{m,n,p} = [m, m+1]\times[n,n+1]\times\{p\}, 
-\\\mathbf{B}_{m,n,p} = \{m+1\}\times[n,n+1]\times[p-1,p],
-\\\mathbf{C}_{m,n,p}= [m,m+1]\times\{n+1\}\times[p-1,p].
-\end{align*}
-&
-\shiftleft{0.8in}{\raisebox{-1in}{\includegraphics[scale=1]{label.png}}}
-\end{tabular}
-
-\begin{figure}[H]
-\centering
-\includesvg[width=1in]{cubesxyzcut}
-\caption{A section of $\mathbf{S}$}
-\end{figure}
-
-\subsection{Flattening $\mathbf{S}$}
-At the moment it is difficult to describe how the parallel transport of a vector tangent to the surface along an arbitrary path on the surface acts on the vector in $\RR^3$.
-To simplify this problem we take $\mathbf{S}$ to an isometric variant embedded in $\RR^2$ by piecewise linear transformations on the sets
-\begin{align*}
-\mathbf{A}=\bigcup\big{\{}\mathbf{A}_{m,n,p}: m+n+p=0\big{\}},
-\\\mathbf{B}=\bigcup\big{\{}\mathbf{B}_{m,n,p}: m+n+p=0\big{\}},
-\\\mathbf{C}=\bigcup\big{\{}\mathbf{C}_{m,n,p}: m+n+p=0\big{\}}.
-\end{align*}
-
-\begin{Def}
-Let $\Psi:s^*(\mathbf{S})\rightarrow\RR^3$ be given as
-\begin{equation}
-\Psi\left[\begin{array}{c}
-	x\\y\\z
-\end{array}\right] 
-= 
-\begin{cases}
-	\left[ \hspace{2mm} \begin{matrix}
-		1 & 0 & 0 \\
-		0 & 1 & 0 \\
-		0 & 0 & 1
-	\end{matrix}\hspace{3mm}\right]
-
-	\left[\begin{array}{c}
-	x - \left\lfloor x \right\rfloor
-	\\ y- \left\lfloor y \right\rfloor
-	\\ -(z - \left\lfloor z \right\rfloor)
-	\end{array} \right]
-	+
-	\left[\begin{array}{c}
-		2 \left\lfloor x \right\rfloor - \frac{3}{2}
-		\\ 2\left\lfloor y \right\rfloor - \frac{3}{2}
-		\\ z - \left\lfloor z \right\rfloor
-	\end{array} \right]
-		& \text{if } (x,y,z)\in \mathbf{A}	\vspace{2mm}
-	\\
-
-	\left[ \begin{matrix}
-	0 & 0 & 1 \\
-	0 & 1 & 0 \\
-	-1 & 0 & 0
-	\end{matrix}\hspace{2mm}\right]
-	\left[\begin{array}{c}
-		x - \left\lfloor x \right\rfloor
-		\\ y- \left\lfloor y \right\rfloor
-		\\ -(z - \left\lfloor z \right\rfloor)
-		\end{array} \right]
-	+
-		\left[\begin{array}{c}
-			2 \left\lfloor x \right\rfloor - \frac{3}{2}
-			\\ 2\left\lfloor y \right\rfloor - \frac{3}{2}
-			\\ x - \left\lfloor x \right\rfloor
-		\end{array} \right]
-			& \text{if } (x,y,z)\in \mathbf{B}\backslash(\mathbf{A}\cup\mathbf{C})	\vspace{2mm}
-	\\
-
-		\left[ \begin{matrix}
-		1 & 0 & 0 \\
-		0 & 0 & 1 \\
-		0 & -1 & 0
-		\end{matrix}\hspace{2mm}\right]
-		\left[\begin{array}{c}
-			x - \left\lfloor x \right\rfloor
-			\\ y- \left\lfloor y \right\rfloor
-			\\ -(z - \left\lfloor z \right\rfloor)
-			\end{array} \right]
-		+
-			\left[\begin{array}{c}
-				2 \left\lfloor x \right\rfloor - \frac{3}{2}
-				\\ 2\left\lfloor y \right\rfloor - \frac{3}{2}
-				\\ y - \left\lfloor y \right\rfloor
-			\end{array} \right]
-				& \text{if } (x,y,z)\in \mathbf{C}\backslash(\mathbf{A}\cup\mathbf{B})	\vspace{2mm}
-\end{cases}
-\end{equation}
-\end{Def}
-\noindent This map behaves as such:
-\begin{figure}[H]
-\centering
-\includesvg[width=1.1in]{cubesxyzcut} \hspace{0.1in}\raisebox{1.0in}{\text{$\rightarrow$}}\hspace{0.1in}
-\raisebox{0.4in}{\includesvg[width=1.5in]{unfoldcut}}
-\caption{An isometry of the surface $\mathbf{S}$.}
-\label{fig:Psi}
-\end{figure}
-
-\begin{lem}
-$\Psi\circ s^*$ is an isometry of the surface $\mathbf{S}$
-\begin{proof}
-$\Psi$ is well-defined on its domain $s^*(\mathbf{S})$, and $s^*$ is an embedding of the surface that is a bijection when restricting its co-domain to its image.
-Since $\Psi$ is a linear transformation composed of Euclidean matrices and translations it is also invertible and bijective.
-\end{proof}
-\end{lem}
-
-\subsection{Translation Surface Cover of $\bGs$}
-
-\subsection{A Four-Fold Cover of $\bUs$}
-
-\newpage
-\begin{Def}
-$\bUs$ is the cover of $\bGs$ such that $\pi_1(\bUs)=\ker\varphi$.
-\end{Def}
-
-Our cover of $\bGs$ has the fundamental group, $\pi_1(\bUs)=\<\<a^{-1}b^{-1}ab, c, d\>\>$, the conjugate subgroup of elements that map trivially under $\varphi_1$.
-The covering map $k_{\varphi_1}:\bUs\hookrightarrow\bGs$ takes every point on $\bUs$ to its modular equivalent under these translational symmetries.
-$\bUs$ is realized as an infinite $\ZZ^2$-tiling of $\bGs$ in $\CC$:
-\begin{figure}[H]
-\centering
-\includesvg[width=2in]{uprime}
-\caption{The infinite surface $\bUs$ embedded in $\CC$ with edges identified.}
+\includesvg[width=4.3in]{neckercubepathsurface}
 \end{figure}
@@
 \begin{Def}
-$\bU$ is the completion of $\bUs$ that includes the branched vertices. 
+Let $\varphi_1:\pi_1(\bGs)\rightarrow \ZZ^2$ be the homomorphism on the free group $\<a,b,c,d\>$ given by $c,d\mapsto(0,0)$, $a\mapsto(1,0)$, and $b\mapsto(0,1)$.
 \end{Def}
 
-\begin{rem}
-The surface $\bG$ is homeomorphic to the torus, and its cover $\bU$ is its universal cover as the paths labeled $c,d$ are only non-trivial when $\bG$ is branched at its three conical points.
-\end{rem}
-
-\subsection{}
-
-\subsection{Isometry From $\mathbf{S}$ to $\bU$}
-The Necker cube surface can be \emph{flattened} onto the plane by piecewise isometric maps onto a subspace of $\RR^3$ with (open) unit squares removed at every even integer pair in the plane, what we claim to be $\bU$.
-
-\noindent The red/blue dotted lines represent the edges that are split on the plane. The map from one surface to the other is composed of piecewise isometries, $\Psi:\mathbf{S}\rightarrow\RR^3$.
-
-The flattened surface is contained entirely in $(x,y,0)\in\RR^3$, which is isometric to $\CC$. $\bU$ is recovered as a topological quotient on the domain
-
-\begin{Def} $\mathbf U$ is the surface obtained as the topological quotient $\mathbf{P}/\sim_{\mathbf{P}}$, where $\CC$ is identified with $\RR^2$ in the usual way and $\sim_{\mathbf{P}}$ is a minimal relation on $\mathbf{P}$ defined as follows:
-\begin{gather*}
-\text{Let } x_{0}=(u_{0},v_{0}),x_{1}=(u_{1},v_{1}) \in \mathbf{P}.  \text{ }\sim_{\mathbf{P}} \text { is given as the relation }\\ x_{0}\sim_{\mathbf{P}}x_{1}  \text{ iff } x_{0}=x_{1}
- \\\text{ or, for some }m,n\in\ZZ~x_{0},x_{1} \in {\partial} \left( \left[2m-\frac{1}{2},2m+\frac{1}{2}\right] \times \left[2n-\frac{1}{2},2n+\frac{1}{2}\right] \right)\\
-  \left[\begin{array}{c}
-u_{1} -2m
-\\v_{1}-2n
-\end{array}\right] = \left[\begin{matrix}
-0 && 1\\
-1 && 0
-\end{matrix}\right]
-\left[ \begin{array}{c}u_{0}-2m\\
-v_{0}-2n
-\end{array}\right]
-.\end{gather*}\end{Def}
-
-\begin{rem}
-Our map $\Psi$ induces an isometry between $\mathbf{S}$ and $\bU$, and its restriction to the branched surfaces induces an isometry between $\mathbf{S}^\circ$ and $\bUs$.
-This follows from these piecewise Euclidean transformations that preserves our induced flat metric.
-\end{rem}
-From here on we use $\bUs$ and $\bU$ instead of $\mathbf{S}$ and $\mathbf{S}^\circ$.
-
-\subsection{Four-Fold Cover}
-We denote the \emph{unit tangent bundle} on the surface $\bUs$ by $T^1\bUs$, and construct a cover with trivial holonomy.
-Initial experiments have shown us that any geodesic viewed as discontinuous line segments on $\mathbf{P}$ moves in at most four directions.
-\begin{figure}[H]
-\centering
-\includesvg[width=1.8in]{cubesxyzpaths}
-\includesvg[width=2.8in]{unfoldcutpath}
-\caption{Simple periodic (c) and drift-periodic (a,b) trajectories represented as line segments on $\mathbf{S}$ and $\mathbf{U}$.\compav{RELABEL THESE}}
-\label{fig:holonomypath}
-\end{figure}
-Observe that the parallel transport of a vector around a closed loop on $\bUs$ will act on vectors tangent to the surface by a rotation of an integer multiple of $\frac{\pi}{2}$ radians (since the surface is flat and embedded in $\CC$, we work with vectors in $\RR^2$).
+The cover of $\bGs$ with fundamental group $\ker\varphi_1$ is the space $\bUs=\tilde{\mathbf{G}}^\circ/\ker\varphi_1$ where $\ker\varphi_1$ acts on $\bGs$'s universal cover,  $\widetilde{\mathbf{G}}^\circ$, by deck transformations.
+We can see that $\bGs$ is obtained as the quotient of the Necker cube surface by these $\ZZ^2$ symmetries.
 
 \begin{Def}
-Define $\varphi_2:\pi_1(\bGs)\rightarrow SO(2,\ZZ)$ as the group homomorphism on generators $a,b,c,d$ such that:
-\begin{align*}
-a,b\mapsto \begin{pmatrix}1 && 0 \\ 0 && 1\end{pmatrix}=I_2,~~~
-c\mapsto \begin{pmatrix}0 && -1 \\ 1 && 0\end{pmatrix}=R,~~~
-d\mapsto \begin{pmatrix}0 && 1 \\ -1 && 0\end{pmatrix}=R^3
-\end{align*}
+Let $p_1:\bUs\rightarrow\bGs$ be a projection that induces a monomorphism on the fundamental groups of these surfaces such that $p_{1*}\pi_1(\bUs)=\ker\varphi_1$.
 \end{Def}
-The homomorphism restricted to $\pi_1(\bUs)$ factors through $\pi_1(\bGs)$ and we show that:
-\begin{lem}
-The group $\varphi_2(\pi_1(\bUs,x_0))=Hol_p(\omega)$.
-\begin{proof}
-Since $c\mapsto R$ and $d\mapsto R^3$, the image of $\varphi_2$ is generated by $R,R^3$ and isomorphic to $SO(2,\ZZ)$.
-We see from figure $\ref{fig:holonomypath}$ that parallel transports of vectors by non-trivial paths produce clockwise/counter-clockwise rotations equal to that of of the cone angles of the singularities they loop around, all integer multiples of $\frac{\pi}{2}$ that coincide with paths generated by elements $c,d$.
-Further, if there is an element in $\pi_1(\bUs,x_0)$ conjugated by $[a,b]$, the effect this would have on holonomy would be trivial as the total cone angle sum around each cut out square is an integer multiple of $2\pi$, which agrees with $a$ and $b$'s trivial images under $\phi_2$.
-\end{proof}
-\end{lem}
-
-Hence, we obtain a natural monodromy representation with the map $m:\pi_1(\bUs,x_0)\rightarrow SO(2,\ZZ)\cong Hol_p(\omega)$, where for $[\gamma]\in\pi_1(\bUs,x_0)$ we have that $\gamma^*(1)=(x_0,m([\gamma])v_0)$.
-It follows that since $\omega$ is a flat connection, trivial loops have trivial holonomy and $Hol_p(\omega)$ acts on $H(p)$.
-
+Fix some point $u_0$ in $\bUs$ in the fiber over $x_0$, $p_1^{-1}(x_0)$. These are both contained in smooth sections, i.e. a face of a cube.
 \begin{Def}
-Let $\btUos$ be the cover of $\bUs$ with fundamental group $\pi_1(\btUos)=\ker\varphi_2|_{\pi_1(\bUs)}\trianglelefteq \pi_1(\bUs)$.
+Define $T_{u_0}(\bUs)$ as the vector space tangent to the surface at $u_0$.
+We denote the unit tangent space at this point $T^1_{u_0}(\bUs)$.
+Both belong to their respective tangent bundles, $T(\bUs)$ and $T^1(\bUs)$.
 \end{Def}
 
-\begin{lem}
-Let $\Phi_t:\bUs\times\RR\rightarrow \bUs$ be a unit-speed geodesic flow on $\bUs$, with a parallel transport map induced by $\omega$. Then the following is true:
-\begin{enumerate}[label=(\roman*)]
-\item $\Phi$ is periodic if and only if its lift to $\btUos$ is.
-\item $\Phi$ is drift-periodic if and only if its lift to $\btUos$ is. 
-\end{enumerate}
-\begin{proof}
-Let $x_0=\Phi(0)$ be an initial point in $\bUs$, and let $v_0=\frac{d}{dt}\Phi|_{t=0}\in\RR^2$ be its initial direction.
-Then $\Phi^*:T^1\bUs\times\RR\rightarrow T^1\bUs$ is a well-defined lift to the unit tangent bundle with initial point $p=(x_0,v_0)$.
-\\\emph{(i)}.
-Suppose $\Phi$ is periodic with period $T\in\RR$.
-Then $\Phi$ is reparameterized as $\alpha:[0,1]\rightarrow \bUs$, where $[\alpha]\in\pi_1(\bUs,x_0)$.
-It follows then that if $\alpha$ does not lift to a closed path, then $\alpha$ must have non-trivial holonomy since $[\alpha]\notin\ker\varphi_2$.
-That is when lifted to the unit tangent bundle with base point $p$, $\alpha^*(1)\neq p$ since $m([\alpha])\neq I_2$.
-But this implies that $\Phi^*(T)\neq p$, which is impossible.
-Hence $[\alpha]\in\ker\varphi_2=\pi_1(\btUos,\tilde{x}_0)$, where $\tilde{x}_0$ belongs to the fiber over $x_0$ under the covering map $\btUos\hookrightarrow\bUs$.
-The converse holds trivially.
-\\\emph{(ii)}.
-Suppose $\Phi$ is drift-periodic with period $T\in\RR$ and non-trivial $f:\bUs\rightarrow\bUs \in Trans(\bUs)\cong\pi_1(\bGs)/\pi_1(\bUs)\cong \ZZ^2$ such that $\Phi(T)=f(x_0)$.
-Re-parameterize this as $\alpha:[0,1]\rightarrow\bUs$ with $[\alpha]\in\pi_1(\bGs)$.
-When lifted to $T^1\bUs$, $\Phi^*(0)=p$ and $\Phi^*(T)=(f(x_0),v_0)$.
-Let $\alpha:[0,1]\rightarrow \bUs$ be its reparameterization up to time $T$.
-Thus $[\alpha]\notin \pi_1(\bUs,x_0),\pi_1(\btUos,\tilde{x}_0)$.
-Further, $f$ has a unique, lift to $\tilde{f}\in Trans(\btUos)$ because the space is connected.
-(a bit stuck here).
-\\
-Conversely, ...
-\end{proof}
-\end{lem}
-
-A visual representation of $\btUos$ is as a four-fold cover of $\bUs$ with trivial holonomy on all closed paths. Arbitrary paths do \emph{not} have trivial holonomy:
-
-\begin{figure}[H]
-\centering
-\includesvg[width=1.5in.]{quotient}
-\caption{A $2\times2$ cut out section centered at each missing square. Edges and vertices identified.}
-\label{fig:quotient}
-\end{figure}
-
-\subsection{Translation Surface}
-
-A better picture of $\btUos$ is obtained by making cyclic edge identifications on $\mathbf{P'}=\mathbf{P}\times\ZZ/4\ZZ$.
-This comes from $\btUos$ inheriting the topological properties of $\bUs\times \pi_1(\bUs)/\pi_1(\btUos)\cong \bUs\times SO(2,\ZZ)$.
-
-\begin{figure}[H]
-\centering
-\includesvg[width=4in]{utilda}
-\label{fig:utilda0}
-\caption{Branched cover of $\mathbf{U}$ of degree four.}
-\end{figure}
-
+\subsection{Discussion of Results}
+For every direction in the tangent space at $u_0$ there is a unique geodesic on $\bUs$.
 \begin{Def}
-$\btU_0$ is the surface obtained as the quotient $\mathbf{P}'/\sim_{\mathbf{P}'}$.
-Denote paths $\zeta$ and $\eta$  on $\mathbf{P}$, parameterized by integers $m,n\in\ZZ$ and $t\in[-\frac{1}{4},\frac{1}{4}]$, and defined:\\ $\zeta(t)=(2m+t)+i(2n-\frac{1}{2})$,\\
-$\eta(t)=(2m+t)+i(2n+\frac{1}{2})$.\\
-
-Let $\aleph=2m+i2n\in\CC$. The minimal relation $\sim_{\mathbf{P}'}$ is given as:
-\begin{equation}
-\begin{split}
-(\zeta(t);j)\sim_{\mathbf{P'}}(e^{i\frac{\pi}{2}}~\overline{\zeta(t)-\aleph};j+1)\\
-(\eta(t);j)\sim_{\mathbf{P'}}(e^{i\frac{\pi}{2}}~\overline{\eta(t)-\aleph};j+1)
-\end{split}
-\end{equation}
+Let $\Phi_t:\RR\rightarrow\bUs$ be the unit-speed geodesic with $\Phi(0)=u_0$ and $\frac{d}{dt}\Phi(0)=v_0\in T^1_{u_0}(\bUs)$.
+It is the image of the geodesic flow based at $u_0$ on the unit tangent bundle in direction $v_0$.
+We additionally require that at no point does the geodesic encounter a cone singularity of the surface.
 \end{Def}
-This relation is similar to $\sim_\mathbf{P}$ in that it translates line segments of squares surrounding even integer pairs to the origin and relates points on one edge of a square to points on an adjacent edge.
-Where it differs is that these adjacent edges now belong to a different ``copy" of $\mathbf{P}$.
-It is in this cyclic manner that edges are glued that allows for trivial linear holonomy on arbitrary paths by rotating each copy of $\mathbf{P}$ accordingly.
-For example, here is a path on a section of the surface in the neighborhood of $P=2m+i2n\in\CC$ $(m,n\in\ZZ)$ after rotation:
-
-\noindent and its subsequent projection onto $\bUs:$
-
+We say that $\Phi$ is \emph{periodic} if for some $T>0$, $Phi(t+T)=\Phi(t)$.
+Let $f$ be a non-trivial automorphism of the surface.
+We say that $\Phi$ is \emph{drift-periodic} if for some $T'>0$, $\Phi(t+T')=f(\Phi(t))$.
+The initial trajectory angle is said to be $\emph{rational}$ if there is some $k\in\RR$ such that $kv_0$'s non-zero components are relatively prime integers.
+Although $v_0$ is a vector of $\RR^3$, the tangent spaces are parallel to subspaces generated by exactly two standard basis vectors.
+For that reason, one component is always 0 and we will just treat it as an element of $\RR^2$.
+A rational initial trajectory angle in $\RR^2$ falls into one of two categories:
 \begin{Def}
-Let $r:\mathbf{P}'\rightarrow\mathbf{P}'$ be the isometric map of $\btU_0$'s domain given  $r(z;\bar{j})=(e^{(-j)i\frac{\pi}{2}}z;\bar{j})$.
+Let $v_0$ be a rational unit vector of the form $\frac{1}{k}(x,y)\in\RR^2$ with $x,y\in\mathbb{Z}$ and $k=\sqrt{x^2+y^2}\in\mathbb{R}$.
+We say $kv_0$ is an \textbf{odd-odd} vector if $x,y$ are relatively prime and both odd.
+We denote the \textbf{set of all odd-odd directions} $\mathcal{O}$.
+We say that $kv_0$ is an \textbf{even-odd} vector if $x,y$ are relatively prime and of opposite parity.
+We denote the \textbf{set of all even-odd directions} $\mathcal{E}$.
+Denote $\cS=\cO\sqcup\cE$ as the \textbf{collection of all possible rational directions}.
 \end{Def}
 
-Observe that when $r$ acts on the relations $\eqref{eq:rel2}$ we get the following:
+We use $||~.~||:\RR^2\rightarrow\RR_{\geq 0}$ to denote the standard Euclidean norm in $\RR^2$.
 
-\begin{equation}
-\begin{split}
-(e^{(-j)i\frac{\pi}{2}}\zeta(t);j)\sim_{r(\mathbf{P'})}(e^{(-j)i\frac{\pi}{2}}~{\zeta(t)-\aleph};j+1)\\
-(e^{(-j)i\frac{\pi}{2}}\eta(t);j)\sim_{r(\mathbf{P'})}(e^{(-j)i\frac{\pi}{2}}~{\eta(t)-\aleph};j+1)
-\end{split}
-\end{equation}
-
-$\btU$ is then recovered as $r\cdot(\btU_0)=\mathbf{P'}/\sim_{r(\mathbf{P}')}$, where all identifications are translations made on copies of $\mathbf{P}$.
-
-\subsection{Translation Surfaces and $\ZZ^2$-Covers}
-
-\begin{Def}
-Algebraic intersection number is a non-degenerate bilinear form:
-\begin{align*}
-i:H_1(S,\mathbf{R})\times H_1(S,\mathbf{R})\rightarrow\mathbf{R},
-\end{align*}
-where $\mathbf{R}$ is a ring and for $[\gamma],[\beta]\in H_1(S,\mathbf{R})$, $i([\beta],[\gamma])$ returns the intersection number of two homology classes.
-\end{Def}
-Algebraic intersections are signed and follow some convention such as the right-hand rule.
-
-\begin{Def}
-Let $u,v\in H_1(S,\QQ)$ be linearly independent homology classes of curves on $S$. Then the group homomorphism from $\pi_1(S)$ to $\QQ^2$ is given as:
-\begin{align*}
-\Omega_{u,v}:\pi_1(S)\rightarrow \mathbf{R}^2 \text{; } \beta\mapsto(i(u,[\beta]),i(v,[\beta])).
-\end{align*}
-\end{Def}
-
-The set of all orientation-preserving affine diffeomorphisms of $S$ forms the group Aff$^+(S)$.
-The corresponding \emph{Veech group}, $V(S)$ of $S$ is the image of the group morphism $D:\text{Aff}^+(S)\rightarrow SL(2,\mathbb{R})$ that takes an affine map to its derivative.
-A surface is said to be \emph{Veech} if its Veech group is commensurable to $SL(2,\mathbb{R})$.
-It is well known that origami, or square-tiled, surfaces have Veech groups commensurable to $SL(2,\ZZ)$.
-[cite] When $\mathbf{R}=\ZZ$, $\Omega_{u,v}$ takes an element of $\pi_1(S)$ to $\Delta$.
-Thus, $\gamma\in\pi_1(S)$ lifts to $\tilde{\gamma}\in\pi_1(\tilde{S})$ if and only if $\gamma\in\ker\Omega_{u,v}$.
-
-\begin{Def}
-Let $\alpha:[0,1]\rightarrow S$ be a closed, non-singular geodesic path on $S$.
-The holonomy map $\mathbf{hol}:H_1(S, \mathbf{R})\rightarrow\CC$ returns the holonomy vector of a closed path as a difference of the starting and endpoints of a flow by\\
-\begin{align*}
-\mathbf{hol}([\alpha])=\int_{\alpha}dz.
-\end{align*}
-\end{Def}
-Since $\alpha$ is non-singular, it can be mapped to $S^\circ$ which admits a flat holomorphic one-form $dz$.
-Let $\theta=Arg(\mathbf{hol}([\alpha]))$.
-We say that $\phi_t^\theta:\RR\times S^\circ \rightarrow S^\circ$ is the unit-speed geodesic flow on $S^\circ$ in direction $\theta$ given by the  $[\alpha]$ such that $\phi^\theta_0=\alpha(0)$.
-In local coordinates this corresponds to $z+te^{i\theta}\in \CC$.
-
-\begin{lem}
-$\phi_t^\theta$ has a period of $T=|\mathbf{hol}([\alpha])|$
-\begin{proof}
-This just follows from the fact that $\phi_t^\theta$ flows at unit-speed in the direction of $\frac{\mathbf{hol}([\alpha])}{|\mathbf{hol}([\alpha])|}$. 
-\end{proof}
-\end{lem}
-And so the length of a vector determines the period of a flow on $S$. More importantly, we have the following:
-
-\begin{lem}
-Denote the lifted flow of $\phi_t^\theta$ on $\tilde{S}$ by $\tilde{\phi}_t^\theta$.
-Then $\tilde{\phi}_t^\theta$ is periodic on $\tilde{S}$ and $\tilde{S}^\circ$ if and only if $[\alpha]\in\ker\Omega_{u,v}$.
-Further, $\tilde{\phi}_t^\theta$ has period $T=|\mathbf{hol}([\alpha])|$.
-\begin{proof}
-Suppose that $[\alpha]\notin\ker\Omega_{u,v}$.
-Since $\alpha([0,1])=\phi^\theta_{[0,\hol([\alpha])]}$, their homology classes are equivalent.
-Hence $\tilde{\phi}_t^\theta$ could not close on $\tilde{S}$ or $\tilde{S}^\circ$, or else $\Omega_{u,v}(k[\alpha])=k(i(u,[\alpha]),i(v,[\alpha]))=(0,0)$ for some non-zero $k\in\ZZ$.
-Conversely, suppose $[\alpha]\in\ker\Omega_{u,v}$.
-Then
-\end{proof}
-\end{lem}
-
-\begin{cor}
-If $[\alpha]\notin\ker\Omega_{u,v}$, then $\tilde{\phi}_t^\theta$ is drift-periodic with period $T=\hol([\alpha])$.
-\begin{proof}
-This follows immediately from the previous lemma since $\tilde{S},\tilde{S}^\circ$ have translational $\ZZ^2$ symmetries and a well-defined $\ZZ^2$ action on an element in the fiber of a basepoint in $S$ under $f$.
-The period is $T$ since a geodesic closes on $S$ with period $T$.
-\end{proof}
-\end{cor}
-
-\begin{lem}
-Let $h\in\text{Aff}^+(S)$. If $h(\beta)=\alpha$ for closed geodesics $\alpha,\beta$ on $S$ and $\beta$ lifts to a closed path on $\tilde{S}$, then so does $\alpha$. (not sure about this)
-\begin{proof}
-If the premise is true, then $[\beta]\in\ker\Omega_{u,v}$.
-Denote the group automorphism induced by $h$ as $h_*$.
-Then $\Omega_{u,v}([\alpha])=\Omega_{u,v}([h(\beta)])=\Omega_{u,v}(h_*\cdot[\beta])=(i([\beta],h_*^{-1}\cdot u),i([\beta],h_*^{-1}\cdot v))=(0,0)$ as automorphisms .
-\end{proof}
-\end{lem}
-
-\begin{lem}
-Let $\beta=h(\alpha)$ as before. If $D(h)=h'\in V(S)$ and $S$ is Veech, then $\tilde{\phi}^\theta_t$ has period $T=h'\cdot\hol([\beta])$. (also not sure)
-\begin{proof}
-
-\end{proof}
-\end{lem}
-
-\subsection{Symmetries of $\bM$}
-
-\section{Four-fold Cover of $\mathbf{U}$}
-
-\begin{figure}[H]
-\centering
-\includesvg[width=2.05in.]{fourfold}\hspace{0.5in}\includesvg[width=2.05in.]{fourfoldrotate}
-\caption{Four-fold cover isometry and the preimage of a point in $\mathbf U \backslash Sing(\mathbf U)$.}
-\end{figure}
-
-\begin{figure}[H]
-\centering
-\includesvg[width=3.15in]{utilda}
-\label{fig:utilda0}
-\caption{Branched cover associating every direction with one plane.}
-\end{figure}
-
-\begin{figure}[H]
-\centering
-\includesvg[width=3.15in]{utildaprime}
-\caption{Infinite-type translation surface obtained by rotating each copy of the fundamental domain accordingly.}
-\label{fig:utilda}
-\end{figure}
-
-The quotient under the group action of translational symmetries is isomorphic to $\mathbb{Z}^2$ since the orbit of any point in the fundamental domain is a lattice in the space.
-
-\begin{thm}
-The translational symmetries of $\tilde{\mathbf{U}}$'s fundamental domain induce symmetries on the surface isomorphic to $\mathbb{Z}^2$.
-\begin{proof}
-Let $(z;j)\in\mathbb{C}\times\mathbb{Z}/4\mathbb{Z}$ and define the group action $T_0^{m,n}:\mathbb{C}\times\mathbb{Z}/4\mathbb{Z}\rightarrow\mathbb{C}\times\mathbb{Z}/4\mathbb{Z}$ as $T^{m,n}(z;j)=(z+2e^{-j\frac{i\pi}{2}}(m+in);j)$.
-This translation acts faithfully on the preimages of $\mathbf{U}\backslash Sing(\mathbf{U})$, and respects edge identifications of $\mathbf{\tilde{\mathbf{U}}}$, thereby making it an isometry of the surface.
-Consider a group homomorphism, $T_0^{m,n}\mapsto m+in$ onto the plane of Gaussian integers, $\mathbb{Z}[i]$.
-The exponential function is never zero, so the identity of the translation group is $T_0^{0,0}$.
-This is an isomorphism since it is clearly surjective and any non-trivial element of $T^{m,n}$ could not possibly map to the identity element of $\mathbb{Z}[i]$, regardless of the value of j.
-Since $\mathbb{Z}^2$ is isomorphic to $\mathbb{Z}[i]$, it is isomorphic to  $T_0^{m,n}$ as well.
-\end{proof}
-\label{thm:z2}
-\end{thm}
-
-\begin{Def}
-The automorphism $T^{m,n}:\tilde{\mathbf{U}}\rightarrow\tilde{\mathbf{U}}$ is an  \textbf{induced translation} of $\tilde{\mathbf{U}}$ as a result of the previous theorem.
-\end{Def}
-
-This surface is obtained as a ramified cover of the unit square torus.
-It is a translation surface and is therefore equipped with a \textbf{holomorphic one-form}, a collection of charts from neighborhoods of $\mathbf{M}$ to $\mathbb{C}$ such that any neighborhood away from Sing$(\mathbf{M})$ has a \emph{flat} induced Euclidean metric.
-A theorem of Gutkin and Judge tells us that its Veech group is commensurable to SL$(2,\mathbb{Z})$ and is therefore a Veech surface.
-We look at some of its affine maps, and generate a subgroup $\mathbb{X}\subset$ Aff$^+(\mathbf{M})$  by the following transformations:
-
-\begin{enumerate}[label=(\roman*)]
-\item Multi-twists of the surface as global diffeomorphisms given by Dehn-twists of its cylinder decomposition in horizontal and vertical directions with derivatives:
-\begin{equation*}
-\left\{ \left[ \hspace{1mm} \begin{matrix}
-				1 &  \pm 2\\
-				0 & 1
-			\end{matrix}\hspace{1mm}\right] \text{ , }
-			\left[ \hspace{1mm} \begin{matrix}
-							1 & 0\\
-							 \pm 2 & 1
-						\end{matrix}\hspace{1mm}\right] \right\}.
-\end{equation*}
-we call $\mathbf{A}^{\pm 1}$, $\mathbf{B}^{\pm 1}$, respectively.\\
-A Dehn-twist on each cylinder in the cylinder decomposition of $\mathbf{M}$ in horizontal and vertical directions gives way to these global affine diffeomorphisms:
-
-\item Rotation group generated by a $+\frac{\pi}{2}$ rotation of the surface fixed about the center of the second square on the bottom of the staircase, an order four isometry on $\mathbf{M}$ denoted $\mathbf{R}$.
-
-\item  Order 2 translation of the surface that moves the bottom left-most square to the square right next to it, denoted $\mathbf{H}$.
-\item Order 2 translation of the surface that takes the bottom right-most square to the one right above it, denoted $\mathbf{V}$.
-\end{enumerate}
-
-\begin{Def}
-The group $\mathbb{X}$ is the isometry group generated by affine maps $\mathbf{A}$,$\mathbf{B}$,$\mathbf{R}$, $\mathbf{H}$, and $\mathbf{V}$.
-The image of the derivative map on elements in $\mathbb{X}$ is denoted $\mathbb{X}'$ and generated by matrices
-\begin{align*}
-\left[\begin{matrix}
-1 & 2 \\ 0 & 1
-\end{matrix}\right],
-\left[\begin{matrix}
-1 & 0 \\ 2 & 1
-\end{matrix}\right],
-\left[\begin{matrix}
-0 & -1 \\ 1 & 0
-\end{matrix}\right],
-\end{align*}
-denoted $\mathbf{A}'$, $\mathbf{B}'$, and $\mathbf{R}'$ in that order.
-\end{Def}
-
-It is not immediately apparent if these affine maps generate Aff$^{+}(\mathbf{M})$, or if their derivatives generate $V(\mathbf{M})$, its Veech group.
-We use these to induce homomorphisms on $H_1(X, \mathbb Q)$.
-A spanning set of $H_1(\mathbf{M}, \mathbb Q)$ is obtained as the set of homology classes of the core curves of X's cylinder decompositions in both vertical and horizontal directions:
-
-\begin{figure}[H]
-\includegraphics[width=3in]{homologyclass.png}
-\centering
-\caption{Cylinder core curves with u,v, and z homology classes that determines the $\mathbf{Z}^{2}$-cover.}
-\label{fig:homology}
-\end{figure}
-
-\begin{Def}
-The set of abelianized cylinder core curves is denoted as $\Gamma=\{\gamma_i: i = 0,\dots,11\}\subset H_1(\mathbf{M},\mathbb{Q})$.
-\end{Def}
-
-\begin{rem}
-We use 12 elements to span homology, although a basis requires only 10.
-It's not impossible to determine the relations between these core curve classes, but it is not necessary.
-A $12\times12$ matrix of these core curve cylinder decompositions to their intersection numbers with adjacent curves is rank 10, as to be expected.
-\end{rem}
-
-The induced homomorphisms of $H_1(\mathbf{M}, \mathbb Q)$ have come from affine maps that have various effects on these core-curves.
-We use a 12-gon to represent the set of curves, and show how these elements act on them.
-The multi-twists add curves to adjacent curves, and the translation maps permute them.
-The reader is encouraged to check these for themselves.
-
-e.g. for $\mathbf{H,V}\in$Aff$^+(\mathbf{M})$,
-
-\begin{minipage}{0.5\textwidth}
-\vspace{0.3in}
-\begin{itemize}
-\item[\textbf{\emph{$\mathbf{H}$ \& $\mathbf{V}$}}] The effect that these two translations have on the 12-gon is a reflection about these lines.
-Observed by keeping track of the squares and core curves after $\mathbf{H}$ and $\mathbf{V}$ have acted on $\mathbf{X}$.
-\end{itemize}
-\end{minipage}
-\begin{minipage}{0.7\textwidth}
-\begin{figure}[H]
-\hspace{0.2in}\includegraphics[width=1.7in]{12gonHV.pdf}
-\end{figure}
-\end{minipage}
-
-\begin{Def}
-The induced homomorphisms of $H_1(\mathbf{M},\mathbb{Q})$ are obtained from the affine subgroup $\mathbb{X}$ and denoted $\mathbb{X}_*$.
-The associated homomorphisms on the spanning set $\Gamma$ are given as:
-\begin{align*}
-\mathbf{A}^k_*\circ[\gamma_i]=&[\gamma_i] + \frac{k}{2}(1-(-1)^i)([\gamma_{i-1}]+[\gamma_{i+1}])\\
-\mathbf{B}^k_*\circ[\gamma_i]=&[\gamma_i] + \frac{k}{2}(1+(-1)^i)([\gamma_{i-1}]+[\gamma_{i+1}])\\
-\mathbf{R}_*\circ[\gamma_i]=&(-1)^i[\gamma_{1-i\text{ mod }12}]\\
-\mathbf{H}_*\circ[\gamma_i]=&[\gamma_{12-i\text{ mod }12}]\\
-\mathbf{V}_*\circ[\gamma_i]=&[\gamma_{10-i\text{ mod }12}]
-\end{align*}
-\end{Def}
-
-\begin{Def}
-The homology classes $u,v,z$ are given as the following sums of core curves:
-\begin{align*}
-[u] &= -[\gamma_2] +[\gamma_5] + [\gamma_8] - [\gamma_{11}],\\
-[v] &= +[\gamma_0] -[\gamma_3] -[\gamma_6] +[\gamma_9],\\
-[z] &= +[\gamma_1] +[\gamma_4]-[\gamma_7]-[\gamma_{10}].
-\end{align*}
-\end{Def}
-
-\begin{thm}
-The fundamental group of the $\mathbb{Z}^2$-cover is obtained by lifting the kernel of the closed paths of $\mathbf{M}$ of the homomorphism:
-
-\begin{align*}
-\Omega_{u,v}:\pi_1(\mathbf{M},x_0)\rightarrow \mathbb{Z}^2 \text{; } \beta\mapsto(i(u,[\beta]),i(v,[\beta])),\text{ where}
-\end{align*}
-
-\begin{align*}
-i:H_1(\mathbf{M},\mathbb Q)\times H_1(\mathbf{M},\mathbb Q)\rightarrow \mathbb Z.
-\end{align*}is the intersection number of two homology classes.
-\begin{proof}
-We know from Theorem $\ref{thm:z2}$ that the translational symmetries of $\tilde{\mathbf{U}}$ induced by $T^{m,n}$ is isometric to $\mathbb{Z}^2$.
-Since $\mathbf{M}$ is a genus 5 base surface, we know that $\pi_1(\mathbf{M},x_0)\simeq \mathbb{Z}^{10}$, and the associated cover satisfies $\mathbf{M}=\tilde{\mathbf{U}}/(\pi_1(\mathbf{M},x_0)/N)$, such that $N$ is a normal subgroup of $\pi_1(\mathbf{M},x_0)$.
-This means that $N\simeq \mathbb{Z}^8$.
-The eight core curve classes are the abelianized forms of  $\gamma_0,\gamma_2,\gamma_3,\gamma_5,\gamma_6,\gamma_8,\gamma_9,$ and $\gamma_{11}$ that span N.
-The classes and their signs are obtained from Figure $\ref{fig:mtilda}$ as the outer regions identified by the translations of $T^{m,n}$.
-Thus any closed path on $\mathbf{M}$ is lifted to a closed path on the cover under the quotient map only when a path has a trivial intersection number with the classes.
-\end{proof}
-
-\end{thm}
-Two paths are homologous if they return the same intersection number with the classes of closed core cylinder curves of $\mathbf{U}$ that span its homology.
-The classes u and v are obtained from the group group action of $T^{m,n}$ on the cover.
-
-\begin{Def}
-$\mathbf{hol}:\mathbf{M}\backslash\text{Sing}(\mathbf{M})\rightarrow\mathbb{C}$ is the holonomy vector pulled back from a non-singular path $\gamma$ in $\mathbf{M}$ onto the complex plane given by $\mathbf{hol}(\gamma)=\int_{\gamma}dz$.
-\end{Def}
-
-We denote the \textbf{closed path} $\alpha$, such that $\mathbf{hol}(\alpha)=6+6i$, and show it is homologous to the closed geodesic with the same holonomy vector.
-The slope one direction also decomposes $\mathbf{M}$ into two cylinders by a series of saddle connections of length $\sqrt{2}$ between singularities:
-
-\begin{figure}[H]
-\centering
-\includesvg[width=4.4in.]{slopeonecylinder}
-\caption{The two right-most cylinders $C_1$ (labeled) and $C_2$(unlabeled).}
-\end{figure}
-
-The circumferences of these two cylinders are $6\sqrt{2}$. Geodesic flows on this surface are well defined, and rational directions
-
-\begin{Def}
-Let $\omega^{\theta}_t:[0,1]\times\mathbb{R}/2\pi\mathbb{Z}\rightarrow\mathbf{M}$ be the \textbf{maximal geodesic flow} on the surface in direction $\theta$ such that $\omega^{\theta}_0=\omega^{\theta}_1=x_0\in\mathbf{M}\backslash\text{Sing}(\mathbf{M})$.\\
-$\omega^{\frac{\pi}{4}}_t$ is the geodesic flow in the \textbf{slope one direction}, and $\chi$ is its image in $\mathbf{M}$ and element of $\pi_1(\mathbf{M},x_0)$.
-\end{Def}
-
-\begin{lem}
-$\alpha$ is homologous to $\chi$ in  $\mathbf{M}\backslash\text{Sing}(\mathbf{M})$.
-\begin{proof}
-Let $\chi$ be a geodesic contained in either $C_1$ or $C_2$.
-Since a geodesic does not admit singularities, it is the image of a closed path on $X\backslash\text{Sing}(X)$ with initial point $x_0$ on the strips of $C_1$ and $C_2$ with boundaries removed, denoted $C_1',$ $C_2'$.
-Express $[\alpha]$ as $\Sigma^{11}_{j=0}$ $\frac{1}{2}\gamma_j$ (a closed path climbing up the staircase).
-We show that the intersection numbers of $[\alpha]$ and $[\chi]$ are the same for every core cylinder curve $\gamma$, i.e.
-$i([\gamma_k], \Sigma^{11}_{j=0} \text{ } \frac{1}{2}[\gamma_j])=i([\gamma_k],[\chi])\text{ } \forall k=0,\dots,11$.\\
-\textbf{Case one:} k is even.
-If k is even, then every curve $\gamma_k$ is oriented to the right.
-Since $\chi$ intersects every curve once, $i([\gamma_k],[\chi])=1$.
-No even indexed curves intersect eachother, so we need only consider when $j$ is odd.
-Now if $j$ is odd, it is incident (positively crossing) with only two horizontal curves, namely $\gamma_{j+1},\gamma_{j-1}$.
-Therefore $i([\gamma_k],[\alpha])=i([\gamma_{j-1}],\frac{1}{2}[\gamma_{j}])+i([\gamma_{j+1}],\frac{1}{2}[\gamma_{j}])=\frac{1}{2}(i([\gamma_{j-1}],[\gamma_{j}])+i([\gamma_{j+1}],[\gamma_{j}]))=\frac{1}{2}(1+1)=1.$\\
-\textbf{Case two:} k is odd.
-If k is odd, then $[\chi]$ will have an intersection number of $-1$ with $[\gamma_k]$ since odd-indexed core curves are oriented upwards.
-Now since k is odd, we only consider when $j$ is even.
-Similarly, this means that $\gamma_j$ negatively intersects the two vertical core curves with adjacent indices.
-Hence, $i([\gamma_k],[\alpha])=i([\gamma_{j-1}],\frac{1}{2}[\gamma_{j}])+i([\gamma_{j+1}],\frac{1}{2}[\gamma_{j}])=\frac{1}{2}(i([\gamma_{j-1}],[\gamma_{j}])+i([\gamma_{j+1}],[\gamma_{j}]))=\frac{1}{2}(-1-1)=-1.$\\
-We know intersection number to be bilinear and non-degenerate on homology.
-So if $\alpha$ and $\chi$'s abelianizations admit the same intersection numbers for every curve in the spanning set of $H_1(\mathbf{M}, \mathbb{Q})$, then $[\alpha]=[\chi]$.
-\end{proof}
-\end{lem}
-
-\begin{thm}
-$\chi\in\pi_1(\mathbf{M},x_0)$ lifts to $\tilde{\chi}\in\pi_1(\tilde{\mathbf{U}},\tilde{x}_0)$
-\begin{proof}
-From Lemma 2, $[\chi]=[\alpha]$, so $\Omega_{u,v}(\chi)=\Omega_{u,v}(\alpha)$. Since
-$i([u],[\alpha])=-i([\gamma_2],[\alpha] +i([\gamma_5],[\alpha]) + i([\gamma_8],[\alpha]) - i([\gamma_{11}],[\alpha])=-1+(-1)+1-(-1)=0$ and $i([v],[\alpha])=1-(-1)-1+(-1)=0$, it follows that $\alpha,\chi\in\text{Ker }\Omega_{u,v}$, and $\chi$ lifts to a closed geodesic on $\tilde{\mathbf{U}}$.
-\end{proof}
-\end{thm}
-
-\begin{cor}
-content...
-\end{cor}
-
-From here, we use $\alpha$ to show that the \emph{only} trajectories that close on the Necker cube surface are those that are in vector direction $(a,b)$ such that $\gcd(a,b)=1$ and $a,b$ are both odd.
-We call these \textbf{odd-odd} directions.
-We can make this claim because the group generated by the matrices
-
-\begin{equation*}
-\left[ \hspace{1mm} \begin{matrix}
-				1 &  \pm 2\\
-				0 & 1
-			\end{matrix}\hspace{1mm}\right] \text{ and }
-			\left[ \hspace{1mm} \begin{matrix}
-							1 & 0\\
-							 \pm 2 & 1
-						\end{matrix}\hspace{1mm}\right]
-\end{equation*}
-is the \emph{Sanov subgroup} of $SL(2,\mathbb{Z})$ and only sends elements in the odd-odd set to itself.
-There are dualizations made between how these matrices skew a geodesic direction,  and how their original affine transformations induce an effect homology.
-In a sense the kernel is obtained by the orbit of $\chi$ under $\mathbb{X}$ and its holonomy vector under $\mathbb{X}'$.
-
-\begin{lem}
-The actions of $\mathbb{X}'$ on $\mathcal{O}$ and $\mathcal{E}$ are closed in their respective sets.
-\begin{proof}
-Since $\mathbb{X}'$ is generated by the elements $\mathbf{A}'$, $\mathbf{B}'$, and $\mathbf{R}'$, any matrix $G'\in\mathbb{X}'$ is of the form $G' = (\mathbf{A}')^{1_1}\circ(\mathbf{B}')^{1_2}\circ(\mathbf{R}')^{1_3}\circ(\mathbf{A}')^{2_1}\circ\dots(\mathbf{A}')^{n_1}\circ(\mathbf{B}')^{n_2}\circ(\mathbf{R}')^{n_3}$, where $i_k\in\mathbb{Z}$ for $i=1,\dots,n$ and $k=1,2,3.$ Let $x=\left(\begin{matrix}p \\ q  \end{matrix}\right),y\in\mathcal{O}$, and consider the equation $G'x=y$.
-Observe that $\left(\begin{matrix}1 && 2 \\ 0 && 1\end{matrix}\right)^l x=\left(\begin{matrix}p+2jq \\ q  \end{matrix}\right),$ and $ \left(\begin{matrix}1 && 0 \\ 2 && 1\end{matrix}\right)^m x=\left(\begin{matrix}p \\ q+2mp  \end{matrix}\right)$ for any $l,m\in\mathbb{Z}$.
-Also note that for any $j\in\mathbb{Z}$, $(\mathbf{R}')^{m}x=\left(\begin{matrix}p \\ q  \end{matrix}\right),\left(\begin{matrix}-q\\ p  \end{matrix}\right),\left(\begin{matrix}-p \\ -q  \end{matrix}\right),\left(\begin{matrix}q \\ -p  \end{matrix}\right)$ when $j \mod{4}\equiv0,1,2,3$, respectively.
-In any case, the product of any power of a generator of $\mathbb{X}'$ and any $x\in\mathcal{O}$ is an element of $\mathcal{O}$ .
-By letting $l=i_1,m=i_2,$ and $j=i_3$, we first consider the base case when $i=n$.
-Let $G'=G'_1\circ\dots\circ G'_n$, such that $G'_i=(\mathbf{A}')^{i_1}\circ(\mathbf{B}')^{i_2}\circ(\mathbf{R}')^{i_3}$.
-Since $n_1,n_2,n_3$ are arbitrary integers, $G'_n x \in\mathcal{O}$.
-Suppose for some $b< n-1$, $G'_{n-b}\circ\dots\circ G'_n x =y'\in\mathcal{O}$.
-Therefore $y'=(G'_1\circ\dots\circ G'_{b})^{-1}y$, which implies that $(G'_1\circ\dots\circ G'_{b})^{-1}$ preserves the set $\mathcal{O}$.
-Otherwise, if $y\in\mathcal{E}$, there exists at least one $G'_i$ for $1< i< b$ and $\tau\in\mathcal{E}$ such that $G'^{-1}_i \tau = (\mathbf{R}')^{-i_3}\circ(\mathbf{B}')^{-i_2}\circ(\mathbf{A}')^{-i_1} \tau \in \mathcal{O}$, a contradiction.
-Since elements in $\mathbb{X}'$ are invertible, $G'_1\circ\dots\circ G'_{b}$ must also map $\mathcal{O}$ to itself.
-Left multiply both sides of the equation to show that $G'_1\circ\dots\circ G'_n x = G' x = y$.
-By the principle of strong induction, this holds for all $0<b\leq n$.
-Since $G'$ is invertible and an arbitrarily chosen element of $\mathbb{X}'$, it follows that $x\in\mathcal{O}$ if and only if $y\in\mathcal{O}$ and $\mathcal{O}$ is closed under $\mathbb{X}'$.\\
-The proof for when $x\in\mathcal{E}$ is made in the same way.
-\end{proof}
-\end{lem}
-
-Now a trajectory in the horizontal direction has a directional vector of (1,0).
-The orbit of this vector by the Veech group is the set of all \textbf{even-odd} vectors.
-We also know that in this direction a geodesic is drift-periodic (See figure 1).
-The Veech group of $\mathbf{M}$ preserves these properties.
-Suppose you had some closed geodesic on $\mathbf{M}\backslash Sing(\mathbf{M})$ called $\beta$ such that $\beta=h(\alpha)$, where $h\in$  Aff$^+(\mathbf{M})$, and $h_*$ is its induced homomorphism.
-Then we want to show that
-
-\begin{align*}
-(i([\beta],[u]), i([\beta],[v]))=(i([\alpha],h^{-1}_*[u]), i([\alpha],h^{-1}_*[v])) = (0,0).
-\end{align*}
-
-But first, we look at some of the properties of the group $\mathbb{X}_*$.
-
-\begin{thm}
-Let $\mathbb{X}_*$ be the group generated by $\mathbf{A}_*, \mathbf{B}_*, \mathbf{R}_*, \mathbf{H}_*,$ and $\mathbf{V}_*$.
-Let $G=\left< \mathbf{A}_*, \mathbf{B}_* \right>$, $T=\left< \mathbf{H}_*, \mathbf{V}_* \right>$, and $R=\left< \mathbf{R}_*\right>$.
+\begin{thm*} \textbf{Periodic and Drift-Periodic Geodesics on the Necker cube surface.} 
+\\Let $\Phi_t:\RR\rightarrow\bUs$ be a non-singular unit-speed geodesic on the Necker cube surface such that $\Phi(0)=u_0$ and $\Phi'(0)=v\in\RR^3$.
+Let $v_0\in\RR^2$ be a distance-preserving projection of $v$ onto the plane tangent to the surface at $u_0$.
+Let $k\in\RR$ such that $kv_0\in\cS$.
 Then the following is true:
 \begin{enumerate}[label=(\roman*)]
-\item $G$ is a free subgroup of $\mathbb{X}_*$ of rank two.
-\item $T$ is a finite cyclic subgroup of $\mathbb{X}_*$ and a centralizer of G.
-\item $R$ is a finite cyclic subgroup of $\mathbb{X}_*$, and a normalizer of $G$.
+\item $\Phi$ is periodic if $kv_0\in\cO$ with period $T=6||kv_0||$.
+\item $\Phi$ is drift-periodic if $v_0\in\cE$ with period $T=2||kv_0||$.
 \end{enumerate}
-\begin{proof} Let $h_*^{j}=\mathbf{A}_*^{k_j}\circ \mathbf{B}_*^{g_j}\in G$ for $k_j, g_j \in\mathbb{Z}$, $j=1,\dots,n$.\\ \emph{(i)}.
-When $\mathbf{A}_*$ and $\mathbf{B}_*$ act on $\gamma_i$, it is only ever trivial if i is even for $\mathbf{A}_*$ or i is odd on $\mathbf{B}_*$.
-Since i cannot be both odd and even at the same time, there is no relation between the two generators and therefore $G$ is free.\\
-\emph{(ii)} It is up to the reader to show that $T$ has the relations $\mathbf{H}_*^2=\mathbf{V}_*^2=(\mathbf{H}_*\mathbf{V}_*)^3=id_*$, and is isomorphic to the rotational group of the hexagon generated by reflections about adjacent vertices of a 12-gon.
-Observe that $\mathbf{H}_*\circ \mathbf{A}_*^{k_j}\circ[\gamma_i]=\mathbf{H}_*\circ[\gamma_i] + \frac{k_j}{2}(1-(-1)^i)(\mathbf{H}_*\circ[\gamma_{i-1}]+\mathbf{H}_*\circ[\gamma_{i+1}])=[\gamma_{-i}] + \frac{k_j}{2}(1-(-1)^i)([\gamma_{1-i}]+[\gamma_{-i-1}])=\mathbf{A}^{k_j}\circ[\gamma_-i]=\mathbf{A}^{k_j}\circ\mathbf{H}_*\circ[\gamma_i],$ and $\mathbf{V}_*\circ \mathbf{A}_*^{k_j}\circ[\gamma_i]=\mathbf{V}_*\circ[\gamma_i] + \frac{k_j}{2}(1-(-1)^i)(\mathbf{V}_*\circ[\gamma_{i-1}]+\mathbf{V}_*\circ[\gamma_{i+1}])=[\gamma_{10-i}] + \frac{k_j}{2}(1-(-1)^i)([\gamma_{11-i}]+[\gamma_{9-i}])=\mathbf{A}^{k_j}\circ[\gamma_10-i]=\mathbf{A}^{k_j}\circ\mathbf{V}_*\circ[\gamma_i].$ In the same way one can show this to be true for $\mathbf{B}_*^{g_j}$, and we can see that $T$ is a centralizer of $G$.
-\\
-\emph{(iii)} $R$ is obviously cyclic and finite since an isomorphism is obtained as $\mathbf{R}_*\mapsto\mathbf{R}'\in SO(2,\mathbb{Z})$.
-\\Note that $\mathbf{R}_*\circ\mathbf{A}_*^{k_j}\circ[\gamma_i]=\mathbf{R}_*\circ[\gamma_i] + \frac{k_j}{2}(1-(-1)^i)(\mathbf{R}_*\circ[\gamma_{i-1}]+\mathbf{R}_*\circ[\gamma_{i+1}])\\
-=(-1)^i[\gamma_{1-i}] + \frac{k_j}{2}(1-(-1)^i)((-1)^{i-1}[\gamma_{2-i}]+(-1)^{i+1}[\gamma_{-i}])\\
-=(-1)^{1-i}([\gamma_{1-i}] - \frac{k_j}{2}(1+(-1)^{1-i})([\gamma_{2-i}]+[\gamma_{-i}]))\\
-=(-1)^{1-i}\mathbf{B}_*^{-k_j}\circ[\gamma_{1-i}]=\mathbf{B}_*^{-k_j}\circ(-1)^{1-i}[\gamma_{1-i}]=\mathbf{B}_*^{-k_j}\circ\mathbf{R}_*\circ[\gamma_{i}].$\\
-Likewise, $\mathbf{R}_*\circ\mathbf{B}_*^{g_j}\circ[\gamma_{i}]=\mathbf{A}_*^{-g_j}\circ\mathbf{R}_*\circ[\gamma_{i}]$.
-\end{proof}
-\end{thm}
-
-\begin{rem}
-It can be easily shown that $\mathbb{X}'$ has similar properties.
-\end{rem}
-
-\begin{lem}
-$Let$ $h_*\in \left<\mathbf{A}_*,\mathbf{B}_*\right>.
-$ Then $h_*\circ[\alpha]\in H_1(\mathbf{M},\mathbb{Q})$ can be expressed as $h_*\circ[\alpha]=\frac{1}{2}(c_1\Sigma^{5}_{j=0}[\gamma_{2j}]+c_2\Sigma^{5}_{j=0}[\gamma_{2j+1}])$ for $c_1,c_2\in\mathbb{Z}$.
-\begin{proof}
-Let $\Sigma^{5}_{j=0}[\gamma_{2j}]=\Sigma\Gamma_{even}$, $\Sigma^{5}_{j=0}[\gamma_{2j+1}]= \Sigma\Gamma_{odd}$, and  $\Sigma^{11}_{j=0}[\gamma_{j}]= \Sigma\Gamma$.
-Let $h_*=h_*^n\circ\dots\circ h_ *^1$, and $h_*^{i}=\mathbf{A}_*^{k_i}\circ \mathbf{B}_*^{g_i}$ for $k_i, g_i \in\mathbb{Z}$, $i=1,\dots,n$.
-Compose these two homomorphisms and obtain $\mathbf{A}_*^{k_i}\circ \mathbf{B}_*^{g_i}(\Sigma\Gamma)=(4g_ik_i+2k_i)\Sigma\Gamma_{even}+2g_i\Sigma\Gamma_{odd}+\Sigma\Gamma$.
-Let $c_i^1=(4g_ik_i+2k_i), c_i^2=2g_i$, and solve for $h_*^{i+1}\circ h_*^i\circ\Sigma\Gamma$:
-\begin{align*}
-h_*^{i+1}\circ h_*^{i}\circ(\Sigma\Gamma)=h_*^{i+1}\circ(c_i^1\Sigma\Gamma_{even}+c_i^2\Sigma\Gamma_{odd}+\Sigma\Gamma)\\ =c_i^{1}h_*^{i+1}\circ(\Sigma\Gamma_{even})+c_i^2h_*^{i+1}\circ(\Sigma\Gamma_{odd})+h_*^{i+1}\circ(\Sigma\Gamma)\\ =2g_{i+1}\Sigma\Gamma_{odd}+(4g_{i+1}k_{i+1}+2k_{i+1})\Sigma\Gamma_{even}+\Sigma\Gamma\\+c_i^{1}(4g_{i+1}k_{i+1}\Sigma\Gamma_{even}+2g_{i+1}\Sigma\Gamma_{odd}+\Sigma\Gamma_{even})\\+c_{i}^2(2k_{i+1}\Sigma\Gamma_{even}+\Sigma\Gamma_{odd})\\
-=\Sigma\Gamma+(c_i^1+(c_{i}^1+1)(4g_{i+1}k_{i+1})+(c_i^2+1)2k_{i+1})\Sigma\Gamma_{even}\\+(c_i^2+(c_i^1+1)2g_{i+1})\Sigma\Gamma_{odd}\\
-\text{Let }c_{i+1}^1:=(c_i^1+(c_{i}^1+1)(4g_{i+1}k_{i+1})+(c_i^2+1)2k_{i+1}),\\
-c_{i+1}^2:=(c_i^2+(c_i^1+1)2g_{i+1}).
-\end{align*}
-From these recursive definitions and a finite sequence of integers, $\{k\}_i,\{g\}_i$, observe then that\\ $h_*\circ[\alpha]=h_*\circ[\frac{1}{2}\Sigma\Gamma]=\frac{1}{2}h_*\circ[\Sigma\Gamma]=\frac{1}{2}[c_n^1\Sigma\Gamma_{even}+c_n^2\Sigma\Gamma_{odd}+\Sigma\Gamma]\\
-=\frac{1}{2}[(c_n^1+1)\Sigma\Gamma_{even}+(c_n^2+1)\Sigma\Gamma_{odd}]$. Further simplify by letting $c_1=c_n^1+1,c_2=c_n^2+1$.
-\end{proof}
-\end{lem}
-
-\begin{lem}
-Let $h_*\circ[\alpha]\in H_1(\mathbf{M},\mathbb{Q})$. Then for $a\in\left<\mathbf{H}_*, \mathbf{V}_*\right>$ and $b\in\left<\mathbf{R}_*\right>$, the following is true:
-\begin{align*}
-& a\circ h_*\circ[\alpha]=h_*\circ[\alpha]\\
-& b\circ h_*\circ[\alpha]=\frac{1}{2}[c'_1\Sigma\Gamma_{even}+c'_2\Sigma\Gamma_{odd}]\\
-& h_*\circ b\circ[\alpha]=\frac{1}{2}[c''_1\Sigma\Gamma_{even}+c''_2\Sigma\Gamma_{odd}]
-\end{align*}
-\begin{proof}
-By Theorem 4, a is a centralizer of the group so $a\circ h_*\circ[\alpha]=h_*\circ a \circ[\alpha]=h_*\circ \frac{1}{2}a \circ[\Sigma\Gamma].$ Since a is a cyclic permutation of the set $\Gamma$, it acts trivially on $\Sigma\Gamma$.
-Therefore, $a\circ h_*\circ[\alpha]=h_*\circ \frac{1}{2}[\Sigma\Gamma]=h_* \circ[\alpha]$.\\
-By theorem 4, $a\circ\mathbf{A}_*^{k_i}\circ\mathbf{B}_*^{g_i}=\mathbf{B}_*^{-k_i}\circ\mathbf{A}_*^{-g_i}\circ a$.
-Extend this property to $h_*$, and denote the normalized element as $h_{**}$, such that $b\circ h_{*}=h_{**}\circ b$.
-Note that $b(\Sigma\Gamma)=b(\Sigma\Gamma_{even}+\Sigma\Gamma_{odd})=\Sigma\Gamma_{odd}-\Sigma\Gamma_{even}$.
-$b\circ h_*\circ[\Sigma\Gamma]=c_1b\circ\Sigma\Gamma_{even}+c_2b\circ\Sigma\Gamma_{odd}=c_1\Sigma\Gamma_{odd}-c_2\Sigma\Gamma_{even}.$ So, $c_1'=-c_2$ and $c_2'=c_1$.\\
-Since $h_*$ is arbitrary, let $h_{**}=g_{*}$ be generated by an integer sequence that defines the word and consider $h_*\circ b\circ[\Sigma\Gamma]=b\circ g_{*}\circ[\Sigma\Gamma]=c^*_1b\circ\Sigma\Gamma_{even}+c^*_2b\circ\Sigma\Gamma_{odd}=c^*_1\Sigma\Gamma_{odd}-c_2^*\Sigma\Gamma_{even}.$ So, $c_1''=-c_2^*$ and $c_2''=c_1^*$.
-\end{proof}
-\end{lem}
-
-\noindent Now that every element in the orbit of $[\alpha]$ can be expressed as a linear combination of integers, it is simple to show they lift to a closed trajectory in the cover.
-
-\begin{Def}
-Let $\mathbf{dir}:UT(\mathbf{M}\backslash\text{Sing}(\mathbf{M}))\rightarrow\mathcal{O}\cup\mathcal{E}$ be the injective map from $\mathbb{R}/2\pi\mathbb{Z}$ to $\mathbb{Z}^2$ given as $\mathbf{dir}(\theta)=(k_1\cos(\theta),k_2\sin(\theta)), k_1,k_2\in\mathbb{R}$ such that $\gcd(k_1\cos(\theta),k_2\sin(\theta))=1$.
-\end{Def}
-
-\begin{thm}(Sketch)\\
-Any geodesic, $\beta$, in $\mathbf{M}$ lifts to a closed geodesic $\tilde{\beta}$ on $\tilde{\mathbf{U}}$ if and only if $\mathbf{dir}(Arg(\mathbf{hol}(\beta)))\in\mathcal{O}$.
-\begin{proof}
-Call the quotient cover $p:\tilde{\mathbf{U}}\rightarrow\mathbf{M}$, and fix a point $\tilde{x_0}\in p^{-1}(x_0)$.
-Let $\beta=h(\chi)$, where $h\in\mathbb{X}$.
-We also obtain $[\beta]=h_*\circ[\alpha]$ from Lemma 2.
-Since $h$ sends geodesics to geodesics, h induces the following: $\mathbf{hol}(h(\chi))=h'(\mathbf{hol}(\chi))=h'(6+6i)$ for $h'=\left[\begin{matrix} a & b \\ c & d\end{matrix}\right]\in\mathbb{X}'$.
-So, Arg$(h'(\mathbf{hol}(\chi)))=$Arg$(6[(a+b)+i(c+d)])=$Arg$(6h'(1+i))$.
-Lemma 3 states that for any $h'\in\mathbb{X}'$, $h'(\mathcal{O})=\mathcal{O}$.
-Therefore there is no such geodesic of \textbf{even-odd} slope in the orbit of $\chi$.
-Otherwise $h',h\notin\mathbb{X}',\mathbb{X}$.
-Consequently, $\mathbf{dir}(Arg(\mathbf{hol}(\beta)))\in\mathcal{O}$.\\
-From Lemma 5 we see that $[h(\chi)]=h_*\circ[\chi]=h_*\circ[\alpha]=\frac{1}{2}(c_1\Sigma^{5}_{j=0}[\gamma_{2j}]+c_2\Sigma^{5}_{j=0}[\gamma_{2j+1}])$ for $c_1,c_2\in\mathbb{Z}$.
-Denote the sums as $\Sigma\Gamma_{even}$ and $\Sigma\Gamma_{odd}.$\\
-Therefore, $2i([u],h_*\circ[\alpha])=c_1i([u],\Sigma\Gamma_{even})+c_2i([u],\Sigma\Gamma_{odd})\\
-=c_1(-i([\gamma_2],0) +i([\gamma_5],[\gamma_6]+[\gamma_4]) + i([\gamma_8],0) - i([\gamma_{11}],[\gamma_{10}]+[\gamma_{0}]))\\
-+c_2(-i([\gamma_2],[\gamma_1]+[\gamma_3]) +i([\gamma_5],0) + i([\gamma_8],[\gamma_7]+[\gamma_9]) - i([\gamma_{11}],0))\\
-=c_1(-(0)+(-1-1)+(0)-(-1-1))+c_2(-(1+1)+(0)+(1+1)-(0))=0.$\\\vspace{0.1in}\\
-Similarly, $2i([v],h_*\circ[\alpha])=c_1i([v],\Sigma\Gamma_{even})+c_2i([v],\Sigma\Gamma_{odd})\\
-=c_1(i([\gamma_0],0)-i([\gamma_3],[\gamma_2]+[\gamma_4])-i([\gamma_6],0)+i([\gamma_9],[\gamma_8]+[\gamma_{10}])\\
-+c_2(i([\gamma_0],[\gamma_{11}]+[\gamma_1])-i([\gamma_3],0)-i([\gamma_6],[\gamma_5]+[\gamma_7])+i([\gamma_9],0)\\
-=c_1((0)-(-2)-(0)+(-2))+c_2((2)-(0)-(2)+(0))=0.$\\
-Therefore, $\Omega_{u,v}(h(\chi))=(0,0)$, and $h(\chi)=\beta\in\text{Ker }\Omega_{u,v}$ for all $h\in\mathbb{X}$.
-By Theorem 2, $\beta$ lifts to $\tilde{\beta}\in\pi_1(\tilde{\mathbf{U}},\tilde{x}_0)$.
-Let $\theta=$Arg$(\mathbf{hol}(\beta))$.
-Then $\omega_t^{\theta}$ at $x_0$ lifts to $\tilde{\omega}_t^{p^{-1}(\theta)}\in\tilde{\mathbf{U}}\backslash\text{Sing}(\tilde{\mathbf{U}})$.\\
-Now suppose instead that $\beta=h(\gamma_{i})$.
-Then $\mathbf{dir}(\beta)=\frac{1}{2}(1+(-1)^i,1-(-1)^i)$.
-According to Lemma 3, $h'(\mathcal{E})=\mathcal{E}$.
-Thus we have no geodesic in the \textbf{odd-odd} directions obtained from the orbits of $(1,0)$ and $(0,1)$.
-For contradiction, suppose that $h(\gamma_{i})\in\textbf{Ker }\Omega{u,v}$.
-Then $(i(h_*\circ[\gamma_i], [u]),i(h_*\circ[\gamma_i], [v]))=(i([\gamma_i], h^{-1}_*\circ[u]),i([\gamma_i], h^{-1}_*\circ[v]))=(0,0).$ Let $h^{-1}_*\circ[u]=\Sigma^{11}_{j=0}x_j[\gamma_j]$, and $h^{-1}_*\circ[v]=\Sigma^{11}_{j=0}y_j[\gamma_j]$.
-Note that since $\gamma_i$ intersects $\gamma_{i\pm1}$, $i([\gamma_i], h^{-1}_*\circ[u])=(-1)^{i+1}(x_{i-1}+x_{i+1})$ and $i([\gamma_i], h^{-1}_*\circ[v])=(-1)^{i+1}(y_{i-1}+y_{i+1}).$
-\\
-$\mathbf{Unfinished.}.$
-\end{proof}
-\end{thm}
-
-\begin{conj}{\textbf{Dynamics of Geodesic Flow on the Necker cube surface.}}\\ Obtain $\theta$ and $\vec{\theta}$ as described in Definition 3.
-Denote the non-singular unit-speed geodesic flow with initial point $s\in(\mathbf{U}\backslash Sing(U))$ in direction $[\theta]\sim\phi\in UT(\mathbf{U}\backslash Sing(\mathbf U))$ by $F_t^{\phi}:\mathbf{U}\times\mathbb{R}_0^+\rightarrow\mathbf{U}$ on $(\mathbf{U},\mu)$, where $\mu$ is a flow-invariant measure.
-Then the following is true:
-\begin{enumerate}[label=(\roman*)]
-\item (Periodic) There exists a $t_0 > 0$ such that $F^{\phi}_{t+t_{0}}(s)=F^{\phi}_{t}(s)$ if and only if $\vec{\theta}\in\mathcal{O}$.
-\item (Drift-Periodic) There exists a $t_0 > 0$ such that $F^{\phi}_{t+t_{0}}(s)= F^{\phi}_{t}(s)+c$, where $c\in \mathbf{U}$ is a non-trivial translation of a point in $\mathbf{U}$, if and only if $\vec{\theta}\in\mathcal{E}$.
-\end{enumerate}
-\begin{proof}(Sketch)\\
-Denote the covering maps $f:\tilde{\mathbf{U}}\rightarrow\mathbf{U}$, $p:\tilde{\mathbf{U}}\rightarrow\mathbf{M}$, and fix a point $\tilde{x}_0\in f^{-1}(s),p^{-1}(x_0),$ for $x_0\in\mathbf{M}$.
-$f^{-1}([\theta])=\{x:x=\theta+n\frac{\pi}{2}, n\in\mathbb{Z}\}=[\theta]\subset UT(\tilde{\mathbf{U}}\backslash\text{Sing}(\tilde{\mathbf{U}}))$ given by the four-fold cover and rotations of each individual plane.
-This gives us a relation between the two tangent bundles, where the translation four-fold cover has the standard $\mathbb{R}/2\pi\mathbb{Z}$ unit tangent fiber.
-$\theta'$ is the direction associated to the flow $\omega_t^{\theta'}:[0,1]\rightarrow\mathbf{M}$ on the translation surface.
-Since the cover is translation, $p^{-1}(\theta')=\theta'=\theta+n\frac{\pi}{2}$.
-First suppose that $\vec{\theta}\in\mathcal{O}$.
-Then $\theta$ is identified with the set of directions that close on $\tilde{\mathbf{U}}$.
-From Theorem 5, $\omega_t^{\theta'}$ lifts to a closed geodesic $\tilde{\omega}_t^{\theta'}$.
-Given $\mathbf{hol}(\omega)=\int_{\omega}dz$, we obtain a period for the unit-speed flow, $t_0=|\mathbf{hol}(\omega)|$.
-That is, $\tilde{F}_t^{\theta'}:\mathbb{R}^+_0\rightarrow\tilde{\mathbf{U}}$ such that  $\frac{d}{dt}\tilde{F}_t=\frac{1}{|\mathbf{hol}(\omega)|}$.
-Then $F_t^{\phi}=F_t^{[\theta']}=f\circ \tilde{F}^{\theta'}_t$.
-The period carries over since there is no concern over a trajectory returning to $\tilde{x_0}$ in a different direction.
-Otherwise, the geodesic $\omega_t^{\theta'}$ on $\mathbf{M}$ would have closed in $0<t<1$.
-Now suppose that $\vec{\theta}\in\mathcal{E}$.
-Identifying it with $\theta'$, we see that $\omega$ in direction $\theta'$ is not an element of $\textbf{Ker }\Omega_{u,v}$ from Theorem 5.
-Therefore, $\Omega_{u,v}(\omega)=(m,n)\simeq T^{m,n}$ and lifting the terminal point $\omega(1)$, $\tilde{\omega}(1)=T^{m,n}(\tilde{\omega}(0))=T^{m,n}(\tilde{x}_0).$ The period remains unchanged, in that $\tilde{F}^{\theta'}_{t+\mathbf{hol}(\omega)}=\tilde{F}^{\theta'}_{t}+T^{m,n}(\tilde{x}_0)$.
-Therefore, $F^{\phi}_{t+t_{0}}(s)=f\circ\tilde{F}^{\theta'}_{t}(\tilde{x}_0)+f\circ T^{m,n}(\tilde{x}_0).$\\
-Conversely, suppose $F_t$ is periodic.
-Then $[\theta]=\phi=[\theta']$, which defines directional flows $\tilde{F}^{\phi}_t$.
-According to Theorem 5, $\tilde{F}^{\phi}_t$ will close if and only if $\phi\subset\mathcal{O}$.
-$\phi$ is the orbit of $\vec{\theta}'$ under the 90 degree rotational matrix.
-This matrix does not alter the length or period of a geodesic.
-Thus, $F^{\phi}_t$ is exactly one of the flows $\tilde{F}^{\phi}_t$.
-Likewise, if $F^\phi_t$ is drift-periodic then $F^\phi_{t+t_0}=f\circ \tilde{F}^\phi_t+f\circ T^{m,n}$.
-$T^{m,n}$ is trivial if and only if $\theta'\in\mathcal{O}$.
-Therefore, $\theta'\in\mathcal{E}$, and $[\theta']=\phi$.
-\end{proof}
-\end{conj}
-
-There is still much work to do in terms of cleaning up the proofs and organizing the final paper.
-
+\end{thm*}
 \newpage
-\section*{Conclusion}
-What I ultimately aim to do is port these results on X's homology back to the Necker cube surface.
-I want do it in such a way that the final theorem is bi-conditional.
-To do so, I imagine I can take a vector image of a small segment of a geodesic in $\mathbb{R}^3$ and project it onto the isometric flattening of the Necker Cube surface to obtain a direction (or classes of equivalent directions), and relate it to the unit tangent bundle of $\mathbf{M}$.
-
-In addition, I would also like to find a formula for the arc-length of a geodesic based on direction alone.
-Knowing that hol$(\alpha)=6+6i$ means that the induced Euclidean metric on $\mathbf{M}\backslash$Sing($\mathbf{M}$) gives the geodesic an arc-length of $6\sqrt{2}$.
-I would like to show that:
-
-\begin{align*}
-\int_{\beta}|dz|=|hol(h(\alpha))|=|h'(hol(\alpha))|,
-\end{align*}
-\noindent where $h'\in V(\mathbf{M})$ is the derivative of h, and $\beta=h(\alpha)$.
-We know that is true on the translation surface, but it's a matter of then showing the translation quotient, branch-cover, and the Necker cube surface have the same induced Euclidean metric of these non-singular geodesics.
-(It would not be surprising considering that the surface is built out of subsets of planes.) Even more of a problem is finding a way to solve for a matrix in the Sanov subgroup that brings (1,1) to the desired odd-odd slope.
-
 \newpage
@@
 \bibitem{Escher2} Escher, M.C., \emph{Convex and Concave}, 1955.
+\bibitem{flatsurf} include github link
 \end{thebibliography}

### 2018-01-04 newpaper.tex
010318

--- 2017-12-04 newpaper.tex
@@
 
-\title{Periodicity of Geodesics on the Necker Cube Surface}
+\setlength{\topmargin}{1in} %top/bot margins
+\setlength{\oddsidemargin}{\topmargin} %sidemargins
+
+\setlength{\textheight}{11in} \setlength{\textwidth}{8.5in}
+\setlength{\hoffset}{-1in} \setlength{\voffset}{-1in} \setlength{\evensidemargin}{\oddsidemargin} \addtolength{\textheight}{-2 \topmargin}\addtolength{\textwidth}{-2\oddsidemargin}
+\setlength{\headheight}{0pt} \setlength{\headsep}{20pt} \setlength{\footskip}{20pt}
+\addtolength{\textheight}{-\footskip} \addtolength{\textheight}{-\headheight} \addtolength{\textheight}{-\headsep}
+
+\title{Periodic Geodesics on the Infinite Necker Cube Surface}
 \author{Pavel Javornik}
@@
 \begin{abstract}
-\noindent Although it is most recognizable as M.C.
-Escher's adaptation of an optical illusion popularized by Louis Albert Necker, the Necker cube surface has a complex structure and is, in analaytic language, described as an infinite, flat (locally isometric to $\RR^2$ on smooth sections), $\RR^3$-embeddable Euclidean cone surface with countably many singularities of cone angles $3\pi$ and $\frac{3\pi}{2}$.
-Its geometric construction is described as the gluing of infinitely many unit cubes along their edges so that every face shares an edge with exactly four others.
-A unit-speed geodesic on this surface has an initial trajectory angle $\theta\in\RR/2\pi\ZZ$ given up by the surface's local plane isometry.
-Given a rational vector direction in $\RR^2$ (components are relatively prime integers) obtained from $\theta$, one can classify every periodic/drift-periodic rational geodesic on the surface.
-In addition its total period/length and, in the case where it is drift-periodic, translational drift as Euclidean distance in its ambient space is given by solving a system of simple linear Diophantine equations.
+\noindent We classify all of the periodic and drift-periodic trajectories on the surface periodically constructed out of unit cubes lying on a plane in $\RR^3$.
+The Necker cube surface is a piecewise smooth Euclidean cone surface with neighborhoods locally isometric to the plane that gives up a natural piecewise isometry, or \emph{flattening}, onto a subset of $\CC$ with $\ZZ^2$ translational symmetries.
+We use the surface's topological properties to draw a connection between translation covers and rotational holonomy of closed curves by inducing fundamental group actions on the discrete space of attainable vectors via parallel transport along lifted paths.
+We then use the symmetries of this surface and its (branched) translation cover to show a correspondence between the space of attainable (rational) geodesics of the Necker cube surface, their dynamical properties, and the translation cover's \emph{Veech group}.
+The ``slope" of a geodesic is represented by a relatively prime pair of integers, which is used to tell not only whether or not a geodesic with this initial direction (relative to choice of basis in its initial tangent space) closes, but the overall length of the trajectory as a function of its components.
 \end{abstract}
@@
 \label{fig:front}
-\caption{Periodic and drift-periodic flows on the Necker cube surface}
-\end{figure}
-
+\caption{Periodic and drift-periodic geodesics on the Necker cube surface}
+\end{figure}
 \newpage
 \section{Introduction}
-
-Put your Problem statement here! Example of a Citation\cite[p.219]{Robotics}. Here's Another Citation \cite{Flueck}
-
-The Necker cube has made numerous appearances in the work of mathematicians, crystallographers, and scientists interested in human visual systems before and after appearing in the works of M.C.
-Escher [cite].
-Louis Albert Necker was credited for having discovered the optical illusion, and studying its geometric properties [cite].
-The solid presentation of the cube, when rendered as a flat surface with three rear faces removed, achieves a similar effect when its three visible faces are shaded in a particular way.
-The Necker cube surface is then obtained by gluing infinitely many copies along its outer edges as in Figure 1.
-
-Dynamical classification of geodesics on flat, periodic surfaces such as the Necker cube surface ultimately boils down to  studying the surface's various symmetries and how the rotational holonomy group acts on the unit tangent bundle of the surface as the unit vector (derivative of the geodesic) is parallel transported along its path.
-Loops on polyhedra punctured at its cone singularities tend to have discrete holonomy groups as its relative holonomy (holonomy of contractible loops) is trivial.
-
+The Necker cube has been studied extensively due to its interesting effects on human visual perception.
+\cite{Albert} A popular adaptation of the Necker cube is the edge-to-edge pairing of multiple solid cubes giving the illusion of a rhombile tiling of the plane when viewed from a fixed angle.
+(Fig 1) It is most recognizable in the works of M.C.
+Escher \cite{Escher1}\cite{Escher2} and as the game board in the 80s cabinet arcade classic, $\text{Q}^*\text{bert}$.\cite{Qbert} The cube was named after the Swiss crystallographer Louis A.
+Necker for having been one of the first people to study the wire-frame cube's properties as an optical illusion.
+
+ The solid presentation of the cube, rendered as a flat surface with three rear faces obscured from view, achieves a similar effect to the ``wire-frame" version when infinitely many are arranged along their outer edges.
+The semi-cube is a construction obtained by removing these obscured faces and making translational identifications on these edges that is homeomorphic to the two-dimensional torus.
+The Necker cube surface is itself obtained by gluing infinitely many copies of the semi-cube surface along those outer edges.
+(Fig ??)
+
+An infinite surface like the Necker cube is piecewise smooth and \emph{flat} on its smooth sections.
+Neighborhoods contained in these faces are isometric to the Euclidean plane.
+The tangent space at a point is the vector field contained in a plane tangent to the surface.
+A vector field translated along a closed loop on this surface induces a phenomena known as \emph{rotational holonomy}.
+This is a linear transformation (element of $\SO(2,\RR)$ for orientable manifolds) between tangent spaces at the initial and terminal point of the loop.
+When a surface is flat, deformations of the loop do not alter its holonomy.
+Furthermore, a contractible (trivial) loop has trivial holonomy due to the local plane isometry on its smooth sections.
+In this case any homotopic path will induce the same linear transformations, and that gives up a well-defined group action on the tangent bundle of the surface by the first fundamental group of the surface.
+
+In studying geodesics on an infinite flat surface, it is important to make this connection early on.
+The rotational holonomy group acts transitively on the holonomy bundle, a principal bundle of tangent spaces over a point on the surface.
+This holonomy group is used to construct a covering space where the only closed loops are those with trivial holonomy.
+In desirable cases, the holonomy group is discrete and this cover has finitely many sheets.
+This surface is known as an infinite \emph{translation surface}.
+It is on this surface that geodesics behave as they would when projected back down to the flat surface.
+In the context of Riemann surfaces, a translation surface is a surface with a maximal translational atlas, i.e.
+every transition function on images of neighborhoods in $\CC$ is an affine translation.
+The surface inherits a flat metric from pulling back the Euclidean metric on $\CC$.
+On such a surface the principal holonomy bundle of tangent spaces over a point consists of a single vector space.
+Combining these two facts, in local coordinates any geodesic in direction $\theta\in\RR/2\pi\ZZ$ on this surface is isometric to $te^{i\theta}$ for $t\in\RR$.
+[cite Zorich]
 \begin{rem}
-Familiarity is assumed on the part of the reader with covering space theory, translation (Veech) surfaces, and their associated Veech groups. For general surveys on these topics: [cite], [cite].
+Familiarity is assumed on the part of the reader with covering space theory, translation (particularly Veech) surfaces, and their associated Veech groups.
+For general surveys on these topics we refer readers to: [cite], [cite].
 \end{rem}
 
+\subsection{Directions and Trajectories}
+What originally motivated this research was the appearance and symmetries of computer simulated geodesics on the surface.
+These symmetries are related to the reflective and rotational symmetries of its translation cover.
+\begin{figure}[H]
+\centering
+\includesvg[width=4in]{oddoddtrajectory}
+\caption{A closed trajectory on the Necker cube surface modeled using sage-flatsurf \cite{flatsurf} and various sage utilities. [cite]}
+\end{figure}
+We began finding patterns between the choice of initial direction and periodicity.
+What we aimed to find out was what geometric properties of the surface relate a geodesic's dynamical characteristics to its initial trajectory angle.
+The answer to the question of what directions determine periodicity and drift periodicity turned out remarkably simple.
+Along the way we learned that the symmetries of this surface and its various covers gave up not only a clear distinction between periodic and drift-periodic trajectories, but the total arclengths of these paths after a single period.
+For example the previous figure's geodesic has an initial trajectory angle $\tan^{-1}(51/31)\approx58.71^\circ$, and total total length of $6\sqrt{3562}\approx 358.09$.
+Appendix [include appendix] showcases various geodesics, their initial trajectory angles, and total lengths.
+
 \subsection{Discussion of Results}
-Our initial experiments strongly supported the theory that there would be a correlation between a choice of trajectory angle and dynamical properties of a geodesic on such a symmetric object.
-Rightfully so, a surface composed of infinitely many cubes an induced flat metric where every neighborhood is locally isometric to the Euclidean plane.
-Via the parallel transport of a unit tangent vector over the sharp edges of the surface denoted $\bUs$, it becomes necessary to think of the geodesic as a sequence of line segments contained in the faces of each cube on the surface's embedded form in $\RR^3$.
-$\ref{fig:front}$.
-\\
-\begin{wrapfigure}{l}{2in.}
-\includesvg{vectorplane}
+A surface composed of infinitely many cubes is endowed with a flat metric and local plane isometries.
+Via the parallel transport of a unit vector tangent to the surface along a geodesic, we view its image as a sequence of straight lines on the smooth faces of the surface.
+(Fig 2) A geodesic is no doubt a piecewise curve in space.
+Let $\bUs$ be the connected, topological space known as the Necker cube surface.
+\\\newpage
+\begin{wrapfigure}{l}{1.9in}
+\includesvg[width=1.7in]{vectorplane}
 \end{wrapfigure}
-\noindent Let $u_0$ be a point on the surface that is contained in the interior of a face and consider a tangent unit vector $v\in\RR^3$ based at point $u_0$.
-Each face is parallel to a $2$-dimensional subspace of $\RR^3$ spanned by exactly two basis vectors, so one component of $v$ is 0.
-Call this projected vector $v_0\in\RR^2$.
-\\\\\\
-The angle $v_0$ makes relative to a choice of basis is the \emph{initial trajectory angle} of the unit-speed geodesic on the surface, $\Phi_t:\RR\rightarrow\bUs$.
+\noindent Let $u_0$ be a point on the surface that is contained in the interior of a face and consider a tangent unit vector $v\in T_{u_0}(\bUs)$, the tangent vector field over $u_0$.
+A flattening map acts as an atlas on the smooth sections of $\bUs$ (see section [find section]) is an affine map from the surface to $\RR^2$ with a derivative in $\SO(3,\RR)$ that   (See section [?] in Appendix [Include section or appendix on this map].)
+\\\\
+Let the angle $v_0$ makes relative to a choice of basis (orthogonal vectors parallel to $\bUs$'s faces) be the \emph{initial trajectory angle} of a \emph{unit-speed geodesic} based at $u_0$, $\Phi_t:\RR\rightarrow\bUs$.
+The initial trajectory angle is said to be $\emph{rational}$ if there is some $k\in\RR$ such that $kv_0$'s nonzero components are relatively prime integers.
 A rational initial trajectory angle in $\RR^2$ falls into one of two categories:
 \begin{Def}
-Let $v_0$ be a unit vector of the form $\frac{1}{k}(x,y)\in\RR^2$ with $x,y\in\mathbb{Z}$ and $k=\sqrt{x^2+y^2}\in\mathbb{R}$.
-We say $v_0$ is an \textbf{odd-odd} vector if its components are relatively prime and both odd.
-We denote the \textbf{set of all odd-odd directions} by $\mathcal{O}$.
-We say that $v_0$ is an \textbf{even-odd} vector if its components are relatively prime and of opposite parity.
-We denote the \textbf{set of all even-odd directions} by $\mathcal{E}$.
-\end{Def}
-\noindent For convenience we call the disjoint union of these sets $\cS$.
-This paper aims to depict the relationship between initial trajectory angles and dynamical properties of a geodesic on the surface.
-Denote $\bUs$'s translational isometries by $\text{Trans}(\bUs)$.
-Recall that a geodesic on the surface is \emph{periodic} if there exists some $T>0$ such that $\Phi(t+T)=\Phi(t)$ for all $t\in\RR$ and a geodesic is \emph{drift-periodic} if there exists some $T'>0$ and non-trivial $p\in \text{Trans}(\bUs)$ such that $\Phi(t+T')=p(\Phi(t))$ for all $t\in\RR$.
-In either case we call this value the \emph{period} of $\Phi$ if it is the \emph{smallest} possible value that satisfies our definition.
-We additionally require that at no point does $\Phi$ encounter a cone singularity of the surface, in which case the geodesic behavior is non-deterministic.
-
-\begin{align}
-\text{Let }\XX'=\left\< \left[ \hspace{1mm} \begin{matrix}
-							1 & 2\\
-							 0 & 1
-						\end{matrix}\hspace{1mm}\right],~
-\left[ \hspace{1mm} \begin{matrix}
-							1 & 0\\
-							 2 & 1
-						\end{matrix}\hspace{1mm}\right]\right\>
-\end{align}
-
-\noindent Section [blank] explains the significance of this group. Let $M=\left[ \hspace{1mm} \begin{matrix}
-							s & u\\
-							 w & z
-						\end{matrix}\hspace{1mm}\right]\in\XX'$.
-
-\begin{thm*}
-(Dynamical Properties of a Geodesic on $\bUs$) \\Let $\Phi_t:\RR\rightarrow\bUs$ be a non-singular unit-speed geodesic on the Necker cube surface such that $\Phi(0)=u_0$ and $\Phi'(0)=v\in\RR^3$, isometric to $v_0\in\RR^2$.
-Let $\overline{u}_t\in\RR^3$ be $\Phi_t$'s coordinate in $\bUs$'s ambient space at time $t$.
+Let $v_0$ be a rational unit vector of the form $\frac{1}{k}(x,y)\in\RR^2$ with $x,y\in\mathbb{Z}$ and $k=\sqrt{x^2+y^2}\in\mathbb{R}$.
+We say $kv_0$ is an \textbf{odd-odd} vector if $x,y$ are relatively prime and both odd.
+We denote the \textbf{set of all odd-odd directions} $\mathcal{O}$.
+We say that $kv_0$ is an \textbf{even-odd} vector if $x,y$ are relatively prime and of opposite parity.
+We denote the \textbf{set of all even-odd directions} $\mathcal{E}$.
+Finally, $\cS=\cO\sqcup\cE$ is the \textbf{collection of all possible rational directions}.
+\end{Def}
+\noindent It may be more helpful to think of an element in $\cS$ as the slope of $\Phi$, given $\bUs$'s flat structure.
+Recall that a geodesic on the surface is \emph{periodic} if there exists some $T>0$ such that $\Phi(t+T)=\Phi(t)$ for all $t\in\RR$, and \emph{drift-periodic} if there exists some $T'>0$ and non-trivial translation, $f$, such that $\Phi(t+T')=f(\Phi(t))$ for all $t\in\RR$.
+In any case we call this value the \emph{period} of $\Phi$ if it is the \emph{smallest} possible value that satisfies this.
+We additionally require that at no point does $\Phi$ encounter a cone singularity of the surface, where geodesic behavior is undefined.
+We call these well behaved geodesics \emph{non-singular}.
+We use $||~||:\RR^2\rightarrow\RR_{\geq 0}$ to denote the standard Euclidean norm in $\RR^2$.
+
+\begin{thm*} \textbf{Periodic and Drift-Periodic Geodesics on the Necker cube surface.} 
+\\Let $\Phi_t:\RR\rightarrow\bUs$ be a non-singular unit-speed geodesic on the Necker cube surface such that $\Phi(0)=u_0$ and $\Phi'(0)=v\in\RR^3$.
+Let $v_0\in\RR^2$ be a distance-preserving projection of $v$ onto the plane tangent to the surface at $u_0$.
+Let $k\in\RR$ such that $kv_0\in\cS$.
 Then the following is true:
 \begin{enumerate}[label=(\roman*)]
-\item $\Phi$ is periodic if $v_0\in\cO$.
-\item Suppose $(i)$ is true. If $M (kv_0)=(1,1)$ then $\Phi$ has period $T=6\sqrt{2}||kv_0||$.
-\item $\Phi$ is drift-periodic if $v_0\in\cE$. 
-\item Suppose $(iii)$ is true. If $M (kv_0)=(1,0)$ or $M (kv_0)=(0,1)$ then $\Phi$ has period $T=2||kv_0||$, and $d_E(\overline{u}_t,\overline{u}_{t+T})=||kv_0||\sqrt{2}$.
+\item $\Phi$ is periodic if $kv_0\in\cO$ with period $T=6||kv_0||$.
+\item $\Phi$ is drift-periodic if $v_0\in\cE$ with period $T=2||kv_0||$.
 \end{enumerate}
 \end{thm*}
-\compav{Remember to prove that such an M exists.}
-
-Here $||~||$ is the standard Euclidean norm in $\RR^2$, and $d_E:\RR^3\times\RR^3\rightarrow\RR$ is the Euclidean distance.
-The proof for this theorem can be found in XXX and follows from Theorems YYYY and ZZZZZ.
+
+There is a very clear dynamical dichotomy at work when it comes to $\Phi$'s slope, not unlike the well-known Ehrenfest Wind-Tree model with $\frac{1}{2}\times\frac{1}{2}$ square obstacles distributed on a $\ZZ^2$ lattice.
+[cite]
 
@@
 
-\section{Periodic Tiling of Necker Cubes}
-This section will detail how the Necker Cube surface is constructed as a covering space of a thrice-punctured torus. Make the following identifications on 3 unit squares:
+\section{Periodic Construction of the Surface}
+This section will detail how the Necker Cube surface is constructed as a covering space of a thrice-punctured torus. We begin by making the following identifications on 3 unit squares:
 \begin{figure}[H]
@@
 We will call this ``half-cube" $\bG$.
-It is a genus one piecewise smooth surface homeomorphic to $\TT^2$.
-Every vertex has a cone angle of either $3\pi$ or $\frac{3\pi}{2}$.
-Let $\Sigma_{\bG}\subset\bG$ be the set of these three vertices.
-These are conical singularities of the surface.
-We denote the surface punctured at these points by $\bG\backslash\Sigma_{\bG} = \bGs$.
-The paths labeled $a,b,c,d$ are independent of one another and span $\bGs$'s first fundamental group, $\pi_1(\bGs)\cong \<a,b,c,d \>$, the free group of four generators.
-The unpunctured Necker cube surface $\bU$ is embedded in $\RR^3$ by tiling along $\bG$'s outer edges:
+It is a piecewise smooth surface homeomorphic to the torus.
+Every vertex is a cone singularity of the surface with cone angle $3\pi$ or $\frac{3\pi}{2}$.
+Let $\Sigma_{\bG}\subset\bG$ be the set of these singularities.
+We use $\bG\backslash\Sigma_{\bG} = \bGs$ to denote the surface punctured at these points.
+The paths labeled $a,b,c,d$ are independent of one another and generate $\bGs$'s fundamental group, $\pi_1(\bGs)\cong \<a,b,c,d \>$, the free group of four generators.
+(Figure ?) The unpunctured Necker cube surface $\bU$ is embedded in $\RR^3$ by gluing along $\bG$'s outer edges (marked with solid triangles):
 
@@
 \end{figure}
-From this diagram we see that $\bU$ is a universal cover of $\TT^2$.
-In this paper we are studying geodesic behavior on $\bU$ and in order to do so, it is necessary to determine how the parallel transport of a unit vector around a cone singularity acts on the unit tangent bundle of the surface.
-But to study the path monodromy we require that these loops (see $c,d$ in Fig 3) aren't contractible on the cube's sharp corners.
+From this diagram we see that $\bU$ is a universal cover of $\bG$.
+In this paper we are studying geodesic behavior on $\bU$ and, in order to do so properly, it is necessary to determine how the parallel transport of a unit vector around a cone singularity acts on the unit tangent bundle of the surface.
+But to study a given path's rotational holonomy, we require that these loops (see $c,d$ in Fig 3) aren't contractible through the cube's sharp corners.
+To that effect, we puncture the surface at these points.
 Denote the countably infinite set of singularities of $\bU$ by $\Sigma_{\bU}$.
 The \emph{Necker cube surface} is the punctured surface $\bUs=\bU\backslash\Sigma_{\bU}$.
-Observe from the previous diagram that $\pi_1(\bUs)$ is the kernel of the following group homomorphism:
-
-\begin{Def}
-Denote $\varphi_1:\pi_1(\bGs)\rightarrow \ZZ^2$ as the group homomorphism such that $c,d\mapsto(0,0)$, $a\mapsto(1,0)$, and $b\mapsto(0,1)$.
-\end{Def}
-
-Take $\pi_1(\bUs)$ to be $\ker\varphi_1$.
-Then $\Delta_{\varphi_1}=\pi_1(\bGs)/\pi_1(\bUs)\cong\ZZ^2$ is the \emph{deck group} of the covering map $\varphi_1^*:\bUs\rightarrow\bGs$.
-Now $\bGs$ is unfolded and flattened in order to study monodromy as it transports a vector in $\RR^2$ along a path.
-Apply the following cutting and gluing operations to the surface: \\
+Observe from the previous diagram that $\pi_1(\bUs)$ is the kernel of the following group homomorphism on $\pi_1(\bGs)$:
+
+\begin{Def}
+Let $\varphi_1:\pi_1(\bGs)\rightarrow \ZZ^2$ be the homomorphism on a free group given by $c,d\mapsto(0,0)$, $a\mapsto(1,0)$, and $b\mapsto(0,1)$.
+\end{Def}
+
+The cover of $\bGs$ with fundamental group $\ker\varphi_1$ is the space $\bUs=\tilde{\mathbf{G}}^\circ/\ker\varphi_1$ where $\ker\varphi_1$ acts on $\bGs$'s universal cover $\tilde{\mathbf{G}}^\circ$ by deck transformations.
+
+\begin{Def}
+Let $p_1:\bUs\rightarrow\bGs$ be a projection that induces a monomorphism on the fundamental groups of these surfaces such that $p_{1*}\pi_1(\bUs)=\ker\varphi_1$.
+\end{Def}
+
+It follows that $(\bUs,p_1)$ is a covering space of $\bGs$.
+We say $\Delta_{p_1}=\pi_1(\bGs)/p_{1*}\pi_1(\bUs)\cong\ZZ^2$ is the \emph{deck group} of the covering isomorphic to the group of deck transformations of $p_1$ $-~\text{Deck}_{p1}(\bUs)$.
+Using the fact that this cover is regular, we say that $p_1(\bUs)=\bUs/\ZZ^2$ and $\bGs$ is a quotient of the Necker cube surface.
+Apply the following cutting and gluing operations to $\bGs$: \\
 \begin{figure}[H]
@@
 \end{figure}
-\noindent This L-shaped torus is then taken apart and recovered as $2\times 2$ torus missing a unit square with the following identifications:\\
+\noindent This L-shaped torus is then taken apart and recovered as $2\times 2$ torus missing a unit square with the following edges identified:\\
 \begin{figure}[H]
@@
 \includesvg[width=2in]{BranchTorus}
-\caption{$\bGs$ and images of paths after reconstruction. The outermost vertex is \emph{not} a puncture.}
+\caption{$\bGs$ and its paths after reconstruction.}
 \end{figure}
@@
 Consider a $2\times2$ sheet missing an open unit square and that square's vertices (punctures).
-Center this structure around $0\in\CC$ and recover $\bGs$ with the identifications made in Fig 3.
+Center this structure around $0\in\CC$ and recover $\bGs$ with the identifications made in Fig ??.
 Let $m,n\in\ZZ$.
@@
 
-\noindent Let $\aleph_{m,n}=2m+i2n\in\CC$ be the center of a removed square. Now let $\zeta_{m,n}(t)=(2m+t)+i(2n-\frac{1}{2})$ and 
-$\xi_{m,n}(t)=(2m+t)+i(2n+\frac{1}{2})$ be paths on $\mathbf{P}$ that run along the edges of the squares, parameterized by $t\in (-\frac{1}{2},\frac{1}{2})$ and $(m,n)\in\ZZ^2$.
-$\bUs$ is recovered as a topological quotient $\mathbf{P}/\sim$, where $\sim$ is the minimal relation on $\mathbf{P}$:
+\begin{figure}[H]
+\centering
+\includegraphics[width=2.8in]{P.png}
+\end{figure}
+\compav{shade the surface}
+\noindent Let $m,n\in\ZZ$.
+The coordinate $c_{m,n}=(2m+i2n)\in\CC$ is the center of a removed square.
+The set of all centers form a lattice in the plane that uniquely identifies every section in the fiber over $\bGs$ under $p_1$ with an element of $2\ZZ[i]$.
+Now let $\zeta_{m,n}(t)=(2m+t)+i(2n-\frac{1}{2})$ and
+$\xi_{m,n}(t)=(2m+t)+i(2n+\frac{1}{2})$ be paths on $\mathbf{P}$ that run along the top and bottom edges of the squares, parameterized by $t\in (-\frac{1}{2},\frac{1}{2})$ and $(m,n)\in\ZZ^2$.
+These paths are oriented so that edge identifications can be made by translations and rotations relative to $(m,n)$.
+(see figure above) $\bUs$ is recovered as the topological quotient $\mathbf{P}/\sim$, where $\sim$ is the following minimal relation on $\mathbf{P}$:
 
@@
 \begin{split}
-\zeta(t)\sim e^{i\frac{\pi}{2}}~\overline{\zeta(t)-\aleph}\\
-\eta(t)\sim e^{i\frac{\pi}{2}}~\overline{\eta(t)-\aleph}
+\zeta_{m,n}(t)\sim i[\zeta_{m,n}(t)-c_{m,n}]+c_{m,n}-1\\
+\xi_{m,n}(t)\sim i[\xi_{m,n}(t)-c_{m,n}]+c_{m,n}+1
 \label{eq:rel2}
@@
 \end{equation}
-The translations, $\pm 2,\pm 2i$, on $\mathbf{P}$ are isometries that preserve these identifications and induce isometries of $\bUs$.
-The group generated by these horizontal and vertical translations form the \emph{group of Deck transformations}, $\text{Deck}(\varphi_1^*)$, that  is isomorphic to $\ZZ^2$:
+The translations, $\pm 2,\pm 2i$, on $\mathbf{P}$ are isometries that preserve these identifications and induce isomorphisms of $\bUs$.
+We denote the isomorphism induced by $+2$ as $f_h$, for horizontal, and the isomorphism induced by $+2i$ as $f_v$, for vertical:
 
@@
 
-Identify 2 with $f_h$, and $2i$ with $f_v$.
-Let $x_0\in\bGs$ and fix a point in the fiber,  $u_0\in\varphi_1^*{-1}(x_0)\subset\bUs$.
-The action of $\ZZ^2$ on the fiber over $x_0$ is given by $(m,n)\cdot u_0=(f_h^m\circ f_v^n)(u_0)$.
-The distance between $u_0$ and $(f_h^m\circ f_v^n)(u_0)$ is measured in its ambient space and taken to be the magnitude of $2m+i2n\in\CC$.
+From this image it's clear that the group generated by these horizontal and vertical isomorphisms under function composition, $\<f_h,f_v~\vert~[f_h,f_v] \>$, form $\text{Deck}_{p_1}(\bUs)$.
+Let $x_0\in\bGs$ and fix a point in the fiber, $u_0\in p_{1}^{-1}(x_0)\subset\bUs$.
+We consider the groups $\pi_1(\bUs,u_0)$ and $\pi_1(\bGs,x_0)$.
+There is a well-defined transitive action of the group $\pi_1(\bGs,x_0)$ on the set $p_{1}^{-1}(x_0)$ by lifting a loop on $\bGs$ to a curve on $\bUs$ based at a point in the fiber, and sending it to the terminal point of the lift.
+[cite that algebraic top book 5.7, beginning pg.133].
+A curve is a continuous map from $[0,1]$ to the surface, and a loop is a closed curve where the terminal and initial point coincide.
+
+\begin{thm}
+Let $a,b\in\pi_1(\bGs,x_0)$ be the homotopy class generators as in Figure [??].
+Let $\tilde{a},\tilde{b}$ be $u_0$-based lifts of a loop in $a,b$ onto $\bUs$.
+Then $\tilde{a}(1)=f_h(u_0)$ and $\tilde{b}(1)=f_v(u_0)$.
+\end{thm}
+
+\begin{cor}
+Let $\alpha\in\pi_1(\bGs,x_0)$. Denote the unique $u_0$-based lift of a loop in $\alpha$ to $\bUs$ as $\tilde{\alpha}$. If $\varphi_1(\alpha)=(m,n)$, then 
+\begin{align}
+\tilde{\alpha}(1)=(f_h^m\circ f_v^n)(u_0).
+\end{align}
+\end{cor}
+This just follows from the previous theorem as $c,d$ are mapped trivially, and under $\varphi_1$ a word in $\pi_1(\bUs)$ is taken to commutative automorphisms of $\bUs$.
+For shorthand, we will use $f_\alpha$ to denote the \emph{induced deck transformation.}
 
 \begin{lem}
-Let $[\alpha]$ be homotopy class not in the kernel of $\varphi_1$. It will lift to an unclosed path terminating on $\varphi_1([\alpha])\cdot u_0$.
+$f_\alpha$ is a local diffeomorphism of the surface $\bUs$. 
+\begin{proof}
+Consider a 
+\end{proof}
 \end{lem}
 
-\subsection{Monodromy Group Representation}
-A \emph{rotational monodromy map} is a morphism from the first fundamental group of a surface to the \emph{monodromy group} acting on the unit tangent bundle via parallel transport of a vector along its path.
-Existence of a parallel transport map is a consequence of the metric connection on a surface with complex structure.
-The monodromy map is well defined when the \emph{holonomy group} is \emph{discrete}.
-When the \emph{relative holonomy} of contractible loops is trivial, the holonomy group is the set of attainable rotations of a unit vector upon returning to the base point after transporting it along a non-trivial loop.
-A \emph{flat surface} like $\bGs$ has trivial relative holonomy as its neighborhoods are isometric to the plane.
-Denote the \emph{unit tangent bundles} of $\bUs$ and $\bGs$ by $T^1\bUs$ and $T^1\bGs$.
-If $\alpha$ is a loop on $\bGs$, then $[\alpha]$ acts on a fixed vector $v_0\in\RR^2$ by rotation.
-In the case where $\alpha$ is \emph{not} a geodesic segment on the surface of length $L$, $v_0=\alpha'(0)\in\RR^2$ may differ from $\alpha'(L)$ by some rotation.
-
-\begin{Def}
-Define $\varphi_2:\pi_1(\bGs)\rightarrow SO(2,\ZZ)$ to be the group homomorphism where $a,b \mapsto I_2$, $c  \mapsto R$, and $d  \mapsto R^3$, where $I_2$ is the identity matrix and $R$ is a positive $90^\circ$ rotation matrix.
-\end{Def}
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=2.5in]{monodromy.png}
-\caption{Effect a nontrivial loop has on basis vectors $(i,j)\in\RR^2$.}
+\subsection{Embedding the Necker cube surface}
+In this section we define a map from $\bUs$'s $\RR^3$ embedded form to $\mathbf{P}$.
+The isomorphisms induced from translations on $\mathbf{P}$ are local diffeomorphisms on the smooth sections of $\bUs$, $S(\bUs)$, that relate the fundamental group of $\bGs$ to the tangent spaces of $p_1^{-1}(x_0)$.
+$S(\bUs)$ is the set of faces of cubes without boundary that are isometric to the plane.
+Let $m,n,l\in\ZZ$.
+A topologically equivalent notion of the (unpunctured) Necker cube surface is as the union of sets:  \\
+\begin{align*}
+\mathbf{A}=\bigcup\big{\{}\mathbf{A}_{m,n,l}: m+n+l=0\big{\}},
+\\\mathbf{B}=\bigcup\big{\{}\mathbf{B}_{m,n,l}: m+n+l=0\big{\}},
+\\\mathbf{C}=\bigcup\big{\{}\mathbf{C}_{m,n,l}: m+n+l=0\big{\}},
+\end{align*}
+which correspond to the following families of ``half-cubes" based integer points on a plane in $\RR^3$:
+
+\vspace{0.2in}
+\begin{tabular}{p{10cm}c}
+\begin{align*}
+\mathbf{A}_{m,n,l} = [m, m+1]\times[n,n+1]\times\{l\}, 
+\\\mathbf{B}_{m,n,l} = \{m+1\}\times[n,n+1]\times[l-1,l],
+\\\mathbf{C}_{m,n,l}= [m,m+1]\times\{n+1\}\times[l-1,l].
+\end{align*}
+&
+\shiftleft{0.3in}{\raisebox{-1in}{\includegraphics[scale=1]{label.png}}}
+\end{tabular}
+
+\begin{figure}[H]
+\centering
+\includesvg[width=1in]{cubesxyzcut}
+\caption{Embedding the Necker cube surface, i.e. $\bU=\mathbf{A}\cup\mathbf{B}\cup\mathbf{C}$.}
+\end{figure}
+
+\compav{use different symbol for embedding maybe $\bUs_{\RR^3}$}\\
+Of course $\bUs$ is just the exclusion of the semi-cube vertices.
+Then by removing the edges of each cube, we obtain $S(\bUs)$.
+Identify $\mathbf{P}\subset\CC$ with $\RR^2$ in the usual way, $(1,0)\sim 1$ and $(0,1)\sim i$.
+We say $\Psi:\RR^3\rightarrow\RR^2$ is a piecewise linear map from $\bUs$ to $\mathbf{P}$ given as
+\begin{equation}
+\Psi\left[\begin{array}{c}
+	x\\y\\z
+\end{array}\right] 
+= 
+\begin{cases}
+	\left[ \hspace{2mm} \begin{matrix}
+		1 & 0 & 0 \\
+		0 & 1 & 0 \\
+	\end{matrix}\hspace{3mm}\right]
+
+	&\left[\begin{array}{c}
+	x - m
+	\\ y- n
+	\\ z - l
+	\end{array} \right]
+	+
+	\left[\begin{array}{c}
+		2m - \frac{3}{2}
+		\\ 2n - \frac{3}{2}
+	\end{array} \right]
+		\text{if } (x,y,z)\in \mathbf{A}_{m,n,l}	\vspace{2mm}
+	\\
+
+	\left[ \begin{matrix} \hspace{2mm}
+	0 & 0 & -1 \\
+	0 & 1 & 0 
+	\end{matrix}\hspace{2mm}\right]
+	&\left[\begin{array}{c}
+		x - m
+		\\ y- n
+		\\ z - l
+		\end{array} \right]
+	+
+		\left[\begin{array}{c}
+			2m - \frac{1}{2}
+			\\ 2n - \frac{3}{2}
+		\end{array} \right]
+		 \text{if } (x,y,z)\in \mathbf{B}_{m,n,l}	\vspace{2mm}
+	\\
+
+		\left[ \begin{matrix} \hspace{2mm}
+		1 & 0 & 0 \\
+		0 & 0 & -1 
+		\end{matrix}\hspace{2mm}\right]
+		&\left[\begin{array}{c}
+			x - m
+			\\ y- n
+			\\ z - l
+			\end{array} \right]
+		+
+			\left[\begin{array}{c}
+				2m - \frac{3}{2}
+				\\ 2n - \frac{1}{2}
+			\end{array} \right]
+				 \text{if } (x,y,z)\in \mathbf{C}_{m,n,l}	\vspace{2mm}
+\end{cases}
+\end{equation}
+
+The inverse of this map would have to recover one of the coordinates as a function of the other two.
+To recover a section of $\Psi(\mathbf{A})$, for example, the z-coordinate would be $-(m+n)$.
+The rest are not difficult to determine.
+To see this particular setup used in modeling geodesics, we refer the reader to a flatsurf demonstration.
+[cite]
+
+\begin{thm}
+Let $\Psi=\{\Psi_{A_{m,n,l}},\Psi_{B_{m,n,l}},\Psi_{C_{m,n,l}}~:~m+n+l=0 \}$. Then $\Psi$ is a complex atlas on $\bUs_{\RR^3}$.
+\begin{proof}
+Clearly these charts cover the surface.
+Observe from the following diagram that there are finitely many intersections of any two neighborhoods that coincide with an edge on the cube centered at $(m,n,l)$.
+There are only 9 edges to consider since every cube is punctured at its vertices.
+\begin{figure}[H]
+\centering
+\includesvg[width=4in]{edgegraph}
+\end{figure}
+A transition function on the four edges of the $\mathbf{A}$ face is obtained as the identity just by setting $x$ or $y$ to $m\pm1$ or $n\pm1$.
+Similarly, from $\mathbf{B}_{m,n,l}$, $\mathbf{C}_{m,n,l}$ to $\mathbf{A}_{m+1,n,l-1}$ and $\mathbf{A}_{m,n+1,l-1}$, these values coincide.
+The trickier ones to show are the transition functions on the 3 vertical edges.
+These are the edges that split when mapped onto $\mathbf{P}$.
+\\
+Let $t\in(-\frac{1}{2},\frac{1}{2})$. Consider the following parameterization of these edges:
+\begin{align*}
+a(t):=&(m+1,n,l-\frac{1}{2}+t)\\
+b(t):=&(m,n+1,l-\frac{1}{2}+t)\\
+c(t):=&(m+1,n+1,l-\frac{1}{2}+t)\\
+\end{align*}
+
+\compav{Show that the transition functions are given by the identification $\sim$ on $\mathbf{P}$.}
+
+By induction on these integers, it naturally follows that this is true of the entire Necker cube surface.
+\end{proof}
+
+\end{thm}
+
+Therefore this surface is Riemann. In addition, $(\bUs,\Psi)$ respects the edge identifications made on $\mathbf{P}$ and the two are topologically equivalent.
+
+\compav{come back to the discussion of holonomy}
+
+Let $\gamma:[0,1]\rightarrow\bUs$ be a curve on the surface with $\gamma(0),\gamma(1)\in S(\bUs)$.
+If $\gamma(0)$ and $\gamma(1)$ are not contained in the same face, then $\gamma$ is a piecewise smooth curve.
+
+Since this surface is obtained as a regular cover of $\bGs$, it has a complex structure and Riemannian connection, $\triangledown$, on the tangent bundle of the surface, $T\bUs=\bigcup_{p\in\bUs}T_{p}(\bUs)$.
+This implies the existence of parallel translation maps of vector fields along $\gamma$ and a linear isomorphism between tangent spaces $T_{\gamma(0)}(\bUs)$ and $T_{\gamma(1)}(\bUs)$.
+[cite Loring W.
+Tu Dif Geometry] The Necker cube surface is \emph{flat} and admits zero curvature throughout.
+Let $\gamma$ be a closed loop on $\bUs$ based at $u_0\in p_1^{-1}(x_0)$.
+If $u_0$ is contractible to a point, then it would act trivially on $T_{u_0}$.
+[Sharpe,1997] Let $v_0\in T\bUs_{\gamma(0)}$.
+Let $P_\gamma:T\bUs_{u_0}\rightarrow T\bUs_{u_0}$ be the parallel transport map given up by the connection $\triangledown$ along the loop $\gamma$.
+The \emph{rotational holonomy group} is the set of all parallel transport maps given up by closed loops $\gamma$, $\text{Hol}_{u_0}(\triangledown)$.
+This forms a subgroup of $\SO(2,\RR)$ that acts on $T_{u_0}(\bUs)$.
+[cite Santoro]
+
+\subsection{Rotational Monodromy Group}
+A \emph{rotational monodromy map} is a homomorphism from the fundamental group of a surface to the \emph{rotational holonomy group} acting on the unit tangent space via parallel transport of a vector along a curve in its homotopy class.
+Existence of a parallel transport map is a consequence of an affine connection on a Riemann surface.
+On an oriented surface, the rotational holonomy group is an element of $\SO(2,\RR)$ that associates a loop with a linear isomorphism of the tangent space upon returning to the initial point.
+When the surface is \emph{flat}, contractible loops have trivial linear holonomy and will parallel transport a vector on a tangent vector field along any contractible loop without rotating it.
+[cite Zorich] In this case, rotational monodromy is well-defined and maps the fundamental group of a surface to a subgroup of $\SO(2,\RR)$.
+
+\begin{Def}
+Let $I_2=\begin{bmatrix}1 & 0\\0 & 1\end{bmatrix}$ and $R=\begin{bmatrix}0 & -1 \\1 & 0\end{bmatrix}$. \\
+The map $\varphi_2:\pi_1(\bGs)\rightarrow SO(2,\ZZ)$ is the group homomorphism given by \\$a,b \mapsto I_2$ \\$c  \mapsto R$\\$d  \mapsto R^3$.
+\end{Def}
+
+\begin{figure}[H]
+\centering
+\includesvg[width=2.5in]{neckercubeholonomy}
+\includegraphics[width=2.9in]{monodromy.png}
+\caption{Parallel transporting a unit vector along loops homologous to $c$ and $d$ (left) and corresponding effect $c$ and $d$ loops have on $i,j\in T_{x_0}(\bGs)$ (right).}
 \label{fig:loop}
@@
 
+\begin{thm}
+$\varphi_2:\pi_1(\bGs)\rightarrow \mathrm{SO}(2,\ZZ)$ is the rotational monodromy map into $\bGs$'s holonomy group.
+\begin{proof}
+Let $\gamma:[0,1]\rightarrow\bGs$ be a closed loop on the surface based at $x_0$, and $\alpha\in\pi_1(\bGs,x_0)$ its homotopy class.
+$\alpha$ is a word in $\<a,b,c,d\>$, so it suffices to show how the vector field along each loop acts on $T_{x_0}$.
+Because $\bGs$ is flat, trivial loops act trivially on $T_{x_0}$.
+From figure [??], we can tell this is true of classes $c$ and $d$.
+Similarly, parallel transportation of a vector field along an element of $a$ or $b$ acts trivially.
+Since this holds for its generators, it holds for $\alpha$ as well.
+\end{proof}
+\end{thm}
+
 \begin{lem}
-$\varphi_2:\pi_1(\bGs)\rightarrow \mathrm{SO}(2,\ZZ)$ is the rotational monodromy representation of $\bGs$.
-\begin{proof}
-From Figure 4 it's clear that $c,d$ will rotate a unit vector by $\pm 90^\circ$ by parallel transport along their paths.
-These values are determined by the cone angles of the singularities they loop around.
-The parallel transport of a vector along $a$ or $b$ would be trivial since the outer edges they pass through are glued by translation.
+Let $\alpha\in\pi_1(\bGs)$, and let $\gamma:[0,1]\rightarrow\bUs$ be a $u_0$ based homotopy lift into $\btUs$. Then $P_{\gamma}=I_2$ if and only if $\varphi_2(\alpha)=I_2$.
+\begin{proof}
+Suppose $P_{\gamma}=I_2$. Then $T_{\gamma(0)}(\bUs)=T_{\gamma(1)}(\bUs)$. Since \\\\
+FINISH THIS\\\\\\
 \end{proof}
@@
 
-\begin{lem}
-Any element in $\text{Deck}(\varphi_1^*)$ acting on $T^1\bUs$ has a trivial effect on holonomy.
-\begin{proof}
-Since $\bGs=\bUs/\ZZ^2$, the deck group action lifts naturally to an action on their tangent vector spaces and unit tangent bundles via pushforward of deck transformations $f_h,f_v$ on the bundle maps.
-i.e.
-$T^1\bGs=T^1\bUs/\ZZ^2$.
-This effect must be trivial since $\varphi_1(a)\sim f_h$ and $\varphi_1(b)\sim f_v$, and both homotopy classes $a,b$ have trivial holonomy.
-\\
-How do $f_h,f_v$ parallel transport a vector?\\
-Well if $\varphi_1([a])\cdot u_0=f_h(u_0)$, then for 
-\end{proof}
-\end{lem}
-
-\begin{thm}
-Let $m=\varphi_2\vert_{\pi_1(\bUs)}$. Then $m$ is the rotational monodromy map of $\bUs$.
-\begin{proof}
-
-\end{proof}
-\end{thm}
-
-\begin{cor}
-Let $\alpha:[0,1]\rightarrow\bGs$ be a loop on $\bGs$ with trivial holonomy and $\tilde{\alpha}$ its lift to $\bUs$. Then $\tilde{\alpha}$ has trivial holonomy as well.
-\begin{proof}
-
-\end{proof}
-\end{cor}
-
-\begin{Def}
-Let $\bMs$ be the regular cover of $\bGs$ with fundamental group $\pi_1(\bMs)=\ker\varphi_2$.
-\end{Def}
-Showing that $\bMs$ is a degree four cover of $\bGs$ with trivial holonomy and deck group $\SO(2,\ZZ)$ follows immediately from its construction.
+\begin{Def}
+Let $\bMs$ be the regular cover of $\bGs$ with deck group $\SO(2,\ZZ)$.
+Let $p_2:\bMs\rightarrow\bGs$ be the projection of this cover that induces a monomorphism $p_{2*}:\pi_1(\bMs,\xtild)\rightarrow\pi_1(\bGs,x_0)$ such that $p_{2*}\pi_1(\bMs,\xtild)=\ker\varphi_2$.
+\end{Def}
+
 \begin{figure}[H]
@@
 \label{fig:arbitrarylift}
-\caption{Lift of paths $c,d$ to $\bMs$ (\compav{RELABEL EDGES})}
-\end{figure}
+\caption{Lift of paths $c,d$ to $\bMs$}
+\end{figure}
+The surface $\bMs$ is a degree four cover of $\bGs$ that forces its loops to have trivial rotational holonomy.
+The metric completion of $\bMs$ is the compact translation surface $\bM$ obtained by reinserting its cone singularities.
+A \emph{translation surface} is a topological surface with a discrete set of cone singularities and a maximal translation atlas on the surface with these singularities removed.
+Neighborhoods of the surface are naturally isometric to $\CC$, and a metric on the translation surface is obtained by pulling back $\CC$'s Euclidean metric.
+We later use $\bMs$ to show that a geodesic in a rational direction on $\bGs$ closes and is unique.
 
 \subsection{A Four-Fold Cover}
-We now take these covers we built to construct a third one that covers them all. This is a four-fold cover of $\bUs$ and a $\ZZ^2$ translation cover of $\bMs$. 
-\begin{Def}
-Let $\btUs$ be the cover of $\bUs$ with fundamental group $\pi_1(\btUs)=\ker\varphi_2\big\vert_{\pi_1(\bUs)}$.
-\end{Def}
-
-It follows that $\ker\varphi_1\big\vert_{\pi_1(\bMs)}=\ker\varphi_2\big\vert_{\pi_1(\bUs)}$ and that $\btUs$ is a cover of these two surfaces.
-Edge identifications are made on four copies of $\mathbf{P}$, possibly indexed by the finite component of the direct product $\mathbf{P}\times\ZZ/4\ZZ$.
-The Deck group of the covering map from $\btUs$ to $\bGs$ is the set $\ZZ^2\times \SO(2,\ZZ)$.
-Consider a loop on $\bUs$:
+We now take these covers we built to construct a third one that covers them all.
+This is a four-fold cover of $\bUs$ and a $\ZZ^2$ translation cover of $\bMs$ with important properties that allow us to study geodesics on $\bUs$.
+\begin{Def}
+Let $\btUs$ be the regular cover of $\bUs$ with fundamental group $\pi_1(\btUs)$ and covering map $q_1:\btUs\rightarrow\bUs$ that induces a monomorphism between their fundamental groups such that $q_{1*}\pi_1(\btUs)=\ker\varphi_2\vert_{\ker\varphi_1}$.
+\end{Def}
+
+It follows that $\ker\varphi_2\vert_{\ker\varphi_1}=\ker\varphi_1\vert_{\ker\varphi_2}$ since they are taken to abelian groups, and that $\btUs$ is a cover of $\bMs$ as well.
+We call this projection $q_2:\btUs\rightarrow\bMs$, where it induces a monomorphism such that $q_{2*}\pi_1(\btUs)=\pi_1(\bMs)$.
+
+\begin{thm}
+Let $\varphi=\varphi_1\times\varphi_2:\pi_1(\bGs)\rightarrow\ZZ^2\times\SO(2,\ZZ)$.
+Then $\varphi$ is surjective, and the deck group of the cover $(p_1\circ q_1):\btUs\rightarrow\bGs$ is isomorphic to $\ZZ^2\times\SO(2,\ZZ)$.
+\begin{proof}
+This follows from the surjectivity of $\varphi_1$, and $\varphi_2$ as generators of both co-domains are mapped to.
+Since these are compositions of regular covers, $\btUs$ is a regular cover of $\bGs$.
+The sets of loops that lift to closed curves on $\btUs$ must close on both $\bMs$ and $\bGs$ and their homotopy classes have to belong to the kernel of both homomorphisms.
+As such, they would belong to the kernel of $\varphi$ as well.
+Therefore, $\pi_1(\bGs)/(p_{1*}\circ q_{1*})\pi_1(\btUs)=\pi_1(\bGs)/(p_{2*}\circ q_{2*})\pi_1(\btUs)=\varphi(\pi_1(\bGs))=\ZZ^2\times\SO(2,\ZZ)\cong \text{Deck}_{p_1\circ q_1}(\btUs)$.
+\end{proof}
+\end{thm}
+
+\compav{Might remove this:}
 
@@
 The key connection made between surfaces $\bUs$ and $\btUs$ is that geodesics on $\bUs$ and $\btUs$ behave identically.
-Unless stated otherwise, a \emph{geodesic} is a \emph{unit-speed} locally distance minimizing curve on a smooth topological surface as a function parameterized by an element in $\RR$.
+For us, a \emph{geodesic}, $\gamma$, is a \emph{unit-speed} locally distance minimizing curve on a smooth surface as a continuous function parameterized by an element in $\RR$.
 A \emph{geodesic segment} is a restriction of the geodesic to some closed interval in $\RR$.
-A geodesic has an initial trajectory defined at time 0, represented by a unit vector $v_0\in\RR^2$.
-If the vector can be scaled by $k\in\RR$ such that $kv_0\in\cS$, the geodesic has a \emph{rational} initial trajectory.
-For example, the following is a periodic geodesic on the surface with initial direction $kv_0\in\cO$:
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=4in.]{closed2.png}
-\caption{Periodic geodesic path on $\bUs$ modeled with sage-flatsurf.}
+$\gamma$'s derivative is an element of $\RR^2$ in the tangent space at time $t$, denoted $\gamma'(t)\in T_{\gamma(t)}$.
+To avoid confusion, we say a geodesic segment is \emph{closed} if it returns to the initial point with the same derivative.
+Otherwise it may self-intersect in a different direction.
+If the vector can be scaled by $k\in\RR$ such that $kv_0\in\cS$ (defined in the introduction), the geodesic has a \emph{rational} initial trajectory angle.
+For example, the following is a periodic geodesic on the surface with initial direction $kv_0=(51,31)\in\cO$:
+\begin{figure}[H]
+\centering
+\includegraphics[width=4in]{closed2.png}
+\caption{Closed geodesic on $\bUs$ modeled with sage-flatsurf.}
 \label{fig:complicated}
 \end{figure}
-
-We refer later to the following commuting diagram of branched covering maps and inclusion maps. It is meant to illustrate the relationship between all of these spaces.
-
-\begin{equation}
-\begin{tikzcd}
-\arrow[two heads, bend left]{rrr}{f^\circ}[swap]{} \arrow[hook]{d}{pr_b} \btUs \arrow[two heads]{r}{r^\circ} & \bUs \arrow[hook]{d}{pr_c} \arrow[two heads]{r}{q^\circ} & \bGs \arrow[hook]{d}{pr_d} &\arrow[two heads]{l}{w^\circ} \bMs \arrow[hook]{d}{pr_a}  \\
-\arrow[two heads, bend right]{rrr}{}[swap]{f} \btU \arrow[two heads]{r}{r} & \bU \arrow[two heads]{r}{q} & \bG & \arrow[two heads]{l}{w}\bM  \\
-\end{tikzcd}
-\label{eq:commute}
-\end{equation}
-
-Let $x_0\in G^\circ$ and fix $u_0\in\bUs$, an element in the fiber of the covering map.
-A \emph{geodesic} on $G^\circ$ is a closed unit-speed straight-line path given as $\eta:\RR\rightarrow\bGs$.
-The \emph{period}/\emph{length} of the geodesic is some strictly positive $b\in\RR$.
-Let $\eta|_{[0,b]}=\alpha:[0,b]\rightarrow\bGs$ be a closed geodesic segment on $\bGs$ with initial and terminal point $x_0$.
-Denote $\alpha$'s unique lift to $\bUs$ by $\gamma:[0,b]\rightarrow\bUs$, where $\gamma(0)=u_0$, and $\gamma(b)$ in the fiber over $x_0$.
-Fix a point $\utild\in\btUs$ in the fiber over $u_0$.
-We define the unique lift of $\gamma$ to be the curve $\tilde{\gamma}:[0,b]\rightarrow\btUs$ with $\tilde{\gamma}(0)=\utild$.
-Denote their homotopy class representatives by $[\alpha]$ and (if they exist) $[\gamma],~[\tilde{\gamma}]$.
-
-The geodesic segment on $\bGs$ has the property that it returns to $x_0$ with a trivial effect on the holonomy of the curve.
-On the other hand, the lifted paths will close only if their homotopy classes belong in their fundamental groups, i.e.
-the kernel of homomorphisms $\varphi_1$, $\varphi_2|_{\ker\varphi_1}$.
-Otherwise they are mapped non-trivially to $\ZZ^2$ or $\ZZ^2\times\SO(2,\ZZ)$.
-Let $v_0\in\RR^2$ be a unit vector such that $\alpha's$ derivative is $v_0$ at times $0$ and $b$.
-The derivative of $\gamma$ is also $v_0$ at these times as per Corollary 1, as $\alpha$ is a geodesic with trivial holonomy.
-$\gamtild'$ are also $v_0$ at their endpoints.
-
-The Deck group of the map $\btUs\rightarrow\bUs$ is $\SO(2,\ZZ)$.
-This deck group acts naturally on the tangent bundles via push-forward of rotational matrices.
-We obtain the identity $T^1\btUs/\SO(2,\ZZ)\cong T^1\bUs$.
-This action is \emph{non-trivial}.
-Essentially, this means that there are 4 choices of a vector on the four-fold cover given a vector $v_0$ on $\bUs$.
-The choice of vector does not matter in the case where we flow it along a geodesic path, as a geodesic must return to a point in the fiber of $u_0$ in the same direction.
-That is to say that no matter where we start by lifting this path, we will always end up on the same plane (copy of $\mathbf{P}$) that we started in.
-In the case where a geodesic is periodic on $\bUs$, it would return to the same point in the fiber of $u_0$ as well.
-For simplicity's sake we will just lift to the vector mod the identity matrix, and call it $v_0$ as well.
-
-\begin{Def}
+This is in fact the same geodesic as in Figure [??].
+When flattened out, it becomes evident that once a choice of angle has been made, there are only four possible directions (when viewed locally) this path can travel in.
+
+\begin{Def}
+Let $\rho_t:\RR\rightarrow\bGs$ be a unit-speed geodesic on the \emph{Necker cube surface} with $\rho(0)=x_0$ and $\rho'(0)=v_0$.\\
 Let $\Phi_t:\RR\rightarrow\bUs$ be a unit-speed geodesic on the \emph{Necker cube surface} with $\Phi(0)=u_0$ and $\Phi'(0)=v_0$.
-\\Similarly, let $\tilde{\Phi}_t:\RR\rightarrow\btUs$ be a unit-speed geodesic on $\bUs$'s four-fold cover with $\phitild(0)=\utild$ and $\phitild'(0)=v_0$.
-\end{Def}
-Let $\varphi=\varphi_1\times\varphi_2:\pi_1(\bGs)\rightarrow \ZZ^2\times\SO(2,\ZZ)$ be given by $[g]\mapsto(\varphi_1([g]),\varphi_2([g]))$. $\varphi$ is a
-product of group homomorphisms that takes a loop to 
+\\Similarly, let $\tilde{\Phi}_t:\RR\rightarrow\btUs$ be a unit-speed geodesic on $\bUs$'s four-fold cover with $\phitild(0)=\utild$ and $\phitild'(0)=v_0$.\\
+Here $u_0$ is a fixed point in $p_1^{-1}(x_0)$ and $\utild$ is a fixed point in $q_1^{-1}(u_0)$. $v_0\in\RR^2$ is a unit-vector in the tangent spaces $T_{x_0}$, $T_{u_0}$, and $T_{\utild}$.
+\end{Def}
+
+\begin{thm}[Existence of Rational Closed Geodesics on $\bGs$]
+Let $\rho_t:\RR\rightarrow\bGs$ be a geodesic with $k\rho'(0)\in\cS$ for some $k\in\RR$. Then $\rho$ is periodic.
+\begin{proof}
+Let $\rho_(0)=x_0$.
+Let $m_0\in p_2^{-1}(x_0)\subset\bMs$, and $\tilde{\rho}_t$ be a geodesic on $\bMs$ based at $m_0$ with derivative $\rho'(0)$ lifted from $\rho$.
+Since $\bMs$ is a square-tiled translation surface, it is obtainable as the cover of a square torus ramified over a single point.
+[Veech dichotomy] Therefore the surface is Veech, and $\tilde{\rho}_t$ is either periodic or uniquely ergodic.
+The initial angle is rational, so it must close.
+Since it closes on $\bMs$, it must close on $\bGs$ as well.
+\end{proof}
+\end{thm}
+
+Since $\rho_t$ is periodic, we will denote its restriction to a single period by $\eta(t)=\rho\vert_{[0,T]}(t)$ for $T\in\RR^+$.
+Here $\eta$ is a closed geodesic segment that returns with the same derivative.
+
 \begin{thm}
-Let $\delta=\varphi_1([\alpha])\in\ZZ^2$ be an element of $\Delta_{\varphi_1}$ and $\tilde{\delta}=\varphi([\alpha])\in\ZZ^2\times \SO(2,\ZZ)$ be an element of $\Delta_{\varphi}$. Then:\\
-$\Phi(t+b)=\delta\cdot\Phi(t)$ for all $t\in\RR$, and 
-$\phitild(t+b)=\tilde{\delta}\cdot\phitild(t)$ for all $t\in\RR$.
-\end{thm}
+Let $\eta$ be as above. Let $\alpha\in\pi_1(\bGs)$ be its homotopy class representative. Then the following is true:
+\begin{enumerate}
+\item $\Phi(t+T)=\varphi_1(\alpha)\cdot\Phi(t)$ for all $t\in\RR$
+\item $\phitild(t+T)=\varphi(\alpha)\cdot\phitild(t)$ for all $t\in\RR$.
+\end{enumerate}
+\begin{proof}
+Let $\tilde{\alpha}$ be a lifted geodesic segment on $\bUs$ based at $u_0$ such that $p_1(\tilde{\alpha}([0,T]))=\eta([0,T])$.
+Let $\tilde{\tilde{\alpha}}:\RR\rightarrow\bUs\times\ZZ$ be defined as $\tilde{\tilde{\alpha}}(t)=(\alpha^n\cdot\tilde{\alpha}(t~(mod~T)),n)$, where $n=\left\lfloor \frac{t}{T}\right\rfloor$ and $\alpha$ acts by $\varphi_1(\alpha)\in\ZZ^2$.
+Let $\iota:\bUs\times\ZZ\twoheadrightarrow\bUs$ be the projection given by $(x,z)\mapsto x$.
+By induction on $\ZZ$, we can show this is a path that extends to the real numbers and that, by uniqueness of geodesics on manifolds, $\iota\circ\tilde{\tilde{\alpha}}=\Phi(t)$.
+Therefore, $\Phi(t+T)=\alpha^{n+1}\cdot\tilde{\alpha}((t+T)~(mod~T))=\alpha\alpha^n\cdot\tilde{\alpha}(t~(mod~T))=\alpha\cdot\Phi(t)$.
+\\
+The extension to $\RR$ for $\tilde{\Phi}$ is done the same way, and the proof in that case is analogous.
+\end{proof}
+\end{thm}
+
+\begin{lem}
+Let $\alpha,\eta$ be as above. Then $\varphi_2(\alpha)=I_2$.
+\begin{proof}
+Otherwise $\eta$ would not be a geodesic.
+\end{proof}
+\end{lem}
 
 \begin{cor}
-The following are equivalent:
+Let $\alpha$ be as above. If $\varphi_1(\alpha)\neq 0$, then $\Phi$ and $\phitild$ are drift-periodic.
+\begin{proof}
+Since $\varphi_1(\alpha)$ is isomorphic to an element $f_\alpha\in\text{Deck}_{p_1}(\bUs)$, we need only show that $f_\alpha$ is a local diffeomorphism of $\bUs$.
+But that much is obvious as neighborhoods are isometric to $\RR^2$ on its smooth sections and $p_1\circ f_\alpha=p_1$.
+Since $f_\alpha$ is a diffeomorphism,
+\compav{not sure how to do this}
+\end{proof}
+\end{cor}
+
+\begin{cor}
+Let $\alpha$, $\eta$ be as above. Then following are equivalent:
 \begin{enumerate}
@@
 \item $\phitild$ is periodic.
-\item $\delta=(0,0)$.
+\item $\varphi(\alpha)=(0,0)$.
 \end{enumerate}
+\begin{proof}
+Suppose $\Phi$ is periodic.
+Then $\Phi\vert_{[0,T]}$ is a closed segment with trivial holonomy lifted from $\eta$.
+Let $\alpha\in\pi_1(\bGs)$ be the homotopy class of $p_1(\Phi)=\eta$.
+From theorems [??] and [??], and lemma [??], $\varphi(\alpha)=((0,0),I_2)$ and so $\phitild(t+T)=\varphi(\alpha)\cdot\phitild(t)=\phitild(t)$ for all $t\in\RR$.\\
+Now suppose $\phitild$ is periodic.
+Then $\varphi_1(\alpha)=(0,0)$ since $\varphi(\alpha)$ is trivial.\\ Now suppose $\varphi(\alpha)=(0,0)$.
+Then $\Phi(t+T)=\Phi(t)$ for all $t\in\RR$, and so $\Phi$ is periodic.
+\end{proof}
 \end{cor}
@@
 \begin{cor}
-The following are equivalent:
+Let $\alpha$, $\eta$ be as above. The following are equivalent:
 \begin{enumerate}
@@
 \item $\phitild$ is drift-periodic.
-\item $\delta\neq(0,0)$.
+\item $\varphi_1(\alpha)\neq(0,0)$.
 \end{enumerate}
+\begin{proof}
+Suppose $\Phi$ is drift-periodic. Then $p_1(\Phi\vert_{[0,T]})=\eta$ belongs to $\alpha\in\pi_1(\bGs)$ with $\varphi_1(\alpha)\neq (0,0)$, and so from theorem [??] $\phitild$ is drift-periodic.\\ 
+Now suppose $\phitild$ is drift-periodic. From lemma [??], we have that $\phi_2(\alpha)=I_2$. Therefore, $\phi_1(\alpha)\neq (0,0)$ or else $\phitild$ would be closed.\\
+Suppose $\varphi_1(\alpha)\neq (0,0)$.
+Then $\varphi_1(\alpha)$ induces some nontrivial local diffeomorphism $f_\alpha\in\text{Deck}_{p_1}(\btU)$ such that $\Phi(t+T)=\varphi_1(\alpha)\cdot\Phi(t)=f_\alpha(\Phi(t))$ for all $t\in\RR$, and the geodesic drifts by some constant automorphism of the surface.
+\end{proof}
 \end{cor}
-
-These theorems essentially tell us that these types of geodesics behave identically on the Necker cube surface and its cover.
-Now we show that this is true for $\btUs$ and its metric completion, $\btU$.
-
-\begin{equation}
-\begin{tikzcd}
-\arrow[hookrightarrow, bend left]{rrr}{f_*^\circ}[swap]{} \arrow[two heads]{d}{PR_b} \pi_1(\btUs) \arrow[hook]{r}{r_*^\circ} & \pi_1(\bUs) \arrow[two heads]{d}{PR_c} \arrow[hook]{r}{q_*^\circ} & \pi_1(\bGs) \arrow[two heads]{d}{PR_d} & \ar[hookrightarrow]{l}{}[swap]{w_*^\circ} \pi_1(\bMs) \arrow[two heads]{d}{PR_a} \\
-\arrow[hookrightarrow, bend right]{rrr}{}[swap]{f_*}  \pi_1(\btU) \arrow[hook]{r}{r_*} & \pi_1(\bU) \arrow[hook]{r}{q_*} & \pi_1(\bG) & \ar[hookrightarrow]{l}{}[swap]{w_*} \pi_1(\bM)\\
-\end{tikzcd}
-\label{eq:homomorphism}
-\end{equation}
 
@@
 
-This follows readily from the fact that geodesics are undefined on cone singularities. A geodesic on $\btUs$ can be mapped bijectively to a geodesic on  $\btU$ and vice-a-versa.
+This follows readily from the fact that geodesics are undefined on cone singularities.
+A geodesic on $\btUs$ can be mapped bijectively to a geodesic on  $\btU$ and vice-a-versa.
+Any automorphisms of $\btU$ can be restricted to $\btUs$, and the notions of periodicity/drift-periodicity of the geodesics are equivalent.
 
 \subsection{Translation Surface}
-Consider the projection $\btU\rightarrow\mathbf{M}$ onto a quotient of the surface under its translational symmetries.
-What we obtain is the following compact translation surface belonging to the stratum $\cH(2,2,2,2):$
-
-\begin{figure}[H]
-\centering
-\includesvg[width=2.6in]{mtildabw}
+The surface $\bMs$ was obtained as a cover of $\bGs$ from the rotational monodromy map on $\pi_1(\bGs)$.
+$\bM$ is its metric completion by reinsertion of its cone singularities.
+There are four cone singularities on this surface, as opposed to the three on $\bG$, each of cone angle $6\pi$.
+The individual sections of $\bM$, labeled $I,II,III,$ and $IV$ are rotated in such a way to make it evident that this is indeed a translation surface.
+(Fig ??) We call the maximal translational $\bM$ atlas $\cA$.
+Though not immediately apparent, $(\bM,\cA)$ is a square-tiled origami surface.
+(Fig ??)
+
+\begin{figure}[H]
+\centering
+\includesvg[width=2.9in]{mtildabw}
 \caption{Compact translation surface, $\mathbf{M}$ with edges and cone singularities (1,2,3,4) identified.
-The Roman numerals are meant to identify every ``copy" of $\bG$ with a direction differing from $v_0$ by rotation of some element in $\SO(2,\ZZ)$.
-The individual copies are then rotated so we have a canonical notion of direction }
+The Roman numerals are meant to identify every ``copy" of $\bG$ with a vector differing from $v_0$ by rotation of some element in $\SO(2,\ZZ)$.
+The individual copies are then rotated so that every edge identification is made by translation.}
 \label{fig:mtilda}
@@
 
-We rearrange $\bM$ as such:
-
-\begin{figure}[H]
-\centering
-\includesvg[width=3.4in]{mtildastaircasebw}
-\caption{The staircase surface with directional planes and vertices identified.
+\begin{figure}[H]
+\centering
+\includesvg[width=3.8in]{mtildastaircasebw}
+\caption{The staircase surface with roman numerals identifying each section of $\bGs$ in the cover.
+Edges and cone singularities are labeled.
 All edges are paired by translation.
-Two adjacent squares have opposite edges (A-M) identified.
-The top edge of the bottom-left square is glued to the bottom edge of the top-right square (both labeled I).
-Likewise, the bottom edge of the bottom-left square is identified with the top edge of the top-right square.}
+Two adjacent squares have opposite edges (A through M) identified.}
 \label{fig:staircase}
@@
 
-\noindent We recall some basic definitions and theorems concerning $\ZZ^2$covers of translation surfaces as they apply to $\bM$.
-
-\begin{Def}
-Algebraic intersection number is a non-degenerate bilinear form:
-\begin{align*}
-i:H_1(\bM,\QQ)\times H_1(\bM,\QQ)\rightarrow\QQ,
-\end{align*}
-for $[\gamma],[\beta]\in H_1(\mathbf{M},\QQ)$, $i([\beta],[\gamma])$ returns the signed intersection number of two homology classes.
-We say a crossing at the moment of intersection is positive if $\gamma$ makes a positive angle relative to $\beta$.
-\end{Def}
-The set of all affine diffeomorphisms of $\bM$ form the group $\text{Aff}^+(\bM)$.
-The $\emph{Veech group}$ of $\bM$ is the image in the co-domain of the homomorphism $D:\text{Aff}^+(\bM)\rightarrow SL(2,\RR)$ that takes every affine map to its derivative, which we denote $V(\bM)$.
-$\bM$ is a square-tiled translation surface whose Veech group is a finite index subgroup of $SL(2,\ZZ)$, thereby making it a \emph{Veech surface}.
-We use automorphisms of the fundamental group in $\text{Aut}(\pi_1(\bM))$ induced by affine maps to prove our main results.
-But first, we define a group homomorphism from $\pi_1(\bM)$ to $\pi_1(\bM)/\pi_1(\btU)$ by algebraic intersection numbers over a sum of linearly independent homology classes.
-Consider the following cylinder core curves on $\bM$ labeled $\gamma_i$ for $0\leq i\leq 11$:
-\begin{figure}[H]
-\includegraphics[width=3in]{homologyclass.png}
-\centering
-\caption{$\bM$'s cylinder core curves with u,v, and z homology class labels.}
+The lines contained in the squares are the outermost edges in the original image and play an important role in determining the cover, $\btU$.
+We will orient $\bM$ so that the orientation of its image in $\CC$ under $\cA$ is respected.
+An orientation preserving affine diffeomorphism of $\bM$ has a derivative in $\CC$ uniform throughout the surface as an element of $\SL(2,\RR)$.
+The set of these affine diffeomorphisms form the group $\text{Aff}_+(\bM)$.
+The map $D:\text{Aff}_+(\bM)\rightarrow\SL(2,\RR)$ factors through $\cA$ and is a homomorphism between these groups.
+The affine group's image in $\SL(2,\RR)$ is itself a subgroup known as the \emph{Veech group} of $\bM$, which we denote $V(\bM)$.
+If $V(\bM)$ is of finite index in $\SL(2,\RR)$, we say that $\bM$ is a \emph{Veech surface}.
+A theorem of Gutkin-Judge states that any compact, square-tiled translation surface is in fact Veech.
+$\bM$  has am obvious construction as a branch cover of the square torus, ramified over a single point.\\
+
+\noindent There are two transverse cylinder decompositions of this surface in vertical and horizontal directions.
+In each decomposition, there are is a cylinder waist curve running along the circumference of the cylinder.
+We use $\gamma_i$ to indicate which waist curve we are referring to.
+(Fig [??])
+
+\begin{figure}[H]
+\includegraphics[width=4in]{homologyclass.png}
+\centering
+\caption{$\bM$'s waist curves with u,v, and z homology class labels.}
 \label{fig:homology}
@@
 
-The rank of a $12\times12$ intersection matrix of these curves is 10. Therefore these curves span $H_1(\bM,\QQ)$.
-
-\begin{Def}
-The homology classes $u,v,z$ are given as the following sums of core curves:
+We will use the first homology of the surface, $H_1(\bM,\QQ)$, to determine the cover.
+These curves associate an even indexed waist curve with a horizontal cylinder and an odd indexed waist curve with a vertical cylinder, a construction that becomes useful later on.
+Let $[\gamma_i]\in H_1(\bMs,\QQ)$ be the homology class corresponding to $\gamma_i$.
+We use $\Gamma=\{[\gamma_i]~:~0\leq i \leq 11\}$ to denote the set of these classes.
+
+\begin{Def}
+The homology classes $u,v,z$ are given as the following weighted sums of elements of $\Gamma$:
 \begin{align*}
@@
 \begin{Def}
-Define the group homomorphism $\Omega_{u,v}:\pi_1(\bM)\rightarrow \ZZ^2$, where $\beta\mapsto(i(u,[\beta]),i(v,[\beta]))$.
-\end{Def}
+Algebraic intersection number on a differentiable, orientable manifold is an integer value obtained as a scalar product of two elements in its first homology.
+\begin{align*}
+\<\<\cdot~,~\cdot\>\>:H_1(\bM,\QQ)\otimes H_1(\bM,\QQ)\rightarrow\ZZ
+\end{align*}
+Let $\alpha,\beta$ be two closed curves on $\bM$, and let $\overline{\alpha},\overline{\beta}$ be their respective homology classes.
+For $\overline{\alpha},\overline{\beta}\in H_1(\mathbf{M},\QQ)$, $\<\<\alpha,\beta\>\>$ returns the signed intersection number of two homology classes as a sum of intersections of the two curves, where  an intersection is positive (+1) if $\beta$ makes a positive angle relative to the direction $\alpha$ is traveling in.
+A crossing is negative (-1) if $\beta$ makes a negative angle relative to $\alpha$.
+$\bM$ is oriented so thats its image in $\CC$ under translational charts agrees with the orientation of $\CC$.
+\end{Def}
+
+It's known that the intersection number is a non-degenerate, bilinear form. We use this fact to find a basis for $H_1(\bM,\QQ)$.
 
 \begin{lem}
-\compav{show somehow that u,v determine the cover}
+$\Gamma$ generates $H_1(\bM,\QQ)$.
+\begin{proof}
+Let $E_{i,j}:=\<\<[\gamma_j],[\gamma_i]\>\>$ for $i,j\in \{0,1,2,\cdots,11 \}$.
+$E$ is an intersection matrix of rank 10.
+Since intersection number is non-degenerate, $\Gamma$ must span all of $H_1(\bM,\QQ)$.
+\end{proof}
 \end{lem}
 
+\begin{cor}
+The set $\Gamma'=\Gamma\backslash\{[\gamma_7],~[\gamma_{10}] \}$ forms a basis for $H_1(\bM,\QQ)$. Further,
+\begin{align}
+[\gamma_7]&=[\gamma_1]-[\gamma_3]+[\gamma_5]+[\gamma_9]-[\gamma_{11}],\\~and~
+[\gamma_{10}]&=[\gamma_0]-[\gamma_2]+[\gamma_4]-[\gamma_6]+[\gamma_8].
+\end{align}
+\begin{proof}
+Let $M_{i,j}:=\<\<[\gamma_j],[\gamma_i]\>\>$ for $i,j\in \{0,1,2,\cdots,11 \}\backslash\{7,10\}$.
+Then $\det M=10$, proving the first statement.
+\\\\Now since intersection is  non-degenerate, if we intersect $[\gamma_7]$ and $[\gamma_{10}]$ with $\sum\Gamma$, a spanning set of $H_1(\bM,\QQ)$, we obtain $\<\<\sum\Gamma,[\gamma_7]\>\>=$\\\\
+(Might not use (5) and (6)).\\\\
+To prove the second claim it suffices to show that passing the RHS of (5) and (6) through the matrix $E_{i,j}$ will equate them to $[\gamma_7]$ and $[\gamma_{10}]$ respectively.
+\end{proof}
+\end{cor}
+
+\begin{cor}
+The sets $u$ and $v$ are linearly independent.
+\begin{proof}
+Obviously follows.
+\end{proof}
+\end{cor}
+
+\begin{Def}
+Fix $m_0\in\bMs$ and let $\beta\in\pi_1(\bM,m_0)$. The map $\Omega_{u,v}:\pi_1(\bM)\rightarrow \ZZ^2$ is given as
+\begin{align*}
+\Omega_{u,v}([\beta])=(~\<\<u,\beta\>\>~,~\<\<v,\beta\>\>~)
+\end{align*}
+\end{Def}
+
+This homomorphism takes $\pi_1(\bM)$ to an abelian group, so it can be factored through $H_1(\bM,\QQ)$, and a cover of $\bM$ is obtained from the kernel of this map.
+This is a special case when it comes to regular covers with abelian Deck groups.
+
+\begin{thm}
+Let $\pi_1(\btU,\utild)$ be the fundamental group of $\btU$ with $\utild\in (q_1\circ p_1)^{-1}(x_0)$ and $x_0\in\bGs$, and $m_0\in p_2^{-1}(x_0)$.
+Let $p_r:\btU\twoheadrightarrow\bM$ be a projection of surfaces by taking a $\ZZ^2$ quotient of its translational symmetries.
+Then $p_r$ induces a map $p_{r*}:\pi_1(\btU,\utild)\rightarrow\pi_1(\bM,m_0)$ such that $p_{r*}\pi_1(\btU)=\ker\Omega_{u,v}$.
+\begin{proof}
+The curves making up $u$ and $v$ are the outermost edges of sections $I,~II,~III,~IV$.
+(Fig 9) By observation, these classes are determined and weighted according to which section in the fiber over $\bM$ a lifted path ends up on after a positive or negative crossing.\\\\\\
+Might need to redo this, factor through null space of $\Omega$ matrix.
+\end{proof}
+\end{thm}
+
 \subsection{Induced Automorphisms of $H_1(\bM,\QQ)$}
-We look at some important automorphisms induced by Affine maps on $\bM$.
-Observe that $\bM$ has a uniform cylinder decomposition in both horizontal and vertical directions as in figure $15$.
-We define the \emph{modulus}, $\mu$, of a cylinder to be the ratio of the cylinder's width to its circumference, $\frac{w}{c}$.
-The $\emph{Dehn-twist}$ of a cylinder is an affine diffeomorphism that skews the cylinder and sends every vertex to itself.
+We have previously remarked on the uniformity of $\bM$'s cylinder decompositions, $\mathfrak{H},~\mathfrak{V}$, but have yet to show the correlation between these cylinders and $\text{Aff}^+(\bM)$.
+We say the \emph{modulus}, $\mu_i$, of the cylinder containing the $i^{th}$ waist curve is the ratio of its width to its circumference, $\frac{w_i}{c_i}$.
+The $\emph{Dehn-twist}$ of a cylinder in either set is an affine diffeomorphism that skews the cylinder and preserves edges and vertices.
+The skew is in the direction of the waist curve.
 
@@
 \label{fig:skew}
-\caption{Dehn-twist of a cylinder in $\bM$'s cylinder decomposition.}
-\end{figure}
-
-Such a map has derivative $\left[~\begin{matrix}1 && \pm\mu^{-1}\\0 && 1\end{matrix}~\right]$.
-On $\bM$ all cylinders in both vertical and horizontal decompositions have moduli $\mu=\frac{1}{2}$.
-These give way to global diffeomorphisms as $\emph{multi-twists}$ of $\bM$.
-
-\begin{Def}
-We define the global affine diffeomorphisms of $\text{Aff}^+(\bM)$ obtained as \textbf{multi-twists of the surface in horizontal and vertical directions} $\psi_h$ and $\psi_v$, respectively.
-\end{Def}
-
-The \emph{derivative map} of the affine group is the homomorphism $D:\text{Aff}^+(\bM)\rightarrow \SL(2,\RR)$.
-The \emph{Veech group} of the surface is the image of the affine group under this map, $D(\text{Aff}^+(\bM))$, denoted $V(\bM)$.
-The twists $\psi_h$ and $\psi_v$ under $D$ are mapped to the same matrices as the Dehn-twists on its individual cylinders. i.e.
+\caption{Dehn-twist of a cylinder in $\bM$'s horizontal cylinder decomposition.}
+\end{figure}
+
+In local coordinates this Dehn-twist has derivative $\left[~\begin{matrix}1 && \pm\mu_{2i}^{-1}\\0 && 1\end{matrix}~\right]$.
+On $\bM$ the cylinders in both vertical and horizontal decompositions have moduli $\mu_i=\frac{1}{2}$.
+Dehn-twists on $\mathfrak{H}$ and $\mathfrak{V}$ give up global diffeomorphisms as $\emph{multi-twists}$ of $\bM$ with the same derivatives.
+[cite]
+
+\begin{Def}
+We denote the global affine diffeomorphisms of $\text{Aff}^+(\bM)$ obtained as \textbf{multi-twists} of the surface in \textbf{horizontal and vertical directions} $\psi_h$ and $\psi_v$, respectively.
+\end{Def}
+
+The images of the twists $\psi_h$ and $\psi_v$ under the derivative map $D$ are taken to the same matrices as the Dehn-twists in local coordinates. i.e.
 \begin{align*}
@@
 \begin{Def}
-We call $\XX$ the generated parabolic subgroup of $\text{Aff}^+(\bM)$, where $\XX=\<\psi_h,\psi_v \>$.
-\end{Def}
-
-The image of this group under $D$ is itself a subgroup of $V(\bM)$ and denoted $D(\XX)=\XX'=\<\psi_h',\psi_v' \>$ as in the introduction.
-Note that when skewing the surface and piecing it back together, the cylinder curves intersect each other:
+Let $\XX$ be the generated parabolic subgroup of $\text{Aff}^+(\bM)$, where $\XX=\<\psi_h,\psi_v \>$.
+\end{Def}
+
+The image of this group under $D$ is itself a subgroup of $V(\bM)$, $\XX'=D(\XX)=D(\<\psi_h,\psi_v \>)=\<D(\psi_h),D(\psi_v)\>=\<\psi_h',\psi_v' \>$.
+When skewing the surface and piecing it back together under $\psi_h$ or $\psi_v$, the cylinder curves intersect each other in pairs:
 
@@
 By skewing in the horizontal direction, $\psi_h$ preserves the even indexed cylinder curves, but the odd indexed vertical curves $\gamma_i$ intersect the two horizontal curves (positive intersection), $\gamma_{i-1}$ and $\gamma_{i+1}$.
-A similar observation can be made from applying $\psi_v$ where the vertical curves are preserved, but the horizontal curves intersect with adjacent vertical curves (negative intersection).
+A similar observation can be made from applying $\psi_v$, where the vertical curves are preserved but the horizontal curves intersect with adjacent vertical curves (negative intersection).
+Let $\Aut(H_1(\bM,\QQ))$ be the group of automorphisms of $\bM$'s homology.
+By extension of these affine maps to the automorphism group, we may express the \emph{induced} automorphisms as they affect the homology generating set $\Gamma$.
+In turn, these automorphisms act linearly on a cycle $\mathfrak{c}=\sum_{i=0}^{11}k_i[\gamma_i]\in H_1(\bM,\QQ)$.
 Let $i\in\ZZ/12\ZZ$.
-Then this convenient indexing scheme gives us the following formulas for $\psi_h$ and $\psi_v$'s counterparts in $\Aut(H_1(\bM,\QQ))$ raised to $k\in\ZZ\backslash\{0\}$:
+Then the following formulas hold for $\psi_h$ and $\psi_v$'s counterparts in $\Aut(H_1(\bM,\QQ))$ raised to $k\in\ZZ$:
 
@@
 
-We call this generated group $\XX^*=\<\psi^{*}_h,\psi^{*}_v \>$.
-There are no relations between its generators since $i$ being even or odd has one act trivially and the other non-trivially.
-Therefore this group is also free and rank 2.
-Hence $\XX^*\cong\XX'\cong\XX$.
-Let $g\in\XX$, $g'\in\XX'$, $g^*\in\XX^*$ be isomorphic elements in their respective groups.
-Let $g=\prod_{j=0}^{n}\psi_h^{k_j}\circ\psi_v^{m_j}$ be a word in $\XX$ as a finite left product of multi-twist pairs with $k_j,m_j\in\ZZ$.
-Let $\psi_j=\psi_h^{k_j}\circ\psi_v^{m_j}$ be the $j^{th}$ word pair.
-Elements $g',g^*$ have identical representations.
+These are used to generate the group $\XX^*=\<\psi^{*}_h,\psi^{*}_v \>$.
+There are no relations between its generators since $i$, being either even or odd, has one automorphism act trivially while the other acts non-trivially.
+Therefore this group is also a free group of rank 2.
+All of our related groups are rank 2 free groups, so they are isomorphic.
+
+Let $\tau\in\XX$.
+Let $D(\tau)=\tau'\in\XX'$ and $\tau^*\in\XX^*$ be $\tau$'s isomorphic counterparts in their respective groups.
+We say $\tau=\prod_{j=0}^{n}\psi_h^{k_j}\circ\psi_v^{m_j}$ is a word in $\XX$ as a finite left concatenated product of multi-twist pairs with $k_j,m_j\in\ZZ$.
+Let $\psi_j=\psi_h^{k_j}\circ\psi_v^{m_j}$ be the $j^{th}$ multi-twist pair.
+Since $\tau',\tau^*$ are isomorphic, we can use these expressions interchangeably.
 \\\\
-The $j^{th}$ element pair that is mapped into $\Aut(H_1(\bM,\QQ))$ can be expressed as
+The corresponding $j^{th}$ automorphism pair is an element $\psi_j^*\in\Aut(H_1(\bM,\QQ))$ and is expressed as
 
@@
 
-\noindent Let $\sum\Gamma=\sum_{i=0}^{11}[\gamma_i]$, and denote the odd and even indexed sums as $\sum\Gamma_o$ and $\sum\Gamma_e$, respectively.
-We consider three kinds of closed geodesic segments on $\bM$ and don't consider initial/terminal points for the time being.
-These are the slope one, horizontal, and vertical loops.\\
-
-\noindent Call the geodesic in direction $\theta=\frac{\pi}{4}$ that runs up the staircase intersecting every $\gamma_i$ once $\chi$.
-Denote its homotopy class by $[\chi]$.
-Observe from Figure 10 that $\chi$ is homologous to the sum of cylinder curves divided by 2.
-In Figure 1, this is the closed geodesic that lifts to a closed loop that looks like a 12-gon on $\bUs$.
+\noindent Similarly,
+
+\begin{align*}
+\psi_j'=\left[ \hspace{1mm} \begin{matrix}
+				1 &   2\\
+				0 & 1
+			\end{matrix}\hspace{1mm}\right]^{k_j}\left[ \hspace{1mm} \begin{matrix}
+							1 &   0\\
+							2 & 1
+						\end{matrix}\hspace{1mm}\right]^{m_j}=\left[ \hspace{1mm} \begin{matrix}
+										1+4m_jk_j &   2k_j\\
+										2m_j & 1
+									\end{matrix}\hspace{1mm}\right].
+\end{align*}
+
+\noindent Let $\sum\Gamma=\sum_{i=0}^{11}[\gamma_i]$, and denote the odd and even indexed sums of homological waist curves $\sum\Gamma_o$ and $\sum\Gamma_e$, respectively.
+We look at three kinds of closed geodesic segments on $\bM$, not considering initial/terminal points for the time being.
+These are the geodesics in directions $\theta=0,~\frac{\pi}{2},$ and $\frac{\pi}{4}$.
+The aim is to show that these linear automorphisms take $\sum\Gamma\mapsto c_1\sum\Gamma_e+c_2\sum\Gamma_o$, and acts faithfully on the kernel of $\Omega_{u,v}$.
+\\
+
+\noindent Let $\chi$ be the geodesic segment in direction $\theta=\frac{\pi}{4}$ that runs up the staircase intersecting every $\gamma_i$ exactly once.
+Denote its homology class $[\chi]$.
+If we flow $\bM$ in this direction we obtain a two cylinder decomposition of the surface, where $\chi$ is contained in or the other cylinder.
+\begin{figure}[H]
+\centering
+\includesvg[width=5.9in]{slopeonecylinder}
+\caption{Cylinder decomposition with $[\chi]$ being the dashed or dotted geodesic. $[\chi]$ is homologous to $\frac{1}{2}\sum\Gamma$ (solid/curvy)}
+\end{figure}
+Observe from the previous figure that $\chi$ is homologous to $\frac{1}{2}\sum\Gamma$, a curve that climbs up the staircase by traversing each waist curve exactly halfway ( their intersection number is 0).
+When lifted to $\btU$ and projected onto $\bUs$, this is a closed geodesic on the Necker cube surface (Fig 1).
 
@@
 \begin{proof}
-Observe that $\Omega_{u,v}([\chi])=\Omega_{u,v}(\frac{1}{2}\sum\Gamma)=\frac{1}{2}(i(u,\sum\Gamma),i(v,\sum\Gamma)).$ Since only adjacent curves intersect, $i(u,\sum\Gamma)=i(-[\gamma_2],[\gamma_{1}]+[\gamma_{3}]) +i([\gamma_5],[\gamma_{4}]+[\gamma_{6}]) + i([\gamma_8],[\gamma_{7}]+[\gamma_{9}]) - i([\gamma_{11}],[\gamma_{10}]+[\gamma_{0}])=-2+(-2)+2-(-2)=0$.\\
-Similarly, $i(u,\sum\Gamma)=i([\gamma_0],[\gamma_{11}]+[\gamma_{1}]) -i([\gamma_3],[\gamma_{2}]+[\gamma_{4}]) -i([\gamma_6],[\gamma_{5}]+[\gamma_{7}]) +i([\gamma_9],[\gamma_{8}]+[\gamma_{10}])=2-(-2)-2+(-2)=0$.
+Consider $\Omega_{u,v}([\chi])=\Omega_{u,v}(\frac{1}{2}\sum\Gamma)=\frac{1}{2}(i(u,\sum\Gamma),i(v,\sum\Gamma))$.
+\\Since only adjacent curves intersect,\\ $i(u,\sum\Gamma)=i(-[\gamma_2],[\gamma_{1}]+[\gamma_{3}]) +i([\gamma_5],[\gamma_{4}]+[\gamma_{6}]) + i([\gamma_8],[\gamma_{7}]+[\gamma_{9}]) - i([\gamma_{11}],[\gamma_{10}]+[\gamma_{0}])\\=-2+(-2)+2-(-2)=0$.\\
+Similarly,\\ $i(u,\sum\Gamma)=i([\gamma_0],[\gamma_{11}]+[\gamma_{1}]) -i([\gamma_3],[\gamma_{2}]+[\gamma_{4}]) -i([\gamma_6],[\gamma_{5}]+[\gamma_{7}]) +i([\gamma_9],[\gamma_{8}]+[\gamma_{10}])\\=2-(-2)-2+(-2)=0$.
 Therefore, $\Omega_{u,v}([\chi])=(0,0)$.
@@
 \end{lem} 
+Since $\chi$ is homologous to this sum of waist curves divided by two, there's a bit of freedom in how we can express a cycle in the orbit of $[\chi]$ in $H_1(\bM,\QQ)$ by seeing how an element in $\XX^*$ acts on $\sum\Gamma$.
 
 \begin{lem}
-Let $g$ be as above. If $\alpha=g(\chi)$, then $[\alpha]=\frac{1}{2}(c_1\sum_e+c_2\sum_o)$ for $c_1,c_2\in\ZZ$.
-\begin{proof}(By Induction on $\XX^*$).\\
+Let $\tau\in\XX$. If $\alpha=\tau(\chi)$ is a closed geodesic segment on $\bM$, then $[\alpha]=\frac{1}{2}(c_1\sum_e+c_2\sum_o)$ for $c_1,c_2\in\ZZ$.
+\begin{proof}(By Induction on $\psi^*_j\in\XX^*$).\\
 Consider $\psi_0(\chi)$.\\
-$[\psi_0(\chi)]=\psi_0^*[\chi]=\frac{1}{2}\psi_0^*(\sum\Gamma)=\frac{1}{2}\psi_0^*(\sum\Gamma_e+\sum\Gamma_o)=\frac{1}{2}(\psi_0^*(\sum\Gamma_e)+\psi_0^*(\sum\Gamma_o))$.
-Now\\ $\psi_0^*(\sum\Gamma_e)=\sum_{i=0}^{11}\psi_0^*([\gamma_{2i}])=\sum_{i=0}^{11}( [\gamma_{2i}]-2m_0k_0[\gamma_{2i}]-m_0k_0[\gamma_{2i-2}] $\\$ -m_0k_0[\gamma_{2i+2}]-m_0[\gamma_{2i+1}]-m_0[\gamma_{2i-1}] )=(1-4m_0k_0)\sum\Gamma_e+(-2m_0)\sum\Gamma_o$.\\
-Now\\ $\psi_0^*(\sum\Gamma_o)=\sum_{i=0}^{11}\psi_0([\gamma_{2i+1}])=\sum_{i=0}^{11}([\gamma_{2i+1}] +k_0([\gamma_{2i}]+[\gamma_{2i+2}]))=\sum\Gamma_o+2k_0\sum\Gamma_e$.\\
-And so $\frac{1}{2}\psi_0^*(\sum\Gamma)=\frac{1}{2}((1+2k_0-4k_0j_0)\sum\Gamma_e +(1-2m_0)\sum\Gamma_o )$.\\\\
-Induction: Suppose $\frac{1}{2}\psi^*_j(\sum\Gamma)=c_1\sum\Gamma_e+c_2\sum\Gamma_o$.\\
-Then $\psi^*_{j+1}(\psi^*_{j}(\sum\Gamma))=c_1\psi^*_{j+1}(\sum\Gamma_e)+c_2\psi^*_{j+1}(\sum\Gamma_o)=\\
-c_1((1-4m_{j+1}k_{j+1})\sum\Gamma_e+(-2m_{j+1})\sum\Gamma_o)+c_2(\sum\Gamma_o+2k_{j+1}\sum\Gamma_e)=\\
-(c_1(1-4m_{j+1}k_{j+1}) + 2k_{j+1}c_2)\sum\Gamma_e+(-2m_{j+1}c_1+c_2)\sum\Gamma_o$.
+$[\psi_0(\chi)]=\psi_0^*([\chi])=\frac{1}{2}\psi_0^*(\sum\Gamma)=\frac{1}{2}\psi_0^*(\sum\Gamma_e+\sum\Gamma_o)=\frac{1}{2}(\psi_0^*(\sum\Gamma_e)+\psi_0^*(\sum\Gamma_o))$.
+\\Now $\psi_0^*(\sum\Gamma_e)\\=\sum_{i=0}^{11}\psi_0^*([\gamma_{2i}])\\=\sum_{i=0}^{11}( [\gamma_{2i}]-2m_0k_0[\gamma_{2i}]-m_0k_0[\gamma_{2i-2}] -m_0k_0[\gamma_{2i+2}]-m_0[\gamma_{2i+1}]-m_0[\gamma_{2i-1}] )\\=(1-4m_0k_0)\sum\Gamma_e+(-2m_0)\sum\Gamma_o$.\\
+Similarly,\\ $\psi_0^*(\sum\Gamma_o)\\=\sum_{i=0}^{11}\psi_0([\gamma_{2i+1}])=\sum_{i=0}^{11}([\gamma_{2i+1}] +k_0([\gamma_{2i}]+[\gamma_{2i+2}]))\\=\sum\Gamma_o+2k_0\sum\Gamma_e$.\\
+And so,\\ $\frac{1}{2}\psi_0^*(\sum\Gamma)=\frac{1}{2}((1+2k_0-4m_0k_0)\sum\Gamma_e +(1-2m_0)\sum\Gamma_o )$.\\\\
+Induction: Suppose $\psi^*_j(\sum\Gamma)=c_1\sum\Gamma_e+c_2\sum\Gamma_o$.\\
+Then $\psi^*_{j+1}(\psi^*_{j}(\sum\Gamma))\\=c_1\psi^*_{j+1}(\sum\Gamma_e)+c_2\psi^*_{j+1}(\sum\Gamma_o)\\=
+c_1((1-4m_{j+1}k_{j+1})\sum\Gamma_e+(-2m_{j+1})\sum\Gamma_o)+c_2(\sum\Gamma_o+2k_{j+1}\sum\Gamma_e)\\=
+(c_1(1-4m_{j+1}k_{j+1}) + 2k_{j+1}c_2)\sum\Gamma_e+(-2m_{j+1}c_1+c_2)\sum\Gamma_o$.\\
+Hence, $[\alpha]=\frac{1}{2}(c_1'\sum_e+c_2'\sum_o)$.
 
@@
 \begin{thm}
-Let $g$ be as above. Suppose $\alpha=g(\chi)$. Then $\Omega_{u,v}([\alpha])=(0,0)$. 
-\begin{proof}
-$\Omega_{u,v}([\alpha])=\Omega_{u,v}([g(\chi)])=\Omega_{u,v}(g^{*}([\chi]))=\frac{1}{2}(i(u,g^*(\sum\Gamma)) i(v,g^*(\sum\Gamma))$.\\
-$i(u,g^*(\sum\Gamma)=c_1 i(u,\sum
-\Gamma_e)+c_2 i(u,\sum\Gamma_o)=\\ c_1(i([\gamma_5],[\gamma_{4}] - i([\gamma_{11}],[\gamma_{10}]+[\gamma_{0}]))\\+c_2(i(-[\gamma_2],[\gamma_{1}]+[\gamma_{3}]) + i([\gamma_8],[\gamma_{7}]+[\gamma_{9}]) )=\\c_1((-2)-(-2))+c_2(-2+ 2 )=0\\$
-Holds similarly for $i(v,g^*(\sum\Gamma)).$\\
-$\therefore \Omega_{u,v}([\alpha])=(0,0)$.
-\end{proof}
-\end{thm}
+Let $\tau$ be as above. Suppose $\alpha=\tau(\chi)$. Then $\Omega_{u,v}([\alpha])=(0,0)$. 
+\begin{proof}
+$\Omega_{u,v}([\alpha])=\Omega_{u,v}([\tau(\chi)])=\Omega_{u,v}(\tau^{*}([\chi]))=\frac{1}{2}(i(u,\tau^*(\sum\Gamma)), i(v,\tau^*(\sum\Gamma))$.\\
+Now by the previous lemma, \\$i(u,\tau^*(\sum\Gamma)=c_1 i(u,\sum
+\Gamma_e)+c_2 i(u,\sum\Gamma_o)$ $\\= c_1(i([\gamma_5],[\gamma_{4}]+[\gamma_6]) - i([\gamma_{11}],[\gamma_{10}]+[\gamma_{0}]))\\+c_2(-i([\gamma_2],[\gamma_{1}]+[\gamma_{3}]) + i([\gamma_8],[\gamma_{7}]+[\gamma_{9}]) )\\=c_1((-2)-(-2))+c_2(-2+ 2 )=0\\$
+It's shown similarly that $i(v,\tau^*(\sum\Gamma))=0$.
+Therefore $\Omega_{u,v}([\alpha])=(0,0)$.
+\end{proof}
+\end{thm}
+
+Now we know that $\chi$ serves as something of a basis for a class of lifted closed segments.
+Because the sums of even and odd curves belong individually in the kernel of $\Omega_{u,v}$, anything in the orbit of $\chi$ will also lift to a closed segment on $\btU$.
+To show that these are the only segments that lift to closed paths on $\btU$, we look at the orbit of segments in directions $\theta=0$ and $\frac{\pi}{2}$ that are homologous to some $\gamma_{2i}$ and $\gamma_{2i+1}$, respectively.
+These will then span the rest of the homology classes of closed geodesic segments of $\bM$.
 
@@
 $\Omega_{u,v}([\gamma_i])\neq (0,0)$.
-\begin{proof}
-every curve intersects nontrivially with exactly one of either $u$ or $v$.
-\end{proof}
+\begin{figure}[H]
+\begin{wrapfigure}{l}{2.9in}
+\includegraphics[width=2.9in]{12gon.png}
+\end{wrapfigure}
+~\\\vspace{0.5in}\\Every curve, $\gamma_i$, intersects exactly one adjacently indexed curve that is an element of exactly one of either $u$ or $v$, or with one of both $u$ \emph{and} $v$.
+The image to the left shows every curves associated image under $\Omega_{u,v}$.
+Therefore, these classes are always taken to a non-trivial element of $\ZZ^2$ under $\Omega_{u,v}$.
+\end{figure}
+
 \end{lem}
+\vspace{0.8in}
+
+\begin{lem}
+$\left[\begin{matrix}
+1 & -1 & 0 & -1 & -1 & 0 & -1 & 0 & 1 & 0\\
+0 & 1 & 1 & 0 & 1 & -1 & 0 & -1 & 0 & 1\\
+\end{matrix}\right]$. Further its rank is 2 and nullity is 8.
+\end{lem}
 
 \begin{thm}
-Suppose $\beta=g(\gamma_i)$. Then $\Omega_{u,v}([\beta])\neq (0,0)$.
-\begin{proof}
-$g^*([\gamma_i])=[\gamma_i]+C$.
-If $i(u,g^*([\gamma_i]))=i(u,[\gamma_i]+C)=i(u,[\gamma_i])+i(u,C)=0$, then 
-\end{proof}
-\end{thm}
+Let $\sigma\in\XX$. Suppose $\beta=\sigma(\gamma_i)$. Let $[\beta]$ be its homology class. Then $\Omega_{u,v}([\beta])\neq (0,0)$.
+\begin{proof}
+Since $\beta=\sigma(\gamma_i)$, $[\beta]=[\sigma(\gamma_i)]=\sigma^*\cdot[\gamma_i]$.
+Express $\psi_j^*$ as a $10\times10$ matrix acting on standard basis coefficients of $\Gamma'$.
+This is a rank 10 matrix.
+The product $\psi_{j+1}^*\cdot\psi_j^*$ is also a rank 10 matrix.
+By induction on $j$, $\sigma^*$ is a finite left concatenated product of these matrices and is rank 10 as well.
+If $[\gamma_i]\in\Gamma'$ it is rank 1.
+If it is $[\gamma_7]$ or $[\gamma_{11}]$, by [LEM] it is a subspace of $H_1(\bM,\QQ)$ spanned by four other curves.
+Therefore the rank of $\sigma^*\cdot[\gamma_i]$ is less than or equal to 4.
+The matrix representation of $\Omega_{u,v}$ on $\Gamma'$ has nullity 8.
+Therefore $[\beta]$ is in a subspace of dimension always less than the dimension of the kernel of $\Omega_{u,v}$ and so $\Omega_{u,v}([\beta])\neq (0,0)$.
+\end{proof}
+\end{thm}
+
+Now since $\Omega$ and $\varphi_1$ are maps into abelian groups, they factor through the homology of these surfaces.
+There is a linear transformation between their homologies that ensures every closed geodesic on both surfaces are taken to the same $\ZZ^2$ element in the cover.
+
+\begin{Def}
+Let $\alpha\in H_1(\bM, \QQ)$ with coordinates relative to $\Gamma'$ (order of coefficients still ascending).
+Let $v_\alpha\in\QQ^{10}$ be the vector of rational coefficients.
+We use the fact that rational homology is the abelianization and field extension of the free group $\pi_1(\bGs)$ to say an element in $H_1(\bGs,\QQ)$ is of the form $x_1a+x_2b+x_3c+x_4d$.
+Let $L:H_1(\bM,\QQ)\twoheadrightarrow H_1(\bGs,\QQ)$ be the linear transformation between these vector spaces given by:
+\begin{align}
+L(\alpha):=\left[\begin{matrix}
+1 & -1 & 0 & -1 & -1 & 0 & -1 & 0 & 1 & 0\\
+0 & 1 & 1 & 0 & 1 & -1 & 0 & -1 & 0 & 1\\
+0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
+0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0\\
+\end{matrix}\right]v_{\alpha}\cdot \left[\begin{matrix}
+a \\ b \\ c\\ d\\
+\end{matrix}\right]^{\perp}
+\end{align}
+\end{Def}
+
+A quick calculation shows us that this matrix is rank 4, making $L$ is surjective.
+
+\begin{thm}
+Let $\alpha\in\pi_1(\bGs,x_0)$ be the homotopy class of a closed geodesic segment.
+Then its lifted image is homotopic to a non-trivial element $\beta\in\pi_1(\bM,m_0)$.
+Furthermore, these homotopy classes have the property that $\varphi_1(\alpha)=\Omega_{u,v}(\beta)$.
+\begin{proof}
+Observe that the lifted homotopy  $\tilde{\alpha}\in\pi_1(\bMs,m_0)$ as $\alpha\in\ker\varphi_2$.
+[lem] Since a periodic geodesic is non-singular, $\tilde{\alpha}$'s unique geodesic can be mapped to the unpunctured surface $\bM$ and vice-versa (where $m_0$ is \emph{not} a cone singularity).
+Call its homotopy class $\beta$.
+A geodesic on a compact, orientable hyperbolic surface could not be contracted to a single point.
+If that were true, the geodesic would have had to have curvature $\pi$ according to the Gauss-Bonnet Theorem, which is a contradiction.
+Therefore $\beta$ is non-trivial and its homology class has a coordinate $v_{\beta}\in \QQ^{10}$.
+Since $\varphi_1$ factors through homology, we may write $\varphi_1(\alpha)=\left[\begin{matrix} 1 & 0 & 0 & 0\\0 & 1 & 0 & 0\end{matrix}\right]v_\alpha\in\ZZ^2$.
+Now if we combine these two we get $\varphi_1\circ L=\left[\begin{matrix}
+1 & -1 & 0 & -1 & -1 & 0 & -1 & 0 & 1 & 0\\
+0 & 1 & 1 & 0 & 1 & -1 & 0 & -1 & 0 & 1\\
+\end{matrix}\right]$ as a function from $H_1(\bM,\QQ)$ to $\ZZ^2$.
+Let $v_\beta=(k_1,k_2,\cdots,k_{10})$.
+Now if we pass $\beta$'s abelianized form, $[\beta]$, along to $\varphi_1\circ L$, we have the first coordinate equal to $k_1-k_2-k_4-k_5-k_7+k_9\\
+=i(-[\gamma_2],k_2[\gamma_1]+k_3[\gamma_2])+i([\gamma_5], k_6[\gamma_5]+k_5[\gamma_4]+k_7[\gamma_6])\\+i([\gamma_8],k_8[\gamma_8]+k_9[\gamma_9])+i(-[\gamma_{11}],k_1[\gamma_{0}]+k_4[\gamma_3])\\=i(u,[\beta])$.\\
+Similarly, the second coordinate is just $i(v,[\beta])$.
+Thus $\Omega_{u,v}=\varphi_1\circ L$.
+Since $\alpha$ and $\beta$ are in one-to-one accordance, $L([\beta])=[\alpha]$, which means $\Omega_{u,v}(\beta)=\varphi_1(L([\beta]))=\varphi_1([\alpha])=\varphi_1(\alpha)$.
+\end{proof}
+\end{thm}
+
+Now projecting back onto the space allows us to identify an element of the deck group by some $\ZZ^2$ action on $\bUs$.
 
 \section{Proof of Main Theorem}
-
-$S=\cO\cup \cE$.
-Let $\XX'$ act on $\cS$ by linear transformations.
-
+Existence of an element in $\XX$ that sends a geodesic to one that is homologous to $\chi$ or some waist curve relies on the isomorphic properties of groups $\XX$, $\XX'$, and $\XX^*$.
+In the end the faithfulness and transitivity of one group implies the faithful action on the kernel of $\Omega_{u,v}$.
+The following is a well known fact of the group $\XX'$:
 \begin{lem}
-
+$\XX'=\ker m$ where $m:M(2,\ZZ)\rightarrow M(2,\ZZ)(mod~2)$.
 \end{lem}
+This lemma is used primarily to establish faithfulness.
+Recall that $(x,y)\in S=\cO\cup \cE$ is an element of $\ZZ^2$ with $x,y$ relatively prime.
+Further partition $\cE$ into sets $\cE_1$ and $\cE_2$, where integer pairs $(x,y)\in\cE$ are in $\cE_1$ if $x$ is odd or in $\cE_2$ if $x$ is even.
+These are disjoint since $x$ and $y$ cannot both be even or odd.
+Let $\XX'$ act on $\cS$ by linear transformations on $\ZZ^2$.
+In these proofs we make use of Bezout's identity often to establish transitivity on $\cO,\cE_1,$ and $\cE_2$:
+
+\begin{align}
+\text{For any relatively prime }x,y\text{ there exist }m,n\in\ZZ\text{ such that }mx+ny=1.
+\end{align}
+
+\subsection{Orbit of Rational Directions}
+Here are some results we use in the next section that are used to prove existence of geodesics in any given rational direction.
 
 \begin{lem}
-$\XX'$ acts faithfully and transitively on integer pairs in $\cO$, partitions $\cE$ into two sets and acts faithfully and transitively on them, and acts faithfully on $\cS$.
-\begin{proof}
-Let $v=kv_0=\begin{pmatrix}x \\ y\end{pmatrix}\in\cS$ with $||v_0||=1$ and $k\in\RR$, and let $A=\begin{pmatrix}a & b \\ c & d
+$\XX'$ acts faithfully and transitively on $\cO$.
+\begin{proof}
+Let $v_0=\begin{pmatrix}x \\ y\end{pmatrix}\in\cO$, where $x,y$ are both odd and relatively prime integers. Let $M=\begin{pmatrix}a & b \\ c & d
 \end{pmatrix}\in\XX'$.
-The kernel of the homomorphism $SL(2, \ZZ)\rightarrow \SL(2,\ZZ/2\ZZ)$ that maps a matrix to its $(mod~2)$ equivalent is precisely $\XX'$ since it maps the generators trivially.
-[quote this paper for the proof of this maybe?
-http://www.fysik.su.se/~ingemar/SL2.pdf ]\\
-$x,y$ are relatively prime, so Bezout's identity states there exist $m,n\in\ZZ$ such that $mx+ny=1~(*)$.
-It's obvious that this action is faithful on the sets since the stabilizer of any vector in $\RR^2$ is trivial.
-Suppose $v\in\cO$.
-Consider the orbit of $\begin{pmatrix}1 \\1 \end{pmatrix}$.
-Then there is some matrix $A\in\XX'$ such that $A\begin{pmatrix}1\\1\end{pmatrix}=v$.
-Since $A\in\SL(2,\ZZ)$, $\det A = ad-bc=1$.
-Obtain from $1=(x-b)(y-c)-bc=xy-xc-by$ that $xy-1=xc+by$.
-From $(*)$ we see that $xy-1=x(u(xy-1))+y(v(xy-1))$.
-Let $c=u(xy-1)$ and $b=v(xy-1)$.
-Since $xy-1$ is even, $c,b$ are even.
-It follows then that $a=x-b$ and $d=y-c$ are odd.
-Therefore $A\in \XX'$ and the action is transitive on $\cO$.
-If $v\in\cE$, then either $x$ or $y$ is even.
-In either case this would mean that either $a$ or $d$ is even, which would mean $A\notin\XX'$, a contradiction.\\
-Now consider the orbit of $\begin{pmatrix}1\\0\end{pmatrix}$.
-Left multiplying by $A$ would mean $a=x$ and $c=y$.
-Split $\cE$ into sets $\cE_1=\{(x,y)\in\cE :x\in2\ZZ,~y\in(2\ZZ+1) \}$ and $\cE_2=\{(x,y)\in\cE:x\in(2\ZZ+1),~y\in2\ZZ \}$.
-Since $\det A=1=da-bc=dx-by=mx+ny$, we have that $d=m$ and $b=-n$.
-We rule out when $v\in\cO,\cE_1$ since $c\neq 0 (\mod 2)$ in those cases.
-Consider $v\in\cE_2$.
-Then $x$ is odd and $y$ is even.
-$d$ must be odd since $mx+ny=1$ and $y$ is even.
-And $b$ must be even since $ad-bc=1$\\
-\compav{finish this and say the proof for $\cE_1$ is analogous.}
-
+Let $u=\begin{pmatrix}1 \\ 1 \end{pmatrix}\in\cO$.
+It follows that this action is faithful since $ax+by$ and $cx+dy$ cannot sum up to an element in $\cE$, as $a,d$ are odd and $b,c$ are even.
+We show that there is a matrix $M$ such that $Mu=v_0$, i.e.
+$a+b=x$ and $c+d=y$.
+We know $\det M=ad-bc=1$ and from $(*)$ that there exist $m,n\in\ZZ$ such that $1=mx+ny$.
+Let $a=(x-b)$, $b=n(xy-1)$, $c=m(xy-1)$, and $d=(y-c)$.
+Then
+\begin{align*}
+\det M&=(x-n(xy-1))(y-m(xy-1))-n(xy-1)m(xy-1)\\
+&=xy-mx(xy-1)-ny(xy-1)+mn(xy-1)^2-mn(xy-1)^2\\
+&=xy-x^2my-y^2nx+(mx+ny)=xy-xy(mx+ny)+1\\
+&=xy-xy+1=1.
+\end{align*}
+Therefore $M\in\SL(2,\ZZ)$.
+Now its clear that $a+b=(x-b)+b=x$ and $c+d=c+(y-c)=y$.
+Further, $b=n(xy-1)$ and $c=m(xy-1)$ are even since $xy-1$ is even, and so $a=x-b$ and $d=y-c$ are odd.
+Hence, $M\in\ker m$ and $Mu=v_0$.
+$M$ is invertible, so $M^{-1}v_0=u.$ Apply the same argument for some other $v_0'\in\cO$, and obtain that $N^{-1}v_0'=u.$ We conclude then that $NM^{-1}v_0=v_0'$ for any $v_0,v_0'\in\cO$.
+Therefore $\XX'$ acts transitively on $\cO$.
 \end{proof}
@@
 
-\begin{Def}
-Let $\phi_t^\theta:\RR\rightarrow\bM$ be a unit-speed geodesic on $\bM$ in direction $\theta$.
-\end{Def}
-
-\subsection{empty}
-\noindent Consider the following families of unit squares in $\RR^3$:\\
-\vspace{0.2in}
-\begin{tabular}{p{10cm}c}
+\begin{lem}
+$\XX'$ acts faithfully and transitively on $\cE_1$ and $\cE_2$.
+\begin{proof}
+Let $v_0=\begin{pmatrix}x \\ y\end{pmatrix}\in\cE$, where $x,y$ are relatively prime integers. Let $M=\begin{pmatrix}a & b \\ c & d
+\end{pmatrix}\in\XX'$. Let $u_1=\begin{pmatrix}1 \\ 0 \end{pmatrix}\in\cE_1$ and $u_2=\begin{pmatrix}0 \\ 1 \end{pmatrix}\in\cE_2$.\\\\
+Case I: $v_0\in \cE_1$. Then let $a=x$, $b=-n$, $c=y$, and $d=m$, where $mx+ny=1$ by $(*)$.  Then
 \begin{align*}
-\mathbf{A}_{m,n,p} = [m, m+1]\times[n,n+1]\times\{p\}, 
-\\\mathbf{B}_{m,n,p} = \{m+1\}\times[n,n+1]\times[p-1,p],
-\\\mathbf{C}_{m,n,p}= [m,m+1]\times\{n+1\}\times[p-1,p].
+\det M &= ad-bc=xm-(-ny)=mx+ny=1.
 \end{align*}
-&
-\shiftleft{0.8in}{\raisebox{-1in}{\includegraphics[scale=1]{label.png}}}
-\end{tabular}
-
-\begin{figure}[H]
-\centering
-\includesvg[width=1in]{cubesxyzcut}
-\caption{A section of $\mathbf{S}$}
-\end{figure}
-
-\subsection{Flattening $\mathbf{S}$}
-At the moment it is difficult to describe how the parallel transport of a vector tangent to the surface along an arbitrary path on the surface acts on the vector in $\RR^3$.
-To simplify this problem we take $\mathbf{S}$ to an isometric variant embedded in $\RR^2$ by piecewise linear transformations on the sets
-\begin{align*}
-\mathbf{A}=\bigcup\big{\{}\mathbf{A}_{m,n,p}: m+n+p=0\big{\}},
-\\\mathbf{B}=\bigcup\big{\{}\mathbf{B}_{m,n,p}: m+n+p=0\big{\}},
-\\\mathbf{C}=\bigcup\big{\{}\mathbf{C}_{m,n,p}: m+n+p=0\big{\}}.
-\end{align*}
-
-\begin{Def}
-Let $\Psi:s^*(\mathbf{S})\rightarrow\RR^3$ be given as
-\begin{equation}
-\Psi\left[\begin{array}{c}
-	x\\y\\z
-\end{array}\right] 
-= 
-\begin{cases}
-	\left[ \hspace{2mm} \begin{matrix}
-		1 & 0 & 0 \\
-		0 & 1 & 0 \\
-		0 & 0 & 1
-	\end{matrix}\hspace{3mm}\right]
-
-	\left[\begin{array}{c}
-	x - \left\lfloor x \right\rfloor
-	\\ y- \left\lfloor y \right\rfloor
-	\\ -(z - \left\lfloor z \right\rfloor)
-	\end{array} \right]
-	+
-	\left[\begin{array}{c}
-		2 \left\lfloor x \right\rfloor - \frac{3}{2}
-		\\ 2\left\lfloor y \right\rfloor - \frac{3}{2}
-		\\ z - \left\lfloor z \right\rfloor
-	\end{array} \right]
-		& \text{if } (x,y,z)\in \mathbf{A}	\vspace{2mm}
-	\\
-
-	\left[ \begin{matrix}
-	0 & 0 & 1 \\
-	0 & 1 & 0 \\
-	-1 & 0 & 0
-	\end{matrix}\hspace{2mm}\right]
-	\left[\begin{array}{c}
-		x - \left\lfloor x \right\rfloor
-		\\ y- \left\lfloor y \right\rfloor
-		\\ -(z - \left\lfloor z \right\rfloor)
-		\end{array} \right]
-	+
-		\left[\begin{array}{c}
-			2 \left\lfloor x \right\rfloor - \frac{3}{2}
-			\\ 2\left\lfloor y \right\rfloor - \frac{3}{2}
-			\\ x - \left\lfloor x \right\rfloor
-		\end{array} \right]
-			& \text{if } (x,y,z)\in \mathbf{B}\backslash(\mathbf{A}\cup\mathbf{C})	\vspace{2mm}
-	\\
-
-		\left[ \begin{matrix}
-		1 & 0 & 0 \\
-		0 & 0 & 1 \\
-		0 & -1 & 0
-		\end{matrix}\hspace{2mm}\right]
-		\left[\begin{array}{c}
-			x - \left\lfloor x \right\rfloor
-			\\ y- \left\lfloor y \right\rfloor
-			\\ -(z - \left\lfloor z \right\rfloor)
-			\end{array} \right]
-		+
-			\left[\begin{array}{c}
-				2 \left\lfloor x \right\rfloor - \frac{3}{2}
-				\\ 2\left\lfloor y \right\rfloor - \frac{3}{2}
-				\\ y - \left\lfloor y \right\rfloor
-			\end{array} \right]
-				& \text{if } (x,y,z)\in \mathbf{C}\backslash(\mathbf{A}\cup\mathbf{B})	\vspace{2mm}
-\end{cases}
-\end{equation}
-\end{Def}
-\noindent This map behaves as such:
-\begin{figure}[H]
-\centering
-\includesvg[width=1.1in]{cubesxyzcut} \hspace{0.1in}\raisebox{1.0in}{\text{$\rightarrow$}}\hspace{0.1in}
-\raisebox{0.4in}{\includesvg[width=1.5in]{unfoldcut}}
-\caption{An isometry of the surface $\mathbf{S}$.}
-\label{fig:Psi}
-\end{figure}
-
-\begin{lem}
-$\Psi\circ s^*$ is an isometry of the surface $\mathbf{S}$
-\begin{proof}
-$\Psi$ is well-defined on its domain $s^*(\mathbf{S})$, and $s^*$ is an embedding of the surface that is a bijection when restricting its co-domain to its image.
-Since $\Psi$ is a linear transformation composed of Euclidean matrices and translations it is also invertible and bijective.
+Hence $Mu_1=v_0$, and $M\in\XX'$ since $p(M)=I_2$.
+Repeating this argument for $v_0'\in \cE_1$, we get that $Nu_1=v_0'$.
+Therefore, $v_0=MN^{-1}v_0'$ and this action is transitive.\\\\\compav{Dont forget to prove faithfulness.}
+Case II: $v_0\in \cE_1$. The proof for this is analogous to the previous case.
 \end{proof}
@@
 
-\subsection{Translation Surface Cover of $\bGs$}
-
-\subsection{A Four-Fold Cover of $\bUs$}
-
-\newpage
-\begin{Def}
-$\bUs$ is the cover of $\bGs$ such that $\pi_1(\bUs)=\ker\varphi$.
-\end{Def}
-
-Our cover of $\bGs$ has the fundamental group, $\pi_1(\bUs)=\<\<a^{-1}b^{-1}ab, c, d\>\>$, the conjugate subgroup of elements that map trivially under $\varphi_1$.
-The covering map $k_{\varphi_1}:\bUs\hookrightarrow\bGs$ takes every point on $\bUs$ to its modular equivalent under these translational symmetries.
-$\bUs$ is realized as an infinite $\ZZ^2$-tiling of $\bGs$ in $\CC$:
-\begin{figure}[H]
-\centering
-\includesvg[width=2in]{uprime}
-\caption{The infinite surface $\bUs$ embedded in $\CC$ with edges identified.}
-\end{figure}
-
-\begin{Def}
-$\bU$ is the completion of $\bUs$ that includes the branched vertices. 
-\end{Def}
-
-\begin{rem}
-The surface $\bG$ is homeomorphic to the torus, and its cover $\bU$ is its universal cover as the paths labeled $c,d$ are only non-trivial when $\bG$ is branched at its three conical points.
-\end{rem}
-
-\subsection{}
-
-\subsection{Isometry From $\mathbf{S}$ to $\bU$}
-The Necker cube surface can be \emph{flattened} onto the plane by piecewise isometric maps onto a subspace of $\RR^3$ with (open) unit squares removed at every even integer pair in the plane, what we claim to be $\bU$.
-
-\noindent The red/blue dotted lines represent the edges that are split on the plane. The map from one surface to the other is composed of piecewise isometries, $\Psi:\mathbf{S}\rightarrow\RR^3$.
-
-The flattened surface is contained entirely in $(x,y,0)\in\RR^3$, which is isometric to $\CC$. $\bU$ is recovered as a topological quotient on the domain
-
-\begin{Def} $\mathbf U$ is the surface obtained as the topological quotient $\mathbf{P}/\sim_{\mathbf{P}}$, where $\CC$ is identified with $\RR^2$ in the usual way and $\sim_{\mathbf{P}}$ is a minimal relation on $\mathbf{P}$ defined as follows:
-\begin{gather*}
-\text{Let } x_{0}=(u_{0},v_{0}),x_{1}=(u_{1},v_{1}) \in \mathbf{P}.  \text{ }\sim_{\mathbf{P}} \text { is given as the relation }\\ x_{0}\sim_{\mathbf{P}}x_{1}  \text{ iff } x_{0}=x_{1}
- \\\text{ or, for some }m,n\in\ZZ~x_{0},x_{1} \in {\partial} \left( \left[2m-\frac{1}{2},2m+\frac{1}{2}\right] \times \left[2n-\frac{1}{2},2n+\frac{1}{2}\right] \right)\\
-  \left[\begin{array}{c}
-u_{1} -2m
-\\v_{1}-2n
-\end{array}\right] = \left[\begin{matrix}
-0 && 1\\
-1 && 0
-\end{matrix}\right]
-\left[ \begin{array}{c}u_{0}-2m\\
-v_{0}-2n
-\end{array}\right]
-.\end{gather*}\end{Def}
-
-\begin{rem}
-Our map $\Psi$ induces an isometry between $\mathbf{S}$ and $\bU$, and its restriction to the branched surfaces induces an isometry between $\mathbf{S}^\circ$ and $\bUs$.
-This follows from these piecewise Euclidean transformations that preserves our induced flat metric.
-\end{rem}
-From here on we use $\bUs$ and $\bU$ instead of $\mathbf{S}$ and $\mathbf{S}^\circ$.
-
-\subsection{Four-Fold Cover}
-We denote the \emph{unit tangent bundle} on the surface $\bUs$ by $T^1\bUs$, and construct a cover with trivial holonomy.
-Initial experiments have shown us that any geodesic viewed as discontinuous line segments on $\mathbf{P}$ moves in at most four directions.
-\begin{figure}[H]
-\centering
-\includesvg[width=1.8in]{cubesxyzpaths}
-\includesvg[width=2.8in]{unfoldcutpath}
-\caption{Simple periodic (c) and drift-periodic (a,b) trajectories represented as line segments on $\mathbf{S}$ and $\mathbf{U}$.\compav{RELABEL THESE}}
-\label{fig:holonomypath}
-\end{figure}
-Observe that the parallel transport of a vector around a closed loop on $\bUs$ will act on vectors tangent to the surface by a rotation of an integer multiple of $\frac{\pi}{2}$ radians (since the surface is flat and embedded in $\CC$, we work with vectors in $\RR^2$).
-
-\begin{Def}
-Define $\varphi_2:\pi_1(\bGs)\rightarrow SO(2,\ZZ)$ as the group homomorphism on generators $a,b,c,d$ such that:
-\begin{align*}
-a,b\mapsto \begin{pmatrix}1 && 0 \\ 0 && 1\end{pmatrix}=I_2,~~~
-c\mapsto \begin{pmatrix}0 && -1 \\ 1 && 0\end{pmatrix}=R,~~~
-d\mapsto \begin{pmatrix}0 && 1 \\ -1 && 0\end{pmatrix}=R^3
-\end{align*}
-\end{Def}
-The homomorphism restricted to $\pi_1(\bUs)$ factors through $\pi_1(\bGs)$ and we show that:
-\begin{lem}
-The group $\varphi_2(\pi_1(\bUs,x_0))=Hol_p(\omega)$.
-\begin{proof}
-Since $c\mapsto R$ and $d\mapsto R^3$, the image of $\varphi_2$ is generated by $R,R^3$ and isomorphic to $SO(2,\ZZ)$.
-We see from figure $\ref{fig:holonomypath}$ that parallel transports of vectors by non-trivial paths produce clockwise/counter-clockwise rotations equal to that of of the cone angles of the singularities they loop around, all integer multiples of $\frac{\pi}{2}$ that coincide with paths generated by elements $c,d$.
-Further, if there is an element in $\pi_1(\bUs,x_0)$ conjugated by $[a,b]$, the effect this would have on holonomy would be trivial as the total cone angle sum around each cut out square is an integer multiple of $2\pi$, which agrees with $a$ and $b$'s trivial images under $\phi_2$.
-\end{proof}
-\end{lem}
-
-Hence, we obtain a natural monodromy representation with the map $m:\pi_1(\bUs,x_0)\rightarrow SO(2,\ZZ)\cong Hol_p(\omega)$, where for $[\gamma]\in\pi_1(\bUs,x_0)$ we have that $\gamma^*(1)=(x_0,m([\gamma])v_0)$.
-It follows that since $\omega$ is a flat connection, trivial loops have trivial holonomy and $Hol_p(\omega)$ acts on $H(p)$.
-
-\begin{Def}
-Let $\btUos$ be the cover of $\bUs$ with fundamental group $\pi_1(\btUos)=\ker\varphi_2|_{\pi_1(\bUs)}\trianglelefteq \pi_1(\bUs)$.
-\end{Def}
-
-\begin{lem}
-Let $\Phi_t:\bUs\times\RR\rightarrow \bUs$ be a unit-speed geodesic flow on $\bUs$, with a parallel transport map induced by $\omega$. Then the following is true:
-\begin{enumerate}[label=(\roman*)]
-\item $\Phi$ is periodic if and only if its lift to $\btUos$ is.
-\item $\Phi$ is drift-periodic if and only if its lift to $\btUos$ is. 
-\end{enumerate}
-\begin{proof}
-Let $x_0=\Phi(0)$ be an initial point in $\bUs$, and let $v_0=\frac{d}{dt}\Phi|_{t=0}\in\RR^2$ be its initial direction.
-Then $\Phi^*:T^1\bUs\times\RR\rightarrow T^1\bUs$ is a well-defined lift to the unit tangent bundle with initial point $p=(x_0,v_0)$.
-\\\emph{(i)}.
-Suppose $\Phi$ is periodic with period $T\in\RR$.
-Then $\Phi$ is reparameterized as $\alpha:[0,1]\rightarrow \bUs$, where $[\alpha]\in\pi_1(\bUs,x_0)$.
-It follows then that if $\alpha$ does not lift to a closed path, then $\alpha$ must have non-trivial holonomy since $[\alpha]\notin\ker\varphi_2$.
-That is when lifted to the unit tangent bundle with base point $p$, $\alpha^*(1)\neq p$ since $m([\alpha])\neq I_2$.
-But this implies that $\Phi^*(T)\neq p$, which is impossible.
-Hence $[\alpha]\in\ker\varphi_2=\pi_1(\btUos,\tilde{x}_0)$, where $\tilde{x}_0$ belongs to the fiber over $x_0$ under the covering map $\btUos\hookrightarrow\bUs$.
-The converse holds trivially.
-\\\emph{(ii)}.
-Suppose $\Phi$ is drift-periodic with period $T\in\RR$ and non-trivial $f:\bUs\rightarrow\bUs \in Trans(\bUs)\cong\pi_1(\bGs)/\pi_1(\bUs)\cong \ZZ^2$ such that $\Phi(T)=f(x_0)$.
-Re-parameterize this as $\alpha:[0,1]\rightarrow\bUs$ with $[\alpha]\in\pi_1(\bGs)$.
-When lifted to $T^1\bUs$, $\Phi^*(0)=p$ and $\Phi^*(T)=(f(x_0),v_0)$.
-Let $\alpha:[0,1]\rightarrow \bUs$ be its reparameterization up to time $T$.
-Thus $[\alpha]\notin \pi_1(\bUs,x_0),\pi_1(\btUos,\tilde{x}_0)$.
-Further, $f$ has a unique, lift to $\tilde{f}\in Trans(\btUos)$ because the space is connected.
-(a bit stuck here).
-\\
-Conversely, ...
-\end{proof}
-\end{lem}
-
-A visual representation of $\btUos$ is as a four-fold cover of $\bUs$ with trivial holonomy on all closed paths. Arbitrary paths do \emph{not} have trivial holonomy:
-
-\begin{figure}[H]
-\centering
-\includesvg[width=1.5in.]{quotient}
-\caption{A $2\times2$ cut out section centered at each missing square. Edges and vertices identified.}
-\label{fig:quotient}
-\end{figure}
-
-\subsection{Translation Surface}
-
-A better picture of $\btUos$ is obtained by making cyclic edge identifications on $\mathbf{P'}=\mathbf{P}\times\ZZ/4\ZZ$.
-This comes from $\btUos$ inheriting the topological properties of $\bUs\times \pi_1(\bUs)/\pi_1(\btUos)\cong \bUs\times SO(2,\ZZ)$.
-
-\begin{figure}[H]
-\centering
-\includesvg[width=4in]{utilda}
-\label{fig:utilda0}
-\caption{Branched cover of $\mathbf{U}$ of degree four.}
-\end{figure}
-
-\begin{Def}
-$\btU_0$ is the surface obtained as the quotient $\mathbf{P}'/\sim_{\mathbf{P}'}$.
-Denote paths $\zeta$ and $\eta$  on $\mathbf{P}$, parameterized by integers $m,n\in\ZZ$ and $t\in[-\frac{1}{4},\frac{1}{4}]$, and defined:\\ $\zeta(t)=(2m+t)+i(2n-\frac{1}{2})$,\\
-$\eta(t)=(2m+t)+i(2n+\frac{1}{2})$.\\
-
-Let $\aleph=2m+i2n\in\CC$. The minimal relation $\sim_{\mathbf{P}'}$ is given as:
-\begin{equation}
-\begin{split}
-(\zeta(t);j)\sim_{\mathbf{P'}}(e^{i\frac{\pi}{2}}~\overline{\zeta(t)-\aleph};j+1)\\
-(\eta(t);j)\sim_{\mathbf{P'}}(e^{i\frac{\pi}{2}}~\overline{\eta(t)-\aleph};j+1)
-\end{split}
-\end{equation}
-\end{Def}
-This relation is similar to $\sim_\mathbf{P}$ in that it translates line segments of squares surrounding even integer pairs to the origin and relates points on one edge of a square to points on an adjacent edge.
-Where it differs is that these adjacent edges now belong to a different ``copy" of $\mathbf{P}$.
-It is in this cyclic manner that edges are glued that allows for trivial linear holonomy on arbitrary paths by rotating each copy of $\mathbf{P}$ accordingly.
-For example, here is a path on a section of the surface in the neighborhood of $P=2m+i2n\in\CC$ $(m,n\in\ZZ)$ after rotation:
-
-\noindent and its subsequent projection onto $\bUs:$
-
-\begin{Def}
-Let $r:\mathbf{P}'\rightarrow\mathbf{P}'$ be the isometric map of $\btU_0$'s domain given  $r(z;\bar{j})=(e^{(-j)i\frac{\pi}{2}}z;\bar{j})$.
-\end{Def}
-
-Observe that when $r$ acts on the relations $\eqref{eq:rel2}$ we get the following:
-
-\begin{equation}
-\begin{split}
-(e^{(-j)i\frac{\pi}{2}}\zeta(t);j)\sim_{r(\mathbf{P'})}(e^{(-j)i\frac{\pi}{2}}~{\zeta(t)-\aleph};j+1)\\
-(e^{(-j)i\frac{\pi}{2}}\eta(t);j)\sim_{r(\mathbf{P'})}(e^{(-j)i\frac{\pi}{2}}~{\eta(t)-\aleph};j+1)
-\end{split}
-\end{equation}
-
-$\btU$ is then recovered as $r\cdot(\btU_0)=\mathbf{P'}/\sim_{r(\mathbf{P}')}$, where all identifications are translations made on copies of $\mathbf{P}$.
-
-\subsection{Translation Surfaces and $\ZZ^2$-Covers}
-
-\begin{Def}
-Algebraic intersection number is a non-degenerate bilinear form:
-\begin{align*}
-i:H_1(S,\mathbf{R})\times H_1(S,\mathbf{R})\rightarrow\mathbf{R},
-\end{align*}
-where $\mathbf{R}$ is a ring and for $[\gamma],[\beta]\in H_1(S,\mathbf{R})$, $i([\beta],[\gamma])$ returns the intersection number of two homology classes.
-\end{Def}
-Algebraic intersections are signed and follow some convention such as the right-hand rule.
-
-\begin{Def}
-Let $u,v\in H_1(S,\QQ)$ be linearly independent homology classes of curves on $S$. Then the group homomorphism from $\pi_1(S)$ to $\QQ^2$ is given as:
-\begin{align*}
-\Omega_{u,v}:\pi_1(S)\rightarrow \mathbf{R}^2 \text{; } \beta\mapsto(i(u,[\beta]),i(v,[\beta])).
-\end{align*}
-\end{Def}
-
-The set of all orientation-preserving affine diffeomorphisms of $S$ forms the group Aff$^+(S)$.
-The corresponding \emph{Veech group}, $V(S)$ of $S$ is the image of the group morphism $D:\text{Aff}^+(S)\rightarrow SL(2,\mathbb{R})$ that takes an affine map to its derivative.
-A surface is said to be \emph{Veech} if its Veech group is commensurable to $SL(2,\mathbb{R})$.
-It is well known that origami, or square-tiled, surfaces have Veech groups commensurable to $SL(2,\ZZ)$.
-[cite] When $\mathbf{R}=\ZZ$, $\Omega_{u,v}$ takes an element of $\pi_1(S)$ to $\Delta$.
-Thus, $\gamma\in\pi_1(S)$ lifts to $\tilde{\gamma}\in\pi_1(\tilde{S})$ if and only if $\gamma\in\ker\Omega_{u,v}$.
-
-\begin{Def}
-Let $\alpha:[0,1]\rightarrow S$ be a closed, non-singular geodesic path on $S$.
-The holonomy map $\mathbf{hol}:H_1(S, \mathbf{R})\rightarrow\CC$ returns the holonomy vector of a closed path as a difference of the starting and endpoints of a flow by\\
-\begin{align*}
-\mathbf{hol}([\alpha])=\int_{\alpha}dz.
-\end{align*}
-\end{Def}
-Since $\alpha$ is non-singular, it can be mapped to $S^\circ$ which admits a flat holomorphic one-form $dz$.
-Let $\theta=Arg(\mathbf{hol}([\alpha]))$.
-We say that $\phi_t^\theta:\RR\times S^\circ \rightarrow S^\circ$ is the unit-speed geodesic flow on $S^\circ$ in direction $\theta$ given by the  $[\alpha]$ such that $\phi^\theta_0=\alpha(0)$.
-In local coordinates this corresponds to $z+te^{i\theta}\in \CC$.
-
-\begin{lem}
-$\phi_t^\theta$ has a period of $T=|\mathbf{hol}([\alpha])|$
-\begin{proof}
-This just follows from the fact that $\phi_t^\theta$ flows at unit-speed in the direction of $\frac{\mathbf{hol}([\alpha])}{|\mathbf{hol}([\alpha])|}$. 
-\end{proof}
-\end{lem}
-And so the length of a vector determines the period of a flow on $S$. More importantly, we have the following:
-
-\begin{lem}
-Denote the lifted flow of $\phi_t^\theta$ on $\tilde{S}$ by $\tilde{\phi}_t^\theta$.
-Then $\tilde{\phi}_t^\theta$ is periodic on $\tilde{S}$ and $\tilde{S}^\circ$ if and only if $[\alpha]\in\ker\Omega_{u,v}$.
-Further, $\tilde{\phi}_t^\theta$ has period $T=|\mathbf{hol}([\alpha])|$.
-\begin{proof}
-Suppose that $[\alpha]\notin\ker\Omega_{u,v}$.
-Since $\alpha([0,1])=\phi^\theta_{[0,\hol([\alpha])]}$, their homology classes are equivalent.
-Hence $\tilde{\phi}_t^\theta$ could not close on $\tilde{S}$ or $\tilde{S}^\circ$, or else $\Omega_{u,v}(k[\alpha])=k(i(u,[\alpha]),i(v,[\alpha]))=(0,0)$ for some non-zero $k\in\ZZ$.
-Conversely, suppose $[\alpha]\in\ker\Omega_{u,v}$.
-Then
-\end{proof}
-\end{lem}
+\subsection{Directions \& Geodesic Orbits}
+Let $\hol:H_1(\bM,\QQ)\rightarrow\RR^2$ be the translational holonomy vector map that associates each cycle with a vector in $\RR^2$.
+The holonomy map preserves distances and is an accurate representation of a geodesic segment on $\bM$.
+In local coordinates an element of $V(\bM)$ factors through $\hol$.
+For example $\chi$ is a geodesic segment in direction $\theta=\frac{\pi}{4}$ with $\hol([\chi])=(6,6)$.
+After applying one multi-twist in the horizontal direction, we obtain $\hol([\psi_h(\chi)])=\hol(\psi_h^*([\chi])=\psi_h'\cdot\hol([\chi])=\psi_h'\cdot(6,6)=(18,6)=6(3,1)$.
+
+We denote the set of closed non-singular geodesics (separatrices) based at $x\in\bMs$ as $V_{x}$.
+The collection of these geodesics is the set $\cV=\bigcup_{x\in\bMs} V_{x}$.
+Let $\tau\in\XX\subseteq \text{Aff}^{+}(\bM)$.
+Then $\tau$ induces an automorphism of $\cV$ that maps $V_{x}$ to $V_{\tau(x)}$.
+In addition to this, it is a parabolic isometry of the surface that takes a geodesic in $V_{x}$ in direction $\theta$ to a geodesic in $V_{\tau(x)}$, not necessarily in the same direction.
+$V_{x}$ forms a set of homology classes, $[V_{x}]$, such that $\hol([V_{x}])$ and $\hol([V_{\tau(x)}])$ are indistinguishable, discrete subsets of $\RR^2$.
+The faithful, transitive action of $D(\tau)=\tau'$ on $\cO$ and $\cE$ extends to $\hol([V_{x}])$.
+Because $\tau$ is isometric to $\tau^*\in\XX^*\subseteq \Aut(H_1(\bM,\QQ))$, the relationship between a choice of direction of a given geodesic, $\alpha$, based at $m_0$ and the orbit of its homology class becomes clear.
+
+\begin{Def}
+Let $\phi_t^\theta:\RR\rightarrow\bM$ be a unit-speed geodesic with base point $m_0$ in direction $\theta\in \RR/2\pi\ZZ$.
+Let $v_\theta=(\cos\theta,\sin\theta)\in\RR^2$ be a unit-vector in direction $\theta$.
+If this direction is rational, then $v_\theta$ is scaled by some $k\in\RR$ where $kv_\theta\in\cS$.
+\end{Def}
+
+The geodesic on $\btU$ in the same direction is $\tildphi_t^\theta:\RR\rightarrow\btU$, where $\tildphi^\theta(0)=\utild$ is a fixed base point in the fiber over $m_0$ under the projection map.
+We will use $\alpha,\beta$ to refer to two kinds of geodesic segments obtained from $\phi_t^\theta$.
+We will call their $\btUs$ lifts based at $\utild$ $\tilde{\alpha}$ and $\tilde{\beta}$, and denote their homology class representatives as $[\alpha]$ and $[\beta]$.
+Similar notation will be used for $\chi$, and horizontal and vertical direction geodesics $\gamma_h$, $\gamma_v$.
+
+\begin{thm}
+Let $kv_\theta\in\cO$.
+Let $\chi$ be a geodesic contained in the two cylinder decomposition obtained by flowing $\bM$ in the direction $\frac{\pi}{4}$.
+Every unit-speed, $m_0$ based, geodesic segment $\alpha:[0,T]\rightarrow\bM$ in direction $\theta$ is homologous to an element in the orbit of $\chi$ under $\XX$.
+Further, the length of $\alpha$ is given as $T=6||kv_\theta||$.
+\begin{proof}
+Let $\alpha:[0,T]\rightarrow\bM$ be a closed unit-speed geodesic segment based at $m_0$ with $\alpha'=v_\theta$ and $T=||\hol([\alpha])||$.
+By transitivity of $\XX'$ acting on $\cO$, we can find some $\tau'$ such that $6kv_\theta=6\tau'\cdot(1,1)=\tau'\cdot(6,6)=\tau'\cdot\hol([\chi])$.
+$\tau'$ is isomorphic to $\tau\in\XX$, so $\tau'\cdot\hol([\chi])=\hol([\tau(\chi)])$.
+Let $\chi:[0,T']\rightarrow\bM$ be a geodesic in one of these cylinders based at $\tau^{-1}(m_0)$.
+Then $\tau(\chi)$ is a geodesic in the same direction as $\alpha$, based at $m_0$.
+$\tau$ is an isometry of $\bM$ that sends geodesics to geodesics, so $\tau(\chi)([0,T'])=\alpha([0,T])$.
+Therefore $[\tau(\chi)]=[\alpha]\in H_1(\bM,\QQ)$.\\
+Show $T=||6kv_\theta||$.
+\end{proof}
+\end{thm}
 
 \begin{cor}
-If $[\alpha]\notin\ker\Omega_{u,v}$, then $\tilde{\phi}_t^\theta$ is drift-periodic with period $T=\hol([\alpha])$.
-\begin{proof}
-This follows immediately from the previous lemma since $\tilde{S},\tilde{S}^\circ$ have translational $\ZZ^2$ symmetries and a well-defined $\ZZ^2$ action on an element in the fiber of a basepoint in $S$ under $f$.
-The period is $T$ since a geodesic closes on $S$ with period $T$.
+Let $\alpha$ be as before. Then $\alpha$ lifts to a closed geodesic $\tilde{\alpha}$ on $\btU$ of the same period.
+\begin{proof}
+Since $\alpha$ is homologous to $\tau(\chi)$, we have that $\Omega_{u,v}([\alpha])=\Omega_{u,v}([\tau(\chi)])=(0,0)$ by [lem].
+Therefore, $[\alpha]$ is in the kernel of $\Omega_{u,v}$ and its lift is an element of $\pi_1(\btU,\mtild)$.
+The equality of their periods follows from $\btU$ being a $\ZZ^2$ cover of a translation surface.
 \end{proof}
@@
 
-\begin{lem}
-Let $h\in\text{Aff}^+(S)$. If $h(\beta)=\alpha$ for closed geodesics $\alpha,\beta$ on $S$ and $\beta$ lifts to a closed path on $\tilde{S}$, then so does $\alpha$. (not sure about this)
-\begin{proof}
-If the premise is true, then $[\beta]\in\ker\Omega_{u,v}$.
-Denote the group automorphism induced by $h$ as $h_*$.
-Then $\Omega_{u,v}([\alpha])=\Omega_{u,v}([h(\beta)])=\Omega_{u,v}(h_*\cdot[\beta])=(i([\beta],h_*^{-1}\cdot u),i([\beta],h_*^{-1}\cdot v))=(0,0)$ as automorphisms .
-\end{proof}
-\end{lem}
-
-\begin{lem}
-Let $\beta=h(\alpha)$ as before. If $D(h)=h'\in V(S)$ and $S$ is Veech, then $\tilde{\phi}^\theta_t$ has period $T=h'\cdot\hol([\beta])$. (also not sure)
-\begin{proof}
-
-\end{proof}
-\end{lem}
-
-\subsection{Symmetries of $\bM$}
-
-\section{Four-fold Cover of $\mathbf{U}$}
-
-\begin{figure}[H]
-\centering
-\includesvg[width=2.05in.]{fourfold}\hspace{0.5in}\includesvg[width=2.05in.]{fourfoldrotate}
-\caption{Four-fold cover isometry and the preimage of a point in $\mathbf U \backslash Sing(\mathbf U)$.}
-\end{figure}
-
-\begin{figure}[H]
-\centering
-\includesvg[width=3.15in]{utilda}
-\label{fig:utilda0}
-\caption{Branched cover associating every direction with one plane.}
-\end{figure}
-
-\begin{figure}[H]
-\centering
-\includesvg[width=3.15in]{utildaprime}
-\caption{Infinite-type translation surface obtained by rotating each copy of the fundamental domain accordingly.}
-\label{fig:utilda}
-\end{figure}
-
-The quotient under the group action of translational symmetries is isomorphic to $\mathbb{Z}^2$ since the orbit of any point in the fundamental domain is a lattice in the space.
-
 \begin{thm}
-The translational symmetries of $\tilde{\mathbf{U}}$'s fundamental domain induce symmetries on the surface isomorphic to $\mathbb{Z}^2$.
-\begin{proof}
-Let $(z;j)\in\mathbb{C}\times\mathbb{Z}/4\mathbb{Z}$ and define the group action $T_0^{m,n}:\mathbb{C}\times\mathbb{Z}/4\mathbb{Z}\rightarrow\mathbb{C}\times\mathbb{Z}/4\mathbb{Z}$ as $T^{m,n}(z;j)=(z+2e^{-j\frac{i\pi}{2}}(m+in);j)$.
-This translation acts faithfully on the preimages of $\mathbf{U}\backslash Sing(\mathbf{U})$, and respects edge identifications of $\mathbf{\tilde{\mathbf{U}}}$, thereby making it an isometry of the surface.
-Consider a group homomorphism, $T_0^{m,n}\mapsto m+in$ onto the plane of Gaussian integers, $\mathbb{Z}[i]$.
-The exponential function is never zero, so the identity of the translation group is $T_0^{0,0}$.
-This is an isomorphism since it is clearly surjective and any non-trivial element of $T^{m,n}$ could not possibly map to the identity element of $\mathbb{Z}[i]$, regardless of the value of j.
-Since $\mathbb{Z}^2$ is isomorphic to $\mathbb{Z}[i]$, it is isomorphic to  $T_0^{m,n}$ as well.
-\end{proof}
-\label{thm:z2}
-\end{thm}
-
-\begin{Def}
-The automorphism $T^{m,n}:\tilde{\mathbf{U}}\rightarrow\tilde{\mathbf{U}}$ is an  \textbf{induced translation} of $\tilde{\mathbf{U}}$ as a result of the previous theorem.
-\end{Def}
-
-This surface is obtained as a ramified cover of the unit square torus.
-It is a translation surface and is therefore equipped with a \textbf{holomorphic one-form}, a collection of charts from neighborhoods of $\mathbf{M}$ to $\mathbb{C}$ such that any neighborhood away from Sing$(\mathbf{M})$ has a \emph{flat} induced Euclidean metric.
-A theorem of Gutkin and Judge tells us that its Veech group is commensurable to SL$(2,\mathbb{Z})$ and is therefore a Veech surface.
-We look at some of its affine maps, and generate a subgroup $\mathbb{X}\subset$ Aff$^+(\mathbf{M})$  by the following transformations:
-
-\begin{enumerate}[label=(\roman*)]
-\item Multi-twists of the surface as global diffeomorphisms given by Dehn-twists of its cylinder decomposition in horizontal and vertical directions with derivatives:
-\begin{equation*}
-\left\{ \left[ \hspace{1mm} \begin{matrix}
-				1 &  \pm 2\\
-				0 & 1
-			\end{matrix}\hspace{1mm}\right] \text{ , }
-			\left[ \hspace{1mm} \begin{matrix}
-							1 & 0\\
-							 \pm 2 & 1
-						\end{matrix}\hspace{1mm}\right] \right\}.
-\end{equation*}
-we call $\mathbf{A}^{\pm 1}$, $\mathbf{B}^{\pm 1}$, respectively.\\
-A Dehn-twist on each cylinder in the cylinder decomposition of $\mathbf{M}$ in horizontal and vertical directions gives way to these global affine diffeomorphisms:
-
-\item Rotation group generated by a $+\frac{\pi}{2}$ rotation of the surface fixed about the center of the second square on the bottom of the staircase, an order four isometry on $\mathbf{M}$ denoted $\mathbf{R}$.
-
-\item  Order 2 translation of the surface that moves the bottom left-most square to the square right next to it, denoted $\mathbf{H}$.
-\item Order 2 translation of the surface that takes the bottom right-most square to the one right above it, denoted $\mathbf{V}$.
-\end{enumerate}
-
-\begin{Def}
-The group $\mathbb{X}$ is the isometry group generated by affine maps $\mathbf{A}$,$\mathbf{B}$,$\mathbf{R}$, $\mathbf{H}$, and $\mathbf{V}$.
-The image of the derivative map on elements in $\mathbb{X}$ is denoted $\mathbb{X}'$ and generated by matrices
-\begin{align*}
-\left[\begin{matrix}
-1 & 2 \\ 0 & 1
-\end{matrix}\right],
-\left[\begin{matrix}
-1 & 0 \\ 2 & 1
-\end{matrix}\right],
-\left[\begin{matrix}
-0 & -1 \\ 1 & 0
-\end{matrix}\right],
-\end{align*}
-denoted $\mathbf{A}'$, $\mathbf{B}'$, and $\mathbf{R}'$ in that order.
-\end{Def}
-
-It is not immediately apparent if these affine maps generate Aff$^{+}(\mathbf{M})$, or if their derivatives generate $V(\mathbf{M})$, its Veech group.
-We use these to induce homomorphisms on $H_1(X, \mathbb Q)$.
-A spanning set of $H_1(\mathbf{M}, \mathbb Q)$ is obtained as the set of homology classes of the core curves of X's cylinder decompositions in both vertical and horizontal directions:
-
-\begin{figure}[H]
-\includegraphics[width=3in]{homologyclass.png}
-\centering
-\caption{Cylinder core curves with u,v, and z homology classes that determines the $\mathbf{Z}^{2}$-cover.}
-\label{fig:homology}
-\end{figure}
-
-\begin{Def}
-The set of abelianized cylinder core curves is denoted as $\Gamma=\{\gamma_i: i = 0,\dots,11\}\subset H_1(\mathbf{M},\mathbb{Q})$.
-\end{Def}
-
-\begin{rem}
-We use 12 elements to span homology, although a basis requires only 10.
-It's not impossible to determine the relations between these core curve classes, but it is not necessary.
-A $12\times12$ matrix of these core curve cylinder decompositions to their intersection numbers with adjacent curves is rank 10, as to be expected.
-\end{rem}
-
-The induced homomorphisms of $H_1(\mathbf{M}, \mathbb Q)$ have come from affine maps that have various effects on these core-curves.
-We use a 12-gon to represent the set of curves, and show how these elements act on them.
-The multi-twists add curves to adjacent curves, and the translation maps permute them.
-The reader is encouraged to check these for themselves.
-
-e.g. for $\mathbf{H,V}\in$Aff$^+(\mathbf{M})$,
-
-\begin{minipage}{0.5\textwidth}
-\vspace{0.3in}
-\begin{itemize}
-\item[\textbf{\emph{$\mathbf{H}$ \& $\mathbf{V}$}}] The effect that these two translations have on the 12-gon is a reflection about these lines.
-Observed by keeping track of the squares and core curves after $\mathbf{H}$ and $\mathbf{V}$ have acted on $\mathbf{X}$.
-\end{itemize}
-\end{minipage}
-\begin{minipage}{0.7\textwidth}
-\begin{figure}[H]
-\hspace{0.2in}\includegraphics[width=1.7in]{12gonHV.pdf}
-\end{figure}
-\end{minipage}
-
-\begin{Def}
-The induced homomorphisms of $H_1(\mathbf{M},\mathbb{Q})$ are obtained from the affine subgroup $\mathbb{X}$ and denoted $\mathbb{X}_*$.
-The associated homomorphisms on the spanning set $\Gamma$ are given as:
-\begin{align*}
-\mathbf{A}^k_*\circ[\gamma_i]=&[\gamma_i] + \frac{k}{2}(1-(-1)^i)([\gamma_{i-1}]+[\gamma_{i+1}])\\
-\mathbf{B}^k_*\circ[\gamma_i]=&[\gamma_i] + \frac{k}{2}(1+(-1)^i)([\gamma_{i-1}]+[\gamma_{i+1}])\\
-\mathbf{R}_*\circ[\gamma_i]=&(-1)^i[\gamma_{1-i\text{ mod }12}]\\
-\mathbf{H}_*\circ[\gamma_i]=&[\gamma_{12-i\text{ mod }12}]\\
-\mathbf{V}_*\circ[\gamma_i]=&[\gamma_{10-i\text{ mod }12}]
-\end{align*}
-\end{Def}
-
-\begin{Def}
-The homology classes $u,v,z$ are given as the following sums of core curves:
-\begin{align*}
-[u] &= -[\gamma_2] +[\gamma_5] + [\gamma_8] - [\gamma_{11}],\\
-[v] &= +[\gamma_0] -[\gamma_3] -[\gamma_6] +[\gamma_9],\\
-[z] &= +[\gamma_1] +[\gamma_4]-[\gamma_7]-[\gamma_{10}].
-\end{align*}
-\end{Def}
+Let $kv_\theta\in\cE$.
+Every unit-speed, $m_0$ based geodesic segment $\beta:[0,T]\rightarrow\bM$ in direction $\theta$ is homologous to some cylinder waist curve $\gamma_i$.
+Furthermore, the length of $\beta$ is given as $T=2||kv_\theta||$.
+\begin{proof}
+Case I: $kv_\theta\in\cE_1$.
+By transitivity, there exists some $\sigma_h'\in\XX'$ such that $2kv_\theta=2\sigma_h'\cdot(1,0)=\sigma_h'\cdot(2,0)=\sigma_h'\cdot\hol([\gamma_{2i}])=\hol([\sigma_h(\gamma_{2i})])$.
+Let $\gamma_h:[0,T_h]\rightarrow\bM$ be a horizontal geodesic based at $\sigma_h^{-1}(m_0)$ contained in a horizontal cylinder in $\mathfrak{H}$ that is homologous to some $\gamma_{2i}$.
+Then $\sigma_h(\gamma_h)$ is a geodesic based at $m_0$ with $\hol([\sigma_h(\gamma_h)])=\sigma_h'\cdot\hol([\gamma_h])=\sigma'\cdot\hol([\gamma_{2i}])=2kv_\theta$.
+Therefore, $\sigma_h(\gamma_h)$ is a geodesic based at the same point and in the same direction as $\beta$, and the two are homologous.\\
+Case II: $kv_\theta\in\cE_1$.
+The proof for this is analogous with the exception that the vertical curve $\beta$ is mapped to is homologous to the waist curve of a vertical cylinder in $\mathfrak{V}$.\\
+Show $T=2||kv_0||$.
+\end{proof}
+\end{thm}
+
+\begin{cor}
+Let $\beta$ be as before. Then $\beta$ lifts to a  geodesic $\tilde{\beta}$ on $\btU$ of the same period that terminates at a point differing from $\mtild$ by some translation of $\btU$.
+\begin{proof}
+Since $\beta$ is homologous to $\sigma(\gamma_i)$ for some $\sigma\in\XX$, from [lem] we obtain that $\Omega_{u,v}([\beta])\neq(0,0)$.
+It follows that $\tilde{\beta}(T)=\Omega_{u,v}([\beta])\cdot(\tilde{\beta}(0))=\Omega_{u,v}([\beta])\cdot\mtild$ is some non-trivial $\ZZ^2$ translation of $\mtild$ that $\tilde{\beta}$ terminates at.
+The equality of the period follows from $\btU$ being a $\ZZ^2$ cover of $\bM$.
+\end{proof}
+\end{cor}
 
 \begin{thm}
-The fundamental group of the $\mathbb{Z}^2$-cover is obtained by lifting the kernel of the closed paths of $\mathbf{M}$ of the homomorphism:
-
-\begin{align*}
-\Omega_{u,v}:\pi_1(\mathbf{M},x_0)\rightarrow \mathbb{Z}^2 \text{; } \beta\mapsto(i(u,[\beta]),i(v,[\beta])),\text{ where}
-\end{align*}
-
-\begin{align*}
-i:H_1(\mathbf{M},\mathbb Q)\times H_1(\mathbf{M},\mathbb Q)\rightarrow \mathbb Z.
-\end{align*}is the intersection number of two homology classes.
-\begin{proof}
-We know from Theorem $\ref{thm:z2}$ that the translational symmetries of $\tilde{\mathbf{U}}$ induced by $T^{m,n}$ is isometric to $\mathbb{Z}^2$.
-Since $\mathbf{M}$ is a genus 5 base surface, we know that $\pi_1(\mathbf{M},x_0)\simeq \mathbb{Z}^{10}$, and the associated cover satisfies $\mathbf{M}=\tilde{\mathbf{U}}/(\pi_1(\mathbf{M},x_0)/N)$, such that $N$ is a normal subgroup of $\pi_1(\mathbf{M},x_0)$.
-This means that $N\simeq \mathbb{Z}^8$.
-The eight core curve classes are the abelianized forms of  $\gamma_0,\gamma_2,\gamma_3,\gamma_5,\gamma_6,\gamma_8,\gamma_9,$ and $\gamma_{11}$ that span N.
-The classes and their signs are obtained from Figure $\ref{fig:mtilda}$ as the outer regions identified by the translations of $T^{m,n}$.
-Thus any closed path on $\mathbf{M}$ is lifted to a closed path on the cover under the quotient map only when a path has a trivial intersection number with the classes.
-\end{proof}
-
-\end{thm}
-Two paths are homologous if they return the same intersection number with the classes of closed core cylinder curves of $\mathbf{U}$ that span its homology.
-The classes u and v are obtained from the group group action of $T^{m,n}$ on the cover.
-
-\begin{Def}
-$\mathbf{hol}:\mathbf{M}\backslash\text{Sing}(\mathbf{M})\rightarrow\mathbb{C}$ is the holonomy vector pulled back from a non-singular path $\gamma$ in $\mathbf{M}$ onto the complex plane given by $\mathbf{hol}(\gamma)=\int_{\gamma}dz$.
-\end{Def}
-
-We denote the \textbf{closed path} $\alpha$, such that $\mathbf{hol}(\alpha)=6+6i$, and show it is homologous to the closed geodesic with the same holonomy vector.
-The slope one direction also decomposes $\mathbf{M}$ into two cylinders by a series of saddle connections of length $\sqrt{2}$ between singularities:
-
-\begin{figure}[H]
-\centering
-\includesvg[width=4.4in.]{slopeonecylinder}
-\caption{The two right-most cylinders $C_1$ (labeled) and $C_2$(unlabeled).}
-\end{figure}
-
-The circumferences of these two cylinders are $6\sqrt{2}$. Geodesic flows on this surface are well defined, and rational directions
-
-\begin{Def}
-Let $\omega^{\theta}_t:[0,1]\times\mathbb{R}/2\pi\mathbb{Z}\rightarrow\mathbf{M}$ be the \textbf{maximal geodesic flow} on the surface in direction $\theta$ such that $\omega^{\theta}_0=\omega^{\theta}_1=x_0\in\mathbf{M}\backslash\text{Sing}(\mathbf{M})$.\\
-$\omega^{\frac{\pi}{4}}_t$ is the geodesic flow in the \textbf{slope one direction}, and $\chi$ is its image in $\mathbf{M}$ and element of $\pi_1(\mathbf{M},x_0)$.
-\end{Def}
-
-\begin{lem}
-$\alpha$ is homologous to $\chi$ in  $\mathbf{M}\backslash\text{Sing}(\mathbf{M})$.
-\begin{proof}
-Let $\chi$ be a geodesic contained in either $C_1$ or $C_2$.
-Since a geodesic does not admit singularities, it is the image of a closed path on $X\backslash\text{Sing}(X)$ with initial point $x_0$ on the strips of $C_1$ and $C_2$ with boundaries removed, denoted $C_1',$ $C_2'$.
-Express $[\alpha]$ as $\Sigma^{11}_{j=0}$ $\frac{1}{2}\gamma_j$ (a closed path climbing up the staircase).
-We show that the intersection numbers of $[\alpha]$ and $[\chi]$ are the same for every core cylinder curve $\gamma$, i.e.
-$i([\gamma_k], \Sigma^{11}_{j=0} \text{ } \frac{1}{2}[\gamma_j])=i([\gamma_k],[\chi])\text{ } \forall k=0,\dots,11$.\\
-\textbf{Case one:} k is even.
-If k is even, then every curve $\gamma_k$ is oriented to the right.
-Since $\chi$ intersects every curve once, $i([\gamma_k],[\chi])=1$.
-No even indexed curves intersect eachother, so we need only consider when $j$ is odd.
-Now if $j$ is odd, it is incident (positively crossing) with only two horizontal curves, namely $\gamma_{j+1},\gamma_{j-1}$.
-Therefore $i([\gamma_k],[\alpha])=i([\gamma_{j-1}],\frac{1}{2}[\gamma_{j}])+i([\gamma_{j+1}],\frac{1}{2}[\gamma_{j}])=\frac{1}{2}(i([\gamma_{j-1}],[\gamma_{j}])+i([\gamma_{j+1}],[\gamma_{j}]))=\frac{1}{2}(1+1)=1.$\\
-\textbf{Case two:} k is odd.
-If k is odd, then $[\chi]$ will have an intersection number of $-1$ with $[\gamma_k]$ since odd-indexed core curves are oriented upwards.
-Now since k is odd, we only consider when $j$ is even.
-Similarly, this means that $\gamma_j$ negatively intersects the two vertical core curves with adjacent indices.
-Hence, $i([\gamma_k],[\alpha])=i([\gamma_{j-1}],\frac{1}{2}[\gamma_{j}])+i([\gamma_{j+1}],\frac{1}{2}[\gamma_{j}])=\frac{1}{2}(i([\gamma_{j-1}],[\gamma_{j}])+i([\gamma_{j+1}],[\gamma_{j}]))=\frac{1}{2}(-1-1)=-1.$\\
-We know intersection number to be bilinear and non-degenerate on homology.
-So if $\alpha$ and $\chi$'s abelianizations admit the same intersection numbers for every curve in the spanning set of $H_1(\mathbf{M}, \mathbb{Q})$, then $[\alpha]=[\chi]$.
-\end{proof}
-\end{lem}
-
+Let $\tildphi_t^\theta$ be a geodesic on $\btU$.
+Suppose there is some $k\in\RR$ such that $kv_\theta\in\cS$.
+Then $\tildphi_t^\theta$ is periodic if and only if $kv_\theta\in\cO$.
+Further, $\tildphi$ has period $T=6||kv_\theta||$.
+\begin{proof}
+Suppose $\tildphi_t^\theta$ is periodic with period $T>0$.
+Let $\tilde{\alpha}=\tildphi_t^\theta\vert_{[0,T]}$ be a closed segment on $\btU$.
+Let $\alpha:[0,T]\rightarrow\bM$ be an $m_0$ based geodesic in the direction of $\theta$ with the same image as $\tilde{\alpha}$'s projection onto $\bM$.
+We've shown in [lem] that $T=6||kv_\theta||$ in both cases.
+Suppose for contradiction that $kv_\theta\notin\cO$.
+Then $kv_\theta\in\cE$.
+By [lem], this would imply $\alpha$ lifts to a geodesic that is not closed.
+If the lift is based at $\mtild=\tilde{\alpha}(0)$, then these are two identical geodesics on the surface of the same length.
+This would mean that $\tilde{\alpha}(T)\neq \mtild$, which is a contradiction.
+Therefore, $kv_\theta\in\cO$.\\
+Conversely suppose $kv_\theta\in\cO$.
+Then it follows from [lem] that a geodesic $\alpha$ on $\bM$ lifts to a closed geodesic on $\btU$ of the length $T=6||kv_\theta||$, a restriction of $\tildphi$ to $[0,T]$.
+It follows that $T$ must be the period of $\tildphi$ since it is the smallest possible value such that $\alpha$'s lift closes up.
+\end{proof}
+\end{thm}
+~\\\\\\\\\\\\\\\\
 \begin{thm}
-$\chi\in\pi_1(\mathbf{M},x_0)$ lifts to $\tilde{\chi}\in\pi_1(\tilde{\mathbf{U}},\tilde{x}_0)$
-\begin{proof}
-From Lemma 2, $[\chi]=[\alpha]$, so $\Omega_{u,v}(\chi)=\Omega_{u,v}(\alpha)$. Since
-$i([u],[\alpha])=-i([\gamma_2],[\alpha] +i([\gamma_5],[\alpha]) + i([\gamma_8],[\alpha]) - i([\gamma_{11}],[\alpha])=-1+(-1)+1-(-1)=0$ and $i([v],[\alpha])=1-(-1)-1+(-1)=0$, it follows that $\alpha,\chi\in\text{Ker }\Omega_{u,v}$, and $\chi$ lifts to a closed geodesic on $\tilde{\mathbf{U}}$.
-\end{proof}
-\end{thm}
-
-\begin{cor}
-content...
-\end{cor}
-
-From here, we use $\alpha$ to show that the \emph{only} trajectories that close on the Necker cube surface are those that are in vector direction $(a,b)$ such that $\gcd(a,b)=1$ and $a,b$ are both odd.
-We call these \textbf{odd-odd} directions.
-We can make this claim because the group generated by the matrices
-
-\begin{equation*}
-\left[ \hspace{1mm} \begin{matrix}
-				1 &  \pm 2\\
-				0 & 1
-			\end{matrix}\hspace{1mm}\right] \text{ and }
-			\left[ \hspace{1mm} \begin{matrix}
-							1 & 0\\
-							 \pm 2 & 1
-						\end{matrix}\hspace{1mm}\right]
-\end{equation*}
-is the \emph{Sanov subgroup} of $SL(2,\mathbb{Z})$ and only sends elements in the odd-odd set to itself.
-There are dualizations made between how these matrices skew a geodesic direction,  and how their original affine transformations induce an effect homology.
-In a sense the kernel is obtained by the orbit of $\chi$ under $\mathbb{X}$ and its holonomy vector under $\mathbb{X}'$.
-
-\begin{lem}
-The actions of $\mathbb{X}'$ on $\mathcal{O}$ and $\mathcal{E}$ are closed in their respective sets.
-\begin{proof}
-Since $\mathbb{X}'$ is generated by the elements $\mathbf{A}'$, $\mathbf{B}'$, and $\mathbf{R}'$, any matrix $G'\in\mathbb{X}'$ is of the form $G' = (\mathbf{A}')^{1_1}\circ(\mathbf{B}')^{1_2}\circ(\mathbf{R}')^{1_3}\circ(\mathbf{A}')^{2_1}\circ\dots(\mathbf{A}')^{n_1}\circ(\mathbf{B}')^{n_2}\circ(\mathbf{R}')^{n_3}$, where $i_k\in\mathbb{Z}$ for $i=1,\dots,n$ and $k=1,2,3.$ Let $x=\left(\begin{matrix}p \\ q  \end{matrix}\right),y\in\mathcal{O}$, and consider the equation $G'x=y$.
-Observe that $\left(\begin{matrix}1 && 2 \\ 0 && 1\end{matrix}\right)^l x=\left(\begin{matrix}p+2jq \\ q  \end{matrix}\right),$ and $ \left(\begin{matrix}1 && 0 \\ 2 && 1\end{matrix}\right)^m x=\left(\begin{matrix}p \\ q+2mp  \end{matrix}\right)$ for any $l,m\in\mathbb{Z}$.
-Also note that for any $j\in\mathbb{Z}$, $(\mathbf{R}')^{m}x=\left(\begin{matrix}p \\ q  \end{matrix}\right),\left(\begin{matrix}-q\\ p  \end{matrix}\right),\left(\begin{matrix}-p \\ -q  \end{matrix}\right),\left(\begin{matrix}q \\ -p  \end{matrix}\right)$ when $j \mod{4}\equiv0,1,2,3$, respectively.
-In any case, the product of any power of a generator of $\mathbb{X}'$ and any $x\in\mathcal{O}$ is an element of $\mathcal{O}$ .
-By letting $l=i_1,m=i_2,$ and $j=i_3$, we first consider the base case when $i=n$.
-Let $G'=G'_1\circ\dots\circ G'_n$, such that $G'_i=(\mathbf{A}')^{i_1}\circ(\mathbf{B}')^{i_2}\circ(\mathbf{R}')^{i_3}$.
-Since $n_1,n_2,n_3$ are arbitrary integers, $G'_n x \in\mathcal{O}$.
-Suppose for some $b< n-1$, $G'_{n-b}\circ\dots\circ G'_n x =y'\in\mathcal{O}$.
-Therefore $y'=(G'_1\circ\dots\circ G'_{b})^{-1}y$, which implies that $(G'_1\circ\dots\circ G'_{b})^{-1}$ preserves the set $\mathcal{O}$.
-Otherwise, if $y\in\mathcal{E}$, there exists at least one $G'_i$ for $1< i< b$ and $\tau\in\mathcal{E}$ such that $G'^{-1}_i \tau = (\mathbf{R}')^{-i_3}\circ(\mathbf{B}')^{-i_2}\circ(\mathbf{A}')^{-i_1} \tau \in \mathcal{O}$, a contradiction.
-Since elements in $\mathbb{X}'$ are invertible, $G'_1\circ\dots\circ G'_{b}$ must also map $\mathcal{O}$ to itself.
-Left multiply both sides of the equation to show that $G'_1\circ\dots\circ G'_n x = G' x = y$.
-By the principle of strong induction, this holds for all $0<b\leq n$.
-Since $G'$ is invertible and an arbitrarily chosen element of $\mathbb{X}'$, it follows that $x\in\mathcal{O}$ if and only if $y\in\mathcal{O}$ and $\mathcal{O}$ is closed under $\mathbb{X}'$.\\
-The proof for when $x\in\mathcal{E}$ is made in the same way.
-\end{proof}
-\end{lem}
-
-Now a trajectory in the horizontal direction has a directional vector of (1,0).
-The orbit of this vector by the Veech group is the set of all \textbf{even-odd} vectors.
-We also know that in this direction a geodesic is drift-periodic (See figure 1).
-The Veech group of $\mathbf{M}$ preserves these properties.
-Suppose you had some closed geodesic on $\mathbf{M}\backslash Sing(\mathbf{M})$ called $\beta$ such that $\beta=h(\alpha)$, where $h\in$  Aff$^+(\mathbf{M})$, and $h_*$ is its induced homomorphism.
-Then we want to show that
-
-\begin{align*}
-(i([\beta],[u]), i([\beta],[v]))=(i([\alpha],h^{-1}_*[u]), i([\alpha],h^{-1}_*[v])) = (0,0).
-\end{align*}
-
-But first, we look at some of the properties of the group $\mathbb{X}_*$.
-
-\begin{thm}
-Let $\mathbb{X}_*$ be the group generated by $\mathbf{A}_*, \mathbf{B}_*, \mathbf{R}_*, \mathbf{H}_*,$ and $\mathbf{V}_*$.
-Let $G=\left< \mathbf{A}_*, \mathbf{B}_* \right>$, $T=\left< \mathbf{H}_*, \mathbf{V}_* \right>$, and $R=\left< \mathbf{R}_*\right>$.
+Let $\tildphi_t^\theta$ be a geodesic on $\btU$.
+Suppose there is some $k\in\RR$ such that $kv_\theta\in\cS$.
+Then $\tildphi_t^\theta$ is drift-periodic if and only if $kv_\theta\in\cE$.
+Further, $\tildphi_t^\theta$ has period $T=2||kv_\theta||$.
+\begin{proof}
+
+\end{proof}
+\end{thm}
+
+\section{Conclusion}
+
+\begin{thm*} \textbf{Periodic and Drift-Periodic Geodesics on the Necker cube surface.} 
+\\Let $\Phi_t:\RR\rightarrow\bUs$ be a non-singular unit-speed geodesic on the Necker cube surface such that $\Phi(0)=u_0$ and $\Phi'(0)=v\in\RR^3$.
+Let $v_0\in\RR^2$ be a distance-preserving projection of $v$ onto the plane tangent to the surface at $u_0$.
+Let $k\in\RR$ such that $kv_0\in\cS$.
 Then the following is true:
 \begin{enumerate}[label=(\roman*)]
-\item $G$ is a free subgroup of $\mathbb{X}_*$ of rank two.
-\item $T$ is a finite cyclic subgroup of $\mathbb{X}_*$ and a centralizer of G.
-\item $R$ is a finite cyclic subgroup of $\mathbb{X}_*$, and a normalizer of $G$.
+\item $\Phi$ is periodic if $kv_0\in\cO$ with period $T=6||kv_0||$.
+\item $\Phi$ is drift-periodic if $v_0\in\cE$ with period $T=2||kv_0||$.
 \end{enumerate}
-\begin{proof} Let $h_*^{j}=\mathbf{A}_*^{k_j}\circ \mathbf{B}_*^{g_j}\in G$ for $k_j, g_j \in\mathbb{Z}$, $j=1,\dots,n$.\\ \emph{(i)}.
-When $\mathbf{A}_*$ and $\mathbf{B}_*$ act on $\gamma_i$, it is only ever trivial if i is even for $\mathbf{A}_*$ or i is odd on $\mathbf{B}_*$.
-Since i cannot be both odd and even at the same time, there is no relation between the two generators and therefore $G$ is free.\\
-\emph{(ii)} It is up to the reader to show that $T$ has the relations $\mathbf{H}_*^2=\mathbf{V}_*^2=(\mathbf{H}_*\mathbf{V}_*)^3=id_*$, and is isomorphic to the rotational group of the hexagon generated by reflections about adjacent vertices of a 12-gon.
-Observe that $\mathbf{H}_*\circ \mathbf{A}_*^{k_j}\circ[\gamma_i]=\mathbf{H}_*\circ[\gamma_i] + \frac{k_j}{2}(1-(-1)^i)(\mathbf{H}_*\circ[\gamma_{i-1}]+\mathbf{H}_*\circ[\gamma_{i+1}])=[\gamma_{-i}] + \frac{k_j}{2}(1-(-1)^i)([\gamma_{1-i}]+[\gamma_{-i-1}])=\mathbf{A}^{k_j}\circ[\gamma_-i]=\mathbf{A}^{k_j}\circ\mathbf{H}_*\circ[\gamma_i],$ and $\mathbf{V}_*\circ \mathbf{A}_*^{k_j}\circ[\gamma_i]=\mathbf{V}_*\circ[\gamma_i] + \frac{k_j}{2}(1-(-1)^i)(\mathbf{V}_*\circ[\gamma_{i-1}]+\mathbf{V}_*\circ[\gamma_{i+1}])=[\gamma_{10-i}] + \frac{k_j}{2}(1-(-1)^i)([\gamma_{11-i}]+[\gamma_{9-i}])=\mathbf{A}^{k_j}\circ[\gamma_10-i]=\mathbf{A}^{k_j}\circ\mathbf{V}_*\circ[\gamma_i].$ In the same way one can show this to be true for $\mathbf{B}_*^{g_j}$, and we can see that $T$ is a centralizer of $G$.
-\\
-\emph{(iii)} $R$ is obviously cyclic and finite since an isomorphism is obtained as $\mathbf{R}_*\mapsto\mathbf{R}'\in SO(2,\mathbb{Z})$.
-\\Note that $\mathbf{R}_*\circ\mathbf{A}_*^{k_j}\circ[\gamma_i]=\mathbf{R}_*\circ[\gamma_i] + \frac{k_j}{2}(1-(-1)^i)(\mathbf{R}_*\circ[\gamma_{i-1}]+\mathbf{R}_*\circ[\gamma_{i+1}])\\
-=(-1)^i[\gamma_{1-i}] + \frac{k_j}{2}(1-(-1)^i)((-1)^{i-1}[\gamma_{2-i}]+(-1)^{i+1}[\gamma_{-i}])\\
-=(-1)^{1-i}([\gamma_{1-i}] - \frac{k_j}{2}(1+(-1)^{1-i})([\gamma_{2-i}]+[\gamma_{-i}]))\\
-=(-1)^{1-i}\mathbf{B}_*^{-k_j}\circ[\gamma_{1-i}]=\mathbf{B}_*^{-k_j}\circ(-1)^{1-i}[\gamma_{1-i}]=\mathbf{B}_*^{-k_j}\circ\mathbf{R}_*\circ[\gamma_{i}].$\\
-Likewise, $\mathbf{R}_*\circ\mathbf{B}_*^{g_j}\circ[\gamma_{i}]=\mathbf{A}_*^{-g_j}\circ\mathbf{R}_*\circ[\gamma_{i}]$.
-\end{proof}
-\end{thm}
-
-\begin{rem}
-It can be easily shown that $\mathbb{X}'$ has similar properties.
-\end{rem}
-
-\begin{lem}
-$Let$ $h_*\in \left<\mathbf{A}_*,\mathbf{B}_*\right>.
-$ Then $h_*\circ[\alpha]\in H_1(\mathbf{M},\mathbb{Q})$ can be expressed as $h_*\circ[\alpha]=\frac{1}{2}(c_1\Sigma^{5}_{j=0}[\gamma_{2j}]+c_2\Sigma^{5}_{j=0}[\gamma_{2j+1}])$ for $c_1,c_2\in\mathbb{Z}$.
-\begin{proof}
-Let $\Sigma^{5}_{j=0}[\gamma_{2j}]=\Sigma\Gamma_{even}$, $\Sigma^{5}_{j=0}[\gamma_{2j+1}]= \Sigma\Gamma_{odd}$, and  $\Sigma^{11}_{j=0}[\gamma_{j}]= \Sigma\Gamma$.
-Let $h_*=h_*^n\circ\dots\circ h_ *^1$, and $h_*^{i}=\mathbf{A}_*^{k_i}\circ \mathbf{B}_*^{g_i}$ for $k_i, g_i \in\mathbb{Z}$, $i=1,\dots,n$.
-Compose these two homomorphisms and obtain $\mathbf{A}_*^{k_i}\circ \mathbf{B}_*^{g_i}(\Sigma\Gamma)=(4g_ik_i+2k_i)\Sigma\Gamma_{even}+2g_i\Sigma\Gamma_{odd}+\Sigma\Gamma$.
-Let $c_i^1=(4g_ik_i+2k_i), c_i^2=2g_i$, and solve for $h_*^{i+1}\circ h_*^i\circ\Sigma\Gamma$:
-\begin{align*}
-h_*^{i+1}\circ h_*^{i}\circ(\Sigma\Gamma)=h_*^{i+1}\circ(c_i^1\Sigma\Gamma_{even}+c_i^2\Sigma\Gamma_{odd}+\Sigma\Gamma)\\ =c_i^{1}h_*^{i+1}\circ(\Sigma\Gamma_{even})+c_i^2h_*^{i+1}\circ(\Sigma\Gamma_{odd})+h_*^{i+1}\circ(\Sigma\Gamma)\\ =2g_{i+1}\Sigma\Gamma_{odd}+(4g_{i+1}k_{i+1}+2k_{i+1})\Sigma\Gamma_{even}+\Sigma\Gamma\\+c_i^{1}(4g_{i+1}k_{i+1}\Sigma\Gamma_{even}+2g_{i+1}\Sigma\Gamma_{odd}+\Sigma\Gamma_{even})\\+c_{i}^2(2k_{i+1}\Sigma\Gamma_{even}+\Sigma\Gamma_{odd})\\
-=\Sigma\Gamma+(c_i^1+(c_{i}^1+1)(4g_{i+1}k_{i+1})+(c_i^2+1)2k_{i+1})\Sigma\Gamma_{even}\\+(c_i^2+(c_i^1+1)2g_{i+1})\Sigma\Gamma_{odd}\\
-\text{Let }c_{i+1}^1:=(c_i^1+(c_{i}^1+1)(4g_{i+1}k_{i+1})+(c_i^2+1)2k_{i+1}),\\
-c_{i+1}^2:=(c_i^2+(c_i^1+1)2g_{i+1}).
-\end{align*}
-From these recursive definitions and a finite sequence of integers, $\{k\}_i,\{g\}_i$, observe then that\\ $h_*\circ[\alpha]=h_*\circ[\frac{1}{2}\Sigma\Gamma]=\frac{1}{2}h_*\circ[\Sigma\Gamma]=\frac{1}{2}[c_n^1\Sigma\Gamma_{even}+c_n^2\Sigma\Gamma_{odd}+\Sigma\Gamma]\\
-=\frac{1}{2}[(c_n^1+1)\Sigma\Gamma_{even}+(c_n^2+1)\Sigma\Gamma_{odd}]$. Further simplify by letting $c_1=c_n^1+1,c_2=c_n^2+1$.
-\end{proof}
-\end{lem}
-
-\begin{lem}
-Let $h_*\circ[\alpha]\in H_1(\mathbf{M},\mathbb{Q})$. Then for $a\in\left<\mathbf{H}_*, \mathbf{V}_*\right>$ and $b\in\left<\mathbf{R}_*\right>$, the following is true:
-\begin{align*}
-& a\circ h_*\circ[\alpha]=h_*\circ[\alpha]\\
-& b\circ h_*\circ[\alpha]=\frac{1}{2}[c'_1\Sigma\Gamma_{even}+c'_2\Sigma\Gamma_{odd}]\\
-& h_*\circ b\circ[\alpha]=\frac{1}{2}[c''_1\Sigma\Gamma_{even}+c''_2\Sigma\Gamma_{odd}]
-\end{align*}
-\begin{proof}
-By Theorem 4, a is a centralizer of the group so $a\circ h_*\circ[\alpha]=h_*\circ a \circ[\alpha]=h_*\circ \frac{1}{2}a \circ[\Sigma\Gamma].$ Since a is a cyclic permutation of the set $\Gamma$, it acts trivially on $\Sigma\Gamma$.
-Therefore, $a\circ h_*\circ[\alpha]=h_*\circ \frac{1}{2}[\Sigma\Gamma]=h_* \circ[\alpha]$.\\
-By theorem 4, $a\circ\mathbf{A}_*^{k_i}\circ\mathbf{B}_*^{g_i}=\mathbf{B}_*^{-k_i}\circ\mathbf{A}_*^{-g_i}\circ a$.
-Extend this property to $h_*$, and denote the normalized element as $h_{**}$, such that $b\circ h_{*}=h_{**}\circ b$.
-Note that $b(\Sigma\Gamma)=b(\Sigma\Gamma_{even}+\Sigma\Gamma_{odd})=\Sigma\Gamma_{odd}-\Sigma\Gamma_{even}$.
-$b\circ h_*\circ[\Sigma\Gamma]=c_1b\circ\Sigma\Gamma_{even}+c_2b\circ\Sigma\Gamma_{odd}=c_1\Sigma\Gamma_{odd}-c_2\Sigma\Gamma_{even}.$ So, $c_1'=-c_2$ and $c_2'=c_1$.\\
-Since $h_*$ is arbitrary, let $h_{**}=g_{*}$ be generated by an integer sequence that defines the word and consider $h_*\circ b\circ[\Sigma\Gamma]=b\circ g_{*}\circ[\Sigma\Gamma]=c^*_1b\circ\Sigma\Gamma_{even}+c^*_2b\circ\Sigma\Gamma_{odd}=c^*_1\Sigma\Gamma_{odd}-c_2^*\Sigma\Gamma_{even}.$ So, $c_1''=-c_2^*$ and $c_2''=c_1^*$.
-\end{proof}
-\end{lem}
-
-\noindent Now that every element in the orbit of $[\alpha]$ can be expressed as a linear combination of integers, it is simple to show they lift to a closed trajectory in the cover.
-
-\begin{Def}
-Let $\mathbf{dir}:UT(\mathbf{M}\backslash\text{Sing}(\mathbf{M}))\rightarrow\mathcal{O}\cup\mathcal{E}$ be the injective map from $\mathbb{R}/2\pi\mathbb{Z}$ to $\mathbb{Z}^2$ given as $\mathbf{dir}(\theta)=(k_1\cos(\theta),k_2\sin(\theta)), k_1,k_2\in\mathbb{R}$ such that $\gcd(k_1\cos(\theta),k_2\sin(\theta))=1$.
-\end{Def}
-
-\begin{thm}(Sketch)\\
-Any geodesic, $\beta$, in $\mathbf{M}$ lifts to a closed geodesic $\tilde{\beta}$ on $\tilde{\mathbf{U}}$ if and only if $\mathbf{dir}(Arg(\mathbf{hol}(\beta)))\in\mathcal{O}$.
-\begin{proof}
-Call the quotient cover $p:\tilde{\mathbf{U}}\rightarrow\mathbf{M}$, and fix a point $\tilde{x_0}\in p^{-1}(x_0)$.
-Let $\beta=h(\chi)$, where $h\in\mathbb{X}$.
-We also obtain $[\beta]=h_*\circ[\alpha]$ from Lemma 2.
-Since $h$ sends geodesics to geodesics, h induces the following: $\mathbf{hol}(h(\chi))=h'(\mathbf{hol}(\chi))=h'(6+6i)$ for $h'=\left[\begin{matrix} a & b \\ c & d\end{matrix}\right]\in\mathbb{X}'$.
-So, Arg$(h'(\mathbf{hol}(\chi)))=$Arg$(6[(a+b)+i(c+d)])=$Arg$(6h'(1+i))$.
-Lemma 3 states that for any $h'\in\mathbb{X}'$, $h'(\mathcal{O})=\mathcal{O}$.
-Therefore there is no such geodesic of \textbf{even-odd} slope in the orbit of $\chi$.
-Otherwise $h',h\notin\mathbb{X}',\mathbb{X}$.
-Consequently, $\mathbf{dir}(Arg(\mathbf{hol}(\beta)))\in\mathcal{O}$.\\
-From Lemma 5 we see that $[h(\chi)]=h_*\circ[\chi]=h_*\circ[\alpha]=\frac{1}{2}(c_1\Sigma^{5}_{j=0}[\gamma_{2j}]+c_2\Sigma^{5}_{j=0}[\gamma_{2j+1}])$ for $c_1,c_2\in\mathbb{Z}$.
-Denote the sums as $\Sigma\Gamma_{even}$ and $\Sigma\Gamma_{odd}.$\\
-Therefore, $2i([u],h_*\circ[\alpha])=c_1i([u],\Sigma\Gamma_{even})+c_2i([u],\Sigma\Gamma_{odd})\\
-=c_1(-i([\gamma_2],0) +i([\gamma_5],[\gamma_6]+[\gamma_4]) + i([\gamma_8],0) - i([\gamma_{11}],[\gamma_{10}]+[\gamma_{0}]))\\
-+c_2(-i([\gamma_2],[\gamma_1]+[\gamma_3]) +i([\gamma_5],0) + i([\gamma_8],[\gamma_7]+[\gamma_9]) - i([\gamma_{11}],0))\\
-=c_1(-(0)+(-1-1)+(0)-(-1-1))+c_2(-(1+1)+(0)+(1+1)-(0))=0.$\\\vspace{0.1in}\\
-Similarly, $2i([v],h_*\circ[\alpha])=c_1i([v],\Sigma\Gamma_{even})+c_2i([v],\Sigma\Gamma_{odd})\\
-=c_1(i([\gamma_0],0)-i([\gamma_3],[\gamma_2]+[\gamma_4])-i([\gamma_6],0)+i([\gamma_9],[\gamma_8]+[\gamma_{10}])\\
-+c_2(i([\gamma_0],[\gamma_{11}]+[\gamma_1])-i([\gamma_3],0)-i([\gamma_6],[\gamma_5]+[\gamma_7])+i([\gamma_9],0)\\
-=c_1((0)-(-2)-(0)+(-2))+c_2((2)-(0)-(2)+(0))=0.$\\
-Therefore, $\Omega_{u,v}(h(\chi))=(0,0)$, and $h(\chi)=\beta\in\text{Ker }\Omega_{u,v}$ for all $h\in\mathbb{X}$.
-By Theorem 2, $\beta$ lifts to $\tilde{\beta}\in\pi_1(\tilde{\mathbf{U}},\tilde{x}_0)$.
-Let $\theta=$Arg$(\mathbf{hol}(\beta))$.
-Then $\omega_t^{\theta}$ at $x_0$ lifts to $\tilde{\omega}_t^{p^{-1}(\theta)}\in\tilde{\mathbf{U}}\backslash\text{Sing}(\tilde{\mathbf{U}})$.\\
-Now suppose instead that $\beta=h(\gamma_{i})$.
-Then $\mathbf{dir}(\beta)=\frac{1}{2}(1+(-1)^i,1-(-1)^i)$.
-According to Lemma 3, $h'(\mathcal{E})=\mathcal{E}$.
-Thus we have no geodesic in the \textbf{odd-odd} directions obtained from the orbits of $(1,0)$ and $(0,1)$.
-For contradiction, suppose that $h(\gamma_{i})\in\textbf{Ker }\Omega{u,v}$.
-Then $(i(h_*\circ[\gamma_i], [u]),i(h_*\circ[\gamma_i], [v]))=(i([\gamma_i], h^{-1}_*\circ[u]),i([\gamma_i], h^{-1}_*\circ[v]))=(0,0).$ Let $h^{-1}_*\circ[u]=\Sigma^{11}_{j=0}x_j[\gamma_j]$, and $h^{-1}_*\circ[v]=\Sigma^{11}_{j=0}y_j[\gamma_j]$.
-Note that since $\gamma_i$ intersects $\gamma_{i\pm1}$, $i([\gamma_i], h^{-1}_*\circ[u])=(-1)^{i+1}(x_{i-1}+x_{i+1})$ and $i([\gamma_i], h^{-1}_*\circ[v])=(-1)^{i+1}(y_{i-1}+y_{i+1}).$
-\\
-$\mathbf{Unfinished.}.$
-\end{proof}
-\end{thm}
-
-\begin{conj}{\textbf{Dynamics of Geodesic Flow on the Necker cube surface.}}\\ Obtain $\theta$ and $\vec{\theta}$ as described in Definition 3.
-Denote the non-singular unit-speed geodesic flow with initial point $s\in(\mathbf{U}\backslash Sing(U))$ in direction $[\theta]\sim\phi\in UT(\mathbf{U}\backslash Sing(\mathbf U))$ by $F_t^{\phi}:\mathbf{U}\times\mathbb{R}_0^+\rightarrow\mathbf{U}$ on $(\mathbf{U},\mu)$, where $\mu$ is a flow-invariant measure.
-Then the following is true:
-\begin{enumerate}[label=(\roman*)]
-\item (Periodic) There exists a $t_0 > 0$ such that $F^{\phi}_{t+t_{0}}(s)=F^{\phi}_{t}(s)$ if and only if $\vec{\theta}\in\mathcal{O}$.
-\item (Drift-Periodic) There exists a $t_0 > 0$ such that $F^{\phi}_{t+t_{0}}(s)= F^{\phi}_{t}(s)+c$, where $c\in \mathbf{U}$ is a non-trivial translation of a point in $\mathbf{U}$, if and only if $\vec{\theta}\in\mathcal{E}$.
+\end{thm*}
+
+\subsection{Calculating Drift Distance}
+Given a drift-periodic trajectory on the surface, it's possible to attain the total Euclidean distance after a single period by finding the element that its corresponding homology class is mapped to under $\Omega_{u,v}$.
+This is identical to the value given by an element in the deck group $\pi_1(\bGs)/\pi_1(\bUs)\cong\ZZ^2$ once projected back onto $\bUs$.
+Its total drift distance can be measured in $\bUs$'s ambient space, $\mathbf{P}$.
+\begin{figure}[H]
+\centering
+\includegraphics[width=5in]{drift1.png}
+\end{figure}
+\noindent A more formal take on this distance is by finding the initial and terminal points in $\RR^3$, the ambient space of $\bUs$'s embedded form.
+Consider the following families of unit squares in $\RR^3$:\\
+
+This piecewise linear map takes the embedded form to the plane $(z=0)$ and preserves path metric.
+It embeds the surface into $\mathbf{P}$, where these edge identifications are made.
+It's the same map that is used to model trajectories with flatsurf.\cite{flatsurf} A demonstration can be found online.
+[cite demo page]
+\begin{figure}[H]
+\centering
+\includegraphics[width=5in]{drift2.png}
+\end{figure}
+
+In any case, the inverse of this map will take these initial and terminal points on $\bUs$ to where they belong in $\RR^3$.
+From the initial point, $u_0$, the geodesic in ambient space drifts by some amount, $2m+i2n\in\CC$.
+On $\mathbf{P}$ this is taken to $\RR^3$ under $\Phi^{-1}$, where the distance between these points is calculated.
+A simple example is the geodesic in the horizontal direction.
+On $\mathbf{P}$ it drifts by $(2,0)$, but in $\RR^3$ the difference is $(1,0,-1)$.
+
+\begin{cor}
+Let $\Phi_t$ be a geodesic on $\bUs$ in direction $v_0$ with $kv_0\in\cE$.
+Since $\Phi_t$ is drift periodic, let $T=2||kv_0||$ be its period.
+Let $\alpha\in\pi_1(\bGs,x_0)$ be the homotopy class containing $p_1(\Phi([0,T]))$ and suppose $\varphi_1(\alpha)=\begin{pmatrix}x\\y \end{pmatrix}$.
+Then the Euclidean distance in $\RR^3$ between $\Phi(0)$ and $\Phi(T)$ is $\sqrt{2}\sqrt{x^2+y^2+xy}$.
+\begin{proof}
+Suppose $\Phi(0)$ is in a smooth section of $\bUs$. Then under $\Psi$
+\end{proof}
+\end{cor}
+
+\subsection{Open Problems}
+Certain questions remain open in the study of this surface, or families of this type of surface.
+I would like to leave a few possible directions to go in regards to further understanding its various properties.
+These kinds of problems are generally pretty approachable, and motivates the study and application of these techniques.
+\begin{enumerate}
+\item Can one find an equation that takes an element in $\cE$ to the total distance that $\Phi$ drifts in $\RR^3$ after a single period?
+Being able to find the total path length was a consequence of creating the translation surface $\bM$ and studying its various geodesics, but it might not be possible to find this $\ZZ^2$ element under $\Omega_{u,v}$ without solving a system of $\ZZ$-linear equations.
+It's possible to solve for a matrix in $SL(2,\RR)$ that sends $(1,0)$ or $(0,1)$ to your element in $\cE$, but there isn't a unique matrix that does this.
+This matrix might be unique in $\XX'$, but it's hard to say for sure.
+We conjecture that its components would give up information on the total drift amount of the geodesic in this $\cE$ direction.
+\item What happens to $\Phi$'s dynamical properties when we parameterize the length, width, and height of $\bG$, which is then used to construct an infinite cube surface.
+We refer to the collection of surfaces that the Necker cube surface belongs to as the \emph{Family of Infinite Step Surfaces}.
+There is no doubt a similar method leads to a translation surface that does not have uniform transverse cylinder decompositions.
+Simulations have shown us that the geodesic $\Phi_t$ with initial direction $\frac{1}{\sqrt{2}}(\pm1,\pm1)$ closes on a variety of step surfaces.
+Classes of closed trajectories might still be in the orbit of the geodesic in the same direction under the translation surface's symmetries.
+There may even be many more intricate collection of classes of geodesics that drift/close than the split between the $\cO$ and $\cE$ directions we've found here.
+\item Can one calculate the entire Veech group of $\bM$?
+There are some rotational symmetries of this surface that we've neglected in favor of a simpler subgroup of $V(\bM)$.
+These rotational symmetries might lend a hand in solving the previous problem.
+\item Are geodesics recurrent on the Necker cube surface or other step surfaces?
+A geodesic on $\bU$ with an irrational initial trajectory angle exhibits strange patterns that suggest this, but its overall properties are yet to be determined.
 \end{enumerate}
-\begin{proof}(Sketch)\\
-Denote the covering maps $f:\tilde{\mathbf{U}}\rightarrow\mathbf{U}$, $p:\tilde{\mathbf{U}}\rightarrow\mathbf{M}$, and fix a point $\tilde{x}_0\in f^{-1}(s),p^{-1}(x_0),$ for $x_0\in\mathbf{M}$.
-$f^{-1}([\theta])=\{x:x=\theta+n\frac{\pi}{2}, n\in\mathbb{Z}\}=[\theta]\subset UT(\tilde{\mathbf{U}}\backslash\text{Sing}(\tilde{\mathbf{U}}))$ given by the four-fold cover and rotations of each individual plane.
-This gives us a relation between the two tangent bundles, where the translation four-fold cover has the standard $\mathbb{R}/2\pi\mathbb{Z}$ unit tangent fiber.
-$\theta'$ is the direction associated to the flow $\omega_t^{\theta'}:[0,1]\rightarrow\mathbf{M}$ on the translation surface.
-Since the cover is translation, $p^{-1}(\theta')=\theta'=\theta+n\frac{\pi}{2}$.
-First suppose that $\vec{\theta}\in\mathcal{O}$.
-Then $\theta$ is identified with the set of directions that close on $\tilde{\mathbf{U}}$.
-From Theorem 5, $\omega_t^{\theta'}$ lifts to a closed geodesic $\tilde{\omega}_t^{\theta'}$.
-Given $\mathbf{hol}(\omega)=\int_{\omega}dz$, we obtain a period for the unit-speed flow, $t_0=|\mathbf{hol}(\omega)|$.
-That is, $\tilde{F}_t^{\theta'}:\mathbb{R}^+_0\rightarrow\tilde{\mathbf{U}}$ such that  $\frac{d}{dt}\tilde{F}_t=\frac{1}{|\mathbf{hol}(\omega)|}$.
-Then $F_t^{\phi}=F_t^{[\theta']}=f\circ \tilde{F}^{\theta'}_t$.
-The period carries over since there is no concern over a trajectory returning to $\tilde{x_0}$ in a different direction.
-Otherwise, the geodesic $\omega_t^{\theta'}$ on $\mathbf{M}$ would have closed in $0<t<1$.
-Now suppose that $\vec{\theta}\in\mathcal{E}$.
-Identifying it with $\theta'$, we see that $\omega$ in direction $\theta'$ is not an element of $\textbf{Ker }\Omega_{u,v}$ from Theorem 5.
-Therefore, $\Omega_{u,v}(\omega)=(m,n)\simeq T^{m,n}$ and lifting the terminal point $\omega(1)$, $\tilde{\omega}(1)=T^{m,n}(\tilde{\omega}(0))=T^{m,n}(\tilde{x}_0).$ The period remains unchanged, in that $\tilde{F}^{\theta'}_{t+\mathbf{hol}(\omega)}=\tilde{F}^{\theta'}_{t}+T^{m,n}(\tilde{x}_0)$.
-Therefore, $F^{\phi}_{t+t_{0}}(s)=f\circ\tilde{F}^{\theta'}_{t}(\tilde{x}_0)+f\circ T^{m,n}(\tilde{x}_0).$\\
-Conversely, suppose $F_t$ is periodic.
-Then $[\theta]=\phi=[\theta']$, which defines directional flows $\tilde{F}^{\phi}_t$.
-According to Theorem 5, $\tilde{F}^{\phi}_t$ will close if and only if $\phi\subset\mathcal{O}$.
-$\phi$ is the orbit of $\vec{\theta}'$ under the 90 degree rotational matrix.
-This matrix does not alter the length or period of a geodesic.
-Thus, $F^{\phi}_t$ is exactly one of the flows $\tilde{F}^{\phi}_t$.
-Likewise, if $F^\phi_t$ is drift-periodic then $F^\phi_{t+t_0}=f\circ \tilde{F}^\phi_t+f\circ T^{m,n}$.
-$T^{m,n}$ is trivial if and only if $\theta'\in\mathcal{O}$.
-Therefore, $\theta'\in\mathcal{E}$, and $[\theta']=\phi$.
-\end{proof}
-\end{conj}
-
-There is still much work to do in terms of cleaning up the proofs and organizing the final paper.
-
-\newpage
-\section*{Conclusion}
-What I ultimately aim to do is port these results on X's homology back to the Necker cube surface.
-I want do it in such a way that the final theorem is bi-conditional.
-To do so, I imagine I can take a vector image of a small segment of a geodesic in $\mathbb{R}^3$ and project it onto the isometric flattening of the Necker Cube surface to obtain a direction (or classes of equivalent directions), and relate it to the unit tangent bundle of $\mathbf{M}$.
-
-In addition, I would also like to find a formula for the arc-length of a geodesic based on direction alone.
-Knowing that hol$(\alpha)=6+6i$ means that the induced Euclidean metric on $\mathbf{M}\backslash$Sing($\mathbf{M}$) gives the geodesic an arc-length of $6\sqrt{2}$.
-I would like to show that:
-
-\begin{align*}
-\int_{\beta}|dz|=|hol(h(\alpha))|=|h'(hol(\alpha))|,
-\end{align*}
-\noindent where $h'\in V(\mathbf{M})$ is the derivative of h, and $\beta=h(\alpha)$.
-We know that is true on the translation surface, but it's a matter of then showing the translation quotient, branch-cover, and the Necker cube surface have the same induced Euclidean metric of these non-singular geodesics.
-(It would not be surprising considering that the surface is built out of subsets of planes.) Even more of a problem is finding a way to solve for a matrix in the Sanov subgroup that brings (1,1) to the desired odd-odd slope.
-
 \newpage
